### 2020-11-09 1_Introduction.tex
Initial Overleaf Import

\section{Introduction}
\label{intro}

% The source code of a program can be naively understood as a collection of sentences in a language that would be interpreted by a computer (or compiled into a lower level representation and then executed by a computer). Nonetheless, this understanding is far from being complete. While it is not possible to execute programs without writing them in a way that the machine is able to interpret, it is also very difficult to develop software that performs complex tasks if developers do not care about how easily other programmers are going to understand its source code \cite{DBLP:journals/cj/Knuth84}.

% \rb{acho que esse proximo paragrafo pode ser excluido}
% \diego{também acho}

% {\color{blue}That means programming should, in various ways, be regarded as an act of communication. In this context, the programmer is often faced with a dilemma. When initially confronted with a problem they have to solve, the programmer will proceed to employ their cognition into developing the logic of the solution, whilst simultaneously having to worry about writing ``syntactically correct'' code for the machine to parse and translate. The problem is that, during the time spent solving the problem and communicating with the computer, developers often do not have enough cognitive resources to also make considerations about whether their code is understandable by other programmers. Therefore, it takes a change of track to make the code easier to understand \cite{DBLP:books/daglib/0019908}. Not proceeding with such change of track is one of the reasons why ``confusing'' language constructs and idioms arise and remain in a program.}

Source code misunderstandings can occur due to diverse reasons.
A poorly understood requirement might lead to poorly implemented code.
An intricate algorithm may require extra time to be fully comprehended.
Nevertheless, programming languages can also be at fault.
Specific syntax and semantics of a language can cause misunderstandings. The same applies to some code idioms and stylistic choices~\cite{DBLP:conf/msr/GopsteinZFC18}.

% \diego{acho que o prox. parágrafo deve ser adaptado ou até removido}
% {\color{blue}A direct consequence of the presence of these idioms in the code base is that it becomes even more difficult to develop new features for the software. 
% According to Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} ``\emph{the ability to understand pre-existing source code is one of the most important elements of a continuously successful software project}'', it follows that, if one cannot properly understand the \emph{purpose} and \emph{meaning} of a piece of code, they are not going to be able to build upon its existing parts.
% Other consequences of the presence of confusing code include: 
% (a) increased probability of bug introduction, due to a lack of proper understanding of what the program does; 
% and (b) loss of internal quality measures, such as cohesion and coupling. 
% This phenomenon usually leads to an architectural erosion. 
% All these situations, which usually compound each other, expose the necessity of developing recommendations, tools, and techniques that help developers to avoid writing confusing code, identify potentially confusing code, and transform confusing code into clear code. 
% Therefore, it is necessary to understand what programming language constructs and idioms might lead to code that is hard to understand.}

JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}. At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''\footnote{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}. It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST, ClientSideBugsJS:TSE}, and asynchronous operations. All these 
%diverse characteristics and capabilities
features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST, ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.

Our research aims to help practitioners identify potentially confusing snippets of code and replace them with clearer code.
% Developers could improve their codebases' understandability by following our recommendations.
To achieve that, we characterize a set of 9 JavaScript \emph{atoms of confusion}; small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.\castor{I think it would be really nice to have an example here. Maybe even a figure with a left-hand and a right-hand side. What do you say? If yes, I
'd go for post increment or assignment as value. If not a figure, the text could be as follow: An example of atom of confusion identified by previous work~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} is the use \texttt{if} statements without the accompanying curly braces. The alternative, in this case, is simply to always use the curly braces.}
Among the atoms in this work, X are shared between JavaScript and \clang/\cpplang (e.g., pre/post-increment), while Y others are specific to JavaScript, e.g., the atom Assignment as Value.
Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/He19,DBLP:conf/sigsoft/GopsteinIYDZYC17} has already explored the impact of atoms of confusion on program understanding,
and confirmed they negatively impact source code understandability.
However, they were constrained to statically typed languages.
To the best of our knowledge, we are the first to explore atoms of confusion in the context of a dynamically typed language (JavaScript).\castor{Review the number of atoms, X, and Y.}

\diego{"Understanding source code comments at large-scale"\cite{DBLP:conf/sigsoft/He19} (acima): 
embora relacionado não me parece um estudo de átomos de confusão}

To understand whether an atom of confusion introduces misunderstandings, we surveyed 150 developers from the Reddit community. Their task was to predict the output of short snippets of code, with half of them containing atoms of confusion. We found that the presence of the Comma Operator atom can increase the average time to predict the output by 65\%, while the behavior of snippets containing the Pre-Increment atom were predicted 44\% faster.  An atom's presence---or lack of---in open-source repositories may reveal cases where developers should be more or less careful when writing code. To identify such cases, we mined the 100 most popular JavaScript projects from GitHub and found a total of X atoms in the analyzed projects. To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15 experienced professional developers. Participants preferred code snippets \textit{without} an atom in 70\% of the cases, therefore corroborating to the survey findings.\castor{It would be nice to have results about correctness.}

Altogether, the main contributions of this paper are:
\rb{revisar as contribuicoes quando o artigo estiver mais maduro. por enquanto, vamos ignorar esses bullets.}

\begin{itemize}
    \item A study about the impact of atoms of confusion on a
    dynamic language.
    \item A list of additional atoms of confusion addressing particular 
    JavaScript constructs and idioms.
    \item A library and infrastructure to mine atoms of confusion using 
    code queries (implemented in the .QL language). This library and 
    infrastructure are freely available and might help researchers and 
    practitioners to automate program analysis at a large scale.
\end{itemize}

% Our research aims to identify and characterize JavaScript \textit{atoms of confusion}: small pieces of code that hinder developers' understanding of the source code of a program~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
% Previous research has confirmed the negative impact of atoms of confusion when considering statically typed languages (e.g., \clang and \cpplang)~\cite{DBLP:journals/ese/MedeirosLAAKRG19, DBLP:conf/msr/GopsteinZFC18}.

% Recent works have explored the impact of \emph{atoms of confusion}~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/He19,DBLP:conf/msr/GopsteinZFC18} on program understanding.
% Atoms of confusion correspond to small pieces of code that hinder developers' understanding of the source code of a program. 
% Previous research has confirmed the negative impact of atoms of confusion when considering statically typed languages (e.g., \clang and \cpplang)~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/msr/GopsteinZFC18}.

% In this work, to the best of our knowledge, we are the first to explore this issue in the context of a dynamically typed language (JavaScript), answering the following research questions:
% \emph{How do JavaScript developers  identify atoms of confusion as contributing to program misunderstanding?}
% \emph{What is the impact of atoms of confusion on misunderstanding JavaScript code?},
% and \emph{What is the frequency of occurrence of  atoms of confusion in practice?} 
% To answer these research questions, we conduct this investigation using a mixed-methods approach, involving a survey, a set of interviews, and an activity of mining source code repositories. Altogether, the main contributions of this paper are

% \rb{revisar as contribuicoes quando o artigo estiver mais maduro. por enquanto, vamos ignorar esses bullets.}

% \begin{itemize}
%     \item A first study about the impact of atoms of confusion on a
%     dynamic language. {\color{red}present some of the results here.}
%     \item A library and infrastructure to mine atoms of confusion using 
%     code queries (implemented in the .QL language). This library and 
%     infrastructure is freely available and might help other researcher and 
%     practitioners to automate program analysis at a large scale. 
%     \item A list of additional atoms of confusion addressing particular 
%     JavaScript constructs and idioms. {\color{red}present some insights 
%     about these additional atoms here}
% \end{itemize}

### 2020-11-09 2_Background.tex
Initial Overleaf Import

\section{Background and Related Work}
\label{back}
\subsection{Program Comprehension}\castor{It would be nice to have a table with the atom candidates similar to the one that appears in Gopstein et al. 2017.}

The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96}. Regardless of whether they are performing maintenance on legacy software or adding new features, programmers have to understand what the source code does before making any improvements. A natural consequence of this is that a programmer must also have in mind that other programmers (or their future selves) are going to have to spend time trying to understand the code being currently developed before doing any work themselves.

As program comprehension is regarded less as a systematic process than as an objective \cite{DBLP:journals/ibmsj/OHareT94}, there is no single approach or framework that is capable of yielding easily understandable programs. Factors such as problem domain, size and complexity of the code base, and available tools all vary significantly between different working settings. These aspects, combined with differences between individuals, which can include experience, working memory capacity, and other cognitive traits, make it clear that program comprehension should be approached in different ways in different contexts.

However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions--- when these exist- allowed by the language or framework they are working with. Although this can be, to an extent, a personal matter, since different individuals might hold different opinions on whether a given syntax is difficult to understand, it is possible to empirically investigate if there exist programming languages' constructs or idioms that are considerably harder to understand, leading readers of the code to frequently make incorrect predictions of its behavior.

\subsection{Atoms of Confusion}\label{sec:aoc}

The concept of atoms of confusion was introduced by Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings. In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion. Their work laid an empirical framework to identify such atoms. In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atoms were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.\castor{Mudei um pouco o protocolo para identificar átomos. Tinha achado a explicação anterior confusa.}
%By having subjects evaluate code where in the expected output relied exclusively on evaluating the result of previously known confusing programs that had been written in \clang or \cpplang, and asking the participants to predict the output of a set of code snippets where the atoms had been removed, the authors were able to isolate and measure the impact of such confusing atoms.

One main example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a number literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 potential atoms, the authors were able to identify 15 snippets that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase. Our first study replicates this study, for all atoms that have a counterpart in JavaScript.\castor{Is ``replicates'' the best word here? From what I could grasp, there are some differences between the protocols.} Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on the different nature of confusion amongst the atoms, noting that, for some atoms, a particular incorrect answer was prevalent, whilst there was a distribution of incorrect ones in other atoms.

The work of Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} presented the results of comprehensive research on atoms of confusion that can be found in the \clang and \cpplang~languages. Even though JavaScript is a dynamic language that differs from \clang in a variety of aspects, it also has a number of constructs in common. In addition, the imperative aspects of the language are syntactically similar to \clang, e.g., assignments, conditional expressions, \texttt{if}-statements, pre and post increments and decrements, among others. On the one hand, this means that some of the atoms of confusion that exist in \clang programs may also occur in JavaScript code. On the other hand, differences between the languages and the surrounding programming cultures may lead to code patterns that are atoms of confusion in one language not being atoms in the other one.  
%Although JavaScript allows us to program in different paradigms from the one C follows, most of the JavaScript code found in popular repositories are written under the imperative paradigm. Also, both languages share many syntactical rules, which allows us to replicate some of the C atoms in JavaScript.

Medeiros et al. \cite{DBLP:journals/ese/MedeirosLAAKRG19} analyzed 50 open source projects written in the C language with the goal of evaluating 12 code patterns called ``misunderstanding patterns'' by the authors. Many of these misunderstanding patterns are either atoms of confusion or atom candidates~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. This study shows that these patterns are prevalent; among these 50 projects, there are more than 109K occurrences of misunderstanding patterns. 
In order to gauge the relevance of these misunderstanding patterns, the authors sent 35 pull requests removing occurrences of these patterns to randomly-selected open source projects. The authors of the study received feedback for 21 of these pull requests and the maintainers of the projects accepted 8 of them (22.86\%). The authors also analyzed 36 coding style guides from these projects, but found few guidelines specifically addressing misunderstanding patterns.

% Para compreender a relevância desses misunderstanding patterns os autores enviaram 35 pull requests selecionados aleatoriamente, para remover as ocorrências desses padrões encontrados no código dos projetos avaliados. Os autores receberam feedback para 21 pull requests, e os desenvolvedores aceitaram 8 pull requests (38\%). Os autores também analisaram 36 Guidelines dos open-source projects provide for developers regarding misunderstanding code patterns.

Castor~\cite{castor2018} presented a preliminary catalog of six atom candidates for the Swift programming language. Unlike JavaScript, in Swift most of the atoms identified by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} are avoided by construction, e.g., it does not have assignments as values, increment operators, or macros. The paper also presents a structured definition for what it means for a code pattern to be an atom of confusion and a set of general principles to help in the identification of new atoms.

%Castor \cite{castor2018} apresentou um catalógo preliminar de átomos de confusão na linguagem de programação Swift. O autor identificou 6 estruturas candidatas a átomos de confusão em Swift. Foram omitidos os candidatos que tinham uma correspondência direta com os átomos previamente identificados por Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.

According to Arnaoudova et al.~\cite{Arnaoudova:2016:LAW}, Linguistic Antipatterns (LA) are \textit{``recurring poor practices in the naming, documentation, and choice of identifiers in the implementation of an entity [..] that may impair program understanding''}. These antipatterns are design issues that go against developer's  intuition about a program, for example, a \texttt{set} method that returns a result. Similarly to atoms of confusion, LAs occur at the source code level and hinder code readability. Differently from atoms of confusion, LAs are more akin to design principles than code idioms. In addition, they are language-agnostic, unlike atoms of confusion.

Santos and Gerosa~\cite{Santos:2018:ICP} analyze the impact of eleven coding practices on program readability. Some of them, such as avoiding more than three levels of code
block nesting, are atom candidates. This work evaluates the coding practices solely based on personal opinions of the study subjects. Furthermore, it focuses on the Java language. Finally, most of the practices are more related to code formatting than language constructs or idioms.

\castor{Provavelmente devemos reduzir o espaço tomado pelos quatro trabalhos acima, a depender da existência de outros. O de Linguistic Antipatterns talvez até devesse sair.}
\rb{Incluir os trabalhos do grupo do Marcio Ribeiro e do Fernando Castor. Quais outros trabalhos deveriamos incluir aqui?}

### 2020-11-09 3_Method.tex
Initial Overleaf Import

\section{Study Settings}
\label{method}

The main goal of this research is to investigate the impact of atoms of confusion in JavaScript code comprehension. As such, in this paper we answer the following general research questions:

\begin{enumerate}[(RQ1)]
    \item Do JavaScript developers identify atoms of confusion as contributing to program misunderstanding? 
    \item What is the impact of atoms of confusion on developers' comprehension of JavaScript code? 
    \item What are the particular JavaScript idioms and language constructs that might cause source code misunderstanding?
    \item What is the frequency of occurrence of atoms of confusion in practice (i.e., in open-source JavaScript projects)?
\end{enumerate}

Answering these research questions has several implications. For instance, we can either generalize or refute the perceptions about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice. Our motivation to detect the presence and impact of atoms of confusion is also motivated by the possibility of laying the foundations for the implementation of libraries that can automatically transform code into its simpler versions---though we postpone these results to a future research work. This might lead to a new catalog of refactoring for removing atoms of confusion.  To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and an effort of mining JavaScript software repositories.

 \subsection{Settings for the first study: a survey}\label{sec:survey-settings}\castor{Eu mudaria os títulos desta sub-seção e das duas próximas para "First study: Survey", "Second study: Mining open source repositories", etc. Vale a pena manter "First study", etc.?}
% \subsubsection{The Atoms Considered In This Work}

In the first study, our goal was to collect evidence to answer two specific questions: 
\begin{itemize}
    \item \emph{Do code snippets that contain atoms of confusion produce a higher misunderstanding rate when programmers try to predict their outcome?}
    \item \emph{Do code snippets with atoms of confusion require programmers to take longer to predict their output?}
\end{itemize}

To address these questions we conducted a survey to assess the impact of atoms of confusion on understanding JavaScript code. Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that may also exist in JavaScript programs. {\color{red}In addition, we consider a particular atom that is exclusive to JavaScript in this first study}. Appendix~\ref{sec:appendix-atoms} summarizes all atoms we consider in our research.

\subsubsection{Survey design}

The design of our first study aims to block two variables (developer experience and the code snippets) and considers two treatments: the presence or absence of 
atoms of confusion within the code snippets. 
To achieve such a design goal, controlling the effect of experience and individual code snippets, we resorted to the \textit{Latin Square Design} \cite{Hunter-Experimenters}. Using this design we create a $2 x 2$ matrix in which each row represents a subject and each column indicates the set of code snippets. The design of each square (a replica) is such that no treatment is repeated in the same row or column. For example, if a given subject (P1) is asked to predict the output of code snippets that contain atom candidates \texttt{\{1,2,3,4,5\}}, then, when answering questions about non-confusing code snippets, they will only be presented with non-confusing versions of atom candidates \texttt{\{6,7,8,9,10\}}. Furthermore, a given subject (P2), which constitutes the second row of our example square, will be asked questions about the non-confusing versions of \texttt{\{1,2,3,4,5\}}, and will answer questions about confusing snippets of \texttt{\{6,7,8,9,10\}}. By doing that, we guarantee that all of our 20 snippets are contained within each square, and that each configuration occurs only once within a square. Figure \ref{fig:latinsquare} offers a visual representation of the concept.

  \begin{figure}[htb!]
      \noindent
      \centering
      \includegraphics[scale=.50]{images/latin-square.pdf}
      \caption{Latin square design. Each ``square'' corresponds to 
      a replica in our study. Each replica comprises two students (square rows) 
      and two sets of code snippets (CS-1 and CS-2). We randomly apply the 
      treatments (atom or non-atom code) to the cells of the squares.} 
      \label{fig:latinsquare}
  \end{figure}

Having selected 10 atoms in our first study, we wrote small pieces of code that contained each atom. We also wrote their equivalent snippet without the confusing idioms and constructs, leading to a total of 20 code snippets. In order to reduce the cognitive effort, we decided that each subject would be asked to predict the output of a subset of 10 listings, wherein each subset contained 5 blocks that contained atoms of confusion, whilst the remaining 5 had the atoms removed. The order in which the questions were presented was randomized. By doing this, we were seeking to minimize the chances of subjects being aware that the current listing they were analyzing contained (or not) atoms of confusion.
That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets did not). 
We measured answer correctness and total time each participant needed to answer reach question of their survey.

\subsubsection{Survey Instrument}

We implemented our survey as a web application. As part of this 
effort, we carried out an informal pilot whose main objectives were: to spot bugs in the application and in the data collection mechanism; to gain feedback from respondents about the user experience of the application; and to formulate an estimate about how long answering the survey would, on average, take. Fellow undergraduate students, professional colleagues, and friends took the pilot survey. Some users reported layout defects, and many reported that the landing page did not explain the survey well enough. We also spotted minor issues with our routines to create and populate the Latin Squares.

We organized the survey in {\color{red}three} sections. The first section aims to characterize the subjects, asking for their age, education level, and programming experience. We also included a check button, whose checking meant users agreed that all collected data would be used solely for research purposes. In the second section, we presented to the participants a small set of instructions, where we explained how the survey worked and asked them to dedicate their attention to it. We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters. For each question page, we kept track of whether or not the subjects switched windows.

The next section of the survey presented a sequence of 10 questions, each containing a code snippet. For each question, there was a text box where the answer should be written. There was also an ``I do not know'' button, which, when clicked, led the subject to the next question. In our setting, ``I do not know'' was treated as a wrong answer. The code snippets were presented as images copied from a text editor, so as to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter. Upon submitting their answer for a particular question, the subject was automatically led to a similar page, containing another snippet.

We decided not to provide feedback about the time students took to answer each question. Nor did we tell them whether their answers were correct or not. Our main concern was to avoid introducing bias for future respondents. Since we posted the survey in a social media platform, if we gave respondents instant feedback, they might post comments on particular atoms, therefore interfering with future participants' thought process.

% \rb{nao acho esse paragrafo necessario} 
% \adriano{o de baixo ne? eu tambem acho}

% As we mentioned before, we first wrote the code listings in a text editor, and took pictures of it. In the case of an atom of confusion that was exclusive to JavaScript, which we called \textit{Automatic Semicolon Insertion} (see Appendix~\ref{sec:appendix-atoms}), it was necessary to remove the syntax highlighter. Even though semicolons at the end of statements are optional to programmers in JavaScript, the interpreter automatically inserts them into the code. Our text editor was incorrectly highlighting a line break after a return statement, even thought it was valid JavaScript syntax. We had thus to turn the highlighter off to take the picture of this atom.

\subsubsection{Survey audience} 
\diego{Esses detalhes são possíveis candidatos para remoção, se precisarmos de espaço}
{\color{blue}We invited developers to answer our survey by sending invitations to communities of JavaScript programmers on the Internet. Our initial plan was to try to engage contributors for two major JavaScript projects, namely Node.js and NPM. None of the projects, though, offered a direct way to interact with the community, such as an online discussion forums or mailing list. We would be required to first make contact with the projects' leaders, and only then would we have a chance to approach potential respondents. Due to time constraints, we decided to look for respondents elsewhere.}
We then posted the survey on Reddit.\footnote{Reddit is a North American online discussion platform. Its discussion threads, often called subreddits, are sorted by subject. For our research, we posted information and links to the survey in two JavaScript subreddits.}. We explained our research purposes, and asked developers of any level of expertise to take the survey. To incentivize serious engagement, we proposed to raffle \$50 gift cards on Amazon products at the end of the survey. Within twelve hours we collected more than 150 answers, populating more than 70 replicas of the Latin Squares. We were able to collect significant data on time taken and discrepancies in answer correctness between confusing and non-confusing versions of the snippets.

\rb{acho que podemos mover esse proximo paragrafo para a secao de resultados, ou de ameacas}
\adriano{acho que pode ir para resultados. nao considero uma ameaca pois eu e o Caio tratamos manualmente as entradas a serem removidas e garantimos a integridade dos latin squares}

Inconsistencies may arise when each square is being built. The main source of inconsistency we faced was when a user quit in the middle of the survey. When this happened, his row in the square was left incomplete. We considered all squares which contained incomplete rows to be invalid, and discarded them. Since we had a large enough number of samples, the squares we had to discard did not impact our results.

% \subsubsection{Pilot Survey}

% To validate the web application we developed to conduct the survey, we ran an informal pilot survey whose main objectives were:
%     \begin{itemize}
%         \item To spot bugs in the application and in the data collection mechanism;
%         \item To gain feedback from respondents about the user experience of the application;
%         \item To formulate an estimate about how long the survey would, on average, take.
%     \end{itemize}

% We had fellow undergraduate students, work colleagues and friends take the pilot survey. We discovered that the aspect that needed most improvement was the user experience. Some users reported layout defects, and many reported that the landing page did not explain the survey well enough. We also spotted minor issues with our routines to create and populate the Latin Squares. After performing all the necessary changes, we were ready to conduct another experiment.

\rb{Comentei o survey com alunos}

% \subsubsection{A Survey With Undergraduate Students}
%     Our second attempt was due to courtesy of a professor of our department. During the semester of the writing of this text, he was lecturing a Data Structures course, which is usually taken during the second semester of the undergraduate course. The professor agreed to take the students to a laboratory during one of his classes, and all the students who attended that day took the survey.

%     No issues were reported with the web application, and all answers were appropriately collected. This time, all the Latin Squares were being populated correctly, making the data we collected eligible for use in the final analysis. However, we opted to dispose of all the data we gathered in this occasion. What motivated this decision were the two following aspects:
%     \begin{enumerate}
%         \item The students were not focusing hard enough. Most students were not very serious about the survey. While we watched them answer the questions, we could see many students peeking at each other's screens. We could also clearly see students not taking nearly enough time to reason about the code, and answering the questions without giving it proper thought.
%         \item We needed a more diverse spectrum of respondents. In the beginning of our research, we established that we would need at least 25 Latin Squares in order to be able to have significant results. By conducting this survey, we were able to populate 11 Latin Squares. Compounded by the problem reported above, we were concerned that our spectrum of respondents might end up skewed towards undergraduate students with little or no JavaScript programming experience.
%     \end{enumerate}

%     These two facts led us to conclude we needed to conduct a survey in which partakers engaged voluntarily, whilst also having an incentive to think hard enough about the problems they faced. To address these points, we prepared our final setting.

\subsection{Settings for the second study: Mining open source repositories}

To understand how often atoms of confusion appear in real settings, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories. To this end, we first collected the most popular GitHub repositories that are primarily written is JavaScript. We measured popularity using the project's stargazers, and selected only projects with stargazers\_count > 100. This metric, available on  GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs}.

After filtering out JavaScript project candidates, in the second step we built a curate dataset comprising the top 100 most popular repositories. Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}. Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research. The size of the projects range from small ones (883 liens of code) to complex systems with more than 10 MLOC. All projects in our dataset have at least \num{1161} forks and \num{23643} stars. We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.

In the third step we mined atoms of confusion from the repositories in our curate dataset. We written source code queries usng the CodeQL language~\cite{moor:gttse2007}, in order to filter the atoms of confusion. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as Java and JavaScript). 
We also automate the process of running the queries and exporting the results to a format that simplifies our analysis and the reproduction of this study. In the final step we carry out basic descriptive statistics, in order to understand how often the atoms of confusion appear in practice.

\begin{table*}[ht]
 \centering
 \begin{tabular}{rrrrrrr}
   \hline
             & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ \hline
 Lines of Code           & \num{883}  & \num{14939} & \num{39520} & \num{281088.46} & \num{145990.75} & \num{10149685} \\
 Num. of Contributors    & \num{6}   & \num{121} & \num{243} & \num{431.82} & \num{436.75} & \num{4042} \\
 Num. of Forks           & \num{1161}   & \num{3365} & \num{5711} & \num{8201.96} & \num{8438.25} & \num{68826} \\
 Num. of Stars        & \num{23643} & \num{27542.25} & \num{35687.5} & \num{45284.86} & \num{47963.25} & \num{310725} \\
    \hline
 \end{tabular}
 \caption{Some descriptive statistics about the projects used in the study}
 \label{tab:projects-statistics} 
 \end{table*}

\subsection{Settings for the third study: interview with practitioners}

To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion. We also asked each participant if they knew of any other JavaScript-specific construct that they regarded as confusing. In this section we details the protocol we followed to conduct the interviews, and we also describe how their results were analysed.

% Nós realizamos entrevistas semi-estruturas com o objetivo de identificar a percepção dos desenvolvedores com experiência em JavaScript sobre algumas questões relacionadas à compreensão de código em JavaScript. Assim, nesta Seção nós descrevemos os procedimentos adotados para selecionar os participantes para as entrevistas e detalhamos como as entrevistas foram conduzidas. Além disso, detalhamos como os resultados das entrevistas foram analisados.

\subsubsection{Participants Selection}: The participants of the interviews were invited personally by the authors, and the pool of potential participants was based on our network of contacts. Our main criterion of selection was that all participants had to be working professionally with JavaScript. We invited a total of 17 developers, and 15 of these agreed to participate, and throughout a two week period, we conducted all the interviews.

% Os participantes das entrevistas foram selecionados a partir da nossa rede de contatos. Nós convidamos 17 pessoas e 2 não responderam ao nosso convite para participar das entrevistas. Durante 2 semanas, nós entrevistamos 15 desenvolvedores. Iniciamente nós conduzimos duas entrevistas como piloto. Após o piloto com dois desenvolvedores com mais de 03 anos de experiência, nós adicionamos uma nova pergunta na entrevista: Você conhece alguma construção em JavaScript que leve a um mal entendimento do código? Ao finalizar as entrevistas, nós perguntamos ao desenvolvedor se ele conhecia algum outro desenvolvedor com experiência em JavaScript que pudesse nos indicar para participar das entrevistas. Nós entrevistamos 02 desenvolvedores a partir da indicação dos entrevistados. A Tabela \ref{pinterview} apresenta o perfil dos 15 participantes das entrevistas.

\subsubsection{Interview Process}

We conducted semi-structured interviews using a web conferencing software. All the interviews were recorded with the consent of participants. On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes, and the longest one 43.06 minutes. The main goal of the interviews was to have developers assess in real time whether two versions of code with the same behavior differed in readability, and, if so, which version they regarded as the easier one to understand. The interviews were conducted by two of the authors of this text, and a third one listened to all the recordings to cross-validate the collected data.

% Nós conduzimos entrevistas semi-estruturadas utilizando um software de conferência de audio/video. Todas as entrevistas foram gravadas com o consenso dos entrevistados. As entrevistas tiveram uma duração de aproximadamente 26.29 minutos. A entrevista com menor duração foi de 14.59 minutos e de maior duração foi de 43.06 minutos. O objetivo ao realizar as entrevistas era identificar quais as construções em JavaScript podem corresponder a confusões de código. As entrevistas foram realizadas por dois dos autores deste artigo.

The interviews had three main parts. In the first one, we asked the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience and other programming languages they have worked with. Table \ref{pinterview} presents the profile of each participant.

\begin{table*}[htb!]
\centering
\begin{tabular}
{|p{0.4cm}|p{0.9cm}|p{2.0cm}|p{7.0cm}|p{1.9cm}|p{3.0cm}|}
\hline
ID & Gender & Level & Role & JS Experience & Other Languages \\ \hline  
P1 & Male & Graduate & Developer at Federal Court of Accounts - TCU Brazil & 9 years & Java, PHP, C, Go  
\\ \hline
P2 & Male & High School & Developer at Luma Health & 3.5 years & Python, Go, Dart, Lua, C++, C\#
\\ \hline
P3 & Male & Graduate & Developer at Superior Labour Court - TST Brazil & 4 years & Java, C
\\ \hline
P4 & Male & Undergraduate & Developer at Brazilian Association of Research and Industrial Innovation - EMBRAPII & 3 years & Python, C, C++, Java, Go
\\ \hline
P5 & Male & Master Student & Developer na Qubo Technology & 3 years & Python, SQL
\\ \hline
P6 & Male & Graduate & Software Architect at Superior Labour Court - TST Brazil & >15 years & Java, PHP, C, Python , Ruby, C\# and R
\\ \hline
P7 & Male & Graduate & Developer na Stefanini & 6 years & C, C++, Java, Assembly, Kotlin
\\ \hline
P8 & Male & PhD & Developer Regional Labour Court - TRT Goiás, Brazil & 2 years & Java, Python
\\ \hline
P9 & Male & Graduate & IT Analyst at Regional Bank of Brasilia - BRB Brazil & 5 years & PHP
\\ \hline
P10 & Male & Master Student & Business Analyst and Developer at National Confederation of Industry - CNI Brazil & 4 years & Java, C\# and Python
\\ \hline
P11 & Male & Master Student & Software Architect at the University of Brasilia's Informatics Centre & 4 years & Java, Erlang, C\#, Cobol
\\ \hline
P12 & Female & Master & Software Engineers at Novatics & 1 year & C
\\ \hline
P13 & Male & Master & Developer at the Military Police of the Federal District - PMDF Brazil & 13 years & Java, PHP
\\ \hline
P14 & Female & Master Student & Developer at Neoway & 3 years & C, Python, Ruby
\\ \hline
P15 & Female &  Graduate & Developer at SICOOB & 2 years & Java, PHP
\\ \hline
\end{tabular}
    \caption{Demographic information of the participants}
    \label{pinterview}
\end{table*}

The second part of the interview was more open, and our aim here was to allow the subjects to describe their JavaScript experience, as well as to allow them to make suggestions about any JavaScript's constructs they regarded as innately confusing, as this would allow us to include their suggestions in future work. The questions asked in this section were:
\begin{itemize}
    \item In general, do you implement new JavaScript components, or do you maintain components developed by other programmers?;
    \item In your opinion, what are the main pros and cons of the language?;
    \item Do you think JavaScript is a language that produces code that is hard to understand?
    \item Do you regard any particular construct of the language as especially confusing?
\end{itemize}

\rb{acho melhor mover os code snippets da entrevista para um apendice online. deixei comentado aqui.}
\adriano{como os snippets sao os mesmos do TCC, a ja temos o apendice, nao poderiamos colocar no apendice B deste texto? posso escrever os snippets no repositorio tambem. deixa-los no apendice pode tornar o paragrafo abaixo mais compreensivel/verificavel.}

In the third part of the interview, participants were shown the 20 code snippets that were used to build the aforementioned Latin Squares. The participants were told both sides of each pair had the same output, and were asked to evaluate which construct was easier to understand. The interviewers restrained themselves from introducing bias into the answers by not explaining that one of the sides of each pair contained the atom under investigation. Subjects were just presented the snippets and allowed to take the necessary time to decide on the most readable snippet.

% As entrevistas foram divididas em três etapas. Na primeira etapa da entrevista nós perguntamos aos desenvolvedores as seguintes informações: Nome; E-mail; Gender; Level; Role; Experience; Quais as outras linguagens de programação ele/ela conhecia. Na segunda etapa da entrevista,  perguntamos aos desenvolvedores: a) No geral você escreve novos componentes em JavaScript, ou você faz manutenção em componentes desenvolvidos por outra pessoa (explique o seu dia-a-dia); b) Na sua opinião, quais são os pontos fortes e fracos da linguagem JavaScript. c) Você acha que a linguagem JavaScript é uma linguagem que leva a um código difícil de ser compreendido? d) Tem alguma construção em JavaScript que você conhece que pode levar a uma dificuldade de entendimento. Na terceira etapa da entrevista nós apresentamos alguns exemplos de trechos de código em JavaScript, contendo os 10 tipos de construção apresentados na Table \ref{tab:atoms} com e sem átomos de confusão (Lado A e Lado B), conforme apresentado na Table \ref{snippets}. Nós perguntamos aos desenvolvedores: 1) Você acha que o código do lado B prejudica de alguma forma a compreensão do código? Os dois códigos apresentados nos dois lados de alguma maneira eram equivalentes.

% \begin{longtable}
% {| p{.50\linewidth} | p{.50\linewidth} |} 
% % \begin{center}
% \caption*{Code Snippets presented on the interview}
% \label{snippets}
% % \begin{tabular}{ | p{8.2cm} | p{8.2cm} | } \hline
% \hline
% \multicolumn{2}{| c |}{Arithmetic as Logic}  \\ \hline
% Side A  &  Side B \\ \hline
% \begin{lstlisting}[language=JavaScript]
% let array_1 = [1,2,3];
% let array_2 = [2,3,2];
% if(array_1.length - array_2.length){
%   console.log(true);
% }
% else{
%   console.log(false);
% } \end{lstlisting}
% & 
% \begin{lstlisting}[language=JavaScript]
% let array_1 = [1,2,3];
% let array_2 = [2,3,2];
% if(array_1.length != array_2.length){
%   console.log(true);
% }
% else{
%   console.log(false);
% } \end{lstlisting} \\ \hline
% \multicolumn{2}{| c |}{Assignment as value}  \\ \hline
% Side A  &  Side B \\ \hline
% \begin{lstlisting}[language=JavaScript]
% function resetSchedulerState () {
%   let activatedChildren = {length: 10};
%   let index = length = activatedChildren.length = 0;
%   console.log(index);
% }
% resetSchedulerState(); \end{lstlisting}
% & 
% \begin{lstlisting}[language=JavaScript]
% function resetSchedulerState () {
%   let activatedChildren = {length: 10};
%   activatedChildren.length = 0;
%   let length = activatedChildren.length;
%   let index = length;
%   console.log(index);
% }
% resetSchedulerState(); \end{lstlisting} \\ \hline
% %\pagebreak 
% \hline
% \multicolumn{2}{| c |}{Automatic Semicolon Insertion}  \\ \hline
% Side A  &  Side B \\ \hline
% \begin{lstlisting}[language=JavaScript]
% function example(){
%   return
%       10
% }
% if(example() == 10){
%   console.log(true);
% }
% else{
%   console.log(false);
% } \end{lstlisting}
% & 
% \begin{lstlisting}[language=JavaScript]
% function example(){
%   return 10;
% }
% if(example() == 10){
%   console.log(true);
% }
% else{
%   console.log(false);
% } \end{lstlisting} \\ \hline
% \multicolumn{2}{| c |}{Comma Operator}  \\ \hline
% Side A  &  Side B \\ \hline
% \begin{lstlisting}[language=JavaScript]
% let V1 = 5, V2 = 10;
% V1 = (V2 = 1, 2);
% console.log(V1+V2);
% \end{lstlisting}
% & 
% \begin{lstlisting}[language=JavaScript]
% let V1 = 5, V2 = 10;
% V2 = 1;
% V1 = 2;
% console.log(V1+V2);
% \end{lstlisting} \\ \hline
% %\pagebreak
% \hline
% \multicolumn{2}{| c |}{Ternary Operator}  \\ \hline
% Side A  &  Side B \\ \hline
% \begin{lstlisting}[language=JavaScript]
% let config = {size: 3, isActive: false};
% const _config = config.isActive === true ? config : {size: 10};
% console.log(_config.size); \end{lstlisting}
% & 
% \begin{lstlisting}[language=JavaScript]
% let config = {size: 3, isActive: false}
% let _config;
% if(config.isActive === true){
%   _config = config;
% }
% else{
%   _config = {size: 10};
% }
% console.log(_config.size); \end{lstlisting} \\ \hline

% \multicolumn{2}{| c |}{Implicit Predicate}  \\ \hline
% Side A  &  Side B \\ \hline
% \begin{lstlisting}[language=JavaScript]
% let V1 = 10, V2 = 3;
% if (!(V1 % V2)){
%   console.log(true);
% }
% else{
%   console.log(false);
% } \end{lstlisting}
% & 
% \begin{lstlisting}[language=JavaScript]
% let V1 = 1, V2 = 2;
% if ((V2 - V1) == 0){
%   console.log(true);
% }
% else{
%   console.log(false);
% } \end{lstlisting} \\ \hline
% \multicolumn{2}{| c |}{Logic Control Flow}  \\ \hline
% Side A  &  Side B \\ \hline
% \begin{lstlisting}[language=JavaScript]
% let V1 = 3;
% let V2 = 5;
% let V3 = 0;
% while (V1 != V2 && ++V1) {
%   V3++;
% }
% console.log(V1 + V3); \end{lstlisting}
% & 
% \begin{lstlisting}[language=JavaScript]
% let V1 = 1;
% let V2 = 11;
% let V3 = 0;
% while (V1 != V2) {
%   V1++;
%   V3++;
% }
% console.log(V1+V3); \end{lstlisting} \\ \hline
% \multicolumn{2}{| c |}{Omitted Curly Braces \& Indentation}  \\ \hline
% Side A  &  Side B \\ \hline
% \begin{lstlisting}[language=JavaScript]
% let V1 = 1, V2 = 2
% if (V1 > V2)
% V2 = 1
% V1 = 2
% console.log(V2+V1) \end{lstlisting}
% & 
% \begin{lstlisting}[language=JavaScript]
% let V1 = 1, V2 = 2;
% if (V1 > V2) {
%   V2 = 1;
% }
% V1 = 2;
% console.log(V2+V1); \end{lstlisting} \\ \hline
% \multicolumn{2}{| c |}{Post Increment}  \\ \hline
% Side A  &  Side B \\ \hline
% \begin{lstlisting}[language=JavaScript]
% let V1 = 0;
% if (V1++ == 0) {
%   console.log(true);
% }
% else {
%   console.log(false);
% } \end{lstlisting}
% & 
% \begin{lstlisting}[language=JavaScript]
% let V1 = 0;
% if (V1 == 0) {
%   console.log(true);
% }
% else {
%   console.log(false);
% }
% V1 = V1 + 1; \end{lstlisting} \\ \hline
% \multicolumn{2}{| c |}{Pre Increment}  \\ \hline
% Side A  &  Side B \\ \hline
% \begin{lstlisting}[language=JavaScript]
% var index = 1;
% while (++index < 10) {
%   console.log(index);
%   break;
% } \end{lstlisting}
% & 
% \begin{lstlisting}[language=JavaScript]
% var index = 0;
% index = index + 1;
% while (index < 10) {
%   console.log(index);
%   index = index + 1;
%   break;
% } \end{lstlisting} \\ \hline
% % \end{tabular}
% % \end{center}
% % \end{table}
% \end{longtable}

\subsubsection{Interview Analysis}

Nós realizamos a análise das entrevistas em pares. Inicialmente um dos participantes de todas as entrevistas transcreveu os resultados. Posteriormente, um dos autores do trabalho ouviu os áudios das entrevistas e transcreveu os resultados.

\textcolor{red}{Adriano teria algo mais para adicionar aqui na  análise das entrevistas?}

\todo[inline]{adriano:estou escrevendo uma tabela em que cada linha e um dos atomos, e colunas representam \% preferencia com atomo, \% preferencia sem atomo, \% indiferente. Mas acredito que esta tabela deve estar na section results, certo? O resto das informacoes deste paragrafo eu citei acima no Interview Process, entao acho que podemos abolir esta section. -- Edna: Concordo contigo Adriano}

### 2020-11-09 4_Result.tex
Initial Overleaf Import

\section{Results and Discussion}
\label{sec:results}

In this section we present the main findings of our research. We first present the results of each study (Section~\ref{sec:survey-resuts}, Sections~\ref{sec:msr-results}, and Section~\ref{sec:interview-results}). After that, in Section~\ref{sec:discussion} we consolidate our findings and present some implications of our study.

\subsection{Results of the first study: a survey}
\label{sec:survey-resuts}

Our first study aims to investigate the impact of atoms of confusion while developers try to understand JavaScript code. We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers) and \emph{cognitive effort} (time necessary to provide an answer). As discussed in section~\ref{sec:survey-settings}, 140 participants answered the survey in full. Figure~\ref{fig:degree} and Figure~\ref{fig:xp} show the distributions of the subjects, according to their education level and years of programming experience, respectively.

\begin{figure}[htb]
      \centering
      \includegraphics[scale=.25]{images/degree.png}
      \caption{Participants' Education Level}\label{fig:degree}
  \end{figure}

  \begin{figure}[htb]
      \centering
      \includegraphics[scale=.25]{images/experience.png}
      \caption{Participants' Experience} \label{fig:xp}
  \end{figure}

It is possible to realize in Figure \ref{fig:xp} that more than a half of the participants hold either a BS degree or had taken some university course, meaning the participants have had some level of formal education on programming. No subject reported to have never attended university. Out audience comprises experience developers, and 93\% of the respondents have more than one year of experience with programming (see Figure~\ref{fig:xp}). Actually, 56\% of the participants have either between four and ten years of experience (37\%) or more than ten years of experience (19\%). Other 37\% of the respondents have between one and four years of experience. In this way, we can characterize the effect of atoms of confusion in JavaScript code taking into account the perceptions of both novice and experience developers.

Each participant in this study evaluated {\color{red}10} code snippets, from which 5 were in their confusing versions, whilst the other 5 contained simplified versions of the code snippets (i.e., without the confusing constructs and idioms). The participants should provide the expected outcomes of the code snippets. We collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{cognitive workload} (the time taken to answer the question). Our final dataset consists of 70 complete Latin Squares, which means that our assessment consider a total of 140 answers to each pair of code snippets.

Regarding \emph{correctness}, Table~\ref{results_correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey. Accordingly, the version of seven code snippets without atoms of confusion present at least a 15\% improvement in answer correctness. In particular, \emph{Comma Operator} atom presents the highest impact on misunderstanding. 
Curiously, frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also introduce high degrees of confusion. As the boxplot in Figure \ref{fig:boxplotcorrectness} shows, there is a considerable decrease in the average number of incorrect answers when the atoms were not present. Also, the sample of answers where there was no atoms had almost no dispersion, which is a sign that the non-confusing code is easier to evaluate correctly.

% \begin{table}[htbp]
% \caption{Difference in answer correctness between confusing and non-confusing pairs}
% \begin{center}
% \begin{small}
% \begin{tabularx}
% {{\linewidth}}{l p{1.5cm} p{1.1cm} p{1.1cm} p{1.2cm} }
% \textbf{Atom} & \textbf{\%Correct} & \textbf{\%Correct} \\
% &  \multicolumn{1}{l}{With AOC} \multicolumn{2}{l}{Without AOC}  & \Delta (\%)    \\
%  \hline
% Comma Operator & 40 & 93 & +132\%                 \\     
% Automatic Semicolon  Insertion & 46 & 97 & +110\% \\
% Post Increment & 69 & 91 & +31\%         \\
% Omitted Curly Braces & 67 & 83 & +23\%   \\ 
% Assignment as Value & 80 & 97 & +21\%    \\
% Implicit Predicate & 83 & 97 & +16\%     \\
% Logic as Control Flow & 59 & 68 & +15\%  \\
% Ternary Operator & 86 & 94 & +9\%        \\
% Pre-Increment & 71 & 76 & +7\%           \\
% Arithmetic as Logic & 91 & 90 & -1\%    \\
% \end{tabularx}
% \end{small}
% \end{center}
% \label{results_correctness}
% \end{table}

\begin{table}[htbp]
\caption{Average number of correct answers.}
\label{tab:difference-correctness}
\begin{tabular}{lccc} \toprule
 & \multicolumn{2}{c}{Average Number of Correct Answers} & \\ [0.1cm]
 Atom & Code with Atoms & Code without Atoms & $\Delta$(\%) \\ \midrule 
 Comma Operator          & 40 & 93  & +132 \\
 Post Increment          & 69 & 91  & + 31  \\
 Omitted Curly Braces    & 67 & 83  & +23 \\
 Assignment as Value     & 80 & 97  & +21 \\
 Implicit Predicate      & 83 & 97  & +16 \\
 Logic as Control Flow   & 59 & 68  & +15 \\
 Ternary Operator        & 86 & 94  & +9  \\
 Pre-Increment           & 71 & 76  & +7  \\
 Arithmetic as Logic     & 91 & 90  & +1  \\ \bottomrule
\end{tabular}
\end{table}

\begin{figure}[htb!]
\noindent
 \centering
 \includegraphics[scale=.20]{images/wrong.png}
 \caption{Number of wrong answers of each subject}
 \label{fig:boxplotcorrectness}
 \end{figure}

\rb{verificar a consistencia nas questoes de pesquisa mais especificas}
Regarding the first question we 
address in the survey (\emph{Do code snippets that contain atoms of confusion produce a higher error rate than snippets where the atom is removed?}), we found evidence that the atoms of confusion lead programmers to misunderstand JavaScript code. We also realized that just one atom whose correction has a non-significant improvement in the percentage of correct answers---we found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of ten atoms we consider in the survey).

\rb{acho que valeria a pena colocar o teste de hipotese aqui}
\adriano{como fazemos/descrevemos o teste?}

Table \ref{tab:difference-time-taken} displays the results of the measurements of the average time taken to answer with and without atoms of confusion. Once again, the \emph{Comma Operator} was the construct whose removal had the greatest impact. The more expressive results in this measurement are the ones regarding \emph{Logic as Control Flow}, \emph{Implicit Predicate} and \emph{Omitted Curly Braces}, {\color{blue}as they are frequent in JavaScript libraries}.

\rb{n\~{a}o sei se devemos antecipar o resultado do estudo de minera\c c\~{a}o 
nessa se\c c\~{a}o. talvez nao comprometa tanto.}

Figure \ref{fig:timetoanswer} shows a box plot of the difference in time taken the predict the output of average time to assess the code snippets. Although the dispersion is smaller for non-confusing code blocks, the median time did not vary much.

\begin{table}[htbp]
\caption{Average number of correct answers.}
\label{tab:difference-time-taken}
\begin{tabular}{lccc} \toprule
 & \multicolumn{2}{c}{Average Time in Seconds to Answer} & \\ [0.1cm]
 Atom & Code with Atoms & Code without Atoms & $\Delta$(\%) \\ \midrule 
Comma Operator        & 60 & 21  & -65  \\
Logic as Control Flow & 85 & 49  & -42  \\
Implicit Predicate    & 33 & 24  & -27  \\
Omitted Curly Braces  & 43 & 31  & -27  \\
Assignment as Value   & 53 & 49  & -7   \\
Ternary Operator      & 42 & 42  & 0    \\
Post Increment        & 27 & 27  & 0    \\
Arithmetic as Logic   & 29 & 36  & +24  \\
Pre-Increment         & 34 & 49  & +44  \\
         \bottomrule
    \end{tabular}
\end{table}

\begin{figure}[htb!]
      \noindent
      \centering
      \includegraphics[scale=.20]{images/timeeach.png}
      \caption{Box plot of time to answer each snippet}
      \label{fig:timetoanswer}
  \end{figure}

Recalling our second research question of this survey (\emph{Do code snippets with atoms of confusion require programmers to take longer to predict their output?}), the variance in the time to correctly predict the output of a snippet is significantly smaller. Moreover, since the number of answers is the same for each inspected atom, data from table \ref{results_time} shows that the time taken to answer questions without atoms of confusion is 20\% smaller. Not all atoms, though, took less time to predict the answer. In fact, for the \emph{Arithmetic as Logic} and the \emph{Pre-Increment} atoms, there was an increase of 24\% and 44\% in the time taken to answer questions wherein the atoms had been removed. For other two atoms, there is no difference whatsoever in the time taken. We only find in half of our atoms an increase in the time taken to answer. Therefore, even though we are not able to confirm hypotheses H.2 defined in Section 1, further inquiry would be necessary to discover the reason for some atoms taking even longer.

\rb{talvez tenhamos que descrever e testar as hipoteses. a descricao pode ser feita na secao de study settings, e o resultado dos testes pode aparecer aqui.}

% \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}

\subsection{Results of the second study: Interview}
\label{sec:interview-results}

In this section we present the results of two rounds of interviews with practitioners who agreed to participate in the interviews.In the first interview, a total of 15 practitioners were surveyed. In this phase, we collected information regarding programming experience, familiarity with JavaScript, and we also collected their opinions regarding nine candidates to atoms of confusion. This was done by presenting them pairs of code snippets whose behavior was identical, and ask them to classify each snippets readability.

In the second round of interviews, fourteen of the fifteen developers who participated in the first round were shown another 8 candidates to atom of confusion whose frequency we observed during the repository mining phase. The same procedure of showing code that behaves exactly the same and surveying for how easy to understand each snippet was applied in this phase.

Supporting the results of the survey,Tables \ref{tab:interview-results1} and \ref{tab:interview-results2} show that for twelve out of the seventeen scenarios surveyed, the respondents prefer the version of the code without the atom of confusion, and in no case the \emph{neutral} ratio was higher than the option for the version without atoms of confusion. An example entry is contained in Appendix~\ref{sec:appendix-atoms}. 
For a full listing of the code snippets, visit the paper repository.

\subsubsection*{Participants views of the atoms} Interesting, only for the \emph{ternary operator} scenario the participants prefer the version of the code with the atom of confusion, instead of the more cleaner version without the atom of confusion. Even in this case, some participants who opted for the \emph{left-hand-side} (with the atom of confusion) version still believed that the \emph{right-hand-side} (without the atom of confusion) version was more readable (see Figure~\ref{code:ternary}). 
The following quotes were extracted from three interviews:

\begin{figure*}

\noindent\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Ternary Operator} atom)]
let config = {size: 3, isActive: false};
const_config = config.isActive === true 
             ? config 
             : {size: 10};
console.log(_config.size);
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript, caption=\emph{Right-hand side} (without the atom)]
let config = {size: 3, isActive: false}
let _config;
if(config.isActive === true) {
  _config = config;
}
else{
   _config = {size: 10};
}
console.log(_config.size);
\end{lstlisting}
\end{minipage}
\caption{Example of pair of code snippets used in the interview. This case explores the use of the \emph{Ternary Operator atom of confusion}}
\label{code:ternary}
\end{figure*}

\begin{mq}
\emph{``I prefer to write [code using the \lhs version], but I think [the \rhs version]  is easier to read, especially for newer programmers".}
\end{mq}

\begin{mq}
\emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the [the code using the \lhs version] is easier to understand"}.
\end{mq}

\begin{mq}
\emph{``I think [the \lhs version] is easier to understand, but [the \rhs version] is what I would write"}
\end{mq}

This particular atom of confusion also opens up the possibility for a derivative construct that JavaScript allows which can be rather confusing, and that is the nested ternaries construct, in which the right-hand side of the a ternary operator can be another ternary construct. While nested ternaries remove the number of lines that would be necessary to construct using nested if-then-else statements, they can become quite taxing, cognitively speaking, to understand. Therefore, nested ternaries are a choice of atoms to be analysed in future work.

\rb{Poder\'{i}amos quantificar a existencia de emph{nesting} com code queries?}

The \emph{Pre-Increment} atom of confusion also caused such divide in one of the interviewees, as they regarded the \emph{\rhs version} (without atom of confusion) as simpler to understand, but would still opt to write code that contained the atom. In contrast to such opinion, one of the participants found the version that contained the atom of confusion more elegant, but recognized it was less readable, and was willing to sacrifice elegance for readability.

When analyzing the \emph{Logic as Control Flow} atom of confusion, one of the interviewees gave an example of his own experience on why one should avoid writing code with such atom:

\begin{mq}
\emph{``This one is interesting, because I have written code that looks like the [\lhs version (with the atom of confusion)], and my colleagues complained that it was difficult to understand. So nowadays I prefer to write code using the [\rhs version]."}
\end{mq}

Two of the atoms had unanimous preference for the versions that did not contain the atoms: \emph{Comma Operator} and \emph{Omitted Curly Braces and Indentation}. Regarding the first one, we could often notice during the interviews that the \emph{\lhs} (with the atom of confusion) caused significant confusion among the participants. Two remarks about the \emph{Comma Operator} atom of confusion are listed below:

\begin{mq}
\emph{``I just learned that [the \lhs version of this code] is possible. I did not even know it worked"}
\end{mq}

\begin{mq}
\emph{``The code in [the \lhs version] is unlikely to be understood unless the programmer knows \clang or \cpplang"}
\end{mq}

As for the \emph{Omitted Curly Braces and Indentation} atom of confusion, which, as we shall see in Section~\ref{sec:msr-results}, is frequent in JS code bases, one of the interviewees mentioned that, although they understand why one would opt to not use braces for simple if-then-else statements, he still advised against it, on grounds that:

\begin{mq}
\emph{``[I prefer the \rhs version of the code \ldots] If I want to see well-written, easily understandable code, then I also have to do my job. Therefore I believe that, since I do not know who is on the other end maintaining this code, and it could be any person with any level of expertise, then I try to write readable, easy-to-understand code"}
\end{mq}

This view is central to our research on why such small, yet tricky, constructs should be avoided whenever possible, since maintenance is one of the longer and most costly parts of the software development life cycle, and one cannot make any assumptions about the other programmers' experience.

\rb{que conclus\~{a}o extra\'{i}mos com essa parte do estudo?}

\subsection{Common Remarks From the Open-ended Questions} The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confuse code.

One of the participants mentioned that the use of JavaScript's prototype-based inheritance can make it difficult to understand code that, particularly when involving  deep prototype chains. This is a core feature of the language, and most high-level tools and frameworks abstract it away. Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.

When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate. One of the respondents regarded nested callbacks difficult even to write:

\begin{mq}
\emph{``Nested callbacks are very confusing. Even writing them can be confusing, let alone understanding them."}
\end{mq}

Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand 'asynchronous' programming in JavaScript. Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.

% citar uninitialized objects' types. floating point precision issues, 'this', type coercion

% \subsection{Common Misunderstandings About the Language}

A total of 40\% (n = 6) practitioners regarded the JavaScript language as quick and/or easy to learn, and 26.7\%(n = 4) mentioned that the abundance of frameworks, coupled with the abstractions they provide, make it so that one can develop quickly in the language. However, as mentioned by one of the interviewees, ``\emph{Developers who have started programming in JavaScript might miss out on important lower-level aspects, such as garbage collection, memory management, and processor management.}''.

Moreover, such abstractions can make it so that one starts to program using the language without actually understanding some of its basic aspects. For instance, 26.7\%(n = 4) said that the language is not typed, 13.3\%(n = 2) believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements. For the multi-purpose language that is one of the most, if not the most used one by practitioners
~\cite{}, a lack of understanding of the languages core features may be a problem worth investigating in future studies.

\rb{seria importante apresentar uma conclus\~{a}o em rela\c c\~{a}o a essa parte do estudo. Acredito que seja interessante discutir um pouco essa ``faca de dois gumes'' sobre a flexibilidade, mencionada abaixo. }

\adriano{Fiquei em duvida se vale citar os contras mais interessantes que os entrevistados falaram sobre JS: (1x) o jeito como ela lida com 'this', (que esta ligado a ela ser puramente orientada a objetos, sem classe); (2x)falta de precisao para tratamento de ponto flutuante, o que pode a tornar inadequada para tratamento de dados/ciencias puras; e a faca de dois gumes que e a flexibilidade da linguagem: muitos falavam isso como ponto positivo e negativo ao mesmo tempo. \todo[inline]{eu acho que podemos mencionar sim}}

\begin{table}[!htb]
    \centering
    \caption{Interviews Round 1 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}). Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
    \label{tab:interview-results1}
    \begin{tabular}{lrrr}\toprule
      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
      \cmidrule(lr){2-4}
         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
         Arithmetic as Logic     & 0  & 93.33  & 6.67  \\
         Assignment as Value     & 20 & 60     & 20    \\
         Comma Operator          & 0  & 100    & 0     \\
         Ternary Operator        & 60 & 26.67  & 13.3  \\
         Implicit Predicate      & 20 & 73.33  & 6.67  \\
         Logic as Control Flow   & 20 & 60     & 20    \\
         Omitted Curly Braces    & 0  & 100    & 0     \\
         Post Increment          & 20 & 73.33  & 6.67  \\
         Pre Increment           & 40 & 46.67  & 13.33 \\ \midrule
         \textsc{overall}        & 20 & 70.37  & 9.63  \\
         \bottomrule
    \end{tabular}
\end{table}

\begin{table}[!htb]
    \centering
    \caption{Interviews Round 2 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}). Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
    \label{tab:interview-results2}
    \begin{tabular}{lrrr}\toprule
      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
      \cmidrule(lr){2-4}
         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
         Property access         & 10  & 60  & 30  \\
         Object spread           & 50 & 50     & 0    \\
         Array spread            & 60  & 40    & 0     \\
         Arrow function          & 40 & 40     & 20  \\
         Array destructuring     & 40 & 40  & 20  \\
         Object destructuring    & 40 & 40     & 20    \\
         Type conversion         & 30  & 60    & 10     \\
         Change Literal encoding & 40 & 50  & 10  \\ \midrule
         \textsc{overall}        & 42.8 & 47.6  & 9.6  \\
         \bottomrule
    \end{tabular}
\end{table}

\subsection{Results of the third study: Mining open source repositories}
\label{sec:msr-results}

We mined \minedprojects open source JavaScript repositories to understand how often atoms of confusion arise in real software. Similarly to a previous study~\cite{DBLP:conf/msr/GopsteinZFC18}, which investigates the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion frequently arise in JavaScript open source systems.

The 3 most frequently found atoms occur in at least 80\% of the projects.
Considering the extremes, atom Ternary Operator was found in 89.69\% of the repositories, while atom Comma Operator occurred in only 11.34\% of them (as seen in Figure~\ref{fig:rate}).

% Figure~\ref{fig:rate} summarizes this finding, showing that the occurrence of atoms of confusion in JavaScript systems range from 11.34\% (Comma Operator) to 89.69\% (Ternary Operator).

%Figure~\ref{fig:atoms-occurrence} shows that the occurrence of in JavaScript systems range from

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{images/rate-1.pdf}
    \caption{Incidence of atoms of confusion across the \minedprojects JavaScript repositories we mined}
    \label{fig:rate}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{images/chart_ocurrence_kloc.png}
    \caption{The rate of each atom per thousand lines of code (KLOC) across all \minedprojects repositories}
    \label{fig:atoms-occurrence}
\end{figure}

Considering all JavaScript projects in our dataset, we found 
a total of \num{206317} atoms of confusion, though three atoms are responsible for 87.47\% of this total: Post Increment Expression, Omitted Curly Braces, and the Ternary Operator. The remaining atoms accounts for \num{25851} (12.53\%) of the total number of occurrences, and the Comma Operator and and Arithmetic as Control Flow are the atoms that arise less frequently (0.25\% in total with 373 and 164 occurrences, respectively).
%In our dataset of JavaScript projects, 
We found the Ternary Operator atom occurring more frequently than the Omitted Curly Braces, differently from what has been reported in a previous work~\cite{DBLP:conf/msr/GopsteinZFC18}.

The results of our survey and interviews suggest that the Ternary Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to our survey). Nonetheless, the Post Increment Expression and Omitted Curly Braces atoms are listed in the top three sources of misunderstanding 
(see Table~\ref{tab:difference-correctness}). As such, fixing the the Post Increment Expression and Omitted Curly Braces might solve most of the unclear code caused by atoms of confusion.

\rb{continuar desse ponto}

\subsection{Discussion}
\label{sec:discussion}

### 2020-11-09 PaperICSME.tex
Initial Overleaf Import

\documentclass[10pt,conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{supertabular}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{colortbl}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage{listings}
\usepackage{color}
\usepackage{scalefnt}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xspace}
\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{tabularx} 
\usepackage{multicol}
\usepackage{listings}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{textcomp}
\usepackage{siunitx}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{array}
\usepackage{listings}
\usepackage{color}
\usepackage{scalefnt}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{appendix}
\usepackage[framemethod=TikZ]{mdframed}
\definecolor{oldlace}{rgb}{0.99, 0.96, 0.9}
\newmdenv [ %
 skipabove=\topsep,
 skipbelow=\topsep,
 leftmargin       = 2,
 rightmargin      = 2,
 splittopskip     = \topskip]{mh}

\newmdenv [ %
 skipabove=\topsep,
 skipbelow=\topsep,
 roundcorner = 5pt,
 leftmargin = 2,
 rightmargin = 2,
 % backgroundcolor = oldlace,
 innertopmargin = 3,
 splittopskip = 3]{mq}

\definecolor{javared}{rgb}{0.6,0,0} % for strings
\definecolor{javagreen}{rgb}{0.25,0.5,0.35} % comments
\definecolor{javapurple}{rgb}{0.5,0,0.35} % keywords
\definecolor{javadocblue}{rgb}{0.25,0.35,0.75} % javadoc
\lstset{
  basicstyle=\footnotesize\tt,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,                    % adds a frame around the code
  language=Java,                 % the language of the code
  keywordstyle=\bf,
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  tabsize=2,                       % sets default tabsize to 2 spaces
  keywordstyle=\color{javapurple}\bfseries,
  stringstyle=\color{javared},
  commentstyle=\color{javagreen},
  morecomment=[s][\color{javadocblue}]{/**}{*/},
  numbers=left,
  numberstyle=\tiny\color{black},
  stepnumber=1,
}    
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

\newcommand{\clang}{C\nolinebreak\xspace}

\newcommand{\cpplang}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\small\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\small\bf +}\xspace}

% TODO: Completar
\newcommand{\minedprojects}{X\xspace}

\newcommand{\rb}[1]{\todo[inline]{{\bf (rb note) }#1}}
\newcommand{\diego}[1]{\todo[inline,color=green]{{\bf (diego note) }#1}}
\newcommand{\adriano}[1]{\todo[inline,color=yellow]{{\bf (adriano note) }#1}}
\newcommand{\castor}[1]{\todo[inline,color=pink]{{\bf (castor note) }#1}}

\lstset{language=C}
\lstset{
 morekeywords={printf}
}

\newcommand{\lhs}{left-hand side\xspace}
\newcommand{\rhs}{right-hand side\xspace}

\lstdefinelanguage{JavaScript}[]{Java}{
  morekeywords={let, console.log},
  moredelim=[is][\textcolor{darkgray}]{\%\%}{\%\%},
  moredelim=[il][\textcolor{darkgray}]{§§}
}

\begin{document}
\input{atom-mdframe-config}

\title{On The Impact of Atoms of Confusion in JavaScript
Code}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

\maketitle

% Número das páginas. Remover para camera-ready
% útil para nós e Program Chair que pode ver o total de páginas facilmente
\thispagestyle{plain}
\pagestyle{plain}

\begin{abstract}
{\bf Context.} Evolving legacy code is a challenging task. It is even more difficult when the code has been poorly written, imposing significant cognitive load on developers when trying to familiarize themselves with a legacy code base. 
\noindent{\bf Problem.} 
%Besides lack of experience and other factors, 
Previous research works have discussed the impact of small code patterns that contribute to misunderstanding the source code written in statically typed languages (e.g., \clang, \cpplang, and Swift). However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamically typed languages. 
\noindent{\bf Research Method and Goal.} In this work, we fill this research gap by presenting the results of a mixed-methods research:
a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort. Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: 
%removing the atoms of confusion lead to an improvement above 15\% in 6 out of 9 atoms we consider in our study when developers try to predict the outcomes of code examples. 
developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atoms.
Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the \texttt{comma operator} is present in 11\% of the projects in our dataset; while five other atoms appear in more than 60\% of the projects---including the atoms \texttt{assignment as value} and \texttt{omitted curly braces}, which appear \num{8253} and \num{53357} times, respectively. Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code. For instance, even experienced JavaScript developers agree that it is hard to understand code that uses callback functions, even though we found widespread use of \texttt{callbacks} in our projects dataset.  
\noindent{\bf Implications.} Altogether, our findings could help developers write cleaner JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code. 
\end{abstract}

\begin{IEEEkeywords}
datasets, neural networks, gaze detection, text tagging
\end{IEEEkeywords}

\input{1_Introduction}
\input{2_Background}
\input{3_Method}
\input{4_Result}
\input{5_Threats}
\input{6_Conclusion}
\input{7_Atoms}

\section*{Acknowledgment}

\section*{References}
\bibliographystyle{IEEEtran}
\bibliography{reference}

\vspace{12pt}
\end{document}

### 2020-11-09 0_Abstract.tex
initial commit

--- 2020-11-09 PaperICSME.tex
@@
-\documentclass[10pt,conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{supertabular}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{colortbl}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\usepackage{listings}
-\usepackage{color}
-\usepackage{scalefnt}
-\usepackage{multirow}
-\usepackage{pifont}
-\usepackage{xspace}
-\usepackage{todonotes}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{tabularx} 
-\usepackage{multicol}
-\usepackage{listings}
-\usepackage{xspace}
-\usepackage{booktabs}
-\usepackage{siunitx}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{graphics}
-\usepackage{textcomp}
-\usepackage{siunitx}
-\usepackage{xcolor}
-\usepackage{csquotes}
-\usepackage{array}
-\usepackage{listings}
-\usepackage{color}
-\usepackage{scalefnt}
-\usepackage{multirow}
-\usepackage{pifont}
-\usepackage{appendix}
-\usepackage[framemethod=TikZ]{mdframed}
-\definecolor{oldlace}{rgb}{0.99, 0.96, 0.9}
-\newmdenv [ %
- skipabove=\topsep,
- skipbelow=\topsep,
- leftmargin       = 2,
- rightmargin      = 2,
- splittopskip     = \topskip]{mh}
-
-\newmdenv [ %
- skipabove=\topsep,
- skipbelow=\topsep,
- roundcorner = 5pt,
- leftmargin = 2,
- rightmargin = 2,
- % backgroundcolor = oldlace,
- innertopmargin = 3,
- splittopskip = 3]{mq}
-
-\definecolor{javared}{rgb}{0.6,0,0} % for strings
-\definecolor{javagreen}{rgb}{0.25,0.5,0.35} % comments
-\definecolor{javapurple}{rgb}{0.5,0,0.35} % keywords
-\definecolor{javadocblue}{rgb}{0.25,0.35,0.75} % javadoc
-\lstset{
-  basicstyle=\footnotesize\tt,        % the size of the fonts that are used for the code
-  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
-  breaklines=true,                 % sets automatic line breaking
-  captionpos=b,                    % sets the caption-position to bottom
-  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
-  frame=single,                    % adds a frame around the code
-  language=Java,                 % the language of the code
-  keywordstyle=\bf,
-  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
-  showstringspaces=false,          % underline spaces within strings only
-  showtabs=false,                  % show tabs within strings adding particular underscores
-  tabsize=2,                       % sets default tabsize to 2 spaces
-  keywordstyle=\color{javapurple}\bfseries,
-  stringstyle=\color{javared},
-  commentstyle=\color{javagreen},
-  morecomment=[s][\color{javadocblue}]{/**}{*/},
-  numbers=left,
-  numberstyle=\tiny\color{black},
-  stepnumber=1,
-}    
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-\newcommand{\clang}{C\nolinebreak\xspace}
-
-\newcommand{\cpplang}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\small\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\small\bf +}\xspace}
-
-% TODO: Completar
-\newcommand{\minedprojects}{X\xspace}
-
-\newcommand{\rb}[1]{\todo[inline]{{\bf (rb note) }#1}}
-\newcommand{\diego}[1]{\todo[inline,color=green]{{\bf (diego note) }#1}}
-\newcommand{\adriano}[1]{\todo[inline,color=yellow]{{\bf (adriano note) }#1}}
-\newcommand{\castor}[1]{\todo[inline,color=pink]{{\bf (castor note) }#1}}
-
-\lstset{language=C}
-\lstset{
- morekeywords={printf}
-}
-
-\newcommand{\lhs}{left-hand side\xspace}
-\newcommand{\rhs}{right-hand side\xspace}
-
-\lstdefinelanguage{JavaScript}[]{Java}{
-  morekeywords={let, console.log},
-  moredelim=[is][\textcolor{darkgray}]{\%\%}{\%\%},
-  moredelim=[il][\textcolor{darkgray}]{§§}
-}
-
-\begin{document}
-\input{atom-mdframe-config}
-
-\title{On The Impact of Atoms of Confusion in JavaScript
-Code}
-
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-}
-
-\maketitle
-
-% Número das páginas. Remover para camera-ready
-% útil para nós e Program Chair que pode ver o total de páginas facilmente
-\thispagestyle{plain}
-\pagestyle{plain}
-
 \begin{abstract}
-{\bf Context.} Evolving legacy code is a challenging task.
-It is even more difficult when the code has been poorly written, imposing significant cognitive load on developers when trying to familiarize themselves with a legacy code base.
+{\bf Context.} Engineering maintainable software is essential and must take into account \emph{how understandable the code is} from a human perspective, and spending much cognitive effort might discourage a developer when trying to familiarize herself with a legacy codebase.
 \noindent{\bf Problem.} 
@@
 Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
-\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: 
+\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead to a code that is hard to understand: 
 %removing the atoms of confusion lead to an improvement above 15\% in 6 out of 9 atoms we consider in our study when developers try to predict the outcomes of code examples. 
-developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atoms.
+developers correctly predict the outcome of at least 15\% more when an atom of confusion is not present for 6 out of 9 atoms.
 Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the \texttt{comma operator} is present in 11\% of the projects in our dataset; while five other atoms appear in more than 60\% of the projects---including the atoms \texttt{assignment as value} and \texttt{omitted curly braces}, which appear \num{8253} and \num{53357} times, respectively.
@@
 
-\begin{IEEEkeywords}
-datasets, neural networks, gaze detection, text tagging
-\end{IEEEkeywords}
+\begin{CCSXML}
+<ccs2012>
+ <concept>
+  <concept_id>10010520.10010553.10010562</concept_id>
+  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+  <concept_significance>500</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010575.10010755</concept_id>
+  <concept_desc>Computer systems organization~Redundancy</concept_desc>
+  <concept_significance>300</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010553.10010554</concept_id>
+  <concept_desc>Computer systems organization~Robotics</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10003033.10003083.10003095</concept_id>
+  <concept_desc>Networks~Network reliability</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+</ccs2012>
+\end{CCSXML}
 
-\input{1_Introduction}
-\input{2_Background}
-\input{3_Method}
-\input{4_Result}
-\input{5_Threats}
-\input{6_Conclusion}
-\input{7_Atoms}
+\ccsdesc[500]{Computer systems organization~Embedded systems}
+\ccsdesc[300]{Computer systems organization~Redundancy}
+\ccsdesc{Computer systems organization~Robotics}
+\ccsdesc[100]{Networks~Network reliability}
 
-\section*{Acknowledgment}
-
-\section*{References}
-\bibliographystyle{IEEEtran}
-\bibliography{reference}
-
-\vspace{12pt}
-\end{document}
+%%
+%% Keywords. The author(s) should pick words that accurately describe
+%% the work being presented. Separate the keywords with commas.
+\keywords{datasets, neural networks, gaze detection, text tagging}

### 2020-11-09 4_Result.tex
initial commit

@@
 
-In this section we present the results of two rounds of interviews with practitioners who agreed to participate in the interviews.In the first interview, a total of 15 practitioners were surveyed.
+In this section we present the results of two rounds of two rounds of interviews with practitioners who agreed to participate in the interviews.In the first interview, a total of 15 practitioners were surveyed.
 In this phase, we collected information regarding programming experience, familiarity with JavaScript, and we also collected their opinions regarding nine candidates to atoms of confusion.
@@
 
-In the second round of interviews, fourteen of the fifteen developers who participated in the first round were shown another 8 candidates to atom of confusion whose frequency we observed during the repository mining phase.
+In the second round of interviews, ten of the fifteen developers who participated in the first round were shown another 8 candidates to atom of confusion whose frequency we observed during the repository mining phase.
 The same procedure of showing code that behaves exactly the same and surveying for how easy to understand each snippet was applied in this phase.
 
-Supporting the results of the survey,Tables \ref{tab:interview-results1} and \ref{tab:interview-results2} show that for twelve out of the seventeen scenarios surveyed, the respondents prefer the version of the code without the atom of confusion, and in no case the \emph{neutral} ratio was higher than the option for the version without atoms of confusion.
-An example entry is contained in Appendix~\ref{sec:appendix-atoms}.
+Supporting the results of the survey,Tables \ref{tab:interview-results1} and \ref{tab:interview-results2} show that for twelve out of the seventeen scenarios surveyed, the respondents prefer the version of the code without the atom of confusion, and in no case the \emph{neutral} ratio was higher than the option for the version without atoms of confusion.An example entry is contained in Appendix~\ref{}.
 For a full listing of the code snippets, visit the paper repository.

### 2020-11-09 5_Threats.tex
initial commit

\section{Threats to Validity}
\label{threat}

Since our research was conducted online with unknown participants, we had no way of confirming their levels of education and experience. Also, as we briefly discussed in Section 4.3, we did not have a way to prevent respondents from cheating, such as running the code on an interpreter, or consulting other people. This presents a threat to the validity of the histograms presented in Figures \ref{fig:degree} and \ref{fig:degree}.

Another threat to our results lies in the fact that some constructs might not be frequent, or present at all, in JavaScript code bases. The Automatic Semicolon Insertion and the Comma Operator atoms were not found in the code samples we consulted when formulating the questions, and we had to create blocks of code containing the atoms. The work done in Chapter 5 could serve as a starting point for detecting these atoms in a large data set of JavaScript files.

Regarding hypotheses H.2 \todo[inline]{Linkar as Hipoteses} and research question RQ.2, a potential problem with our method is that there was no strong incentive for subjects to think thoroughly about the questions. We observed a lack of engagement when we ran the survey with undergraduate students. Although our final subjects were voluntarily partaking in the survey, we could not be sure that, after some time taking the survey, respondents would become tired and stop thinking clearly about the code. As questions were assigned randomly, and contained different levels of difficulty, comparing a subject's answer with their other questions would not provide a reliable metric.

In one of the atoms, namely the Omitted Curly Braces, we intentionally removed indentation from the code, which is highly unusual, given that many programmers use automatic formatting in their code editors. This can introduce some level of artificiality to this atom's question. On the other hand, while describing JavaScript's grammar to develop the parser we used for the work in Chapter 5, we discovered that the language allows the programmer to omit the curly braces after \textit{if} statements, and insert multiple statements in the following line. This fact itself might constitute a source of confusion, which we leave to analyse in our future endeavors.

### 2020-11-09 6_Conclusion.tex
initial commit

\section{Conclusions}
\label{conclusion}

\rb{draft. apenas movi esse texto que estava em settings para ca. pode servir de inspiracao para responder algumas das questoes de pesquisa.}

{\color{blue}Based on these results, we provide evidence that, although the atoms of confusion considered in our study do not take a much higher amount of time to predict the output of the code, their impact on program comprehension is, in almost all of our cases, highly significant.}

### 2020-11-09 3_Method.tex
update the repository

@@
 
-The main goal of this research is to investigate the impact of atoms of confusion in JavaScript code comprehension. As such, in this paper we answer the following general research questions:
+The main goal of this research is to investigate the impact of atoms of confusion in JavaScript code comprehension. As such, in this paper we answer the following research questions:
 
@@
     \item Do JavaScript developers identify atoms of confusion as contributing to program misunderstanding? 
-    \item What is the impact of atoms of confusion on developers' comprehension of JavaScript code? 
+    \item What is the impact of atoms of confusion on the comprehension of JavaScript code? 
     \item What are the particular JavaScript idioms and language constructs that might cause source code misunderstanding?
@@
 In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice.
-Our motivation to detect the presence and impact of atoms of confusion is also motivated by the possibility of laying the foundations for the implementation of libraries that can automatically transform code into its simpler versions---though we postpone these results to a future research work.
+Our motivation to detect the presence and impact of atoms of confusion is also motivated by the possibility of laying the foundations for the implementation of libraries that can automatically transform code into its cleaner versions---though we postpone these results to a future research work.
 This might lead to a new catalog of refactoring for removing atoms of confusion.
@@
 
- \subsection{Settings for the first study: a survey}\label{sec:survey-settings}\castor{Eu mudaria os títulos desta sub-seção e das duas próximas para "First study: Survey", "Second study: Mining open source repositories", etc.
+ \subsection{First Study: Survey}\label{sec:survey-settings}\castor{Eu mudaria os títulos desta sub-seção e das duas próximas para "First study: Survey", "Second study: Mining open source repositories", etc.
 Vale a pena manter "First study", etc.?}
@@
 To achieve such a design goal, controlling the effect of experience and individual code snippets, we resorted to the \textit{Latin Square Design} \cite{Hunter-Experimenters}.
-Using this design we create a $2 x 2$ matrix in which each row represents a subject and each column indicates the set of code snippets.
+Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of code snippets.
 The design of each square (a replica) is such that no treatment is repeated in the same row or column.
@@
 \subsubsection{Survey audience}
+
 \diego{Esses detalhes são possíveis candidatos para remoção, se precisarmos de espaço}
@@
 
-\subsection{Settings for the second study: Mining open source repositories}
+\subsection{Second Study: Mining Software Repositories}
 
@@
 
-\subsection{Settings for the third study: interview with practitioners}
+\subsection{Third Study: Interviews with Practitioners}
 

### 2020-11-11 1_Introduction.tex
mais uma revisao da secao de introducao

@@
 At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''\footnote{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}.
-It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST, ClientSideBugsJS:TSE}, and asynchronous operations.
+It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.
 All these
 %diverse characteristics and capabilities
-features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST, ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.
+features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.
 
-Our research aims to help practitioners identify potentially confusing snippets of code and replace them with clearer code.
-% Developers could improve their codebases' understandability by following our recommendations.
-To achieve that, we characterize a set of 9 JavaScript \emph{atoms of confusion}; small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.\castor{I think it would be really nice to have an example here.
+Our research aims to help practitioners identify potentially confusing snippets of code, so that
+developers could replace them with clearer code.
+To achieve that, we investigate {\color{red}XYZ} JavaScript idioms that
+might compromize program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in
+JavaScript---that is, small
+pieces of code that hinder developers' understanding of the source code of a program
+for which there are functionally equivalent alternatives that are easier
+to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
+\castor{I think it would be really nice to have an example here.
 Maybe even a figure with a left-hand and a right-hand side.
@@
 Among the atoms in this work, X are shared between JavaScript and \clang/\cpplang (e.g., pre/post-increment), while Y others are specific to JavaScript, e.g., the atom Assignment as Value.
-Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/He19,DBLP:conf/sigsoft/GopsteinIYDZYC17} has already explored the impact of atoms of confusion on program understanding,
+Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17} has
+already explored the impact of atoms of confusion on program understanding,
 and confirmed they negatively impact source code understandability.
-However, they were constrained to statically typed languages.
-To the best of our knowledge, we are the first to explore atoms of confusion in the context of a dynamically typed language (JavaScript).\castor{Review the number of atoms, X, and Y.}
+However, they were constrained to statically typed languages (e.g., C and C++).
+To the best of our knowledge, we are the first to explore atoms of confusion
+in the context of a dynamically typed language (JavaScript).\castor{Review the number of atoms, X, and Y.}
 
-\diego{"Understanding source code comments at large-scale"\cite{DBLP:conf/sigsoft/He19} (acima): 
-embora relacionado não me parece um estudo de átomos de confusão}
-
-To understand whether an atom of confusion introduces misunderstandings, we surveyed 150 developers from the Reddit community.
+To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 developers from the Reddit community.
 Their task was to predict the output of short snippets of code, with half of them containing atoms of confusion.
-We found that the presence of the Comma Operator atom can increase the average time to predict the output by 65\%, while the behavior of snippets containing the Pre-Increment atom were predicted 44\% faster.
+We found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of 9 atoms).
+The presence of the Comma Operator atom, for instance, can increase the average time to predict the output by 65\%, while the behavior of snippets containing the Pre-Increment atom were predicted 44\% faster.
 An atom's presence---or lack of---in open-source repositories may reveal cases where developers should be more or less careful when writing code.
@@
 To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15 experienced professional developers.
-Participants preferred code snippets \textit{without} an atom in 70\% of the cases, therefore corroborating to the survey findings.\castor{It would be nice to have results about correctness.}
+Participants preferred code snippets \textit{without} an atom in 70\% of the cases, therefore corroborating to the survey findings.
+During the MSR and interview studies, we also explored additional atom candidates that are specific to the
+JavaScript language.
 

### 2020-11-11 2_Background.tex
additional review on sections 1, 2

@@
 
-However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions--- when these exist- allowed by the language or framework they are working with.
+However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions---when these exist---allowed by the language or framework they are working with.
 Although this can be, to an extent, a personal matter, since different individuals might hold different opinions on whether a given syntax is difficult to understand, it is possible to empirically investigate if there exist programming languages' constructs or idioms that are considerably harder to understand, leading readers of the code to frequently make incorrect predictions of its behavior.
@@
 Their work laid an empirical framework to identify such atoms.
-In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atoms were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.\castor{Mudei um pouco o protocolo para identificar átomos.
-Tinha achado a explicação anterior confusa.}
+In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atoms were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.
 %By having subjects evaluate code where in the expected output relied exclusively on evaluating the result of previously known confusing programs that had been written in \clang or \cpplang, and asking the participants to predict the output of a set of code snippets where the atoms had been removed, the authors were able to isolate and measure the impact of such confusing atoms.
@@
 This occurs because a leading \texttt{0} in a number literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers.
-After formulating a list of 19 potential atoms, the authors were able to identify 15 snippets that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed.
+After formulating a list of 19 potential atoms, Gopstein et al.
+\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 snippets that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed.
 The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase.
-Our first study replicates this study, for all atoms that have a counterpart in JavaScript.\castor{Is ``replicates'' the best word here?
-From what I could grasp, there are some differences between the protocols.} Gopstein et al.
+Our first study follows the empirical framework discussed in~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, for all atoms that have a counterpart in JavaScript.
+Gopstein et al.
 \cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on the different nature of confusion amongst the atoms, noting that, for some atoms, a particular incorrect answer was prevalent, whilst there was a distribution of incorrect ones in other atoms.
@@
 
-\castor{Provavelmente devemos reduzir o espaço tomado pelos quatro trabalhos acima, a depender da existência de outros. O de Linguistic Antipatterns talvez até devesse sair.}
-\rb{Incluir os trabalhos do grupo do Marcio Ribeiro e do Fernando Castor. Quais outros trabalhos deveriamos incluir aqui?}
+%% \castor{Provavelmente devemos reduzir o espaço tomado pelos quatro trabalhos acima, a depender da existência de outros. O de Linguistic Antipatterns talvez até devesse sair.}
+%% \rb{Incluir os trabalhos do grupo do Marcio Ribeiro e do Fernando Castor. Quais outros trabalhos deveriamos incluir aqui?}

### 2020-11-12 1_Introduction.tex
additional review on sections 1, 2
add a code snippet to make the notion of atoms-of-confusion clearer

@@
 To achieve that, we investigate {\color{red}XYZ} JavaScript idioms that
-might compromize program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in
+might compromise program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in
 JavaScript---that is, small
@@
 for which there are functionally equivalent alternatives that are easier
-to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
-\castor{I think it would be really nice to have an example here.
-Maybe even a figure with a left-hand and a right-hand side.
-What do you say?
-If yes, I
-'d go for post increment or assignment as value.
-If not a figure, the text could be as follow: An example of atom of confusion identified by previous work~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} is the use \texttt{if} statements without the accompanying curly braces.
-The alternative, in this case, is simply to always use the curly braces.}
+to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For instance, consider
+the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}
+for presentation purpose only).
+The code in the left of the figure contains two atoms of confusion discussed in a previous work~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}.
+On the right we present a cleaned version of the code, that is, without atoms of confusion.
+According to
+Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} removing
+the atom \emph{Logic as Control Flow} and \emph{Post-increment} improves the accuracy of
+program understanding tasks in  41\% and 34\%, respectively.
+
 Among the atoms in this work, X are shared between JavaScript and \clang/\cpplang (e.g., pre/post-increment), while Y others are specific to JavaScript, e.g., the atom Assignment as Value.
@@
 and confirmed they negatively impact source code understandability.
-However, they were constrained to statically typed languages (e.g., C and C++).
+However, they were constrained to statically typed languages (e.g., \clang and \cpplang).
 To the best of our knowledge, we are the first to explore atoms of confusion
 in the context of a dynamically typed language (JavaScript).\castor{Review the number of atoms, X, and Y.}
+
+\begin{figure*}
+\noindent\begin{minipage}{.45\textwidth}  
+\begin{lstlisting}[language=JavaScript]
+res = properties.reduce((s, b) => {
+  b && s++
+  return s
+}
+\end{lstlisting}
+\end{minipage}\hfill
+\begin{minipage}{.45\textwidth}
+\begin{lstlisting}[language=JavaScript]
+res = properties.reduce((s, b) => {
+  if(b) {
+    s = s + 1
+  }
+  return s
+}
+\end{lstlisting}
+\end{minipage}
+\label{fig:lst01}
+\captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project. Line 2 in the left hand side of the figure contains two
+  atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic
+  as Control Flow} and \emph{Post-increment}.
+In the right hand side, we present the alternative code without atoms of confusion.
+We could simplify the code on the left even more, by replacing the \texttt{reduce} recursive pattern by a functional composition between the \texttt{filter} and
+\texttt{length} patterns (e.g., \texttt{res = properties.filter(p => p).length}).}
+\end{figure*}
 

### 2020-11-13 1_Introduction.tex
text improvements

@@
 In the right hand side, we present the alternative code without atoms of confusion.
-We could simplify the code on the left even more, by replacing the \texttt{reduce} recursive pattern by a functional composition between the \texttt{filter} and
-\texttt{length} patterns (e.g., \texttt{res = properties.filter(p => p).length}).}
+We could simplify the code on the left even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p).length}).}
 \end{figure*}

### 2020-11-16 3_Method.tex
updates na secao de MSR

@@
 To this end, we first collected the most popular GitHub repositories that are primarily written is JavaScript.
-We measured popularity using the project's stargazers, and selected only projects with stargazers\_count > 100.
+We measured popularity using the project's stargazers, and selected only projects with stargazers\_count greater than 100.
 This metric, available on  GitHub API, represents the number of stars a project received from users of the platform.
 The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs}.
+We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.
 
-After filtering out JavaScript project candidates, in the second step we built a curate dataset comprising the top 100 most popular repositories.
+After filtering out JavaScript project candidates, in the second step we built a curate dataset comprising the top \minedprojects repositories.
 Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}.
 Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research.
-The size of the projects range from small ones (883 liens of code) to complex systems with more than 10 MLOC.
-All projects in our dataset have at least \num{1161} forks and \num{23643} stars.
+The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
+All projects in our dataset have at least \num{1244} forks and \num{23672} stars.
 We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.
@@
 In the third step we mined atoms of confusion from the repositories in our curate dataset.
-We written source code queries usng the CodeQL language~\cite{moor:gttse2007}, in order to filter the atoms of confusion.
+We written source code queries using the CodeQL language~\cite{moor:gttse2007}, in order to filter the atoms of confusion.
 CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as Java and JavaScript).
@@
              & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ \hline
- Lines of Code           & \num{883}  & \num{14939} & \num{39520} & \num{281088.46} & \num{145990.75} & \num{10149685} \\
- Num. of Contributors    & \num{6}   & \num{121} & \num{243} & \num{431.82} & \num{436.75} & \num{4042} \\
- Num. of Forks           & \num{1161}   & \num{3365} & \num{5711} & \num{8201.96} & \num{8438.25} & \num{68826} \\
- Num. of Stars        & \num{23643} & \num{27542.25} & \num{35687.5} & \num{45284.86} & \num{47963.25} & \num{310725} \\
+ Lines of Code           & \num{5543}  & \num{16060.25} & \num{36161.5} & \num{111432.33} & \num{111182.50} & \num{1278405} \\
+ Num. of Forks     & \num{1244}   & \num{3029.75} & \num{6078} & \num{8906.24} & \num{9260.50} & \num{68849} \\
+ Num. of Contributors  & \num{6}   & \num{184} & \num{285.5} & \num{533.44} & \num{515.75} & \num{4047} \\
+ Num. of Stars        & \num{23672} & \num{27055} & \num{34990} & \num{46919.51} & \num{48518} & \num{310935} \\
+
     \hline

### 2020-11-16 4_Result.tex
updates na secao de MSR

@@
 We mined \minedprojects open source JavaScript repositories to understand how often atoms of confusion arise in real software.
-Similarly to a previous study~\cite{DBLP:conf/msr/GopsteinZFC18}, which investigates the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion frequently arise in JavaScript open source systems.
+Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18}~\cite{DBLP:journals/ese/MedeirosLAAKRG19}, which investigates the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion frequently arise in JavaScript open source systems.
 
-The 3 most frequently found atoms occur in at least 80\% of the projects.
-Considering the extremes, atom Ternary Operator was found in 89.69\% of the repositories, while atom Comma Operator occurred in only 11.34\% of them (as seen in Figure~\ref{fig:rate}).
+The 5 most frequently found atoms occur in at least 80\% of the projects.
+Considering the extremes, atoms Ternary Operator and Access Property were found in all repositories, while atom Comma Operator occurred only in 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
 
@@
 
-\begin{figure}
+% latex table generated in R 4.0.3 by xtable 1.8-4 package
+% Fri Nov 13 11:28:56 2020
+\begin{table}[!htb]
 \centering
-    \includegraphics[width=\columnwidth]{images/rate-1.pdf}
-    \caption{Incidence of atoms of confusion across the \minedprojects JavaScript repositories we mined}
-    \label{fig:rate}
-\end{figure}
+ \caption{Summary of atoms occurrences on our dataset}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{images/chart_ocurrence_kloc.png}
-    \caption{The rate of each atom per thousand lines of code (KLOC) across all \minedprojects repositories}
-    \label{fig:atoms-occurrence}
-\end{figure}
+\setlength\tabcolsep{2pt} % default value: 6pt
+\label{tab:occurrences-summary}
+\begin{tabular}{lrcc}
+
+  \hline
+Atom & Projects & Ocurrences/KLOC & Category \\ 
+ \hline
+Ternary Operator & 100\% & 10.16 &  highly used \\ 
+  Access Property & 100\% & 3.63 &  highly used \\ 
+  Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
+  Post Increment & 90.28\% & 5.62 &  highly used \\ 
+  Arrow Function & 83.33\% & 21.10 &  highly used \\ 
+  Change Literal Encoding & 76.39\% & 21.43 &  highly used \\ 
+  Object Destructing & 72.22\% & 2.86 &  highly used \\ 
+  Pre-Increment & 83.33\% & 1.04 & commonly used \\ 
+  Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
+  Implicit Operand Conversion & 79.17\% & 0.80 & commonly used \\ 
+  Array Destructing & 62.50\% & 0.47 & commonly used \\ 
+  Array Spread & 59.72\% & 0.31 & commonly used \\ 
+  Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
+  Object Spread & 55.56\% & 0.64 & commonly used \\ 
+  Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
+  Comma Operator & 15.28\% & 0.03 & little used \\ 
+   \hline
+\end{tabular}
+\end{table}
 
 Considering all JavaScript projects in our dataset, we found 
-a total of \num{206317} atoms of confusion, though three atoms are responsible for 87.47\% of this total: Post Increment Expression, Omitted Curly Braces, and the Ternary Operator.
-The remaining atoms accounts for \num{25851} (12.53\%) of the total number of occurrences, and the Comma Operator and and Arithmetic as Control Flow are the atoms that arise less frequently (0.25\% in total with 373 and 164 occurrences, respectively).
+a total of \num{615364} atoms of confusion, though five atoms are responsible for 84.65\% of this total: Arrow Function, Change Literal Encoding, Ternary Operator, Omitted Curly Braces and the Post Increment.
+The remaining atoms accounts for \num{94450} (15.35\%) of the total number of occurrences, and the Comma Operator and Arithmetic as Logic are the atoms that arise less frequently (0.0064\% in total with 232 and 165 occurrences, respectively).
+
+Regarding frequency, we proceeded with a classification of atom occurrences according to the metrics established by a previous study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}.
+As such, we found 7 atoms as highly used (43.75\%), 7 as commonly used (43.75\%) and only 2 atoms as little used (12.50\%).
+We also calculated the frequency of each atom per KLOC and found atoms Change Literal Encoding and Arrow Function as the most occurring in practice (21.43 and 21.10 occurrences per KLOC, respectively).
 %In our dataset of JavaScript projects, 
-We found the Ternary Operator atom occurring more frequently than the Omitted Curly Braces, differently from what has been reported in a previous work~\cite{DBLP:conf/msr/GopsteinZFC18}.
+%We found the Ternary Operator atom occurring more frequently than the Omitted Curly Braces, differently from what has been reported in a previous work~\cite{DBLP:conf/msr/GopsteinZFC18}. %
 
@@
 Nonetheless, the Post Increment Expression and Omitted Curly Braces atoms are listed in the top three sources of misunderstanding
-(see Table~\ref{tab:difference-correctness}). As such, fixing the the Post Increment Expression and Omitted Curly Braces might solve most of the unclear code caused by atoms of confusion.
+(see Table~\ref{tab:difference-correctness}). As such, fixing the the Post Increment and Omitted Curly Braces might solve most of the unclear code caused by atoms of confusion.
 

### 2020-11-16 3_Method.tex
revisao da secao de metodologia

@@
 In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice.
-Our motivation to detect the presence and impact of atoms of confusion is also motivated by the possibility of laying the foundations for the implementation of libraries that can automatically transform code into its cleaner versions---though we postpone these results to a future research work.
-This might lead to a new catalog of refactoring for removing atoms of confusion.
-To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and an effort of mining JavaScript software repositories.
+Answering these research questions also lay the foundations for the implementation of libraries that can automatically transform code into its cleaner versions---though we postpone these results to a future research work.
+To answer these research questions we conduct a mixed-methods study, including {\color{red}two surveys}, a set of interviews, and an effort of mining JavaScript software repositories.
 
- \subsection{First Study: Survey}\label{sec:survey-settings}\castor{Eu mudaria os títulos desta sub-seção e das duas próximas para "First study: Survey", "Second study: Mining open source repositories", etc.
-Vale a pena manter "First study", etc.?}
-% \subsubsection{The Atoms Considered In This Work}
+\subsection{First Study: Survey-I}
 
-In the first study, our goal was to collect evidence to answer two specific questions: 
-\begin{itemize}
-    \item \emph{Do code snippets that contain atoms of confusion produce a higher misunderstanding rate when programmers try to predict their outcome?}
-    \item \emph{Do code snippets with atoms of confusion require programmers to take longer to predict their output?}
-\end{itemize}
-
-To address these questions we conducted a survey to assess the impact of atoms of confusion on understanding JavaScript code.
-Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that may also exist in JavaScript programs.
-{\color{red}In addition, we consider a particular atom that is exclusive to JavaScript in this first study}.
+In the firt study we explore the impact of atoms of confusion on
+understanding JavaScript code, by answering two questions: (a) \emph{Do code snippets that contain atoms of confusion produce a higher misunderstanding rate when programmers try to predict their outcome?} and (b) \emph{Do code snippets with atoms of confusion require programmers to take longer to predict their output?}
+Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} whose idioms also exist in JavaScript programs.
 Appendix~\ref{sec:appendix-atoms} summarizes all atoms we consider in our research.
@@
 
-The design of our first study aims to block two variables (developer experience and the code snippets) and considers two treatments: the presence or absence of 
+The design of our first study blocks two variables
+(developer experience and the code snippets) and
+considers two treatments: the presence or absence of 
 atoms of confusion within the code snippets. 
@@
 The design of each square (a replica) is such that no treatment is repeated in the same row or column.
-For example, if a given subject (P1) is asked to predict the output of code snippets that contain atom candidates \texttt{\{1,2,3,4,5\}}, then, when answering questions about non-confusing code snippets, they will only be presented with non-confusing versions of atom candidates \texttt{\{6,7,8,9,10\}}.
-Furthermore, a given subject (P2), which constitutes the second row of our example square, will be asked questions about the non-confusing versions of \texttt{\{1,2,3,4,5\}}, and will answer questions about confusing snippets of \texttt{\{6,7,8,9,10\}}.
-By doing that, we guarantee that all of our 20 snippets are contained within each square, and that each configuration occurs only once within a square.
+For example, considering that we have a set of 10
+code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about non-confusing code snippets, they will only be presented with non-confusing versions of the code samples $s_6, s_7,..., s_{10}$.
+Furthermore, a given subject (P2), which constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing snippets of for $s_6, s_7,..., s_{10}$.
+By doing that, we guarantee that all versions of code snippets are contained within each square, and that each configuration occurs only once within a square.
 Figure \ref{fig:latinsquare} offers a visual representation of the concept.
@@
 That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets did not). 
-We measured answer correctness and total time each participant needed to answer reach question of their survey.
+We measured answer correctness and the total time each participant needed to answer reach question of the survey.
 
@@
 We implemented our survey as a web application. As part of this 
-effort, we carried out an informal pilot whose main objectives were: to spot bugs in the application and in the data collection mechanism; to gain feedback from respondents about the user experience of the application; and to formulate an estimate about how long answering the survey would, on average, take.
+effort, we carried out an informal pilot whose main objectives were: to spot bugs in the application and in the data collection mechanism; to gain feedback from respondents about the user experience of the application; and to formulate an estimate about how long answering the survey would take on average.
 Fellow undergraduate students, professional colleagues, and friends took the pilot survey.
@@
 
-\diego{Esses detalhes são possíveis candidatos para remoção, se precisarmos de espaço}
-{\color{blue}We invited developers to answer our survey by sending invitations to communities of JavaScript programmers on the Internet.
+%% \diego{Esses detalhes são possíveis candidatos para remoção, se precisarmos de espaço}
+%% {\color{blue}We invited developers to answer our survey by sending invitations to communities of JavaScript programmers on the Internet.
 Our initial plan was to try to engage contributors for two major JavaScript projects, namely Node.js and NPM.
@@
 Due to time constraints, we decided to look for respondents elsewhere.}
-We then posted the survey on Reddit.\footnote{Reddit is a North American online discussion platform.
+We posted the survey on Reddit.\footnote{Reddit is a North American online discussion platform.
 Its discussion threads, often called subreddits, are sorted by subject.
@@
 We were able to collect significant data on time taken and discrepancies in answer correctness between confusing and non-confusing versions of the snippets.
+Some inconsistencies arose while building the squares, for instannce, when a user quit in the middle of the survey.
+When this happened, the corresponding row in the square was left incomplete.
+During data analysis, we considered all squares which contained incomplete rows to be invalid, and discarded them.
+Since we had a large enough number of samples, the squares we had to discard did not impact our analysis.
 
-\rb{acho que podemos mover esse proximo paragrafo para a secao de resultados, ou de ameacas}
-\adriano{acho que pode ir para resultados. nao considero uma ameaca pois eu e o Caio tratamos manualmente as entradas a serem removidas e garantimos a integridade dos latin squares}
+\subsection{Third Study: Interviews with Practitioners}
 
-Inconsistencies may arise when each square is being built.
-The main source of inconsistency we faced was when a user quit in the middle of the survey.
-When this happened, his row in the square was left incomplete.
-We considered all squares which contained incomplete rows to be invalid, and discarded them.
-Since we had a large enough number of samples, the squares we had to discard did not impact our results.
+To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion.
+We also asked each participant if they knew of any other JavaScript-specific construct that they regarded as confusing.
+In this section we details the protocol we followed to conduct the interviews and analyze the results.
 
-% \subsubsection{Pilot Survey}
+% Nós realizamos entrevistas semi-estruturas com o objetivo de identificar a percepção dos desenvolvedores com experiência em JavaScript sobre algumas questões relacionadas à compreensão de código em JavaScript.
+Assim, nesta Seção nós descrevemos os procedimentos adotados para selecionar os participantes para as entrevistas e detalhamos como as entrevistas foram conduzidas.
+Além disso, detalhamos como os resultados das entrevistas foram analisados.
 
-% To validate the web application we developed to conduct the survey, we ran an informal pilot survey whose main objectives were:
-%     \begin{itemize}
-%         \item To spot bugs in the application and in the data collection mechanism;
-%         \item To gain feedback from respondents about the user experience of the application;
-%         \item To formulate an estimate about how long the survey would, on average, take.
-%     \end{itemize}
+\subsubsection{Participants Selection} We invited the participants of the interviews using a snow balling technique.
+That is, starting from our network of contacts, we invited an initial set of candidates to take part of our survey.
+From this initial list, we asked for indication of additional candidates.
+Our main criterion of selection was that all participants shoukd have to been working professionally with JavaScript.
+We invited a total of 17 developers, and 15 of these agreed to participate.
+We conducted the interviews during a period of two weeks.
 
-% We had fellow undergraduate students, work colleagues and friends take the pilot survey.
-We discovered that the aspect that needed most improvement was the user experience.
-Some users reported layout defects, and many reported that the landing page did not explain the survey well enough.
-We also spotted minor issues with our routines to create and populate the Latin Squares.
-After performing all the necessary changes, we were ready to conduct another experiment.
+\subsubsection{Interview Process} We conducted semi-structured interviews using a web conferencing software.
+All the interviews were recorded with the consent of the participants.
+On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes, and the longest one 43.06 minutes.
+The main goal of the interviews was to have developers assess in real time whether two versions of code (one with a single atom and another one without the atom) with the same behavior differed in readability, and, if so, which version they regarded as the easier one to understand.
+The interviews were conducted by two of the authors of this paper, and a third one listened to all the recordings to cross-validate the collected data.
 
-\rb{Comentei o survey com alunos}
+The interviews had three main parts.
+In the first one, we asked the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years and other programming languages they have worked with.
+Table \ref{pinterview} summarizes this demographic information.
 
-% \subsubsection{A Survey With Undergraduate Students}
-%     Our second attempt was due to courtesy of a professor of our department.
-During the semester of the writing of this text, he was lecturing a Data Structures course, which is usually taken during the second semester of the undergraduate course.
-The professor agreed to take the students to a laboratory during one of his classes, and all the students who attended that day took the survey.
+\begin{table*}[htb!]
+\centering
+\begin{tabular}
+{p{0.4cm}p{0.9cm}p{2.0cm}p{6cm}p{1.9cm} p{4.5cm}}
+\toprule
+ID & Gender & Level & Role & JS Experience & Other Languages \\ \midrule 
+P1 & Male & Graduate & Software Developer & 9 years & Java, PHP, C, Go  
+\\ 
+P2 & Male & High School & Software Developer & 3 years & Python, Go, Dart, Lua, C++, C\#
+\\ 
+P3 & Male & Graduate & Software Developer & 4 years & Java, C
+\\ 
+P4 & Male & Undergraduate & Software Developer & 3 years & Python, C, C++, Java, Go
+\\ 
+P5 & Male & Master Student & Software Developer & 3 years & Python, SQL
+\\ 
+P6 & Male & Graduate & Software Architect  & 15  years  & Java, PHP, C, Python , Ruby, C\# and R
+\\ 
+P7 & Male & Graduate & Software Developer  & 6 years & C, C++, Java, Assembly, Kotlin
+\\ 
+P8 & Male & PhD & Software Developer & 2 years & Java, Python
+\\ 
+P9 & Male & Graduate & Business Analyst and Software Developer & 5 years & PHP
+\\ 
+P10 & Male & Master Student & Business Analyst and Software Developer  & 4 years & Java, C\# and Python
+\\ 
+P11 & Male & Master Student & Software Architect & 4 years & Java, Erlang, C\#, Cobol
+\\ 
+P12 & Female & Master & Software Developer  & 1 year & C
+\\ 
+P13 & Male & Master & Software Developer & 13 years & Java, PHP
+\\ 
+P14 & Female & Master Student & Software Developer  & 3 years & C, Python, Ruby
+\\ 
+P15 & Female &  Graduate & Software Developer & 2 years & Java, PHP
+\\ \bottomrule
+\end{tabular}
+    \caption{Demographic information of the participants}
+    \label{pinterview}
+\end{table*}
 
-%     No issues were reported with the web application, and all answers were appropriately collected.
-This time, all the Latin Squares were being populated correctly, making the data we collected eligible for use in the final analysis.
-However, we opted to dispose of all the data we gathered in this occasion.
-What motivated this decision were the two following aspects:
-%     \begin{enumerate}
-%         \item The students were not focusing hard enough.
-Most students were not very serious about the survey.
-While we watched them answer the questions, we could see many students peeking at each other's screens.
-We could also clearly see students not taking nearly enough time to reason about the code, and answering the questions without giving it proper thought.
-%         \item We needed a more diverse spectrum of respondents.
-In the beginning of our research, we established that we would need at least 25 Latin Squares in order to be able to have significant results.
-By conducting this survey, we were able to populate 11 Latin Squares.
-Compounded by the problem reported above, we were concerned that our spectrum of respondents might end up skewed towards undergraduate students with little or no JavaScript programming experience.
-%     \end{enumerate}
+In the second part of the interview our aim was to allow the subjects to describe
+their JavaScript experience, as well as to allow them to reveal any JavaScript's constructs they regarded as innately confusing.
+This would allow us to identify potential atom candidates that are more specific to the JavaScript language.
+The questions we explored in this section include:
 
-%     These two facts led us to conclude we needed to conduct a survey in which partakers engaged voluntarily, whilst also having an incentive to think hard enough about the problems they faced.
-To address these points, we prepared our final setting.
+\begin{itemize}
+\item In general, do you implement new JavaScript components,
+  or do you maintain components developed by other programmers?;
 
-\subsection{Second Study: Mining Software Repositories}
+\item In your opinion, what are the main pros and cons of the JavaScript language?;
+
+\item Do you think JavaScript is a language that produces code that is hard to understand?
+
+\item Do you regard any particular construct or idiom of the language as
+  especially confusing?
+\end{itemize}
+
+In the third part of the interview, participants were shown the 20 code snippets that were used in the first study.
+The participants were told both sides of each pair had the same output, and were asked to evaluate which version of the code was easier to understand.
+The interviewers restrained themselves from introducing bias into the answers by not explaining that one of the sides of each pair contained the atom under investigation.
+Subjects were just presented the snippets and allowed to take the necessary time to decide on the most readable snippet.
+
+\subsubsection{Interview Analysis}
+
+{\color{red}Nós realizamos a análise das entrevistas em pares.
+Inicialmente um dos participantes de todas as entrevistas transcreveu os resultados.
+Posteriormente, um dos autores do trabalho ouviu os áudios das entrevistas e transcreveu os resultados.}
+
+\textcolor{red}{Adriano teria algo mais para adicionar aqui na  análise das entrevistas?}
+
+\todo[inline]{adriano:estou escrevendo uma tabela em que cada linha e um dos atomos, e colunas representam \% preferencia com atomo, \% preferencia sem atomo, \% indiferente.
+Mas acredito que esta tabela deve estar na section results, certo?
+O resto das informacoes deste paragrafo eu citei acima no Interview Process, entao acho que podemos abolir esta section.
+-- Edna: Concordo contigo Adriano}
+
+\subsection{Third Study: Additional Survey and Interviews}
+
+\rb{acho que aqui podemos incluir o segundo survey e a segunda
+  rodada das entrevistas, explorando os \'{a}tomos adicionais}
+
+\subsection{Fourth Study: Mining Software Repositories}
 
 To understand how often atoms of confusion appear in real settings, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories.
-To this end, we first collected the most popular GitHub repositories that are primarily written is JavaScript.
-We measured popularity using the project's stargazers, and selected only projects with stargazers\_count greater than 100.
+To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript.
+We measured popularity using the project's stargazers.
 This metric, available on  GitHub API, represents the number of stars a project received from users of the platform.
-The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs}.
+The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}.
 We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.
@@
 
-In the third step we mined atoms of confusion from the repositories in our curate dataset.
-We written source code queries using the CodeQL language~\cite{moor:gttse2007}, in order to filter the atoms of confusion.
+In the third step we mined atoms of confusion from the repositories in our curate dataset, using source code queries that we had written using the CodeQL language~\cite{moor:gttse2007}.
 CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as Java and JavaScript).
-We also automate the process of running the queries and exporting the results to a format that simplifies our analysis and the reproduction of this study.
-In the final step we carry out basic descriptive statistics, in order to understand how often the atoms of confusion appear in practice.
+We also automate the process o
+f running the queries and exporting the results to a format that simplifies our
+analysis (and also the reproduction of this study). In the final step we carry
+out basic descriptive statistics, in order to understand how often the
+atoms of confusion appear in practice.
 
@@
  \end{table*}
-
-\subsection{Third Study: Interviews with Practitioners}
-
-To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion.
-We also asked each participant if they knew of any other JavaScript-specific construct that they regarded as confusing.
-In this section we details the protocol we followed to conduct the interviews, and we also describe how their results were analysed.
-
-% Nós realizamos entrevistas semi-estruturas com o objetivo de identificar a percepção dos desenvolvedores com experiência em JavaScript sobre algumas questões relacionadas à compreensão de código em JavaScript.
-Assim, nesta Seção nós descrevemos os procedimentos adotados para selecionar os participantes para as entrevistas e detalhamos como as entrevistas foram conduzidas.
-Além disso, detalhamos como os resultados das entrevistas foram analisados.
-
-\subsubsection{Participants Selection}: The participants of the interviews were invited personally by the authors, and the pool of potential participants was based on our network of contacts.
-Our main criterion of selection was that all participants had to be working professionally with JavaScript.
-We invited a total of 17 developers, and 15 of these agreed to participate, and throughout a two week period, we conducted all the interviews.
-
-% Os participantes das entrevistas foram selecionados a partir da nossa rede de contatos.
-Nós convidamos 17 pessoas e 2 não responderam ao nosso convite para participar das entrevistas.
-Durante 2 semanas, nós entrevistamos 15 desenvolvedores.
-Iniciamente nós conduzimos duas entrevistas como piloto.
-Após o piloto com dois desenvolvedores com mais de 03 anos de experiência, nós adicionamos uma nova pergunta na entrevista: Você conhece alguma construção em JavaScript que leve a um mal entendimento do código?
-Ao finalizar as entrevistas, nós perguntamos ao desenvolvedor se ele conhecia algum outro desenvolvedor com experiência em JavaScript que pudesse nos indicar para participar das entrevistas.
-Nós entrevistamos 02 desenvolvedores a partir da indicação dos entrevistados.
-A Tabela \ref{pinterview} apresenta o perfil dos 15 participantes das entrevistas.
-
-\subsubsection{Interview Process}
-
-We conducted semi-structured interviews using a web conferencing software.
-All the interviews were recorded with the consent of participants.
-On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes, and the longest one 43.06 minutes.
-The main goal of the interviews was to have developers assess in real time whether two versions of code with the same behavior differed in readability, and, if so, which version they regarded as the easier one to understand.
-The interviews were conducted by two of the authors of this text, and a third one listened to all the recordings to cross-validate the collected data.
-
-% Nós conduzimos entrevistas semi-estruturadas utilizando um software de conferência de audio/video.
-Todas as entrevistas foram gravadas com o consenso dos entrevistados.
-As entrevistas tiveram uma duração de aproximadamente 26.29 minutos.
-A entrevista com menor duração foi de 14.59 minutos e de maior duração foi de 43.06 minutos.
-O objetivo ao realizar as entrevistas era identificar quais as construções em JavaScript podem corresponder a confusões de código.
-As entrevistas foram realizadas por dois dos autores deste artigo.
-
-The interviews had three main parts.
-In the first one, we asked the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience and other programming languages they have worked with.
-Table \ref{pinterview} presents the profile of each participant.
-
-\begin{table*}[htb!]
-\centering
-\begin{tabular}
-{|p{0.4cm}|p{0.9cm}|p{2.0cm}|p{7.0cm}|p{1.9cm}|p{3.0cm}|}
-\hline
-ID & Gender & Level & Role & JS Experience & Other Languages \\ \hline  
-P1 & Male & Graduate & Developer at Federal Court of Accounts - TCU Brazil & 9 years & Java, PHP, C, Go  
-\\ \hline
-P2 & Male & High School & Developer at Luma Health & 3.5 years & Python, Go, Dart, Lua, C++, C\#
-\\ \hline
-P3 & Male & Graduate & Developer at Superior Labour Court - TST Brazil & 4 years & Java, C
-\\ \hline
-P4 & Male & Undergraduate & Developer at Brazilian Association of Research and Industrial Innovation - EMBRAPII & 3 years & Python, C, C++, Java, Go
-\\ \hline
-P5 & Male & Master Student & Developer na Qubo Technology & 3 years & Python, SQL
-\\ \hline
-P6 & Male & Graduate & Software Architect at Superior Labour Court - TST Brazil & >15 years & Java, PHP, C, Python , Ruby, C\# and R
-\\ \hline
-P7 & Male & Graduate & Developer na Stefanini & 6 years & C, C++, Java, Assembly, Kotlin
-\\ \hline
-P8 & Male & PhD & Developer Regional Labour Court - TRT Goiás, Brazil & 2 years & Java, Python
-\\ \hline
-P9 & Male & Graduate & IT Analyst at Regional Bank of Brasilia - BRB Brazil & 5 years & PHP
-\\ \hline
-P10 & Male & Master Student & Business Analyst and Developer at National Confederation of Industry - CNI Brazil & 4 years & Java, C\# and Python
-\\ \hline
-P11 & Male & Master Student & Software Architect at the University of Brasilia's Informatics Centre & 4 years & Java, Erlang, C\#, Cobol
-\\ \hline
-P12 & Female & Master & Software Engineers at Novatics & 1 year & C
-\\ \hline
-P13 & Male & Master & Developer at the Military Police of the Federal District - PMDF Brazil & 13 years & Java, PHP
-\\ \hline
-P14 & Female & Master Student & Developer at Neoway & 3 years & C, Python, Ruby
-\\ \hline
-P15 & Female &  Graduate & Developer at SICOOB & 2 years & Java, PHP
-\\ \hline
-\end{tabular}
-    \caption{Demographic information of the participants}
-    \label{pinterview}
-\end{table*}
-
-The second part of the interview was more open, and our aim here was to allow the subjects to describe their JavaScript experience, as well as to allow them to make suggestions about any JavaScript's constructs they regarded as innately confusing, as this would allow us to include their suggestions in future work.
-The questions asked in this section were:
-\begin{itemize}
-    \item In general, do you implement new JavaScript components, or do you maintain components developed by other programmers?;
-    \item In your opinion, what are the main pros and cons of the language?;
-    \item Do you think JavaScript is a language that produces code that is hard to understand?
-    \item Do you regard any particular construct of the language as especially confusing?
-\end{itemize}
-
-\rb{acho melhor mover os code snippets da entrevista para um apendice online. deixei comentado aqui.}
-\adriano{como os snippets sao os mesmos do TCC, a ja temos o apendice, nao poderiamos colocar no apendice B deste texto?
-posso escrever os snippets no repositorio tambem.
-deixa-los no apendice pode tornar o paragrafo abaixo mais compreensivel/verificavel.}
-
-In the third part of the interview, participants were shown the 20 code snippets that were used to build the aforementioned Latin Squares.
-The participants were told both sides of each pair had the same output, and were asked to evaluate which construct was easier to understand.
-The interviewers restrained themselves from introducing bias into the answers by not explaining that one of the sides of each pair contained the atom under investigation.
-Subjects were just presented the snippets and allowed to take the necessary time to decide on the most readable snippet.
-
-% As entrevistas foram divididas em três etapas.
-Na primeira etapa da entrevista nós perguntamos aos desenvolvedores as seguintes informações: Nome; E-mail; Gender; Level; Role; Experience; Quais as outras linguagens de programação ele/ela conhecia.
-Na segunda etapa da entrevista,  perguntamos aos desenvolvedores: a) No geral você escreve novos componentes em JavaScript, ou você faz manutenção em componentes desenvolvidos por outra pessoa (explique o seu dia-a-dia); b) Na sua opinião, quais são os pontos fortes e fracos da linguagem JavaScript.
-c) Você acha que a linguagem JavaScript é uma linguagem que leva a um código difícil de ser compreendido?
-d) Tem alguma construção em JavaScript que você conhece que pode levar a uma dificuldade de entendimento.
-Na terceira etapa da entrevista nós apresentamos alguns exemplos de trechos de código em JavaScript, contendo os 10 tipos de construção apresentados na Table \ref{tab:atoms} com e sem átomos de confusão (Lado A e Lado B), conforme apresentado na Table \ref{snippets}.
-Nós perguntamos aos desenvolvedores: 1) Você acha que o código do lado B prejudica de alguma forma a compreensão do código?
-Os dois códigos apresentados nos dois lados de alguma maneira eram equivalentes.
-
-% \begin{longtable}
-% {| p{.50\linewidth} | p{.50\linewidth} |} 
-% % \begin{center}
-% \caption*{Code Snippets presented on the interview}
-% \label{snippets}
-% % \begin{tabular}{ | p{8.2cm} | p{8.2cm} | } \hline
-% \hline
-% \multicolumn{2}{| c |}{Arithmetic as Logic}  \\ \hline
-% Side A  &  Side B \\ \hline
-% \begin{lstlisting}[language=JavaScript]
-% let array_1 = [1,2,3];
-% let array_2 = [2,3,2];
-% if(array_1.length - array_2.length){
-%   console.log(true);
-% }
-% else{
-%   console.log(false);
-% } \end{lstlisting}
-% & 
-% \begin{lstlisting}[language=JavaScript]
-% let array_1 = [1,2,3];
-% let array_2 = [2,3,2];
-% if(array_1.length != array_2.length){
-%   console.log(true);
-% }
-% else{
-%   console.log(false);
-% } \end{lstlisting} \\ \hline
-% \multicolumn{2}{| c |}{Assignment as value}  \\ \hline
-% Side A  &  Side B \\ \hline
-% \begin{lstlisting}[language=JavaScript]
-% function resetSchedulerState () {
-%   let activatedChildren = {length: 10};
-%   let index = length = activatedChildren.length = 0;
-%   console.log(index);
-% }
-% resetSchedulerState(); \end{lstlisting}
-% & 
-% \begin{lstlisting}[language=JavaScript]
-% function resetSchedulerState () {
-%   let activatedChildren = {length: 10};
-%   activatedChildren.length = 0;
-%   let length = activatedChildren.length;
-%   let index = length;
-%   console.log(index);
-% }
-% resetSchedulerState(); \end{lstlisting} \\ \hline
-% %\pagebreak 
-% \hline
-% \multicolumn{2}{| c |}{Automatic Semicolon Insertion}  \\ \hline
-% Side A  &  Side B \\ \hline
-% \begin{lstlisting}[language=JavaScript]
-% function example(){
-%   return
-%       10
-% }
-% if(example() == 10){
-%   console.log(true);
-% }
-% else{
-%   console.log(false);
-% } \end{lstlisting}
-% & 
-% \begin{lstlisting}[language=JavaScript]
-% function example(){
-%   return 10;
-% }
-% if(example() == 10){
-%   console.log(true);
-% }
-% else{
-%   console.log(false);
-% } \end{lstlisting} \\ \hline
-% \multicolumn{2}{| c |}{Comma Operator}  \\ \hline
-% Side A  &  Side B \\ \hline
-% \begin{lstlisting}[language=JavaScript]
-% let V1 = 5, V2 = 10;
-% V1 = (V2 = 1, 2);
-% console.log(V1+V2);
-% \end{lstlisting}
-% & 
-% \begin{lstlisting}[language=JavaScript]
-% let V1 = 5, V2 = 10;
-% V2 = 1;
-% V1 = 2;
-% console.log(V1+V2);
-% \end{lstlisting} \\ \hline
-% %\pagebreak
-% \hline
-% \multicolumn{2}{| c |}{Ternary Operator}  \\ \hline
-% Side A  &  Side B \\ \hline
-% \begin{lstlisting}[language=JavaScript]
-% let config = {size: 3, isActive: false};
-% const _config = config.isActive === true ? config : {size: 10};
-% console.log(_config.size); \end{lstlisting}
-% & 
-% \begin{lstlisting}[language=JavaScript]
-% let config = {size: 3, isActive: false}
-% let _config;
-% if(config.isActive === true){
-%   _config = config;
-% }
-% else{
-%   _config = {size: 10};
-% }
-% console.log(_config.size); \end{lstlisting} \\ \hline
-
-% \multicolumn{2}{| c |}{Implicit Predicate}  \\ \hline
-% Side A  &  Side B \\ \hline
-% \begin{lstlisting}[language=JavaScript]
-% let V1 = 10, V2 = 3;
-% if (!(V1 % V2)){
-%   console.log(true);
-% }
-% else{
-%   console.log(false);
-% } \end{lstlisting}
-% & 
-% \begin{lstlisting}[language=JavaScript]
-% let V1 = 1, V2 = 2;
-% if ((V2 - V1) == 0){
-%   console.log(true);
-% }
-% else{
-%   console.log(false);
-% } \end{lstlisting} \\ \hline
-% \multicolumn{2}{| c |}{Logic Control Flow}  \\ \hline
-% Side A  &  Side B \\ \hline
-% \begin{lstlisting}[language=JavaScript]
-% let V1 = 3;
-% let V2 = 5;
-% let V3 = 0;
-% while (V1 != V2 && ++V1) {
-%   V3++;
-% }
-% console.log(V1 + V3); \end{lstlisting}
-% & 
-% \begin{lstlisting}[language=JavaScript]
-% let V1 = 1;
-% let V2 = 11;
-% let V3 = 0;
-% while (V1 != V2) {
-%   V1++;
-%   V3++;
-% }
-% console.log(V1+V3); \end{lstlisting} \\ \hline
-% \multicolumn{2}{| c |}{Omitted Curly Braces \& Indentation}  \\ \hline
-% Side A  &  Side B \\ \hline
-% \begin{lstlisting}[language=JavaScript]
-% let V1 = 1, V2 = 2
-% if (V1 > V2)
-% V2 = 1
-% V1 = 2
-% console.log(V2+V1) \end{lstlisting}
-% & 
-% \begin{lstlisting}[language=JavaScript]
-% let V1 = 1, V2 = 2;
-% if (V1 > V2) {
-%   V2 = 1;
-% }
-% V1 = 2;
-% console.log(V2+V1); \end{lstlisting} \\ \hline
-% \multicolumn{2}{| c |}{Post Increment}  \\ \hline
-% Side A  &  Side B \\ \hline
-% \begin{lstlisting}[language=JavaScript]
-% let V1 = 0;
-% if (V1++ == 0) {
-%   console.log(true);
-% }
-% else {
-%   console.log(false);
-% } \end{lstlisting}
-% & 
-% \begin{lstlisting}[language=JavaScript]
-% let V1 = 0;
-% if (V1 == 0) {
-%   console.log(true);
-% }
-% else {
-%   console.log(false);
-% }
-% V1 = V1 + 1; \end{lstlisting} \\ \hline
-% \multicolumn{2}{| c |}{Pre Increment}  \\ \hline
-% Side A  &  Side B \\ \hline
-% \begin{lstlisting}[language=JavaScript]
-% var index = 1;
-% while (++index < 10) {
-%   console.log(index);
-%   break;
-% } \end{lstlisting}
-% & 
-% \begin{lstlisting}[language=JavaScript]
-% var index = 0;
-% index = index + 1;
-% while (index < 10) {
-%   console.log(index);
-%   index = index + 1;
-%   break;
-% } \end{lstlisting} \\ \hline
-% % \end{tabular}
-% % \end{center}
-% % \end{table}
-% \end{longtable}
-
-\subsubsection{Interview Analysis}
-
-Nós realizamos a análise das entrevistas em pares.
-Inicialmente um dos participantes de todas as entrevistas transcreveu os resultados.
-Posteriormente, um dos autores do trabalho ouviu os áudios das entrevistas e transcreveu os resultados.
-
-\textcolor{red}{Adriano teria algo mais para adicionar aqui na  análise das entrevistas?}
-
-\todo[inline]{adriano:estou escrevendo uma tabela em que cada linha e um dos atomos, e colunas representam \% preferencia com atomo, \% preferencia sem atomo, \% indiferente.
-Mas acredito que esta tabela deve estar na section results, certo?
-O resto das informacoes deste paragrafo eu citei acima no Interview Process, entao acho que podemos abolir esta section.
--- Edna: Concordo contigo Adriano}

### 2021-01-20 3_Method.tex
revisao da secao de study settings e resultados

@@
 That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets did not). 
-We measured answer correctness and the total time each participant needed to answer reach question of the survey.
+We measured answer correctness and the total time each participant needed to answer the survey's questions.
 
@@
 Due to time constraints, we decided to look for respondents elsewhere.}
+
 We posted the survey on Reddit.\footnote{Reddit is a North American online discussion platform.
-Its discussion threads, often called subreddits, are sorted by subject.
-For our research, we posted information and links to the survey in two JavaScript subreddits.}.
-We explained our research purposes, and asked developers of any level of expertise to take the survey.
-To incentivize serious engagement, we proposed to raffle \$50 gift cards on Amazon products at the end of the survey.
-Within twelve hours we collected more than 150 answers, populating more than 70 replicas of the Latin Squares.
-We were able to collect significant data on time taken and discrepancies in answer correctness between confusing and non-confusing versions of the snippets.
-Some inconsistencies arose while building the squares, for instannce, when a user quit in the middle of the survey.
+Its discussion threads, often called subreddits, are sorted by subject. For our research, we po\
+sted information and links to the survey in two JavaScript subreddits.} We explained our research purposes and asked developers of any level of expertise to take the survey.
+Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares.
+We collected significant data on time taken and discrepancies in answer correctness between confusing and non-confusing versions of the snippets.
+Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the survey.
 When this happened, the corresponding row in the square was left incomplete.
-During data analysis, we considered all squares which contained incomplete rows to be invalid, and discarded them.
-Since we had a large enough number of samples, the squares we had to discard did not impact our analysis.
+We considered all squares that contained incomplete rows to be invalid during data analysis and discarded them.
+Since we had\
+ a large enough number of samples, the squares we had to discard did not impact our analysis.
 
-\subsection{Third Study: Interviews with Practitioners}
+\subsection{Second Study: Interviews with Practitioners}
 
@@
 From this initial list, we asked for indication of additional candidates.
-Our main criterion of selection was that all participants shoukd have to been working professionally with JavaScript.
+Our main criterion of selection was that all participants shoukd havebeen working professionally with JavaScript.
 We invited a total of 17 developers, and 15 of these agreed to participate.
@@
 The main goal of the interviews was to have developers assess in real time whether two versions of code (one with a single atom and another one without the atom) with the same behavior differed in readability, and, if so, which version they regarded as the easier one to understand.
-The interviews were conducted by two of the authors of this paper, and a third one listened to all the recordings to cross-validate the collected data.
+The interviews were conducted by two of the researchers, and a third one listened to all the recordings to cross-validate the collected data.
 
@@
 This would allow us to identify potential atom candidates that are more specific to the JavaScript language.
-The questions we explored in this section include:
+Examples of questions we explored in this section include:\emph{Do you think JavaScript is a language that produces code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as
+  especially confusing?}
 
-\begin{itemize}
-\item In general, do you implement new JavaScript components,
-  or do you maintain components developed by other programmers?;
-
-\item In your opinion, what are the main pros and cons of the JavaScript language?;
-
-\item Do you think JavaScript is a language that produces code that is hard to understand?
-
-\item Do you regard any particular construct or idiom of the language as
-  especially confusing?
-\end{itemize}
-
-In the third part of the interview, participants were shown the 20 code snippets that were used in the first study.
+In the third part of the interview, participants were shown the code snippets that were used in the first study.
 The participants were told both sides of each pair had the same output, and were asked to evaluate which version of the code was easier to understand.
@@
  \end{tabular}
- \caption{Some descriptive statistics about the projects used in the study}
+ \caption{Some descriptive statistics about the projects used in the MSR study}
  \label{tab:projects-statistics} 

### 2021-01-20 4_Result.tex
revisao da secao de study settings e resultados

@@
-\section{Results and Discussion}
-\label{sec:results}
+%% \section{Results and Discussion}
+%% \label{sec:results}
 
-In this section we present the main findings of our research.
+%% In this section we present the main findings of our research.
 We first present the results of each study (Section~\ref{sec:survey-resuts}, Sections~\ref{sec:msr-results}, and Section~\ref{sec:interview-results}).
@@
 
-\subsection{Results of the first study: a survey}
+\section{Results of the Survey}
 \label{sec:survey-resuts}
 
-Our first study aims to investigate the impact of atoms of confusion while developers try to understand JavaScript code.
-We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers) and \emph{cognitive effort} (time necessary to provide an answer).
-As discussed in section~\ref{sec:survey-settings}, 140 participants answered the survey in full.
-Figure~\ref{fig:degree} and Figure~\ref{fig:xp} show the distributions of the subjects, according to their education level and years of programming experience, respectively.
+Our first study investigates the impact of atoms of confusion while developers try to understand JavaScript code.
+We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
+and \emph{cognitive effort} (time necessary to provide a correct answer). We receive full answers to our survey
+from 140 participants. Figure~\ref{fig:degree} and Figure~\ref{fig:xp} show the distributions of the subjects,
+according to their education level and years of programming experience, respectively.
 
@@
       \centering
-      \includegraphics[scale=.25]{images/degree.png}
+      \includegraphics[width=\columnwidth]{images/dem-education-1.pdf}
       \caption{Participants' Education Level}\label{fig:degree}
@@
       \centering
-      \includegraphics[scale=.25]{images/experience.png}
+      \includegraphics[width=\columnwidth]{images/dem-experience-1.pdf}
       \caption{Participants' Experience} \label{fig:xp}
@@
 
-It is possible to realize in Figure \ref{fig:xp} that more than a half of the participants hold either a BS degree or had taken some university course, meaning the participants have had some level of formal education on programming.
+Figure~\ref{fig:xp} shows that more than a half of the participants hold either a BS degree or had taken some university course, meaning the participants have had some level of formal education on programming.
 No subject reported to have never attended university.
-Out audience comprises experience developers, and 93\% of the respondents have more than one year of experience with programming (see Figure~\ref{fig:xp}).
+Our audience comprises skilful developers, and 93\% of the respondents have more than one year of experience with programming (see Figure~\ref{fig:xp}).
 Actually, 56\% of the participants have either between four and ten years of experience (37\%) or more than ten years of experience (19\%).
 Other 37\% of the respondents have between one and four years of experience.
-In this way, we can characterize the effect of atoms of confusion in JavaScript code taking into account the perceptions of both novice and experience developers.
+In this way, we characterize the effect of atoms of confusion in JavaScript code taking into account the perceptions of both novice and experience developers.
 
-Each participant in this study evaluated {\color{red}10} code snippets, from which 5 were in their confusing versions, whilst the other 5 contained simplified versions of the code snippets (i.e., without the confusing constructs and idioms).
+\subsection{Misunderstanding Analysis}
+
+\subsubsection*{Exploratory Data Analysis}
+
+Each participant in this study evaluated {\color{red}10} code snippets, from which 5 were in their confusing versions, whilst the other 5 contained cleaned versions of the code snippets (i.e., without the confusing constructs and idioms).
 The participants should provide the expected outcomes of the code snippets.
 We collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{cognitive workload} (the time taken to answer the question).
-Our final dataset consists of 70 complete Latin Squares, which means that our assessment consider a total of 140 answers to each pair of code snippets.
+Our final dataset consists of 70 complete Latin Squares.
 
@@
 Curiously, frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also introduce high degrees of confusion.
-As the boxplot in Figure \ref{fig:boxplotcorrectness} shows, there is a considerable decrease in the average number of incorrect answers when the atoms were not present.
+
+Moreover, as the boxplot in Figure \ref{fig:boxplotcorrectness} shows, there is a considerable decrease in the average number of incorrect answers when observing the cleaned version of the code snippets.
 Also, the sample of answers where there was no atoms had almost no dispersion, which is a sign that the non-confusing code is easier to evaluate correctly.
@@
 
-\begin{table}[htbp]
-\caption{Average number of correct answers.}
+\begin{table*}[htbp]
+\caption{Total number of correct answers.}
 \label{tab:difference-correctness}
-\begin{tabular}{lccc} \toprule
- & \multicolumn{2}{c}{Average Number of Correct Answers} & \\ [0.1cm]
- Atom & Code with Atoms & Code without Atoms & $\Delta$(\%) \\ \midrule 
- Comma Operator          & 40 & 93  & +132 \\
- Post Increment          & 69 & 91  & + 31  \\
- Omitted Curly Braces    & 67 & 83  & +23 \\
- Assignment as Value     & 80 & 97  & +21 \\
- Implicit Predicate      & 83 & 97  & +16 \\
- Logic as Control Flow   & 59 & 68  & +15 \\
- Ternary Operator        & 86 & 94  & +9  \\
- Pre-Increment           & 71 & 76  & +7  \\
- Arithmetic as Logic     & 91 & 90  & +1  \\ \bottomrule
+\centering{
+\begin{tabular}{lccc|c} \toprule
+ % & \multicolumn{2}{c}{Total Number of Correct Answers} & \\ [0.1cm]
+  Atom & Confuse Code & Clean Code & $\Delta$(\%) & Effect Size (Odds Ratio) \\ \midrule
+  Comma Operator &  28 &  65 & +132 & 19.02 \\ 
+  {\color{red}Automatic Semicolon Insertion} &  32 &  68 & +112 & 39.32 \\ 
+  Post Increment &  48 &  64 & +33 & 4.83 \\ 
+  Ommitted Curly Braces &  47 &  58 &  +23 & 2.35 \\
+  Assignment as Value &  56 &  68 &  +21 & 8.38 \\ 
+  Implict Predicate &  58 &  68 &  +17 & 6.94 \\ 
+  Logic as Control Flow &  41 &  48 &  +17 & 1.53 \\ 
+  Ternary Operator &  60 &  66 &  +10 & 2.73 \\ 
+  Pre Increment &  50 &  53 &  +6 & 1.24 \\ 
+  Arithmetic as Logic &  64 &  63 &  -2 & 0.84 \\ \bottomrule 
 \end{tabular}
-\end{table}
+}
+\end{table*}
+
+ %% Comma Operator          & 40 & 93  & +132 \\
+ %% Post Increment          & 69 & 91  & + 31  \\
+ %% Omitted Curly Braces    & 67 & 83  & +23 \\
+ %% Assignment as Value     & 80 & 97  & +21 \\
+ %% Implicit Predicate      & 83 & 97  & +16 \\
+ %% Logic as Control Flow   & 59 & 68  & +15 \\
+ %% Ternary Operator        & 86 & 94  & +9  \\
+ %% Pre-Increment           & 71 & 76  & +7  \\
+ %% Arithmetic as Logic     & 91 & 90  & +1  \\ \bottomrule
 
@@
  \centering
- \includegraphics[scale=.20]{images/wrong.png}
+ \includegraphics[width=\columnwidth]{images/wrong-answers-plot-1.pdf}
  \caption{Number of wrong answers of each subject}
@@
 \rb{verificar a consistencia nas questoes de pesquisa mais especificas}
-Regarding the first question we 
-address in the survey (\emph{Do code snippets that contain atoms of confusion produce a higher error rate than snippets where the atom is removed?}), we found evidence that the atoms of confusion lead programmers to misunderstand JavaScript code.
+
+\subsubsection*{Hypothesis Testing and Effect Size}
+
+We first use the \emph{Pearson's Chi-squared test}
+to investigate if there is a statistically significant difference
+in the frequency of correct and incorrect answers---due to the versions
+of the code snippets (confuse and clean code).
+The result shows that the confusing version of the code snippets impacts code understanding (p-value $< 0.05$).
+We also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant
+\emph{education} or participant \emph{experience} impact on correctness.
+Although the results suggest that experience impact correctness, the education level does not.
+We then replicate the \emph{Pearson's Chi-squared test} for each experience group (see Figure~\ref{fig:xp}) and found that the statistic difference is less significant for those developers with more than ten years of experience.
+Finally, we measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR)
+method (see Table~\ref{tab:difference-correctness}).
+We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre Increment.
+Nonetheless, for the remaining candidates, the effect size is considerable.
+For instance, the Comma Operators leads to an \emph{Odds Ratio} of
+\num{19.02}, meaning that the participants are \num{19.02} times
+more likely to correctly interpret the clean version of a code in comparison with the corresponding, confuse version of the code snippet.
+
+%% Regarding the first question we 
+%% address in the survey (\emph{Do code snippets that contain atoms of confusion produce a higher error rate than snippets where the atom is removed?}), we found evidence that the atoms of confusion lead programmers to misunderstand JavaScript code.
 We also realized that just one atom whose correction has a non-significant improvement in the percentage of correct answers---we found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of ten atoms we consider in the survey).
 
-\rb{acho que valeria a pena colocar o teste de hipotese aqui}
-\adriano{como fazemos/descrevemos o teste?}
+\subsection{Cognitive Workload Analysis}
 
-Table \ref{tab:difference-time-taken} displays the results of the measurements of the average time taken to answer with and without atoms of confusion.
-Once again, the \emph{Comma Operator} was the construct whose removal had the greatest impact.
-The more expressive results in this measurement are the ones regarding \emph{Logic as Control Flow}, \emph{Implicit Predicate} and \emph{Omitted Curly Braces}, {\color{blue}as they are frequent in JavaScript libraries}.
+\subsubsection*{Exploratory Data Analysis}
+Table \ref{tab:difference-time-taken} shows the average time necessary for the participants give a correct answer about the expected outcomes of a code snippet, considering both confuse and clean versions.
+Eight atom candidates require an increasing cognitive workload.
+For these eight atom candidates, developers take at least 5.90\% less time on average to find a correct answer---when considering the clean version of a code snippet.
+In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
+time on average to find the correct answer for the clean version of the code snippets. 
+Not all atom candidates, though, take less time to predict the answer. In fact, for the Arithmetic as Logic and the Pre-Increment atoms, there was an increase
+of 29.06\% and 38.19\% in the average time taken to give a correct answwer---wherein the clean code versions.
 
-\rb{n\~{a}o sei se devemos antecipar o resultado do estudo de minera\c c\~{a}o 
-nessa se\c c\~{a}o. talvez nao comprometa tanto.}
+\begin{table*}[htbp]
+\caption{Time in seconds to submit a {\color{red}correct} answer}
+\label{tab:difference-time-taken}
+\centering{
+\begin{tabular}{lccc|cc} \toprule
+Atom & Confuse Code & Clean Code & $\Delta$(\%) & Mann-Whitney test & Cliff's Delta \\ \midrule 
+Comma Operator & 87.67 & 20.84 & -76.23 &  \colorbox{lightgray}{$<$ 0.01} & -0.53 \\ 
+Logic as Control Flow & 108.94 & 51.07 & -53.12 &  \colorbox{lightgray}{$<$ 0.01} & -0.39 \\ 
+{\color{red}Automatic Semicolon Insertion} & 46.08 & 22.04 & -52.17 & 0.09 & -0.16 \\ 
+Ommitted Curly Braces & 48.85 & 30.00 & -38.58 & 0.92 & -0.01 \\ 
+Implict Predicate & 36.24 & 24.01 & -33.75 & \colorbox{lightgray}{$<$ 0.01} & -0.29 \\ 
+Post Increment & 28.70 & 25.67 & -10.56 & 0.12 & 0.15 \\ 
+Assignment as Value & 52.47 & 48.95 & -6.71 & 0.82 & 0.02 \\ 
+Ternary Operator & 41.80 & 39.34 & -5.90 & 0.24 & -0.12 \\ 
+Arithmetic as Logic & 28.82 & 37.20 & +29.06 & \colorbox{lightgray}{0.01} & 0.23 \\ 
+Pre Increment & 30.71 & 42.45 & +38.19 & \colorbox{lightgray}{$<$ 0.01} & 0.27 \\ 
+         \bottomrule
+    \end{tabular}}
+\end{table*}
 
-Figure \ref{fig:timetoanswer} shows a box plot of the difference in time taken the predict the output of average time to assess the code snippets.
-Although the dispersion is smaller for non-confusing code blocks, the median time did not vary much.
+\subsubsection*{Hypotheis Testing}
+We use the non-parametric \emph{Mann-Whitney test} to
+investigates the null hypothesis that developers 
+spend the same ammount of time to correctly
+predict the outcome of a code snippet, regardless
+of evaluating the confuse or the clean version
+of the code. Table~\ref{tab:difference-time-taken}
+shows the results, which suggest that we
+should refute the null hypothes for
+five atom candidates (Comma Operator, Logic
+as Control Flow, Implicit Predicate, Arithmetic
+as Logic, and Pre Increment). For the first three,
+the analysis suggest that the participants
+need more time to predict the outcome of
+the code snippet in the confuse code version. Interesting,
+for the atom candidates Arithmetic as Logic and
+Pre Increment, the participants take less time
+to predict the outcome of the code snippets in
+the confuse version.
+We also computed the effect size using the Cliff's Delta method.
+We found a large effect for the Comma Operator atom
+candidate; and a moderate effect for the atom candidates
+Logic as Control Flow, Implicit Predicate, Arithmetic
+as Logic, and Pre Increment.
 
-\begin{table}[htbp]
-\caption{Average number of correct answers.}
-\label{tab:difference-time-taken}
-\begin{tabular}{lccc} \toprule
- & \multicolumn{2}{c}{Average Time in Seconds to Answer} & \\ [0.1cm]
- Atom & Code with Atoms & Code without Atoms & $\Delta$(\%) \\ \midrule 
-Comma Operator        & 60 & 21  & -65  \\
-Logic as Control Flow & 85 & 49  & -42  \\
-Implicit Predicate    & 33 & 24  & -27  \\
-Omitted Curly Braces  & 43 & 31  & -27  \\
-Assignment as Value   & 53 & 49  & -7   \\
-Ternary Operator      & 42 & 42  & 0    \\
-Post Increment        & 27 & 27  & 0    \\
-Arithmetic as Logic   & 29 & 36  & +24  \\
-Pre-Increment         & 34 & 49  & +44  \\
-         \bottomrule
-    \end{tabular}
-\end{table}
+\rb{Acho que dever\'{i}mos justificar o uso do Cliff's Delta aqui e o Odds Ratio na
+  se\c c\~{a}o anterior.}
 
-\begin{figure}[htb!]
-      \noindent
-      \centering
-      \includegraphics[scale=.20]{images/timeeach.png}
-      \caption{Box plot of time to answer each snippet}
-      \label{fig:timetoanswer}
-  \end{figure}
+\subsection{Discussion}
 
-Recalling our second research question of this survey (\emph{Do code snippets with atoms of confusion require programmers to take longer to predict their output?}), the variance in the time to correctly predict the output of a snippet is significantly smaller.
-Moreover, since the number of answers is the same for each inspected atom, data from table \ref{results_time} shows that the time taken to answer questions without atoms of confusion is 20\% smaller.
-Not all atoms, though, took less time to predict the answer.
-In fact, for the \emph{Arithmetic as Logic} and the \emph{Pre-Increment} atoms, there was an increase of 24\% and 44\% in the time taken to answer questions wherein the atoms had been removed.
-For other two atoms, there is no difference whatsoever in the time taken.
-We only find in half of our atoms an increase in the time taken to answer.
-Therefore, even though we are not able to confirm hypotheses H.2 defined in Section 1, further inquiry would be necessary to discover the reason for some atoms taking even longer.
+\rb{Acho que podemos incluir uma discuss\~{a}o ao
+  t\'{e}ermino de cada uma das se\c c\~{o}es com
+  resultados. Ou separamos em uma \'{u}nica
+se\c c\~{a}o de discuss\~{a}o?}
 
-\rb{talvez tenhamos que descrever e testar as hipoteses. a descricao pode ser feita na secao de study settings, e o resultado dos testes pode aparecer aqui.}
+%%  Comma Operator        & 60 & 21  & -65  \\
+%% Logic as Control Flow & 85 & 49  & -42  \\
+%% Implicit Predicate    & 33 & 24  & -27  \\
+%% Omitted Curly Braces  & 43 & 31  & -27  \\
+%% Assignment as Value   & 53 & 49  & -7   \\
+%% Ternary Operator      & 42 & 42  & 0    \\
+%% Post Increment        & 27 & 27  & 0    \\
+%% Arithmetic as Logic   & 29 & 36  & +24  \\
+%% Pre-Increment         & 34 & 49  & +44  \\
 
+\
 % \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}
 
-\subsection{Results of the second study: Interview}
+\section{Results of the Interview Study}
 \label{sec:interview-results}
 
-In this section we present the results of two rounds of two rounds of interviews with practitioners who agreed to participate in the interviews.In the first interview, a total of 15 practitioners were surveyed.
-In this phase, we collected information regarding programming experience, familiarity with JavaScript, and we also collected their opinions regarding nine candidates to atoms of confusion.
-This was done by presenting them pairs of code snippets whose behavior was identical, and ask them to classify each snippets readability.
+\rb{acho que essa separa\c c\~{a}o em dois rounds est\'{a}
+  gerando uma complexidade desnecess\'{a}ria. vou tentar
+  remover isso}.
 
-In the second round of interviews, ten of the fifteen developers who participated in the first round were shown another 8 candidates to atom of confusion whose frequency we observed during the repository mining phase.
+In this section we present the results of the 
+interviews with the practitioners who agreed to participate
+in the study. A total of 15 practitioners
+took part of the interviews.
+We collected information regarding programming
+experience, familiarity with JavaScript, and their opinion
+about the {\color{red}nine} atom candidates---by
+(a) presenting them pairs of code snippets whose behavior were identical
+and then (b) asking them to compare the code snippets readability.
+
+%% In the second round of interviews, ten of the fifteen developers
+%% who participated in the first round were shown another 8 atom candidates to atom of confusion whose frequency we observed during the repository mining phase.
 The same procedure of showing code that behaves exactly the same and surveying for how easy to understand each snippet was applied in this phase.
 
-Supporting the results of the survey,Tables \ref{tab:interview-results1} and \ref{tab:interview-results2} show that for twelve out of the seventeen scenarios surveyed, the respondents prefer the version of the code without the atom of confusion, and in no case the \emph{neutral} ratio was higher than the option for the version without atoms of confusion.An example entry is contained in Appendix~\ref{}.
-For a full listing of the code snippets, visit the paper repository.
+Supporting the results of the survey, Tables~\ref{tab:interview-results1}
+and~\ref{tab:interview-results2} show that for twelve out of the
+seventeen scenarios surveyed, the respondents prefer the version of the
+code without the atom of confusion, and in no case the \emph{neutral}
+ratio was higher than the option for the version without atoms of confusion.
+%% An example entry is contained in Appendix~\ref{}. 
+%% For a full listing of the code snippets, visit the paper repository.
 
-\subsubsection*{Participants views of the atoms} Interesting, only for the \emph{ternary operator} scenario the participants prefer the version of the code with the atom of confusion, instead of the more cleaner version without the atom of confusion.
-Even in this case, some participants who opted for the \emph{left-hand-side} (with the atom of confusion) version still believed that the \emph{right-hand-side} (without the atom of confusion) version was more readable (see Figure~\ref{code:ternary}).
-The following quotes were extracted from three interviews:
+\subsubsection*{Participants views of the atoms} Interesting,
+only for the \emph{ternary operator} scenario the participants
+prefer the version of the code with the atom of confusion,
+instead of the more cleaner version without the atom of confusion.
+Even in this case, some participants who opted for
+the \emph{left-hand-side} (with the atom of confusion) version
+still believed that the \emph{right-hand-side} (without the atom of
+confusion) version was more readable (see Figure~\ref{code:ternary}). 
+The following quotes were extracted from the transcripts with
+three inteviwes:
 
@@
 \begin{mq}
-\emph{``I prefer to write [code using the \lhs version], but I think [the \rhs version]  is easier to read, especially for newer programmers".}
+\emph{``I prefer to write [code using the \lhs version], but I think [the \rhs version]  is easier to read, especially for newer programmers''}
 \end{mq}
@@
 \begin{mq}
-\emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the [the code using the \lhs version] is easier to understand"}.
+\emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the [the code using the \lhs version] is easier to understand''}.
 \end{mq}
@@
 
-This particular atom of confusion also opens up the possibility for a derivative construct that JavaScript allows which can be rather confusing, and that is the nested ternaries construct, in which the right-hand side of the a ternary operator can be another ternary construct.
-While nested ternaries remove the number of lines that would be necessary to construct using nested if-then-else statements, they can become quite taxing, cognitively speaking, to understand.
+The atom candidate Ternary Operator also opens up
+the possibility for a derivative construct that JavaScript
+allows which can be rather confusing, and that is the nested
+ternaries construct, in which the right-hand side of the a
+ternary operator can be another ternary construct.
+While nested ternaries remove the number of lines that
+would be necessary to construct using nested if-then-else statements,
+they can become quite taxing, cognitively speaking, to understand.
 Therefore, nested ternaries are a choice of atoms to be analysed in future work.
 
-\rb{Poder\'{i}amos quantificar a existencia de emph{nesting} com code queries?}
+The Pre-Increment atom candidate also caused such
+divide in one of the interviewees, as they
+regarded the \emph{\rhs version} (without atom of confusion)
+as simpler to understand, but would still opt to write
+code that contained the atom. In contrast to such opinion,
+one of the participants found the version that contained the atom
+of confusion more elegant, but recognized it was less readable,
+and was willing to sacrifice elegance for readability.
 
-The \emph{Pre-Increment} atom of confusion also caused such divide in one of the interviewees, as they regarded the \emph{\rhs version} (without atom of confusion) as simpler to understand, but would still opt to write code that contained the atom.
-In contrast to such opinion, one of the participants found the version that contained the atom of confusion more elegant, but recognized it was less readable, and was willing to sacrifice elegance for readability.
+\rb{pela discuss\~{a}o anterior, parece-me que alguns
+  desenvolvedores reconhecem que esses candidatos
+  a \'{a}tomos introduzem certa dificuldade de compreens\~{a}o;
+  mas mesmo assim optam por usar a constru\c c\~{a}o correspondente.
+  Acho que cabe um box aqui com essa discuss\~{a}o sobre isso.}
 
-When analyzing the \emph{Logic as Control Flow} atom of confusion, one of the interviewees gave an example of his own experience on why one should avoid writing code with such atom:
+When analyzing the Logic as Control Flow atom candidate, one of
+the interviewees gave an example of his own experience
+on why one should avoid writing code using this construct:
 
 \begin{mq}
-\emph{``This one is interesting, because I have written code that looks like the [\lhs version (with the atom of confusion)], and my colleagues complained that it was difficult to understand.
+  \emph{``This one is interesting, because I have written
+  code that looks like the [\lhs version (with the atom of confusion)],
+  and my colleagues complained that it was difficult to understand.
   So nowadays I prefer to write code using the [\rhs version]."}
@@
 
-Two of the atoms had unanimous preference for the versions that did not contain the atoms: \emph{Comma Operator} and \emph{Omitted Curly Braces and Indentation}.
-Regarding the first one, we could often notice during the interviews that the \emph{\lhs} (with the atom of confusion) caused significant confusion among the participants.
-Two remarks about the \emph{Comma Operator} atom of confusion are listed below:
+Two of the atom candidates had unanimous preference for the
+versions that did not contain the atoms: Comma Operator and
+Omitted Curly Braces. Regarding the first one, we could
+often notice during the interviews that
+the \emph{\lhs} (with the atom of confusion) caused significant
+confusion among the participants. Two remarks about
+the \emph{Comma Operator} atom of confusion are listed below:
 
 \begin{mq}
-\emph{``I just learned that [the \lhs version of this code] is possible. I did not even know it worked"}
+\emph{``I just learned that [the \lhs version of this code] is possible. I did not even know it worked''}
 \end{mq}
@@
 \begin{mq}
-\emph{``The code in [the \lhs version] is unlikely to be understood unless the programmer knows \clang or \cpplang"}
+\emph{``The code in [the \lhs version] is unlikely to be understood unless the programmer knows \clang or \cpplang''}
 \end{mq}
 
-As for the \emph{Omitted Curly Braces and Indentation} atom of confusion, which, as we shall see in Section~\ref{sec:msr-results}, is frequent in JS code bases, one of the interviewees mentioned that, although they understand why one would opt to not use braces for simple if-then-else statements, he still advised against it, on grounds that:
+As for the atom candidate Omitted Curly Braces, one of the interviewees mentioned that, although they understand why one would opt not to use braces for simple if-then-else statements, he still advised against it, on grounds that:
 
@@
 \emph{``[I prefer the \rhs version of the code \ldots] If I want to see well-written, easily understandable code, then I also have to do my job.
-Therefore I believe that, since I do not know who is on the other end maintaining this code, and it could be any person with any level of expertise, then I try to write readable, easy-to-understand code"}
+Therefore I believe that, since I do not know who is on the other end maintaining this code, and it could be any person with any level of expertise, then I try to write readable, easy-to-understand code''}
 \end{mq}
 
-This view is central to our research on why such small, yet tricky, constructs should be avoided whenever possible, since maintenance is one of the longer and most costly parts of the software development life cycle, and one cannot make any assumptions about the other programmers' experience.
+This is an interesting perspective, because such small, yet tricky, constructs should be avoided, since maintenance is one of the longer and most costly parts of the software development life cycle, and one cannot make any assumptions about the other programmers' experience.
 
-\rb{que conclus\~{a}o extra\'{i}mos com essa parte do estudo?}
+% \subsection{Common Remarks From the Open-ended Questions}
 
-\subsection{Common Remarks From the Open-ended Questions} The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confuse code.
+The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confuse code.
 
-One of the participants mentioned that the use of JavaScript's prototype-based inheritance can make it difficult to understand code that, particularly when involving  deep prototype chains.
+One of the participants mentioned that the use of JavaScript's prototype-based inheritance can make it difficult to understand code, particularly when involving  deep prototype chains.
 This is a core feature of the language, and most high-level tools and frameworks abstract it away.
@@
 
+\rb{seria importante incluir uma cita\c c\~{a}o para a \'{u}ltima afirma\c c\~{a}o acima.}
+
 When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate.
-One of the respondents regarded nested callbacks difficult even to write:
+One of the respondents stated:
 
 \begin{mq}
-\emph{``Nested callbacks are very confusing. Even writing them can be confusing, let alone understanding them."}
+\emph{``Nested callbacks are very confusing. Even writing them can be confusing, let alone understanding them.''}
 \end{mq}
@@
 
-A total of 40\% (n = 6) practitioners regarded the JavaScript language as quick and/or easy to learn, and 26.7\%(n = 4) mentioned that the abundance of frameworks, coupled with the abstractions they provide, make it so that one can develop quickly in the language.
-However, as mentioned by one of the interviewees, ``\emph{Developers who have started programming in JavaScript might miss out on important lower-level aspects, such as garbage collection, memory management, and processor management.}''.
+Six practitioners regarded the JavaScript language as easy to learn, while four
+participants mentioned that the abundance of frameworks,
+coupled with the abstractions they provide, make it so that one
+can develop quickly in the language. However, as mentioned by one
+of the interviewees, ``\emph{Developers who have started
+programming in JavaScript might miss out on important lower-level
+aspects, such as garbage collection, memory management,
+and process management.}''.
 
-Moreover, such abstractions can make it so that one starts to program using the language without actually understanding some of its basic aspects.
-For instance, 26.7\%(n = 4) said that the language is not typed, 13.3\%(n = 2) believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
+\rb{esse pr\'{o}ximo par\'{a}grafo eh interessante,
+  se refor\c carmos que os desenvolvedores tem
+  experi\^{e}ncia com a linguagem. caso contr\'{a}rio,
+  parece que nao fizemos uma boa selecao de participantes.
+  algo para discutirmos.}
+
+Moreover, such abstractions might lead a developer to start to program
+using the language without actually understanding some of its basic aspects.
+For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
 For the multi-purpose language that is one of the most, if not the most used one by practitioners
-~\cite{}, a lack of understanding of the languages core features may be a problem worth investigating in future studies.
+~\cite{}, a lack of understanding of the languages core
+features may be a problem worth investigating in future studies.
 
@@
 
-\begin{table}[!htb]
-    \centering
-    \caption{Interviews Round 2 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}).
+%% \begin{table}[!htb]
+%%     \centering
+%%     \caption{Interviews Round 2 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}).
 Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
-    \label{tab:interview-results2}
-    \begin{tabular}{lrrr}\toprule
-      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
-      \cmidrule(lr){2-4}
-         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
-                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
-                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
-         Property access         & 10  & 60  & 30  \\
-         Object spread           & 50 & 50     & 0    \\
-         Array spread            & 60  & 40    & 0     \\
-         Arrow function          & 40 & 40     & 20  \\
-         Array destructuring     & 40 & 40  & 20  \\
-         Object destructuring    & 40 & 40     & 20    \\
-         Type conversion         & 30  & 60    & 10     \\
-         Change Literal encoding & 40 & 50  & 10  \\ \midrule
-         \textsc{overall}        & 42.8 & 47.6  & 9.6  \\
-         \bottomrule
-    \end{tabular}
-\end{table}
+%%     \label{tab:interview-results2}
+%%     \begin{tabular}{lrrr}\toprule
+%%       & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
+%%       \cmidrule(lr){2-4}
+%%          \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
+%%                                       &  \multicolumn{1}{c}{\textsc{without aoc}}
+%%                                                & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
+%%          Property access         & 10  & 60  & 30  \\
+%%          Object spread           & 50 & 50     & 0    \\
+%%          Array spread            & 60  & 40    & 0     \\
+%%          Arrow function          & 40 & 40     & 20  \\
+%%          Array destructuring     & 40 & 40  & 20  \\
+%%          Object destructuring    & 40 & 40     & 20    \\
+%%          Type conversion         & 30  & 60    & 10     \\
+%%          Change Literal encoding & 40 & 50  & 10  \\ \midrule
+%%          \textsc{overall}        & 42.8 & 47.6  & 9.6  \\
+%%          \bottomrule
+%%     \end{tabular}
+%% \end{table}
 
-\subsection{Results of the third study: Mining open source repositories}
+\section{Results of the MSR Effort}
 \label{sec:msr-results}

### 2021-01-20 dataset/analysis.Rmd
adicionando os datasets

---
title: "atoms-of-confusion"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(".")
require(sqldf)
```

## Load the Dataset

```{r}
ds <- read.csv("survey-results.csv")
colnames(ds)
nrow(ds)
```

## Clean up the Dataset

```{r}
ds <- ds[ds$ref=="reddit" & ds$total == "YES",]
ds.count <- tapply(ds$time,ds$rid,length)
ds.full <- names(ds.count)[ds.count==20]
finalds <- ds[is.element(el = ds$rid, set = ds.full),]
dim(finalds)
```

## Number of Latin squares

```{r}
sqldf("select count(distinct rid) from finalds")
```

## Effect of atoms of confusion in correctness

```{r}
prop.table(table(finalds$correct,finalds$atom),margin = 2)
chisq.test(finalds$correct,finalds$atom)
```

### 2021-01-20 dataset/survey-results-final.Rmd
adicionando os datasets

---
title: "survey-analysis-final"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='images/', dev=c('png', 'pdf'))

library(questionr)
library(fmsb)
library(sqldf)
library(xtable)
library(effsize)
setwd(".")

```

## Load the datasets

```{r cars}
ds <- read.csv("survey-results.csv")
nrow(ds)
colnames(ds)

atomId <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 0) 
atomDescription <- c("Arithmetic as Logic", "Assignment as Value", "Automatic Semicolon Insertion",
           "Comma Operator", "Ternary Operator", "Implict Predicate", 
           "Logic as Control Flow", "Ommitted Curly Braces", "Post Increment", 
           "Pre Increment")

atom_ds <- data.frame(atomId, atomDescription)
```

## Filter the datasets

```{r}
subSet <- ds[ds$ref=="reddit" & ds$total == "YES",]
summary(subSet$total)
squareLengths <- tapply(subSet$time,subSet$rid,length)
completeCases <- names(squareLengths)[squareLengths==20]
ds <- ds[is.element(el = ds$rid, set = completeCases),]
dim(ds)
```

## Exploratory Data Analysis

### Demographics

#### Education

```{r dem-education}
educationLevelIds <- c(1, 2, 3, 4, 5)
educationLevelLabels <- c("High school degree or equivalent", "Some university course but not a degree", "Bacharelor degree", "Master degree", "Doctor degree")

education_ds <- data.frame(educationLevelIds, educationLevelLabels)
colnames(education_ds) <- c("id", "description")

tmp <- sqldf("select distinct sid, education from ds")

demEducation <- sqldf("select b.description, count(*) total 
                      from tmp a, education_ds b
                      where a.education = b.id
                      group by b.description 
                      order by 2 desc")

barplot(demEducation$total, col=topo.colors(5), ylim = c(0, 70), cex=0.7, cex.lab = 0.7, cex.axis=0.7, ylab="Number of participants")

legend("topright", legend=demEducation$description, fill=topo.colors(5), cex=0.6)
```

#### Experience

```{r dem-experience}
experienceLevelIds <- c(1, 2, 3, 4)
experienceLevelLabels <- c("Under one year of experience", "One year and under four years of experience", "Four years and under ten years of experience", "More than ten years of experience")

experience_ds <- data.frame(experienceLevelIds, experienceLevelLabels)

colnames(experience_ds) <- c("id", "description")

tmp <- sqldf("select distinct sid, experience from ds")
demExperience <- sqldf("select experience, count(*) total from tmp group by experience order by 1")

demExperience <- sqldf("select a.experience, b.description, a.total
                        from demExperience a, experience_ds b
                        where a.experience = b.id 
                        order by 3 desc")
demExperience

barplot(demExperience$total, col=topo.colors(4), ylim = c(0, 60), cex=0.7, cex.lab = 0.7, cex.axis=0.7, ylab="Number of participants")

legend("topright", legend=demExperience$description, fill=topo.colors(4), cex=0.6)
```

### Total number of correct answers (Table III)

```{r correct-answers}
codeWithAtoms <- sqldf("select qid, count(*) confuseCode, avg(time) timeConfuseCode
                        from ds 
                        where atom == 'YES' and correct = 'CORRECT'
                        group by qid")

codeWithoutAtoms <- sqldf("select qid, count(*) cleanCode, avg(time) timeCleanCode
                        from ds 
                        where atom == 'NO' and correct = 'CORRECT'
                        group by qid")

codeWithAtoms["atomId"] = codeWithAtoms$qid %% 10
codeWithoutAtoms["atomId"] = codeWithoutAtoms$qid %% 10

merged <- sqldf("select c.atomDescription as Atom, 
                        a.confuseCode as 'Confuse Code', b.cleanCode as 'Clean Code',  
                        (b.cleanCode * 100 / a.confuseCode) -100 as 'Delta (%)'
                 from codeWithAtoms a, codeWithoutAtoms b, atom_ds c
                 where a.atomId = b.atomId and a.atomId = c.atomId
                 order by 4 desc")

xtable(merged)
```

```{r wrong-answers-plot}
wrongAnswers = ds[ds$correct == 'WRONG', ]

wrongAnswersByStudentTreatment <- aggregate(rid~sid+atom, 
                                            data = wrongAnswers, 
                                            FUN=length)

boxplot(wrongAnswersByStudentTreatment$rid ~ wrongAnswersByStudentTreatment$atom
        , ylab = "", xlab = "", main = "", 
        names = c("Clean Code", "Confuse Code"))
```

### Average time for correct answers (Table IV)

```{r time}
merged <- sqldf("select c.atomDescription as Atom, 
                        a.timeConfuseCode as 'Confuse Code', b.timeCleanCode as 'Clean Code',
                        (b.timeCleanCode * 100 / a.timeConfuseCode) -100 as 'Delta (%)'
                 from codeWithAtoms a, codeWithoutAtoms b, atom_ds c
                 where a.atomId = b.atomId and a.atomId = c.atomId
                 order by 4")  
xtable(merged)

correctAnswers = ds[ds$correct == 'CORRECT', ]

correctAnswersByStudentTreatment <- aggregate(time~sid+atom, 
                                            data = correctAnswers, 
                                            FUN=mean)

boxplot(log(correctAnswersByStudentTreatment$time) ~ correctAnswersByStudentTreatment$atom
        , ylab = "", xlab = "", main = "Number of wrong answers of each participant", 
        names = c("Clean Code", "Confuse Code"))

```

```{r}
prop.table(table(ds$correct,ds$atom),margin = 2)
chisq.test(ds$correct,ds$atom)
```

## Regression Analysis (correctness~atom + experience + education)

```{r}
experience <- as.factor(ds$experience)
education <- as.factor(ds$education) 
contrr<-matrix(c(rep(1,4),c(1/2,1/2,-1/2,-1/2),c(1,-1,0,0),c(0,0,1,-1)),byrow=TRUE,nrow=4)
contrasts(experience)<-solve(contrr)[,2:4]

ds$atom <- as.factor(ds$atom)
ds$correct <- as.factor(ds$correct)

mod <- glm(ds$correct~ds$atom+experience+education,family = "binomial")
summary(mod)
```

## Impact considering Developer Experience

```{r experience}
for(i in 1:4){
  print(paste("Experience ", i ))
  print(prop.table(table(ds$correct[ds$experience==i],ds$atom[ds$experience==i]),margin=2))
  print(chisq.test(table(ds$correct[ds$experience==i],ds$atom[ds$experience==i])))
}
```

## Impact considering Developer Education

```{r education}
for(i in 1:5){
  print(paste("Education ", i ))
  print(prop.table(table(ds$correct[ds$education==i],ds$atom[ds$education==i]),margin=2))
  print(chisq.test(table(ds$correct[ds$education==i],ds$atom[ds$education==i])))
}
```

## Impact of Individual Atoms

```{r}

chiTest <- c()
oddsRatio <- c()
mannWhitneyTest <- c()
cliffDelta <- c()

for(i in 1:10){
  subSet <- ds[is.element(el = ds$qid, set = c(i,i+10)),]
  print(atomDescription[i])

  tableCorrectness <- table(subSet$correct,subSet$atom)
  tableTime <- aggregate(ds$time, by=list(atom=ds$atom), FUN=sum)

  print(tableCorrectness)
  print(tableTime)

  test <- chisq.test(tableCorrectness)
  print(test)

  chiTest[i] <- format.pval(test$p.value) 
  oddsRatio[i] <- odds.ratio(tableCorrectness)$OR

  mannWhitneyTest[i] <- format.pval(wilcox.test(subSet$time~as.factor(subSet$atom))$p.value)
  cliffDelta[i] <- cliff.delta(subSet$time~as.factor(subSet$atom))$estimate

  experience <- as.factor(subSet$experience)
  contrr<-matrix(c(rep(1,4),c(1/2,1/2,-1/2,-1/2),c(1,-1,0,0),c(0,0,1,-1)),byrow=TRUE,nrow=4)
  contrasts(experience)<-solve(contrr)[,2:4]

  mod <- glm(subSet$correct ~ subSet$atom +  experience,family="binomial")
  print(summary(mod))
  print(prop.table(table(subSet$correct,subSet$experience),margin = 2))
}

analysis_df <- data.frame(atomDescription, chiTest, oddsRatio, mannWhitneyTest, cliffDelta)
analysis_df <- analysis_df[order(-oddsRatio),]

colnames(analysis_df) <- c("Atom", "ChiTest", "Odds Ratio Correctness", "Wilcox Test (Time)", "Cliff Delta")

analysis_df

xtable(analysis_df)
```

### 2021-01-25 2_Background.tex
sincronizando o github

@@
 Regardless of whether they are performing maintenance on legacy software or adding new features, programmers have to understand what the source code does before making any improvements.
-A natural consequence of this is that a programmer must also have in mind that other programmers (or their future selves) are going to have to spend time trying to understand the code being currently developed before doing any work themselves.
+A natural consequence of this is that a programmer must also have in mind that other programmers (or their future selves)
+%are going to have to
+will
+spend time trying to understand the code being currently developed before doing any work themselves.
 
@@
 block nesting, are atom candidates. 
-This work evaluates the coding practices solely based on personal opinions of the study subjects.
+% Diego: Troquei "this work" para "their work". Removi o "more" de "are more related"
+Their work evaluates the coding practices solely based on personal opinions of the study subjects.
 Furthermore, it focuses on the Java language.
-Finally, most of the practices are more related to code formatting than language constructs or idioms.
+Finally, most of the practices are related to code formatting than language constructs or idioms.
 

### 2021-01-25 3_Method.tex
sincronizando o github

@@
 
-In the firt study we explore the impact of atoms of confusion on
+In the first study we explore the impact of atoms of confusion on
 understanding JavaScript code, by answering two questions: (a) \emph{Do code snippets that contain atoms of confusion produce a higher misunderstanding rate when programmers try to predict their outcome?} and (b) \emph{Do code snippets with atoms of confusion require programmers to take longer to predict their output?}

### 2021-01-25 main.tex
initial commit
update the repository
add a code snippet to make the notion of atoms-of-confusion clearer
updates na secao de MSR
revisao da secao de metodologia
revisao da secao de study settings e resultados
incluindo alguns plots

--- 2020-11-09 PaperICSME.tex
@@
 \documentclass[10pt,conference]{IEEEtran}
+
 \IEEEoverridecommandlockouts
+
 % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+
+\usepackage{enumerate}
+\usepackage{caption}
 \usepackage{cite}
@@
 \usepackage{pifont}
-\usepackage{appendix}
+% \usepackage{appendix}
 \usepackage[framemethod=TikZ]{mdframed}
@@
 % TODO: Completar
-\newcommand{\minedprojects}{X\xspace}
+\newcommand{\minedprojects}{72\xspace}
 
@@
 \begin{abstract}
-{\bf Context.} Evolving legacy code is a challenging task.
-It is even more difficult when the code has been poorly written, imposing significant cognitive load on developers when trying to familiarize themselves with a legacy code base.
+  {\bf Context.} Evolving legacy code is a challenging task, in particular when the code has been poorly written---
+  which imposes significant cognitive load on developers when trying to familiarize themselves with a legacy code base. 
 \noindent{\bf Problem.} 
 %Besides lack of experience and other factors, 
-Previous research works have discussed the impact of small code patterns that contribute to misunderstanding the source code written in statically typed languages (e.g., \clang, \cpplang, and Swift).
-However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamically typed languages.
-\noindent{\bf Research Method and Goal.} In this work, we fill this research gap by presenting the results of a mixed-methods research:
+Previous research have investigatedthe impact of small code patterns
+that contribute to misunderstanding the source code written in \clang and \cpplang.
+However, little is known about the influence of these small code patterns (hereafter atoms of confusion)
+on understanding a dynamic languages (such as JavaScript). 
+\noindent{\bf Research Method and Goal.} In this work, we present the results of a mixed-methods research:
 a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
-Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
-\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: 
-%removing the atoms of confusion lead to an improvement above 15\% in 6 out of 9 atoms we consider in our study when developers try to predict the outcomes of code examples. 
-developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atoms.
-Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the \texttt{comma operator} is present in 11\% of the projects in our dataset; while five other atoms appear in more than 60\% of the projects---including the atoms \texttt{assignment as value} and \texttt{omitted curly braces}, which appear \num{8253} and \num{53357} times, respectively.
-Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code.
-For instance, even experienced JavaScript developers agree that it is hard to understand code that uses callback functions, even though we found widespread use of \texttt{callbacks} in our projects dataset.
-\noindent{\bf Implications.} Altogether, our findings could help developers write cleaner JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code.
+Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language
+whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
+\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead
+to code that is hard to understand: 
+developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not
+present in 6 out of 9 atom candidates.
+Also, the results of our MSR effort reveal that atoms of confusion frequently arise
+in open-source JavaScript systems: the \texttt{comma operator} is present in 11\% of the projects in our dataset; while five other
+atoms appear in more than 60\% of the projects---including the atoms Assignment as Value and Oomitted Curly Braces, which appear 8253 and 53357 times,
+respectively. Our investigation also reveals additional
+JavaScript constructs and idioms that might hinder developers from understanding
+JavaScript code. For instance, even experienced JavaScript
+developers agree that it is hard to understand code that uses callback functions,
+even though we found widespread use of \texttt{callbacks} in our
+projects dataset.
+\noindent{\bf Implications.} Altogether, our findings might help
+developers write cleaner JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code. 
 \end{abstract}
@@
 \input{6_Conclusion}
-\input{7_Atoms}
+% \input{7_Atoms}
 

### 2021-01-26 3_Method.tex
sincronizando com o GH

@@
 We posted the survey on Reddit.\footnote{Reddit is a North American online discussion platform.
-Its discussion threads, often called subreddits, are sorted by subject. For our research, we po\
-sted information and links to the survey in two JavaScript subreddits.} We explained our research purposes and asked developers of any level of expertise to take the survey.
+Its discussion threads, often called subreddits, are sorted by subject.
+For our research, we posted information and links to the survey in two JavaScript subreddits.} We explained our research purposes and asked developers of any level of expertise to take the survey.
 Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares.
@@
 From this initial list, we asked for indication of additional candidates.
-Our main criterion of selection was that all participants shoukd havebeen working professionally with JavaScript.
+Our main criterion of selection was that all participants should have been working professionally with JavaScript.
 We invited a total of 17 developers, and 15 of these agreed to participate.
@@
 
-After filtering out JavaScript project candidates, in the second step we built a curate dataset comprising the top \minedprojects repositories.
+After filtering out JavaScript project candidates, in the second step we built a curated dataset comprising the top \minedprojects repositories.
 Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}.
@@
 CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as Java and JavaScript).
-We also automate the process o
-f running the queries and exporting the results to a format that simplifies our
+We also automate the process of running the queries and exporting the results to a format that simplifies our
 analysis (and also the reproduction of this study). In the final step we carry

### 2021-01-26 4_Result.tex
sincronizando com o GH

@@
 No subject reported to have never attended university.
-Our audience comprises skilful developers, and 93\% of the respondents have more than one year of experience with programming (see Figure~\ref{fig:xp}).
+Our audience comprises skillful developers, and 93\% of the respondents have more than one year of experience with programming (see Figure~\ref{fig:xp}).
 Actually, 56\% of the participants have either between four and ten years of experience (37\%) or more than ten years of experience (19\%).

### 2021-01-27 3_Method.tex
revisao adicional na secao de study settings

@@
 In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice.
-Answering these research questions also lay the foundations for the implementation of libraries that can automatically transform code into its cleaner versions---though we postpone these results to a future research work.
-To answer these research questions we conduct a mixed-methods study, including {\color{red}two surveys}, a set of interviews, and an effort of mining JavaScript software repositories.
+Answering these research questions also lay the foundations for the implementation of tools that can automatically transform code into its cleaner versions---though we postpone these results to a future research work.
+To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and an effort of mining JavaScript software repositories.
 
-\subsection{First Study: Survey-I}
+\subsection{Survey Study Settings}
 
-In the first study we explore the impact of atoms of confusion on
-understanding JavaScript code, by answering two questions: (a) \emph{Do code snippets that contain atoms of confusion produce a higher misunderstanding rate when programmers try to predict their outcome?} and (b) \emph{Do code snippets with atoms of confusion require programmers to take longer to predict their output?}
+In the first study we explore the impact of atom candidates on
+understanding JavaScript code, by answering two questions:
+(a) \emph{Do code snippets that contain atom candidates
+produce a higher misunderstanding rate when programmers
+try to predict their outcome?}
+and (b) \emph{Do code snippets with atom candidates require
+programmers to take longer to predict their output?}
 Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} whose idioms also exist in JavaScript programs.
-Appendix~\ref{sec:appendix-atoms} summarizes all atoms we consider in our research.
+Our supplementary material
+summarizes all atoms we consider in our research.
 
-\subsubsection{Survey design}
+\subsubsection*{Survey design}
 
@@
       \caption{Latin square design. Each ``square'' corresponds to 
-      a replica in our study. Each replica comprises two students (square rows) 
+      a replica in our study. Each replica comprises two participants (square rows) 
       and two sets of code snippets (CS-1 and CS-2). We randomly apply the 
@@
 We also wrote their equivalent snippet without the confusing idioms and constructs, leading to a total of 20 code snippets.
-In order to reduce the cognitive effort, we decided that each subject would be asked to predict the output of a subset of 10 listings, wherein each subset contained 5 blocks that contained atoms of confusion, whilst the remaining 5 had the atoms removed.
+In order to reduce the cognitive effort, we decided that each subject would be asked to predict the output of a subset of 10 code snippets, wherein each subset contained 5 blocks of code that contained atoms of confusion, whilst the remaining 5 had the atoms removed.
 The order in which the questions were presented was randomized.
@@
 
-\subsubsection{Survey Instrument}
+\subsubsection*{Survey Instrument}
 
@@
 
-We organized the survey in {\color{red}three} sections.
+We organized the survey in three sections.
 The first section aims to characterize the subjects, asking for their age, education level, and programming experience.
@@
 
-The next section of the survey presented a sequence of 10 questions, each containing a code snippet.
+The last section of the survey presented a sequence of 10 questions, each containing a code snippet.
 For each question, there was a text box where the answer should be written.
@@
 
-\subsubsection{Survey audience}
+\subsubsection*{Survey audience}
 
@@
 
-\subsection{Second Study: Interviews with Practitioners}
+\subsection{Interview Study Settings}
 
 To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion.
-We also asked each participant if they knew of any other JavaScript-specific construct that they regarded as confusing.
-In this section we details the protocol we followed to conduct the interviews and analyze the results.
+We also asked each participant if they knew of any other JavaScript-specific construct that they regarded as being likely to make the code hard to understand.
+In this section we detail the protocol we followed to conduct the interviews and to analyze the results.
 
@@
 
-\subsubsection{Participants Selection} We invited the participants of the interviews using a snow balling technique.
-That is, starting from our network of contacts, we invited an initial set of candidates to take part of our survey.
-From this initial list, we asked for indication of additional candidates.
-Our main criterion of selection was that all participants should have been working professionally with JavaScript.
-We invited a total of 17 developers, and 15 of these agreed to participate.
-We conducted the interviews during a period of two weeks.
+\subsubsection*{Participants Selection} We invited the participants of the interviews using a snowballing technique.
+That is, starting from our network of contacts, we invited an initial set of candidates to take part in our survey.
+From this initial list, we asked for an indication of additional candidates.
+Our main selection criterion was that all participants should have been working with JavaScript in their daily professional activities.
+We invited a total of 17 developers, and 15 of them agreed to participate.
+%We conducted the interviews for a period of two weeks.
 
-\subsubsection{Interview Process} We conducted semi-structured interviews using a web conferencing software.
-All the interviews were recorded with the consent of the participants.
-On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes, and the longest one 43.06 minutes.
-The main goal of the interviews was to have developers assess in real time whether two versions of code (one with a single atom and another one without the atom) with the same behavior differed in readability, and, if so, which version they regarded as the easier one to understand.
-The interviews were conducted by two of the researchers, and a third one listened to all the recordings to cross-validate the collected data.
+\subsubsection*{Interview Process} We conducted semi-structured interviews using a web conferencing software.
+We recorded all the interviews with the consent of the participants.
+On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes.
+The main goal of the interviews was to have developers assess whether two versions of code (one with a single atom and another one without the atom) with the same behavior differed concerning program comprehension, and, if so, which version they\
+ regarded as the easier one to understand. Two of the researchers conducted the interviews, and a third one listened to all the recordings to cross-validate the collected data.
 
@@
 
-\subsubsection{Interview Analysis}
+\subsubsection*{Interview Analysis}
 
-{\color{red}Nós realizamos a análise das entrevistas em pares.
-Inicialmente um dos participantes de todas as entrevistas transcreveu os resultados.
-Posteriormente, um dos autores do trabalho ouviu os áudios das entrevistas e transcreveu os resultados.}
+We first transcribed each interview and then examined the broad distribution of the answers.
+Our goal was to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion.
+We next followed with an open-coding	procedure, highlighting	the main themes and quoting the participant answers, which we show the most important ones in
+Section~\ref{sec:interview-results}. Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.
 
-\textcolor{red}{Adriano teria algo mais para adicionar aqui na  análise das entrevistas?}
+%\subsection{Third Study: Additional Survey and Interviews}
 
-\todo[inline]{adriano:estou escrevendo uma tabela em que cada linha e um dos atomos, e colunas representam \% preferencia com atomo, \% preferencia sem atomo, \% indiferente.
-Mas acredito que esta tabela deve estar na section results, certo?
-O resto das informacoes deste paragrafo eu citei acima no Interview Process, entao acho que podemos abolir esta section.
--- Edna: Concordo contigo Adriano}
+% \rb{acho que aqui podemos incluir o segundo survey e a segunda
+%  rodada das entrevistas, explorando os \'{a}tomos adicionais}
 
-\subsection{Third Study: Additional Survey and Interviews}
+\subsection{Mining JavaScript Software Repositories Settings}
 
-\rb{acho que aqui podemos incluir o segundo survey e a segunda
-  rodada das entrevistas, explorando os \'{a}tomos adicionais}
-
-\subsection{Fourth Study: Mining Software Repositories}
-
-To understand how often atoms of confusion appear in real settings, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories.
+To understand how often atoms of confusion appear in open source projects, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories.
 To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript.
@@
 
-In the third step we mined atoms of confusion from the repositories in our curate dataset, using source code queries that we had written using the CodeQL language~\cite{moor:gttse2007}.
-CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as Java and JavaScript).
+We mined the occurrence of atom candidates from the repositories in our curate dataset using source code queries that we had written using the CodeQL language~\cite{moor:gttse2007}.
+CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript).
 We also automate the process of running the queries and exporting the results to a format that simplifies our
-analysis (and also the reproduction of this study). In the final step we carry
+analysis (and also the reproduction of this study). Finally, we carried
 out basic descriptive statistics, in order to understand how often the

### 2021-01-27 main.tex
revisao adicional na secao de study settings

@@
 \begin{abstract}
-  {\bf Context.} Evolving legacy code is a challenging task, in particular when the code has been poorly written---
-  which imposes significant cognitive load on developers when trying to familiarize themselves with a legacy code base. 
-\noindent{\bf Problem.} 
-%Besides lack of experience and other factors, 
-Previous research have investigatedthe impact of small code patterns
-that contribute to misunderstanding the source code written in \clang and \cpplang.
-However, little is known about the influence of these small code patterns (hereafter atoms of confusion)
-on understanding a dynamic languages (such as JavaScript). 
-\noindent{\bf Research Method and Goal.} In this work, we present the results of a mixed-methods research:
-a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
-Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language
-whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
-\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead
-to code that is hard to understand: 
-developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not
-present in 6 out of 9 atom candidates.
-Also, the results of our MSR effort reveal that atoms of confusion frequently arise
-in open-source JavaScript systems: the \texttt{comma operator} is present in 11\% of the projects in our dataset; while five other
-atoms appear in more than 60\% of the projects---including the atoms Assignment as Value and Oomitted Curly Braces, which appear 8253 and 53357 times,
-respectively. Our investigation also reveals additional
-JavaScript constructs and idioms that might hinder developers from understanding
-JavaScript code. For instance, even experienced JavaScript
-developers agree that it is hard to understand code that uses callback functions,
-even though we found widespread use of \texttt{callbacks} in our
-projects dataset.
-\noindent{\bf Implications.} Altogether, our findings might help
-developers write cleaner JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code. 
+Evolving legacy code is a challenging task, particularly when the code has been poorly written---which imposes a significant cognitive load on developers when trying to familiarize themselves with a legacy codebase.
+Previous research has investigated the impact of small code patterns that contribute to misunderstanding the source code written in C and C++.
+However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamic languages (such as JavaScript).
+In this work, we present the results of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
+Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
+From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atom candidates.
+Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates
+appear in more than 60\% of the projects---including the atoms Assignment as Value and Omitted Curly Braces, which appear in 8253 and 53357 times,
+respectively.
+Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code.
+For instance, even experienced JavaScript developers agree that it is hard to understand code that uses callback functions, even though we found widespread use of callbacks in our projects dataset.
+Altogether, our findings might help developers better understand the implications of atoms of confusion on understanding JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code.
 \end{abstract}

### 2021-01-28 4_Result.tex
fix the odds ratio report

@@
 For instance, the Comma Operators leads to an \emph{Odds Ratio} of
-\num{19.02}, meaning that the participants are \num{19.02} times
-more likely to correctly interpret the clean version of a code in comparison with the corresponding, confuse version of the code snippet.
+\num{19.02}, meaning that the the odds of a correct answer are \num{19.02} times
+higher when interpreting the clean version of a code---in comparison with the corresponding, confuse version of the code snippet.
 

### 2021-01-28 4_Result.tex
revisao da secao com resultados do survey

@@
 and \emph{cognitive effort} (time necessary to provide a correct answer). We receive full answers to our survey
-from 140 participants. Figure~\ref{fig:degree} and Figure~\ref{fig:xp} show the distributions of the subjects,
-according to their education level and years of programming experience, respectively.
+from 140 participants (in a total of 70 replicas).
+All participants had taken at least some university course or hold a bachelor
+degree or equivalent, meaning the participants have had some
+level of formal education on programming.
+Twenty four participants hold either a master degree (21 participants) or a doctorate degree (3 participant).
 
-\begin{figure}[htb]
-      \centering
-      \includegraphics[width=\columnwidth]{images/dem-education-1.pdf}
-      \caption{Participants' Education Level}\label{fig:degree}
-  \end{figure}
+%% Figure~\ref{fig:degree} and Figure~\ref{fig:xp} show the distributions of the subjects,
+%% according to their education level and years of programming experience, respectively.
 
-  \begin{figure}[htb]
-      \centering
-      \includegraphics[width=\columnwidth]{images/dem-experience-1.pdf}
-      \caption{Participants' Experience} \label{fig:xp}
-  \end{figure}
+%% \begin{figure}[htb]
+%%       \centering
+%%       \includegraphics[width=\columnwidth]{images/dem-education-1.pdf}
+%%       \caption{Participants' Education Level}\label{fig:degree}
+%%   \end{figure}
 
-Figure~\ref{fig:xp} shows that more than a half of the participants hold either a BS degree or had taken some university course, meaning the participants have had some level of formal education on programming.
-No subject reported to have never attended university.
-Our audience comprises skillful developers, and 93\% of the respondents have more than one year of experience with programming (see Figure~\ref{fig:xp}).
-Actually, 56\% of the participants have either between four and ten years of experience (37\%) or more than ten years of experience (19\%).
-Other 37\% of the respondents have between one and four years of experience.
-In this way, we characterize the effect of atoms of confusion in JavaScript code taking into account the perceptions of both novice and experience developers.
+%%   \begin{figure}[htb]
+%%       \centering
+%%       \includegraphics[width=\columnwidth]{images/dem-experience-1.pdf}
+%%       \caption{Participants' Experience} \label{fig:xp}
+%%   \end{figure}
+%Figure~\ref{fig:xp} shows that more than a half of
+
+Our audience comprises professional developers,
+and 93\% of the respondents have more than one year of experience with programming.
+% (see Figure~\ref{fig:xp}). 
+Actually, 56\% of the participants have either between four and ten years of experience (37\%)
+or more than ten years of experience (19\%). Other 37\% of the respondents have between one and four years of experience.
+Accordingly, we characterize the effect of atoms of confusion in JavaScript
+code taking into account the perceptions of both novice and experience developers.
 
@@
 
-Each participant in this study evaluated {\color{red}10} code snippets, from which 5 were in their confusing versions, whilst the other 5 contained cleaned versions of the code snippets (i.e., without the confusing constructs and idioms).
+Each participant in this study evaluated 10 code snippets, from which 5 were in their confusing versions, whilst the other 5 contained cleaned versions of the code snippets (i.e., without the confusing constructs and idioms).
 The participants should provide the expected outcomes of the code snippets.
-We collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{cognitive workload} (the time taken to answer the question).
-Our final dataset consists of 70 complete Latin Squares.
-
-Regarding \emph{correctness}, Table~\ref{results_correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey.
-Accordingly, the version of seven code snippets without atoms of confusion present at least a 15\% improvement in answer correctness.
-In particular, \emph{Comma Operator} atom presents the highest impact on misunderstanding.
-Curiously, frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also introduce high degrees of confusion.
-
-Moreover, as the boxplot in Figure \ref{fig:boxplotcorrectness} shows, there is a considerable decrease in the average number of incorrect answers when observing the cleaned version of the code snippets.
-Also, the sample of answers where there was no atoms had almost no dispersion, which is a sign that the non-confusing code is easier to evaluate correctly.
+As discussed, we collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{cognitive workload} (the time taken to answer the question).
+Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey w.r.t.
+correctness.
+Accordingly, the version of six code snippets without atoms of confusion present at least a 15\% improvement in answer correctness.
+In particular, the presence of the \emph{Comma Operator} atom presents the highest impact on
+misunderstanding, and  
+frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also introduce high degrees of confusion.
+The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a nonnegligible decrease in the average number of incorrect answers when observing the cleaned version of the code snippets.
+Also, the sample of responses with no atoms had almost no dispersion, which signifies that the non-confusing code is easier to evaluate correctly.
 
@@
   Comma Operator &  28 &  65 & +132 & 19.02 \\ 
-  {\color{red}Automatic Semicolon Insertion} &  32 &  68 & +112 & 39.32 \\ 
+%   {\color{red}Automatic Semicolon Insertion} &  32 &  68 & +112 & 39.32 \\ 
   Post Increment &  48 &  64 & +33 & 4.83 \\ 
-  Ommitted Curly Braces &  47 &  58 &  +23 & 2.35 \\
+  Omitted Curly Braces &  47 &  58 &  +23 & 2.35 \\
   Assignment as Value &  56 &  68 &  +21 & 8.38 \\ 
-  Implict Predicate &  58 &  68 &  +17 & 6.94 \\ 
+  Implicit Predicate &  58 &  68 &  +17 & 6.94 \\ 
   Logic as Control Flow &  41 &  48 &  +17 & 1.53 \\ 
@@
 
-\rb{verificar a consistencia nas questoes de pesquisa mais especificas}
+% \rb{verificar a consistencia nas questoes de pesquisa mais especificas}
 
@@
 of the code snippets (confuse and clean code).
-The result shows that the confusing version of the code snippets impacts code understanding (p-value $< 0.05$).
+The result indicates that the confusing version of the code snippets impacts code understanding (p-value $< 0.05$).
 We also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant
@@
 Although the results suggest that experience impact correctness, the education level does not.
-We then replicate the \emph{Pearson's Chi-squared test} for each experience group (see Figure~\ref{fig:xp}) and found that the statistic difference is less significant for those developers with more than ten years of experience.
+As such, we replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistic difference is less significant for those developers with more than ten years of experience.
 Finally, we measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR)
-method (see Table~\ref{tab:difference-correctness}).
+method (see the Effect Size column of Table~\ref{tab:difference-correctness}).
 We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre Increment.
-Nonetheless, for the remaining candidates, the effect size is considerable.
-For instance, the Comma Operators leads to an \emph{Odds Ratio} of
+Nonetheless, for the remaining candidates, the effect size is significant.
+For instance, the Comma Operator leads to an \emph{Odds Ratio} of
 \num{19.02}, meaning that the the odds of a correct answer are \num{19.02} times
 higher when interpreting the clean version of a code---in comparison with the corresponding, confuse version of the code snippet.
+
+\begin{mh}
+  The results of our exploratory data analysis and
+  hypothesis testing suggest that the atom
+  candidates Comma Operator, Post Increment,
+  Omitted Curly Braces, Assignment as Value,
+  Implicit Predicate, and Ternary Operator
+  introduce some degree of misunderstanding
+  in JavaScript code. 
+\end{mh}
 
@@
 Table \ref{tab:difference-time-taken} shows the average time necessary for the participants give a correct answer about the expected outcomes of a code snippet, considering both confuse and clean versions.
-Eight atom candidates require an increasing cognitive workload.
-For these eight atom candidates, developers take at least 5.90\% less time on average to find a correct answer---when considering the clean version of a code snippet.
+Seven atom candidates require an increasing cognitive workload.
+For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer---when considering the clean version of a code snippet.
 In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
@@
 Not all atom candidates, though, take less time to predict the answer. In fact, for the Arithmetic as Logic and the Pre-Increment atoms, there was an increase
-of 29.06\% and 38.19\% in the average time taken to give a correct answwer---wherein the clean code versions.
+of 29.06\% and 38.19\% in the average time taken to give a correct answer---wherein the clean code versions.
 
@@
 Logic as Control Flow & 108.94 & 51.07 & -53.12 &  \colorbox{lightgray}{$<$ 0.01} & -0.39 \\ 
-{\color{red}Automatic Semicolon Insertion} & 46.08 & 22.04 & -52.17 & 0.09 & -0.16 \\ 
-Ommitted Curly Braces & 48.85 & 30.00 & -38.58 & 0.92 & -0.01 \\ 
-Implict Predicate & 36.24 & 24.01 & -33.75 & \colorbox{lightgray}{$<$ 0.01} & -0.29 \\ 
+% {\color{red}Automatic Semicolon Insertion} & 46.08 & 22.04 & -52.17 & 0.09 & -0.16 \\ 
+Omitted Curly Braces & 48.85 & 30.00 & -38.58 & 0.92 & -0.01 \\ 
+Implicit Predicate & 36.24 & 24.01 & -33.75 & \colorbox{lightgray}{$<$ 0.01} & -0.29 \\ 
 Post Increment & 28.70 & 25.67 & -10.56 & 0.12 & 0.15 \\ 
@@
 
-\subsubsection*{Hypotheis Testing}
+\subsubsection*{Hypothesis Testing}
 We use the non-parametric \emph{Mann-Whitney test} to
 investigates the null hypothesis that developers 
-spend the same ammount of time to correctly
+spend the same amount of time to correctly
 predict the outcome of a code snippet, regardless
@@
 shows the results, which suggest that we
-should refute the null hypothes for
+should refute the null hypotheses for
 five atom candidates (Comma Operator, Logic
@@
 
-\rb{Acho que dever\'{i}mos justificar o uso do Cliff's Delta aqui e o Odds Ratio na
-  se\c c\~{a}o anterior.}
-
-\subsection{Discussion}
-
-\rb{Acho que podemos incluir uma discuss\~{a}o ao
-  t\'{e}ermino de cada uma das se\c c\~{o}es com
-  resultados. Ou separamos em uma \'{u}nica
-se\c c\~{a}o de discuss\~{a}o?}
+\begin{mh}
+  The results suggest that developers spend
+  significantly more time to predict the
+  correct answer in the presence of the
+  atom candidates Comma Operator, Logic as
+  Control Flow, and Implicit Predicate. Conversely,
+  the version of the code snippets with the
+  atom candidates Arithmetic as Logic and
+  Pre Increment demands less time to
+  predict the correct answers. 
+\end{mh}
 
@@
 
-\rb{acho que essa separa\c c\~{a}o em dois rounds est\'{a}
-  gerando uma complexidade desnecess\'{a}ria. vou tentar
-  remover isso}.
+%% \rb{acho que essa separa\c c\~{a}o em dois rounds est\'{a}
+%%   gerando uma complexidade desnecess\'{a}ria. vou tentar
+%%   remover isso}.
 
@@
 The following quotes were extracted from the transcripts with
-three inteviwes:
+three interviewees:
 
@@
 they can become quite taxing, cognitively speaking, to understand.
-Therefore, nested ternaries are a choice of atoms to be analysed in future work.
+Therefore, nested ternaries are a choice of atoms to be analyzed in future work.
 
@@
   \hline
-Atom & Projects & Ocurrences/KLOC & Category \\ 
+Atom & Projects & Occurrences/KLOC & Category \\ 
  \hline

### 2021-01-28 dataset/survey-results-final.Rmd
incluindo alguns plots
revisao da secao com resultados do survey

@@
 title: "survey-analysis-final"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 
+demEducation
 barplot(demEducation$total, col=topo.colors(5), ylim = c(0, 70), cex=0.7, cex.lab = 0.7, cex.axis=0.7, ylab="Number of participants")
@@
 oddsRatio <- c()
+oddsRatioTest <- c()
 mannWhitneyTest <- c()
@@
   oddsRatio[i] <- odds.ratio(tableCorrectness)$OR
+  oddsRatioTest[i] <- format.pval(odds.ratio(tableCorrectness)$p)
 
@@
 
-analysis_df <- data.frame(atomDescription, chiTest, oddsRatio, mannWhitneyTest, cliffDelta)
+analysis_df <- data.frame(atomDescription, chiTest, oddsRatio, oddsRatioTest, mannWhitneyTest, cliffDelta)
 analysis_df <- analysis_df[order(-oddsRatio),]
 
-colnames(analysis_df) <- c("Atom", "ChiTest", "Odds Ratio Correctness", "Wilcox Test (Time)", "Cliff Delta")
+colnames(analysis_df) <- c("Atom", "ChiTest", "Odds Ratio Correctness", "p-value", "Wilcox Test (Time)", "Cliff Delta")
 

### 2021-01-29 4_Result.tex
revisao da secao de resultados

@@
   Ternary Operator &  60 &  66 &  +10 & 2.73 \\ 
-  Pre Increment &  50 &  53 &  +6 & 1.24 \\ 
+  Pre-Increment &  50 &  53 &  +6 & 1.24 \\ 
   Arithmetic as Logic &  64 &  63 &  -2 & 0.84 \\ \bottomrule 
@@
 Finally, we measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR)
-method (see the Effect Size column of Table~\ref{tab:difference-correctness}).
-We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre Increment.
-Nonetheless, for the remaining candidates, the effect size is significant.
-For instance, the Comma Operator leads to an \emph{Odds Ratio} of
+method (see the Effect Size column of Table~\ref{tab:difference-correctness}). We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and
+Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. For instance, the Comma Operator leads to an \emph{Odds Ratio} of
 \num{19.02}, meaning that the the odds of a correct answer are \num{19.02} times
@@
 \begin{table*}[htbp]
-\caption{Time in seconds to submit a {\color{red}correct} answer}
+\caption{Time in seconds to submit a correct answer}
 \label{tab:difference-time-taken}
@@
 Arithmetic as Logic & 28.82 & 37.20 & +29.06 & \colorbox{lightgray}{0.01} & 0.23 \\ 
-Pre Increment & 30.71 & 42.45 & +38.19 & \colorbox{lightgray}{$<$ 0.01} & 0.27 \\ 
+Pre-Increment & 30.71 & 42.45 & +38.19 & \colorbox{lightgray}{$<$ 0.01} & 0.27 \\ 
          \bottomrule
@@
 as Control Flow, Implicit Predicate, Arithmetic
-as Logic, and Pre Increment). For the first three,
+as Logic, and Pre-Increment). For the first three,
 the analysis suggest that the participants
@@
 for the atom candidates Arithmetic as Logic and
-Pre Increment, the participants take less time
+Pre-Increment, the participants take less time
 to predict the outcome of the code snippets in
@@
 Logic as Control Flow, Implicit Predicate, Arithmetic
-as Logic, and Pre Increment.
+as Logic, and Pre-Increment.
 
@@
   atom candidates Arithmetic as Logic and
-  Pre Increment demands less time to
+  Pre-Increment demands less time to
   predict the correct answers. 
@@
 experience, familiarity with JavaScript, and their opinion
-about the {\color{red}nine} atom candidates---by
+about the nine atom candidates---by
 (a) presenting them pairs of code snippets whose behavior were identical
@@
 
-Supporting the results of the survey, Tables~\ref{tab:interview-results1}
-and~\ref{tab:interview-results2} show that for twelve out of the
-seventeen scenarios surveyed, the respondents prefer the version of the
-code without the atom of confusion, and in no case the \emph{neutral}
+\subsubsection*{The Participants' perceptions of the atom candidates}
+Supporting the results of the survey, Table~\ref{tab:interview-results1}
+ shows that for eight out of the
+nine scenarios surveyed, the maijority of the respondents agree that the version of the
+code without the atom of confusion is easier to understand, and in no case the \emph{neutral}
 ratio was higher than the option for the version without atoms of confusion.
@@
 %% For a full listing of the code snippets, visit the paper repository. 
-
-\subsubsection*{Participants views of the atoms} Interesting,
-only for the \emph{ternary operator} scenario the participants
+Only for the Ternary Operator atom candidate the participants
 prefer the version of the code with the atom of confusion,
@@
 would be necessary to construct using nested if-then-else statements,
-they can become quite taxing, cognitively speaking, to understand.
-Therefore, nested ternaries are a choice of atoms to be analyzed in future work.
+they can become quite taxing to understand.
+Nested ternaries are a choice of atoms to be analyzed in future work.
 
@@
 
-\rb{pela discuss\~{a}o anterior, parece-me que alguns
-  desenvolvedores reconhecem que esses candidatos
-  a \'{a}tomos introduzem certa dificuldade de compreens\~{a}o;
-  mas mesmo assim optam por usar a constru\c c\~{a}o correspondente.
-  Acho que cabe um box aqui com essa discuss\~{a}o sobre isso.}
+%% \rb{pela discuss\~{a}o anterior, parece-me que alguns
+%%   desenvolvedores reconhecem que esses candidatos
+%%   a \'{a}tomos introduzem certa dificuldade de compreens\~{a}o;
+%%   mas mesmo assim optam por usar a constru\c c\~{a}o correspondente.
+%%   Acho que cabe um box aqui com essa discuss\~{a}o sobre isso.}
 
@@
 the interviewees gave an example of his own experience
-on why one should avoid writing code using this construct:
+that might motivate one avoid writing code using this construct:
 
@@
 
-This is an interesting perspective, because such small, yet tricky, constructs should be avoided, since maintenance is one of the longer and most costly parts of the software development life cycle, and one cannot make any assumptions about the other programmers' experience.
+This is an interesting perspective, because such small, yet tricky program constructs and idioms, might hinder novice developers in the task of maintaining code; and one cannot make any assumption about the programmers' experience of who is going to maintain a code in a long run.
 
-% \subsection{Common Remarks From the Open-ended Questions}
+\begin{table}[!htb]
+    \centering
+    \caption{Interviews Round 1 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}).
+Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
+    \label{tab:interview-results1}
+    \begin{tabular}{lrrr}\toprule
+      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
+      \cmidrule(lr){2-4}
+         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
+                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
+                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
+         Arithmetic as Logic     & 0  & 93.33  & 6.67  \\
+         Assignment as Value     & 20 & 60     & 20    \\
+         Comma Operator          & 0  & 100    & 0     \\
+         Ternary Operator        & 60 & 26.67  & 13.3  \\
+         Implicit Predicate      & 20 & 73.33  & 6.67  \\
+         Logic as Control Flow   & 20 & 60     & 20    \\
+         Omitted Curly Braces    & 0  & 100    & 0     \\
+         Post Increment          & 20 & 73.33  & 6.67  \\
+         Pre-Increment           & 40 & 46.67  & 13.33 \\ \midrule
+         \textsc{overall}        & 20 & 70.37  & 9.63  \\
+         \bottomrule
+    \end{tabular}
+\end{table}
+
+\subsubsection*{Confusion in JavaScript Code}
 
@@
 Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.
-
-\rb{seria importante incluir uma cita\c c\~{a}o para a \'{u}ltima afirma\c c\~{a}o acima.}
 
@@
 
-\rb{esse pr\'{o}ximo par\'{a}grafo eh interessante,
-  se refor\c carmos que os desenvolvedores tem
-  experi\^{e}ncia com a linguagem. caso contr\'{a}rio,
-  parece que nao fizemos uma boa selecao de participantes.
-  algo para discutirmos.}
-
 Moreover, such abstractions might lead a developer to start to program
@@
 For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
-For the multi-purpose language that is one of the most, if not the most used one by practitioners
-~\cite{}, a lack of understanding of the languages core
+For the multi-purpose language that is one of the most, if not the most used one by practitioners,
+a lack of understanding of the languages core
 features may be a problem worth investigating in future studies.
-
-\rb{seria importante apresentar uma conclus\~{a}o em rela\c c\~{a}o a essa parte do estudo.
-Acredito que seja interessante discutir um pouco essa ``faca de dois gumes'' sobre a flexibilidade, mencionada abaixo.
-}
-
-\adriano{Fiquei em duvida se vale citar os contras mais interessantes que os entrevistados falaram sobre JS: (1x) o jeito como ela lida com 'this', (que esta ligado a ela ser puramente orientada a objetos, sem classe); (2x)falta de precisao para tratamento de ponto flutuante, o que pode a tornar inadequada para tratamento de dados/ciencias puras; e a faca de dois gumes que e a flexibilidade da linguagem: muitos falavam isso como ponto positivo e negativo ao mesmo tempo.
-\todo[inline]{eu acho que podemos mencionar sim}}
-
-\begin{table}[!htb]
-    \centering
-    \caption{Interviews Round 1 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}).
-Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
-    \label{tab:interview-results1}
-    \begin{tabular}{lrrr}\toprule
-      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
-      \cmidrule(lr){2-4}
-         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
-                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
-                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
-         Arithmetic as Logic     & 0  & 93.33  & 6.67  \\
-         Assignment as Value     & 20 & 60     & 20    \\
-         Comma Operator          & 0  & 100    & 0     \\
-         Ternary Operator        & 60 & 26.67  & 13.3  \\
-         Implicit Predicate      & 20 & 73.33  & 6.67  \\
-         Logic as Control Flow   & 20 & 60     & 20    \\
-         Omitted Curly Braces    & 0  & 100    & 0     \\
-         Post Increment          & 20 & 73.33  & 6.67  \\
-         Pre Increment           & 40 & 46.67  & 13.33 \\ \midrule
-         \textsc{overall}        & 20 & 70.37  & 9.63  \\
-         \bottomrule
-    \end{tabular}
-\end{table}
 
@@
 (see Table~\ref{tab:difference-correctness}). As such, fixing the the Post Increment and Omitted Curly Braces might solve most of the unclear code caused by atoms of confusion.
-
-\rb{continuar desse ponto}
-
-\subsection{Discussion}
-\label{sec:discussion}

### 2021-01-29 4_Result.tex
sincronizando com o GH

@@
 We mined \minedprojects open source JavaScript repositories to understand how often atoms of confusion arise in real software.
+However, it was not possible to detect the atom Implicit Predicate using the mining techniques employed in this study because it depends on the semantics.
+As such, we were able to mine only the eight atoms listed in Table~\ref{tab:occurrences-summary}.
+
 Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18}~\cite{DBLP:journals/ese/MedeirosLAAKRG19}, which investigates the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion frequently arise in JavaScript open source systems.
@@
 The 5 most frequently found atoms occur in at least 80\% of the projects.
-Considering the extremes, atoms Ternary Operator and Access Property were found in all repositories, while atom Comma Operator occurred only in 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
+Considering the extremes, atoms Ternary Operator was found in all repositories, while atom Comma Operator occurred only in 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
 
@@
 Ternary Operator & 100\% & 10.16 &  highly used \\ 
-  Access Property & 100\% & 3.63 &  highly used \\ 
 Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
 Post Increment & 90.28\% & 5.62 &  highly used \\ 
-  Arrow Function & 83.33\% & 21.10 &  highly used \\ 
-  Change Literal Encoding & 76.39\% & 21.43 &  highly used \\ 
-  Object Destructing & 72.22\% & 2.86 &  highly used \\ 
 Pre-Increment & 83.33\% & 1.04 & commonly used \\ 
 Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
-  Implicit Operand Conversion & 79.17\% & 0.80 & commonly used \\ 
-  Array Destructing & 62.50\% & 0.47 & commonly used \\ 
-  Array Spread & 59.72\% & 0.31 & commonly used \\ 
 Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
-  Object Spread & 55.56\% & 0.64 & commonly used \\ 
 Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
@@
 Considering all JavaScript projects in our dataset, we found 
-a total of \num{615364} atoms of confusion, though five atoms are responsible for 84.65\% of this total: Arrow Function, Change Literal Encoding, Ternary Operator, Omitted Curly Braces and the Post Increment.
-The remaining atoms accounts for \num{94450} (15.35\%) of the total number of occurrences, and the Comma Operator and Arithmetic as Logic are the atoms that arise less frequently (0.0064\% in total with 232 and 165 occurrences, respectively).
+a total of \num{204229} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces and the Post Increment.
+The remaining atoms accounts for \num{24558} (12.03\%) of the total number of occurrences, and the Comma Operator and Arithmetic as Logic are the atoms that arise less frequently (0.001\% in total with 232 and 165 occurrences, respectively).
 
 Regarding frequency, we proceeded with a classification of atom occurrences according to the metrics established by a previous study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}.
-As such, we found 7 atoms as highly used (43.75\%), 7 as commonly used (43.75\%) and only 2 atoms as little used (12.50\%).
-We also calculated the frequency of each atom per KLOC and found atoms Change Literal Encoding and Arrow Function as the most occurring in practice (21.43 and 21.10 occurrences per KLOC, respectively).
+As such, we found 3 atoms as highly used (37.50\%), 3 as commonly used (37.50\%) and 2 atoms as little used (25.00\%).
+We also calculated the frequency of each atom per KLOC and found atom Ternary Operator as the most occurring in practice (10.16 occurrences per KLOC).
 %In our dataset of JavaScript projects, 

### 2021-01-29 discussion.tex
inclusao da secao de discussao

\section{Discussion}
\label{sec:discussion}

Our work leads to several implications.
First, we generalize the
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17},
which investigates
the impact of atom candidates on
understanding \clang and \cpplang code. That is,
similarly to their work, the atom candidates
Comma Operator, Post/Pre Increment, Omitted Curly Braces,
Assignment as Value, Implicit Predicate, Logic as
Control Flow, and Ternary Operator make the 
JavaScript code hard to understand. Our results also 
refute the hypothesis that Arithmetic as Logic is an atom of
confusion (i.e., a source of misunderstanding).
We just find some differences in the effect size
of the atom candidates, in comparison to the original
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.

The results of the interview study complement the understanding of atoms of confusion because the participants make clear the existence of a tradeoff between code comprehension and other quality attributes. For instance, most of the participants prefer the versions of the code with the Ternary Operator, even though they agree that its uses might contribute to the misunderstanding of JavaScript code, particularly when novices are maintaining the codebase. The participants of the interview study also mentioned other possible sources of misunderstanding in JavaScript,
including the use of prototype-inheritance and nested call-backs (as discussed in Section~\ref{sec:interview-results}). Other JavaScript atom candidates include
Object Destructuring, Array Spread, Object Spread, and Type Conversion. We conducted a non-exact replication of the three studies (survey, interview, and mining software repositories) considering these more specific JavaScript atom candidates. We confirmed that they truly correspond to sources of misunderstanding. Due to lack of space,
we cannot present all the results here, and we postpone the presentation of these results as future work.

### 2021-01-29 5_Threats.tex
revisao inicial da secao de ameacas

@@
 
-Since our research was conducted online with unknown participants, we had no way of confirming their levels of education and experience.
-Also, as we briefly discussed in Section 4.3, we did not have a way to prevent respondents from cheating, such as running the code on an interpreter, or consulting other people.
-This presents a threat to the validity of the histograms presented in Figures \ref{fig:degree} and \ref{fig:degree}.
+Since our research was conducted online with unknown participants, and thus we had no way of confirming their
+levels of education and experience. Also,  we did not have a way to prevent respondents from cheating,
+such as running the code on an interpreter, or consulting other people. This presents a threat to the validity to the
+conclusions of our work, even though we leverage some means to try to mitigate this threat (as discussed
+in Section~\ref{method}).
 
-Another threat to our results lies in the fact that some constructs might not be frequent, or present at all, in JavaScript code bases.
-The Automatic Semicolon Insertion and the Comma Operator atoms were not found in the code samples we consulted when formulating the questions, and we had to create blocks of code containing the atoms.
-The work done in Chapter 5 could serve as a starting point for detecting these atoms in a large data set of JavaScript files.
+Another threat to our results lies in the fact that some constructs
+might not be frequent, or present at all, in JavaScript code bases.
+The atom candidates Arithmetic as Logic and
+the Comma Operator were rarely found in the
+open source JavaScript code we mined in our
+third study.
 
-Regarding hypotheses H.2 \todo[inline]{Linkar as Hipoteses} and research question RQ.2, a potential problem with our method is that there was no strong incentive for subjects to think thoroughly about the questions.
-We observed a lack of engagement when we ran the survey with undergraduate students.
-Although our final subjects were voluntarily partaking in the survey, we could not be sure that, after some time taking the survey, respondents would become tired and stop thinking clearly about the code.
-As questions were assigned randomly, and contained different levels of difficulty, comparing a subject's answer with their other questions would not provide a reliable metric.
+A potential problem with our method is that
+there little incentive for subjects to think
+thoroughly about the questions. We observed a lack of
+engagement when we ran the survey with undergraduate students during
+a pilot study. Although our final subjects were voluntarily
+partaking in the survey, we could not be sure that, after some
+time taking the survey, respondents would become tired and stop thinking
+clearly about the code.
 
-In one of the atoms, namely the Omitted Curly Braces, we intentionally removed indentation from the code, which is highly unusual, given that many programmers use automatic formatting in their code editors.
-This can introduce some level of artificiality to this atom's question.
-On the other hand, while describing JavaScript's grammar to develop the parser we used for the work in Chapter 5, we discovered that the language allows the programmer to omit the curly braces after \textit{if} statements, and insert multiple statements in the following line.
-This fact itself might constitute a source of confusion, which we leave to analyse in our future endeavors.
+Finally, in one of the atoms, namely the Omitted Curly Braces,
+we intentionally removed indentation from the original code,
+which is highly unusual, given that many programmers use automatic
+formatting in their code editors. This can introduce some level
+of artificiality to this atom's question. Nonetheless,
+we discovered that JavaScript does allow the programmer
+to omit the curly braces after \textit{if} statements,
+and insert multiple statements in the following line.
+This fact itself might constitute
+a source of confusion, which we leave to analyse
+in our future endeavors.

### 2021-04-05 4_Result.tex
Updates from Overleaf

@@
  shows that for eight out of the
-nine scenarios surveyed, the maijority of the respondents agree that the version of the
+nine scenarios surveyed, the majority of the respondents agree that the version of the
 code without the atom of confusion is easier to understand, and in no case the \emph{neutral}

### 2021-04-08 3_Method.tex
ajustes nas issues apontadas por Marcio

@@
 \begin{enumerate}[(RQ1)]
-    \item Do JavaScript developers identify atoms of confusion as contributing to program misunderstanding? 
-    \item What is the impact of atoms of confusion on the comprehension of JavaScript code? 
-    \item What are the particular JavaScript idioms and language constructs that might cause source code misunderstanding?
-    \item What is the frequency of occurrence of atoms of confusion in practice (i.e., in open-source JavaScript projects)?
+\item \rqa 
+\item \rqb
+\item \rqc  
+\item \rqd
 \end{enumerate}
@@
 
-In the first study we explore the impact of atom candidates on
-understanding JavaScript code, by answering two questions:
-(a) \emph{Do code snippets that contain atom candidates
-produce a higher misunderstanding rate when programmers
-try to predict their outcome?}
-and (b) \emph{Do code snippets with atom candidates require
-programmers to take longer to predict their output?}
+In the first study, we explore the impact of atom candidates on
+understanding JavaScript code (goal of research question RQ1).	
+To this end, we investigate whether or not code snippets that
+contain atom candidates (a) produce a higher
+misunderstanding rate when programmers
+try to predict their outcome and (b) require
+programmers to take longer to predict their output.
 Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} whose idioms also exist in JavaScript programs.
@@
 We also wrote their equivalent snippet without the confusing idioms and constructs, leading to a total of 20 code snippets.
-In order to reduce the cognitive effort, we decided that each subject would be asked to predict the output of a subset of 10 code snippets, wherein each subset contained 5 blocks of code that contained atoms of confusion, whilst the remaining 5 had the atoms removed.
+In order to reduce the cognitive effort, we decided that each subject would be asked to predict the output of a subset of 10 code snippets, wherein each subset contained five blocks of code that contained atoms of confusion, whilst the remaining five had the atoms removed.
 The order in which the questions were presented was randomized.
@@
 
-To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion.
-We also asked each participant if they knew of any other JavaScript-specific construct that they regarded as being likely to make the code hard to understand.
+To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion (research question RQ2).
+We also asked each participant if they knew of any other JavaScript-specific construct that they regarded as being likely to make the code hard to understand (research question RQ3).
 In this section we detail the protocol we followed to conduct the interviews and to analyze the results.
@@
 This would allow us to identify potential atom candidates that are more specific to the JavaScript language.
-Examples of questions we explored in this section include:\emph{Do you think JavaScript is a language that produces code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as
+Examples of questions we explored in this section include: \emph{Do you think JavaScript is a language that produces code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as
   especially confusing?}
@@
 The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
-All projects in our dataset have at least \num{1244} forks and \num{23672} stars.
+All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars.
 We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.

### 2021-04-08 4_Result.tex
ajustes nas issues apontadas por Marcio

@@
 level of formal education on programming.
-Twenty four participants hold either a master degree (21 participants) or a doctorate degree (3 participant).
+Twenty four participants hold either a master degree (21 participants) or a doctorate degree
+(three participants).
 
@@
 
-Each participant in this study evaluated 10 code snippets, from which 5 were in their confusing versions, whilst the other 5 contained cleaned versions of the code snippets (i.e., without the confusing constructs and idioms).
+Each participant in this study evaluated 10 code snippets, from which five were in their confusing versions, whilst the other five contained cleaned versions of the code snippets (i.e., without the confusing constructs and idioms).
 The participants should provide the expected outcomes of the code snippets.
 As discussed, we collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{cognitive workload} (the time taken to answer the question).
-Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey w.r.t.
-correctness.
-Accordingly, the version of six code snippets without atoms of confusion present at least a 15\% improvement in answer correctness.
+Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey w.r.t. correctness.
+
+Considering Table~\ref{tab:difference-correctness}, the version of six code snippets without atoms of confusion present at least a 15\% improvement in answer correctness.
 In particular, the presence of the \emph{Comma Operator} atom presents the highest impact on
@@
 The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a nonnegligible decrease in the average number of incorrect answers when observing the cleaned version of the code snippets.
-Also, the sample of responses with no atoms had almost no dispersion, which signifies that the non-confusing code is easier to evaluate correctly.
+Also, the sample of responses with no atoms had almost no dispersion, which supports the
+argument that the non-confusing versions of the code snippets are
+easier to evaluate correctly.
 
@@
 
-Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand 'asynchronous' programming in JavaScript.
+Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript.
 Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.
@@
 
-We mined \minedprojects open source JavaScript repositories to understand how often atoms of confusion arise in real software.
-However, it was not possible to detect the atom Implicit Predicate using the mining techniques employed in this study because it depends on the semantics.
-As such, we were able to mine only the eight atoms listed in Table~\ref{tab:occurrences-summary}.
+We mined \minedprojects open source JavaScript repositories to understand how often atoms of confusion arise in real software. We discarded in this analysis the atom Implicit Predicate,
+because writing a CodeQL query for mining this particular atom is everything but trival.
+As such, in this study
+we only mined the eight atoms listed in Table~\ref{tab:occurrences-summary}.
 
 Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18}~\cite{DBLP:journals/ese/MedeirosLAAKRG19}, which investigates the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion frequently arise in JavaScript open source systems.
-
-The 5 most frequently found atoms occur in at least 80\% of the projects.
-Considering the extremes, atoms Ternary Operator was found in all repositories, while atom Comma Operator occurred only in 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
+For instance, the 5 most frequently found atoms occur in at least 80\% of the projects.
+Considering the extremes, atom Ternary Operator was found in all repositories,
+while atom Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
 
@@
 Considering all JavaScript projects in our dataset, we found 
-a total of \num{204229} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces and the Post Increment.
-The remaining atoms accounts for \num{24558} (12.03\%) of the total number of occurrences, and the Comma Operator and Arithmetic as Logic are the atoms that arise less frequently (0.001\% in total with 232 and 165 occurrences, respectively).
+a total of \num{204229} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces and the Post Increment. The remaining atoms
+account for \num{24558} (12.03\%) of the total number of occurrences, and the Comma Operator and Arithmetic as Logic are the atoms that arise less frequently (0.001\% in total with 232 and 165 occurrences, respectively).
 
-Regarding frequency, we proceeded with a classification of atom occurrences according to the metrics established by a previous study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}.
-As such, we found 3 atoms as highly used (37.50\%), 3 as commonly used (37.50\%) and 2 atoms as little used (25.00\%).
-We also calculated the frequency of each atom per KLOC and found atom Ternary Operator as the most occurring in practice (10.16 occurrences per KLOC).
+Regarding frequency, we proceeded with a classification of \emph{atom occurrence}
+according to the thresholds of a previous
+study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. As such, we found three
+atoms as highly used (Ternary Operator, Omitted Curly Braces, and Post Increment),
+three as commonly used (Pre-Increment, Assignment as Value, and Logic as Control Flow)
+and two atoms as little used (Arithmetic as Logic and Comma Operator).
+We also calculated the frequency of each atom per KLOC and found the atom
+Ternary Operator as the most occurring in practice (10.16 occurrences per KLOC).
 %In our dataset of JavaScript projects, 
@@
 Nonetheless, the Post Increment Expression and Omitted Curly Braces atoms are listed in the top three sources of misunderstanding
-(see Table~\ref{tab:difference-correctness}). As such, fixing the the Post Increment and Omitted Curly Braces might solve most of the unclear code caused by atoms of confusion.
+(see Table~\ref{tab:difference-correctness}). As such, fixing the the Post Increment and Omitted Curly Braces might solve most of the unclear code caused by atoms of confusion
+in JavaScript projects.

### 2021-04-08 5_Threats.tex
ajustes nas issues apontadas por Marcio

@@
 
-Since our research was conducted online with unknown participants, and thus we had no way of confirming their
-levels of education and experience. Also,  we did not have a way to prevent respondents from cheating,
-such as running the code on an interpreter, or consulting other people. This presents a threat to the validity to the
-conclusions of our work, even though we leverage some means to try to mitigate this threat (as discussed
-in Section~\ref{method}).
+%% Since our research was conducted online with unknown participants, and thus we had no way of confirming their
+%% levels of education and experience. Also,  we did not have a way to prevent respondents from cheating,
+%% such as running the code on an interpreter, or consulting other people. This presents a threat to the validity to the
+%% conclusions of our work, even though we leverage some means to try to mitigate this threat (as discussed
+%% in Section~\ref{method}).
 
-Another threat to our results lies in the fact that some constructs
-might not be frequent, or present at all, in JavaScript code bases.
-The atom candidates Arithmetic as Logic and
-the Comma Operator were rarely found in the
-open source JavaScript code we mined in our
-third study.
+%% Another threat to our results lies in the fact that some constructs
+%% might not be frequent, or present at all, in JavaScript code bases.
+%% The atom candidates Arithmetic as Logic and
+%% the Comma Operator were rarely found in the
+%% open source JavaScript code we mined in our
+%% third study.
 
-A potential problem with our method is that
-there little incentive for subjects to think
-thoroughly about the questions. We observed a lack of
-engagement when we ran the survey with undergraduate students during
-a pilot study. Although our final subjects were voluntarily
-partaking in the survey, we could not be sure that, after some
-time taking the survey, respondents would become tired and stop thinking
-clearly about the code.
+%% A potential problem with our method is that
+%% there little incentive for subjects to think
+%% thoroughly about the questions. We observed a lack of
+%% engagement when we ran the survey with undergraduate students during
+%% a pilot study. Although our final subjects were voluntarily
+%% partaking in the survey, we could not be sure that, after some
+%% time taking the survey, respondents would become tired and stop thinking
+%% clearly about the code.
 
-Finally, in one of the atoms, namely the Omitted Curly Braces,
-we intentionally removed indentation from the original code,
-which is highly unusual, given that many programmers use automatic
-formatting in their code editors. This can introduce some level
-of artificiality to this atom's question. Nonetheless,
-we discovered that JavaScript does allow the programmer
-to omit the curly braces after \textit{if} statements,
-and insert multiple statements in the following line.
-This fact itself might constitute
-a source of confusion, which we leave to analyse
-in our future endeavors.
+%% Finally, in one of the atoms, namely the Omitted Curly Braces,
+%% we intentionally removed indentation from the original code,
+%% which is highly unusual, given that many programmers use automatic
+%% formatting in their code editors. This can introduce some level
+%% of artificiality to this atom's question. Nonetheless,
+%% we discovered that JavaScript does allow the programmer
+%% to omit the curly braces after \textit{if} statements,
+%% and insert multiple statements in the following line.
+%% This fact itself might constitute
+%% a source of confusion, which we leave to analyse
+%% in our future endeavors.

### 2021-04-08 discussion.tex
ajustes nas issues apontadas por Marcio

@@
 confusion (i.e., a source of misunderstanding).
-We just find some differences in the effect size
+In comparison to the previous work, our study
+led to some differences in the effect size
 of the atom candidates, in comparison to the original
 work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
+Altogether, we answer our first research question
+\emph{\rqa}:
 
-The results of the interview study complement the understanding of atoms of confusion because the participants make clear the existence of a tradeoff between code comprehension and other quality attributes.
-For instance, most of the participants prefer the versions of the code with the Ternary Operator, even though they agree that its uses might contribute to the misunderstanding of JavaScript code, particularly when novices are maintaining the codebase.
-The participants of the interview study also mentioned other possible sources of misunderstanding in JavaScript,
+\begin{mh}
+  {\bf Answer to RQ1:} The first study (survey) give evidence that
+  the atom candidates for \clang and \cpplang
+  that also exist for JavaScript correspond to
+  a source of misunderstanding in
+  JavaScript code. 
+\end{mh}
+
+The results of the interview study complement the understanding
+of atoms of confusion because the participants make
+clear the existence of a trade-off between code comprehension
+and other quality attributes. For instance, most of the participants
+prefer the version of the code with the Ternary Operator,
+even though they agree that its uses might contribute to the
+misunderstanding of JavaScript code, particularly when novices are maintaining
+the codebase. The participants of the interview study also
+mentioned other possible sources of misunderstanding in JavaScript,
 including the use of prototype-inheritance and nested call-backs (as discussed in Section~\ref{sec:interview-results}). Other JavaScript atom candidates include
 Object Destructuring, Array Spread, Object Spread, and Type Conversion.
-We conducted a non-exact replication of the three studies (survey, interview, and mining software repositories) considering these more specific JavaScript atom candidates.
-We confirmed that they truly correspond to sources of misunderstanding.
-Due to lack of space,
-we cannot present all the results here, and we postpone the presentation of these results as future work.
+In summary, the results of the second study (interviews) allow
+us to answer the second (\rqb) and third research questions
+(\emph{\rqc}).
+
+\begin{mh}
+  {\bf Answer to RQ2:} The qualitative analysis of the
+  interviews supports the results of the first study,
+  indicating that atoms of confusion compromise
+  source code understanding. 
+\end{mh}
+
+\begin{mh}
+  {\bf Answer to RQ3:} The qualitative analysis of the
+  interviews suggest that specific constructs of
+  JavaScript might also correspond to atoms of
+  confusion, including prototype inheritance and
+  nested callbacks. 
+\end{mh}
+
+The result of the third study (mining open source
+Java script repositories) give evidence that,
+although atoms of confusion compromise program
+comprehension, they frequently appear in open
+source JavaScript projects. In particular,
+{\color{red}six, out of nine atoms considered
+  in our study,} appear in more than 50\% of
+the projects. To sum up, the third study
+allows us to answer the fourth research
+question (\emph{\rqd}).
+
+\begin{mh}
+  {\bf Answer to RQ4:} The MSR study reveals that
+  the atoms of confusion explored in our research
+  appear frequently in practice. 
+\end{mh}
+
+%% We conducted a non-exact replication of the three
+%% studies (survey, interview, and mining software repositories)
+%% considering these more specific JavaScript atom candidates.
+%% We confirmed that they truly correspond to sources of misunderstanding.
+%% Due to lack of space,
+%% we cannot present all the results here, and we postpone the presentation of these results as future work.

### 2021-04-08 main.tex
inclusao da secao de discussao
revisao inicial da secao de ameacas
ajustes nas issues apontadas por Marcio

@@
-\documentclass[10pt,conference]{IEEEtran}
+\documentclass[10pt,conference,blind]{IEEEtran}
 
@@
 
+\newcommand{\rqa}{What is the impact of atoms of confusion on the comprehension of
+  JavaScript code?}
+
+\newcommand{\rqb}{Do JavaScript developers identify atoms of confusion as contributing to program misunderstanding?}
+
+\newcommand{\rqc}{What are the particular JavaScript idioms and language constructs that might cause source code misunderstanding?}
+
+\newcommand{\rqd}{What is the frequency of occurrence of atoms of confusion in practice (i.e., in open-source JavaScript projects)?}
+
 \lstset{language=C}
@@
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-}
+\author{\IEEEauthorblockN{Anonymous}}
+
+%% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% \and
+%% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% \and
+%% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% \and
+%% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% \and
+%% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% \and
+%% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% }
 
@@
 \input{4_Result}
+\input{discussion}
 \input{5_Threats}
@@
 
-\section*{Acknowledgment}
+% \section*{Acknowledgment}
 
-\section*{References}
 \bibliographystyle{IEEEtran}

### 2021-04-19 1_Introduction.tex
text improvements
revisão do Prof. Fernando Castor

--- 2020-11-11 1_Introduction.tex
@@
 JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}.
-At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''\footnote{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}.
-It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.
+At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''.\footnote{\url{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}} It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.
 All these
@@
 
-Our research aims to help practitioners identify potentially confusing snippets of code, so that
-developers could replace them with clearer code.
+Our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer.
 To achieve that, we investigate {\color{red}XYZ} JavaScript idioms that
-might compromize program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in
+might negatively impact program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in
 JavaScript---that is, small
@@
 for which there are functionally equivalent alternatives that are easier
-to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
-\castor{I think it would be really nice to have an example here.
-Maybe even a figure with a left-hand and a right-hand side.
-What do you say?
-If yes, I
-'d go for post increment or assignment as value.
-If not a figure, the text could be as follow: An example of atom of confusion identified by previous work~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} is the use \texttt{if} statements without the accompanying curly braces.
-The alternative, in this case, is simply to always use the curly braces.}
-Among the atoms in this work, X are shared between JavaScript and \clang/\cpplang (e.g., pre/post-increment), while Y others are specific to JavaScript, e.g., the atom Assignment as Value.
-Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17} has
-already explored the impact of atoms of confusion on program understanding,
-and confirmed they negatively impact source code understandability.
-However, they were constrained to statically typed languages (e.g., C and C++).
-To the best of our knowledge, we are the first to explore atoms of confusion
-in the context of a dynamically typed language (JavaScript).\castor{Review the number of atoms, X, and Y.}
+to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For instance, consider
+the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
+%for presentation purpose only). 
+The code on the left-hand side of the figure contains two atoms of confusion discussed in previous work focusing on the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}.
+On the right-hand side, we present a cleaned version of the code, that is, without atoms of confusion.
+In the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in  41\% and 34\%, respectively.
 
-To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 developers from the Reddit community.
+Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability.
+However, these previous studies focus exclusively on statically typed languages (e.g., \clang and \cpplang).
+To the best of our knowledge, we are the first to explore atoms of confusion in the context of a dynamically typed language (JavaScript).\castor{Review the number of atoms, X, and Y.}  JavaScript's programming culture differs from that of languages such as C and Java, e.g., because there are no types and due to the pervading use of callbacks.
+Thus, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa.
+\castor{The following should probably be left out.
+I don't think we investigate atoms outside of those that were already studied in C.
+Or maybe say explicitly that we are studying a group of atoms that stems from the previous study?}
+Furthermore, among the atom candidates we investigate in this work, {\color{red}X} are shared between JavaScript and \clang/\cpplang (e.g., pre/post-increment), while {\color{red}Y} others are specific to JavaScript, e.g., the atom Assignment as Value.
+
+\begin{figure*}[thb]
+\noindent\begin{minipage}{.45\textwidth}  
+\begin{lstlisting}[language=JavaScript]
+res = properties.reduce((s, b) => {
+  b && s++
+  return s
+}
+\end{lstlisting}
+\end{minipage}\hfill
+\begin{minipage}{.45\textwidth}
+\begin{lstlisting}[language=JavaScript]
+res = properties.reduce((s, b) => {
+  if(b) {
+    s = s + 1
+  }
+  return s
+}
+\end{lstlisting}
+\end{minipage}
+\captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project. Line 2 on the left-hand side of the figure contains two
+  atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic
+  as Control Flow} and \emph{Post-increment}.
+On the right-hand side, we present an alternative implementation without atoms of confusion.
+We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p).length}).}\label{fig:lst01}
+\end{figure*}
+
+To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 developers from the Reddit community.\footnote{\url{https://www.reddit.com/r/javascript/comments/bw36h6/academic_research_on_confusing_code_on_javascript/}}
 Their task was to predict the output of short snippets of code, with half of them containing atoms of confusion. 
-We found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of 9 atoms).
+\castor{Improvement of 15\% in terms of what.
+This is too loose.} We found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of 9 atoms).
 The presence of the Comma Operator atom, for instance, can increase the average time to predict the output by 65\%, while the behavior of snippets containing the Pre-Increment atom were predicted 44\% faster.
-An atom's presence---or lack of---in open-source repositories may reveal cases where developers should be more or less careful when writing code.
-To identify such cases, we mined the 100 most popular JavaScript projects from GitHub and found a total of X atoms in the analyzed projects.
+The presence or absence of atoms in open-source repositories may reveal cases where developers should be more or less careful when writing code.
+To identify such cases, we mined the 100 most popular JavaScript projects from GitHub and found a total of {\color{red}X} atoms in the analyzed projects.
 To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15 experienced professional developers.
-Participants preferred code snippets \textit{without} an atom in 70\% of the cases, therefore corroborating to the survey findings.
+Participants preferred code snippets \textit{without} an atom in 70\% of the cases, therefore corroborating with the survey findings.
 During the MSR and interview studies, we also explored additional atom candidates that are specific to the

### 2021-04-19 2_Background.tex
inclusao da secao de discussao
revisão do Prof. Fernando Castor

@@
 \label{back}
-\subsection{Program Comprehension}\castor{It would be nice to have a table with the atom candidates similar to the one that appears in Gopstein et al. 2017.}
+\subsection{Program Comprehension}
+
+% \castor{It would be nice to have a table with the atom candidates
+% similar to the one that appears in Gopstein et al. 2017.}
 
@@
 
-However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions---when these exist---allowed by the language or framework they are working with.
-Although this can be, to an extent, a personal matter, since different individuals might hold different opinions on whether a given syntax is difficult to understand, it is possible to empirically investigate if there exist programming languages' constructs or idioms that are considerably harder to understand, leading readers of the code to frequently make incorrect predictions of its behavior.
+However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions---when these exist---that are potentially less complex.
+%allowed by the language or framework they are working with.
+This can be a matter of personal taste, since different individuals might hold different opinions on whether a given syntax is difficult to understand.
+Nevertheless, it is possible to empirically investigate if there exist programming language constructs or idioms that are considerably harder to understand, leading readers of the code to frequently make incorrect predictions of its behavior.
 
@@
 This occurs because a leading \texttt{0} in a number literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers.
-After formulating a list of 19 potential atoms, Gopstein et al.
-\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 snippets that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed.
+After formulating a list of 19 atom candidates, Gopstein et al.
+\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 code idioms and language constructs that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed.
 The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase.
@@
 Gopstein et al.
-\cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on the different nature of confusion amongst the atoms, noting that, for some atoms, a particular incorrect answer was prevalent, whilst there was a distribution of incorrect ones in other atoms.
+\cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on how confusion manifested when developers inspected code exhibiting different atoms.\castor{I'd rather have two concrete examples here, instead of this.} They note that, for some atoms, a particular incorrect answer was prevalent, whilst there was a variety of incorrect answers for other atoms.
 
@@
 Also, both languages share many syntactical rules, which allows us to replicate some of the C atoms in JavaScript.
+
+The first study of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} inspired other studies to further investigate the impact of atoms of confusion on code comprehension.
+Oliveira et al.~\cite{TheEyesDoNotLie} conducted an experiment using an eye-tracker with 30 participants involving three atoms.
+They found that code with atoms of confusion requires more time from participants to predict the output and more visual effort to comprehend.
+Gopstein et al., in another paper~\cite{ThinkingAloud} further explored the original atoms by additionally conducting interviews and having programmers discuss between themselves the studied atoms.
+The authors argue that this method is important to complement a quantitative analysis.
+Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''
+}. \castor{We use interviews, but did we "test" the interviewees to verify if they understand code?} We follow a similar approach as we augment our quantitative analysis with interviews.
+
+Gopstein et al. also influenced researchers to look into atoms in languages other than \clang and \cpplang.
+Langhout and Aniche~\cite{Langhout:2021:ACJ} derived a set of 14 atoms for Java and performed an experiment with 132 students.
+They found that participants are more likely to make mistakes (ranging from 2.7 up to 56 times) when atoms are present in half of their studied atoms.
+Castor~\cite{castor2018} presented a preliminary catalog of six atom candidates for the Swift programming language.
+Unlike JavaScript, in Swift, most of the atoms identified by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} are avoided by construction, e.g., it does not have assignments as values, increment operators, or macros.
+The paper also presents a structured definition for what it means for a code pattern to be an atom of confusion and a set of general principles to help in the identification of new atoms.
 
@@
 
-Castor~\cite{castor2018} presented a preliminary catalog of six atom candidates for the Swift programming language.
+% Castor~\cite{castor2018} presented a preliminary catalog of six atom candidates for the Swift programming language.
 Unlike JavaScript, in Swift most of the atoms identified by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} are avoided by construction, e.g., it does not have assignments as values, increment operators, or macros.
@@
 
-Santos and Gerosa~\cite{Santos:2018:ICP} analyze the impact of eleven coding practices on program readability. Some of them, such as avoiding more than three levels of code
-block nesting, are atom candidates. 
+Santos and Gerosa~\cite{Santos:2018:ICP} analyze the impact of eleven coding practices on program readability.
+Some of them, such as avoiding more than three levels of code block nesting, are atom candidates.
 % Diego: Troquei "this work" para "their work". Removi o "more" de "are more related"
@@
 Furthermore, it focuses on the Java language.
-Finally, most of the practices are related to code formatting than language constructs or idioms.
+Finally, most of the practices analyzed in this study are more related to code formatting than language constructs or idioms.
 

### 2021-04-19 3_Method.tex
revisão do Prof. Fernando Castor

@@
-\section{Study Settings}
+\section{Methodology}
 \label{method}
 
-The main goal of this research is to investigate the impact of atoms of confusion in JavaScript code comprehension. As such, in this paper we answer the following research questions:
+The main goal of this research is to investigate the impact of atoms of confusion on JavaScript code comprehension. As such, in this paper we answer the following research questions:
 
@@
 Answering these research questions also lay the foundations for the implementation of tools that can automatically transform code into its cleaner versions---though we postpone these results to a future research work.
-To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and an effort of mining JavaScript software repositories.
+To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and a JavaScript software repository mining study.
 
-\subsection{Survey Study Settings}
+\subsection{Survey Study Setting}\label{sec:meth:survey}
 
@@
 code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about non-confusing code snippets, they will only be presented with non-confusing versions of the code samples $s_6, s_7,..., s_{10}$.
-Furthermore, a given subject (P2), which constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing snippets of for $s_6, s_7,..., s_{10}$.
+Furthermore, a given subject (P2), which constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing snippets for $s_6, s_7,..., s_{10}$.
 By doing that, we guarantee that all versions of code snippets are contained within each square, and that each configuration occurs only once within a square.
@@
       \caption{Latin square design. Each ``square'' corresponds to 
-      a replica in our study. Each replica comprises two participants (square rows) 
+      a replica in our study. Each replica comprises two participants (square rows, e.g., P1 and P2) 
       and two sets of code snippets (CS-1 and CS-2). We randomly apply the 
@@
 
-Having selected 10 atoms in our first study, we wrote small pieces of code that contained each atom.
-We also wrote their equivalent snippet without the confusing idioms and constructs, leading to a total of 20 code snippets.
-In order to reduce the cognitive effort, we decided that each subject would be asked to predict the output of a subset of 10 code snippets, wherein each subset contained five blocks of code that contained atoms of confusion, whilst the remaining five had the atoms removed.
+Having selected 10 atom candidates in our first study, we wrote 10 short programs, one containing each atom candidate.
+We call them the confusing versions of the programs.
+We also wrote corresponding short programs without the confusing idioms and constructs and we call them the clean versions of the programs.
+Overall, the survey employed 20 code snippets, 10 confusing and 10 clean.
+In order to reduce the cognitive effort, each subject was asked to predict the output of 10 code snippets, five confusing and five clean.
 The order in which the questions were presented was randomized.
@@
 
-We implemented our survey as a web application. As part of this 
-effort, we carried out an informal pilot whose main objectives were: to spot bugs in the application and in the data collection mechanism; to gain feedback from respondents about the user experience of the application; and to formulate an estimate about how long answering the survey would take on average.
+We implemented our survey as a web application.
+As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
 Fellow undergraduate students, professional colleagues, and friends took the pilot survey.
@@
 We organized the survey in three sections.
-The first section aims to characterize the subjects, asking for their age, education level, and programming experience.
+The first section aims to characterize the subjects, asking their age, education level, and programming experience.
 We also included a check button, whose checking meant users agreed that all collected data would be used solely for research purposes.
@@
 
-The last section of the survey presented a sequence of 10 questions, each containing a code snippet.
+The last section of the survey presented a sequence of ten questions, each containing a code snippet.
 For each question, there was a text box where the answer should be written.
@@
 
-\subsection{Interview Study Settings}
+\subsection{Interview Study Setting}
 
 To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion (research question RQ2).
-We also asked each participant if they knew of any other JavaScript-specific construct that they regarded as being likely to make the code hard to understand (research question RQ3).
+We also asked each participant if they knew of any other JavaScript-specific construct or idiom that they regarded as being likely to make the code hard to understand (research question RQ3).
 In this section we detail the protocol we followed to conduct the interviews and to analyze the results.
@@
 
-\subsubsection*{Participants Selection} We invited the participants of the interviews using a snowballing technique.
+\subsubsection*{Participant Selection} We invited the participants of the interviews using a snowballing technique.
 That is, starting from our network of contacts, we invited an initial set of candidates to take part in our survey.
@@
 On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes.
-The main goal of the interviews was to have developers assess whether two versions of code (one with a single atom and another one without the atom) with the same behavior differed concerning program comprehension, and, if so, which version they\
- regarded as the easier one to understand. Two of the researchers conducted the interviews, and a third one listened to all the recordings to cross-validate the collected data.
 
+%In the interviews we presented developers with two functionally equivalent short programs, one with a single atom and another one without the atom, differed concerning program comprehension.
+We then asked them to tell us which program was easier to understand.
+%The main goal of the interviews was to have developers assess whether two functionally equivalent short programs, one with a single atom and another one without the atom, differed concerning program comprehension, and, if so, which version they regarded as the easier one to understand.
+
+Two of the researchers conducted the interviews, and a third one listened to all the recordings to cross-validate the collected data.
 The interviews had three main parts.
@@
 
-In the second part of the interview our aim was to allow the subjects to describe
-their JavaScript experience, as well as to allow them to reveal any JavaScript's constructs they regarded as innately confusing.
+In the second part of the interview our aim was to allow the subjects to describe their JavaScript experience, as well as to allow them to reveal any JavaScript constructs they regarded as innately confusing.
 This would allow us to identify potential atom candidates that are more specific to the JavaScript language.
@@
 The participants were told both sides of each pair had the same output, and were asked to evaluate which version of the code was easier to understand.
-The interviewers restrained themselves from introducing bias into the answers by not explaining that one of the sides of each pair contained the atom under investigation.
+The interviewers avoided introducing bias in the answers by not explaining that one of the sides of each pair contained the atom under investigation.
 Subjects were just presented the snippets and allowed to take the necessary time to decide on the most readable snippet.
@@
 We first transcribed each interview and then examined the broad distribution of the answers.
-Our goal was to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion.
-We next followed with an open-coding	procedure, highlighting	the main themes and quoting the participant answers, which we show the most important ones in
-Section~\ref{sec:interview-results}. Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.
+Our goal was to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular.
+We next followed with an open-coding procedure, highlighting the main themes and quoting the answers of the participants.
+We present these results in Section~\ref{sec:interview-results}.
+Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.
 
@@
 
-\subsection{Mining JavaScript Software Repositories Settings}
+\subsection{Mining JavaScript Software Repositories Setting}
 
@@
 We measured popularity using the project's stargazers.
-This metric, available on  GitHub API, represents the number of stars a project received from users of the platform.
+This metric, available through the GitHub API, represents the number of stars a project received from users of the platform.
 The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}.
@@
 
-We mined the occurrence of atom candidates from the repositories in our curate dataset using source code queries that we had written using the CodeQL language~\cite{moor:gttse2007}.
+We mined the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}.
 CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript).
-We also automate the process of running the queries and exporting the results to a format that simplifies our
-analysis (and also the reproduction of this study). Finally, we carried
-out basic descriptive statistics, in order to understand how often the
+We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study).
+Finally, we carried out basic descriptive statistics in order to understand how often the
 atoms of confusion appear in practice.

### 2021-04-19 4_Result.tex
revisão do Prof. Fernando Castor

@@
 All participants had taken at least some university course or hold a bachelor
-degree or equivalent, meaning the participants have had some
-level of formal education on programming.
-Twenty four participants hold either a master degree (21 participants) or a doctorate degree
-(three participants).
+degree or equivalent, meaning the participants have had some level of formal education in programming.
+Twenty four participants hold either a master's degree (21 participants) or a doctorate degree (three participants).
+Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.
+Accordingly, we characterize the effect of atoms of confusion in JavaScript
+code taking into account the perceptions of both novice and experienced developers.
+
+%Our audience comprises professional developers, and 93\% of the respondents have more than one year of experience with programming.
+% (see Figure~\ref{fig:xp}). 
+%Actually, 56\% of the participants have either between four and ten years of experience (37\%) or more than ten years of experience (19\%).
+Other 37\% of the respondents have between one and four years of experience.
 
@@
 
-Our audience comprises professional developers,
-and 93\% of the respondents have more than one year of experience with programming.
-% (see Figure~\ref{fig:xp}). 
-Actually, 56\% of the participants have either between four and ten years of experience (37\%)
-or more than ten years of experience (19\%). Other 37\% of the respondents have between one and four years of experience.
-Accordingly, we characterize the effect of atoms of confusion in JavaScript
-code taking into account the perceptions of both novice and experience developers.
-
 \subsection{Misunderstanding Analysis}
@@
 
-Each participant in this study evaluated 10 code snippets, from which five were in their confusing versions, whilst the other five contained cleaned versions of the code snippets (i.e., without the confusing constructs and idioms).
+As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten.
+%code snippets, from which five were in their confusing versions, whilst the other five contained cleaned versions of the code snippets (i.e., without the confusing constructs and idioms). 
 The participants should provide the expected outcomes of the code snippets.
@@
 
-Considering Table~\ref{tab:difference-correctness}, the version of six code snippets without atoms of confusion present at least a 15\% improvement in answer correctness.
+Considering Table~\ref{tab:difference-correctness}, the versions of six code snippets without atoms of confusion present at least a 15\% improvement in answer correctness.
 In particular, the presence of the \emph{Comma Operator} atom presents the highest impact on
@@
 frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also introduce high degrees of confusion.
-The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a nonnegligible decrease in the average number of incorrect answers when observing the cleaned version of the code snippets.
+The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the cleaned version of the code snippets.
 Also, the sample of responses with no atoms had almost no dispersion, which supports the
@@
  % & \multicolumn{2}{c}{Total Number of Correct Answers} & \\ [0.1cm]
-  Atom & Confuse Code & Clean Code & $\Delta$(\%) & Effect Size (Odds Ratio) \\ \midrule
+  Atom & Confusing Code & Clean Code & $\Delta$(\%) & Effect Size (Odds Ratio) \\ \midrule
   Comma Operator &  28 &  65 & +132 & 19.02 \\ 
@@
 in the frequency of correct and incorrect answers---due to the versions
-of the code snippets (confuse and clean code).
-The result indicates that the confusing version of the code snippets impacts code understanding (p-value $< 0.05$).
+of the code snippets (confusing and clean code).
+The result indicates that the confusing versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$).
 We also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant
-\emph{education} or participant \emph{experience} impact on correctness.
-Although the results suggest that experience impact correctness, the education level does not.
-As such, we replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistic difference is less significant for those developers with more than ten years of experience.
+\emph{education} or participant \emph{experience} impact correctness.
+Although the results suggest that experience impacts correctness, the education level does not.
+As such, we replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.
 Finally, we measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR)
@@
 \num{19.02}, meaning that the the odds of a correct answer are \num{19.02} times
-higher when interpreting the clean version of a code---in comparison with the corresponding, confuse version of the code snippet.
+higher when interpreting the clean version of a code---in comparison with the corresponding, confusing version of the code snippet.
 
@@
 \subsubsection*{Exploratory Data Analysis}
-Table \ref{tab:difference-time-taken} shows the average time necessary for the participants give a correct answer about the expected outcomes of a code snippet, considering both confuse and clean versions.
+Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both confusing and clean versions.
 Seven atom candidates require an increasing cognitive workload.
-For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer---when considering the clean version of a code snippet.
+For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet.
 In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
 time on average to find the correct answer for the clean version of the code snippets. 
-Not all atom candidates, though, take less time to predict the answer. In fact, for the Arithmetic as Logic and the Pre-Increment atoms, there was an increase
-of 29.06\% and 38.19\% in the average time taken to give a correct answer---wherein the clean code versions.
+Not all atom candidates, though, require less time for the participants to predict the answer.
+In fact, for the Arithmetic as Logic and the Pre-Increment atoms, the time to give a correct answer increased
+29.06\% and 38.19\% for the clean versions.
 
@@
 \begin{tabular}{lccc|cc} \toprule
-Atom & Confuse Code & Clean Code & $\Delta$(\%) & Mann-Whitney test & Cliff's Delta \\ \midrule 
+Atom & Confusing Code & Clean Code & $\Delta$(\%) & Mann-Whitney test & Cliff's Delta \\ \midrule 
 Comma Operator & 87.67 & 20.84 & -76.23 &  \colorbox{lightgray}{$<$ 0.01} & -0.53 \\ 
@@
 We use the non-parametric \emph{Mann-Whitney test} to
-investigates the null hypothesis that developers 
+investigate the null hypothesis that developers 
 spend the same amount of time to correctly
 predict the outcome of a code snippet, regardless
-of evaluating the confuse or the clean version
-of the code. Table~\ref{tab:difference-time-taken}
-shows the results, which suggest that we
-should refute the null hypotheses for
+of evaluating the confusing or  clean version
+of the code. The results of Table~\ref{tab:difference-time-taken}
+suggest that we should refute the null hypotheses for
 five atom candidates (Comma Operator, Logic
 as Control Flow, Implicit Predicate, Arithmetic
-as Logic, and Pre-Increment). For the first three,
-the analysis suggest that the participants
+as Logic, and Pre-Increment). This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%. 
+For the first three, the analysis suggests that the participants
 need more time to predict the outcome of
-the code snippet in the confuse code version. Interesting,
+the code snippet in the confusing code version. Interesting,
 for the atom candidates Arithmetic as Logic and
@@
 to predict the outcome of the code snippets in
-the confuse version.
+the confusing version.
 We also computed the effect size using the Cliff's Delta method.
@@
 
-The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confuse code.
+The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confusing code.
 
@@
 % Fri Nov 13 11:28:56 2020
-\begin{table}[!htb]
+%%\begin{table}[!htb]
+%%\centering
+%% \caption{Summary of atoms occurrences on our dataset}
+%% 
+%%\setlength\tabcolsep{2pt} % default value: 6pt
+%%\label{tab:occurrences-summary}
+%%\begin{tabular}{lrcc}
+%%
+%%  \hline
+%%Atom & Projects & Occurrences/KLOC & Category \\ 
+%% \hline
+%%Ternary Operator & 100\% & 10.16 &  highly used \\ 
+%%Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
+%%Post Increment & 90.28\% & 5.62 &  highly used \\ 
+%%Pre-Increment & 83.33\% & 1.04 & commonly used \\ 
+%%Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
+%%Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
+%%Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
+%%Comma Operator & 15.28\% & 0.03 & little used \\ 
+%%   \hline
+%%\end{tabular}
+%%\end{table}
+
+% latex table generated in R 4.0.3 by xtable 1.8-4 package
+% Tue Apr 13 09:34:35 2021
+\begin{table}[ht]
 \centering
 \caption{Summary of atoms occurrences on our dataset}
-
 \setlength\tabcolsep{2pt} % default value: 6pt
@@
 \begin{tabular}{lrcc}
-
   \hline
@@
   \hline
-Ternary Operator & 100\% & 10.16 &  highly used \\ 
-Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
-Post Increment & 90.28\% & 5.62 &  highly used \\ 
-Pre-Increment & 83.33\% & 1.04 & commonly used \\ 
-Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
-Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
-Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
-Comma Operator & 15.28\% & 0.03 & little used \\ 
+Implicit Predicate & 100.00 & 19.89 &  highly used \\ 
+  Ternary Operator & 100.00 & 10.16 &  highly used \\ 
+  Omitted Curly Braces & 91.67 & 6.61 &  highly used \\ 
+  Post Increment & 90.28 & 5.62 &  highly used \\ 
+  Pre Increment & 83.33 & 1.04 & commonly used \\ 
+  Assignment as Value & 81.94 & 1.02 & commonly used \\ 
+  Logic as Control Flow & 56.94 & 0.95 & commonly used \\ 
+  Arithmetic as Logic & 23.61 & 0.02 & little used \\ 
+  Comma Operator & 15.28 & 0.03 & little used \\ 
    \hline
@@
 Considering all JavaScript projects in our dataset, we found 
-a total of \num{204229} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces and the Post Increment. The remaining atoms
+a total of {\color{red}\num{363810}} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces and the Post Increment.
+The remaining atoms
 account for \num{24558} (12.03\%) of the total number of occurrences, and the Comma Operator and Arithmetic as Logic are the atoms that arise less frequently (0.001\% in total with 232 and 165 occurrences, respectively).

### 2021-04-19 main.tex
revisão do Prof. Fernando Castor

@@
 \usepackage{pifont}
+\usepackage{xurl}
 % \usepackage{appendix}
@@
 In this work, we present the results of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
-Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
+Our goal is to investigate the impact of atoms of confusion on understanding code written in JavaScript --- a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
 From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atom candidates.
 Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates
-appear in more than 60\% of the projects---including the atoms Assignment as Value and Omitted Curly Braces, which appear in 8253 and 53357 times,
-respectively.
+appear in more than 60\% of the projects---including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
 Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code.
@@
 \begin{IEEEkeywords}
-datasets, neural networks, gaze detection, text tagging
+code readability, program comprehension, program understanding, atoms of confusion
+%datasets, neural networks, gaze detection, text tagging
 \end{IEEEkeywords}

### 2021-04-20 1_Introduction.tex
additional review of the Introduction

@@
 Source code misunderstandings can occur due to diverse reasons.
-A poorly understood requirement might lead to poorly implemented code.
+For instance, a poorly understood requirement might lead to poorly implemented code.
 An intricate algorithm may require extra time to be fully comprehended.
-Nevertheless, programming languages can also be at fault.
-Specific syntax and semantics of a language can cause misunderstandings. The same applies to some code idioms and stylistic choices~\cite{DBLP:conf/msr/GopsteinZFC18}.
+Besides that, programming languages can also be at fault
+and specific syntax and semantics of a language can cause misunderstandings. The same applies to some code
+idioms and stylistic choices~\cite{DBLP:conf/msr/GopsteinZFC18}.
 
@@
 
-JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}.
+This research focus on existing JavaScript idioms that a previous 
+research work have proved to introduce misunderstanding in \clang.
+JavaScript the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}.
 At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''.\footnote{\url{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}} It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.
@@
 
-Our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer.
-To achieve that, we investigate {\color{red}XYZ} JavaScript idioms that
+In particular, our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer.
+To achieve that, we investigate JavaScript idioms that
 might negatively impact program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in
@@
 %for presentation purpose only). 
-The code on the left-hand side of the figure contains two atoms of confusion discussed in previous work focusing on the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}.
+The code on the left-hand side of the figure contains two atoms of confusion discussed in previous work focusing on the \clang
+language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}.
 On the right-hand side, we present a cleaned version of the code, that is, without atoms of confusion.
@@
 Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability.
-However, these previous studies focus exclusively on statically typed languages (e.g., \clang and \cpplang).
-To the best of our knowledge, we are the first to explore atoms of confusion in the context of a dynamically typed language (JavaScript).\castor{Review the number of atoms, X, and Y.}  JavaScript's programming culture differs from that of languages such as C and Java, e.g., because there are no types and due to the pervading use of callbacks.
+However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java).
+However, JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic
+capabilities and weaker type system.
 Thus, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa.
-\castor{The following should probably be left out.
-I don't think we investigate atoms outside of those that were already studied in C.
-Or maybe say explicitly that we are studying a group of atoms that stems from the previous study?}
-Furthermore, among the atom candidates we investigate in this work, {\color{red}X} are shared between JavaScript and \clang/\cpplang (e.g., pre/post-increment), while {\color{red}Y} others are specific to JavaScript, e.g., the atom Assignment as Value.
+Here we investigate a group of atom candidates
+that stems from the Gopstein et al. study and detail additional JavaScript constructs
+that might introduce misunderstanding---according to the opinion of JavaScript
+developers.
 
@@
 
-To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 developers from the Reddit community.\footnote{\url{https://www.reddit.com/r/javascript/comments/bw36h6/academic_research_on_confusing_code_on_javascript/}}
-Their task was to predict the output of short snippets of code, with half of them containing atoms of confusion. 
-\castor{Improvement of 15\% in terms of what.
-This is too loose.} We found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of 9 atoms).
-The presence of the Comma Operator atom, for instance, can increase the average time to predict the output by 65\%, while the behavior of snippets containing the Pre-Increment atom were predicted 44\% faster.
-The presence or absence of atoms in open-source repositories may reveal cases where developers should be more or less careful when writing code.
-To identify such cases, we mined the 100 most popular JavaScript projects from GitHub and found a total of {\color{red}X} atoms in the analyzed projects.
-To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15 experienced professional developers.
-Participants preferred code snippets \textit{without} an atom in 70\% of the cases, therefore corroborating with the survey findings.
-During the MSR and interview studies, we also explored additional atom candidates that are specific to the
-JavaScript language.
+To understand whether an atom of confusion introduces misunderstandings,
+we surveyed 140 JavaScript developers. In the survey, 
+their task was to predict the output of short code
+snippets, with half of them containing atoms of confusion. 
+\castor{Improvement of 15\% in terms of what. This is too loose.}
+       {\color{red}The number of correct answers for the
+         cleaned version of the code snippets for seven (out of nine)
+         atom candidates is at least 15\% higher, in comparison to the version
+         of the code with the atoms. The presence of an atom candidate
+       might also affects the time necessary for submitting a correct answer.}
+       For instance, the presence of the Comma Operator atom increases
+       the average time to predict the output by 65\%, though the output of snippets containing
+       the Pre-Increment atom were correctly predicted 44\% faster.
+       To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15
+       experienced professional developers. Participants preferred code snippets \textit{without} an atom in 70\% of the cases,
+       therefore corroborating with the survey findings. The participants also discussed additional 
+       idioms that might lead developers to incorrectly comprehend the JavaScript code.
+       The occurrence of atoms in open-source repositories may reveal cases where developers
+       should be more or less careful when writing code. To identify such cases, we mined popular JavaScript projects
+       from GitHub and found that four atoms are highly used in JavaScript
+       code: the Implicit Predicate atom, for instance, occurs 19.89 times per KLOC of
+       code.
 

### 2021-04-20 1_Introduction.tex
additional review of the Introduction

@@
 snippets, with half of them containing atoms of confusion. 
-\castor{Improvement of 15\% in terms of what. This is too loose.}
 {\color{red}The number of correct answers for the
-         cleaned version of the code snippets for seven (out of nine)
-         atom candidates is at least 15\% higher, in comparison to the version
-         of the code with the atoms. The presence of an atom candidate
+  \emph{cleaned version} of the code snippets for seven (out of nine)
+  atom candidates is at least 15\% higher than the corresponding
+  \emph{confuse version
+  of the code snippet (i.e., with an atom candidate). The presence of an atom candidate
   might also affects the time necessary for submitting a correct answer.}
@@
   the Pre-Increment atom were correctly predicted 44\% faster.
+
   To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15
-       experienced professional developers. Participants preferred code snippets \textit{without} an atom in 70\% of the cases,
-       therefore corroborating with the survey findings. The participants also discussed additional 
-       idioms that might lead developers to incorrectly comprehend the JavaScript code.
-       The occurrence of atoms in open-source repositories may reveal cases where developers
-       should be more or less careful when writing code. To identify such cases, we mined popular JavaScript projects
-       from GitHub and found that four atoms are highly used in JavaScript
+  experienced professional developers. Participants considered the code snippets without
+  the atom candidates easier to understand in 70\% of the cases,
+  therefore corroborating the survey findings. The participants also discussed other 
+  idioms that might lead developers to misunderstand JavaScript code.
+  The occurrence of atoms in open-source repositories might reveal cases where developers
+  should be more or less careful when writing code---and open the possibility to implement
+  new program transformations for cleaning up confuse code. To identify such cases, we mined popular JavaScript projects
+  from GitHub and found four atom candidates that are highly used in JavaScript
   code: the Implicit Predicate atom, for instance, occurs 19.89 times per KLOC of
-       code.
+  code.}
 

### 2021-04-20 1_Introduction.tex
minor

@@
   atom candidates is at least 15\% higher than the corresponding
-  \emph{confuse version
+  \emph{confuse version}
   of the code snippet (i.e., with an atom candidate). The presence of an atom candidate
-  might also affects the time necessary for submitting a correct answer.}
+  might also affects the time necessary for submitting a correct answer.
   For instance, the presence of the Comma Operator atom increases

### 2021-04-20 1_Introduction.tex
fix the introduction with correct numbers

@@
   \emph{cleaned version} of the code snippets for seven (out of nine)
-  atom candidates is at least 15\% higher than the corresponding
+  atom candidates is at least 10\% higher than the corresponding
   \emph{confuse version}
@@
   For instance, the presence of the Comma Operator atom increases
-  the average time to predict the output by 65\%, though the output of snippets containing
-  the Pre-Increment atom were correctly predicted 44\% faster.
+  the average time to predict the output by 76.23\%, though the output of snippets containing
+  the Pre-Increment atom were correctly predicted 38.19\% faster.
 

### 2021-04-21 dataset/survey-results-final.Rmd
Added a few lines to dataset/survey-results-final.Rmd in order for it to also report the 95% confidence interval.

@@
 ```{r}
-
 chiTest <- c()
@@
 oddsRatioTest <- c()
+ci25 <- c()
+ci975 <- c()
 mannWhitneyTest <- c()
 cliffDelta <- c()
-
 for(i in 1:10){
@@
   oddsRatioTest[i] <- format.pval(odds.ratio(tableCorrectness)$p)
+  ci25[i] <- odds.ratio(tableCorrectness, level=0.95)$"2.5 %"
+  ci975[i] <- odds.ratio(tableCorrectness, level=0.95)$"97.5 %"
 
@@
 }
-
-analysis_df <- data.frame(atomDescription, chiTest, oddsRatio, oddsRatioTest, mannWhitneyTest, cliffDelta)
+analysis_df <- data.frame(atomDescription, chiTest, oddsRatio, oddsRatioTest, ci25, ci975, mannWhitneyTest, cliffDelta)
 analysis_df <- analysis_df[order(-oddsRatio),]
-
-colnames(analysis_df) <- c("Atom", "ChiTest", "Odds Ratio Correctness", "p-value", "Wilcox Test (Time)", "Cliff Delta")
-
+colnames(analysis_df) <- c("Atom", "ChiTest", "Odds Ratio Correctness", "p-value", "CI 2.5%", "CI 97.5%", "Wilcox Test (Time)", "Cliff Delta")
 analysis_df
-
 xtable(analysis_df)

### 2021-04-23 4_Result.tex

@@
 
-\begin{table*}[htbp]
+% \begin{table*}[htbp]
+% \caption{Time in seconds to submit a correct answer}
+% \label{tab:difference-time-taken}
+% \centering{
+% \begin{tabular}{lccc|cc} \toprule
+% Atom & Confusing Code & Clean Code & $\Delta$(\%) & Mann-Whitney test & Cliff's Delta \\ \midrule 
+% Comma Operator & 87.67 & 20.84 & -76.23 &  \colorbox{lightgray}{$<$ 0.01} & -0.53 \\ 
+% Logic as Control Flow & 108.94 & 51.07 & -53.12 &  \colorbox{lightgray}{$<$ 0.01} & -0.39 \\ 
+% % {\color{red}Automatic Semicolon Insertion} & 46.08 & 22.04 & -52.17 & 0.09 & -0.16 \\ 
+% Omitted Curly Braces & 48.85 & 30.00 & -38.58 & 0.92 & -0.01 \\ 
+% Implicit Predicate & 36.24 & 24.01 & -33.75 & \colorbox{lightgray}{$<$ 0.01} & -0.29 \\ 
+% Post Increment & 28.70 & 25.67 & -10.56 & 0.12 & 0.15 \\ 
+% Assignment as Value & 52.47 & 48.95 & -6.71 & 0.82 & 0.02 \\ 
+% Ternary Operator & 41.80 & 39.34 & -5.90 & 0.24 & -0.12 \\ 
+% Arithmetic as Logic & 28.82 & 37.20 & +29.06 & \colorbox{lightgray}{0.01} & 0.23 \\ 
+% Pre-Increment & 30.71 & 42.45 & +38.19 & \colorbox{lightgray}{$<$ 0.01} & 0.27 \\ 
+%          \bottomrule
+%     \end{tabular}}
+% \end{table*}
+
+ % latex table generated in R 4.0.4 by xtable 1.8-4 package
+ % Thu Apr 22 07:58:21 2021
+ \begin{table*}[ht]
 \caption{Time in seconds to submit a correct answer}
 \label{tab:difference-time-taken}
-\centering{
-\begin{tabular}{lccc|cc} \toprule
-Atom & Confusing Code & Clean Code & $\Delta$(\%) & Mann-Whitney test & Cliff's Delta \\ \midrule 
-Comma Operator & 87.67 & 20.84 & -76.23 &  \colorbox{lightgray}{$<$ 0.01} & -0.53 \\ 
-Logic as Control Flow & 108.94 & 51.07 & -53.12 &  \colorbox{lightgray}{$<$ 0.01} & -0.39 \\ 
-% {\color{red}Automatic Semicolon Insertion} & 46.08 & 22.04 & -52.17 & 0.09 & -0.16 \\ 
-Omitted Curly Braces & 48.85 & 30.00 & -38.58 & 0.92 & -0.01 \\ 
-Implicit Predicate & 36.24 & 24.01 & -33.75 & \colorbox{lightgray}{$<$ 0.01} & -0.29 \\ 
-Post Increment & 28.70 & 25.67 & -10.56 & 0.12 & 0.15 \\ 
-Assignment as Value & 52.47 & 48.95 & -6.71 & 0.82 & 0.02 \\ 
-Ternary Operator & 41.80 & 39.34 & -5.90 & 0.24 & -0.12 \\ 
-Arithmetic as Logic & 28.82 & 37.20 & +29.06 & \colorbox{lightgray}{0.01} & 0.23 \\ 
-Pre-Increment & 30.71 & 42.45 & +38.19 & \colorbox{lightgray}{$<$ 0.01} & 0.27 \\ 
-         \bottomrule
-    \end{tabular}}
+
+ \centering
+ \begin{tabular}{rllrlrrlr}
+   \hline
+  & Atom & ChiTest & Odds Ratio Correctness & p-value & CI 2.5\% & CI 97.5\% & Wilcox Test (Time) & Cliff Delta \\ 
+   \hline
+ 3 & Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & 2.4351e-12 & 9.21 & 356.62 & 0.09802 & -0.16 \\ 
+   4 & Comma Operator & 1.1727e-10 & 19.02 & 1.5812e-11 & 6.60 & 68.22 & 5.8249e-08 & -0.53 \\ 
+   2 & Assignment as Value & 0.0034775 & 8.39 & 0.0024331 & 1.81 & 79.12 & 0.82681 & 0.02 \\ 
+   6 & Implict Predicate & 0.01123 & 6.95 & 0.008974 & 1.46 & 66.56 & 0.0029039 & -0.29 \\ 
+   9 & Post Increment & 0.0015279 & 4.83 & 0.0012328 & 1.73 & 15.74 & 0.12006 & 0.15 \\ 
+   5 & Ternary Operator & 0.15896 & 2.73 & 0.15712 & 0.74 & 12.57 & 0.24073 & -0.12 \\ 
+   8 & Ommitted Curly Braces & 0.050962 & 2.35 & 0.050056 & 1.00 & 5.77 & 0.92529 & -0.01 \\ 
+   7 & Logic as Control Flow & 0.292 & 1.54 & 0.29198 & 0.73 & 3.28 & 7.5862e-05 & -0.39 \\ 
+   10 & Pre Increment & 0.70147 & 1.25 & 0.70181 & 0.55 & 2.85 & 0.0062981 & 0.27 \\ 
+   1 & Arithmetic as Logic & 1 & 0.84 & 1 & 0.22 & 3.12 & 0.01723 & 0.23 \\ 
+    \hline
+ \end{tabular}
  \end{table*}

### 2021-04-25 4_Result.tex
Updates from Overleaf

@@
 \subsubsection*{Hypothesis Testing}
-We use the non-parametric \emph{Mann-Whitney test} to
+We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
 investigate the null hypothesis that developers 
@@
   \hline
-Implicit Predicate & 100.00 & 19.89 &  highly used \\ 
-  Ternary Operator & 100.00 & 10.16 &  highly used \\ 
-  Omitted Curly Braces & 91.67 & 6.61 &  highly used \\ 
-  Post Increment & 90.28 & 5.62 &  highly used \\ 
-  Pre Increment & 83.33 & 1.04 & commonly used \\ 
-  Assignment as Value & 81.94 & 1.02 & commonly used \\ 
-  Logic as Control Flow & 56.94 & 0.95 & commonly used \\ 
-  Arithmetic as Logic & 23.61 & 0.02 & little used \\ 
-  Comma Operator & 15.28 & 0.03 & little used \\ 
+Implicit Predicate & 100\% & 19.89 &  highly used \\ 
+  Ternary Operator & 100\% & 10.16 &  highly used \\ 
+  Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
+  Post Increment & 90.28\% & 5.62 &  highly used \\ 
+  Pre Increment & 83.33\% & 1.04 & commonly used \\ 
+  Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
+  Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
+  Automatic Semicolon Insertion & 33.33\% & 0.13 & commonly used \\ 
+  Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
+  Comma Operator & 15.28\% & 0.03 & little used \\ 
    \hline
@@
 Considering all JavaScript projects in our dataset, we found 
-a total of {\color{red}\num{363810}} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces and the Post Increment.
+a total of {\color{red}\num{364873}} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces and the Post Increment.
 The remaining atoms

### 2021-04-25 5_Threats.tex
Updates from Overleaf

@@
 
-%% Since our research was conducted online with unknown participants, and thus we had no way of confirming their
-%% levels of education and experience. Also,  we did not have a way to prevent respondents from cheating,
-%% such as running the code on an interpreter, or consulting other people. This presents a threat to the validity to the
-%% conclusions of our work, even though we leverage some means to try to mitigate this threat (as discussed
-%% in Section~\ref{method}).
+Conclusion validity is connected with how well it is possible to establish relationships between treatments and outcomes.
+Threats to conclusion validity often come from inappropriate use of statistics.
+In the context of our survey study, we apply different statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test).
+Furthermore, besides reporting p-values, we also report effect size measures appropriate for each scenario (odds-ratio and Cliff's delta) and apply a p-value correction technique to avoid the multiple testing problem.
+Finally, it could be argued that the size of the samples is insufficient to make conclusions for some of the atoms, a common problem in empirical software engineering.
+We estimate the sample size for each one of the atom candidates, considering that each one had a different number of samples (Table~\ref{tab:difference-correctness}).
+To that end, we use the $\phi$ measure of effect size for each atom (based on the Chi-square statistic), the standard $\alpha$ coefficient of 0.05, and set the expected statistical power to 0.8, as usually employed in the literature~\cite{Ellis:2010:EGE}.
+As a result of this analysis, we discover that for all but one of the atoms where we had a statistically significant result for correctness, Implicit Predicate, the sample size is sufficient.
+For the ones where we did not find statistically significant results for correctness, in fact the sample sizes we have analyzed are insufficient for such small effect sizes.
+This indicates that further studies on these atoms are required.
 
-%% Another threat to our results lies in the fact that some constructs
-%% might not be frequent, or present at all, in JavaScript code bases.
-%% The atom candidates Arithmetic as Logic and
-%% the Comma Operator were rarely found in the
-%% open source JavaScript code we mined in our
-%% third study.
+%One threat to validity we have not properly addressed is lack of power, a common problem in Software Engineering research.
 
-%% A potential problem with our method is that
-%% there little incentive for subjects to think
-%% thoroughly about the questions. We observed a lack of
-%% engagement when we ran the survey with undergraduate students during
-%% a pilot study. Although our final subjects were voluntarily
-%% partaking in the survey, we could not be sure that, after some
-%% time taking the survey, respondents would become tired and stop thinking
-%% clearly about the code.
+Construct validity is connected with how well the selected measures actually represent the concept of interest.
+We use correctness as a proxy for program comprehension.
+Furthermore, we use predicted outcomes of small programs as a measure of correctness.
+As discussed elsewhere~\cite{Oliveira:2020:ECR}, this approach is a test of the developers' ability to trace programs.
+Although this is a common approach in studies about atoms of confusion~\cite{TheEyesDoNotLie,Langhout:2021:ACJ,DBLP:conf/sigsoft/GopsteinIYDZYC17}, other measures of correctness could have been employed, potentially yielding different results.
+As a complement to correctness in the survey, we have measured the time it took for the participants to finish.
+Furthermore, we have  asked the interviewees about their preferences when comparing confusing and clean versions of the programs.
+
+%A potential problem with our method is that there little incentive for subjects to think thoroughly about the questions.
+We observed a lack of engagement when we ran the survey with undergraduate students during a pilot study.
+Although our final subjects were voluntarily partaking in the survey, we could not be sure that, after some time taking the survey, respondents would become tired and stop thinking clearly about the code.
+
+Internal validity is concerned with how well the study isolates the variables of interest and accounts for confounding factors.
+Since the survey was conducted online with unknown participants, we have no way of confirming their levels of education and experience.
+We mitigate this threat by, in the data analysis, explicitly accounting for programmer experience and using an experimental design that allows us to isolate the impact of the treatment from factors such as experience and formal education level.
+Also, we did not have a way to prevent respondents from cheating, such as running the code on an interpreter, or consulting other people.
+We partially mitigate this threat by presenting images with source code, instead of text.
+This creates an obstacle for participants to run the code while taking the survey.
+
+External validity is linked to whether it is possible to extrapolate the results of the study.
+Although our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, it is not clear whether that result extrapolates to other scenarios.
+Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect.
+Another possible threat to the generalizability of our results, in particular for the survey and interviews, lies in the fact that the analyzed atoms may rarely occur in real JavaScript code bases.
+To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are fairly frequent, occurring at least once per 1,000 lines of code.
+Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code.
 

### 2021-04-25 main.tex
Updates from Overleaf

@@
 \begin{abstract}
-Evolving legacy code is a challenging task, particularly when the code has been poorly written---which imposes a significant cognitive load on developers when trying to familiarize themselves with a legacy codebase.
+Evolving legacy code is a challenging task, particularly when the code has been poorly written, increasing effort and imposing a significant cognitive load on developers.
+Previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in C and C++.
+However, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript).
+In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
+From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atom candidates.
+Our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
+Altogether, our findings might help developers better understand the implications of atoms of confusion on understanding JavaScript code and contribute to the design of program transformation tools to simplify JavaScript code.
+
+% Our investigation also reveals additional JavaScript constructs that might hinder even experienced developers from understanding JavaScript code, such as the use of callbacks functions.
+Altogether...
+
+%Evolving legacy code is a challenging task, particularly when the code has been poorly written---which imposes a significant cognitive load on developers when trying to familiarize themselves with a legacy codebase.
 Previous research has investigated the impact of small code patterns that contribute to misunderstanding the source code written in C and C++.
@@
 From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atom candidates.
-Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates
-appear in more than 60\% of the projects---including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
+Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects---including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
 Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code.

### 2021-04-25 main.tex
additional review in the abstract

@@
 
+\newcommand{\na}{9\xspace}
+
 \begin{document}
@@
   Evolving legacy code is a challenging task, particularly when the code has been poorly written, increasing effort and imposing a significant cognitive load on developers.
-Previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in C and C++.
-However, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript).
+Although previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in \clang, \cpplang, and Java, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript).
 In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
-From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atom candidates.
-Our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
-Altogether, our findings might help developers better understand the implications of atoms of confusion on understanding JavaScript code and contribute to the design of program transformation tools to simplify JavaScript code.
+From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in {\color{red}6} out of \na atom candidates.
+In addition, our MSR effort reveals that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
+Altogether, our findings might help practitioners (a) better understand the implications of atoms of confusion on understanding JavaScript code, (b) avoid writing code that is unnecessarily difficult to maintain, and (c) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
 
@@
 \begin{IEEEkeywords}
-code readability, program comprehension, program understanding, atoms of confusion
+code readability, program comprehension, program understanding, atoms of confusion, JavaScript code
 %datasets, neural networks, gaze detection, text tagging

### 2021-04-25 5_Threats.tex
minor review

@@
 Although this is a common approach in studies about atoms of confusion~\cite{TheEyesDoNotLie,Langhout:2021:ACJ,DBLP:conf/sigsoft/GopsteinIYDZYC17}, other measures of correctness could have been employed, potentially yielding different results.
-As a complement to correctness in the survey, we have measured the time it took for the participants to finish.
+As a complement to correctness in the survey, we have measured the time it took for the participants to correctly predict the outcome of the code snippets (either with or without atom candidates).
 Furthermore, we have  asked the interviewees about their preferences when comparing confusing and clean versions of the programs.

### 2021-04-25 4_Result.tex
review of the survey results

@@
 
-\subsection{Misunderstanding Analysis}
+\subsection{Misunderstanding Rate Analysis}
 
@@
 
-As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten.
-%code snippets, from which five were in their confusing versions, whilst the other five contained cleaned versions of the code snippets (i.e., without the confusing constructs and idioms). 
+As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten
+code snippets, from which five were in their confusing versions, whilst the other five contained cleaned versions of the
+code snippets (i.e., without the confusing constructs and idioms). 
 The participants should provide the expected outcomes of the code snippets.
@@
 
-\begin{table*}[htbp]
+\begin{table}[htbp]
 \caption{Total number of correct answers.}
@@
 \centering{
-\begin{tabular}{lccc|c} \toprule
- % & \multicolumn{2}{c}{Total Number of Correct Answers} & \\ [0.1cm]
-  Atom & Confusing Code & Clean Code & $\Delta$(\%) & Effect Size (Odds Ratio) \\ \midrule
-  Comma Operator &  28 &  65 & +132 & 19.02 \\ 
-%   {\color{red}Automatic Semicolon Insertion} &  32 &  68 & +112 & 39.32 \\ 
-  Post Increment &  48 &  64 & +33 & 4.83 \\ 
-  Omitted Curly Braces &  47 &  58 &  +23 & 2.35 \\
-  Assignment as Value &  56 &  68 &  +21 & 8.38 \\ 
-  Implicit Predicate &  58 &  68 &  +17 & 6.94 \\ 
-  Logic as Control Flow &  41 &  48 &  +17 & 1.53 \\ 
-  Ternary Operator &  60 &  66 &  +10 & 2.73 \\ 
-  Pre-Increment &  50 &  53 &  +6 & 1.24 \\ 
-  Arithmetic as Logic &  64 &  63 &  -2 & 0.84 \\ \bottomrule 
+\begin{scriptsize}
+\begin{tabular}{lccc} \toprule
+  Atom & Confusing Code & Clean Code & $\Delta$(\%)  \\ \midrule
+  Comma Operator                &  28 &  65 & +132   \\ 
+  Automatic Semicolon Insertion &  32 &  68 & +112   \\ 
+  Post-Increment                &  48 &  64 & +33    \\ 
+  Omitted Curly Braces          &  47 &  58 &  +23   \\
+  Assignment as Value           &  56 &  68 &  +21   \\ 
+  Implicit Predicate            &  58 &  68 &  +17   \\ 
+  Logic as Control Flow         &  41 &  48 &  +17   \\ 
+  Ternary Operator              &  60 &  66 &  +10   \\ 
+  Pre-Increment                 &  50 &  53 &  +6    \\ 
+  Arithmetic as Logic           &  64 &  63 &  -2    \\ \bottomrule
+
 \end{tabular}
+\end{scriptsize}
 }
-\end{table*}
+\end{table}
 
@@
 Finally, we measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR)
-method (see the Effect Size column of Table~\ref{tab:difference-correctness}). We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and
+method (see the Odds Ratio column of Table~\ref{tab:hypothesis-testing}). We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and
 Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. For instance, the Comma Operator leads to an \emph{Odds Ratio} of
@@
 
-% \begin{table*}[htbp]
-% \caption{Time in seconds to submit a correct answer}
-% \label{tab:difference-time-taken}
-% \centering{
-% \begin{tabular}{lccc|cc} \toprule
-% Atom & Confusing Code & Clean Code & $\Delta$(\%) & Mann-Whitney test & Cliff's Delta \\ \midrule 
-% Comma Operator & 87.67 & 20.84 & -76.23 &  \colorbox{lightgray}{$<$ 0.01} & -0.53 \\ 
-% Logic as Control Flow & 108.94 & 51.07 & -53.12 &  \colorbox{lightgray}{$<$ 0.01} & -0.39 \\ 
-% % {\color{red}Automatic Semicolon Insertion} & 46.08 & 22.04 & -52.17 & 0.09 & -0.16 \\ 
-% Omitted Curly Braces & 48.85 & 30.00 & -38.58 & 0.92 & -0.01 \\ 
-% Implicit Predicate & 36.24 & 24.01 & -33.75 & \colorbox{lightgray}{$<$ 0.01} & -0.29 \\ 
-% Post Increment & 28.70 & 25.67 & -10.56 & 0.12 & 0.15 \\ 
-% Assignment as Value & 52.47 & 48.95 & -6.71 & 0.82 & 0.02 \\ 
-% Ternary Operator & 41.80 & 39.34 & -5.90 & 0.24 & -0.12 \\ 
-% Arithmetic as Logic & 28.82 & 37.20 & +29.06 & \colorbox{lightgray}{0.01} & 0.23 \\ 
-% Pre-Increment & 30.71 & 42.45 & +38.19 & \colorbox{lightgray}{$<$ 0.01} & 0.27 \\ 
-%          \bottomrule
-%     \end{tabular}}
-% \end{table*}
+\begin{table}[htbp]
+\caption{Time in seconds to submit a correct answer}
+\label{tab:difference-time-taken}
+\centering{
+\begin{scriptsize}
+\begin{tabular}{lccc} \toprule
+Atom & Confusing Code & Clean Code & $\Delta$(\%) \\ \midrule
+Comma Operator & 87.67 & 20.84 & -76.23 \\ 
+Automatic Semicolon Insertion & 46.08 & 22.04 & -52.17 \\ 
+Post-Increment & 28.70 & 25.67 & -10.56 \\ 
+Ommitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
+Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
+Implict Predicate & 36.24 & 24.01 & -33.75 \\ 
+Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
+Ternary Operator & 41.80 & 39.34 & -5.90 \\ 
+Pre-Increment & 30.71 & 42.45 & 38.19 \\  
+Arithmetic as Logic & 28.82 & 37.20 & 29.06 \\ \bottomrule
+\end{tabular}
+\end{scriptsize}
+}
+\end{table}
 
@@
  \begin{table*}[ht]
-\caption{Time in seconds to submit a correct answer}
-\label{tab:difference-time-taken}
+\caption{Hipothesis Testing (misunderstanding and cognitive workload)}
+\label{tab:hypothesis-testing}
 
  \centering
- \begin{tabular}{rllrlrrlr}
-   \hline
-  & Atom & ChiTest & Odds Ratio Correctness & p-value & CI 2.5\% & CI 97.5\% & Wilcox Test (Time) & Cliff Delta \\ 
-   \hline
- 3 & Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & 2.4351e-12 & 9.21 & 356.62 & 0.09802 & -0.16 \\ 
-   4 & Comma Operator & 1.1727e-10 & 19.02 & 1.5812e-11 & 6.60 & 68.22 & 5.8249e-08 & -0.53 \\ 
-   2 & Assignment as Value & 0.0034775 & 8.39 & 0.0024331 & 1.81 & 79.12 & 0.82681 & 0.02 \\ 
-   6 & Implict Predicate & 0.01123 & 6.95 & 0.008974 & 1.46 & 66.56 & 0.0029039 & -0.29 \\ 
-   9 & Post Increment & 0.0015279 & 4.83 & 0.0012328 & 1.73 & 15.74 & 0.12006 & 0.15 \\ 
-   5 & Ternary Operator & 0.15896 & 2.73 & 0.15712 & 0.74 & 12.57 & 0.24073 & -0.12 \\ 
-   8 & Ommitted Curly Braces & 0.050962 & 2.35 & 0.050056 & 1.00 & 5.77 & 0.92529 & -0.01 \\ 
-   7 & Logic as Control Flow & 0.292 & 1.54 & 0.29198 & 0.73 & 3.28 & 7.5862e-05 & -0.39 \\ 
-   10 & Pre Increment & 0.70147 & 1.25 & 0.70181 & 0.55 & 2.85 & 0.0062981 & 0.27 \\ 
-   1 & Arithmetic as Logic & 1 & 0.84 & 1 & 0.22 & 3.12 & 0.01723 & 0.23 \\ 
-    \hline
+ \begin{tabular}{lcccc|cc}
+   \toprule
+Atom & ChiTest & Odds Ratio          & CI 2.5\% & CI 97.5\% & Wilcox Test & Cliff Delta \\ 
+     &         & (misunderstanding)  &          &           & (workload)             & \\ \midrule
+Comma Operator & 1.1727e-10 & 19.02 & 6.60 & 68.22 & 5.8249e-08 & -0.53 \\ 
+Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & 9.21 & 356.62 & 0.09802 & -0.16 \\ 
+Post-Increment & 0.0015279 & 4.83 & 1.73 & 15.74 & 0.12006 & 0.15 \\ 
+Ommitted Curly Braces & 0.050962 & 2.35 & 1.00 & 5.77 & 0.92529 & -0.01 \\ 
+Assignment as Value & 0.0034775 & 8.39 & 1.81 & 79.12 & 0.82681 & 0.02 \\ 
+Implict Predicate & 0.01123 & 6.95 & 1.46 & 66.56 & 0.0029039 & -0.29 \\ 
+Logic as Control Flow & 0.292 & 1.54 & 0.73 & 3.28 & 7.5862e-05 & -0.39 \\ 
+Ternary Operator & 0.15896 & 2.73 & 0.74 & 12.57 & 0.24073 & -0.12 \\ 
+Pre-Increment & 0.70147 & 1.25 & 0.55 & 2.85 & 0.0062981 & 0.27 \\ 
+Arithmetic as Logic & 1 & 0.84 & 0.22 & 3.12 & 0.01723 & 0.23 \\ 
+    \bottomrule
  \end{tabular}
@@
 suggest that we should refute the null hypotheses for
-five atom candidates (Comma Operator, Logic
-as Control Flow, Implicit Predicate, Arithmetic
-as Logic, and Pre-Increment). This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%. 
-For the first three, the analysis suggests that the participants
+six atom candidates (Comma Operator, Automatic Semicolong Insertion,
+Implicit Predicate, Logic
+as Control Flow, Pre-Increment, and Arithmetic
+as Logic). This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%. 
+For the first four, the analysis suggests that the participants
 need more time to predict the outcome of
@@
   correct answer in the presence of the
-  atom candidates Comma Operator, Logic as
+  atom candidates Comma Operator, Automatic Semicolon Insertion, Logic as
   Control Flow, and Implicit Predicate. Conversely,

### 2021-04-25 3_Method.tex
reduced the size of the tables

@@
 
-\begin{table*}[htb!]
+\begin{table}[htb!]
   \centering
-\begin{tabular}
-{p{0.4cm}p{0.9cm}p{2.0cm}p{6cm}p{1.9cm} p{4.5cm}}
+\begin{scriptsize}  
+\begin{tabular}{clcl}
 \toprule
-ID & Gender & Level & Role & JS Experience & Other Languages \\ \midrule 
-P1 & Male & Graduate & Software Developer & 9 years & Java, PHP, C, Go  
+ID & Education & JS Experience & Other Languages \\ \midrule 
+P1 & BSc Degree & 9 years & Java, PHP, C, Go  
 \\ 
-P2 & Male & High School & Software Developer & 3 years & Python, Go, Dart, Lua, C++, C\#
+P2 & HS Degree & 3 years & Python, Go, Dart, Lua, C++, C\#
 \\ 
-P3 & Male & Graduate & Software Developer & 4 years & Java, C
+P3 & BSc Degree & 4 years & Java, C
 \\ 
-P4 & Male & Undergraduate & Software Developer & 3 years & Python, C, C++, Java, Go
+P4 & Undergraduate & 3 years & Python, C, C++, Java, Go
 \\ 
-P5 & Male & Master Student & Software Developer & 3 years & Python, SQL
+P5 & Master Student & 3 years & Python, SQL
 \\ 
-P6 & Male & Graduate & Software Architect  & 15  years  & Java, PHP, C, Python , Ruby, C\# and R
+P6 & Bsc Degree & 15  years  & Java, PHP, C, Python , Ruby, C\# and R
 \\ 
-P7 & Male & Graduate & Software Developer  & 6 years & C, C++, Java, Assembly, Kotlin
+P7 & BSc Degree  & 6 years & C, C++, Java, Assembly, Kotlin
 \\ 
-P8 & Male & PhD & Software Developer & 2 years & Java, Python
+P8 & PhD Degree & 2 years & Java, Python
 \\ 
-P9 & Male & Graduate & Business Analyst and Software Developer & 5 years & PHP
+P9 & BSc Degree  & 5 years & PHP
 \\ 
-P10 & Male & Master Student & Business Analyst and Software Developer  & 4 years & Java, C\# and Python
+P10 & Master Student & 4 years & Java, C\# and Python
 \\ 
-P11 & Male & Master Student & Software Architect & 4 years & Java, Erlang, C\#, Cobol
+P11 & Master Student & 4 years & Java, Erlang, C\#, Cobol
 \\ 
-P12 & Female & Master & Software Developer  & 1 year & C
+P12 & Master Degree & 1 year & C
 \\ 
-P13 & Male & Master & Software Developer & 13 years & Java, PHP
+P13 & Master Degree & 13 years & Java, PHP
 \\ 
-P14 & Female & Master Student & Software Developer  & 3 years & C, Python, Ruby
+P14 & Master Student & 3 years & C, Python, Ruby
 \\ 
-P15 & Female &  Graduate & Software Developer & 2 years & Java, PHP
+P15 &  BSc degree  & 2 years & Java, PHP
 \\ \bottomrule
 \end{tabular}
+\end{scriptsize}
     \caption{Demographic information of the participants}
     \label{pinterview}
-\end{table*}
+\end{table}
 
@@
 
-\begin{table*}[ht]
+\begin{table}[ht]
   \centering
- \begin{tabular}{rrrrrrr}
+  \begin{scriptsize}
+ \begin{tabular}{rrrrr}
    \hline
-             & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ \hline
- Lines of Code           & \num{5543}  & \num{16060.25} & \num{36161.5} & \num{111432.33} & \num{111182.50} & \num{1278405} \\
- Num. of Forks     & \num{1244}   & \num{3029.75} & \num{6078} & \num{8906.24} & \num{9260.50} & \num{68849} \\
- Num. of Contributors  & \num{6}   & \num{184} & \num{285.5} & \num{533.44} & \num{515.75} & \num{4047} \\
- Num. of Stars        & \num{23672} & \num{27055} & \num{34990} & \num{46919.51} & \num{48518} & \num{310935} \\
+                       & Min.             & Median        & Mean             & Max. \\ \hline
+ Lines of Code         & \num{5543}       & \num{36161.5} & \num{111432.33}  & \num{1278405} \\
+ Num. of Forks         & \num{1244}       & \num{6078}    & \num{8906.24}    & \num{68849} \\
+ Num. of Contributors  & \num{6}          & \num{285.5}   & \num{533.44}     & \num{4047} \\
+ Num. of Stars         & \num{23672}      & \num{34990}   & \num{46919.51}   & \num{310935} \\
 
@@
  \end{tabular}
+ \end{scriptsize}
  \caption{Some descriptive statistics about the projects used in the MSR study}
  \label{tab:projects-statistics} 
- \end{table*}
+ \end{table}

### 2021-04-25 4_Result.tex
additional review in the survey / MSR results

@@
 experience, familiarity with JavaScript, and their opinion
-about the nine atom candidates---by
+about the \na atom candidates---by
 (a) presenting them pairs of code snippets whose behavior were identical
@@
 instead of the more cleaner version without the atom of confusion.
-Even in this case, some participants who opted for
-the \emph{left-hand-side} (with the atom of confusion) version
-still believed that the \emph{right-hand-side} (without the atom of
-confusion) version was more readable (see Figure~\ref{code:ternary}). 
+We show the code snippets for the Ternary Operator in Figure~\ref{code:ternary}.
+Even for this atom candidate, some participants who preffered 
+the \emph{left-hand-side} version (with the atom of confusion) 
+still believed that the \emph{right-hand-side} version (without the atom of
+confusion) was more readable.  
 The following quotes were extracted from the transcripts with
@@
 
-\begin{mq}
-\emph{``I think [the \lhs version] is easier to understand, but [the \rhs version] is what I would write"}
-\end{mq}
+%% \begin{mq}
+%% \emph{``I think [the \lhs version] is easier to understand, but [the \rhs version] is what I would write"}
+%% \end{mq}
 
-The atom candidate Ternary Operator also opens up
-the possibility for a derivative construct that JavaScript
-allows which can be rather confusing, and that is the nested
-ternaries construct, in which the right-hand side of the a
-ternary operator can be another ternary construct.
-While nested ternaries remove the number of lines that
-would be necessary to construct using nested if-then-else statements,
-they can become quite taxing to understand.
-Nested ternaries are a choice of atoms to be analyzed in future work.
+%% The atom candidate Ternary Operator also opens up
+%% the possibility for a derivative construct that JavaScript
+%% allows which can be rather confusing, and that is the nested
+%% ternaries construct, in which the right-hand side of the a
+%% ternary operator can be another ternary construct.
+%% While nested ternaries remove the number of lines that
+%% would be necessary to construct using nested if-then-else statements,
+%% they can become quite taxing to understand.
+%% Nested ternaries are a choice of atoms to be analyzed in future work.
 
@@
   and my colleagues complained that it was difficult to understand.
-  So nowadays I prefer to write code using the [\rhs version]."}
+  Nowadays I prefer to write code using the [\rhs version]."}
 \end{mq}
@@
 the \emph{\lhs} (with the atom of confusion) caused significant
-confusion among the participants. Two remarks about
-the \emph{Comma Operator} atom of confusion are listed below:
+confusion among the participants. One remark about
+the \emph{Comma Operator} atom of confusion is listed below:
 
-\begin{mq}
-\emph{``I just learned that [the \lhs version of this code] is possible. I did not even know it worked''}
-\end{mq}
+%% \begin{mq}
+%% \emph{``I just learned that [the \lhs version of this code] is possible. I did not even know it worked''}
+%% \end{mq}
 
@@
 
-This is an interesting perspective, because such small, yet tricky program constructs and idioms, might hinder novice developers in the task of maintaining code; and one cannot make any assumption about the programmers' experience of who is going to maintain a code in a long run.
+This is an interesting perspective, because such a perhaps simple decision,  might hinder novice developers in the task of maintaining code; and one cannot make any assumption about the programmers' experience of who is going to maintain a code in a long run.
 
@@
     \centering
-    \caption{Interviews Round 1 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}).
+    \caption{Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}).
 Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
@@
 For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
-For the multi-purpose language that is one of the most, if not the most used one by practitioners,
-a lack of understanding of the languages core
-features may be a problem worth investigating in future studies.
+%% For the multi-purpose language that is one of the most, if not the most used one by practitioners,
+%% a lack of understanding of the languages core
+%% features may be a problem worth investigating in future studies.
 
@@
 
-We mined \minedprojects open source JavaScript repositories to understand how often atoms of confusion arise in real software. We discarded in this analysis the atom Implicit Predicate,
-because writing a CodeQL query for mining this particular atom is everything but trival.
-As such, in this study
-we only mined the eight atoms listed in Table~\ref{tab:occurrences-summary}.
-
-Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18}~\cite{DBLP:journals/ese/MedeirosLAAKRG19}, which investigates the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion frequently arise in JavaScript open source systems.
-For instance, the 5 most frequently found atoms occur in at least 80\% of the projects.
-Considering the extremes, atom Ternary Operator was found in all repositories,
+We mined \minedprojects open source JavaScript repositories to understand how
+often atoms of confusion arise in real software. 
+Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19}, which investigte
+the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
+frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
+occur in at least 80\% of the projects. Considering the extremes, atom Ternary Operator was found in all repositories we mined,
 while atom Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
@@
   \hline
-Atom & Projects & Occurrences/KLOC & Category \\ 
+Atom & Projects & Occurr./KLOC & Category \\ 
   \hline
@@
 Considering all JavaScript projects in our dataset, we found 
-a total of {\color{red}\num{364873}} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces and the Post Increment.
+a total of {\color{red}\num{364873}} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces, and the Post Increment.
 The remaining atoms
@@
 atoms as highly used (Ternary Operator, Omitted Curly Braces, and Post Increment),
-three as commonly used (Pre-Increment, Assignment as Value, and Logic as Control Flow)
+four as commonly used (Pre-Increment, Assignment as Value,
+Logic as Control Flow, and Automatic Semicolon Insertion),
 and two atoms as little used (Arithmetic as Logic and Comma Operator).
@@
 
-The results of our survey and interviews suggest that the Ternary Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to our survey).
-Nonetheless, the Post Increment Expression and Omitted Curly Braces atoms are listed in the top three sources of misunderstanding
-(see Table~\ref{tab:difference-correctness}). As such, fixing the the Post Increment and Omitted Curly Braces might solve most of the unclear code caused by atoms of confusion
-in JavaScript projects.
+The results of our survey and interviews suggest that the
+Ternary Operator does not contribute significantly as a source of
+misunderstanding (increasing the number of wrong answers in 9\% of the cases,
+according to our survey). Nonetheless, the Post-Increment Expression
+and Omitted Curly Braces atoms are listed in the top four sources of misunderstanding 
+(see Table~\ref{tab:difference-correctness}). As such, fixing the
+the Post-Increment and Omitted Curly Braces might solve most
+of the unclear code caused by atoms of confusion in
+JavaScript projects.

### 2021-04-26 1_Introduction.tex
Updates from Overleaf

--- 2021-04-20 1_Introduction.tex
@@
 
-This research focus on existing JavaScript idioms that a previous 
-research work have proved to introduce misunderstanding in \clang.
-JavaScript the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}.
+%This research focuses on existing JavaScript idioms that a previous research work have proved to introduce misunderstanding in \clang.
+JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}.
 At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''.\footnote{\url{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}} It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.
@@
 
-In particular, our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer.
-To achieve that, we investigate JavaScript idioms that
-might negatively impact program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in
-JavaScript---that is, small
-pieces of code that hinder developers' understanding of the source code of a program
-for which there are functionally equivalent alternatives that are easier
-to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For instance, consider
-the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
+%In particular, 
+Our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer.
+To achieve that, we investigate JavaScript idioms that might negatively impact program understanding.
+These idioms correspond to a set of candidates for \emph{atoms of confusion} in JavaScript---that is, small
+pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier
+to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For instance, consider the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
 %for presentation purpose only). 
@@
 language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}.
-On the right-hand side, we present a cleaned version of the code, that is, without atoms of confusion.
+On the right-hand side, we present a clean version of the code, that is, without atoms of confusion.
 In the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in  41\% and 34\%, respectively.
@@
 Thus, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa.
-Here we investigate a group of atom candidates
-that stems from the Gopstein et al. study and detail additional JavaScript constructs
-that might introduce misunderstanding---according to the opinion of JavaScript
+Here we investigate a group of atom candidates that stems from the study of Gopstein and colleagues and detail additional JavaScript constructs that might introduce misunderstanding---according to the opinions of JavaScript
 developers.
@@
 
-To understand whether an atom of confusion introduces misunderstandings,
-we surveyed 140 JavaScript developers. In the survey, 
-their task was to predict the output of short code
-snippets, with half of them containing atoms of confusion. 
+To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 JavaScript developers.
+In the survey,  their task was to predict the output of short code snippets, with half of them containing atoms of confusion.
 {\color{red}The number of correct answers for the
-  \emph{cleaned version} of the code snippets for seven (out of nine)
+  \emph{clean version} of the code snippets for 7 out of \na
   atom candidates is at least 15\% higher than the corresponding
-  \emph{confuse version}
+  \emph{confusing version}
   of the code snippet (i.e., with an atom candidate). The presence of an atom candidate
-  might also affects the time necessary for submitting a correct answer.
+  might also affect the time necessary for submitting a correct answer.
   For instance, the presence of the Comma Operator atom increases
-  the average time to predict the output by 65\%, though the output of snippets containing
-  the Pre-Increment atom were correctly predicted 44\% faster.
+  the average time to predict the output by 76.23\%, though the outputs of snippets containing
+  the Pre-Increment atom were correctly predicted 38.19\% faster.
 
@@
   should be more or less careful when writing code---and open the possibility to implement
-  new program transformations for cleaning up confuse code. To identify such cases, we mined popular JavaScript projects
+  new program transformations for cleaning up confusing code. To identify such cases, we mined popular JavaScript projects
   from GitHub and found four atom candidates that are highly used in JavaScript
-  code: the Implicit Predicate atom, for instance, occurs 19.89 times per KLOC of
-  code.}
+  code. The Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.}
 

### 2021-04-26 2_Background.tex
Updates from Overleaf

@@
 As program comprehension is regarded less as a systematic process than as an objective \cite{DBLP:journals/ibmsj/OHareT94}, there is no single approach or framework that is capable of yielding easily understandable programs.
-Factors such as problem domain, size and complexity of the code base, and available tools all vary significantly between different working settings.
+Factors such as problem domain, size and complexity of the code base, and available tools all vary significantly between different settings.
 These aspects, combined with differences between individuals, which can include experience, working memory capacity, and other cognitive traits, make it clear that program comprehension should be approached in different ways in different contexts.
@@
 Their work laid an empirical framework to identify such atoms.
-In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atoms were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.
+In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atom candidates were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.
 %By having subjects evaluate code where in the expected output relied exclusively on evaluating the result of previously known confusing programs that had been written in \clang or \cpplang, and asking the participants to predict the output of a set of code snippets where the atoms had been removed, the authors were able to isolate and measure the impact of such confusing atoms.
 
-One main example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
+One example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
 often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}.
-This occurs because a leading \texttt{0} in a number literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers.
+This occurs because a leading \texttt{0} in a numerical literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers.
 After formulating a list of 19 atom candidates, Gopstein et al.
@@
 They found that code with atoms of confusion requires more time from participants to predict the output and more visual effort to comprehend.
-Gopstein et al., in another paper~\cite{ThinkingAloud} further explored the original atoms by additionally conducting interviews and having programmers discuss between themselves the studied atoms.
+Gopstein et al., in another paper~\cite{ThinkingAloud} further explored the original atoms by additionally conducting interviews and having programmers discuss among themselves the studied atoms.
 The authors argue that this method is important to complement a quantitative analysis.
 Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''
-}. \castor{We use interviews, but did we "test" the interviewees to verify if they understand code?} We follow a similar approach as we augment our quantitative analysis with interviews.
+}. We follow a similar approach as we augment our quantitative analysis with interviews.
 

### 2021-04-26 3_Method.tex
Updates from Overleaf

@@
 In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice.
-Answering these research questions also lay the foundations for the implementation of tools that can automatically transform code into its cleaner versions---though we postpone these results to a future research work.
+Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions---though we postpone these results to future research work.
 To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and a JavaScript software repository mining study.
@@
 programmers to take longer to predict their output.
-Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} whose idioms also exist in JavaScript programs.
+Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs.
 Our supplementary material
@@
 considers two treatments: the presence or absence of 
-atoms of confusion within the code snippets. 
+atoms of candidates within the code snippets. 
 To achieve such a design goal, controlling the effect of experience and individual code snippets, we resorted to the \textit{Latin Square Design} \cite{Hunter-Experimenters}.
@@
 Furthermore, a given subject (P2), which constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing snippets for $s_6, s_7,..., s_{10}$.
-By doing that, we guarantee that all versions of code snippets are contained within each square, and that each configuration occurs only once within a square.
+By doing that, we guarantee that all versions of the code snippets are contained within each square, and that each configuration occurs only once within a square.
 Figure \ref{fig:latinsquare} offers a visual representation of the concept.
@@
 The order in which the questions were presented was randomized.
-By doing this, we were seeking to minimize the chances of subjects being aware that the current listing they were analyzing contained (or not) atoms of confusion.
-That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets did not).
+By doing this, we seek to minimize the chances of subjects being aware that the current listing they are analyzing contains (or not) atoms of confusion.
+That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets do not).
 We measured answer correctness and the total time each participant needed to answer the survey's questions.
@@
 As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
-Fellow undergraduate students, professional colleagues, and friends took the pilot survey.
+Undergraduate students and professional colleagues took the pilot survey.
 Some users reported layout defects, and many reported that the landing page did not explain the survey well enough.
@@
 We decided not to provide feedback about the time students took to answer each question.
-Nor did we tell them whether their answers were correct or not.
+We also did not tell them whether their answers were correct or not.
 Our main concern was to avoid introducing bias for future respondents.
@@
 We considered all squares that contained incomplete rows to be invalid during data analysis and discarded them.
-Since we had\
- a large enough number of samples, the squares we had to discard did not impact our analysis.
+%Since we had a large enough number of samples, the squares we had to discard did not impact our analysis.
 
@@
 In the third part of the interview, participants were shown the code snippets that were used in the first study.
-The participants were told both sides of each pair had the same output, and were asked to evaluate which version of the code was easier to understand.
-The interviewers avoided introducing bias in the answers by not explaining that one of the sides of each pair contained the atom under investigation.
-Subjects were just presented the snippets and allowed to take the necessary time to decide on the most readable snippet.
+The snippets were presented in pairs, where each pair contained a clean and a confusing version.
+The participants
+%were told both sides of each pair had the same output, and 
+were asked to evaluate which version of the code was easier to understand.
+To avoid introducing bias in the answers, the interviewers did not explain that one of the versions in each pair contained the atom under investigation.
+Subjects were just presented the pairs of snippets and allowed to take the necessary time to decide on the most readable snippet.
 
@@
 We present these results in Section~\ref{sec:interview-results}.
-Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.
+%Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.
 
@@
 
-To understand how often atoms of confusion appear in open source projects, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories.
+To understand how often the analyzed atom candidates appear in open source projects, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories.
 To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript.
@@
 This metric, available through the GitHub API, represents the number of stars a project received from users of the platform.
-The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}.
+The same metric has been used in a number of previous studies as a proxy to estimate a project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}.
 We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.
@@
 Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research.
-The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
+The sizes of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
 All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars.
@@
 We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study).
-Finally, we carried out basic descriptive statistics in order to understand how often the
-atoms of confusion appear in practice.
+Finally, we carried out basic descriptive statistics in order to understand how often the atoms of confusion appear in practice.
 

### 2021-04-26 4_Result.tex
Updates from Overleaf

@@
 As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten
-code snippets, from which five were in their confusing versions, whilst the other five contained cleaned versions of the
-code snippets (i.e., without the confusing constructs and idioms). 
+code snippets, from which five were in their confusing versions, whilst the other five contained clean versions of the code snippets (i.e., without the confusing constructs and idioms).
 The participants should provide the expected outcomes of the code snippets.
@@
 
-Considering Table~\ref{tab:difference-correctness}, the versions of six code snippets without atoms of confusion present at least a 15\% improvement in answer correctness.
-In particular, the presence of the \emph{Comma Operator} atom presents the highest impact on
-misunderstanding, and  
-frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also introduce high degrees of confusion.
-The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the cleaned version of the code snippets.
+Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the confusing versions.
+In particular, the presence of the \emph{Comma Operator} atom presents the highest impact on misunderstanding, and frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also result in many mistakes.%introduce high degrees of confusion.
+The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the cleaned versions of the code snippets.
 Also, the sample of responses with no atoms had almost no dispersion, which supports the
-argument that the non-confusing versions of the code snippets are
-easier to evaluate correctly.
+argument that the non-confusing versions of the code snippets are easier to evaluate correctly.
 
@@
 We first use the \emph{Pearson's Chi-squared test}
-to investigate if there is a statistically significant difference
-in the frequency of correct and incorrect answers---due to the versions
-of the code snippets (confusing and clean code).
-The result indicates that the confusing versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$).
-We also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant
-\emph{education} or participant \emph{experience} impact correctness.
+to investigate if there is a statistically significant difference in the frequency of correct and incorrect answers---due to the versions of the code snippets (confusing and clean code).
+The p-values for these tests are reported in the ``Chi-square test'' column of Table~\ref{tab:hypothesis-testing}.
+The results indicate that for five atom candidates (Comma Operator, Automatic Semicolon Insertion, Post-Increment, Assignment as Value, Implicit Predicate) the confusing versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$).
+This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
+
+We measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR).
+The results are reported in the ``Odds Ratio'' column of Table~\ref{tab:hypothesis-testing}).
+In the table we also report the confidence interval (CI) for the OR in the ``CI'' column.
+Although many of the intervals are wide, for six of the atom candidates the lower bound of the confidence interval is greater than or equal to 1.
+This indicates that, at a 95\% confidence level, a developer is likely to commit less mistakes when using the clean versions.
+%We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. 
+For instance, when comparing clean and confusing code snippets pertaining to the Comma Operator atom candidate, we observed an \emph{Odds Ratio} of \num{19.02}.
+This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding, confusing version of the code snippet.
+Furthermore, with a 95\% confidence level, the odds  ratio is between 6.6 and 68.22, i.e., although the error margin is wide, it is strongly in favor of the clean version.
+If there is a significant likelihood of a developer committing a mistake when analyzing a clean version, the lower bound of the CI should be a number between 0 and 1 (since the OR is a ratio).
+This is the case for the four atom candidates in the lower part of the table.
+In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a study subject was actually more likely to misunderstand the clean version.
+
+Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant \emph{education} or participant \emph{experience} impact correctness.
 Although the results suggest that experience impacts correctness, the education level does not.
 As such, we replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.
-Finally, we measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR)
-method (see the Odds Ratio column of Table~\ref{tab:hypothesis-testing}). We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and
-Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. For instance, the Comma Operator leads to an \emph{Odds Ratio} of
-\num{19.02}, meaning that the the odds of a correct answer are \num{19.02} times
-higher when interpreting the clean version of a code---in comparison with the corresponding, confusing version of the code snippet.
 
@@
   hypothesis testing suggest that the atom
-  candidates Comma Operator, Post Increment,
-  Omitted Curly Braces, Assignment as Value,
-  Implicit Predicate, and Ternary Operator
+  candidates Comma Operator, Automatic Semicolon Insertion, Post Increment, Omitted Curly Braces, 
+  Assignment as Value, and
+  Implicit Predicate %, and Ternary Operator
   introduce some degree of misunderstanding
-  in JavaScript code. 
+  in JavaScript code. Based on the results of this study, they can be called atoms of confusion for JavaScript.
 \end{mh}
@@
  \centering
- \begin{tabular}{lcccc|cc}
+ \begin{tabular}{lccc|cc}
    \toprule
-Atom & ChiTest & Odds Ratio          & CI 2.5\% & CI 97.5\% & Wilcox Test & Cliff Delta \\ 
-     &         & (misunderstanding)  &          &           & (workload)             & \\ \midrule
-Comma Operator & 1.1727e-10 & 19.02 & 6.60 & 68.22 & 5.8249e-08 & -0.53 \\ 
-Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & 9.21 & 356.62 & 0.09802 & -0.16 \\ 
-Post-Increment & 0.0015279 & 4.83 & 1.73 & 15.74 & 0.12006 & 0.15 \\ 
-Ommitted Curly Braces & 0.050962 & 2.35 & 1.00 & 5.77 & 0.92529 & -0.01 \\ 
-Assignment as Value & 0.0034775 & 8.39 & 1.81 & 79.12 & 0.82681 & 0.02 \\ 
-Implict Predicate & 0.01123 & 6.95 & 1.46 & 66.56 & 0.0029039 & -0.29 \\ 
-Logic as Control Flow & 0.292 & 1.54 & 0.73 & 3.28 & 7.5862e-05 & -0.39 \\ 
-Ternary Operator & 0.15896 & 2.73 & 0.74 & 12.57 & 0.24073 & -0.12 \\ 
-Pre-Increment & 0.70147 & 1.25 & 0.55 & 2.85 & 0.0062981 & 0.27 \\ 
-Arithmetic as Logic & 1 & 0.84 & 0.22 & 3.12 & 0.01723 & 0.23 \\ 
+Atom candidate & Chi-square test & Odds Ratio          & CI & Mann-Whitey U test & Cliff's Delta \\ 
+     &         & (misunderstanding)  &                    & (workload)             & \\ \midrule
+Comma Operator & 1.1727e-10 & 19.02 & (6.60, 68.22) & 5.8249e-08 & -0.53 \\ 
+Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & (9.21, 356.62) & 0.09802 & -0.16 \\ 
+Post-Increment & 0.0015279 & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
+Ommitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
+Assignment as Value & 0.0034775 & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
+Implict Predicate & 0.01123 & 6.95 & (1.46, 66.56) & 0.0029039 & -0.29 \\ 
+Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05 & -0.39 \\ 
+Ternary Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
+Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981 & 0.27 \\ 
+Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & 0.01723 & 0.23 \\ 
     \bottomrule
@@
 as Control Flow, Pre-Increment, and Arithmetic
-as Logic). This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%. 
-For the first four, the analysis suggests that the participants
+as Logic). For the first four, the analysis suggests that the participants
 need more time to predict the outcome of
-the code snippet in the confusing code version. Interesting,
-for the atom candidates Arithmetic as Logic and
-Pre-Increment, the participants take less time
+the code snippet in the confusing code version. Interesting, for the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time
 to predict the outcome of the code snippets in
 the confusing version.
-We also computed the effect size using the Cliff's Delta method.
-We found a large effect for the Comma Operator atom
-candidate; and a moderate effect for the atom candidates
+We also computed the effect size using the Cliff's Delta method. We found a large effect for the Comma Operator atom candidate; and a moderate effect for the atom candidates
 Logic as Control Flow, Implicit Predicate, Arithmetic

### 2021-04-26 5_Threats.tex
Updates from Overleaf

@@
 Furthermore, besides reporting p-values, we also report effect size measures appropriate for each scenario (odds-ratio and Cliff's delta) and apply a p-value correction technique to avoid the multiple testing problem.
-Finally, it could be argued that the size of the samples is insufficient to make conclusions for some of the atoms, a common problem in empirical software engineering.
+Finally, it could be argued that the size of the samples is insufficient to make conclusions for some of the atoms, a common problem in Empirical Software Engineering.
 We estimate the sample size for each one of the atom candidates, considering that each one had a different number of samples (Table~\ref{tab:difference-correctness}).
@@
 Another possible threat to the generalizability of our results, in particular for the survey and interviews, lies in the fact that the analyzed atoms may rarely occur in real JavaScript code bases.
-To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are fairly frequent, occurring at least once per 1,000 lines of code.
+To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are common, occurring at least once per 1,000 lines of code.
 Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code.
+\castor{A forma como os átomos foram detectados é uma ameaça que não discutimos.}
 

### 2021-04-26 main.tex
Updates from Overleaf

@@
 
-\newcommand{\na}{9\xspace}
+\newcommand{\na}{10\xspace}
 
@@
 Although previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in \clang, \cpplang, and Java, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript).
-In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
-From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in {\color{red}6} out of \na atom candidates.
-In addition, our MSR effort reveals that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
+In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study.
+From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in {\color{red}7} out of \na atom candidates.
+In addition, our MSR effort reveals that atoms of confusion frequently arise in 72 popular open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
 Altogether, our findings might help practitioners (a) better understand the implications of atoms of confusion on understanding JavaScript code, (b) avoid writing code that is unnecessarily difficult to maintain, and (c) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.

### 2021-04-26 4_Result.tex
update the survey + spell checker

@@
 Post-Increment & 28.70 & 25.67 & -10.56 \\ 
-Ommitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
+Omitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
 Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
-Implict Predicate & 36.24 & 24.01 & -33.75 \\ 
+Implicit Predicate & 36.24 & 24.01 & -33.75 \\ 
 Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
@@
  \begin{table*}[ht]
-\caption{Hipothesis Testing (misunderstanding and cognitive workload)}
+\caption{Hypotheses Testing (misunderstanding and cognitive workload)}
 \label{tab:hypothesis-testing}
@@
 Post-Increment & 0.0015279 & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
-Ommitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
+Omitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
 Assignment as Value & 0.0034775 & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
-Implict Predicate & 0.01123 & 6.95 & (1.46, 66.56) & 0.0029039 & -0.29 \\ 
+Implicit Predicate & 0.01123 & 6.95 & (1.46, 66.56) & 0.0029039 & -0.29 \\ 
 Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05 & -0.39 \\ 
@@
 suggest that we should refute the null hypotheses for
-six atom candidates (Comma Operator, Automatic Semicolong Insertion,
+six atom candidates (Comma Operator, Automatic Semicolon Insertion,
 Implicit Predicate, Logic
@@
 experience, familiarity with JavaScript, and their opinion
-about the \na atom candidates---by
-(a) presenting them pairs of code snippets whose behavior were identical
-and then (b) asking them to compare the code snippets readability.
+about the \na atom candidates we explored in our research.
+To this end we first present them pairs of code snippets whose behavior were similar
+and then ask the participants to discuss their preferences (i.e.,
+code with or without atom).
 
@@
  shows that for eight out of the
-nine scenarios surveyed, the majority of the respondents agree that the version of the
-code without the atom of confusion is easier to understand, and in no case the \emph{neutral}
+10 scenarios surveyed, the majority of the respondents prefer the version of the
+code without the atom of confusion, and in no case the \emph{neutral}
 ratio was higher than the option for the version without atoms of confusion.
@@
 prefer the version of the code with the atom of confusion,
-instead of the more cleaner version without the atom of confusion.
+instead of the version without the atom of confusion.
 We show the code snippets for the Ternary Operator in Figure~\ref{code:ternary}.
-Even for this atom candidate, some participants who preffered 
+Even for this atom candidate, some participants who preferred 
 the \emph{left-hand-side} version (with the atom of confusion) 
@@
 as simpler to understand, but would still opt to write
-code that contained the atom. In contrast to such opinion,
-one of the participants found the version that contained the atom
+code with the atom. In contrast to such opinion,
+one of the participants found the version with the atom
 of confusion more elegant, but recognized it was less readable,
@@
 the interviewees gave an example of his own experience
-that might motivate one avoid writing code using this construct:
+that might motivate one to avoid writing code using this construct:
 
@@
                                                & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
-         Arithmetic as Logic     & 0  & 93.33  & 6.67  \\
+         Comma Operator                  & 0  & 100    & 0     \\
+         Automatic Semicolon Insertion  & 0  & 80     & 0     \\
+         Post-Increment                  & 20 & 73.33  & 6.67  \\
+         Omitted Curly Braces            & 0  & 100    & 0     \\
          Assignment as Value             & 20 & 60     & 20    \\
-         Comma Operator          & 0  & 100    & 0     \\
-         Ternary Operator        & 60 & 26.67  & 13.3  \\
          Implicit Predicate              & 20 & 73.33  & 6.67  \\
          Logic as Control Flow           & 20 & 60     & 20    \\
-         Omitted Curly Braces    & 0  & 100    & 0     \\
-         Post Increment          & 20 & 73.33  & 6.67  \\
-         Pre-Increment           & 40 & 46.67  & 13.33 \\ \midrule
-         \textsc{overall}        & 20 & 70.37  & 9.63  \\
+         Ternary Operator                & 60 & 26.67  & 13.3  \\
+         Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
+         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \ midrule
+         \textsc{overall}                & 18 & 71.33  & 8.64  \\
          \bottomrule
@@
 often atoms of confusion arise in real software. 
-Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19}, which investigte
+Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19}, which investigate
 the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
@@
 Considering all JavaScript projects in our dataset, we found 
-a total of {\color{red}\num{364873}} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces, and the Post Increment.
-The remaining atoms
-account for \num{24558} (12.03\%) of the total number of occurrences, and the Comma Operator and Arithmetic as Logic are the atoms that arise less frequently (0.001\% in total with 232 and 165 occurrences, respectively).
+a total of {\color{red}\num{364873}} atoms of confusion, though three atoms are responsible
+for {\color{red}87.97\% of this total}: Ternary Operator, Omitted Curly Braces, and the Post Increment. The remaining atoms
+account for \num{24558} (12.03\%) of the total number of occurrences, and the
+Comma Operator and Arithmetic as Logic are the atoms that arise less
+frequently (0.001\% in total with 232 and 165 occurrences, respectively).
 
@@
 according to the thresholds of a previous
-study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. As such, we found three
-atoms as highly used (Ternary Operator, Omitted Curly Braces, and Post Increment),
+study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. As such, we found four
+atoms as highly used (Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment),
 four as commonly used (Pre-Increment, Assignment as Value,
@@
 and two atoms as little used (Arithmetic as Logic and Comma Operator).
-We also calculated the frequency of each atom per KLOC and found the atom
-Ternary Operator as the most occurring in practice (10.16 occurrences per KLOC).
+We also calculated the frequency of each atom per KLOC and found the atoms
+Implicit Predicate (19.89 occurrences per KLOC) and Ternary Operator (10.16
+occurrences per KLOC) as the most occurring
+in practice.
 %In our dataset of JavaScript projects, 
@@
 and Omitted Curly Braces atoms are listed in the top four sources of misunderstanding 
-(see Table~\ref{tab:difference-correctness}). As such, fixing the
-the Post-Increment and Omitted Curly Braces might solve most
-of the unclear code caused by atoms of confusion in
+(see Table~\ref{tab:difference-correctness}), and also frequently appear
+in open source systems. As such, fixing the
+the Post-Increment and Omitted Curly Braces might solve a considerable
+amount unclear code caused by atoms of confusion in
 JavaScript projects.

### 2021-04-26 4_Result.tex
Updates from Overleaf

@@
          Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
-         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \ midrule
+         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \midrule
          \textsc{overall}                & 18 & 71.33  & 8.64  \\
@@
 frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
-occur in at least 80\% of the projects. Considering the extremes, atom Ternary Operator was found in all repositories we mined,
+occur in at least 80\% of the projects. Considering the extremes, atom Ternary Operator and Implicit Predicate were found in all repositories we mined,
 while atom Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
@@
 Considering all JavaScript projects in our dataset, we found 
-a total of {\color{red}\num{364873}} atoms of confusion, though three atoms are responsible
-for {\color{red}87.97\% of this total}: Ternary Operator, Omitted Curly Braces, and the Post Increment. The remaining atoms
-account for \num{24558} (12.03\%) of the total number of occurrences, and the
+a total of \num{364873} atoms of confusion, though four atoms are responsible
+for 92.97\% of this total: Implicit Predicate, Ternary Operator, Omitted Curly Braces, and the Post Increment. The remaining atoms
+account for \num{25621} (7.03\%) of the total number of occurrences, and the
 Comma Operator and Arithmetic as Logic are the atoms that arise less

### 2021-04-28 1_Introduction.tex
Updates from Overleaf

@@
 
-% The source code of a program can be naively understood as a collection of sentences in a language that would be interpreted by a computer (or compiled into a lower level representation and then executed by a computer).
-Nonetheless, this understanding is far from being complete.
-While it is not possible to execute programs without writing them in a way that the machine is able to interpret, it is also very difficult to develop software that performs complex tasks if developers do not care about how easily other programmers are going to understand its source code \cite{DBLP:journals/cj/Knuth84}.
+% Source code misunderstandings
+% Recent work... For instance, consider the pair...
+% However, focus on C, C++ and Java.
+% In this paper ==> Javascript!
 
-% \rb{acho que esse proximo paragrafo pode ser excluido}
-% \diego{também acho}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% INICIO INTRODUCAO - MARCIO
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% {\color{blue}That means programming should, in various ways, be regarded as an act of communication.
-In this context, the programmer is often faced with a dilemma.
-When initially confronted with a problem they have to solve, the programmer will proceed to employ their cognition into developing the logic of the solution, whilst simultaneously having to worry about writing ``syntactically correct'' code for the machine to parse and translate.
-The problem is that, during the time spent solving the problem and communicating with the computer, developers often do not have enough cognitive resources to also make considerations about whether their code is understandable by other programmers.
-Therefore, it takes a change of track to make the code easier to understand \cite{DBLP:books/daglib/0019908}.
-Not proceeding with such change of track is one of the reasons why ``confusing'' language constructs and idioms arise and remain in a program.}
-
-Source code misunderstandings can occur due to diverse reasons.
+Source code misunderstandings can occur due to many reasons.
 For instance, a poorly understood requirement might lead to poorly implemented code.
-An intricate algorithm may require extra time to be fully comprehended.
-Besides that, programming languages can also be at fault
-and specific syntax and semantics of a language can cause misunderstandings. The same applies to some code
-idioms and stylistic choices~\cite{DBLP:conf/msr/GopsteinZFC18}.
-
-% \diego{acho que o prox. parágrafo deve ser adaptado ou até removido}
-% {\color{blue}A direct consequence of the presence of these idioms in the code base is that it becomes even more difficult to develop new features for the software. 
-% According to Gopstein et al.
-\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} ``\emph{the ability to understand pre-existing source code is one of the most important elements of a continuously successful software project}'', it follows that, if one cannot properly understand the \emph{purpose} and \emph{meaning} of a piece of code, they are not going to be able to build upon its existing parts.
-% Other consequences of the presence of confusing code include: 
-% (a) increased probability of bug introduction, due to a lack of proper understanding of what the program does; 
-% and (b) loss of internal quality measures, such as cohesion and coupling. 
-% This phenomenon usually leads to an architectural erosion. 
-% All these situations, which usually compound each other, expose the necessity of developing recommendations, tools, and techniques that help developers to avoid writing confusing code, identify potentially confusing code, and transform confusing code into clear code.
-% Therefore, it is necessary to understand what programming language constructs and idioms might lead to code that is hard to understand.}
-
-%This research focuses on existing JavaScript idioms that a previous research work have proved to introduce misunderstanding in \clang.
-JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}.
-At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''.\footnote{\url{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}} It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.
-All these
-%diverse characteristics and capabilities
-features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.
-
-%In particular, 
-Our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer.
-To achieve that, we investigate JavaScript idioms that might negatively impact program understanding.
-These idioms correspond to a set of candidates for \emph{atoms of confusion} in JavaScript---that is, small
-pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier
-to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For instance, consider the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
-%for presentation purpose only). 
-The code on the left-hand side of the figure contains two atoms of confusion discussed in previous work focusing on the \clang
-language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}.
-On the right-hand side, we present a clean version of the code, that is, without atoms of confusion.
-In the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in  41\% and 34\%, respectively.
+Besides that, programming languages can also have specific constructs that may lead to misunderstandings.
+In this context, the same applies to \emph{atoms of confusion}---that is, small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.
 
 Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability.
-However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java).
-However, JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic
-capabilities and weaker type system.
-Thus, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa.
-Here we investigate a group of atom candidates that stems from the study of Gopstein and colleagues and detail additional JavaScript constructs that might introduce misunderstanding---according to the opinions of JavaScript
-developers.
+For instance, consider the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
+The code at the left-hand side contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}.
+At the right-hand side, we present a clean version of the code, that is, without atoms of confusion.
+According to the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in 41\% and 34\%, respectively.
 
@@
 
-To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 JavaScript developers.
+However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java).
+This way, there is a knowledge gap with respect to the impact of such atoms in dynamic languages such as JavaScript.
+In this sense, because JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic capabilities and weaker type system, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa.
+Also, besides being the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}, JavaScript contains features that combined make the source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}, such as weakly typed language~\cite{DynamicBehaviorJS:PLDI}, object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, and features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.
+
+In this way, to better understand the impact of atoms of confusion in JavaScript, in this paper we present a mixed-method research effort based on a survey, on interviews, and on a mining software repositories study.
+First, we surveyed 140 JavaScript developers.
 In the survey, their task was to predict the output of short code snippets, with half of them containing atoms of confusion.
-{\color{red}The number of correct answers for the
-  \emph{clean version} of the code snippets for 7 out of \na
-  atom candidates is at least 15\% higher than the corresponding
-  \emph{confusing version}
-  of the code snippet (i.e., with an atom candidate). The presence of an atom candidate
-  might also affect the time necessary for submitting a correct answer.
-  For instance, the presence of the Comma Operator atom increases
-  the average time to predict the output by 76.23\%, though the outputs of snippets containing
-  the Pre-Increment atom were correctly predicted 38.19\% faster.
+The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate).
+The presence of an atom candidate might also affect the time necessary for submitting a correct answer.
+For instance, the presence of the Comma Operator atom increases the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.
 
-  To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15
-  experienced professional developers. Participants considered the code snippets without
-  the atom candidates easier to understand in 70\% of the cases,
-  therefore corroborating the survey findings. The participants also discussed other 
-  idioms that might lead developers to misunderstand JavaScript code.
-  The occurrence of atoms in open-source repositories might reveal cases where developers
-  should be more or less careful when writing code---and open the possibility to implement
-  new program transformations for cleaning up confusing code. To identify such cases, we mined popular JavaScript projects
-  from GitHub and found four atom candidates that are highly used in JavaScript
-  code. The Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.}
+To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers.
+Participants considered the code snippets without the atom candidates easier to understand in 70\% of the cases, therefore corroborating the survey findings.
+The participants also discussed other idioms that might lead developers to misunderstand JavaScript code.
+The occurrence of atoms in open-source repositories might reveal cases where developers should be more or less careful when writing code---and open the possibility to implement new program transformations for cleaning up confusing code.
+To identify such cases, we mined popular JavaScript projects from GitHub and found four atom candidates that are highly used in JavaScript code.
+The Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.
 
 Altogether, the main contributions of this paper are:
-\rb{revisar as contribuicoes quando o artigo estiver mais maduro. por enquanto, vamos ignorar esses bullets.}
 
 \begin{itemize}
-    \item A study about the impact of atoms of confusion on a
-    dynamic language.
+
+    \item A mixed-method research effort---based on a survey, interviews, and mining software repositories---about the impact of atoms of confusion on a
+    dynamic language;
+
     \item A list of additional atoms of confusion addressing particular 
-    JavaScript constructs and idioms.
+    JavaScript constructs and idioms;
+
     \item A library and infrastructure to mine atoms of confusion using 
@@
     practitioners to automate program analysis at a large scale.
+
 \end{itemize}
+
+As implications, our results can help researchers and practitioners to better understand atoms of confusion in JavaScript.
+For example, our findings might alert the JavaScript community to avoid writing code with certain atoms of confusion.
+Moreover, our mixed study might help researchers and practitioners to design program transformation tools to remove atoms of confusion in JavaScript.
+They might prioritize atoms that cause a lot of confusion and are very common in the repositories, for instance.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FINAL INTRODUCAO - MARCIO
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%Here we investigate a group of atom candidates that stems from the study of Gopstein and colleagues and detail additional JavaScript constructs that might introduce misunderstanding---according to the opinions of JavaScript developers.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% The source code of a program can be naively understood as a collection of sentences in a language that would be interpreted by a computer (or compiled into a lower level representation and then executed by a computer).
+Nonetheless, this understanding is far from being complete.
+While it is not possible to execute programs without writing them in a way that the machine is able to interpret, it is also very difficult to develop software that performs complex tasks if developers do not care about how easily other programmers are going to understand its source code \cite{DBLP:journals/cj/Knuth84}.
+
+% \rb{acho que esse proximo paragrafo pode ser excluido}
+% \diego{também acho}
+
+% {\color{blue}That means programming should, in various ways, be regarded as an act of communication.
+In this context, the programmer is often faced with a dilemma.
+When initially confronted with a problem they have to solve, the programmer will proceed to employ their cognition into developing the logic of the solution, whilst simultaneously having to worry about writing ``syntactically correct'' code for the machine to parse and translate.
+The problem is that, during the time spent solving the problem and communicating with the computer, developers often do not have enough cognitive resources to also make considerations about whether their code is understandable by other programmers.
+Therefore, it takes a change of track to make the code easier to understand \cite{DBLP:books/daglib/0019908}.
+Not proceeding with such change of track is one of the reasons why ``confusing'' language constructs and idioms arise and remain in a program.}
+
+% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
+%Source code misunderstandings can occur due to diverse reasons.
+For instance, a poorly understood requirement might lead to poorly implemented code.
+An intricate algorithm may require extra time to be fully comprehended.
+Besides that, programming languages can also be at fault and specific syntax and semantics of a language can cause misunderstandings.
+The same applies to some code idioms and stylistic choices~\cite{DBLP:conf/msr/GopsteinZFC18}.
+
+% \diego{acho que o prox. parágrafo deve ser adaptado ou até removido}
+% {\color{blue}A direct consequence of the presence of these idioms in the code base is that it becomes even more difficult to develop new features for the software. 
+% According to Gopstein et al.
+\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} ``\emph{the ability to understand pre-existing source code is one of the most important elements of a continuously successful software project}'', it follows that, if one cannot properly understand the \emph{purpose} and \emph{meaning} of a piece of code, they are not going to be able to build upon its existing parts.
+% Other consequences of the presence of confusing code include: 
+% (a) increased probability of bug introduction, due to a lack of proper understanding of what the program does; 
+% and (b) loss of internal quality measures, such as cohesion and coupling. 
+% This phenomenon usually leads to an architectural erosion. 
+% All these situations, which usually compound each other, expose the necessity of developing recommendations, tools, and techniques that help developers to avoid writing confusing code, identify potentially confusing code, and transform confusing code into clear code.
+% Therefore, it is necessary to understand what programming language constructs and idioms might lead to code that is hard to understand.}
+
+%This research focuses on existing JavaScript idioms that a previous research work have proved to introduce misunderstanding in \clang.
+
+% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
+%JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}.
+At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''.\footnote{\url{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}} It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.
+All these features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.
+
+% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
+%Our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer.
+To achieve that, we investigate JavaScript idioms that might negatively impact program understanding.
+These idioms correspond to a set of candidates for \emph{atoms of confusion} in JavaScript---that is, small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
+For instance, consider the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
+The code on the left-hand side of the figure contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}.
+On the right-hand side, we present a clean version of the code, that is, without atoms of confusion.
+In the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in  41\% and 34\%, respectively.
+
+% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
+%Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability.
+However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java).
+However, JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic capabilities and weaker type system.
+Thus, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa.
+Here we investigate a group of atom candidates that stems from the study of Gopstein and colleagues and detail additional JavaScript constructs that might introduce misunderstanding---according to the opinions of JavaScript developers.
+
+% \begin{figure*}[thb]
+% \noindent\begin{minipage}{.45\textwidth}  
+% \begin{lstlisting}[language=JavaScript]
+% res = properties.reduce((s, b) => {
+%   b && s++
+%   return s
+% }
+% \end{lstlisting}
+% \end{minipage}\hfill
+% \begin{minipage}{.45\textwidth}
+% \begin{lstlisting}[language=JavaScript]
+% res = properties.reduce((s, b) => {
+%   if(b) {
+%     s = s + 1
+%   }
+%   return s
+% }
+% \end{lstlisting}
+% \end{minipage}
+% \captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project. Line 2 on the left-hand side of the figure contains two
+%   atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic
+%   as Control Flow} and \emph{Post-increment}.
+On the right-hand side, we present an alternative implementation without atoms of confusion.
+We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p).length}).}\label{fig:lst01}
+% \end{figure*}
+
+% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
+%To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 JavaScript developers.
+In the survey,  their task was to predict the output of short code snippets, with half of them containing atoms of confusion.
+{\color{red}The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate).
+The presence of an atom candidate might also affect the time necessary for submitting a correct answer.
+For instance, the presence of the Comma Operator atom increases  the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.
+
+% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
+%To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15 experienced professional developers.
+Participants considered the code snippets without the atom candidates easier to understand in 70\% of the cases, therefore corroborating the survey findings.
+The participants also discussed other idioms that might lead developers to misunderstand JavaScript code.
+The occurrence of atoms in open-source repositories might reveal cases where developers should be more or less careful when writing code---and open the possibility to implement new program transformations for cleaning up confusing code.
+To identify such cases, we mined popular JavaScript projects from GitHub and found four atom candidates that are highly used in JavaScript code.
+The Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.
+
+% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
+%Altogether, the main contributions of this paper are: \rb{revisar as contribuicoes quando o artigo estiver mais maduro. por enquanto, vamos ignorar esses bullets.}
+
+% \begin{itemize}
+%     \item A study about the impact of atoms of confusion on a
+%     dynamic language.
+%     \item A list of additional atoms of confusion addressing particular 
+%     JavaScript constructs and idioms.
+%     \item A library and infrastructure to mine atoms of confusion using 
+%     code queries (implemented in the .QL language). This library and 
+%     infrastructure are freely available and might help researchers and 
+%     practitioners to automate program analysis at a large scale.
+% \end{itemize}
 

### 2021-04-28 2_Background.tex
Updates from Overleaf

@@
-\section{Background and Related Work}
-\label{back}
-\subsection{Program Comprehension}
+%\section{Background and Related Work}
+%\label{back}
+%\subsection{Program Comprehension}
 
@@
 
-The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96}.
+% The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96}.
 Regardless of whether they are performing maintenance on legacy software or adding new features, programmers have to understand what the source code does before making any improvements.
 A natural consequence of this is that a programmer must also have in mind that other programmers (or their future selves)
-%are going to have to
-will
-spend time trying to understand the code being currently developed before doing any work themselves.
+% %are going to have to
+% will
+% spend time trying to understand the code being currently developed before doing any work themselves.
 
-As program comprehension is regarded less as a systematic process than as an objective \cite{DBLP:journals/ibmsj/OHareT94}, there is no single approach or framework that is capable of yielding easily understandable programs.
+% As program comprehension is regarded less as a systematic process than as an objective \cite{DBLP:journals/ibmsj/OHareT94}, there is no single approach or framework that is capable of yielding easily understandable programs.
 Factors such as problem domain, size and complexity of the code base, and available tools all vary significantly between different settings.
@@
 
-However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions---when these exist---that are potentially less complex.
+% However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions---when these exist---that are potentially less complex.
 %allowed by the language or framework they are working with.
-This can be a matter of personal taste, since different individuals might hold different opinions on whether a given syntax is difficult to understand.
+% This can be a matter of personal taste, since different individuals might hold different opinions on whether a given syntax is difficult to understand.
 Nevertheless, it is possible to empirically investigate if there exist programming language constructs or idioms that are considerably harder to understand, leading readers of the code to frequently make incorrect predictions of its behavior.
 
-\subsection{Atoms of Confusion}\label{sec:aoc}
+%\subsection{Atoms of Confusion}\label{sec:aoc}
 
-The concept of atoms of confusion was introduced by Gopstein et al.
+\section{Related Work}
+
+\label{back}
+
+The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96, DBLP:journals/ibmsj/OHareT94}.
+Atoms of confusion is an important field of program comprehension introduced by Gopstein et al.
 \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings.
 In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion.
-Their work laid an empirical framework to identify such atoms.
-In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atom candidates were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.
-%By having subjects evaluate code where in the expected output relied exclusively on evaluating the result of previously known confusing programs that had been written in \clang or \cpplang, and asking the participants to predict the output of a set of code snippets where the atoms had been removed, the authors were able to isolate and measure the impact of such confusing atoms.
-
-One example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
+One example that occurs in many programming languages is the Change of Literal Encoding.
+For instance, the \clang code statement \lstinline{printf("\%d", 013);}
 often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}.
@@
 Gopstein et al.
-\cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on how confusion manifested when developers inspected code exhibiting different atoms.\castor{I'd rather have two concrete examples here, instead of this.} They note that, for some atoms, a particular incorrect answer was prevalent, whilst there was a variety of incorrect answers for other atoms.
+\cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on how confusion manifested when developers inspected code exhibiting different atoms.
+%\castor{I'd rather have two concrete examples here, instead of this.}
+They note that, for some atoms, a particular incorrect answer was prevalent, whilst there was a variety of incorrect answers for other atoms.
 
-The work of Gopstein et al.
-\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} presented the results of comprehensive research on atoms of confusion that can be found in the \clang and \cpplang~languages.
+%Their work laid an empirical framework to identify such atoms.
+In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atom candidates were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.
+%By having subjects evaluate code where in the expected output relied exclusively on evaluating the result of previously known confusing programs that had been written in \clang or \cpplang, and asking the participants to predict the output of a set of code snippets where the atoms had been removed, the authors were able to isolate and measure the impact of such confusing atoms.
+
+Gopstein et al.
+\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} also presented the results of a comprehensive research on atoms of confusion that can be found in the \clang and \cpplang~languages.
 Even though JavaScript is a dynamic language that differs from \clang in a variety of aspects, it also has a number of constructs in common.
@@
 
-The first study of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} inspired other studies to further investigate the impact of atoms of confusion on code comprehension.
-Oliveira et al.~\cite{TheEyesDoNotLie} conducted an experiment using an eye-tracker with 30 participants involving three atoms.
+%The first study of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} inspired other studies to further investigate the impact of atoms of confusion on code comprehension.
+
+Oliveira et al.~\cite{TheEyesDoNotLie} conducted an experiment using an eye-tracker with 30 participants involving three atoms in \clang.
 They found that code with atoms of confusion requires more time from participants to predict the output and more visual effort to comprehend.
@@
 The authors argue that this method is important to complement a quantitative analysis.
-Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''
-}. We follow a similar approach as we augment our quantitative analysis with interviews.
+Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''}.
+We follow a similar approach as we augment our quantitative analysis with interviews.
 
-Gopstein et al. also influenced researchers to look into atoms in languages other than \clang and \cpplang.
+%Gopstein et al. also influenced researchers to look into atoms in languages other than \clang and \cpplang.
 Langhout and Aniche~\cite{Langhout:2021:ACJ} derived a set of 14 atoms for Java and performed an experiment with 132 students.
@@
 Unlike JavaScript, in Swift, most of the atoms identified by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} are avoided by construction, e.g., it does not have assignments as values, increment operators, or macros.
-The paper also presents a structured definition for what it means for a code pattern to be an atom of confusion and a set of general principles to help in the identification of new atoms.
+%The paper also presents a structured definition for what it means for a code pattern to be an atom of confusion and a set of general principles to help in the identification of new atoms.
 
 Medeiros et al.
-\cite{DBLP:journals/ese/MedeirosLAAKRG19} analyzed 50 open source projects written in the C language with the goal of evaluating 12 code patterns called ``misunderstanding patterns'' by the authors.
+\cite{DBLP:journals/ese/MedeirosLAAKRG19} analyzed 50 open source projects written in the \clang language with the goal of evaluating 12 code patterns called ``misunderstanding patterns'' by the authors.
 Many of these misunderstanding patterns are either atoms of confusion or atom candidates~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
@@
 The authors of the study received feedback for 21 of these pull requests and the maintainers of the projects accepted 8 of them (22.86\%).
-The authors also analyzed 36 coding style guides from these projects, but found few guidelines specifically addressing misunderstanding patterns.
+%The authors also analyzed 36 coding style guides from these projects, but found few guidelines specifically addressing misunderstanding patterns.
 

### 2021-04-28 3_Method.tex
Updates from Overleaf

@@
 programmers to take longer to predict their output.
-Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs.
+Since JavaScript and C have some constructs in common (Section~\ref{back}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs.
 Our supplementary material

### 2021-04-28 6_Conclusion.tex
Updates from Overleaf

@@
 
-\rb{draft. apenas movi esse texto que estava em settings para ca. pode servir de inspiracao para responder algumas das questoes de pesquisa.}
+This paper introduced a mixed-method research effort to better understand the impact of atoms of confusion in JavaScript code.
+First, we conducted a survey with 140 JavaScript developers.
+We asked them to predict the output of some code snippets (with and without atoms).
+We provide evidences that, even though the atoms of confusion we considered do not take a much higher amount of time to predict the output of the code, they might still have a significant impact on program comprehension activities in almost all of the cases we studied.
+Afterwards, we interviewed 15 professional developers.
+In general, they considered code without atoms of confusion easier to understand.
+Last but not least, we conducted a mining software repositories study.
+In particular, we found four atoms of confusion highly used in JavaScript code, i.e., Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment.
 
-{\color{blue}Based on these results, we provide evidence that, although the atoms of confusion considered in our study do not take a much higher amount of time to predict the output of the code, their impact on program comprehension is, in almost all of our cases, highly significant.}
+Our efforts have important implications for practice.
+First, we alert developers to avoid writing JavaScript code with certain atoms of confusion ({\color{red} \textbf{EXEMPLO AQUI???}).} Second, we might help tool developers to create a program transformation tool to remove automatically atoms of confusion.
+They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice.
+
+%\rb{draft. apenas movi esse texto que estava em settings para ca. pode servir de inspiracao para responder algumas das questoes de pesquisa.}
+
+%{\color{blue}Based on these results, we provide evidence that, although the atoms of confusion considered in our study do not take a much higher amount of time to predict the output of the code, their impact on program comprehension is, in almost all of our cases, highly significant.}

### 2021-04-29 5_Threats.tex
Updates from Overleaf

@@
 
-Conclusion validity is connected with how well it is possible to establish relationships between treatments and outcomes.
+%Conclusion validity is connected with how well it is possible to establish relationships between treatments and outcomes.
 Threats to conclusion validity often come from inappropriate use of statistics.
-In the context of our survey study, we apply different statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test).
+
+\textbf{Conclusion validity.} In the context of our survey study, we apply different statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test).
 Furthermore, besides reporting p-values, we also report effect size measures appropriate for each scenario (odds-ratio and Cliff's delta) and apply a p-value correction technique to avoid the multiple testing problem.
@@
 For the ones where we did not find statistically significant results for correctness, in fact the sample sizes we have analyzed are insufficient for such small effect sizes.
-This indicates that further studies on these atoms are required.
+This indicates that further studies on these atoms are required, due to the likelihood of type II errors.
 
@@
 
-Construct validity is connected with how well the selected measures actually represent the concept of interest.
-We use correctness as a proxy for program comprehension.
+%Construct validity is connected with how well the selected measures actually represent the concept of interest.
+
+\textbf{Construct validity.} We use correctness as a proxy for program comprehension.
 Furthermore, we use predicted outcomes of small programs as a measure of correctness.
@@
 
-Internal validity is concerned with how well the study isolates the variables of interest and accounts for confounding factors.
-Since the survey was conducted online with unknown participants, we have no way of confirming their levels of education and experience.
+%Internal validity is concerned with how well the study isolates the variables of interest and accounts for confounding factors.
+
+\textbf{Internal validity.} Since the survey was conducted online with unknown participants, we have no way of confirming their levels of education and experience.
 We mitigate this threat by, in the data analysis, explicitly accounting for programmer experience and using an experimental design that allows us to isolate the impact of the treatment from factors such as experience and formal education level.
@@
 
-External validity is linked to whether it is possible to extrapolate the results of the study.
-Although our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, it is not clear whether that result extrapolates to other scenarios.
+%External validity is linked to whether it is possible to extrapolate the results of the study.
+
+\textbf{External validity.} Although our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, it is not clear whether that result extrapolates to other scenarios.
 Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect.

### 2021-04-29 1_Introduction.tex
minor changes (introduction)

@@
 
-Source code misunderstandings can occur due to many reasons.
-For instance, a poorly understood requirement might lead to poorly implemented code.
-Besides that, programming languages can also have specific constructs that may lead to misunderstandings.
-In this context, the same applies to \emph{atoms of confusion}---that is, small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.
+Source code misunderstandings can occur due to many reasons; for instance, the wrong choice of names for arguments, variables, and methods might hinder developers from understanding a piece of code~\cite{avidan:icpc2017}.
+Besides that, programming languages can also have specific idioms and constructs that may lead to misunderstandings.
+That is the case of \emph{atoms of confusion}---small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.
 
 Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability.
-For instance, consider the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
+For instance, consider the pair of JavaScript code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
 The code at the left-hand side contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}.
@@
 In this sense, because JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic capabilities and weaker type system, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa.
-Also, besides being the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}, JavaScript contains features that combined make the source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}, such as weakly typed language~\cite{DynamicBehaviorJS:PLDI}, object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, and features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.
+Also, besides being the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}, JavaScript contains features that combined make the source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.
+%, such as weakly typed language~\cite{DynamicBehaviorJS:PLDI}, object-oriented, functional,
+% and imperative constructs~\cite{JS-CodeSmells:SCAM}, and features such as run-time
+% evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.
 
@@
 First, we surveyed 140 JavaScript developers.
-In the survey, their task was to predict the output of short code snippets, with half of them containing atoms of confusion.
+In the survey, their task was to predict the output of small code snippets, with half of them containing atoms of confusion.
 The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate).
@@
 For instance, the presence of the Comma Operator atom increases the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.
+To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers.
+Participants prefered the code snippets without the atom candidates in 70\% of the cases, therefore corroborating the survey findings.
+The participants also discussed other idioms and constructs that might lead developers to misunderstand JavaScript code.
 
-To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers.
-Participants considered the code snippets without the atom candidates easier to understand in 70\% of the cases, therefore corroborating the survey findings.
-The participants also discussed other idioms that might lead developers to misunderstand JavaScript code.
-The occurrence of atoms in open-source repositories might reveal cases where developers should be more or less careful when writing code---and open the possibility to implement new program transformations for cleaning up confusing code.
-To identify such cases, we mined popular JavaScript projects from GitHub and found four atom candidates that are highly used in JavaScript code.
-The Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.
+The frequency in which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code.
+To understand the incidence of atoms, we mined popular JavaScript projects from GitHub and found that atoms of confusion frequently appear in practice.
+Indeed, four atom candidates are highly used in JavaScript code---the Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.
 
@@
 
-    \item A list of additional atoms of confusion addressing particular 
-    JavaScript constructs and idioms;
+  \item A list of additional, JavaScript-specific
+    constructs and idioms that might introduce code
+    misunderstanding;
 

### 2021-04-29 2_Background.tex
fix some misleading information (related-work)

@@
 The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase.
-Our first study follows the empirical framework discussed in~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, for all atoms that have a counterpart in JavaScript.
-Gopstein et al.
-\cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on how confusion manifested when developers inspected code exhibiting different atoms.
-%\castor{I'd rather have two concrete examples here, instead of this.}
-They note that, for some atoms, a particular incorrect answer was prevalent, whilst there was a variety of incorrect answers for other atoms.
+In a different work, Gopstein et al.
+\cite{DBLP:conf/msr/GopsteinZFC18} presented the results of a comprehensive research on the incidence of atoms of confusion in the wild, considering open-source projects written in \clang and \cpplang~programming languages.
 
@@
 
-Gopstein et al.
-\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} also presented the results of a comprehensive research on atoms of confusion that can be found in the \clang and \cpplang~languages.
+These previous studies~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17,DBLP:conf/msr/GopsteinZFC18}
+have a greate influence in our work.
 Even though JavaScript is a dynamic language that differs from \clang in a variety of aspects, it also has a number of constructs in common.
@@
 
-According to Arnaoudova et al.~\cite{Arnaoudova:2016:LAW}, Linguistic Antipatterns (LA) are \textit{``recurring poor practices in the naming, documentation, and choice of identifiers in the implementation of an entity [..] that may impair program understanding''}.
+%% According to Arnaoudova et al.~\cite{Arnaoudova:2016:LAW}, Linguistic Antipatterns (LA) are \textit{``recurring poor practices in the naming, documentation, and choice of identifiers in the implementation of an entity [..] that may impair program understanding''}.
 These antipatterns are design issues that go against developer's  intuition about a program, for example, a \texttt{set} method that returns a result.
@@
 Santos and Gerosa~\cite{Santos:2018:ICP} analyze the impact of eleven coding practices on program readability.
-Some of them, such as avoiding more than three levels of code block nesting, are atom candidates.
-% Diego: Troquei "this work" para "their work". Removi o "more" de "are more related"
-Their work evaluates the coding practices solely based on personal opinions of the study subjects.
-Furthermore, it focuses on the Java language.
-Finally, most of the practices analyzed in this study are more related to code formatting than language constructs or idioms.
+Although some of these practices, such as avoiding more than three levels of code block nesting, are atom candidates, most of the practices analyzed in their study are more related to code formatting than language constructs or idioms.
+We borrow from their research the methodology for collecting the opinion of practitioners regarding code snippets preference.
 

### 2021-04-29 2_Background.tex
minor fixes

@@
 These previous studies~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17,DBLP:conf/msr/GopsteinZFC18}
-have a greate influence in our work.
+have a great influence in our work.
 Even though JavaScript is a dynamic language that differs from \clang in a variety of aspects, it also has a number of constructs in common.
@@
 Although some of these practices, such as avoiding more than three levels of code block nesting, are atom candidates, most of the practices analyzed in their study are more related to code formatting than language constructs or idioms.
-We borrow from their research the methodology for collecting the opinion of practitioners regarding code snippets preference.
+We borrow from their research the methodology for collecting the practitioners opinion about code snippets preference.
 

### 2021-04-29 4_Result.tex
additional review (conclusions)

@@
     \centering
-    \caption{Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}).
-Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
+    \caption{Summary of participants' preferences for code snippets \emph{with} (LHS) and
+      \emph{without} atoms of confusion (RHS).
+      Participants were also allowed to choose NEUTRAL when they thought both sides were equally readable.}
     \label{tab:interview-results1}
@@
       \cmidrule(lr){2-4}
-         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
-                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
-                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
+         \textsc{atom}           & \multicolumn{1}{c}{\textsc{LHS}}
+                                      &  \multicolumn{1}{c}{\textsc{RHS}}
+                                               & \multicolumn{1}{c}{NEUTRAL} \\ \midrule
          Comma Operator                  & 0  & 100    & 0     \\
@@
 Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.
+We also found other idioms that are JavaScript atom candidates, including
+\emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).
+
+\begin{figure}
+\begin{small}
+\begin{lstlisting}[language=JavaScript,caption=Example of arrow function,label=arr1]
+let inc = (x) => x + 1;
+\end{lstlisting}
+\begin{lstlisting}[language=JavaScript,caption=Alternative version without arrow function,label=arr2]
+let inc = function(x) {
+  return x + 1
+}
+\end{lstlisting}
+\end{small}
+\end{figure}
 
@@
 
-Six practitioners regarded the JavaScript language as easy to learn, while four
-participants mentioned that the abundance of frameworks,
-coupled with the abstractions they provide, make it so that one
-can develop quickly in the language. However, as mentioned by one
-of the interviewees, ``\emph{Developers who have started
-programming in JavaScript might miss out on important lower-level
-aspects, such as garbage collection, memory management,
-and process management.}''.
+% TODO: I don't see that our results support these claims.
+% Let us comment for now.
 
-Moreover, such abstractions might lead a developer to start to program
-using the language without actually understanding some of its basic aspects.
-For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
+%% Six practitioners regarded the JavaScript language as easy to learn, while four
+%% participants mentioned that the abundance of frameworks,
+%% coupled with the abstractions they provide, make it so that one
+%% can develop quickly in the language. However, as mentioned by one
+%% of the interviewees, ``\emph{Developers who have started
+%% programming in JavaScript might miss out on important lower-level
+%% aspects, such as garbage collection, memory management,
+%% and process management.}''.
+
+%% Moreover, such abstractions might lead a developer to start to program
+%% using the language without actually understanding some of its basic aspects.
+%% For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
 %% For the multi-purpose language that is one of the most, if not the most used one by practitioners,

### 2021-04-29 5_Threats.tex
additional review (conclusions)

@@
 As a result of this analysis, we discover that for all but one of the atoms where we had a statistically significant result for correctness, Implicit Predicate, the sample size is sufficient.
-For the ones where we did not find statistically significant results for correctness, in fact the sample sizes we have analyzed are insufficient for such small effect sizes.
-This indicates that further studies on these atoms are required, due to the likelihood of type II errors.
+% For the ones where we did not find statistically significant results for correctness, in fact the sample sizes
+% we have analyzed are insufficient for such small effect sizes. This indicates that further studies on these
+% atoms are required, due to the likelihood of type II errors.
 
@@
 Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code.
-\castor{A forma como os átomos foram detectados é uma ameaça que não discutimos.}
+% \castor{A forma como os átomos foram detectados é uma ameaça que não discutimos.}
 

### 2021-04-29 6_Conclusion.tex
additional review (conclusions)

--- 2020-11-09 6_Conclusion.tex
@@
 
-\rb{draft. apenas movi esse texto que estava em settings para ca. pode servir de inspiracao para responder algumas das questoes de pesquisa.}
+This paper has reported the results of a mixed-method research effort to better understand the impact of atoms of confusion in JavaScript code.
+First, we conducted a survey with 140 JavaScript developers.
+We asked them to predict the output of some code snippets (with and without atoms).
+We provide evidences that seven atom candidates have a significant impact on program comprehension activities.
 
-{\color{blue}Based on these results, we provide evidence that, although the atoms of confusion considered in our study do not take a much higher amount of time to predict the output of the code, their impact on program comprehension is, in almost all of our cases, highly significant.}
+Afterwards, we interviewed 15 professional developers.
+In general, they considered code without atoms of confusion easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding.
+Last but not least, we conducted a mining software repositories study.
+In particular, we found four atoms of confusion highly used in JavaScript code, i.e., Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment.
+
+Our efforts have important implications for practice.
+First, we alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment, and Omitted Curly Braces).
+Second, our results might help tool developers to create program transformation tools to remove atoms of confusion automaticallu.
+They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).
+
+As mentioned, the results of the interviews pointed out additional JavaScript constructs and idioms
+that might introduce misunderstanding. These include not only nested callbacks and prototype-based
+inheritance (which might not have a counterpart version), but also idioms such
+as \emph{property access} and \emph{arrow functions}. As future work, we intend to
+reproduce our study to validate whether or not these \emph{atom candidates} affect
+the understanding of JavaScript code.

### 2021-04-29 main.tex
Updates from Overleaf
additional review (conclusions)

@@
 \usepackage{enumerate}
+\usepackage{balance}
 \usepackage{caption}
@@
   stepnumber=1,
+  keywords={let},
 }    
@@
 In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study.
-From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in {\color{red}7} out of \na atom candidates.
+From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present inseven out of ten atom candidates.
 In addition, our MSR effort reveals that atoms of confusion frequently arise in 72 popular open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
@@
 
+\balance
 \bibliographystyle{IEEEtran}

### 2021-04-30 2_Background.tex
review on the methodology section

@@
 Santos and Gerosa~\cite{Santos:2018:ICP} analyze the impact of eleven coding practices on program readability.
-Although some of these practices, such as avoiding more than three levels of code block nesting, are atom candidates, most of the practices analyzed in their study are more related to code formatting than language constructs or idioms.
+However, most of the practices analyzed in their study relate to code formatting, instead of language constructs or idioms that might introduce misunderstanding.
 We borrow from their research the methodology for collecting the practitioners opinion about code snippets preference.

### 2021-04-30 3_Method.tex
review on the methodology section

@@
 In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice.
-Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions---though we postpone these results to future research work.
-To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and a JavaScript software repository mining study.
+Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions.
+%---though we postpone these results to future research work.
+We conduct a mixed-methods study to anwer these questions, including a survey, a set of interviews, and a study for mining JavaScript software repositories.
 
@@
 programmers to take longer to predict their output.
-Since JavaScript and C have some constructs in common (Section~\ref{back}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs.
-Our supplementary material
-summarizes all atoms we consider in our research.
+Since JavaScript and C have some constructs in common (Section~\ref{back}), we first selected a set of 9+1 atom candidates: 9 atoms of confusion already validated~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} for the \clang language that also exist in JavaScript programs + 1 atom candidate that is specific of the JavaScript language (Automatic Semicolon Insertion).
+Our supplementary material discusses and shows examples of
+the atoms we consider in our research.
 
@@
 considers two treatments: the presence or absence of 
-atoms of candidates within the code snippets. 
+atoms candidates within the code snippets. 
 To achieve such a design goal, controlling the effect of experience and individual code snippets, we resorted to the \textit{Latin Square Design} \cite{Hunter-Experimenters}.
@@
 
-We posted the survey on Reddit.\footnote{Reddit is a North American online discussion platform.
-Its discussion threads, often called subreddits, are sorted by subject.
-For our research, we posted information and links to the survey in two JavaScript subreddits.} We explained our research purposes and asked developers of any level of expertise to take the survey.
+We posted the survey on a JavaScript Reddit channel.\footnote{https://www.reddit.com/r/javascript/} We explained our research purposes and asked developers of any level of expertise to take the survey.
 Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares.
@@
 Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the survey.
-When this happened, the corresponding row in the square was left incomplete.
-We considered all squares that contained incomplete rows to be invalid during data analysis and discarded them.
+We discarded from our study all squares that contained incomplete rows (a common approach for data imputation).
 %Since we had a large enough number of samples, the squares we had to discard did not impact our analysis.

### 2021-05-03 3_Method.tex
Updates from Overleaf

--- 2021-04-28 3_Method.tex
@@
 Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions---though we postpone these results to future research work.
-To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and a JavaScript software repository mining study.
+To answer these research questions, we conduct a mixed-methods study, including a survey, a set of interviews, and a JavaScript software repository mining study.
 
@@
 try to predict their outcome and (b) require
-programmers to take longer to predict their output.
+programmers more time to predict their output.
 Since JavaScript and C have some constructs in common (Section~\ref{back}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs.
@@
 considers two treatments: the presence or absence of 
-atoms of candidates within the code snippets. 
+atom candidates in the code snippets. 
 To achieve such a design goal, controlling the effect of experience and individual code snippets, we resorted to the \textit{Latin Square Design} \cite{Hunter-Experimenters}.
@@
 We call them the confusing versions of the programs.
-We also wrote corresponding short programs without the confusing idioms and constructs and we call them the clean versions of the programs.
+We also wrote corresponding short programs without the confusing idioms and constructs, which we call the clean versions of the programs.
 Overall, the survey employed 20 code snippets, 10 confusing and 10 clean.
@@
 
-We decided not to provide feedback about the time students took to answer each question.
+We decided not to provide feedback about the time participants took to answer each question.
 We also did not tell them whether their answers were correct or not.
@@
 That is, starting from our network of contacts, we invited an initial set of candidates to take part in our survey.
-From this initial list, we asked for an indication of additional candidates.
+From this initial list, we asked for recommendations of additional candidates.
 Our main selection criterion was that all participants should have been working with JavaScript in their daily professional activities.
@@
   \centering
+\caption{Demographic information of the interviewees}
 \begin{scriptsize}  
@@
 \end{scriptsize}
-    \caption{Demographic information of the participants}
+
     \label{pinterview}
@@
 
-In the second part of the interview our aim was to allow the subjects to describe their JavaScript experience, as well as to allow them to reveal any JavaScript constructs they regarded as innately confusing.
+In the second part of the interview our goal was to allow the subjects to describe their JavaScript experience, as well as to allow them to reveal any JavaScript constructs they regarded as innately confusing.
 This would allow us to identify potential atom candidates that are more specific to the JavaScript language.
@@
 We first transcribed each interview and then examined the broad distribution of the answers.
-Our goal was to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular.
+Our goal was to build an initial understanding of the participants' perceptions on the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular.
 We next followed with an open-coding procedure, highlighting the main themes and quoting the answers of the participants.
@@
 
-To understand how often the analyzed atom candidates appear in open source projects, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories.
+% To understand how often the analyzed atom candidates appear in open source projects, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories.
 To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript.
-We measured popularity using the project's stargazers.
+We measured popularity using a project's stargazers.
 This metric, available through the GitHub API, represents the number of stars a project received from users of the platform.
@@
 
-After filtering out JavaScript project candidates, in the second step we built a curated dataset comprising the top \minedprojects repositories.
+% After filtering out JavaScript project candidates, in the second step we built a curated dataset comprising the top \minedprojects repositories.
 Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}.
 Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research.
-The sizes of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
+The sizes of the projects range from small ones (\num{5543} lines of code) to complex systems with more than 1 MLOC.
 All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars.
@@
 
-We mined the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}.
+To understand how often the analyzed atom candidates appear in open source projects and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of popular GitHub open source repositories.
+We measured popularity using a project's number of stars, as done in other studies~\cite{gyimesi2019bugsjs,canedo:esem2020}. 
+We first collected the top-100 most popular GitHub repositories that are primarily written in JavaScript. 
+Then we selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code. 
+We removed 28 repositories, resulting in the final curated dataset of 72 repositories.
+
+Our repository selection includes very popular projects.
+Examples are  \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}.
+Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research.
+The sizes of the projects range from small ones (\num{5543} lines of code) to complex systems with more than 1 MLOC.
+All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars.
+We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.
+
+We mined the occurrence of atom candidates from the repositories in our curated dataset using source code queries wirtten in the CodeQL language~\cite{moor:gttse2007}.
 CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript).
-We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study).
+We also automated the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study).
 Finally, we carried out basic descriptive statistics in order to understand how often the atoms of confusion appear in practice.
@@
   \centering
+  \caption{Some descriptive statistics about the repositories used in the MSR study}
   \begin{scriptsize}
  \begin{tabular}{rrrrr}
-   \hline
-                       & Min.             & Median        & Mean             & Max. \\ \hline
+    \toprule
+                       & Min.             & Median        & Mean             & Max. \\ \midrule
  Lines of Code         & \num{5543}       & \num{36161.5} & \num{111432.33}  & \num{1278405} \\
@@
 
-    \hline
+    \bottomrule
  \end{tabular}
  \end{scriptsize}
- \caption{Some descriptive statistics about the projects used in the MSR study}
  \label{tab:projects-statistics} 

### 2021-05-03 4_Result.tex
Updates from Overleaf

@@
 We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
-and \emph{cognitive effort} (time necessary to provide a correct answer). We receive full answers to our survey
-from 140 participants (in a total of 70 replicas).
+and \emph{cognitive effort} (time necessary to provide a correct answer). 
+We received full answers to our survey from 140 participants (a total of 70 replicas).
 All participants had taken at least some university course or hold a bachelor
@@
 Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the confusing versions.
-In particular, the presence of the \emph{Comma Operator} atom presents the highest impact on misunderstanding, and frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also result in many mistakes.%introduce high degrees of confusion.
-The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the cleaned versions of the code snippets.
+In particular, the presence of the \emph{Comma Operator} atom presents the highest impact on misunderstanding.
+Frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also result in many mistakes.
+%introduce high degrees of confusion.
+The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the clean versions of the code snippets.
 Also, the sample of responses with no atoms had almost no dispersion, which supports the
@@
 \begin{table}[htbp]
-\caption{Total number of correct answers.}
+\caption{Total number of correct answers}
 \label{tab:difference-correctness}
@@
  \includegraphics[width=\columnwidth]{images/wrong-answers-plot-1.pdf}
- \caption{Number of wrong answers of each subject}
+ \caption{Number of wrong answers of each subject.}
  \label{fig:boxplotcorrectness}
@@
 
-We measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR).
-The results are reported in the ``Odds Ratio'' column of Table~\ref{tab:hypothesis-testing}).
-In the table we also report the confidence interval (CI) for the OR in the ``CI'' column.
+We measure the effect size of the clean version of the code snippets into the answers' correctness using the \emph{Odds Ratio} (OR).
+The results are reported in the ``Odds Ratio'' column of Table~\ref{tab:hypothesis-testing}.
+In the table, we also report the confidence interval (CI) for the OR in the ``CI'' column.
 Although many of the intervals are wide, for six of the atom candidates the lower bound of the confidence interval is greater than or equal to 1.
@@
 For instance, when comparing clean and confusing code snippets pertaining to the Comma Operator atom candidate, we observed an \emph{Odds Ratio} of \num{19.02}.
-This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding, confusing version of the code snippet.
+This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding confusing version of the code snippet.
 Furthermore, with a 95\% confidence level, the odds  ratio is between 6.6 and 68.22, i.e., although the error margin is wide, it is strongly in favor of the clean version.
@@
   introduce some degree of misunderstanding
-  in JavaScript code. Based on the results of this study, they can be called atoms of confusion for JavaScript.
+  in JavaScript code. %Based on the results of this study, 
+  Therefore, they can be deemed atoms of confusion for JavaScript.
 \end{mh}
@@
    \toprule
-Atom candidate & Chi-square test & Odds Ratio          & CI & Mann-Whitey U test & Cliff's Delta \\ 
+Atom  & Chi-square test & Odds Ratio          & CI & Mann-Whitey U test & Cliff's Delta \\ 
      &         & (misunderstanding)  &                    & (workload)             & \\ \midrule
@@
 need more time to predict the outcome of
-the code snippet in the confusing code version. Interesting, for the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time
+the code snippet in the confusing code version. Interestingly, for the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time
 to predict the outcome of the code snippets in
@@
   Control Flow, and Implicit Predicate. Conversely,
-  the version of the code snippets with the
+  the code snippets with the
   atom candidates Arithmetic as Logic and
-  Pre-Increment demands less time to
+  Pre-Increment demand less time to
   predict the correct answers. 
@@
 
-In this section we present the results of the 
-interviews with the practitioners who agreed to participate
-in the study. A total of 15 practitioners
-took part of the interviews.
+In this section, we present the results of the 
+interviews with the 15 practitioners who agreed to participate
+in the study. 
+% A total of 15 practitioners took part of the interviews.
 We collected information regarding programming
@@
 about the \na atom candidates we explored in our research.
-To this end we first present them pairs of code snippets whose behavior were similar
-and then ask the participants to discuss their preferences (i.e.,
-code with or without atom).
+We first present them pairs of clean and confusing code snippets 
+and then ask them to discuss their preferences towards one version.
+% ask the participants to discuss their preferences (i.e., code with or without atom).
 
@@
 10 scenarios surveyed, the majority of the respondents prefer the version of the
-code without the atom of confusion, and in no case the \emph{neutral}
+code without the atom of confusion.
+In no case the \emph{neutral}
 ratio was higher than the option for the version without atoms of confusion.
@@
 The following quotes were extracted from the transcripts with
-three interviewees:
+% three interviewees:
+two interviewees:
 
@@
 \noindent\begin{minipage}{.45\textwidth}
-\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Ternary Operator} atom)]
+\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Ternary Operator} atom).]
 let config = {size: 3, isActive: false};
@@
 \begin{minipage}{.45\textwidth}
-\begin{lstlisting}[language=JavaScript, caption=\emph{Right-hand side} (without the atom)]
+\begin{lstlisting}[language=JavaScript, caption=\emph{Right-hand side} (without the atom).]
 let config = {size: 3, isActive: false}
@@
 \begin{mq}
-\emph{``I prefer to write [code using the \lhs version], but I think [the \rhs version]  is easier to read, especially for newer programmers''}
+\emph{``I prefer to write [code using the \lhs version], but I think [the \rhs version] is easier to read, especially for newer programmers''}.
 \end{mq}
@@
 \begin{mq}
-\emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the [the code using the \lhs version] is easier to understand''}.
+\emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the 
+[the code using the \lhs version] 
+is easier to understand''}.
 \end{mq}
@@
 as simpler to understand, but would still opt to write
-code with the atom. In contrast to such opinion,
-one of the participants found the version with the atom
-of confusion more elegant, but recognized it was less readable,
+code with the atom. In contrast, %to such opinion,
+one of the participants found the 
+version with the atom of confusion more elegant,
+but recognized it was less readable,
 and was willing to sacrifice elegance for readability.
@@
   and my colleagues complained that it was difficult to understand.
-  Nowadays I prefer to write code using the [\rhs version]."}
+  Nowadays I prefer to write code using the [\rhs version]"}.
 \end{mq}
@@
 often notice during the interviews that
-the \emph{\lhs} (with the atom of confusion) caused significant
+the \emph{\lhs} (with the atom of confusion) 
+caused significant
 confusion among the participants. One remark about
@@
 \begin{mq}
-\emph{``The code in [the \lhs version] is unlikely to be understood unless the programmer knows \clang or \cpplang''}
+\emph{``The code in 
+[the \lhs version]
+is unlikely to be understood unless the programmer knows \clang or \cpplang''}.
 \end{mq}
@@
 \begin{mq}
-\emph{``[I prefer the \rhs version of the code \ldots] If I want to see well-written, easily understandable code, then I also have to do my job.
-Therefore I believe that, since I do not know who is on the other end maintaining this code, and it could be any person with any level of expertise, then I try to write readable, easy-to-understand code''}
+\emph{``[I prefer the \rhs version of the code \ldots]
+If I want to see well-written, easily understandable code, then I also have to do my job.
+Therefore I believe that, since I do not know who is on the other end maintaining this code, and it could be any person with any level of expertise, then I try to write readable, easy-to-understand code''}.
 \end{mq}
 
-This is an interesting perspective, because such a perhaps simple decision,  might hinder novice developers in the task of maintaining code; and one cannot make any assumption about the programmers' experience of who is going to maintain a code in a long run.
+%This is an interesting perspective, because such a perhaps simple decision,  might hinder novice developers in the task of maintaining code; and one cannot make any assumption about the programmers' experience of who is going to maintain a code in a long run.
+
+This is an interesting perspective because such a perhaps simple decision might hinder novice developers in maintaining code; one cannot assume the programmers' experience of who will maintain a code in the long run.
 
@@
       \cmidrule(lr){2-4}
-         \textsc{atom}           & \multicolumn{1}{c}{\textsc{LHS}}
+         Atom           & \multicolumn{1}{c}{\textsc{LHS}}
                                       &  \multicolumn{1}{c}{\textsc{RHS}}
@@
 
-The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confusing code.
+% The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confusing code.
+
+The final remarks drawn from the interviews are related to potentially confusing constructs suggested by the participants and their perspectives on JavaScript as a language.
+From the latter, we can also uncover some other forms in which the language itself might contribute to writing confusing code.
 
@@
 
-When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate.
+% When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate.
+When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern---potentially leading to several levels of nested function calls---as extremely difficult to assimilate.
 One of the respondents stated:
@@
 \begin{small}
-\begin{lstlisting}[language=JavaScript,caption=Example of arrow function,label=arr1]
+\begin{lstlisting}[language=JavaScript,caption=Example of arrow function.,label=arr1]
 let inc = (x) => x + 1;
 \end{lstlisting}
-\begin{lstlisting}[language=JavaScript,caption=Alternative version without arrow function,label=arr2]
+\begin{lstlisting}[language=JavaScript,caption=Alternative version without arrow function.,label=arr2]
 let inc = function(x) {
@@
 frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
-occur in at least 80\% of the projects. Considering the extremes, atom Ternary Operator and Implicit Predicate were found in all repositories we mined,
+occur in at least 80\% of the projects. Considering the extremes, atoms Ternary Operator and Implicit Predicate were found in all repositories we mined,
 while atom Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
@@
 \begin{tabular}{lrcc}
-  \hline
-Atom & Projects & Occurr./KLOC & Category \\ 
-  \hline
+  \toprule
+Atom & Projects & Occurr./KLOC & Frequency \\ 
+  \midrule
 Implicit Predicate & 100\% & 19.89 &  highly used \\ 
@@
   Comma Operator & 15.28\% & 0.03 & little used \\ 
-   \hline
+   \bottomrule
 \end{tabular}
@@
 for 92.97\% of this total: Implicit Predicate, Ternary Operator, Omitted Curly Braces, and the Post Increment. The remaining atoms
-account for \num{25621} (7.03\%) of the total number of occurrences, and the
-Comma Operator and Arithmetic as Logic are the atoms that arise less
+account for \num{25621} (7.03\%) of the total number of occurrences.
+Moreover, the Comma Operator and Arithmetic as Logic atoms are the ones that arise less
 frequently (0.001\% in total with 232 and 165 occurrences, respectively).
@@
 according to the thresholds of a previous
-study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. As such, we found four
+study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. 
+In particular, atoms with less than \num{1000} occurrences are classified as \textit{little used}, while atoms with more than \num{10000} occurrences are classified as \textit{highly used}; \textit{commonly used} atoms occur in between.
+We found four
 atoms as highly used (Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment),
@@
 misunderstanding (increasing the number of wrong answers in 9\% of the cases,
-according to our survey). Nonetheless, the Post-Increment Expression
+according to our survey). 
+Nonetheless, the Post-Increment Expression
 and Omitted Curly Braces atoms are listed in the top four sources of misunderstanding 
@@
 in open source systems. As such, fixing the
-the Post-Increment and Omitted Curly Braces might solve a considerable
-amount unclear code caused by atoms of confusion in
+the Post-Increment and Omitted Curly Braces might improve a considerable
+amount of unclear code caused by atoms of confusion in
 JavaScript projects.

### 2021-05-03 main.tex
Updates from Overleaf

@@
 In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study.
-From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present inseven out of ten atom candidates.
+From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in seven out of ten atom candidates.
 In addition, our MSR effort reveals that atoms of confusion frequently arise in 72 popular open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.

### 2021-05-03 1_Introduction.tex
Updates from Overleaf

@@
 On the right-hand side, we present an alternative implementation without atoms of confusion.
-We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p).length}).}\label{fig:lst01}
+We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p.length})).}\label{fig:lst01}
 \end{figure*}
@@
 
-  \item A list of additional, JavaScript-specific
+  \item A list of additional JavaScript-specific
     constructs and idioms that might introduce code
@@
     code queries (implemented in the .QL language). This library and 
-    infrastructure are freely available and might help researchers and 
+    infrastructure are publicly available and might help researchers and 
     practitioners to automate program analysis at a large scale.

### 2021-05-03 6_Conclusion.tex
Updates from Overleaf

@@
 We asked them to predict the output of some code snippets (with and without atoms).
-We provide evidences that seven atom candidates have a significant impact on program comprehension activities.
+We provide evidence that seven atom candidates have a significant impact on program comprehension activities.
 
@@
 In general, they considered code without atoms of confusion easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding.
-Last but not least, we conducted a mining software repositories study.
-In particular, we found four atoms of confusion highly used in JavaScript code, i.e., Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment.
+%Last, but not least,
+Finally, we conducted a mining software repositories study.
+In particular, we found four atoms of confusion frequently used in JavaScript code, i.e., Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment.
 
@@
 First, we alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment, and Omitted Curly Braces).
-Second, our results might help tool developers to create program transformation tools to remove atoms of confusion automaticallu.
-They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).
+Second, our results might help tool developers to create program transformation tools to remove atoms of confusion automatically.
+%They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).
+Promising candidates may be the atoms we found to be most confusing and, at the same time, common in practice (e.g., Omitted Curly Braces and Post Increment).
 
-As mentioned, the results of the interviews pointed out additional JavaScript constructs and idioms
+%As mentioned, the results of the interviews
+The results of the interviews pointed out additional JavaScript constructs and idioms
 that might introduce misunderstanding. These include not only nested callbacks and prototype-based

### 2021-05-03 discussion.tex
additional review (conclusions)
Updates from Overleaf

@@
 confusion (i.e., a source of misunderstanding).
-In comparison to the previous work, our study
+In comparison to the original
+work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, 
+our study
 led to some differences in the effect size
-of the atom candidates, in comparison to the original
-work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
+of the atom candidates.
 Altogether, we answer our first research question
@@
 \begin{mh}
-  {\bf Answer to RQ1:} The first study (survey) give evidence that
+  {\bf Answer to RQ1:} The first study (survey) gives evidence that
   the atom candidates for \clang and \cpplang
@@
 prefer the version of the code with the Ternary Operator,
-even though they agree that its uses might contribute to the
+even though they agree that its use might contribute to the
 misunderstanding of JavaScript code, particularly when novices are maintaining
@@
 
-The result of the third study (mining open source
+The results of the third study (mining open source
 Java script repositories) give evidence that,
@@
 source JavaScript projects. In particular,
-{\color{red}six, out of nine atoms considered
-  in our study,} appear in more than 50\% of
-the projects. To sum up, the third study
+seven, out of 10 atoms considered
+  in our study, appear in more than 50\% of
+the projects. 
+In summary, the third study
 allows us to answer the fourth research

### 2021-05-03 1_Introduction.tex
minor review

@@
 For instance, consider the pair of JavaScript code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
-The code at the left-hand side contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}.
+The code at the left-hand side contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: Logic as Control Flow and Post-increment.
 At the right-hand side, we present a clean version of the code, that is, without atoms of confusion.
-According to the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in 41\% and 34\%, respectively.
+According to the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms Logic as Control Flow and Post-increment improved the accuracy of program understanding tasks in 41\% and 34\%, respectively.
 
@@
 \captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project. Line 2 on the left-hand side of the figure contains two
-  atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic
-  as Control Flow} and \emph{Post-increment}.
+  atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: Logic
+  as Control Flow and Post-increment.
 On the right-hand side, we present an alternative implementation without atoms of confusion.
@@
 
-However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java).
+Nonetheless, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java).
 This way, there is a knowledge gap with respect to the impact of such atoms in dynamic languages such as JavaScript.

### 2021-05-03 6_Conclusion.tex
final review

@@
 as \emph{property access} and \emph{arrow functions}. As future work, we intend to
-reproduce our study to validate whether or not these \emph{atom candidates} affect
+reproduce our survey to validate whether or not these additional
+\emph{atom candidates}, which the interviewees pointed out, truly affect
 the understanding of JavaScript code.

### 2021-09-24 0_Abstract.tex
Updates from Overleaf

@@
 %Besides lack of experience and other factors, 
-Previous research works have discussed the impact of small code patterns that contribute to misunderstanding the source code written in statically typed languages (e.g., \clang, \cpplang, and Swift).
+Previous research work has discussed the impact of small code patterns that contribute to misunderstanding the source code written in statically typed languages (e.g., \clang, \cpplang, and Swift).
 However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamically typed languages.
@@
 Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
-\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead to a code that is hard to understand: 
+\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: 
 %removing the atoms of confusion lead to an improvement above 15\% in 6 out of 9 atoms we consider in our study when developers try to predict the outcomes of code examples. 
-developers correctly predict the outcome of at least 15\% more when an atom of confusion is not present for 6 out of 9 atoms.
+developers correctly predict the outcome of at least 15\% more when an atom of confusion is not present for 9 out of 10 atoms.
+For 5 of these atoms, the difference is statistically significant (accounting for p-value correction) and effect size is large for 4 and medium for 1 of them.
 Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the \texttt{comma operator} is present in 11\% of the projects in our dataset; while five other atoms appear in more than 60\% of the projects---including the atoms \texttt{assignment as value} and \texttt{omitted curly braces}, which appear \num{8253} and \num{53357} times, respectively.

### 2021-09-24 1_Introduction.tex
Updates from Overleaf

@@
 The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate).
+For 5 of these atoms, the difference is statistically significant (accounting for p-value correction).
+Effect size is large for 4 atoms and medium for 1 of them.
 The presence of an atom candidate might also affect the time necessary for submitting a correct answer.
@@
 The frequency in which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code.
-To understand the incidence of atoms, we mined popular JavaScript projects from GitHub and found that atoms of confusion frequently appear in practice.
-Indeed, four atom candidates are highly used in JavaScript code---the Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.
+To understand the incidence of atoms, we mined popular JavaScript projects from GitHub and found that atoms frequently appear in practice.
+Our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems.
+Four atom candidates appear in 90\% of the analyzed projects and two of them, Implicit Predicate and Ternary Operator, occur more than once for every 100 lines of code in the dataset.
+%Indeed, four atom candidates are frequently used in JavaScript code---the Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.
 

### 2021-09-24 6_Conclusion.tex
Updates from Overleaf

@@
 
-This paper has reported the results of a mixed-method research effort to better understand the impact of atoms of confusion in JavaScript code.
-First, we conducted a survey with 140 JavaScript developers.
-We asked them to predict the output of some code snippets (with and without atoms).
-We provide evidence that seven atom candidates have a significant impact on program comprehension activities.
-
-Afterwards, we interviewed 15 professional developers.
-In general, they considered code without atoms of confusion easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding.
+This paper reports the results of a mixed-method research effort to better understand the impact of atoms of confusion in JavaScript code.
+First, we conducted a survey with 140 JavaScript developers, asking them to predict the output of some code snippets (with and without atoms).
+We provide evidence that five atom candidates significantly impact program comprehension activities and can be considered atoms of confusion: Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate.
+We also interviewed 15 professional developers.
+In general, they considered code without atom candidates easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding.
 %Last, but not least,
 Finally, we conducted a mining software repositories study.
-In particular, we found four atoms of confusion frequently used in JavaScript code, i.e., Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment.
+We found four atom candidates frequently used in JavaScript code, i.e., Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces.
+%The first two can be considered atoms of confusion.
 
 Our efforts have important implications for practice.
-First, we alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment, and Omitted Curly Braces).
-Second, our results might help tool developers to create program transformation tools to remove atoms of confusion automatically.
+First, we alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate).
+Second, our results might help tool developers to create program transformation tools to remove the atoms automatically.
 %They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).
-Promising candidates may be the atoms we found to be most confusing and, at the same time, common in practice (e.g., Omitted Curly Braces and Post Increment).
+%Promising candidates may be the atoms we found to be most confusing and, at the same time, common in practice (e.g., Implicit Predicate and Post Increment).
 
@@
 The results of the interviews pointed out additional JavaScript constructs and idioms 
-that might introduce misunderstanding. These include not only nested callbacks and prototype-based
-inheritance (which might not have a counterpart version), but also idioms such
-as \emph{property access} and \emph{arrow functions}. As future work, we intend to
+that might introduce misunderstanding. These include nested callbacks, % and prototype-based inheritance (which might not have a counterpart version), but also idioms such
+%as 
+\emph{property access}, and \emph{arrow functions}. As future work, we intend to
 reproduce our survey to validate whether or not these additional
-\emph{atom candidates}, which the interviewees pointed out, truly affect
+\emph{atom candidates}
+%, which the interviewees pointed out, 
+truly affect
 the understanding of JavaScript code.

### 2021-09-24 discussion.tex
Updates from Overleaf

@@
 
-Our work leads to several implications.
-First, we generalize the
+Our work has several implications.
+First, it adds external validity to the
 work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17},
@@
 Assignment as Value, Implicit Predicate, Logic as
-Control Flow, and Ternary Operator make the 
-JavaScript code hard to understand. Our results also 
-refute the hypothesis that Arithmetic as Logic is an atom of
-confusion (i.e., a source of misunderstanding).
+Control Flow, and Ternary Operator seem to make 
+JavaScript code harder to understand. For five of them, the difference is statistically significant, with a large effect size for four atoms. Our results also 
+refute the hypothesis that Arithmetic as Logic is an atom of confusion (i.e., a source of misunderstanding).
 In comparison to the original
@@
 \begin{mh}
-  {\bf Answer to RQ1:} The first study (survey) gives evidence that
-  the atom candidates for \clang and \cpplang
-  that also exist for JavaScript correspond to
-  a source of misunderstanding in
+  {\bf Answer to RQ1:} The first study (survey) gives evidence that some of the atom candidates in \clang and \cpplang programs that also exist in JavaScript correspond to a source of misunderstanding in
   JavaScript code. 
@@
 
-The results of the interview study complement the understanding
-of atoms of confusion because the participants make
-clear the existence of a trade-off between code comprehension
-and other quality attributes. For instance, most of the participants
-prefer the version of the code with the Ternary Operator,
-even though they agree that its use might contribute to the
-misunderstanding of JavaScript code, particularly when novices are maintaining
-the codebase. The participants of the interview study also
+The results of the interview study complement the understanding of atoms of confusion because the participants make clear the existence of a trade-off between code comprehension and other quality attributes.
+For instance, most of the participants prefer the version of the code with the Ternary Operator, even though they agree that its use might contribute to the misunderstanding of JavaScript code, particularly when novices are maintaining the codebase.
+The participants of the interview study also
 mentioned other possible sources of misunderstanding in JavaScript,
-including the use of prototype-inheritance and nested call-backs (as discussed in Section~\ref{sec:interview-results}). Other JavaScript atom candidates include
+including the use of prototype-based inheritance and nested call-backs (as discussed in Section~\ref{sec:interview-results}). Other JavaScript atom candidates include
 Object Destructuring, Array Spread, Object Spread, and Type Conversion.
@@
   {\bf Answer to RQ3:} The qualitative analysis of the
-  interviews suggest that specific constructs of
-  JavaScript might also correspond to atoms of
+  interviews suggests that specific JavaScript constructs might also correspond to atoms of
   confusion, including prototype inheritance and
@@
 The results of the third study (mining open source
-Java script repositories) give evidence that,
+Java script repositories) provides evidence that,
 although atoms of confusion compromise program
@@
   in our study, appear in more than 50\% of
-the projects. 
-In summary, the third study
+the 72 projects we analyzed. Furthermore, at least two of them are used intensively, more than once for every 200 lines of code. In summary, the third study
 allows us to answer the fourth research

### 2021-09-24 main.tex
Updates from Overleaf

--- 2021-04-29 main.tex
@@
 In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study.
-From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present inseven out of ten atom candidates.
-In addition, our MSR effort reveals that atoms of confusion frequently arise in 72 popular open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
+From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: considering 10 atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates.
+For five of these atoms, the difference is statistically significant (accounting for p-value correction).
+Effect size is large for four atoms and medium for one of them.
+In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems.
+Four atom candidates appear in 90\% of the analyzed projects and two of them occur more than once for every 100 lines of code in the dataset.
+  %: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
   Altogether, our findings might help practitioners (a) better understand the implications of atoms of confusion on understanding JavaScript code, (b) avoid writing code that is unnecessarily difficult to maintain, and (c) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.

### 2021-09-27 1_Introduction.tex
small changes

--- 2021-05-03 1_Introduction.tex
@@
 Besides that, programming languages can also have specific idioms and constructs that may lead to misunderstandings.
-That is the case of \emph{atoms of confusion}---small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.
+That is the case of atoms of confusion: small pieces of code that prevent developers from program comprehension, even though functionally equivalent alternatives exist and are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.
 
@@
 
-    \item A mixed-method research effort---based on a survey, interviews, and mining software repositories---about the impact of atoms of confusion on a
-    dynamic language;
+\item A mixed-method research effort---based on a survey, interviews, and mining software
+  repositories---about the impact of atoms of confusion on JavaScript (a
+    dynamic language).
 
-  \item A list of additional JavaScript-specific
-    constructs and idioms that might introduce code
-    misunderstanding;
+  \item A comparison of the impact and prevalence of atoms of confusion in
+    JavaScript with what has been previously reported for other languages
+    (\clang, \cpplang, and Java).
 

### 2021-09-27 dataset/survey-results-final.Rmd
update the survey + spell checker
small changes

@@
 ---
-title: "survey-analysis-final"
+title: "Survey Analysis"
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 ---
@@
 setwd(".")
-
 ```
 
-## Load the datasets
+## (1) Load the datasets
 
@@
 atomId <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 0) 
-atomDescription <- c("Arithmetic as Logic", "Assignment as Value", "Automatic Semicolon Insertion",
+atomDescription <- c("Arithmetic as Logic", "Assignment as Value", 
+            "Automatic Semicolon Insertion",
            "Comma Operator", "Ternary Operator", "Implict Predicate", 
@@
 
-## Filter the datasets
+## (2) Filter the datasets
 
@@
 
-## Exploratory Data Analysis
+## (3) Exploratory Data Analysis
 
@@
 educationLevelIds <- c(1, 2, 3, 4, 5)
-educationLevelLabels <- c("High school degree or equivalent", "Some university course but not a degree", "Bacharelor degree", "Master degree", "Doctor degree")
+educationLevelLabels <- c("High school degree or equivalent", 
+                          "Some university course but not a degree", 
+                          "Bacharelor degree", "Master degree", 
+                          "Doctor degree")
 
@@
 demEducation
-barplot(demEducation$total, col=topo.colors(5), ylim = c(0, 70), cex=0.7, cex.lab = 0.7, cex.axis=0.7, ylab="Number of participants")
+
+barplot(demEducation$total, col=topo.colors(5), 
+        ylim = c(0, 70), cex=0.7, cex.lab = 0.7, cex.axis=0.7, 
+        ylab="Number of participants")
 
@@
 experienceLevelIds <- c(1, 2, 3, 4)
-experienceLevelLabels <- c("Under one year of experience", "One year and under four years of experience", "Four years and under ten years of experience", "More than ten years of experience")
+experienceLevelLabels <- c("Under one year of experience", 
+                           "One year and under four years of experience", 
+                           "Four years and under ten years of experience",
+                           "More than ten years of experience")
 
@@
 tmp <- sqldf("select distinct sid, experience from ds")
-demExperience <- sqldf("select experience, count(*) total from tmp group by experience order by 1")
+demExperience <- sqldf("select experience, count(*) total 
+                       from tmp group by experience order by 1")
 
@@
 
-barplot(demExperience$total, col=topo.colors(4), ylim = c(0, 60), cex=0.7, cex.lab = 0.7, cex.axis=0.7, ylab="Number of participants")
+barplot(demExperience$total, col=topo.colors(4), 
+        ylim = c(0, 60), cex=0.7, cex.lab = 0.7, cex.axis=0.7, 
+        ylab="Number of participants")
 
@@
 merged <- sqldf("select c.atomDescription as Atom, 
-                        a.confuseCode as 'Confuse Code', b.cleanCode as 'Clean Code',  
+                        a.confuseCode as 'Confuse Code', 
+                        b.cleanCode as 'Clean Code',  
                         (b.cleanCode * 100 / a.confuseCode) -100 as 'Delta (%)'
@@
 merged <- sqldf("select c.atomDescription as Atom, 
-                        a.timeConfuseCode as 'Confuse Code', b.timeCleanCode as 'Clean Code',
+                        a.timeConfuseCode as 'Confuse Code', 
+                        b.timeCleanCode as 'Clean Code',
                         (b.timeCleanCode * 100 / a.timeConfuseCode) -100 as 'Delta (%)'
@@
 
-## Regression Analysis (correctness~atom + experience + education)
+## (4) Regression Analysis (correctness~atom + experience + education)
 
@@
 
-## Impact considering Developer Experience
+### Impact considering Developer Experience
 
@@
 
-## Impact considering Developer Education
+### Impact considering Developer Education
 
@@
 
-## Impact of Individual Atoms
+## Impact of Individual Atoms and Hypotheses Testing
 
@@
 }
-analysis_df <- data.frame(atomDescription, chiTest, oddsRatio, oddsRatioTest, ci25, ci975, mannWhitneyTest, cliffDelta)
+analysis_df <- data.frame(atomDescription, chiTest, oddsRatio, 
+                          oddsRatioTest, ci25, ci975, mannWhitneyTest, cliffDelta)
+
 analysis_df <- analysis_df[order(-oddsRatio),]
-colnames(analysis_df) <- c("Atom", "ChiTest", "Odds Ratio Correctness", "p-value", "CI 2.5%", "CI 97.5%", "Wilcox Test (Time)", "Cliff Delta")
-analysis_df
+
+colnames(analysis_df) <- c("Atom", "ChiTest", 
+                           "Odds Ratio Correctness", 
+                           "p-value", "CI 2.5%", "CI 97.5%", 
+                           "Wilcox Test (Time)", 
+                           "Cliff Delta")
+
 xtable(analysis_df)

### 2021-09-27 4_Result.tex
Updates from Overleaf
final review

@@
 We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
-and \emph{cognitive effort} (time necessary to provide a correct answer). 
+and \emph{time} (how long to provide a correct answer).
 We received full answers to our survey from 140 participants (a total of 70 replicas).
@@
 The participants should provide the expected outcomes of the code snippets.
-As discussed, we collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{cognitive workload} (the time taken to answer the question).
+As discussed, we collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{time}.
 Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey w.r.t. correctness.
@@
 \begin{table}[htbp]
-\caption{Total number of correct answers}
+\caption{Summary of the correctness analysis}
 \label{tab:difference-correctness}
@@
 
-\begin{figure}[htb!]
+\begin{figure}[b!]
 \noindent
+\vspace*{-1.4cm}
  \centering
  \includegraphics[width=\columnwidth]{images/wrong-answers-plot-1.pdf}
+ \vspace*{-1.3cm}
  \caption{Number of wrong answers of each subject.}
@@
 This indicates that, at a 95\% confidence level, a developer is likely to commit less mistakes when using the clean versions.
+%Considering only the atoms for which there was a significant difference in correctness for confusing and clean versions and u
+Using the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), for four of the atoms, Comma Operator, Automatic Semicolon Insertion, Assignment as Value, and Implicit Predicate, effect size can be considered large.
+For Implicit Predicate, it is medium.
 %We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. 
@@
 \begin{mh}
-  The results of our exploratory data analysis and
-  hypothesis testing suggest that the atom
-  candidates Comma Operator, Automatic Semicolon Insertion, Post Increment, Omitted Curly Braces, 
-  Assignment as Value, and
-  Implicit Predicate %, and Ternary Operator
-  introduce some degree of misunderstanding
-  in JavaScript code. %Based on the results of this study, 
+  The results of our exploratory data analysis and   hypothesis testing suggest that the atom candidates Comma Operator, Automatic Semicolon Insertion, Post Increment, Omitted Curly Braces, Assignment as Value, and Implicit Predicate %, and Ternary Operator
+  introduce some degree of misunderstanding in JavaScript code. %Based on the results of this study, 
   Therefore, they can be deemed atoms of confusion for JavaScript.
@@
 
-\subsection{Cognitive Workload Analysis}
+\subsection{Time Analysis}
 
@@
 Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both confusing and clean versions.
-Seven atom candidates require an increasing cognitive workload.
+We do not consider cases where the participants made mistakes.
+Seven atom candidates required more time from participants to predict a correct response.
 For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet.
@@
 \caption{Time in seconds to submit a correct answer}
-\label{tab:difference-time-taken}
 \centering{
@@
 \end{tabular}
+\label{tab:difference-time-taken}
 \end{scriptsize}
@@
  \begin{table*}[ht]
-\caption{Hypotheses Testing (misunderstanding and cognitive workload)}
-\label{tab:hypothesis-testing}
+\caption{Hypotheses Testing (misunderstanding and time)}
 
@@
  \end{tabular}
+ \label{tab:hypothesis-testing}
  \end{table*}
@@
 of the code. The results of Table~\ref{tab:difference-time-taken}
-suggest that we should refute the null hypotheses for
-six atom candidates (Comma Operator, Automatic Semicolon Insertion,
-Implicit Predicate, Logic
-as Control Flow, Pre-Increment, and Arithmetic
-as Logic). For the first four, the analysis suggests that the participants
-need more time to predict the outcome of
-the code snippet in the confusing code version. Interestingly, for the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time
-to predict the outcome of the code snippets in
-the confusing version.
-We also computed the effect size using the Cliff's Delta method. We found a large effect for the Comma Operator atom candidate; and a moderate effect for the atom candidates
-Logic as Control Flow, Implicit Predicate, Arithmetic
-as Logic, and Pre-Increment.
+suggests that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
+For the first three, the analysis suggests that the participants need more time to predict the outcome of the code snippet in the confusing code version.
+For the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time to predict the outcome of the code snippets in the confusing version.
+We also computed the effect size using Cliff's Delta statistic.
+In the table, negative effect sizes suggest that it took less time to correctly predict the output of clean versions of the snippets.
+Considering the thresholds established by Romano et al.~\cite{Romano:2006:ASO} (0.147 $<$ $|$d$|$ $<$ 0.33 small, $|$d$|$ $<$ 0.474 medium, otherwise large), we found a large effect for the Comma Operator atom candidate; a medium effect size for Logic as Control Flow, and small effect sizes for Automatic Semicolon Insertion, Implicit Predicate, Arithmetic as Logic, Post-, and Pre-Increment.
 
 \begin{mh}
-  The results suggest that developers spend
-  significantly more time to predict the
-  correct answer in the presence of the
-  atom candidates Comma Operator, Automatic Semicolon Insertion, Logic as
-  Control Flow, and Implicit Predicate. Conversely,
-  the code snippets with the
-  atom candidates Arithmetic as Logic and
-  Pre-Increment demand less time to
-  predict the correct answers. 
+  The results suggest that developers tend to spend
+  more time to predict the correct answer in the presence of the atom candidates Comma Operator, Logic as Control Flow, and Implicit Predicate.
+Conversely, the code snippets with the atom candidates Arithmetic as Logic and Pre-Increment demand less time to predict the correct answers.
 \end{mh}
@@
 
+\vspace*{-0.5cm}
 \section{Results of the Interview Study}
@@
 In this section, we present the results of the 
-interviews with the 15 practitioners who agreed to participate
-in the study. 
+interviews with the 15 practitioners who agreed to participate in the study. 
 % A total of 15 practitioners took part of the interviews.
 We collected information regarding programming
-experience, familiarity with JavaScript, and their opinion
-about the \na atom candidates we explored in our research.
-We first present them pairs of clean and confusing code snippets 
-and then ask them to discuss their preferences towards one version.
+experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in our research.
+We first presented them pairs of clean and confusing code snippets and then asked them to discuss their preferences towards one version.
+We did not indicate in any way whether any version was assumed to be confusing or not.
 % ask the participants to discuss their preferences (i.e., code with or without atom).
@@
 
+\begin{table}[!htb]
+    \centering
+    \caption{Summary of participants' preferences for code snippets \emph{with} and
+      \emph{without} atom candidates. The participants were only presented the code snippets, without any indication about whether one was confusing or not. 
+      Participants were also allowed to choose Neutral when they thought both sides were equally readable.}
+    \begin{tabular}{lrrr}\toprule
+      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
+      \cmidrule(lr){2-4}
+         Atom           & \multicolumn{1}{c}{Confusing}
+                                      &  \multicolumn{1}{c}{Clean}
+                                               & \multicolumn{1}{c}{Neutral} \\ \midrule
+         Comma Operator                  & 0  & 100    & 0     \\
+         Automatic Semicolon Insertion  & 0  & 80     & 0     \\
+         Post-Increment                  & 20 & 73.33  & 6.67  \\
+         Omitted Curly Braces            & 0  & 100    & 0     \\
+         Assignment as Value             & 20 & 60     & 20    \\
+         Implicit Predicate              & 20 & 73.33  & 6.67  \\
+         Logic as Control Flow           & 20 & 60     & 20    \\
+         Ternary Operator                & 60 & 26.67  & 13.3  \\
+         Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
+         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \midrule
+         \textsc{overall}                & 18 & 71.33  & 8.64  \\
+         \bottomrule
+    \end{tabular}
+    \label{tab:interview-results1}
+\end{table}
+
 \subsubsection*{The Participants' perceptions of the atom candidates}
@@
 10 scenarios surveyed, the majority of the respondents prefer the version of the
-code without the atom of confusion.
+code without the atom candidate.
 In no case the \emph{neutral}
-ratio was higher than the option for the version without atoms of confusion.
+ratio was higher than the option for the Clean version.
 %% An example entry is contained in Appendix~\ref{}. 
 %% For a full listing of the code snippets, visit the paper repository. 
-Only for the Ternary Operator atom candidate the participants
-prefer the version of the code with the atom of confusion,
-instead of the version without the atom of confusion.
+Only for the Ternary Operator atom candidate the participants prefer the Confusing version, instead of the Clean one.
 We show the code snippets for the Ternary Operator in Figure~\ref{code:ternary}.
-Even for this atom candidate, some participants who preferred 
-the \emph{left-hand-side} version (with the atom of confusion) 
-still believed that the \emph{right-hand-side} version (without the atom of
-confusion) was more readable.  
+Even for this atom candidate, some participants who preferred the left-hand side version (confusing) still believed that the right-hand side version (clean) was more readable.
 The following quotes were extracted from the transcripts with
@@
 \end{minipage}
-\caption{Example of pair of code snippets used in the interview. This case explores the use of the \emph{Ternary Operator atom of confusion}}
+\caption{Example of a pair of code snippets used in the interview. This case explores the use of the \emph{Ternary Operator atom of confusion}}
 \label{code:ternary}
+\vspace{-0.2cm}
 \end{figure*}
@@
 \emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the 
-[the code using the \lhs version] 
+[the code using the \rhs version] 
 is easier to understand''}.
@@
 The Pre-Increment atom candidate also caused such
-divide in one of the interviewees, as they
-regarded the \emph{\rhs version} (without atom of confusion)
-as simpler to understand, but would still opt to write
+divide in one of the interviewees, who
+regarded the clean version as simpler to understand, but would still opted to write
 code with the atom. In contrast, %to such opinion,
-one of the participants found the 
-version with the atom of confusion more elegant,
-but recognized it was less readable,
-and was willing to sacrifice elegance for readability.
+one of the participants found the version with the atom candidate more elegant, but recognized it was less readable, and was willing to sacrifice elegance for readability.
 
@@
 
-When analyzing the Logic as Control Flow atom candidate, one of
-the interviewees gave an example of his own experience
-that might motivate one to avoid writing code using this construct:
+When analyzing the Logic as Control Flow atom candidate, one of the interviewees gave an example personal experience that might motivate one to avoid writing code using this construct:
 
 \begin{mq}
-  \emph{``This one is interesting, because I have written
-  code that looks like the [\lhs version (with the atom of confusion)],
-  and my colleagues complained that it was difficult to understand.
-  Nowadays I prefer to write code using the [\rhs version]"}.
+  \emph{``This one is interesting, because I have written code that looks like the left-hand side [Confusing version], and my colleagues complained that it was difficult to understand.
+Nowadays I prefer to write code using the version on the right-hand side [Clean]''}.
 \end{mq}
 
-Two of the atom candidates had unanimous preference for the
-versions that did not contain the atoms: Comma Operator and
-Omitted Curly Braces. Regarding the first one, we could
-often notice during the interviews that
-the \emph{\lhs} (with the atom of confusion) 
-caused significant
-confusion among the participants. One remark about
-the \emph{Comma Operator} atom of confusion is listed below:
+For two of the atom candidates, the interviewees were unanimous in their preference for the clean version: Comma Operator and Omitted Curly Braces.
+Regarding the first one, we could often notice during the interviews that the \emph{\lhs} (with the atom of confusion) caused significant confusion among the participants.
+One remark about the \emph{Comma Operator} atom of confusion is listed below:
 
@@
 \begin{mq}
-\emph{``The code in 
-[the \lhs version]
-is unlikely to be understood unless the programmer knows \clang or \cpplang''}.
+\emph{``The code in [the \lhs version] is unlikely to be understood unless the programmer knows \clang or \cpplang''}.
 \end{mq}
 
-As for the atom candidate Omitted Curly Braces, one of the interviewees mentioned that, although they understand why one would opt not to use braces for simple if-then-else statements, he still advised against it, on grounds that:
+As for the atom candidate Omitted Curly Braces, one of the interviewees mentioned that, although they understand why one would opt not to use braces for simple if-then-else statements, they still advised against it, on grounds that:
 
@@
 
-This is an interesting perspective because such a perhaps simple decision might hinder novice developers in maintaining code; one cannot assume the programmers' experience of who will maintain a code in the long run.
-
-\begin{table}[!htb]
-    \centering
-    \caption{Summary of participants' preferences for code snippets \emph{with} (LHS) and
-      \emph{without} atoms of confusion (RHS).
-      Participants were also allowed to choose NEUTRAL when they thought both sides were equally readable.}
-    \label{tab:interview-results1}
-    \begin{tabular}{lrrr}\toprule
-      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
-      \cmidrule(lr){2-4}
-         Atom           & \multicolumn{1}{c}{\textsc{LHS}}
-                                      &  \multicolumn{1}{c}{\textsc{RHS}}
-                                               & \multicolumn{1}{c}{NEUTRAL} \\ \midrule
-         Comma Operator                  & 0  & 100    & 0     \\
-         Automatic Semicolon Insertion  & 0  & 80     & 0     \\
-         Post-Increment                  & 20 & 73.33  & 6.67  \\
-         Omitted Curly Braces            & 0  & 100    & 0     \\
-         Assignment as Value             & 20 & 60     & 20    \\
-         Implicit Predicate              & 20 & 73.33  & 6.67  \\
-         Logic as Control Flow           & 20 & 60     & 20    \\
-         Ternary Operator                & 60 & 26.67  & 13.3  \\
-         Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
-         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \midrule
-         \textsc{overall}                & 18 & 71.33  & 8.64  \\
-         \bottomrule
-    \end{tabular}
-\end{table}
+%This is an interesting perspective because such a simple decision might hinder novice developers in maintaining code; one cannot assume the programmers' experience of who will maintain a code in the long run.
 
@@
 \end{lstlisting}
+\vspace*{-0.6cm}
 \end{small}
@@
 frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
-occur in at least 80\% of the projects. Considering the extremes, atoms Ternary Operator and Implicit Predicate were found in all repositories we mined,
-while atom Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
+occur in at least 80\% of the projects. Considering the extremes, atom candidates Ternary Operator and Implicit Predicate were found in all repositories we mined,
+while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
 
@@
 \centering
-\caption{Summary of atoms occurrences on our dataset}
+\caption{Summary of atom candidate occurrences on our dataset.}
 \setlength\tabcolsep{2pt} % default value: 6pt
-\label{tab:occurrences-summary}
-\begin{tabular}{lrcc}
+\begin{tabular}{lrc}%c}
   \toprule
-Atom & Projects & Occurr./KLOC & Frequency \\ 
+Atom & Projects & Occurr./KLOC \\%& Frequency \\ 
   \midrule
-Implicit Predicate & 100\% & 19.89 &  highly used \\ 
-  Ternary Operator & 100\% & 10.16 &  highly used \\ 
-  Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
-  Post Increment & 90.28\% & 5.62 &  highly used \\ 
-  Pre Increment & 83.33\% & 1.04 & commonly used \\ 
-  Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
-  Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
-  Automatic Semicolon Insertion & 33.33\% & 0.13 & commonly used \\ 
-  Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
-  Comma Operator & 15.28\% & 0.03 & little used \\ 
+Implicit Predicate & 100\% & 19.89 \\%&  intensively used \\ 
+  Ternary Operator & 100\% & 10.16 \\%&  intensively used \\ 
+  Omitted Curly Braces & 91.67\% & 6.61 \\%&  intensively used \\ 
+  Post Increment & 90.28\% & 5.62 \\%&  intensively used \\ 
+  Pre Increment & 83.33\% & 1.04 \\%& commonly used \\ 
+  Assignment as Value & 81.94\% & 1.02 \\%& commonly used \\ 
+  Logic as Control Flow & 56.94\% & 0.95 \\%& commonly used \\ 
+  Automatic Semicolon Insertion & 33.33\% & 0.13 \\%& commonly used \\ 
+  Arithmetic as Logic & 23.61\% & 0.02 \\%& rarely used \\ 
+  Comma Operator & 15.28\% & 0.03 \\%& rarely used \\
    \bottomrule
 \end{tabular}
+\label{tab:occurrences-summary}
 \end{table}
 
-Considering all JavaScript projects in our dataset, we found 
-a total of \num{364873} atoms of confusion, though four atoms are responsible
-for 92.97\% of this total: Implicit Predicate, Ternary Operator, Omitted Curly Braces, and the Post Increment. The remaining atoms
-account for \num{25621} (7.03\%) of the total number of occurrences.
-Moreover, the Comma Operator and Arithmetic as Logic atoms are the ones that arise less
+Considering all JavaScript projects in our dataset, we found a total of \num{364873} atom candidates, though four atoms are responsible
+for 92.97\% of this total: Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces.
+The first two, based on the results of Section~\ref{sec:survey-resuts}, can be considered atoms of confusion.
+The remaining atom
+candidates comprise \num{25621} occurrences combined, 7.03\% of the total number of occurrences.
+The Comma Operator and Arithmetic as Logic atoms are the ones that arise less
 frequently (0.001\% in total with 232 and 165 occurrences, respectively).
 
-Regarding frequency, we proceeded with a classification of \emph{atom occurrence}
-according to the thresholds of a previous
-study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. 
-In particular, atoms with less than \num{1000} occurrences are classified as \textit{little used}, while atoms with more than \num{10000} occurrences are classified as \textit{highly used}; \textit{commonly used} atoms occur in between.
-We found four
-atoms as highly used (Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment),
-four as commonly used (Pre-Increment, Assignment as Value,
-Logic as Control Flow, and Automatic Semicolon Insertion),
-and two atoms as little used (Arithmetic as Logic and Comma Operator).
-We also calculated the frequency of each atom per KLOC and found the atoms
-Implicit Predicate (19.89 occurrences per KLOC) and Ternary Operator (10.16
-occurrences per KLOC) as the most occurring
-in practice.
+% Regarding frequency, we proceeded with a classification of \emph{atom occurrence}
+% based on thresholds used in a previous
+% study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. 
+% Atoms that occur 1000 times are classified as \textit{rarely used} while atoms with more than \num{10000} occurrences are classified as \textit{highly used}; \textit{commonly used} atoms occur in between.
+We calculated the frequency of each atom per KLOC.
+The four atoms mentioned in the previous paragraph,
+ Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces, occur frequently in the analyzed projects.
+All of them have more than 5 occurrences per KLOC on average.
+In particular, two atom candidates occur very frequently, Implicit Predicate (19.89 occurrences per KLOC) and Ternary Operator (10.16 occurrences per KLOC).
+On the other hand, Arithmetic as Logic and Comma Operator occur less than once for every 50 KLOC.
+
 %In our dataset of JavaScript projects, 
@@
 
-The results of our survey and interviews suggest that the
-Ternary Operator does not contribute significantly as a source of
-misunderstanding (increasing the number of wrong answers in 9\% of the cases,
-according to our survey). 
-Nonetheless, the Post-Increment Expression
-and Omitted Curly Braces atoms are listed in the top four sources of misunderstanding 
-(see Table~\ref{tab:difference-correctness}), and also frequently appear
-in open source systems. As such, fixing the
-the Post-Increment and Omitted Curly Braces might improve a considerable
-amount of unclear code caused by atoms of confusion in
-JavaScript projects.
+The results of our survey and interviews suggest that the Ternary Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to our survey).
+Nonetheless, the Post-Increment Expression and Automatic Semicolon Insertion atoms are listed in the top four sources of misunderstanding  (Table~\ref{tab:difference-correctness}), and also frequently appear in open source systems.
+As such, fixing the the Post-Increment and Automatic Semicolon Insertion might improve the readabiliy of large amounts of code in JavaScript projects.

### 2021-09-27 5_Threats.tex

@@
 
-\textbf{Conclusion validity.} In the context of our survey study, we apply different statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test).
+\textbf{Conclusion validity.} 
+%Threats to conclusion validity often come from inappropriate use of statistics. 
+In the context of our survey study, we apply different non-parametric statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test).
 Furthermore, besides reporting p-values, we also report effect size measures appropriate for each scenario (odds-ratio and Cliff's delta) and apply a p-value correction technique to avoid the multiple testing problem.
@@
 To that end, we use the $\phi$ measure of effect size for each atom (based on the Chi-square statistic), the standard $\alpha$ coefficient of 0.05, and set the expected statistical power to 0.8, as usually employed in the literature~\cite{Ellis:2010:EGE}.
-As a result of this analysis, we discover that for all but one of the atoms where we had a statistically significant result for correctness, Implicit Predicate, the sample size is sufficient.
+We find out that the sample size is sufficient for all but one of the atoms where we had a statistically significant result for correctness, Implicit Predicate.
 % For the ones where we did not find statistically significant results for correctness, in fact the sample sizes
@@
 
-%One threat to validity we have not properly addressed is lack of power, a common problem in Software Engineering research.
-
 %Construct validity is connected with how well the selected measures actually represent the concept of interest.
 
-\textbf{Construct validity.} We use correctness as a proxy for program comprehension.
-Furthermore, we use predicted outcomes of small programs as a measure of correctness.
+\textbf{Construct validity.} We use correctness as a proxy for program comprehension and predicted outcomes of small programs as a measure of correctness.
 As discussed elsewhere~\cite{Oliveira:2020:ECR}, this approach is a test of the developers' ability to trace programs.
@@
 
-\textbf{External validity.} Although our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, it is not clear whether that result extrapolates to other scenarios.
+\textbf{External validity.} Our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, but it is not clear if that result extrapolates to other scenarios.
 Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect.

### 2021-10-06 3_Method.tex
Updates from Overleaf
final review
rephrase one of the interview questions

--- 2021-04-30 3_Method.tex
@@
 Since JavaScript and C have some constructs in common (Section~\ref{back}), we first selected a set of 9+1 atom candidates: 9 atoms of confusion already validated~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} for the \clang language that also exist in JavaScript programs + 1 atom candidate that is specific of the JavaScript language (Automatic Semicolon Insertion).
-Our supplementary material discusses and shows examples of
-the atoms we consider in our research.
+Our supplementary material discusses and shows examples of the atoms we consider in our research.
 
@@
 On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes.
-
-%In the interviews we presented developers with two functionally equivalent short programs, one with a single atom and another one without the atom, differed concerning program comprehension.
-We then asked them to tell us which program was easier to understand.
-%The main goal of the interviews was to have developers assess whether two functionally equivalent short programs, one with a single atom and another one without the atom, differed concerning program comprehension, and, if so, which version they regarded as the easier one to understand.
-
 Two of the researchers conducted the interviews, and a third one listened to all the recordings to cross-validate the collected data.
@@
   \centering
+  \caption{Demographic information of the participants}
 \begin{scriptsize}  
@@
 \end{scriptsize}
-    \caption{Demographic information of the participants}
     \label{pinterview}
@@
 This would allow us to identify potential atom candidates that are more specific to the JavaScript language.
-Examples of questions we explored in this section include: \emph{Do you think JavaScript is a language that produces code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as
+Examples of questions we explored in this section include: \emph{Does JavaScript favor developers to produce code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as
   especially confusing?}
@@
   \centering
+   \caption{Some descriptive statistics about the projects used in the MSR study}
   \begin{scriptsize}
@@
  \end{scriptsize}
- \caption{Some descriptive statistics about the projects used in the MSR study}
  \label{tab:projects-statistics} 

### 2021-10-14 4_Result.tex
minor fix

@@
 This is the case for the four atom candidates in the lower part of the table.
-In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a study subject was actually more likely to misunderstand the clean version.
+In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a participant was actually more likely to misunderstand the clean version.
 

### 2021-10-16 4_Result.tex
passada inicial na secao de resultados

@@
 
-Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant \emph{education} or participant \emph{experience} impact correctness.
-Although the results suggest that experience impacts correctness, the education level does not.
-As such, we replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.
+{\color{blue}Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant \emph{education} or participant \emph{experience} impact correctness.
+The findings suggest that \emph{developer experience} impacts on the results related to correctness (see Figure~\ref{fig:correctness-over-experience}).
+Even though the presence of atoms of confusion reduces the number of correct answers to all developer experience groups, this effect is not uniform.
+For instance, novice developers provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers when evaluating the confusing version of the code snippets.
+It is also important to note that, independently of the developer experience group, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.} {\color{red}We replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.} Differently from developer experience, the factor \emph{education level} does not change substantially the correctness of the answers.
+
+\begin{figure}
+  \includegraphics[scale=0.35]{images/correctness-by-experience}
+  \caption{Impact of atoms of confusion on the correctness (over the developer experience groups)}
+  \label{fig:correctness-over-experience}
+\end{figure}
 
@@
 In no case the \emph{neutral}
-ratio was higher than the option for the Clean version.
+ratio was higher than the option for the clean version.
 %% An example entry is contained in Appendix~\ref{}. 
 %% For a full listing of the code snippets, visit the paper repository. 
-Only for the Ternary Operator atom candidate the participants prefer the Confusing version, instead of the Clean one.
+Only for the Ternary Operator atom candidate the participants prefer the Confusing version, instead of the clean one.
 We show the code snippets for the Ternary Operator in Figure~\ref{code:ternary}.
@@
 This is a core feature of the language, and most high-level tools and frameworks abstract it away.
-Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.
-
+%Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.
 % When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate.
@@
 As such, fixing the the Post-Increment and Automatic Semicolon Insertion might improve the readabiliy of large amounts of code in JavaScript projects.
+
+\todo[inline]{Seria legal contrastar esses achados com o que foi reportado anteriormente}

### 2021-10-20 4_Result.tex
comparacao com trabalhos anteriores

--- 2021-10-14 4_Result.tex
@@
   Logic as Control Flow         &  41 &  48 &  +17   \\ 
-  Ternary Operator              &  60 &  66 &  +10   \\ 
+  Conditional Operator              &  60 &  66 &  +10   \\ 
   Pre-Increment                 &  50 &  53 &  +6    \\ 
@@
 
-Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant \emph{education} or participant \emph{experience} impact correctness.
-Although the results suggest that experience impacts correctness, the education level does not.
-As such, we replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.
+{\color{blue}Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impact correctness.
+The findings suggest that \emph{participant experience} impacts on the results related to correctness (see Figure~\ref{fig:correctness-over-experience}).
+Even though the presence of atoms of confusion reduces the number of correct answers to all \emph{experience groups}, this effect is not uniform.
+For instance, novice developers provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets.
+It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.} {\color{red}We replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.} Differently from \emph{participant experience}, the factor \emph{participant education} does not change substantially the correctness of the answers.
+
+\begin{figure}
+  \includegraphics[scale=0.3]{images/correctness-by-experience}
+  \caption{Impact of atoms of confusion on the correctness (over the participant experience groups)}
+  \label{fig:correctness-over-experience}
+\end{figure}
 
@@
 
+{\color{red}Comparing to previous research, three 
+atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed
+for \clang and \cpplang do not lead to a statistically significant impact
+on correctness in our study: Logic as Control Flow,
+Conditional Operator, and Pre-Increment. Our results
+support the Gopstein et al. findings for the remaining
+atoms. Considering the work of Langhout and Aniche~\cite{Langhout:2021:ACJ},
+their work investigates six atom candidates for Java that
+we also explore here for JavaScript. Among these atom
+candidates, we achieve similar conclusions with their
+study for four atom candidates: Post Increment,
+Conditional Operator, Arithmetic as Logic, and
+Omitted Curly Braces---both studies confirmed only the former
+as an atom of confusion. Our findings diverged from
+the work of Langhout and Aniche for two atoms: Pre Increment and
+Logic as Control Flow.}
+
+\begin{mh}
+{\color{red}Comparing the results of different studies,
+  the impact of atoms candidates on source code
+  (mis)understanding differ according to the programming
+  language.}
+\end{mh}
 %% Regarding the first question we 
@@
 Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
-Ternary Operator & 41.80 & 39.34 & -5.90 \\ 
+Conditional Operator & 41.80 & 39.34 & -5.90 \\ 
 Pre-Increment & 30.71 & 42.45 & 38.19 \\  
@@
 Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05 & -0.39 \\ 
-Ternary Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
+Conditional Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
 Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981 & 0.27 \\ 
@@
          Logic as Control Flow           & 20 & 60     & 20    \\
-         Ternary Operator                & 60 & 26.67  & 13.3  \\
+         Conditional Operator                & 60 & 26.67  & 13.3  \\
          Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
@@
 In no case the \emph{neutral}
-ratio was higher than the option for the Clean version.
+ratio was higher than the option for the clean version.
 %% An example entry is contained in Appendix~\ref{}. 
 %% For a full listing of the code snippets, visit the paper repository. 
-Only for the Ternary Operator atom candidate the participants prefer the Confusing version, instead of the Clean one.
-We show the code snippets for the Ternary Operator in Figure~\ref{code:ternary}.
+Only for the Conditional Operator atom candidate the participants prefer the Confusing version, instead of the clean one.
+We show the code snippets for the Conditional Operator in Figure~\ref{code:ternary}.
 Even for this atom candidate, some participants who preferred the left-hand side version (confusing) still believed that the right-hand side version (clean) was more readable.
@@
 \noindent\begin{minipage}{.45\textwidth}
-\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Ternary Operator} atom).]
+\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Conditional Operator} atom).]
 let config = {size: 3, isActive: false};
@@
 \end{minipage}
-\caption{Example of a pair of code snippets used in the interview. This case explores the use of the \emph{Ternary Operator atom of confusion}}
+\caption{Example of a pair of code snippets used in the interview. This case explores the use of the \emph{Conditional Operator atom of confusion}}
 \label{code:ternary}
@@
 frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
-occur in at least 80\% of the projects. Considering the extremes, atom candidates Ternary Operator and Implicit Predicate were found in all repositories we mined,
+occur in at least 80\% of the projects. Considering the extremes, atom candidates Conditional Operator and Implicit Predicate were found in all repositories we mined,
 while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
@@
 Implicit Predicate & 100\% & 19.89 \\%&  intensively used \\ 
-  Ternary Operator & 100\% & 10.16 \\%&  intensively used \\ 
+  Conditional Operator & 100\% & 10.16 \\%&  intensively used \\ 
   Omitted Curly Braces & 91.67\% & 6.61 \\%&  intensively used \\ 
@@
 Considering all JavaScript projects in our dataset, we found a total of \num{364873} atom candidates, though four atoms are responsible
-for 92.97\% of this total: Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces.
+for 92.97\% of this total: Implicit Predicate, Post Increment, Conditional Operator, and Omitted Curly Braces.
 The first two, based on the results of Section~\ref{sec:survey-resuts}, can be considered atoms of confusion.
@@
 The four atoms mentioned in the previous paragraph,
- Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces, occur frequently in the analyzed projects.
+ Implicit Predicate, Post Increment, Conditional Operator, and Omitted Curly Braces, occur frequently in the analyzed projects.
 All of them have more than 5 occurrences per KLOC on average.
-In particular, two atom candidates occur very frequently, Implicit Predicate (19.89 occurrences per KLOC) and Ternary Operator (10.16 occurrences per KLOC).
+In particular, two atom candidates occur very frequently, Implicit Predicate (19.89 occurrences per KLOC) and Conditional Operator (10.16 occurrences per KLOC).
 On the other hand, Arithmetic as Logic and Comma Operator occur less than once for every 50 KLOC.
@@
 
-The results of our survey and interviews suggest that the Ternary Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to our survey).
+The results of our survey and interviews suggest that the Conditional Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to our survey).
 Nonetheless, the Post-Increment Expression and Automatic Semicolon Insertion atoms are listed in the top four sources of misunderstanding  (Table~\ref{tab:difference-correctness}), and also frequently appear in open source systems.
 As such, fixing the the Post-Increment and Automatic Semicolon Insertion might improve the readabiliy of large amounts of code in JavaScript projects.
+
+\todo[inline]{Seria legal contrastar esses achados com o que foi reportado anteriormente}

### 2021-10-20 3_Method.tex
minor change
revisao da secao de resultados e discussao

@@
 \item \rqb
-\item \rqc  
+% \item \rqc  
 \item \rqd
@@
 programmers to take longer to predict their output.
-Since JavaScript and C have some constructs in common (Section~\ref{back}), we first selected a set of 9+1 atom candidates: 9 atoms of confusion already validated~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} for the \clang language that also exist in JavaScript programs + 1 atom candidate that is specific of the JavaScript language (Automatic Semicolon Insertion).
+Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first selected a set of 9+1 atom candidates: 9 atoms of confusion already validated~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} for the \clang language that also exist in JavaScript programs + 1 atom candidate that is specific of the JavaScript language (Automatic Semicolon Insertion).
 Our supplementary material discusses and shows examples of the atoms we consider in our research.
@@
 Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research.
-The sizes of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
+The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
 All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars.

### 2021-10-20 4_Result.tex
revisao da secao de resultados e discussao

@@
 In fact, for the Arithmetic as Logic and the Pre-Increment atoms, the time to give a correct answer increased
-29.06\% and 38.19\% for the clean versions.
+29.06\% and 38.19\% for the clean versions---we did not confirm these
+atoms as atoms of confusion in the previous section.
 
@@
 
-\
 % \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}
 
-\vspace*{-0.5cm}
+Next, we summarize the results of our second study (interview with
+practitioners), which
+we use to contrast the findings discussed here with the opinion
+of software developers about their preferences regarding the confusing
+or clean versions of the code snippets we used in our experiment.
+
 \section{Results of the Interview Study}
@@
 
-In this section, we present the results of the 
-interviews with the 15 practitioners who agreed to participate in the study. 
-% A total of 15 practitioners took part of the interviews.
+%In this section, we present the results of the 
+% interviews with the 15 practitioners who agreed to participate in the study. 
+A total of 15 practitioners took part of the interviews.
 We collected information regarding programming
@@
 \begin{mq}
-\emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the 
+\emph{``When I am programming, I write code with the conditional operator, [...], but, to be honest, I still think that the 
 [the code using the \rhs version] 
@@
 
-When analyzing the Logic as Control Flow atom candidate, one of the interviewees gave an example personal experience that might motivate one to avoid writing code using this construct:
+When analyzing the Logic as Control Flow atom candidate, one of the interviewees gave an example of personal experience that might motivate one to avoid writing code using this construct:
 
 \begin{mq}
-  \emph{``This one is interesting, because I have written code that looks like the left-hand side [Confusing version], and my colleagues complained that it was difficult to understand.
-Nowadays I prefer to write code using the version on the right-hand side [Clean]''}.
+  \emph{``This one is interesting, because I have written code that looks like the left-hand side [confusing version], and my colleagues complained that it was difficult to understand.
+Nowadays I prefer to write code using the version on the right-hand side [clean version]''}.
 \end{mq}
@@
 \emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).
+As future work, we aim at investigating whether or not these atom candidates are more likely to introduce confusion in JavaScript code.
+The following section presents the results of our mining software repository effort, whose goal is to reveal how frequent atom candidates are in open source code.
 
@@
 As such, fixing the the Post-Increment and Automatic Semicolon Insertion might improve the readabiliy of large amounts of code in JavaScript projects.
-
-\todo[inline]{Seria legal contrastar esses achados com o que foi reportado anteriormente}

### 2021-10-20 discussion.tex
revisao da secao de resultados e discussao

@@
 the impact of atom candidates on
-understanding \clang and \cpplang code. That is,
+understanding \clang,\cpplang code. That is,
 similarly to their work, the atom candidates
 Comma Operator, Post/Pre Increment, Omitted Curly Braces,
-Assignment as Value, Implicit Predicate, Logic as
-Control Flow, and Ternary Operator seem to make 
+Assignment as Value, Implicit Predicate, and Logic as
+Control Flow seem to make 
 JavaScript code harder to understand. For five of them, the difference is statistically significant, with a large effect size for four atoms. Our results also 
@@
 The results of the interview study complement the understanding of atoms of confusion because the participants make clear the existence of a trade-off between code comprehension and other quality attributes.
-For instance, most of the participants prefer the version of the code with the Ternary Operator, even though they agree that its use might contribute to the misunderstanding of JavaScript code, particularly when novices are maintaining the codebase.
+For instance, most of the participants prefer the version of the code with the Conditional Operator, even though they agree that its use might contribute to the misunderstanding of JavaScript code, particularly when novices are maintaining the codebase.
 The participants of the interview study also
@@
 In summary, the results of the second study (interviews) allow
-us to answer the second (\rqb) and third research questions
-(\emph{\rqc}).
+us to answer the second (\rqb) research question.
 
@@
   interviews supports the results of the first study,
-  indicating that atoms of confusion compromise
+  since JavaScript developers most often agree that atoms of confusion compromise
   source code understanding. 
@@
 
-\begin{mh}
-  {\bf Answer to RQ3:} The qualitative analysis of the
-  interviews suggests that specific JavaScript constructs might also correspond to atoms of
-  confusion, including prototype inheritance and
-  nested callbacks. 
-\end{mh}
+%% \begin{mh}
+%%   {\bf Answer to RQ3:} The qualitative analysis of the
+%%   interviews suggests that specific JavaScript constructs might also correspond to atoms of
+%%   confusion, including prototype inheritance and
+%%   nested callbacks. 
+%% \end{mh}
 
 The results of the third study (mining open source
-Java script repositories) provides evidence that,
+JavaScript repositories) provides evidence that,
 although atoms of confusion compromise program
@@
 \begin{mh}
-  {\bf Answer to RQ4:} The MSR study reveals that
-  the atoms of confusion explored in our research
-  appear frequently in practice. 
+  {\bf Answer to RQ3:} The MSR study reveals that
+  the several atom candidates explored in our research
+  appear frequently in practice, and cleaning up the use of 
+  Post-Increment/Decrement and the Automatic Semicolon Insertion
+  might improve the readabiliy of JavaScript code substantially. 
+(\emph{\rqc}) 
 \end{mh}

### 2021-10-21 4_Result.tex
remocao de texto nao essencial para ajustar as limitacoes de espaco

@@
 
-\begin{mh}
-  The results of our exploratory data analysis and   hypothesis testing suggest that the atom candidates Comma Operator, Automatic Semicolon Insertion, Post Increment, Omitted Curly Braces, Assignment as Value, and Implicit Predicate %, and Ternary Operator
-  introduce some degree of misunderstanding in JavaScript code. %Based on the results of this study, 
-  Therefore, they can be deemed atoms of confusion for JavaScript.
-\end{mh}
+%% \begin{mh}
+%%   The results of our exploratory data analysis and   hypothesis testing suggest that the atom candidates Comma Operator, Automatic Semicolon Insertion, Post Increment, Omitted Curly Braces, Assignment as Value, and Implicit Predicate %, and Ternary Operator
+%%   introduce some degree of misunderstanding in JavaScript code. %Based on the results of this study, 
+%%   Therefore, they can be deemed atoms of confusion for JavaScript.
+%% \end{mh}
 
@@
 
-\begin{mh}
-{\color{red}Comparing the results of different studies,
-  the impact of atoms candidates on source code
-  (mis)understanding differ according to the programming
-  language.}
-\end{mh}
+%% \begin{mh}
+%% {\color{red}Comparing the results of different studies,
+%%   the impact of atoms candidates on source code
+%%   (mis)understanding differ according to the programming
+%%   language.}
+%% \end{mh}
 %% Regarding the first question we 
@@
 
-\begin{mh}
-  The results suggest that developers tend to spend
-  more time to predict the correct answer in the presence of the atom candidates Comma Operator, Logic as Control Flow, and Implicit Predicate.
+%% \begin{mh}
+%%   The results suggest that developers tend to spend
+%%   more time to predict the correct answer in the presence of the atom candidates Comma Operator, Logic as Control Flow, and Implicit Predicate.
 Conversely, the code snippets with the atom candidates Arithmetic as Logic and Pre-Increment demand less time to predict the correct answers.
-\end{mh}
+%% \end{mh}
 

### 2021-10-21 2_Background.tex
ajustes no texto necessarios para o limite de 10 paginas

@@
 Gopstein et al., in another paper~\cite{ThinkingAloud} further explored the original atoms by additionally conducting interviews and having programmers discuss among themselves the studied atoms.
-The authors argue that this method is important to complement a quantitative analysis.
+The authors argue in favor of complementing a quantitative analysis.
 Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''}.
-We follow a similar approach as we augment our quantitative analysis with interviews.
+Here, we also complement the quantitative analysis with interviews.
 

### 2021-10-21 3_Method.tex
ajustes no texto necessarios para o limite de 10 paginas

@@
 Our main concern was to avoid introducing bias for future respondents.
-Since we posted the survey in a social media platform, if we gave respondents instant feedback, they might post comments on particular atoms, therefore interfering with future participants' thought process.
+Since we posted the survey in a social media platform, possible threats could have arisen if we gave respondents instant feedback.
 
@@
 \\ 
-P6 & Bsc Degree & 15  years  & Java, PHP, C, Python , Ruby, C\# and R
+P6 & Bsc Degree & 15  years  & Java, PHP, C, Python , Ruby, C\#
 \\ 

### 2021-10-21 4_Result.tex
ajustes no texto necessarios para o limite de 10 paginas

@@
 
-{\color{blue}Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impact correctness.
+Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impact correctness.
 The findings suggest that \emph{participant experience} impacts on the results related to correctness (see Figure~\ref{fig:correctness-over-experience}).
@@
 For instance, novice developers provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets.
-It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.} {\color{red}We replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.} Differently from \emph{participant experience}, the factor \emph{participant education} does not change substantially the correctness of the answers.
+It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.
+We replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.
+Differently from \emph{participant experience}, the factor \emph{participant education} does not change substantially the correctness of the answers.
 
@@
 
-{\color{red}Comparing to previous research, three 
+Comparing to previous research, three 
 atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed
@@
 the work of Langhout and Aniche for two atoms: Pre Increment and
-Logic as Control Flow.}
+Logic as Control Flow.
 
@@
 One of the participants mentioned that the use of JavaScript's prototype-based inheritance can make it difficult to understand code, particularly when involving deep prototype chains.
-This is a core feature of the language, and most high-level tools and frameworks abstract it away.
+%This is a core feature of the language, and most high-level tools and frameworks abstract it away.
 %Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.
@@
 Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript.
-Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.
+%Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.
 We also found other idioms that are JavaScript atom candidates, including

### 2021-10-21 main.tex
rephrase one of the interview questions
ajustes no texto necessarios para o limite de 10 paginas

@@
 \begin{abstract}
-  Evolving legacy code is a challenging task, particularly when the code has been poorly written, increasing effort and imposing a significant cognitive load on developers.
-Although previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in \clang, \cpplang, and Java, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript).
-In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study.
-From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: considering 10 atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates.
+  Evolving legacy code is a challenging task, particularly when the code has been poorly written or uses confuse idioms and language constructs, which might increase maintenance efforts and impose a significant cognitive load on developers.
+For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in statically typed languages such as \clang, \cpplang, and Java.
+In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language.
+We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study.
+Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript.
+Considering ten atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates.
 For five of these atoms, the difference is statistically significant (accounting for p-value correction).
@@
 In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems.
-Four atom candidates appear in 90\% of the analyzed projects and two of them occur more than once for every 100 lines of code in the dataset.
-  %: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
-  Altogether, our findings might help practitioners (a) better understand the implications of atoms of confusion on understanding JavaScript code, (b) avoid writing code that is unnecessarily difficult to maintain, and (c) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
-
-% Our investigation also reveals additional JavaScript constructs that might hinder even experienced developers from understanding JavaScript code, such as the use of callbacks functions.
-Altogether...
-
-%Evolving legacy code is a challenging task, particularly when the code has been poorly written---which imposes a significant cognitive load on developers when trying to familiarize themselves with a legacy codebase.
-Previous research has investigated the impact of small code patterns that contribute to misunderstanding the source code written in C and C++.
-However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamic languages (such as JavaScript).
-In this work, we present the results of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
-Our goal is to investigate the impact of atoms of confusion on understanding code written in JavaScript --- a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
-From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atom candidates.
-Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects---including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
-Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code.
-For instance, even experienced JavaScript developers agree that it is hard to understand code that uses callback functions, even though we found widespread use of callbacks in our projects dataset.
-Altogether, our findings might help developers better understand the implications of atoms of confusion on understanding JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code.
+Four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code in the dataset.
+Altogether, our findings might help practitioners: (1) better understand the implications of atoms of confusion on understanding JavaScript code, (2) avoid writing code that is unnecessarily difficult to maintain, and (3) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
 \end{abstract}
@@
 
-\input{1_Introduction}
+\input{1_Introduction_alt}
 \input{2_Background}

### 2021-10-21 2_Background.tex
Passed through Sections 1 and 2. Stopped at the beginning of Page 4. I will resume tomorrow.

@@
 The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96, DBLP:journals/ibmsj/OHareT94}.
-Atoms of confusion is an important field of program comprehension introduced by Gopstein et al.
+The study of program comprehension in the presence of atoms of confusion was introduced by Gopstein et al.
 \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings.
@@
 \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 code idioms and language constructs that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed.
-The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase.
+The least confusing atom produced a 14\% boost in prediction accuracy when it was removed, whereas the most confusing one showed a 60\% accuracy increase.
 In a different work, Gopstein et al.
-\cite{DBLP:conf/msr/GopsteinZFC18} presented the results of a comprehensive research on the incidence of atoms of confusion in the wild, considering open-source projects written in \clang and \cpplang~programming languages.
+\cite{DBLP:conf/msr/GopsteinZFC18} presented the results of a comprehensive study on the incidence of atoms of confusion in the wild, considering open-source projects written in \clang and \cpplang~programming languages.
 
@@
 On the one hand, this means that some of the atoms of confusion that exist in \clang programs may also occur in JavaScript code.
-On the other hand, differences between the languages and the surrounding programming cultures may lead to code patterns that are atoms of confusion in one language not being atoms in the other one.
+On the other hand, differences between the languages and the surrounding programming cultures may lead to code patterns that are atoms of confusion in one language not being atoms in the other one.\castor{Shouldn't we explain here how this work differs from previous work on the C language \textbf{besides} just saying that JavaScript is different?
+I think we should at least emphasize the different experimental setup.
+I think this is very important to emphasize how different this study is from the previous ones.
+}
 %Although JavaScript allows us to program in different paradigms from the one C follows, most of the JavaScript code found in popular repositories are written under the imperative paradigm.
@@
 Langhout and Aniche~\cite{Langhout:2021:ACJ} derived a set of 14 atoms for Java and performed an experiment with 132 students.
-They found that participants are more likely to make mistakes (ranging from 2.7 up to 56 times) when atoms are present in half of their studied atoms.
+They found that participants are more likely to make mistakes (ranging from 2.7 up to 56 times) when atoms are present in half of their studied atoms\castor{I do not understand the meaning of the previous section "atoms are present in half of their studied atoms".
+Furthermore, I do not understand the factors between parentheses.}.
 Castor~\cite{castor2018} presented a preliminary catalog of six atom candidates for the Swift programming language.
@@
 However, most of the practices analyzed in their study relate to code formatting, instead of language constructs or idioms that might introduce misunderstanding.
-We borrow from their research the methodology for collecting the practitioners opinion about code snippets preference.
+In addition, the study participants did not have to perform comprehension tasks.
+We borrow from their research the methodology for collecting the practitioners opinion about code snippet preferences.\castor{I would rather we did not say this about our study because this methodology of presenting two or more code snippets and asking subjects which one they prefer is fairly old, e.g., the following paper already used this approach: Arab M (1992) Enhancing Program Comprehension : Formatting and Documenting.ACM SIGPLAN Notices 27(2) :37–46}.
 

### 2021-10-21 3_Method.tex
Passed through Sections 1 and 2. Stopped at the beginning of Page 4. I will resume tomorrow.

@@
 %---though we postpone these results to future research work.
-We conduct a mixed-methods study to anwer these questions, including a survey, a set of interviews, and a study for mining JavaScript software repositories.
+We conduct a mixed-methods study to answer these questions, including a survey, a set of interviews, and a study for mining JavaScript software repositories.
 
@@
 misunderstanding rate when programmers
-try to predict their outcome and (b) require
-programmers to take longer to predict their output.
-Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first selected a set of 9+1 atom candidates: 9 atoms of confusion already validated~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} for the \clang language that also exist in JavaScript programs + 1 atom candidate that is specific of the JavaScript language (Automatic Semicolon Insertion).
+try to predict their outcome and (b) require more time for programmers to predict their output.
+Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first selected a set of 9+1 atom candidates: nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion).
 Our supplementary material discusses and shows examples of the atoms we consider in our research.
@@
 code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about non-confusing code snippets, they will only be presented with non-confusing versions of the code samples $s_6, s_7,..., s_{10}$.
-Furthermore, a given subject (P2), which constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing snippets for $s_6, s_7,..., s_{10}$.
+Furthermore, a given subject (P2), who constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing snippets for $s_6, s_7,..., s_{10}$.
 By doing that, we guarantee that all versions of the code snippets are contained within each square, and that each configuration occurs only once within a square.

### 2021-10-22 4_Result.tex
Made a pass through Section 4.

@@
 All participants had taken at least some university course or hold a bachelor
-degree or equivalent, meaning the participants have had some level of formal education in programming.
-Twenty four participants hold either a master's degree (21 participants) or a doctorate degree (three participants).
+degree or equivalent, meaning the participants have had some level of formal education in programming. In addition, 21 participants hold a master's degree and three a doctorate degree. 
 Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.
@@
 Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the confusing versions.
-In particular, the presence of the \emph{Comma Operator} atom presents the highest impact on misunderstanding.
+In particular, the presence of the \emph{Comma Operator} atom exhibits the highest impact on misunderstanding.
 Frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also result in many mistakes.
@@
 The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the clean versions of the code snippets.
-Also, the sample of responses with no atoms had almost no dispersion, which supports the
-argument that the non-confusing versions of the code snippets are easier to evaluate correctly.
+Also, the sample of responses with no atoms had almost no dispersion, which supports the argument that the non-confusing versions of the code snippets are easier to evaluate correctly.
 
@@
 \begin{tabular}{lccc} \toprule
-  Atom & Confusing Code & Clean Code & $\Delta$(\%)  \\ \midrule
+  Atom & Confusing & Clean & $\Delta$(\%)  \\ \midrule
   Comma Operator                &  28 &  65 & +132   \\ 
@@
  \label{fig:boxplotcorrectness}
- \end{figure}
+ \end{figure}\castor{Fiz de tudo para ajeitar a figura porque o texto embaixo deveria ser "Clean versions" e "Confusing versions" mas não sei porque não funcionou.
+Até mudei a figura.
+Não sei o que está havendo.}
 
@@
 
-\subsubsection*{Hypothesis Testing and Effect Size}
+\subsubsection*{Statistical analysis}
 
@@
 
-Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impact correctness.
-The findings suggest that \emph{participant experience} impacts on the results related to correctness (see Figure~\ref{fig:correctness-over-experience}).
+Finally, we also use a \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impact correctness.
+The findings suggest that \emph{participant experience} impacts on the results related to correctness (Figure~\ref{fig:correctness-over-experience}).
 Even though the presence of atoms of confusion reduces the number of correct answers to all \emph{experience groups}, this effect is not uniform.
-For instance, novice developers provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets.
+For instance, novice developers (up to four years of programming experience) provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets.
 It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.
-We replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.
-Differently from \emph{participant experience}, the factor \emph{participant education} does not change substantially the correctness of the answers.
+We replicate the \emph{Pearson's Chi-squared test} for each experience group and find that the statistical difference is less significant for those developers with more than ten years of experience.
+Differently from \emph{participant experience}, the factor \emph{participant education} does not significantly change the correctness of the answers.\castor{What does "for each experience group" mean above?}
 
@@
 \end{figure}
+\castor{This figure is missing. Is that expected?}
 
@@
 support the Gopstein et al. findings for the remaining
-atoms. Considering the work of Langhout and Aniche~\cite{Langhout:2021:ACJ},
-their work investigates six atom candidates for Java that
-we also explore here for JavaScript. Among these atom
-candidates, we achieve similar conclusions with their
+atoms. The work of Langhout and Aniche~\cite{Langhout:2021:ACJ} investigates six atom candidates for Java that we also explore here for JavaScript. Among these atom
+candidates, we achieve similar conclusions to their
 study for four atom candidates: Post Increment,
@@
 \begin{tabular}{lccc} \toprule
-Atom & Confusing Code & Clean Code & $\Delta$(\%) \\ \midrule
+Atom & Confusing  & Clean & $\Delta$(\%) \\ \midrule
 Comma Operator & 87.67 & 20.84 & -76.23 \\ 
@@
  \begin{table*}[ht]
-\caption{Hypotheses Testing (misunderstanding and time)}
+\caption{Hypotheses Testing (misunderstanding and time). Asterisks ($^{*}$) indicate a statistically significant difference.}
 
@@
      &         & (misunderstanding)  &                    & (workload)             & \\ \midrule
-Comma Operator & 1.1727e-10 & 19.02 & (6.60, 68.22) & 5.8249e-08 & -0.53 \\ 
-Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & (9.21, 356.62) & 0.09802 & -0.16 \\ 
-Post-Increment & 0.0015279 & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
+Comma Operator & 1.1727e-10$^{*}$ & 19.02 & (6.60, 68.22) & 5.8249e-08$^{*}$ & -0.53 \\ 
+Automatic Semicolon Insertion & 5.8352e-11$^{*}$ & 39.33 & (9.21, 356.62) & 0.09802 & -0.16 \\ 
+Post-Increment & 0.0015279$^{*}$ & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
 Omitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
-Assignment as Value & 0.0034775 & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
-Implicit Predicate & 0.01123 & 6.95 & (1.46, 66.56) & 0.0029039 & -0.29 \\ 
-Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05 & -0.39 \\ 
+Assignment as Value & 0.0034775$^{*}$ & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
+Implicit Predicate & 0.01123$^{*}$ & 6.95 & (1.46, 66.56) & 0.0029039$^{*}$ & -0.29 \\ 
+Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05$^{*}$ & -0.39 \\ 
 Conditional Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
-Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981 & 0.27 \\ 
-Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & 0.01723 & 0.23 \\ 
+Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981$^{*}$ & 0.27 \\ 
+Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & 0.01723$^{*}$ & 0.23 \\ 
     \bottomrule
@@
 
-\subsubsection*{Hypothesis Testing}
+\subsubsection*{Statistical analysis}
 We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
@@
 spend the same amount of time to correctly
-predict the outcome of a code snippet, regardless
-of evaluating the confusing or clean version
-of the code. The results of Table~\ref{tab:difference-time-taken}
-suggests that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
+predict the outcome of clean and confusing versions of a code snippet.
+%, regardless of evaluating the confusing or clean version of the code.
+The results of Table~\ref{tab:difference-time-taken}
+suggest that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
 For the first three, the analysis suggests that the participants need more time to predict the outcome of the code snippet in the confusing code version.
@@
 
-Next, we summarize the results of our second study (interview with
-practitioners), which
-we use to contrast the findings discussed here with the opinion
+\section{Results of the Interview Study}
+\label{sec:interview-results}
+
+In this section we present the results of the interviews with the practitioners. 
+We contrast the findings of the survey presented in Section~\ref{sec:survey-resuts} here with the opinion
 of software developers about their preferences regarding the confusing
 or clean versions of the code snippets we used in our experiment.
-
-\section{Results of the Interview Study}
-\label{sec:interview-results}
 
@@
 
-\subsubsection*{The Participants' perceptions of the atom candidates}
-Supporting the results of the survey, Table~\ref{tab:interview-results1}
+\subsubsection*{The Participants' perceptions of the atom candidates} Supporting the results of the survey, Table~\ref{tab:interview-results1}
  shows that for eight out of the
@@
 %% For a full listing of the code snippets, visit the paper repository. 
-Only for the Conditional Operator atom candidate the participants prefer the Confusing version, instead of the clean one.
-We show the code snippets for the Conditional Operator in Figure~\ref{code:ternary}.
-Even for this atom candidate, some participants who preferred the left-hand side version (confusing) still believed that the right-hand side version (clean) was more readable.
+Only for the Conditional Operator atom candidate the participants prefer the confusing version, instead of the clean one.
+Figure~\ref{code:ternary} shows the code snippets for the corresponding confusing and clean versions.
+For this atom candidate, some participants who preferred the left-hand side version (confusing) still believed that the right-hand side version (clean) was more readable.
 The following quotes were extracted from the transcripts with
@@
 \end{minipage}
-\caption{Example of a pair of code snippets used in the interview. This case explores the use of the \emph{Conditional Operator atom of confusion}}
+\caption{Example of a pair of code snippets used in the interview pertaining to the \emph{Conditional Operator} atom candidate}
 \label{code:ternary}
@@
 \begin{mq}
-\emph{``When I am programming, I write code with the conditional operator, [...], but, to be honest, I still think that the 
+\emph{``When I am programming, I write code with the conditional operator, [...], but, to be honest, I still think that 
 [the code using the \rhs version] 
@@
 divide in one of the interviewees, who
-regarded the clean version as simpler to understand, but would still opted to write
+regarded the clean version as simpler to understand, but would still opt to write
 code with the atom. In contrast, %to such opinion,
@@
 As future work, we aim at investigating whether or not these atom candidates are more likely to introduce confusion in JavaScript code.
-The following section presents the results of our mining software repository effort, whose goal is to reveal how frequent atom candidates are in open source code.
 
@@
 
+In this section we present the results of our mining software repository effort, whose goal is to reveal how prevalent atom candidates are in open source JavaScript projects. 
 We mined \minedprojects open source JavaScript repositories to understand how
@@
 \centering
-\caption{Summary of atom candidate occurrences on our dataset.}
+\caption{Summary of atom candidate occurrences in our dataset.}
 \setlength\tabcolsep{2pt} % default value: 6pt
@@
 Nonetheless, the Post-Increment Expression and Automatic Semicolon Insertion atoms are listed in the top four sources of misunderstanding  (Table~\ref{tab:difference-correctness}), and also frequently appear in open source systems.
-As such, fixing the the Post-Increment and Automatic Semicolon Insertion might improve the readabiliy of large amounts of code in JavaScript projects.
+As such, refactoring existing systems to avoid Post-Increment and Automatic Semicolon Insertion might improve the readabiliy of large amounts of code in JavaScript projects.

### 2021-10-22 main.tex
Made a pass through Section 4.

@@
 In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language.
-We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study.
+We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study\castor{Do we still want to talk about the MSR part as a "study"?
+I remember Rodrigo saying that we did not.}.
 Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript.

### 2021-10-23 4_Result.tex
fixed problem with the x labels of Figure 2

@@
 \noindent
-\vspace*{-1.4cm}
  \centering
  \includegraphics[width=\columnwidth]{images/wrong-answers-plot-1.pdf}
- \vspace*{-1.3cm}
  \caption{Number of wrong answers of each subject.}
  \label{fig:boxplotcorrectness}
- \end{figure}\castor{Fiz de tudo para ajeitar a figura porque o texto embaixo deveria ser "Clean versions" e "Confusing versions" mas não sei porque não funcionou.
+ \end{figure}% \castor{Fiz de tudo para ajeitar a figura porque o texto embaixo deveria ser "Clean versions" e "Confusing versions" mas não sei porque não funcionou.
 Até mudei a figura.

### 2021-10-23 2_Background.tex
addressing some comments from F. Castor

--- 2021-10-21 2_Background.tex
@@
 On the other hand, differences between the languages and the surrounding programming cultures may lead to code patterns that are atoms of confusion in one language not being atoms in the other one.
+Compared to previous research, in this work we leverage the \emph{Latin square} experimental design to control both participant experience and education better, while investigating the effect of the atom candidates in code comprehension.
+We present details about our setup in Section~\ref{method}.
+Differently, previous research works (e.g., \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ}) use a \emph{Randomized Partial Counterbalanced Design}, which does not \emph{block} relevant characteristics of the participants.
+
 %Although JavaScript allows us to program in different paradigms from the one C follows, most of the JavaScript code found in popular repositories are written under the imperative paradigm.
@@
 Langhout and Aniche~\cite{Langhout:2021:ACJ} derived a set of 14 atoms for Java and performed an experiment with 132 students.
-They found that participants are more likely to make mistakes (ranging from 2.7 up to 56 times) when atoms are present in half of their studied atoms.
+For seven atom candidates (out of 14), they report that participants are more likely to make mistakes in the confusing version of the code.
 Castor~\cite{castor2018} presented a preliminary catalog of six atom candidates for the Swift programming language.
@@
 However, most of the practices analyzed in their study relate to code formatting, instead of language constructs or idioms that might introduce misunderstanding.
-We borrow from their research the methodology for collecting the practitioners opinion about code snippets preference.
+In addition, the study participants did not have to perform comprehension tasks.
+We borrow from their research the methodology for collecting the practitioner's opinion about code snippets.
+Indeed, such a methodology is a well-known approach for comparing developers' preferences among pairs of code snippets (e.g., the work of M.
+Arab about the effect of code comments and documentation already uses this approach~\cite{arab:sigplan}).
 

### 2021-10-23 4_Result.tex
addressing some comments from F. Castor

@@
 Even though the presence of atoms of confusion reduces the number of correct answers to all \emph{experience groups}, this effect is not uniform.
-For instance, novice developers (up to four years of programming experience) provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets.
+For instance, novice developers (under one year of experience) provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets.
 It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.
 We replicate the \emph{Pearson's Chi-squared test} for each experience group and find that the statistical difference is less significant for those developers with more than ten years of experience.
-Differently from \emph{participant experience}, the factor \emph{participant education} does not significantly change the correctness of the answers.\castor{What does "for each experience group" mean above?}
+Differently from \emph{participant experience}, the factor \emph{participant education} does not significantly change the correctness of the answers.
 
@@
   \includegraphics[scale=0.3]{images/correctness-by-experience}
-  \caption{Impact of atoms of confusion on the correctness (over the participant experience groups)}
+  \caption{Impact of atoms of confusion on the correctness (over the four participant experience groups)}
   \label{fig:correctness-over-experience}
 \end{figure}
-\castor{This figure is missing. Is that expected?}
 

### 2021-10-25 2_Background.tex
submitted version

@@
 
-Santos and Gerosa~\cite{Santos:2018:ICP} analyze the impact of eleven coding practices on program readability.
+%Santos and Gerosa~\cite{Santos:2018:ICP} analyze the impact of eleven coding practices on program readability.
 However, most of the practices analyzed in their study relate to code formatting, instead of language constructs or idioms that might introduce misunderstanding.

### 2022-01-21 main.tex
JSS format

--- 2021-10-21 main.tex
@@
-\documentclass[10pt,conference,blind]{IEEEtran}
+\documentclass[review]{elsarticle}
 
-\IEEEoverridecommandlockouts
+%%\acmConference[ESEC/FSE 2020]{The 28th ACM Joint European Software %%Engineering Conference and Symposium on the Foundations of Software %%Engineering}{8 - 13 November, 2020}{Sacramento, California, United States}
 
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+% Fonte em português brasileiro
 
 \usepackage{enumerate}
-\usepackage{balance}
 \usepackage{caption}
-\usepackage{cite}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 \usepackage{listings}
-\usepackage{color}
 \usepackage{scalefnt}
@@
 \usepackage{multicol}
-\usepackage{listings}
-\usepackage{xspace}
-\usepackage{booktabs}
+
 \usepackage{siunitx}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{graphics}
 \usepackage{textcomp}
-\usepackage{siunitx}
-\usepackage{xcolor}
 \usepackage{csquotes}
 \usepackage{array}
-\usepackage{listings}
-\usepackage{color}
 \usepackage{scalefnt}
@@
 \usepackage{xurl}
-% \usepackage{appendix}
+
+\usepackage{csvsimple}
+\usepackage{pgfplotstable,filecontents}
+
+% Default fixed font does not support bold face
+\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{12} % for bold
+\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{12}  % for normal
+
+% Identação e Margens
+\usepackage{fullpage}      % Margens
+\usepackage{indentfirst}   % Autoidentar
+
+% Figuras
+\usepackage{graphicx}       % Pictures
+\usepackage{subfig}
+\usepackage{venndiagram}
+
+%%%%%%
+
 \usepackage[framemethod=TikZ]{mdframed}
@@
 
+\newcommand{\na}{10\xspace}
+
 \newcommand{\clang}{C\nolinebreak\xspace}
@@
 
-\newcommand{\na}{10\xspace}
+\begin{document}
 
-\begin{document}
-\input{atom-mdframe-config}
+\begin{frontmatter}
 
@@
 
-\author{\IEEEauthorblockN{Anonymous}}
+\author[1]{Adriano Torres}
+\author[1]{M\'{a}rcio Okimoto}
+\author[1]{Caio Oliveira}
+\author[2]{Diego Marc\'{i}lio}
+\author[3]{Pedro Queiroga}
+\author[4]{M\'{a}rcio Ribeiro}
+\author[1]{Edna Dias Canedo}
+\author[5]{Fernando Castor}
+\author[1]{Rodrigo Bonif\'{a}cio}
+\author[6]{Eduardo Monteiro}
 
-%% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% \and
-%% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% \and
-%% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% \and
-%% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% \and
-%% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% \and
-%% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% }
+\address[1]{Computer Science Department, University of Bras\'{i}lia, Brazil}
+\address[2]{Faculty of Informatics, Universit\`{a} della Svizzera Italiana, Switzerland}
+\address[3]{Informatics Center, Federal University of Pernambuco, Brazil}
+\address[4]{Institute of Computing, Federal University of Alagoas, Brazil}
+\address[5]{Department of Information and Computing Sciences, Utrecht University, The Netherlands}
+\address[6]{Statistics Department, University of Bras\'{i}lia, Brazil}
 
-\maketitle
+\input{abstract}
 
-% Número das páginas. Remover para camera-ready
-% útil para nós e Program Chair que pode ver o total de páginas facilmente
-\thispagestyle{plain}
-\pagestyle{plain}
-
-\begin{abstract}
-  Evolving legacy code is a challenging task, particularly when the code has been poorly written or uses confuse idioms and language constructs, which might increase maintenance efforts and impose a significant cognitive load on developers.
-For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in statically typed languages such as \clang, \cpplang, and Java.
-In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language.
-We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study.
-Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript.
-Considering ten atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates.
-For five of these atoms, the difference is statistically significant (accounting for p-value correction).
-Effect size is large for four atoms and medium for one of them.
-In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems.
-Four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code in the dataset.
-Altogether, our findings might help practitioners: (1) better understand the implications of atoms of confusion on understanding JavaScript code, (2) avoid writing code that is unnecessarily difficult to maintain, and (3) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
-\end{abstract}
-
-\begin{IEEEkeywords}
-code readability, program comprehension, program understanding, atoms of confusion, JavaScript code
-%datasets, neural networks, gaze detection, text tagging
-\end{IEEEkeywords}
+\end{frontmatter}
 
@@
 \input{6_Conclusion}
-% \input{7_Atoms}
 
-% \section*{Acknowledgment}
-
-\balance
-\bibliographystyle{IEEEtran}
+\bibliographystyle{elsarticle-num}
 \bibliography{reference}
 
-\vspace{12pt}
 \end{document}

### 2022-01-21 abstract.tex
moving the abstract to a separate file

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written or uses confuse idioms and language constructs, which might increase maintenance efforts and impose a significant cognitive load on developers. For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in statically typed languages such as \clang, \cpplang, and Java. In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language. We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and an activity of mining open source JavaScript repositories (MSR). Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript. Considering ten atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates. For five of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for four atoms and medium for one of them. In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code in the dataset. Altogether, our findings might help practitioners: (1) better understand the implications of atoms of confusion on understanding JavaScript code, (2) avoid writing code that is unnecessarily difficult to maintain, and (3) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
\end{abstract}

\begin{keyword}
  code readability, program comprehension, program understanding, atoms of confusion, JavaScript code
\end{keyword}

### 2022-03-29 3_Method.tex
Added text describing the methodology of the replication study as a separate file. This commit aims to be an intermediate step. The next one will be to combine the methodologies of both studies, in particular the two experiments.

@@
 Our main concern was to avoid introducing bias for future respondents.
-Since we posted the survey in a social media platform, possible threats could have arisen if we gave respondents instant feedback.
+Since we posted the survey on a social media platform, possible threats could have arisen if we gave respondents instant feedback.
 
@@
 
-\subsubsection*{Interview Process} We conducted semi-structured interviews using a web conferencing software.
+\subsubsection*{Interview Process} We conducted semi-structured interviews using web conferencing software.
 We recorded all the interviews with the consent of the participants.
@@
 The interviews had three main parts.
-In the first one, we asked the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years and other programming languages they have worked with.
+In the first one, we asked the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years, and other programming languages they have worked with.
 Table \ref{pinterview} summarizes this demographic information.
@@
 This would allow us to identify potential atom candidates that are more specific to the JavaScript language.
-Examples of questions we explored in this section include: \emph{Does JavaScript favor developers to produce code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as
-  especially confusing?}
+Examples of questions we explored in this section include: \emph{Does JavaScript favor developers to produce code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as especially confusing?}
 
@@
 We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study).
-Finally, we carried out basic descriptive statistics in order to understand how often the atoms of confusion appear in practice.
+Finally, we computed some descriptive statistics to measure the prevalence of atoms of confusion in practice.
 

### 2022-03-29 6_Conclusion.tex
Added text describing the methodology of the replication study as a separate file. This commit aims to be an intermediate step. The next one will be to combine the methodologies of both studies, in particular the two experiments.

@@
 
-This paper reports the results of a mixed-method research effort to better understand the impact of atoms of confusion in JavaScript code.
+This paper reports the results of a mixed-method research effort that allowed us to better understand the impact of atoms of confusion in JavaScript code.
 First, we conducted a survey with 140 JavaScript developers, asking them to predict the output of some code snippets (with and without atoms).
-We provide evidence that five atom candidates significantly impact program comprehension activities and can be considered atoms of confusion: Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate.
-We also interviewed 15 professional developers.
+We provide evidence that five atom candidates significantly impact program comprehension activities and can be considered atoms of confusion.
+%% : Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate. 
+%% We also interviewed 15 professional developers.
 In general, they considered code without atom candidates easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding.
 %Last, but not least,
-Finally, we conducted a mining software repositories study.
+%Finally, we conducted a mining software repositories study.
 We found four atom candidates frequently used in JavaScript code, i.e., Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces.
 %The first two can be considered atoms of confusion.
-
-Our efforts have important implications for practice.
-First, we alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate).
-Second, our results might help tool developers to create program transformation tools to remove the atoms automatically.
+Our efforts have two implications for practice.
+First, we present evidence that not all atoms of confusion validated to statically typed languages (such as \clang, \cpplang, or Java) led to a statistically significant impact on program understanding for JavaScript (a dynamically typed language).
+Besides that, our findings might be used to alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment/Decrement, Assignment as Value, and Implicit Predicate).
+Finally, our results might help tool developers to create program transformation tools for removing atoms that frequently appear in software.
 %They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).
 %Promising candidates may be the atoms we found to be most confusing and, at the same time, common in practice (e.g., Implicit Predicate and Post Increment).
-
 %As mentioned, the results of the interviews
-The results of the interviews pointed out additional JavaScript constructs and idioms 
+Our research also pointed out additional JavaScript constructs and idioms 
 that might introduce misunderstanding. These include nested callbacks, % and prototype-based inheritance (which might not have a counterpart version), but also idioms such

### 2022-03-29 dataset/leituraDados.Rmd
Added text describing the methodology of the replication study as a separate file. This commit aims to be an intermediate step. The next one will be to combine the methodologies of both studies, in particular the two experiments.

5751137

---
title: "Atoms"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(".")
```

```{r}
dados <- read.csv("survey-results.csv")
```

```{r,echo=FALSE}
##Considerando todos os dados... Refeito apos a selecao dos quadrados completos
# prop.table(table(dados$correct,dados$atom),margin = 2)
# chisq.test(dados$correct,dados$atom)
# dim(dados)
# length(unique(dados$rid))
# length(unique(dados$sid))
# length(unique(dados$qid))
# table(dados$qid,dados$atom)
# summary(as.factor(dados$qid))
# zeroum <- ifelse(dados$correct=="CORRECT",1,0)
# mod <- glm(zeroum~dados$atom+dados$experience+dados$education,family = "binomial")
# mod <- glm(zeroum~dados$atom+dados$education,family = "binomial")
# ?glm
# summary(mod)
# table(predict(mod,type="response")>.5, dados$correct)
# 
# 
# for(i in 1:4){
#   print(paste("experiencia ", i ))
#   print(prop.table(table(dados$correct[dados$experience==i],dados$atom[dados$experience==i]),margin=2))
#   print(chisq.test(table(dados$correct[dados$experience==i],dados$atom[dados$experience==i])))
# }
# 
# for(i in 1:4){
#   print(paste("experiencia ", i ))
#   print(prop.table(table(dados$correct[dados$experience==i],dados$atom[dados$experience==i]),margin=2))
#   print(chisq.test(table(dados$correct[dados$experience==i],dados$atom[dados$experience==i])))
# }
# 
# 
# for(i in 1:5){
#   print(paste("educacao ", i ))
#   print(prop.table(table(final$correct[dados$education==i],dados$atom[dados$education==i]),margin=2))
#   print(chisq.test(table(dados$correct[dados$education==i],dados$atom[dados$education==i])))
# }
```

Fazer a limpeza para ficar só os casos completos

```{r}
subred <- dados[dados$ref=="reddit" & dados$total == "YES",]
summary(subred$total)
contagem <- tapply(subred$time,subred$rid,length)
casoscompletos <- names(contagem)[contagem==20]
final <- subred[is.element(el = subred$rid, set = casoscompletos),]
dim(final)
```

Refazendo a análise considerando apenas os casos completos

```{r}
prop.table(table(final$correct,final$atom),margin = 2)
chisq.test(final$correct,final$atom)
```
```{r}
teste <- as.factor(final$experience)
contrr<-matrix(c(rep(1,4),c(1/2,1/2,-1/2,-1/2),c(1,-1,0,0),c(0,0,1,-1)),byrow=TRUE,nrow=4)
contrasts(teste)<-solve(contrr)[,2:4]

mod <- glm(final$correct~final$atom+teste+as.factor(final$education),family = "binomial")

#mod <- glm(zeroum~final$atom+final$education,family = "binomial")
summary(mod)

for(i in 1:4){
  print(paste("experiencia ", i ))
  print(prop.table(table(final$correct[final$experience==i],final$atom[final$experience==i]),margin=2))
  print(chisq.test(table(final$correct[final$experience==i],final$atom[final$experience==i])))
}

for(i in 1:5){
  print(paste("educacao ", i ))
  print(prop.table(table(final$correct[final$education==i],final$atom[final$education==i]),margin=2))
  print(chisq.test(table(final$correct[final$education==i],final$atom[final$education==i])))
}
```

## Fazendo a análise por questão, juntando pares de questão com atomo e sem atomo

```{r}
valoresqui <- c()
for(i in 1:10){
  subconjunto <- final[is.element(el = final$qid, set = c(i,i+10)),]
  print(paste("Questão ", i))
  print(table(subconjunto$correct,subconjunto$atom))
  teste <- chisq.test(table(subconjunto$correct,subconjunto$atom))
  print(teste)
  valoresqui[i] <- teste$p.value
  experiencia <- as.factor(subconjunto$experience)
  contrr<-matrix(c(rep(1,4),c(1/2,1/2,-1/2,-1/2),c(1,-1,0,0),c(0,0,1,-1)),byrow=TRUE,nrow=4)
  contrasts(experiencia)<-solve(contrr)[,2:4]
  mod <- glm(subconjunto$correct ~ subconjunto$atom +  experiencia,family="binomial")
  print(summary(mod))
  print(prop.table(table(subconjunto$correct,subconjunto$experience),margin = 2))
}

```

### 2022-03-29 dataset/survey.Rmd
Added text describing the methodology of the replication study as a separate file. This commit aims to be an intermediate step. The next one will be to combine the methodologies of both studies, in particular the two experiments.

---
title: "survey-analysis-final"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='images/', dev=c('png', 'pdf'))

library(questionr)
library(fmsb)
library(sqldf)
library(xtable)
library(effsize)
setwd(".")

```

## Load the datasets

```{r}
ds <- read.csv("interview-summary.csv", head=T, sep=";")

meanValues = c("", mean(ds$Left), mean(ds$Right), mean(ds$Neutral))

colnames(ds) <- c("Atom", "LHS (without AOC)", "RHS (with AOC", "Neutral")

ds <- rbind(ds, meanValues)
ds

xtable(ds)
```

### 2022-03-29 method_replication_study.tex
Added text describing the methodology of the replication study as a separate file. This commit aims to be an intermediate step. The next one will be to combine the methodologies of both studies, in particular the two experiments.

We call the first study we performed the \textit{``Replication Study''} because it is a partial replication~\cite{Juristo2012} of the first study on atoms of confusion, conducted by Gopstein et. al~\cite{GopsteinIYDZYC17}, but targeting JavaScript instead of the C language. Gopstein et. al. employed a randomized, partial counterbalancing experimental design on a sample of 73 subjects, mostly students, with three or more months of experience in C or C++. The control variable consisted of a tiny program containing a single atom candidate and the treatment a functionally-equivalent version that does not contain the atom. The main dependent variable is whether subjects are able to correctly determine the output of the programs. The other dependent variable is the time required to correctly vs. incorrectly determine the out of the programs. The study had the null hypothesis that \textit{``code from both control and treatment groups can be hand-evaluated with equal accuracy''} and the alternate hypothesis that \textit{``the existence of an atom of confusion causes more errors than other code in hand-evaluated outputs.''}. For each atom candidate, six programs were built, three versions including the atom candidate and three versions not including it. Each subject was exposed to four programs for every atom candidate, two including an atom and two not including an atom.

In this study we adopt the same experimental design, control and treatment variables, dependent variable, and hypotheses. We have also built six programs for each atom candidate, three with and three without the atom. Considering the elements that can vary in a replication study, defined by Juristo and Gomez~\cite{Juristo2012}, in this study we employ the same experimental protocol and construct operationalizations. We have different population properties, experimenters, and a different site. In addition, since we focus on a different language, some atoms that exist in C cannot by construction exist in JavaScript programs. Conversely, JavaScript exhibits atom candidates that do not exist in C.

The design of the replication study considers two treatments: the presence or absence of atoms candidates within the programs. To identify atom candidates, we start out by selecting every atom candidate from Gopstein's study that can also happen in JavaScript. That excludes atoms related to pointers and preprocessor directives. Furthermore, we conduct an informal search of programming forums, solutions to Code Golf~\footnote{https://codegolf.stackexchange.com/} problems, code style guidelines, and expert knowledge to identify new atom candidates. Overall, the replication study analyzes 24 atoms, 15 from the original study and 9 that are new and specific to JavaScript. Table~\ref{tab:atoms_replication} presents the complete list.

For each atom candidate, we write six short programs, three including and three not including the atom candidate. The former are the control for the study or the obfuscated versions of the programs. The alternative, atom-free, versions of the programs are the treatment, or clean, versions. Each subject is exposed to one obfuscated and one clean version of each atom, totalizing 48 programs, and should determine their outputs. As mentioned before, we measure answer correctness and the time to answer questions correctly vs. incorrectly. We control for learning effects~\cite{Neely:1991:SPE} in three manners. First, by having multiple obfuscated and clean versions for each atom and presenting only one of each per subject. Second, by presenting the programs in a random order. Third, by presenting obfuscated and clean versions corresponding to the same atom candidate with at least 11 other programs in between, in  accordance to the original experimental protocol~\cite{GopsteinIYDZYC17}.

\subsubsection*{Survey Instrument}

The study is conducted by means of a web application. We conducted an initial pilot to find bugs with the application and to get general impressions. Based on the feedback of the participants of the pilot, we decided to present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation. Initially, our plan was to present two of each. 
To reduce the likelihood of a subject simply running the program, the web application verifies whether the subject attempts to change tabs or windows and presents the programs as images, instead of text.

We organize the survey in three sections. In the first one we present instructions and also a check button whose checking meant users agreed that all collected data would be used anonymously and solely for research purposes. The instructions explain how the survey works and asks them to dedicate their attention to it.  We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters. For each question page, we kept track of whether or not the subjects switched windows.

The last section of the survey presented a sequence of ten questions, each containing a code snippet. For each question, there was a text box where the answer should be written. There was also an ``I do not know'' button, which, when clicked, led the subject to the next question. In our setting, ``I do not know'' was treated as a wrong answer. The code snippets were presented as images copied from a text editor, so as to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter. Upon submitting their answer for a particular question, the subject was automatically led to a similar page, containing another snippet.

We decided not to provide feedback about the time subjects took to answer each question. We also did not tell them whether their answers were correct or not. Our main concern was to avoid introducing bias for future respondents. Since we posted the survey on a social media platform, possible threats could have arisen if we gave respondents instant feedback.

\subsubsection*{Survey audience}

We posted the survey on the authors' social media platforms, on Brazilian CS departments' mailing lists, and on two programming subreddits. We explained our research purposes and asked developers to take the survey only if they have some familiarity with JavaScript. We collected 70 responses. The average survey respondent has a mean 26 months of experience in JavaScript, with a median of 12.

### 2022-03-29 3_Method.tex
First version with the combined methodology section. The initial part is written starting from the assumption that the MSR section will be moved. We can review this later. I still need to add a big table with all the atom candidates and make a complete pass to avoid duplication and inconsistencies. I did a first pass and addressed some of that already.

@@
 % \item \rqc  
-\item \rqd
+\item \rqd\castor{This one will probably be removed.}
 \end{enumerate}
 
-Answering these research questions has several implications.
-For instance, we can either generalize or refute the perceptions about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)).
-In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice.
+By answering these research questions, we can either generalize or refute the findings about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)).
+In addition, answer to the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice\castor{This should probably be removed.}.
 Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions.
 %---though we postpone these results to future research work.
-We conduct a mixed-methods study to answer these questions, including a survey, a set of interviews, and a study for mining JavaScript software repositories.
+We conduct a mixed-methods study to answer these questions.
+It includes two independently designed and conducted experiments\castor{I would rather use the term ``experiment''.}: a replication of the first study on atoms of confusion, but targeting JavaScript programs, and a new study using a latin-square counterbalancing experimental design to control for subject experience.
+These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ1).
+They investigate whether or not code snippets that contain atom candidates produce a higher misunderstanding rate when programmers try to predict their outcome.\castor{I've removed the part about time because, in order for the first study to be a replication, we need to compute time differently, accounting for incorrect responses as well.}% and (b) require more time for programmers to predict their output.
+We leverage two independently-developed studies with different designs to more rigorously verify the impact of atoms of confusion in code comprehension activities. 
+We also conduct a set of interviews with developers to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2).
 
-\subsection{Survey Study Setting}\label{sec:meth:survey}
+\subsection{Replication Study Setting}\label{sec:meth:replication}
 
-In the first study, we explore the impact of atom candidates on
-understanding JavaScript code (goal of research question RQ1).	
-To this end, we investigate whether or not code snippets that
-contain atom candidates (a) produce a higher
-misunderstanding rate when programmers
-try to predict their outcome and (b) require more time for programmers to predict their output.
-Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first selected a set of 9+1 atom candidates: nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion).
-Our supplementary material discusses and shows examples of the atoms we consider in our research.
+We call the first study we performed the \textit{``Replication Study''} because it is a partial replication~\cite{Juristo2012} of the first study on atoms of confusion, conducted by Gopstein et.
+al~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, but targeting JavaScript instead of the C language.
+Gopstein et.
+al.
+employed a randomized, partial counterbalancing experimental design on a sample of 73 subjects, mostly students, with three or more months of experience in C or C++.
+The control variable consisted of a tiny program containing a single atom candidate and the treatment a functionally-equivalent version that does not contain the atom.
+The main dependent variable is whether subjects are able to correctly determine the output of the programs.
+The other dependent variable is the time required to correctly vs.
+incorrectly determine the out of the programs.
+The study had the null hypothesis that \textit{``code from both control and treatment groups can be hand-evaluated with equal accuracy''} and the alternate hypothesis that \textit{``the existence of an atom of confusion causes more errors than other code in hand-evaluated outputs.''}.
+For each atom candidate, six programs were built, three versions including the atom candidate and three versions not including it.
+Each subject was exposed to four programs for every atom candidate, two including an atom and two not including an atom.
 
-\subsubsection*{Survey design}
+\subsubsection*{Experimental design}
 
-The design of our first study blocks two variables
+In this study we adopt the same experimental design, control and treatment variables, dependent variable, and hypotheses.
+We have also built six programs for each atom candidate, three with and three without the atom.
+Considering the elements that can vary in a replication study, defined by Juristo and Gomez~\cite{Juristo2012}, in this study we employ the same experimental protocol and construct operationalizations.
+We have different population properties, experimenters, and a different site.
+In addition, since we focus on a different language, some atoms that exist in C cannot by construction exist in JavaScript programs.
+Conversely, JavaScript exhibits atom candidates that do not exist in C.
+
+The design of the replication study considers two treatments: the presence or absence of atoms candidates within the programs.
+To identify atom candidates, we start out by selecting every atom candidate from Gopstein's study that can also happen in JavaScript.
+That excludes atoms related to pointers and preprocessor directives.
+Furthermore, we conduct an informal search of programming forums, solutions to Code Golf~\footnote{https://codegolf.stackexchange.com/} problems, code style guidelines, and expert knowledge to identify new atom candidates.
+Overall, the replication study analyzes 24 atoms, 15 from the original study and 9 that are new and specific to JavaScript.
+Table~\ref{tab:atoms_replication} presents the complete list.
+
+For each atom candidate, we write six short programs, three including and three not including the atom candidate.
+The former are the control for the study or the obfuscated versions of the programs.
+The alternative, atom-free, versions of the programs are the treatment, or clean, versions.
+Each subject is exposed to one obfuscated and one clean version of each atom, totalizing 48 programs, and should determine their outputs.
+As mentioned before, we measure answer correctness and the time to answer questions correctly vs.
+incorrectly.
+We control for learning effects~\cite{Neely:1991:SPE} in three manners.
+First, by having multiple obfuscated and clean versions for each atom and presenting only one of each per subject.
+Second, by presenting the programs in a random order.
+Third, by presenting obfuscated and clean versions corresponding to the same atom candidate with at least 11 other programs in between, in  accordance to the original experimental protocol~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
+
+\subsubsection*{Study instrument}
+
+The study is conducted by means of a web application.
+As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
+Based on the feedback of the participants of the pilot, we decided to present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.
+Initially, our plan was to present two of each.
+
+We organize the study in three sections.
+In the first one we present instructions and also a check button whose checking meant users agreed that all collected data would be used anonymously and solely for research purposes.
+The instructions explain how the survey works and asks them to dedicate their attention to it.
+We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters.
+
+The second section presents the programs, one at a time.
+For each one, there is a text box where the answer should be written.
+There is also an ``I do not know'' button, which, when clicked, leads the subject to the next question.
+In our setting, ``I do not know'' was treated as a wrong answer.
+To reduce the likelihood of a subject leveraging attempting to execute the programs, the web application verifies whether the subject attempts to change tabs or windows and presents the programs as images, instead of text.
+Upon submitting an answer for a particular program, the subject is automatically led to a similar page, containing the next one.
+
+We do not provide feedback about the time subjects take to answer each question.
+We also do not tell them whether their answers are correct or not.
+This aims at avoiding introducing bias for future respondents.
+Since we posted the survey on social media platforms, possible threats could have arisen if we gave respondents instant feedback.
+
+\subsubsection*{Study audience}
+
+We posted the survey on the authors' social media platforms, on Brazilian CS departments' mailing lists, and on two programming subreddits.
+We explained our research purposes and asked developers to take the survey only if they have some familiarity with JavaScript.
+We collected 70 responses.
+The average survey respondent has a mean 26 months of experience in JavaScript, with a median of 12.
+
+\subsection{Latin-Square Study Setting}\label{sec:meth:survey}
+
+This experiment complements the replication study by using a different experimental design and controlling for one additional variable, the experience of the subjects.
+For this study we employ a subset of the atom candidates used in the replication study.
+More specifically, we
+%Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first 
+select a set of 9+1 atom candidates: nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion)\castor{I think the name of this atom is inconsistent between the studies}.
+Our supplementary material discusses and shows examples of the atoms we consider in our research.\castor{We should probably do this here or just refer to the table that I intend to insert for the replication study.}
+
+\subsubsection*{Experimental design}
+
+The design of our second study blocks two variables
 (developer experience and the code snippets) and
@@
 
-\subsubsection*{Survey Instrument}
+\subsubsection*{Study instrument}
 
-We implemented our survey as a web application.
-As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
+We implemented our survey as a web application and carried out a pilot, similarly to the replication study. 
+%As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
 Undergraduate students and professional colleagues took the pilot survey.
@@
 Upon submitting their answer for a particular question, the subject was automatically led to a similar page, containing another snippet.
-
-We decided not to provide feedback about the time students took to answer each question.
-We also did not tell them whether their answers were correct or not.
-Our main concern was to avoid introducing bias for future respondents.
-Since we posted the survey on a social media platform, possible threats could have arisen if we gave respondents instant feedback.
+Similarly to the replication study, we do not provide feedback about time and correctness to the subjects.
 
@@
 
-\subsubsection*{Survey audience}
+\subsubsection*{Study audience}
 

### 2022-03-29 3_Method.tex
For this change, I did three things. First, tried to make the two sections describing the experiments more consistent. Second, changed the tense of the sections on the experiments and the interview to the preset. Third, changed uses of snippet to program. The rationale for the latter is that, to me, snippet gives the idea that they are part of something else. However, they are self-contained, albeit small, programs.

@@
 We conduct a mixed-methods study to answer these questions.
-It includes two independently designed and conducted experiments\castor{I would rather use the term ``experiment''.}: a replication of the first study on atoms of confusion, but targeting JavaScript programs, and a new study using a latin-square counterbalancing experimental design to control for subject experience.
+It includes two independently designed and conducted experiments\castor{I would rather use the term ``experiment''.}: a replication of the first study on atoms of confusion, but targeting JavaScript programs, and a new study using a latin square counterbalancing experimental design to control for subject experience.
 These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ1).
-They investigate whether or not code snippets that contain atom candidates produce a higher misunderstanding rate when programmers try to predict their outcome.\castor{I've removed the part about time because, in order for the first study to be a replication, we need to compute time differently, accounting for incorrect responses as well.}% and (b) require more time for programmers to predict their output.
+They investigate whether or not programs that contain atom candidates produce a higher misunderstanding rate when programmers try to predict their outcome.\castor{I've removed the part about time because, in order for the first study to be a replication, we need to compute time differently, accounting for incorrect responses as well.}% and (b) require more time for programmers to predict their output.
 We leverage two independently-developed studies with different designs to more rigorously verify the impact of atoms of confusion in code comprehension activities. 
@@
 
-We call the first study we performed the \textit{``Replication Study''} because it is a partial replication~\cite{Juristo2012} of the first study on atoms of confusion, conducted by Gopstein et.
+We call the first study we perform the \textit{``Replication Study''} because it is a partial replication~\cite{Juristo2012} of the first study on atoms of confusion, conducted by Gopstein et.
 al~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, but targeting JavaScript instead of the C language.
@@
 al.
-employed a randomized, partial counterbalancing experimental design on a sample of 73 subjects, mostly students, with three or more months of experience in C or C++.
-The control variable consisted of a tiny program containing a single atom candidate and the treatment a functionally-equivalent version that does not contain the atom.
+employ a randomized, partial counterbalancing experimental design on a sample of 73 subjects, mostly students, with three or more months of experience in C or C++.
+The control variable consists of a tiny program containing a single atom candidate and the treatment a functionally-equivalent version that does not contain the atom.
 The main dependent variable is whether subjects are able to correctly determine the output of the programs.
 The other dependent variable is the time required to correctly vs.
-incorrectly determine the out of the programs.
-The study had the null hypothesis that \textit{``code from both control and treatment groups can be hand-evaluated with equal accuracy''} and the alternate hypothesis that \textit{``the existence of an atom of confusion causes more errors than other code in hand-evaluated outputs.''}.
+incorrectly determine the outputs.
+The study has the null hypothesis that \textit{``code from both control and treatment groups can be hand-evaluated with equal accuracy''} and the alternate hypothesis that \textit{``the existence of an atom of confusion causes more errors than other code in hand-evaluated outputs.''}.
 For each atom candidate, six programs were built, three versions including the atom candidate and three versions not including it.
-Each subject was exposed to four programs for every atom candidate, two including an atom and two not including an atom.
+Each subject is exposed to four programs for every atom candidate, two including an atom and two not including an atom.
 
@@
 For each atom candidate, we write six short programs, three including and three not including the atom candidate.
-The former are the control for the study or the obfuscated versions of the programs.
+The former are the control for the study, or the obfuscated versions of the programs.
 The alternative, atom-free, versions of the programs are the treatment, or clean, versions.
-Each subject is exposed to one obfuscated and one clean version of each atom, totalizing 48 programs, and should determine their outputs.
+Each subject is exposed to one obfuscated and one clean version of each atom, totalling 48 programs, and should determine their outputs.
 As mentioned before, we measure answer correctness and the time to answer questions correctly vs.
@@
 As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
-Based on the feedback of the participants of the pilot, we decided to present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.
+Based on the feedback of the participants of the pilot, we present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.
 Initially, our plan was to present two of each.
@@
 We organize the study in three sections.
-In the first one we present instructions and also a check button whose checking meant users agreed that all collected data would be used anonymously and solely for research purposes.
+In the first one we present instructions and also a check button whose checking means users agree that all collected data will be used anonymously and solely for research purposes.
 The instructions explain how the survey works and asks them to dedicate their attention to it.
-We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters.
+We stress to participants the importance of not using any aids during the survey, such as online or console interpreters.
 
 The second section presents the programs, one at a time.
-For each one, there is a text box where the answer should be written.
-There is also an ``I do not know'' button, which, when clicked, leads the subject to the next question.
-In our setting, ``I do not know'' was treated as a wrong answer.
-To reduce the likelihood of a subject leveraging attempting to execute the programs, the web application verifies whether the subject attempts to change tabs or windows and presents the programs as images, instead of text.
+For each one, there is a text box where the answer should be written and a ``Next'' button.
+If the subject leaves the text box empty and presses ``Next'', this is treated as a wrong answer.
+To reduce the likelihood of a subject attempting to execute the programs, the web application verifies whether the subject tries to change tabs or windows and presents a pop-up message if that is the case.
+Furthermore, it presents the programs as images, instead of text.
 Upon submitting an answer for a particular program, the subject is automatically led to a similar page, containing the next one.
@@
 We collected 70 responses.
-The average survey respondent has a mean 26 months of experience in JavaScript, with a median of 12.
+The average survey respondent has a mean 26 months of experience in JavaScript, with a median of 12.\castor{We should not say this here unless we are able to do the same for the Latin square study.
+}
 
-\subsection{Latin-Square Study Setting}\label{sec:meth:survey}
+\subsection{Latin Square Study Setting}\label{sec:meth:survey}
 
@@
 
-The design of our second study blocks two variables
-(developer experience and the code snippets) and
-considers two treatments: the presence or absence of 
-atoms candidates within the code snippets. 
-To achieve such a design goal, controlling the effect of experience and individual code snippets, we resorted to the \textit{Latin Square Design} \cite{Hunter-Experimenters}.
-Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of code snippets.
+The design of our second study blocks two variables (subject experience and the programs) and
+considers two treatments: the presence or absence of atom candidates within the programs. 
+To achieve such a design goal, controlling the effect of experience and individual programs, we resort to the \textit{Latin Square Design} \cite{Hunter-Experimenters}.
+Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of programs.
 The design of each square (a replica) is such that no treatment is repeated in the same row or column.
 For example, considering that we have a set of 10
-code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about non-confusing code snippets, they will only be presented with non-confusing versions of the code samples $s_6, s_7,..., s_{10}$.
-Furthermore, a given subject (P2), who constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing snippets for $s_6, s_7,..., s_{10}$.
-By doing that, we guarantee that all versions of the code snippets are contained within each square, and that each configuration occurs only once within a square.
+code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about non-confusing programs, they will only be presented with non-confusing versions of the code samples $s_6, s_7,..., s_{10}$.
+Furthermore, a given subject (P2), who constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing programs for $s_6, s_7,..., s_{10}$.
+By doing that, we guarantee that all versions of the programs are contained within each square, and that each configuration occurs only once within a square.
 Figure \ref{fig:latinsquare} offers a visual representation of the concept.
@@
       a replica in our study. Each replica comprises two participants (square rows, e.g., P1 and P2) 
-      and two sets of code snippets (CS-1 and CS-2). We randomly apply the 
+      and two sets of programs (CS-1 and CS-2). We randomly apply the 
       treatments (atom or non-atom code) to the cells of the squares.} 
@@
 
-Having selected 10 atom candidates in our first study, we wrote 10 short programs, one containing each atom candidate.
-We call them the confusing versions of the programs.
-We also wrote corresponding short programs without the confusing idioms and constructs and we call them the clean versions of the programs.
-Overall, the survey employed 20 code snippets, 10 confusing and 10 clean.
-In order to reduce the cognitive effort, each subject was asked to predict the output of 10 code snippets, five confusing and five clean.
-The order in which the questions were presented was randomized.
+For each of the 10 selected atom candidates, we write one short program containing the atom.
+As in the replication study, we call them the confusing versions of the programs.
+We also write corresponding short programs without the atom candidates and call them the clean versions of the programs.
+Overall, the experiment employed 20 programs, 10 confusing and 10 clean.
+In order to reduce the cognitive effort, each subject was asked to predict the output of 10 programs, five confusing and five clean.
+The order in which they are presented is randomized.
 By doing this, we seek to minimize the chances of subjects being aware that the current listing they are analyzing contains (or not) atoms of confusion.
-That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets do not).
-We measured answer correctness and the total time each participant needed to answer the survey's questions.
+That is, each subject should indicate what would be the outcomes of the programs, some of them having atoms of confusion (while other programs do not).
+Since participants are not exposed to obfuscated and clean versions corresponding to the same atom candidate, learning effect is not possible.
+We measure answer correctness and the total time each participant requires to participate in the experiment.
 
@@
 Some users reported layout defects, and many reported that the landing page did not explain the survey well enough.
-We also spotted minor issues with our routines to create and populate the Latin Squares.
+We also spotted minor issues with our routines to create and populate the latin squares.
 
-We organized the survey in three sections.
+We organize the survey in three sections.
 The first section aims to characterize the subjects, asking their age, education level, and programming experience.
-We also included a check button, whose checking meant users agreed that all collected data would be used solely for research purposes.
-In the second section, we presented to the participants a small set of instructions, where we explained how the survey worked and asked them to dedicate their attention to it.
-We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters.
-For each question page, we kept track of whether or not the subjects switched windows.
+We also include a check button, whose checking means users agree that all collected data will be used solely for research purposes.
+In the second section, we present to the participants a small set of instructions, where we explain how the survey works and ask them to dedicate their attention to it.
+We stress to participants the importance of not using any aids during the survey, such as online or console interpreters.
+For each question page, we kept track of whether or not the subjects switch windows.
 
-The last section of the survey presented a sequence of ten questions, each containing a code snippet.
-For each question, there was a text box where the answer should be written.
-There was also an ``I do not know'' button, which, when clicked, led the subject to the next question.
-In our setting, ``I do not know'' was treated as a wrong answer.
-The code snippets were presented as images copied from a text editor, so as to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter.
-Upon submitting their answer for a particular question, the subject was automatically led to a similar page, containing another snippet.
+The last section of the survey presents a sequence of ten questions, each containing a program.
+For each question, there is a text box where the answer should be written.
+There is also an ``I do not know'' button, which, when clicked, leads the subject to the next question.
+In our setting, ``I do not know'' is treated as a wrong answer.
+The programs are presented as images copied from a text editor, so as to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter.
+Upon submitting their answer for a particular question, the subjects are automatically led to a similar page, containing another program.
 Similarly to the replication study, we do not provide feedback about time and correctness to the subjects.
@@
 Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares.
-We collected significant data on time taken and discrepancies in answer correctness between confusing and non-confusing versions of the snippets.
+We collected significant data on time taken and discrepancies in answer correctness between confusing and non-confusing versions of the programs.
 Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the survey.
@@
 
-To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion (research question RQ2).
-We also asked each participant if they knew of any other JavaScript-specific construct or idiom that they regarded as being likely to make the code hard to understand (research question RQ3).
+To complement the two experiments, we perform semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding programs containing atoms of confusion (research question RQ2).
+We also ask each participant if they know of any other JavaScript-specific construct or idiom that they regard as being likely to make the code hard to understand.% (research question RQ3).
 In this section we detail the protocol we followed to conduct the interviews and to analyze the results.
@@
 
-\subsubsection*{Participant Selection} We invited the participants of the interviews using a snowballing technique.
-That is, starting from our network of contacts, we invited an initial set of candidates to take part in our survey.
-From this initial list, we asked for an indication of additional candidates.
-Our main selection criterion was that all participants should have been working with JavaScript in their daily professional activities.
-We invited a total of 17 developers, and 15 of them agreed to participate.
+\subsubsection*{Participant Selection} We invite the participants of the interviews using a snowballing technique.
+That is, starting from our network of contacts, we invite an initial set of candidates to take part in our survey.
+From this initial list, we ask for an indication of additional candidates.
+Our main selection criterion is that all participants should have been working with JavaScript in their daily professional activities.
+We invite a total of 17 developers, and 15 of them agree to participate.
 %We conducted the interviews for a period of two weeks.
 
-\subsubsection*{Interview Process} We conducted semi-structured interviews using web conferencing software.
-We recorded all the interviews with the consent of the participants.
-On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes.
-Two of the researchers conducted the interviews, and a third one listened to all the recordings to cross-validate the collected data.
-The interviews had three main parts.
-In the first one, we asked the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years, and other programming languages they have worked with.
+\subsubsection*{Interview Process} We conduct semi-structured interviews using web conferencing software.
+We record all the interviews with the consent of the participants.
+On average, the interviews last 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes.
+Two of the researchers conduct the interviews, and a third one listens to all the recordings to cross-validate the collected data.
+The interviews have three main parts.
+In the first one, we ask the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years, and other programming languages they have worked with.
 Table \ref{pinterview} summarizes this demographic information.
@@
 
-In the second part of the interview our aim was to allow the subjects to describe their JavaScript experience, as well as to allow them to reveal any JavaScript constructs they regarded as innately confusing.
-This would allow us to identify potential atom candidates that are more specific to the JavaScript language.
-Examples of questions we explored in this section include: \emph{Does JavaScript favor developers to produce code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as especially confusing?}
+In the second part of the interview our aim is to allow the subjects to describe their JavaScript experience, as well as to allow them to reveal any JavaScript constructs they regard as innately confusing.
+This allows us to identify potential atom candidates that are more specific to the JavaScript language.
+Examples of questions we explore in this section include: \emph{Does JavaScript favor developers to produce code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as especially confusing?}
 
-In the third part of the interview, participants were shown the code snippets that were used in the first study.
-The snippets were presented in pairs, where each pair contained a clean and a confusing version.
+In the third part of the interview, participants are shown pairs of programs that were used in the first study, where each pair contains a clean and a confusing version of the same atom candidate.
 The participants
 %were told both sides of each pair had the same output, and 
-were asked to evaluate which version of the code was easier to understand.
-To avoid introducing bias in the answers, the interviewers did not explain that one of the versions in each pair contained the atom under investigation.
-Subjects were just presented the pairs of snippets and allowed to take the necessary time to decide on the most readable snippet.
+are asked to evaluate which version of the code is easier to understand.
+To avoid introducing bias in the answers, the interviewers do not explain that one of the versions in each pair contains the atom under investigation.
+Subjects are just presented the pairs of programs and allowed to take the necessary time to decide on the most readable one.
 
@@
 
-We first transcribed each interview and then examined the broad distribution of the answers.
-Our goal was to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular.
-We next followed with an open-coding procedure, highlighting the main themes and quoting the answers of the participants.
+We first transcribe each interview and then examine the broad distribution of the answers.
+Our goal is to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular.
+We follow-up with an open-coding procedure, highlighting the main themes and quoting the answers of the participants.
 We present these results in Section~\ref{sec:interview-results}.

### 2022-03-31 3_Method.tex
Fixed some terminology issues about obfuscated and clean versions where we were employing different terms.

@@
 By answering these research questions, we can either generalize or refute the findings about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)).
-In addition, answer to the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice\castor{This should probably be removed.}.
+In addition, answering the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice\castor{This should probably be removed.}.
 Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions.
@@
 As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
-Based on the feedback of the participants of the pilot, we present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.
-Initially, our plan was to present two of each.
+Based on the feedback of the participants of the pilot, we present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.% Initially, our plan was to present two of each.
 
@@
 
-This experiment complements the replication study by using a different experimental design and controlling for one additional variable, the experience of the subjects.
+This experiment uses a different experimental design that makes it possible to control for the experience of the subjects.
 For this study we employ a subset of the atom candidates used in the replication study.
@@
 For example, considering that we have a set of 10
-code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about non-confusing programs, they will only be presented with non-confusing versions of the code samples $s_6, s_7,..., s_{10}$.
-Furthermore, a given subject (P2), who constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing programs for $s_6, s_7,..., s_{10}$.
+code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about clean programs, they will only be presented with clean versions of the code samples $s_6, s_7,..., s_{10}$.
+Furthermore, a given subject (P2), who constitutes the second row of our example square, will be asked questions about the clean versions for $s_1, s_2, ..., s_5$, and will answer questions about obfuscated programs for $s_6, s_7,..., s_{10}$.
 By doing that, we guarantee that all versions of the programs are contained within each square, and that each configuration occurs only once within a square.
@@
 For each of the 10 selected atom candidates, we write one short program containing the atom.
-As in the replication study, we call them the confusing versions of the programs.
+As in the replication study, we call them the obfuscated versions of the programs.
 We also write corresponding short programs without the atom candidates and call them the clean versions of the programs.
-Overall, the experiment employed 20 programs, 10 confusing and 10 clean.
-In order to reduce the cognitive effort, each subject was asked to predict the output of 10 programs, five confusing and five clean.
+Overall, the experiment employed 20 programs, 10 obfuscated and 10 clean.
+In order to reduce the cognitive effort, each subject was asked to predict the output of 10 programs, five obfuscated and five clean.
 The order in which they are presented is randomized.
@@
 Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares.
-We collected significant data on time taken and discrepancies in answer correctness between confusing and non-confusing versions of the programs.
+We collected significant data on time taken and discrepancies in answer correctness between obfuscated and clean versions of the programs.
 Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the survey.
@@
 
-In the third part of the interview, participants are shown pairs of programs that were used in the first study, where each pair contains a clean and a confusing version of the same atom candidate.
+In the third part of the interview, participants are shown pairs of programs that were used in the first study, where each pair contains a clean and an obfuscated version of the same atom candidate.
 The participants

### 2022-03-31 discussion.tex
remocao de texto nao essencial para ajustar as limitacoes de espaco
Fixed some terminology issues about obfuscated and clean versions where we were employing different terms.

@@
 of the atom candidates.
-Altogether, we answer our first research question
-\emph{\rqa}:
+Altogether, we answer our first research question.
 
@@
 In summary, the results of the second study (interviews) allow
-us to answer the second (\rqb) research question.
+us to answer the second research question.
 
@@
 the 72 projects we analyzed. Furthermore, at least two of them are used intensively, more than once for every 200 lines of code. In summary, the third study
-allows us to answer the fourth research
-question (\emph{\rqd}).
+allows us to answer the third research
+question.
 
@@
   might improve the readabiliy of JavaScript code substantially. 
-(\emph{\rqc}) 
 \end{mh}

### 2022-03-31 main.tex
Fixed some terminology issues about obfuscated and clean versions where we were employing different terms.

--- 2021-10-21 main.tex
@@
 In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language.
-We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study.
+We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and an activity of mining open source JavaScript repositories (MSR).
 Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript.

### 2022-03-31 method_replication_study.tex
Fixed some terminology issues about obfuscated and clean versions where we were employing different terms.

@@
 The study is conducted by means of a web application.
-We conducted an initial pilot to find bugs with the application and to get general impressions.
+We carried out an initial pilot to find bugs with the application and to get general impressions.
 Based on the feedback of the participants of the pilot, we decided to present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.
 Initially, our plan was to present two of each.
-To reduce the likelihood of a subject simply running the program, the web application verifies whether the subject attempts to change tabs or windows and presents the programs as images, instead of text.
 
-We organize the survey in three sections.
+We organize the study in three sections.
 In the first one we present instructions and also a check button whose checking meant users agreed that all collected data would be used anonymously and solely for research purposes.
@@
 We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters.
-For each question page, we kept track of whether or not the subjects switched windows.
 
-The last section of the survey presented a sequence of ten questions, each containing a code snippet.
-For each question, there was a text box where the answer should be written.
-There was also an ``I do not know'' button, which, when clicked, led the subject to the next question.
+The second section presents the programs, one at a time.
+For each one, there is a text box where the answer should be written.
+There is also an ``I do not know'' button, which, when clicked, leads the subject to the next question.
 In our setting, ``I do not know'' was treated as a wrong answer.
-The code snippets were presented as images copied from a text editor, so as to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter.
-Upon submitting their answer for a particular question, the subject was automatically led to a similar page, containing another snippet.
+To reduce the likelihood of a subject leveraging attempting to execute the programs, the web application verifies whether the subject attempts to change tabs or windows and presents the programs as images, instead of text.
+Upon submitting an answer for a particular program, the subject is automatically led to a similar page, containing the next one.
 
-We decided not to provide feedback about the time subjects took to answer each question.
-We also did not tell them whether their answers were correct or not.
-Our main concern was to avoid introducing bias for future respondents.
-Since we posted the survey on a social media platform, possible threats could have arisen if we gave respondents instant feedback.
+We do not provide feedback about the time subjects take to answer each question.
+We also do not tell them whether their answers are correct or not.
+This aims at avoiding introducing bias for future respondents.
+Since we posted the survey on social media platforms, possible threats could have arisen if we gave respondents instant feedback.
 

### 2022-07-02 method_replication_study.tex
Added data for the replication study, specifically about the number of errors and correct guesses for obfuscated and clear versions and also about time require for correct and incorrect responses.

@@
 There is also an ``I do not know'' button, which, when clicked, leads the subject to the next question.
-In our setting, ``I do not know'' was treated as a wrong answer.
-To reduce the likelihood of a subject leveraging attempting to execute the programs, the web application verifies whether the subject attempts to change tabs or windows and presents the programs as images, instead of text.
+In our setting, ``I do not know'' is treated as a wrong answer.
+To reduce the likelihood of a subject attempting to execute the programs, the web application verifies whether the subject attempts to change tabs or windows and presents the programs as images, instead of text.
 Upon submitting an answer for a particular program, the subject is automatically led to a similar page, containing the next one.

### 2022-07-02 calculation_time_replication_study.py

@@ -1,0 +1,16 @@
+import pandas as pd
+import numpy as np
+import scipy.stats as stats
+
+atoms = pd.read_csv('results_JS_replication_study.csv')
+atoms.dropna(thresh=10, inplace=True)
+times = atoms.loc[:, 'Question 1 time':'Question 48 time']
+correct = atoms.loc[:,'Question 1 correct':'Question 48 correct']
+correct.columns = times.columns
+correctTimesTemp = times.where(correct)
+correctTimes = pd.Series(correctTimesTemp.values.flatten()).dropna()
+incorrectTimesTemp = times.where(np.logical_not(correct))
+incorrectTimes = pd.Series(incorrectTimesTemp.values.flatten()).dropna()
+correctTimes = correctTimes[correctTimes.map(lambda t: type(t) != type('ok'))]
+incorrectTimes = incorrectTimes[incorrectTimes.map(lambda t: type(t) != type('ok'))]
+print(stats.ranksums(correctTimes, incorrectTimes))

### 2022-07-11 3_Method.tex
The Jupyter Notebook with the analysis of the Repeated Measures Study is finally ready. I have started including the results in the paper.

@@
 
+\castor{This will become the Repeated Measures Study and we need to introduce the atoms that did not originally appear in Gopstein's study, maybe all of them.}
 We call the first study we perform the \textit{``Replication Study''} because it is a partial replication~\cite{Juristo2012} of the first study on atoms of confusion, conducted by Gopstein et.

### 2022-07-11 4_Result.tex
The Jupyter Notebook with the analysis of the Repeated Measures Study is finally ready. I have started including the results in the paper.

@@
 
-Our first study investigates the impact of atoms of confusion while developers try to understand JavaScript code.
-We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
-and \emph{time} (how long to provide a correct answer).
-We received full answers to our survey from 140 participants (a total of 70 replicas).
-All participants had taken at least some university course or hold a bachelor
-degree or equivalent, meaning the participants have had some level of formal education in programming. In addition, 21 participants hold a master's degree and three a doctorate degree. 
-Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.
-Accordingly, we characterize the effect of atoms of confusion in JavaScript
-code taking into account the perceptions of both novice and experienced developers.
+%-----------------------------------------------------------------------
+\subsection{Repeated measures study}\label{sec:repeated}
+
+In the repeated measures study, we investigate whether 24 atom candidates and their clear counterparts impact readability.
+We examine mainly the cases where participants are able to successfully predict the output of a clear version but fail to do so for the corresponding obfuscated version and vice-versa.
+In addition, similarly to Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, we verify whether there is a trade-off between correctness and time irrespective of the presence of atoms; we check if participants who correctly predict the outputs of the programs tend to do so more slowly than the ones that make incorrect predictions.
+
+Overall, this study had 70 participants. Out of these, 67 have received university-level training in programming, 4 hold a PhD degree, 7 hold a master's degree, and 11 hold a bachelor's degree. 
+The median programming experience of the participants is four years (mean 6.5 years) and the median experience with JavaScript programming is one year (mean 2.5 years). 
+% General programming: median: 48, 1st quartile: 22,5, 3rd quartile: 91,5
+% JavaScript programming: median: 12, 1st quartile: 4, 3rd: 24
+
+%-----------------------------------------------------------------------
+\subsection{Correctness Analysis}
+
+\subsubsection*{Exploratory Data Analysis}
+
+For each of the 24 atom candidates, the participants predicted the outputs of two versions, one clean (control) and one obfuscated (treatment), as per the definition of a repeated measures design~\cite{Hunter-Experimenters}.
+Thus, each participant examined 48 code snippets.
+We also measured the time required by the participants to provide their answers.
+For most cases, participants made more mistakes when attempting to predict the outcomes of the obfuscated versions.
+
+Table~\ref{tab:diff-correctness-repeated} presents the number of correct predictions for the clean and obfuscated versions of each atom candidate, as well as the percentage difference.
+In only three cases the participants predicted results correctly more often for the obfuscated versions, Arithmetic as Logic, Logic as Control Flow, and Property Access.
+The former two were empirically validated as atoms of confusion in the study of Gopstein and colleagues~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
+For seven of the atom candidates, the number of correct predictions for the clean versions is more than twice the corresponding number for the obfuscated versions.
+Some of these atoms are common programming patterns in JavaScript programs, such as pre- and post-increments and object destructuring.
+The difference was even greater for Type Conversion (+440\%) and Change of Literal Encoding (+300\%).
+\begin{table}[htbp]
+\caption{Summary of the correctness analysis}
+\label{tab:diff-correctness-repeated}
+\centering{
+\begin{scriptsize}
+\begin{tabular}{lccc} \toprule
+  Atom & Confusing & Clean & $\Delta$(\%)  \\ \midrule
+Type Conversion           &            5.0 &      27.0 &                 +440.0 \\
+Change of Literal Encoding &           9.0 &      36.0 &                 +300.0 \\
+Comma Operator            &            8.0 &      29.0 &                 +262.0 \\
+Object Destructuring      &           15.0 &      42.0 &                 +180.0 \\
+Assignment As Value       &           18.0 &      43.0 &                 +139.0 \\
+Pre-Increment             &           13.0 &      28.0 &                 +115.0 \\
+Post-Increment            &            8.0 &      17.0 &                 +112.0 \\
+Array Destructuring       &           15.0 &      28.0 &                 +87.0 \\
+Indentation no Braces     &           30.0 &      50.0 &                 +67.0 \\
+Omitted Curly Braces      &           35.0 &      51.0 &                 +46.0 \\
+Object Spread             &           21.0 &      30.0 &                 +43.0 \\
+Array Spread              &           22.0 &      31.0 &                 +41.0 \\
+Automatic Semicolon Insertion &       15.0 &      21.0 &                 +40.0 \\
+Infix Operator Precedence &           47.0 &      53.0 &                 +13.0 \\
+Arrow Function            &           54.0 &      58.0 &                  +7.0 \\
+Implicit Predicate        &           41.0 &      44.0 &                  +7.0 \\
+Indentation With Braces   &           53.0 &      55.0 &                  +4.0 \\
+Conditional Operator      &           60.0 &      62.0 &                  +3.0 \\
+Constant Variables        &           59.0 &      61.0 &                  +3.0 \\
+Dead Unreachable Repeated &           63.0 &      64.0 &                  +2.0 \\
+Repurposed Variables      &           23.0 &      23.0 &                   0.0 \\
+Arithmetic as Logic       &           52.0 &      47.0 &                 -10.0 \\
+Logic as Control Flow     &           33.0 &      28.0 &                 -15.0 \\
+Property Access           &           51.0 &      42.0 &                 -18.0 \\
+\bottomrule  
+\end{tabular}
+\end{scriptsize}
+}
+\end{table}
+
+\subsubsection*{Statistical analysis}
+
+Data is categorical, paired, dependent.
+This is analogous to clinical trials where the same patient receives two different treatments.
+These treatments are spaced to reduce interaction between them.
+For this, McNemar.
+Odds ratio as a measure of effect size.
 
@@
 
+%-----------------------------------------------------------------------
+\subsection{Latin-square study}\label{sec:latin}
+
+Our first study investigates the impact of atoms of confusion while developers try to understand JavaScript code.
+We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
+and \emph{time} (how long to provide a correct answer).
+We received full answers to our survey from 140 participants (a total of 70 replicas).
+All participants had taken at least some university course or hold a bachelor degree or equivalent, meaning the participants have had some level of formal education in programming.
+In addition, 21 participants hold a master's degree and three a doctorate degree.
+Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.
+Accordingly, we characterize the effect of atoms of confusion in JavaScript code taking into account the perceptions of both novice and experienced developers.
+
 \subsection{Misunderstanding Rate Analysis}
+\castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me which rate is being referred to.}
 
@@
 
-As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten
-code snippets, from which five were in their confusing versions, whilst the other five contained clean versions of the code snippets (i.e., without the confusing constructs and idioms).
+As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten code snippets, from which five were in their confusing versions, whilst the other five contained clean versions of the code snippets (i.e., without the confusing constructs and idioms).
 The participants should provide the expected outcomes of the code snippets.

### 2022-07-15 3_Method.tex
First version with the results for the repeated measures study.

--- 2022-03-31 3_Method.tex
@@
 
+\castor{This will become the Repeated Measures Study (aka Within Subject) and we need to introduce the atoms that did not originally appear in Gopstein's study, maybe all of them.}
 We call the first study we perform the \textit{``Replication Study''} because it is a partial replication~\cite{Juristo2012} of the first study on atoms of confusion, conducted by Gopstein et.

### 2022-07-15 4_Result.tex
First version with the results for the repeated measures study.

@@
 
-\section{Results of the Survey}
+%\section{Results of the Survey}
 \label{sec:survey-resuts}
@@
 %-----------------------------------------------------------------------
-\subsection{Repeated measures study}\label{sec:repeated}
+\section{Results of the Repeated Measures Study}\label{sec:repeated}
 
-In the repeated measures study, we investigate whether 24 atom candidates and their clear counterparts impact readability.
-We examine mainly the cases where participants are able to successfully predict the output of a clear version but fail to do so for the corresponding obfuscated version and vice-versa.
-In addition, similarly to Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, we verify whether there is a trade-off between correctness and time irrespective of the presence of atoms; we check if participants who correctly predict the outputs of the programs tend to do so more slowly than the ones that make incorrect predictions.
+In this study, we investigate whether 24 atom candidates impact readability.
+We examine how the performance of the participants differ when predicting the outputs of the clean and obfuscated versions of small code snippets (Section~\ref{sec:repeated:correct}).
+Based on those results we establish which candidates can actually be considered atoms of confusion.
+In addition, similarly to Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, we verify whether there is a trade-off between correctness and time irrespective of the presence of atoms; we check if participants who correctly predict the outputs of the programs tend to do so more slowly than the ones that make incorrect predictions (Section~\ref{sec:repeated:time}).
 
@@
 %-----------------------------------------------------------------------
-\subsection{Correctness Analysis}
+\subsection{Correctness Analysis}\label{sec:repeated:correct}
 
-\subsubsection*{Exploratory Data Analysis}
-
+\subsubsection*{Exploratory Data Analysis. }
 For each of the 24 atom candidates, the participants predicted the outputs of two versions, one clean (control) and one obfuscated (treatment), as per the definition of a repeated measures design~\cite{Hunter-Experimenters}.
@@
 The difference was even greater for Type Conversion (+440\%) and Change of Literal Encoding (+300\%).
-\begin{table}[htbp]
-\caption{Summary of the correctness analysis}
+
+\begin{table}[tb]
+\caption{Summary of the correctness analysis.}
 \label{tab:diff-correctness-repeated}
@@
 
-Data is categorical, paired, dependent.
-This is analogous to clinical trials where the same patient receives two different treatments.
-These treatments are spaced to reduce interaction between them.
-For this, McNemar.
-Odds ratio as a measure of effect size.
+Since this study uses a within-subject design, the data is paired and dependent.
+Our goal is to verify whether there is a significant difference in the performance of the same participants when examining clean and obfuscated versions of programs.
+The analyzed data is also dichotomous, since each participant either correctly predicts the output of a program or does not.
+Considering this scenario, we employ McNemar's test~\cite{McNemar:1947:NSE}, which is aimed at \textit{``judging the significance of the difference between correlated proportions''}.
 
-%Our audience comprises professional developers, and 93\% of the respondents have more than one year of experience with programming.
-% (see Figure~\ref{fig:xp}). 
-%Actually, 56\% of the participants have either between four and ten years of experience (37\%) or more than ten years of experience (19\%).
-Other 37\% of the respondents have between one and four years of experience.
+The ``McNemar test'' column of Table~\ref{tab:repeated:misunderstanding} indicates the p-values for the tests.
+These p-values are compared against an alpha of 0.0021, after applying Bonferroni correction (0.05/24).
+One of the atom candidates, Constant Variables, is omitted from the table because, due to the very low number of mistakes, it is not possible to perform the McNemar test.
+The table shows that eight of the 24 analyzed atom candidates can be considered atoms of confusion based on the results of our study, Object Destructuring, Type Conversion, Change of Literal Encoding, Comma Operator, Indentation No Braces, Assignment as Value, Pre-Increment, and Ommitted Curly Braces.
+Among these eight, only Object Destructuring is JavaScript-specific.
+The other seven atoms have been previously reported for the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and also for Java~\cite{Langhout:2021:ACJ} (with the exception of Comma Operator).
+For some candidates that have been previously identified as atoms, there was not a statistically significant difference.
+This is the case for Repurposed Variables, Logic as Control Flow, Implicit Predicate, and Conditional Operator.
 
-%% Figure~\ref{fig:degree} and Figure~\ref{fig:xp} show the distributions of the subjects,
-%% according to their education level and years of programming experience, respectively.
+We leverage the odds ratio (OR) as a measure of effect size, to quantify the meaningfulness~\cite{Ellis:2010:EGE} of these results.
+The odds ratio is an appropriate measure of effect size for scenarios where McNemar's test is applicable~\cite{Mangiafico:2016:SAE}.
+For example, odds ratio for Type Conversion is 12.
+This suggests that the odds of a participant correctly predicting the output of a clean version while incorrectly predicting the output of the corresponding obfuscated version is 12 times higher than the odds of correctly predicting the output of the obfuscated version and incorrectly predicting the output of the correct one.
 
-%% \begin{figure}[htb]
-%%       \centering
-%%       \includegraphics[width=\columnwidth]{images/dem-education-1.pdf}
-%%       \caption{Participants' Education Level}\label{fig:degree}
-%%   \end{figure}
+According to the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), for most of the atoms where there was a statistically significant difference the odds ratio can be considered high.
+The value of the odds ratio for Object Destructuring is $\infty$ because no participant correctly guessed the output for the obfuscated version while incorrectly guessing the output for the clean one, which results in a denominator of 0.
+Omitted Curly Braces exhibits an OR of 6.3, which is borderline between medium and high~\cite{Chen:2010:HBB}.
+Array Destructuring and Post-Increment are two atom candidates that deserve a special note.
+The difference in the performance of participants examining clean and obfuscated versions of these atoms was not statistically significant, after correction.
+Notwithstanding, effect sizes were large and medium, respectively.
+This is a result that hints at practical relevance in spite of the p-values~\cite{Ellis:2010:EGE}.
+Further investigation is required for these two atom candidates.
 
-%%   \begin{figure}[htb]
-%%       \centering
-%%       \includegraphics[width=\columnwidth]{images/dem-experience-1.pdf}
-%%       \caption{Participants' Experience} \label{fig:xp}
-%%   \end{figure}
-%Figure~\ref{fig:xp} shows that more than a half of
+\begin{table}[tb]
+\caption{Hypotheses Testing (correctness).
+Asterisks ($^{*}$) indicate a statistically significant difference.
+Using Bonferroni correction, a p-value is considered statistically significant if it is lower than 0.0021.}
+ \centering
+ \begin{tabular}{lrr}
+   \toprule
+Atom candidate & McNemar test &      odds ratio \\
+\midrule
+Object Destructuring      &   \textbf{$<$ 0.0001*} & $\infty$ \\
+Type Conversion           &   \textbf{$<$ 0.0001*} & 12.0 \\
+Change of Literal Encoding   &   \textbf{$<$ 0.0001*} & 28.0 \\
+Comma Operator            &   \textbf{$<$ 0.0001*} & 11.5 \\
+Indentation No Braces     &   \textbf{$<$ 0.0001*} &  11.0 \\
+Assignment As Value       &   \textbf{$<$ 0.0001*} &  9.3 \\
+Pre-Increment              &  \textbf{0.0003*} &    16.0 \\
+Omitted Curly Braces      &  \textbf{0.0009*} &  6.3 \\
+Array Destructuring       &  0.0023 &     7.5 \\
+Post-Increment              &  0.0225 &     5.5 \\
+Array Spread              &  0.0490 &    3.3 \\
+Object Spread             &  0.0636 &     2.8 \\
+Property Access           &  0.0636 &     2.8 \\
+Infix Operator Precedence &  0.1796 &     2.5 \\
+Arrow Function            &  0.2891 &     3.0 \\
+Arithmetic As Logic       &  0.3323 &  1.8 \\
+Automated Semicolon Insertion  &  0.3915 &  1.4 \\
+Logic As Control Flow     &  0.4421 &  1.5 \\
+%Constant Variables        &  0.5000 &     0.0 \\
+Implicit Predicate        &  0.6900 &  1.3 \\
+Indentation With Braces   &  0.7266 &  1.7 \\
+Conditional Operator      &  0.7905 &  1.3 \\
+Dead Unreachable Repeated &  1.0000 &     1.5 \\
+Repurposed Variables      &   1.0000 &     1.0 \\
+\bottomrule
+\end{tabular}
+\label{tab:repeated:misunderstanding}
+\end{table}
 
 %-----------------------------------------------------------------------
-\subsection{Latin-square study}\label{sec:latin}
+\subsection{Time Analysis}\label{sec:repeated:time}
+
+\subsubsection*{Exploratory Data Analysis}
+
+In this section we investigate whether there is a trade-off between correctness and time required to predict program outputs, similarly to the analysis performed by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
+For this analysis, we consider only whether participants correctly predict the outputs of programs or not, independently of the presence or absence of atoms, and the time required to do so.
+The null hypothesis we investigate is that there is no significant difference in time between correct and incorrect predictions.
+Intuitively, we would expect correct responses to take longer, on the basis that participants would spend more time carefully analyzing the programs before providing an answer.
+
+Table~\ref{tab:repeated:time} presents the mean and median times required by participants to provide answers.
+The data is based on the 1648 correct and 1572 incorrect answers collected in the study.
+Contrary to our intuition, the mean time required by participants to provide an incorrect answer is 25\% greater than the time required to provide a correct one and the median time is 20\%.
+greater.
+The standard deviations suggest that there is great variation in the times required by different participants, for different code snippets.
+Notwithstanding, the standard deviation for incorrect answers is 62.58\% greater when compared to the same statistic for correct answers.
+
+\begin{table}[tb]
+  \caption{Time in seconds required by the participants to correctly and incorrectly predict the outcomes of the programs.}\label{tab:repeated:time}
+  \centering
+  \begin{tabular}{lrrr}
+    \toprule  
+             & Mean  & Median & Std. Dev. \\
+    \midrule
+  Correct   & 35.44 & 24.36  & 38.43     \\
+  Incorrect & 44.50 & 29.29  & 62.48    \\
+  \bottomrule
+  \end{tabular}
+  \end{table}
+
+\subsubsection*{Statistical Analysis}
+
+We employ the non-parametric Mann-Whitney U test, also known as Wilcoxon's rank-sum test, to verify the aforementioned null hypothesis, i.e., to check whether the times for correct and incorrect answers can be considered identical.
+That test produces a p-value of \textbf{0.00000046}, which shows that the difference it statistically significant and we can reject the null hypothesis.
+To gauge the magnitude of that difference we employ Cliff's Delta as a measure of effect size.
+This produces a result of \textbf{0.1027}, which suggests that, although participants tend to provide correct predictions more quickly, they do that only slightly so.
+
+%-----------------------------------------------------------------------
+\section{Results of the Latin-Square Study}\label{sec:latin}
 

### 2022-07-15 discussion.tex
First version with the results for the repeated measures study.

@@
 \label{sec:discussion}
+
+\castor{Which atoms?
+The weight of experience.
+Statistical significance vs.
+practical relevance.
+Most of them, with the exception of Comma Operator and Object Destructuring, have also been observed for Java programs~\cite{Langhout:2021:ACJ}.
+
+In addition, many of the coding idioms that have been previously identified as atoms of confusion in previous work~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ}, such as Logic as Control Flow
+
+Differences among the two studies, for example, post increment was not significant in the first one but it was significant in the second.
+}
 

### 2022-07-22 4_Result.tex
The results section now has the two studies and the text is (mostly) consistent with the existence of these two studies.

@@
 
-Overall, this study had 70 participants. Out of these, 67 have received university-level training in programming, 4 hold a PhD degree, 7 hold a master's degree, and 11 hold a bachelor's degree. 
+This study had 70 participants. Out of these, 67 have received university-level training in programming, 4 hold a PhD degree, 7 hold a master's degree, and 11 hold a bachelor's degree. 
 The median programming experience of the participants is four years (mean 6.5 years) and the median experience with JavaScript programming is one year (mean 2.5 years). 
@@
 These p-values are compared against an alpha of 0.0021, after applying Bonferroni correction (0.05/24).
-One of the atom candidates, Constant Variables, is omitted from the table because, due to the very low number of mistakes, it is not possible to perform the McNemar test.
+One of the atom candidates, Constant Variables, is omitted from the table because, due to the very low number of mistakes, it is not possible to perform the test.
 The table shows that eight of the 24 analyzed atom candidates can be considered atoms of confusion based on the results of our study, Object Destructuring, Type Conversion, Change of Literal Encoding, Comma Operator, Indentation No Braces, Assignment as Value, Pre-Increment, and Ommitted Curly Braces.
@@
 We employ the non-parametric Mann-Whitney U test, also known as Wilcoxon's rank-sum test, to verify the aforementioned null hypothesis, i.e., to check whether the times for correct and incorrect answers can be considered identical.
-That test produces a p-value of \textbf{0.00000046}, which shows that the difference it statistically significant and we can reject the null hypothesis.
+That test produces a p-value of \textbf{0.00000046}, which shows that the difference is statistically significant and we can reject the null hypothesis.
 To gauge the magnitude of that difference we employ Cliff's Delta as a measure of effect size.
@@
 
-Our first study investigates the impact of atoms of confusion while developers try to understand JavaScript code.
-We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
+In the latin-square study, we estimate the impact of atoms considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
 and \emph{time} (how long to provide a correct answer).
-We received full answers to our survey from 140 participants (a total of 70 replicas).
-All participants had taken at least some university course or hold a bachelor degree or equivalent, meaning the participants have had some level of formal education in programming.
+In particular, in the time analysis, differently from the repeated  measures study, we compare the time required by participants to submit a correct response when analyzing code snippets with and without atom candidates.
+This study is based on the responses of 140 participants (a total of 70 replicas).
+All participants had taken at least some university course or hold a bachelor degree or equivalent.
 In addition, 21 participants hold a master's degree and three a doctorate degree.
@@
 \subsection{Misunderstanding Rate Analysis}
-\castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me which rate is being referred to.}
+\castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me to which rate this refers.}
 
@@
 
-As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten code snippets, from which five were in their confusing versions, whilst the other five contained clean versions of the code snippets (i.e., without the confusing constructs and idioms).
-The participants should provide the expected outcomes of the code snippets.
+As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten code snippets, from which five were in their confusing versions, whilst the other five contained clean versions of the code snippets (i.e., without the atom candidates).
+As in the repeated measures study, the participants should provide the expected outcomes of the code snippets.
+Differently from that study, in this one each participant only predicts the outcome of one version of each code snippet, either clean or obfuscated.
 As discussed, we collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{time}.
@@
 %Considering only the atoms for which there was a significant difference in correctness for confusing and clean versions and u
-Using the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), for four of the atoms, Comma Operator, Automatic Semicolon Insertion, Assignment as Value, and Implicit Predicate, effect size can be considered large.
-For Implicit Predicate, it is medium.
+%Using the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), 
+For four of the atoms, Comma Operator, Automatic Semicolon Insertion, Assignment as Value, and Implicit Predicate, effect size can be considered large. For Implicit Predicate, it is medium. 
 %We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. 
@@
 This is the case for the four atom candidates in the lower part of the table.
-In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a participant was actually more likely to misunderstand the clean version.
+%In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a participant was actually more likely to misunderstand the clean version.
 
@@
 The findings suggest that \emph{participant experience} impacts on the results related to correctness (Figure~\ref{fig:correctness-over-experience}).
-Even though the presence of atoms of confusion reduces the number of correct answers to all \emph{experience groups}, this effect is not uniform.
+Even though the presence of atoms of confusion reduces the number of correct answers for all \emph{experience groups}, this effect is not uniform.
 For instance, novice developers (under one year of experience) provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets.
@@
 %% \end{mh}
-
-Comparing to previous research, three 
-atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed
-for \clang and \cpplang do not lead to a statistically significant impact
-on correctness in our study: Logic as Control Flow,
-Conditional Operator, and Pre-Increment. Our results
-support the Gopstein et al. findings for the remaining
-atoms. The work of Langhout and Aniche~\cite{Langhout:2021:ACJ} investigates six atom candidates for Java that we also explore here for JavaScript. Among these atom
-candidates, we achieve similar conclusions to their
-study for four atom candidates: Post Increment,
-Conditional Operator, Arithmetic as Logic, and
-Omitted Curly Braces---both studies confirmed only the former
-as an atom of confusion. Our findings diverged from
-the work of Langhout and Aniche for two atoms: Pre Increment and
-Logic as Control Flow.
 
@@
 29.06\% and 38.19\% for the clean versions---we did not confirm these
-atoms as atoms of confusion in the previous section.
+atom candidates as atoms of confusion in the previous section.
 
@@
 \subsubsection*{Statistical analysis}
-We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
+We use the non-parametric \emph{Mann-Whitney U test} to
 investigate the null hypothesis that developers 
@@
 In this section we present the results of the interviews with the practitioners.
-We contrast the findings of the survey presented in Section~\ref{sec:survey-resuts} here with the opinion
-of software developers about their preferences regarding the confusing
-or clean versions of the code snippets we used in our experiment.
+We contrast the findings presented in Section~\ref{sec:survey-resuts} with the opinion of software developers about their preferences regarding the confusing or clean versions of the code snippets we used in our experiment.
+For the interview study, we only considered atoms that were analyzed in both the repeated measures (Section~\ref{sec:repeated}) and the latin square (Section~\ref{sec:latin}) studies.
 
@@
 We collected information regarding programming
-experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in our research.
+experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in the studies of the previous sections.
 We first presented them pairs of clean and confusing code snippets and then asked them to discuss their preferences towards one version.
@@
 
-\subsubsection*{The Participants' perceptions of the atom candidates} Supporting the results of the survey, Table~\ref{tab:interview-results1}
+\subsubsection*{The Participants' perceptions of the atom candidates}
+
+Supporting the results of the studies of Sections~\ref{sec:repeated} and~\ref{sec:latin}, Table~\ref{tab:interview-results1}
  shows that for eight out of the
@@
 one of the participants found the version with the atom candidate more elegant, but recognized it was less readable, and was willing to sacrifice elegance for readability.
+In the repeated measures study, we found a significant difference in favor of the clean version, with a large effect size in its favor.
+In the latin square study, no difference could be observed in terms of correctness.
+However, there was a significant difference in terms of the time required predict the output correctly in favor of the obfuscated version.
+This highlights how factors such as the experience of the participants and the selection of code snippets can have considerable impact on the results of studies about code readability~\cite{Feitelson:2021:CPC}.
+\castor{Should the last sentence above be in the discussion?}
 
@@
 \end{mq}
+\noindent
+None of the two studies found a significant difference in correctness between clean and obfuscated versions of snippets related to this atom candidate.
+However, in the latin square study the participants analyzing clean versions of the snippets required significantly less time.
 
 For two of the atom candidates, the interviewees were unanimous in their preference for the clean version: Comma Operator and Omitted Curly Braces.
+Not coincidentally, in both studies there were significant differences in correctness in favor of the clean versions for both atoms.
 Regarding the first one, we could often notice during the interviews that the \emph{\lhs} (with the atom of confusion) caused significant confusion among the participants.
@@
 \emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).
+The former was investigated in the repeated measures study but we did not find a significant difference between clean and obfuscated versions.
 As future work, we aim at investigating whether or not these atom candidates are more likely to introduce confusion in JavaScript code.

### 2022-07-22 discussion.tex
The results section now has the two studies and the text is (mostly) consistent with the existence of these two studies.

@@
 \label{sec:discussion}
+
+% From the paper. I think it makes more sense to discuss this in the context of teh two studies. 
+Comparing the results of the latin square study to previous research, three 
+atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed
+for \clang and \cpplang do not lead to a statistically significant impact
+on correctness in our study: Logic as Control Flow,
+Conditional Operator, and Pre-Increment. Our results
+support the Gopstein et al. findings for the remaining
+atoms. The work of Langhout and Aniche~\cite{Langhout:2021:ACJ} investigates six atom candidates for Java that we also explore here for JavaScript. Among these atom
+candidates, we achieve similar conclusions to their
+study for four atom candidates: Post Increment,
+Conditional Operator, Arithmetic as Logic, and
+Omitted Curly Braces---both studies confirmed only the former
+as an atom of confusion. Our findings diverged from
+the work of Langhout and Aniche for two atoms: Pre Increment and
+Logic as Control Flow.
 

### 2022-07-22 3_Method.tex
Reviewed the methodology section, changing the name of the first expeirment to Repeated Measures Study and adjusting the remainder of the section for consistency. I have also attempted to emphasize that the two studies were designed and conducted independently, by different groups of researchers.

--- 2022-03-31 3_Method.tex
@@
 By answering these research questions, we can either generalize or refute the findings about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). 
+\castor{The following pertains to the third RQ.}
 In addition, answering the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice\castor{This should probably be removed.}.
@@
 We conduct a mixed-methods study to answer these questions. 
-It includes two independently designed and conducted experiments\castor{I would rather use the term ``experiment''.}: a replication of the first study on atoms of confusion, but targeting JavaScript programs, and a new study using a latin square counterbalancing experimental design to control for subject experience.
+It includes two independently designed and conducted experiments: a repeated measures/within subjects study and one using a Latin square, counterbalancing experimental design to control for subject experience.
+By using two different experiments, designed by different researchers, with different designs, and conducted with different participants, we hope to rigorously verify the impact of atoms of confusion in code comprehension activities.
 These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ1).
-They investigate whether or not programs that contain atom candidates produce a higher misunderstanding rate when programmers try to predict their outcome.\castor{I've removed the part about time because, in order for the first study to be a replication, we need to compute time differently, accounting for incorrect responses as well.}% and (b) require more time for programmers to predict their output.
-We leverage two independently-developed studies with different designs to more rigorously verify the impact of atoms of confusion in code comprehension activities. 
+They investigate whether or not programs that contain atom candidates tend to produce more misunderstanding for programmers trying to predict their outcome.% We leverage two independently-developed and conducted experiments with different designs to more
 We also conduct a set of interviews with developers to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2).
 
-\subsection{Replication Study Setting}\label{sec:meth:replication}
+\subsection{Repeated Measures Study}\label{sec:meth:replication}
 
-We call the first study we perform the \textit{``Replication Study''} because it is a partial replication~\cite{Juristo2012} of the first study on atoms of confusion, conducted by Gopstein et.
-al~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, but targeting JavaScript instead of the C language.
-Gopstein et.
-al.
-employ a randomized, partial counterbalancing experimental design on a sample of 73 subjects, mostly students, with three or more months of experience in C or C++.
-The control variable consists of a tiny program containing a single atom candidate and the treatment a functionally-equivalent version that does not contain the atom.
-The main dependent variable is whether subjects are able to correctly determine the output of the programs.
-The other dependent variable is the time required to correctly vs.
-incorrectly determine the outputs.
-The study has the null hypothesis that \textit{``code from both control and treatment groups can be hand-evaluated with equal accuracy''} and the alternate hypothesis that \textit{``the existence of an atom of confusion causes more errors than other code in hand-evaluated outputs.''}.
-For each atom candidate, six programs were built, three versions including the atom candidate and three versions not including it.
-Each subject is exposed to four programs for every atom candidate, two including an atom and two not including an atom.
+\castor{We must have introduced the atoms before.}
+The first experiment we perform uses a repeated measures, or within-subject, design~\cite{Keselman:2001:ARM}.
+This experiment is inspired by the work of Gopstein et.
+al~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, but targeting JavaScript instead of the C language, and including some methodological differences.
+It covers all the atom candidates analyzed in the work of Gopstein et al.
+that can be transposed to JavaScript and adds candidates that are specific to JavaScript.
+Overall, we select 24 atom candidates in the repeated measures study.
 
@@
 
-In this study we adopt the same experimental design, control and treatment variables, dependent variable, and hypotheses.
+In this experiment, the control variable consists of a tiny program containing a single atom candidate and the treatment is a functionally-equivalent version that does not contain the atom.
+The main dependent variable is whether subjects are able to correctly determine the output of the programs.
+The other dependent variable is the time required to correctly vs.
+incorrectly determine the outputs, independently of the presence of atoms.
+We use the same null hypothesis as Gopstein et al.: \textit{``code from both control and treatment groups can be hand-evaluated with equal accuracy''} and the alternate hypothesis that \textit{``the existence of an atom of confusion causes more errors than other code in hand-evaluated outputs.''}.
+For each atom candidate, we built six programs, three versions including the atom candidate and three versions not including it.
+%Each subject is exposed to four programs for every atom candidate, two including an atom and two not including an atom.
 We have also built six programs for each atom candidate, three with and three without the atom.
@@
 
-The design of the replication study considers two treatments: the presence or absence of atoms candidates within the programs.
+The design of the repeated measures study considers two treatments: the presence or absence of atom candidates within the programs.
 To identify atom candidates, we start out by selecting every atom candidate from Gopstein's study that can also happen in JavaScript.
@@
 Furthermore, we conduct an informal search of programming forums, solutions to Code Golf~\footnote{https://codegolf.stackexchange.com/} problems, code style guidelines, and expert knowledge to identify new atom candidates.
-Overall, the replication study analyzes 24 atoms, 15 from the original study and 9 that are new and specific to JavaScript.
+Overall, the repeated measures analyzes 24 atoms, 15 from the original study and 9 that are new and specific to JavaScript.
 Table~\ref{tab:atoms_replication} presents the complete list.
@@
 For each atom candidate, we write six short programs, three including and three not including the atom candidate.
-The former are the control for the study, or the obfuscated versions of the programs.
-The alternative, atom-free, versions of the programs are the treatment, or clean, versions.
+The former are the control for the study, or the \textbf{obfuscated} versions of the programs.
+The alternative, atom-free, versions of the programs are the treatment, or \textbf{clean}, versions.
 Each subject is exposed to one obfuscated and one clean version of each atom, totalling 48 programs, and should determine their outputs.
@@
 To reduce the likelihood of a subject attempting to execute the programs, the web application verifies whether the subject tries to change tabs or windows and presents a pop-up message if that is the case.
-Furthermore, it presents the programs as images, instead of text.
+Furthermore, it presents the programs as images, instead of text, to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter.
 Upon submitting an answer for a particular program, the subject is automatically led to a similar page, containing the next one.
@@
 We collected 70 responses.
-The average survey respondent has a mean 26 months of experience in JavaScript, with a median of 12.\castor{We should not say this here unless we are able to do the same for the Latin square study.
+%The average survey respondent has a mean 26 months of experience in JavaScript, with a median of 12.\castor{We should not say this here unless we are able to do the same for the Latin square study.
 }
@@
 This experiment uses a different experimental design that makes it possible to control for the experience of the subjects.
-For this study we employ a subset of the atom candidates used in the replication study.
+It employs a subset of the atom candidates used in the repeated measures study.
 More specifically, we
 %Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first 
-select a set of 9+1 atom candidates: nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion)\castor{I think the name of this atom is inconsistent between the studies}.
-Our supplementary material discusses and shows examples of the atoms we consider in our research.\castor{We should probably do this here or just refer to the table that I intend to insert for the replication study.}
+select a set of 9+1 atom candidates: nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion).
+%Our supplementary material discusses and shows examples of the atoms we consider in our research.\castor{We should probably do this here or just refer to the table that I intend to insert for the replication study.}
 
@@
 
-The design of our second study blocks two variables (subject experience and the programs) and
-considers two treatments: the presence or absence of atom candidates within the programs. 
-To achieve such a design goal, controlling the effect of experience and individual programs, we resort to the \textit{Latin Square Design} \cite{Hunter-Experimenters}.
+The design of our second study blocks two variables (subject experience and the programs) and considers two treatments: the presence or absence of atom candidates within the programs. 
+To achieve such a design goal, controlling the effect of experience and individual programs, we resort to the \textit{Latin Square Design}~\cite{Hunter-Experimenters}.
 Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of programs.
@@
 For each of the 10 selected atom candidates, we write one short program containing the atom.
-As in the replication study, we call them the obfuscated versions of the programs.
+As in the repeated measures study, we call them the obfuscated versions of the programs.
 We also write corresponding short programs without the atom candidates and call them the clean versions of the programs.
@@
 
-We implemented our survey as a web application and carried out a pilot, similarly to the replication study. 
+We implemented our survey as a web application and carried out a pilot. 
 %As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
@@
 In our setting, ``I do not know'' is treated as a wrong answer.
-The programs are presented as images copied from a text editor, so as to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter.
+The programs are also presented as images copied from a text editor.
 Upon submitting their answer for a particular question, the subjects are automatically led to a similar page, containing another program.
-Similarly to the replication study, we do not provide feedback about time and correctness to the subjects.
+Similarly to the repeated measures study, we do not provide feedback about time and correctness to the subjects.
 
@@
 Our main selection criterion is that all participants should have been working with JavaScript in their daily professional activities.
-We invite a total of 17 developers, and 15 of them agree to participate.
+We invited a total of 17 developers, and 15 of them agreed to participate.
 %We conducted the interviews for a period of two weeks.

### 2022-07-25 3_Method.tex
small fixes in the text

@@
 The first experiment we perform uses a repeated measures, or within-subject, design~\cite{Keselman:2001:ARM}.
-This experiment is inspired by the work of Gopstein et.
-al~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, but targeting JavaScript instead of the C language, and including some methodological differences.
+This experiment is inspired by the work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, but targeting JavaScript instead of the C language, and including some methodological differences.
 It covers all the atom candidates analyzed in the work of Gopstein et al.
 that can be transposed to JavaScript and adds candidates that are specific to JavaScript.
-Overall, we select 24 atom candidates in the repeated measures study.
+%Overall, we select 24 atom candidates in the repeated measures study.
 
@@
 
-The design of the repeated measures study considers two treatments: the presence or absence of atom candidates within the programs.
+% The design of the repeated measures study considers two treatments: the presence or absence of atom candidates within the programs.
 To identify atom candidates, we start out by selecting every atom candidate from Gopstein's study that can also happen in JavaScript.
@@
 Furthermore, it presents the programs as images, instead of text, to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter.
-Upon submitting an answer for a particular program, the subject is automatically led to a similar page, containing the next one.
+Upon submitting an answer for a particular program, the subject is automatically led to a similar page, containing the next program.
 
@@
 
-\subsection{Latin Square Study Setting}\label{sec:meth:survey}
+\subsection{Latin Square Study}\label{sec:meth:survey}
 
@@
 To complement the two experiments, we perform semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding programs containing atoms of confusion (research question RQ2).
-We also ask each participant if they know of any other JavaScript-specific construct or idiom that they regard as being likely to make the code hard to understand.% (research question RQ3).
+We also ask each participant if they know of any other JavaScript-specific construct or idiom that they regard as being likely to make the code hard to understand.
+% (research question RQ3).
  In this section we detail the protocol we followed to conduct the interviews and to analyze the results.
@@
 
-\subsubsection*{Interview Process} We conduct semi-structured interviews using web conferencing software.
+\subsubsection*{Interview Process} We conduct semi-structured interviews using a web conferencing software.
 We record all the interviews with the consent of the participants.
@@
 
-To understand how often the analyzed atom candidates appear in open source projects, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories.
+To understand how often the analyzed atom candidates appear in open source projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories.
 To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript.
@@
  \begin{tabular}{rrrrr}
-   \hline
-                       & Min.             & Median        & Mean             & Max. \\ \hline
+   \toprule
+                       & Min.             & Median        & Mean             & Max. \\ \midrule
  Lines of Code         & \num{5543}       & \num{36161.5} & \num{111432.33}  & \num{1278405} \\
@@
 
-    \hline
+    \bottomrule
  \end{tabular}

### 2022-07-25 4_Result.tex
small fixes in the text

@@
 %\section{Results of the Survey}
-\label{sec:survey-resuts}
+%\label{sec:survey-resuts}
 
@@
 
-\subsubsection*{Exploratory Data Analysis. }
+\subsubsection*{Exploratory Data Analysis}
 For each of the 24 atom candidates, the participants predicted the outputs of two versions, one clean (control) and one obfuscated (treatment), as per the definition of a repeated measures design~\cite{Hunter-Experimenters}.
@@
   Atom & Confusing & Clean & $\Delta$(\%)  \\ \midrule
-Type Conversion           &            5.0 &      27.0 &                 +440.0 \\
-Change of Literal Encoding &           9.0 &      36.0 &                 +300.0 \\
-Comma Operator            &            8.0 &      29.0 &                 +262.0 \\
-Object Destructuring      &           15.0 &      42.0 &                 +180.0 \\
-Assignment As Value       &           18.0 &      43.0 &                 +139.0 \\
-Pre-Increment             &           13.0 &      28.0 &                 +115.0 \\
-Post-Increment            &            8.0 &      17.0 &                 +112.0 \\
-Array Destructuring       &           15.0 &      28.0 &                 +87.0 \\
-Indentation no Braces     &           30.0 &      50.0 &                 +67.0 \\
-Omitted Curly Braces      &           35.0 &      51.0 &                 +46.0 \\
-Object Spread             &           21.0 &      30.0 &                 +43.0 \\
-Array Spread              &           22.0 &      31.0 &                 +41.0 \\
-Automatic Semicolon Insertion &       15.0 &      21.0 &                 +40.0 \\
-Infix Operator Precedence &           47.0 &      53.0 &                 +13.0 \\
-Arrow Function            &           54.0 &      58.0 &                  +7.0 \\
-Implicit Predicate        &           41.0 &      44.0 &                  +7.0 \\
-Indentation With Braces   &           53.0 &      55.0 &                  +4.0 \\
-Conditional Operator      &           60.0 &      62.0 &                  +3.0 \\
-Constant Variables        &           59.0 &      61.0 &                  +3.0 \\
-Dead Unreachable Repeated &           63.0 &      64.0 &                  +2.0 \\
-Repurposed Variables      &           23.0 &      23.0 &                   0.0 \\
-Arithmetic as Logic       &           52.0 &      47.0 &                 -10.0 \\
-Logic as Control Flow     &           33.0 &      28.0 &                 -15.0 \\
-Property Access           &           51.0 &      42.0 &                 -18.0 \\
+Type Conversion           &            5 &      27 &                 +440 \\
+Change of Literal Encoding &           9 &      36 &                 +300 \\
+Comma Operator            &            8 &      29 &                 +262 \\
+Object Destructuring      &           15 &      42 &                 +180 \\
+Assignment As Value       &           18 &      43 &                 +139 \\
+Pre-Increment             &           13 &      28 &                 +115 \\
+Post-Increment            &            8 &      17 &                 +112 \\
+Array Destructuring       &           15 &      28 &                 +87 \\
+Indentation no Braces     &           30 &      50 &                 +67 \\
+Omitted Curly Braces      &           35 &      51 &                 +46 \\
+Object Spread             &           21 &      30 &                 +43 \\
+Array Spread              &           22 &      31 &                 +41 \\
+Automatic Semicolon Insertion &       15 &      21 &                 +40 \\
+Infix Operator Precedence &           47 &      53 &                 +13 \\
+Arrow Function            &           54 &      58 &                  +7 \\
+Implicit Predicate        &           41 &      44 &                  +7 \\
+Indentation With Braces   &           53 &      55 &                  +4 \\
+Conditional Operator      &           60 &      62 &                  +3 \\
+Constant Variables        &           59 &      61 &                  +3 \\
+Dead Unreachable Repeated &           63 &      64 &                  +2 \\
+Repurposed Variables      &           23 &      23 &                   0 \\
+Arithmetic as Logic       &           52 &      47 &                 -10 \\
+Logic as Control Flow     &           33 &      28 &                 -15 \\
+Property Access           &           51 &      42 &                 -18 \\
 \bottomrule  
@@
 
-Table~\ref{tab:repeated:time} presents the mean and median times required by participants to provide answers.
+Table~\ref{tab:repeated:time} presents the mean and median times required by the participants to provide answers.
 The data is based on the 1648 correct and 1572 incorrect answers collected in the study.
-Contrary to our intuition, the mean time required by participants to provide an incorrect answer is 25\% greater than the time required to provide a correct one and the median time is 20\%.
-greater.
+Contrary to our intuition, the mean time required by participants to provide an incorrect answer is 25\% greater than the time required to provide a correct one and the median time is 20\% greater.
 The standard deviations suggest that there is great variation in the times required by different participants, for different code snippets.
@@
 
-In the latin-square study, we estimate the impact of atoms considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
+In the latin-square study, we estimate the impact of atoms considering the same two perspectives of the previous experiment: \emph{correctness} (number of wrong answers)
 and \emph{time} (how long to provide a correct answer).
@@
 
-\subsection{Misunderstanding Rate Analysis}
-\castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me to which rate this refers.}
+\subsection{Correctness Analysis}
+% \castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me to which rate this refers.}
 
@@
 As discussed, we collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{time}.
-Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey w.r.t. correctness.
+Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the correctness experiment.
 
@@
 %, regardless of evaluating the confusing or clean version of the code.
-The results of Table~\ref{tab:difference-time-taken}
+The results of Table~\ref{tab:hypothesis-testing}
 suggest that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
@@
 In this section we present the results of the interviews with the practitioners.
-We contrast the findings presented in Section~\ref{sec:survey-resuts} with the opinion of software developers about their preferences regarding the confusing or clean versions of the code snippets we used in our experiment.
+We contrast the findings presented in Section~\ref{sec:latin} and Section~\ref{sec:repeated} with the opinion of software developers about their preferences regarding the confusing or clean versions of the code snippets we used in our experiment.
 For the interview study, we only considered atoms that were analyzed in both the repeated measures (Section~\ref{sec:repeated}) and the latin square (Section~\ref{sec:latin}) studies.
@@
 In the latin square study, no difference could be observed in terms of correctness.
-However, there was a significant difference in terms of the time required predict the output correctly in favor of the obfuscated version.
+However, there was a significant difference in terms of the time required to predict the output correctly in favor of the obfuscated version.
 This highlights how factors such as the experience of the participants and the selection of code snippets can have considerable impact on the results of studies about code readability~\cite{Feitelson:2021:CPC}.
 \castor{Should the last sentence above be in the discussion?}
-
+\rb{Agree. We can move it to the discussion section.}
 %% \rb{pela discuss\~{a}o anterior, parece-me que alguns
@@
 for 92.97\% of this total: Implicit Predicate, Post Increment, Conditional Operator, and Omitted Curly Braces.
-The first two, based on the results of Section~\ref{sec:survey-resuts}, can be considered atoms of confusion.
+The first two, based on the results of Section~\ref{sec:latin} and Section~\ref{sec:repeated}, can be considered atoms of confusion.
 The remaining atom

### 2022-07-25 discussion.tex
small fixes in the text

@@
 \begin{mh}
-  {\bf Answer to RQ1:} The first study (survey) gives evidence that some of the atom candidates in \clang and \cpplang programs that also exist in JavaScript correspond to a source of misunderstanding in
+  {\bf Answer to RQ1:} Our experiments using repeated measures and the Latin square designs give evidence that some of the atom candidates in \clang and \cpplang programs that also exist in JavaScript correspond to a source of misunderstanding in
   JavaScript code. 
@@
   {\bf Answer to RQ2:} The qualitative analysis of the
-  interviews supports the results of the first study,
-  since JavaScript developers most often agree that atoms of confusion compromise
+  interviews supports the results of our experiments---
+  JavaScript developers most often agree that atoms of confusion compromise
   source code understanding. 
@@
   {\bf Answer to RQ3:} The MSR study reveals that
-  the several atom candidates explored in our research
+  several atom candidates explored in our research
   appear frequently in practice, and cleaning up the use of 

### 2022-07-25 3_Method.tex
Passed through the methodology section, with a focus only on the experiments.

--- 2022-07-22 3_Method.tex
@@
 It employs a subset of the atom candidates used in the repeated measures study.
-More specifically, we
+More specifically, this study investigates ten of the atoms investigated in the repeated measures study:
 %Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first 
-select a set of 9+1 atom candidates: nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion).
+%select a set of 9+1 atom candidates: 
+nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion).
 %Our supplementary material discusses and shows examples of the atoms we consider in our research.\castor{We should probably do this here or just refer to the table that I intend to insert for the replication study.}
@@
 
-The design of our second study blocks two variables (subject experience and the programs) and considers two treatments: the presence or absence of atom candidates within the programs. 
+The design of the Latin square study blocks two variables (subject experience and the programs) and considers two treatments: the presence or absence of atom candidates within the programs.
 To achieve such a design goal, controlling the effect of experience and individual programs, we resort to the \textit{Latin Square Design}~\cite{Hunter-Experimenters}.
@@
 
-We implemented our survey as a web application and carried out a pilot. 
-%As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
+We implemented our survey as a web application and 
+%carried out a pilot. 
+%As part of this effort, we 
+carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
 Undergraduate students and professional colleagues took the pilot survey.

### 2022-07-25 main.tex
The Jupyter Notebook with the analysis of the Repeated Measures Study is finally ready. I have started including the results in the paper.

@@
 
+%\usepackage[english]{babel}
 \usepackage{enumerate}
@@
 \usepackage{booktabs}
-\usepackage{siunitx}
+\usepackage[group-separator = {,}, group-minimum-digits = 4]{siunitx}
 \usepackage{amsmath,amssymb,amsfonts}
@@
   morecomment=[s][\color{javadocblue}]{/**}{*/},
-  numbers=left,
-  numberstyle=\tiny\color{black},
+%  numbers=left,
+%  numberstyle=\tiny\color{black},
   stepnumber=1,

### 2022-07-25 0_Abstract.tex
Reviewed the abstract and intro to make thenm consistent with Sections 3 and 4.

--- 2020-11-09 0_Abstract.tex
@@
 %Besides lack of experience and other factors, 
-Previous research works have discussed the impact of small code patterns that contribute to misunderstanding the source code written in statically typed languages (e.g., \clang, \cpplang, and Swift).
-However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamically typed languages.
-\noindent{\bf Research Method and Goal.} In this work, we fill this research gap by presenting the results of a mixed-methods research:
-a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
+Previous research has discussed the impact of small code patterns, called atoms of confusion, that contribute to misunderstanding the source code and for which there are functionally equivalent alternatives.
+These prior studies have focused on the C and Java languages.
+However, different languages have different features, developer communities, and development cultures.
+This justifies the exploration other languages to verify whether they also exhibit confusion-inducing patterns.
+\noindent{\bf Research Method and Goal.} In this work, we fill this research gap by presenting the results of a mixed-methods research: two experiments, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
 Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
-\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead to a code that is hard to understand: 
+\noindent{\bf Results.} Both our experiments suggest that two code patterns that have been observed to confused C programmers also confuse JavaScript programmers (with statistical significance, after correction): the comma operator and assignments being used as values.
+In addition,
+pre- and post-increment operator, when used as values, ommitted curly braces, changes of literal encoding, and type conversions have led to confusion in at least one of the two experiments.
+For all these cases effect sizes were either medium or high.
+In addition, we have discovered the object destructuring and automatic semicolon insertion are potentially JavaScript-specific atoms that have not been discussed before.
+The interviews we conducted reinforce some of these findings while pointing to other constructs and idioms that merit investigation in the future.
+
+%From our experiments and interviews, we confirm that atoms of confusion lead to code that is hard to understand. 
 %removing the atoms of confusion lead to an improvement above 15\% in 6 out of 9 atoms we consider in our study when developers try to predict the outcomes of code examples. 
-developers correctly predict the outcome of at least 15\% more when an atom of confusion is not present for 6 out of 9 atoms.
+%developers correctly predict the outcome of at least 15\% more when an atom of confusion is not present for 9 out of 10 atoms.
+For 5 of these atoms, the difference is statistically significant (accounting for p-value correction) and effect size is large for 4 and medium for 1 of them.
 Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the \texttt{comma operator} is present in 11\% of the projects in our dataset; while five other atoms appear in more than 60\% of the projects---including the atoms \texttt{assignment as value} and \texttt{omitted curly braces}, which appear \num{8253} and \num{53357} times, respectively.

### 2022-07-25 main.tex
Reviewed the abstract and intro to make thenm consistent with Sections 3 and 4.

@@
 \begin{abstract}
-  Evolving legacy code is a challenging task, particularly when the code has been poorly written or uses confuse idioms and language constructs, which might increase maintenance efforts and impose a significant cognitive load on developers.
-For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in statically typed languages such as \clang, \cpplang, and Java.
-In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language.
-We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and an activity of mining open source JavaScript repositories (MSR).
-Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript.
-Considering ten atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates.
-For five of these atoms, the difference is statistically significant (accounting for p-value correction).
-Effect size is large for four atoms and medium for one of them.
+\castor{This is still too big.}
+Evolving software is a challenging task, particularly when the code has been poorly written or uses confusing idioms and language constructs, which might increase maintenance effort and impose a significant cognitive load on developers.
+For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in languages such as \clang, \cpplang, and Java.
+However, different languages have different features, developer communities, and development cultures.
+This justifies the exploration of other languages to verify whether they also exhibit confusion-inducing patterns.
+In this work, we fill this research gap by presenting the results of a mixed-methods research comprising two experiments, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
+Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in diverse application domains, e.g., web, backend, mobile, and IOT development.
+Both our experiments suggest that two code patterns that have been previously observed to confuse C programmers also confuse JavaScript programmers (with statistical significance, after correction): the comma operator and assignments being used as values.
+In addition, pre- and post-increment operators, when used as values, omitted curly braces, changes of literal encoding, and type conversions have led to confusion in at least one of the two experiments.
+For all these cases effect sizes were either medium or high.
+In addition, we have discovered that object destructuring and automatic semicolon insertion are potential JavaScript-specific atoms that have not been discussed in previous work.
+The interviews we conducted reinforce some of these findings while pointing to other constructs and idioms that merit investigation in the future.
 In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems.
 Four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code in the dataset.
-Altogether, our findings might help practitioners: (1) better understand the implications of atoms of confusion on understanding JavaScript code, (2) avoid writing code that is unnecessarily difficult to maintain, and (3) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
+Our findings provide significant evidence that some code patterns should be avoided if code is expected to be understood by a wide audience.
+Furthermore, they motivate the creation of program transformation tools and the extension of linters to address these potentially harmful idioms in JavaScript codebases.
 \end{abstract}
 
-\begin{IEEEkeywords}
-code readability, program comprehension, program understanding, atoms of confusion, JavaScript code
+%\begin{IEEEkeywords}
+%code readability, program comprehension, program understanding, atoms of confusion, JavaScript code
 %datasets, neural networks, gaze detection, text tagging
-\end{IEEEkeywords}
+%\end{IEEEkeywords}
 

### 2022-07-25 3_Method.tex
Reviewed the abstract and intro to make thenm consistent with Sections 3 and 4.
Re-reviewed the text of the methodology section pertaning to the two experiments. Tried to remove the word survey from it.

@@
 It covers all the atom candidates analyzed in the work of Gopstein et al.
-that can be transposed to JavaScript and adds candidates that are specific to JavaScript.
+that can be transposed to JavaScript and adds candidates that are specific to this language.
 %Overall, we select 24 atom candidates in the repeated measures study.
@@
 Furthermore, we conduct an informal search of programming forums, solutions to Code Golf~\footnote{https://codegolf.stackexchange.com/} problems, code style guidelines, and expert knowledge to identify new atom candidates.
-Overall, the repeated measures analyzes 24 atoms, 15 from the original study and 9 that are new and specific to JavaScript.
+Overall, the repeated measures study analyzes 24 atoms, 15 from the original study and 9 that are new.
 Table~\ref{tab:atoms_replication} presents the complete list.
@@
 
-The study is conducted by means of a web application.
-As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
+The experiment is conducted by means of a questionnaire made available as a web application.
+As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the questionnaire would take on average.
 Based on the feedback of the participants of the pilot, we present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.% Initially, our plan was to present two of each.
@@
 In the first one we present instructions and also a check button whose checking means users agree that all collected data will be used anonymously and solely for research purposes.
-The instructions explain how the survey works and asks them to dedicate their attention to it.
-We stress to participants the importance of not using any aids during the survey, such as online or console interpreters.
+The instructions explain how the study works and asks them to dedicate their attention to it.
+We stress to participants the importance of not using any aids during the experiment, such as online or console interpreters.
 
@@
 This aims at avoiding introducing bias for future respondents.
-Since we posted the survey on social media platforms, possible threats could have arisen if we gave respondents instant feedback.
+Since we posted information about the study on social media platforms, possible threats could have arisen if we gave respondents instant feedback.
 
@@
 
-We posted the survey on the authors' social media platforms, on Brazilian CS departments' mailing lists, and on two programming subreddits.
-We explained our research purposes and asked developers to take the survey only if they have some familiarity with JavaScript.
+We posted the link to the questionnaire on the authors' social media platforms, on Brazilian CS departments' mailing lists, and on two programming subreddits.
+We explained our research purposes and asked developers to take part in the study only if they have some familiarity with JavaScript.
 We collected 70 responses.
@@
 
-We implemented our survey as a web application and 
+We implemented our experiment as a questionnaire in a web application and 
 %carried out a pilot. 
 %As part of this effort, we 
-carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
-Undergraduate students and professional colleagues took the pilot survey.
-Some users reported layout defects, and many reported that the landing page did not explain the survey well enough.
+carried out an informal pilot.% whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
+Undergraduate students and professional colleagues took part in the pilot.
+Some users reported layout defects, and many reported that the landing page did not explain the study well enough.
 We also spotted minor issues with our routines to create and populate the latin squares.
 
-We organize the survey in three sections.
+We organize the questionnaire in three sections.
 The first section aims to characterize the subjects, asking their age, education level, and programming experience.
 We also include a check button, whose checking means users agree that all collected data will be used solely for research purposes.
-In the second section, we present to the participants a small set of instructions, where we explain how the survey works and ask them to dedicate their attention to it.
-We stress to participants the importance of not using any aids during the survey, such as online or console interpreters.
+In the second section, we present to the participants a small set of instructions, where we explain how the study works and ask them to dedicate their attention to it.
+We stress to participants the importance of not using any aids during the experiment, such as online or console interpreters.
 For each question page, we kept track of whether or not the subjects switch windows.
 
-The last section of the survey presents a sequence of ten questions, each containing a program.
+The last section of the questionnaire presents a sequence of ten questions, each containing a program.
 For each question, there is a text box where the answer should be written.
@@
 
-We posted the survey on a JavaScript Reddit channel.\footnote{https://www.reddit.com/r/javascript/} We explained our research purposes and asked developers of any level of expertise to take the survey.
+We posted the invitation to take part in the study on a JavaScript Reddit channel.\footnote{https://www.reddit.com/r/javascript/} We explained our research purposes and asked developers of any level of expertise to take part.
 Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares.
 We collected significant data on time taken and discrepancies in answer correctness between obfuscated and clean versions of the programs.
-Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the survey.
+Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the questionnaire.
 We discarded from our study all squares that contained incomplete rows (a common approach for data imputation).
@@
 \subsubsection*{Participant Selection} We invite the participants of the interviews using a snowballing technique.
-That is, starting from our network of contacts, we invite an initial set of candidates to take part in our survey.
+That is, starting from our network of contacts, we invite an initial set of candidates to take part in our experiments.
 From this initial list, we ask for an indication of additional candidates.
@@
 
-\subsubsection*{Interview Process} We conduct semi-structured interviews using a web conferencing software.
+\subsubsection*{Interview Process} We conduct semi-structured interviews using web conferencing software.
 We record all the interviews with the consent of the participants.

### 2022-07-25 3_Method.tex
Finished reviewing the methodology section. Next is the results section.

@@
 To complement the two experiments, we perform semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding programs containing atoms of confusion (research question RQ2).
-We also ask each participant if they know of any other JavaScript-specific construct or idiom that they regard as being likely to make the code hard to understand.
+We also ask each participant if they know of any other JavaScript-specific construct or idiom they think is likely to make the code hard to understand.
 % (research question RQ3).
@@
 
-In the third part of the interview, participants are shown pairs of programs that were used in the first study, where each pair contains a clean and an obfuscated version of the same atom candidate.
+In the third part of the interview, participants are shown pairs of programs that were used in the experiments, where each pair contains a clean and an obfuscated version of the same atom candidate.
+For this part we only use atoms that appear in both experiments.
 The participants
@@
 
-To understand how often the analyzed atom candidates appear in open source projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories.
-To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript.
-We measured popularity using the project's stargazers.
+To understand how often the analyzed atom candidates appear in open source projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mine a set of GitHub open source repositories.
+To this end, we first collect the most popular GitHub repositories that are primarily written in JavaScript.
+We measure popularity using the project's stargazers.
 This metric, available through the GitHub API, represents the number of stars a project received from users of the platform.
 The same metric has been used in a number of previous studies as a proxy to estimate a project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}.
-We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.
+We then select the top 100 most popular repositories and remove projects that do not reach the first quartile of the distribution of lines of code.
 
-After filtering out JavaScript project candidates, in the second step we built a curated dataset comprising the top \minedprojects repositories.
+After filtering out JavaScript project candidates, in the second step we build a curated dataset comprising the top \minedprojects repositories.
 Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}.
@@
 All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars.
-We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.
+We automate all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.
 
-We mined the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}.
+We mine the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}.
 CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript).
 We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study).
-Finally, we computed some descriptive statistics to measure the prevalence of atoms of confusion in practice.
+Finally, we compute some descriptive statistics to measure the prevalence of atoms of confusion in practice.
 

### 2022-07-25 4_Result.tex
Finished reviewing the results section.

@@
 
-In this study, we investigate whether 24 atom candidates impact readability.
+In this experiment, we investigate whether 24 atom candidates impact readability.
 We examine how the performance of the participants differ when predicting the outputs of the clean and obfuscated versions of small code snippets (Section~\ref{sec:repeated:correct}).
@@
 \begin{tabular}{lccc} \toprule
-  Atom & Confusing & Clean & $\Delta$(\%)  \\ \midrule
+  Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
 Type Conversion           &            5 &      27 &                 +440 \\
@@
 Among these eight, only Object Destructuring is JavaScript-specific.
-The other seven atoms have been previously reported for the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and also for Java~\cite{Langhout:2021:ACJ} (with the exception of Comma Operator).
+Six of the other atoms have been previously reported for the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and five for Java~\cite{Langhout:2021:ACJ}.
+Indentation No Braces is not specific to JavaScript, but has not been studied previously.
 For some candidates that have been previously identified as atoms, there was not a statistically significant difference.
@@
 The odds ratio is an appropriate measure of effect size for scenarios where McNemar's test is applicable~\cite{Mangiafico:2016:SAE}.
-For example, odds ratio for Type Conversion is 12.
+For example, the odds ratio for Type Conversion is 12.
 This suggests that the odds of a participant correctly predicting the output of a clean version while incorrectly predicting the output of the corresponding obfuscated version is 12 times higher than the odds of correctly predicting the output of the obfuscated version and incorrectly predicting the output of the correct one.
 
-According to the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), for most of the atoms where there was a statistically significant difference the odds ratio can be considered high.
+According to the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), for most of the atoms where there is a statistically significant difference the odds ratio can be considered high.
 The value of the odds ratio for Object Destructuring is $\infty$ because no participant correctly guessed the output for the obfuscated version while incorrectly guessing the output for the clean one, which results in a denominator of 0.
@@
 
-We employ the non-parametric Mann-Whitney U test, also known as Wilcoxon's rank-sum test, to verify the aforementioned null hypothesis, i.e., to check whether the times for correct and incorrect answers can be considered identical.
+We employ the non-parametric Mann-Whitney U test, also known as Wilcoxon's rank-sum test, to verify the aforementioned null hypothesis, i.e., whether the times for correct and incorrect answers can be considered identical.
 That test produces a p-value of \textbf{0.00000046}, which shows that the difference is statistically significant and we can reject the null hypothesis.
@@
 
-As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten code snippets, from which five were in their confusing versions, whilst the other five contained clean versions of the code snippets (i.e., without the atom candidates).
+As mentioned in Section~\ref{sec:meth:survey}, each participant in this experiment evaluated ten code snippets, from which five were in their obfuscated versions, whilst the other five contained clean versions of the code snippets (i.e., without the atom candidates).
 As in the repeated measures study, the participants should provide the expected outcomes of the code snippets.
 Differently from that study, in this one each participant only predicts the outcome of one version of each code snippet, either clean or obfuscated.
-As discussed, we collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{time}.
+As discussed, we collect information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{time}.
 Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the correctness experiment.
 
-Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the confusing versions.
+Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the obfuscated versions.
 In particular, the presence of the \emph{Comma Operator} atom exhibits the highest impact on misunderstanding.
@@
 The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the clean versions of the code snippets.
-Also, the sample of responses with no atoms had almost no dispersion, which supports the argument that the non-confusing versions of the code snippets are easier to evaluate correctly.
+Also, the sample of responses with no atoms had almost no dispersion, which supports the argument that the clean versions of the code snippets are easier to evaluate correctly.
 
@@
 \begin{tabular}{lccc} \toprule
-  Atom & Confusing & Clean & $\Delta$(\%)  \\ \midrule
+  Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
   Comma Operator                &  28 &  65 & +132   \\ 
@@
 We first use the \emph{Pearson's Chi-squared test}
-to investigate if there is a statistically significant difference in the frequency of correct and incorrect answers---due to the versions of the code snippets (confusing and clean code).
+to investigate if there is a statistically significant difference in the frequency of correct and incorrect answers---due to the versions of the code snippets (obfuscated and clean code).
 The p-values for these tests are reported in the ``Chi-square test'' column of Table~\ref{tab:hypothesis-testing}.
-The results indicate that for five atom candidates (Comma Operator, Automatic Semicolon Insertion, Post-Increment, Assignment as Value, Implicit Predicate) the confusing versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$).
+The results indicate that for five atom candidates (Comma Operator, Automatic Semicolon Insertion, Post-Increment, Assignment as Value, Implicit Predicate) the obfuscated versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$).
 This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
@@
 %We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. 
-For instance, when comparing clean and confusing code snippets pertaining to the Comma Operator atom candidate, we observed an \emph{Odds Ratio} of \num{19.02}.
-This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding confusing version of the code snippet.
+For instance, when comparing clean and obfuscated code snippets pertaining to the Comma Operator atom candidate, we observed an \emph{Odds Ratio} of \num{19.02}.
+This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding obfuscated version of the code snippet.
 Furthermore, with a 95\% confidence level, the odds ratio is between 6.6 and 68.22, i.e., although the error margin is wide, it is strongly in favor of the clean version.
@@
 
-Finally, we also use a \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impact correctness.
-The findings suggest that \emph{participant experience} impacts on the results related to correctness (Figure~\ref{fig:correctness-over-experience}).
+Finally, we also use a \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impacts correctness.
+The findings suggest that \emph{participant experience} impacts the results related to correctness (Figure~\ref{fig:correctness-over-experience}).
 Even though the presence of atoms of confusion reduces the number of correct answers for all \emph{experience groups}, this effect is not uniform.
-For instance, novice developers (under one year of experience) provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets.
+For instance, novice developers (under one year of experience) provide a higher number of wrong answers for the obfuscated version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the obfuscated version of the code snippets.
 It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.
@@
 Differently from \emph{participant experience}, the factor \emph{participant education} does not significantly change the correctness of the answers.
-
+\castor{I'd like to have the numbers for the results above. What are the p-values, regression coefficients, etc.?}
 \begin{figure}
@@
 \subsubsection*{Exploratory Data Analysis}
-Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both confusing and clean versions.
+Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both obfuscated and clean versions.
 We do not consider cases where the participants made mistakes.
@@
 Not all atom candidates, though, require less time for the participants to predict the answer.
-In fact, for the Arithmetic as Logic and the Pre-Increment atoms, the time to give a correct answer increased
-29.06\% and 38.19\% for the clean versions---we did not confirm these
-atom candidates as atoms of confusion in the previous section.
+In fact, for  Arithmetic as Logic and Pre-Increment, the time to give a correct answer for the clean versions was
+29.06\% and 38.19\% than for the obfuscated versions---we did not confirm these atom candidates as atoms of confusion in the previous section.
 
@@
 \begin{tabular}{lccc} \toprule
-Atom & Confusing  & Clean & $\Delta$(\%) \\ \midrule
+Atom & Obfuscated  & Clean & $\Delta$(\%) \\ \midrule
 Comma Operator & 87.67 & 20.84 & -76.23 \\ 
@@
 \subsubsection*{Statistical analysis}
-We use the non-parametric \emph{Mann-Whitney U test} to
+We use the \emph{Mann-Whitney U test} to
 investigate the null hypothesis that developers 
 spend the same amount of time to correctly
-predict the outcome of clean and confusing versions of a code snippet.
+predict the outcome of clean and obfuscated versions of a code snippet.
 %, regardless of evaluating the confusing or clean version of the code.
@@
 suggest that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
-For the first three, the analysis suggests that the participants need more time to predict the outcome of the code snippet in the confusing code version.
-For the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time to predict the outcome of the code snippets in the confusing version.
+For the first three, the analysis suggests that the participants need more time to predict the outcome of the code snippet in the obfuscated code version.
+For the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time to predict the outcome of the code snippets in the obfuscated version.
 We also computed the effect size using Cliff's Delta statistic.
@@
 In this section we present the results of the interviews with the practitioners.
-We contrast the findings presented in Section~\ref{sec:latin} and Section~\ref{sec:repeated} with the opinion of software developers about their preferences regarding the confusing or clean versions of the code snippets we used in our experiment.
+We contrast the findings of the experiments of Sections~\ref{sec:repeated} and~\ref{sec:latin} with the opinion of software developers about their preferences regarding the obfuscated or clean versions of the code snippets.
 For the interview study, we only considered atoms that were analyzed in both the repeated measures (Section~\ref{sec:repeated}) and the latin square (Section~\ref{sec:latin}) studies.
@@
 We collected information regarding programming
-experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in the studies of the previous sections.
-We first presented them pairs of clean and confusing code snippets and then asked them to discuss their preferences towards one version.
+experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in the both experiments.
+We first presented them pairs of clean and obfuscated code snippets and then asked them to discuss their preferences towards one version.
 We did not indicate in any way whether any version was assumed to be confusing or not.
@@
       \cmidrule(lr){2-4}
-         Atom           & \multicolumn{1}{c}{Confusing}
+         Atom           & \multicolumn{1}{c}{Obfuscated}
                                       &  \multicolumn{1}{c}{Clean}
@@
 
-Supporting the results of the studies of Sections~\ref{sec:repeated} and~\ref{sec:latin}, Table~\ref{tab:interview-results1}
+Supporting the results of Sections~\ref{sec:repeated} and~\ref{sec:latin}, Table~\ref{tab:interview-results1}
  shows that for eight out of the
@@
 %% For a full listing of the code snippets, visit the paper repository. 
-Only for the Conditional Operator atom candidate the participants prefer the confusing version, instead of the clean one.
-Figure~\ref{code:ternary} shows the code snippets for the corresponding confusing and clean versions.
-For this atom candidate, some participants who preferred the left-hand side version (confusing) still believed that the right-hand side version (clean) was more readable.
+Only for the Conditional Operator atom candidate the participants preferred the obfuscated version, instead of the clean one.
+Figure~\ref{code:ternary} shows the code snippets for the corresponding obfuscated and clean versions.
+For this atom candidate, some participants who preferred the left-hand side version (obfuscated) still believed that the right-hand side version (clean) was more readable.
 The following quotes were extracted from the transcripts with
@@
 
-The Pre-Increment atom candidate also caused such
-divide in one of the interviewees, who
+The Pre-Increment atom candidate also caused a
+conflict in one of the interviewees, who
 regarded the clean version as simpler to understand, but would still opt to write
@@
 \begin{mq}
-  \emph{``This one is interesting, because I have written code that looks like the left-hand side [confusing version], and my colleagues complained that it was difficult to understand.
+  \emph{``This one is interesting, because I have written code that looks like the left-hand side [obfuscated version], and my colleagues complained that it was difficult to understand.
 Nowadays I prefer to write code using the version on the right-hand side [clean version]''}.
@@
 For two of the atom candidates, the interviewees were unanimous in their preference for the clean version: Comma Operator and Omitted Curly Braces.
-Not coincidentally, in both studies there were significant differences in correctness in favor of the clean versions for both atoms.
+Not coincidentally, in both experiments there were significant differences in correctness in favor of the clean versions for both atoms.
 Regarding the first one, we could often notice during the interviews that the \emph{\lhs} (with the atom of confusion) caused significant confusion among the participants.
@@
 
-In this section we present the results of our mining software repository effort, whose goal is to reveal how prevalent atom candidates are in open source JavaScript projects. 
+In this section we present the results of the  software repository mining effort, whose goal is to reveal how prevalent atom candidates are in open source JavaScript projects. 
 We mined \minedprojects open source JavaScript repositories to understand how
-often atoms of confusion arise in real software. 
-Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19}, which investigate
+often atoms arise in real software. As in the interview study, this study focused on atoms that were investigated in both experiments.
+
+Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19} which investigate
 the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
@@
 % Atoms that occur 1000 times are classified as \textit{rarely used} while atoms with more than \num{10000} occurrences are classified as \textit{highly used}; \textit{commonly used} atoms occur in between.
-We calculated the frequency of each atom per KLOC.
+We calculate the frequency of each atom per KLOC.
 The four atoms mentioned in the previous paragraph,
@@
 
-The results of our survey and interviews suggest that the Conditional Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to our survey).
+The results of our experiments and interviews suggest that Conditional Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to the Latin-square study).
 Nonetheless, the Post-Increment Expression and Automatic Semicolon Insertion atoms are listed in the top four sources of misunderstanding  (Table~\ref{tab:difference-correctness}), and also frequently appear in open source systems.
-As such, refactoring existing systems to avoid Post-Increment and Automatic Semicolon Insertion might improve the readabiliy of large amounts of code in JavaScript projects.
+As such, refactoring existing systems to avoid Post-Increment and Automatic Semicolon Insertion might improve the readability of large amounts of code in JavaScript projects.

### 2022-07-26 4_Result.tex
Put the paper in the elsarticle template. It still missing authors and affiliations. Furthermore, we still need to add the descriptions of the atoms and the discussion.

@@
 
-\begin{table}[tb]
+\begin{table}[t!]
 \caption{Summary of the correctness analysis.}
@@
 
-\begin{table}[tb]
+\begin{table}[th!]
 \caption{Hypotheses Testing (correctness).
@@
  \centering
+ {\scriptsize 
  \begin{tabular}{lrr}
@@
 \end{tabular}
+}
 \label{tab:repeated:misunderstanding}
@@
   \centering
+  {\scriptsize 
   \begin{tabular}{lrrr}
@@
   \end{tabular}
+}
   \end{table}
@@
 \centering{
+  {\scriptsize
   \begin{scriptsize}
@@
 \end{scriptsize}
-}
+}}
 \end{table}
@@
 \caption{Hypotheses Testing (misunderstanding and time). Asterisks ($^{*}$) indicate a statistically significant difference.}
-
  \centering
+ {\scriptsize
  \begin{tabular}{lccc|cc}
    \toprule
-Atom  & Chi-square test & Odds Ratio          & CI & Mann-Whitey U test & Cliff's Delta \\ 
-     &         & (misunderstanding)  &                    & (workload)             & \\ \midrule
-Comma Operator & 1.1727e-10$^{*}$ & 19.02 & (6.60, 68.22) & 5.8249e-08$^{*}$ & -0.53 \\ 
-Automatic Semicolon Insertion & 5.8352e-11$^{*}$ & 39.33 & (9.21, 356.62) & 0.09802 & -0.16 \\ 
-Post-Increment & 0.0015279$^{*}$ & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
-Omitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
-Assignment as Value & 0.0034775$^{*}$ & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
-Implicit Predicate & 0.01123$^{*}$ & 6.95 & (1.46, 66.56) & 0.0029039$^{*}$ & -0.29 \\ 
-Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05$^{*}$ & -0.39 \\ 
-Conditional Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
-Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981$^{*}$ & 0.27 \\ 
-Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & 0.01723$^{*}$ & 0.23 \\ 
+Atom  & Chi-square & Odds          & CI & Mann-Whitey& Cliff's\\ 
+ &  test &Ratio          & CI &  U test & Delta \\  \midrule
+Comma Operator & \textbf{$<$ 0.0001*} & 19.02 & (6.60, 68.22) & \textbf{$<$ 0.0001*} & -0.53 \\ 
+Automatic Semicolon Insertion & \textbf{$<$ 0.0001*} & 39.33 & (9.21, 356.62) & 0.0980 & -0.16 \\ 
+Post-Increment & \textbf{0.0015*} & 4.83 & (1.73, 15.74) & 0.1201 & 0.15 \\ 
+Omitted Curly Braces & 0.0510 & 2.35 & (1.00, 5.77) & 0.9253 & -0.01 \\ 
+Assignment as Value & \textbf{0.0035*} & 8.39 & (1.81, 79.12) & 0.8268 & 0.02 \\ 
+Implicit Predicate & \textbf{0.0112*} & 6.95 & (1.46, 66.56) & \textbf{0.0029*} & -0.29 \\ 
+Logic as Control Flow & 0.2920 & 1.54 & (0.73, 3.28) & \textbf{$<$ 0.0001*} & -0.39 \\ 
+Conditional Operator & 0.1590 & 2.73 & (0.74, 12.57) & 0.2407 & -0.12 \\ 
+Pre-Increment & 0.7015 & 1.25 & (0.55, 2.85) & \textbf{0.0062*} & 0.27 \\ 
+Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & \textbf{0.0172*} & 0.23 \\ 
     \bottomrule
- \end{tabular}
+ \end{tabular}}
  \label{tab:hypothesis-testing}
@@
     \centering
+    {\scriptsize
     \caption{Summary of participants' preferences for code snippets \emph{with} and
@@
     \end{tabular}
+    }
     \label{tab:interview-results1}
@@
   \centering
+  {\scriptsize 
 \caption{Summary of atom candidate occurrences in our dataset.}
@@
 \end{tabular}
+  }
 \label{tab:occurrences-summary}

### 2022-07-26 main.tex
Put the paper in the elsarticle template. It still missing authors and affiliations. Furthermore, we still need to add the descriptions of the atoms and the discussion.
Fist version of the table with all the atoms. It still sucks. I need to make a lot of adjustmentes to make it fit the page. Maybe I will have to break it in two.
abstract review + gitignore

--- 2022-07-25 main.tex
@@
-\documentclass[10pt,conference,blind]{IEEEtran}
-
-\IEEEoverridecommandlockouts
+\documentclass[preprint,12pt]{elsarticle}
 
@@
 
+%\usepackage{natbib}
 %\usepackage[english]{babel}
@@
 \usepackage{caption}
-\usepackage{cite}
+%\usepackage{cite}
 \usepackage{amsmath,amssymb,amsfonts}
 \usepackage{algorithmic}
+%\usepackage[backend=bibtex]{biblatex}
 \usepackage{graphicx}
-\usepackage{supertabular}
+%\usepackage{supertabular}
 \usepackage{textcomp}
@@
 \usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{tabularx} 
+%\usepackage{longtable}
+%\usepackage{tabularx} 
 \usepackage{multicol}
@@
 \begin{document}
+
+\begin{frontmatter}
 \input{atom-mdframe-config}
 
-\title{On The Impact of Atoms of Confusion in JavaScript
-Code}
+\title{On The Impact of Atoms of Confusion in JavaScriptCode}
 
-\author{\IEEEauthorblockN{Anonymous}}
+\input{0_Abstract}
 
-%% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% \and
-%% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% \and
-%% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% \and
-%% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% \and
-%% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% \and
-%% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address or ORCID}
-%% }
+\end{frontmatter}
 
-\maketitle
-
-% Número das páginas. Remover para camera-ready
-% útil para nós e Program Chair que pode ver o total de páginas facilmente
-\thispagestyle{plain}
-\pagestyle{plain}
-
-\begin{abstract}
-  Evolving legacy code is a challenging task, particularly when the code has been poorly written or uses confuse idioms and language constructs, which might increase maintenance efforts and impose a significant cognitive load on developers.
-For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in statically typed languages such as \clang, \cpplang, and Java.
-In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language.
-We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and an activity of mining open source JavaScript repositories (MSR).
-Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript.
-Considering ten atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates.
-For five of these atoms, the difference is statistically significant (accounting for p-value correction).
-Effect size is large for four atoms and medium for one of them.
-In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems.
-Four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code in the dataset.
-Altogether, our findings might help practitioners: (1) better understand the implications of atoms of confusion on understanding JavaScript code, (2) avoid writing code that is unnecessarily difficult to maintain, and (3) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
-\end{abstract}
-
-\begin{IEEEkeywords}
-code readability, program comprehension, program understanding, atoms of confusion, JavaScript code
+%\begin{IEEEkeywords}
+%code readability, program comprehension, program understanding, atoms of confusion, JavaScript code
 %datasets, neural networks, gaze detection, text tagging
-\end{IEEEkeywords}
+%\end{IEEEkeywords}
 
@@
 % \section*{Acknowledgment}
-
-\balance
-\bibliographystyle{IEEEtran}
+{\small
+\bibliographystyle{elsarticle-num-names} 
 \bibliography{reference}
-
-\vspace{12pt}
+}
 \end{document}

### 2022-07-26 atom_descriptions.tex

\footnote{https://developer.mozilla.org/en-US/docs/Web/JavaScript/Guide/Expressions\_and\_Operators}

\cite{Alves:2022:BPS}

% Please add the following required packages to your document preamble:
% \usepackage[normalem]{ulem}
% \useunder{\uline}{\ul}{}
\begin{table}[]
    \begin{tiny}
    \begin{tabular}{llll}
    \hline
    Atom Name                                                                  & Description                                                                                                                                                                                                                                                                                                                              & Obfuscated                                                                                                       & Clean                                                                                                                                     \\ \hline
    \begin{tabular}[c]{@{}l@{}}Change of Literal\\ Encoding\end{tabular}       & \begin{tabular}[c]{@{}l@{}}All numbers are stored as binary, but for convenience we represent numbers in decimal, hex, or octal.\\ Depending on the circumstance, certain representations are more understandable.\end{tabular}                                                                                                          & \begin{tabular}[c]{@{}l@{}}let V1 = 013\\ console.log(V1)\end{tabular}                                           & \begin{tabular}[c]{@{}l@{}}let V1 = 11\\ console.log(V1)\end{tabular}                                                                     \\
    Logic as Control Flow                                                      & \begin{tabular}[c]{@{}l@{}}Traditionally, the \&\& and || operators are used for logical conjunction and disjunction, respectively. Due to\\ short-circuiting, they can also be used for conditional execution.\end{tabular}                                                                                                             & V1 \&\& F2();                                                                                                    & if (V1) \{ F2() \};                                                                                                                       \\
    Assignment as Value                                                        & \begin{tabular}[c]{@{}l@{}}The assignment expression changes the state of the program when it executes, however it also returns a\\ value. When reading an assignment expression people may forget one of the two effects of the expression\end{tabular}                                                                                 & V1 = V2 = 3                                                                                                      & \begin{tabular}[c]{@{}l@{}}V2 = 3\\ V1 = V2\end{tabular}                                                                                  \\
    \begin{tabular}[c]{@{}l@{}}Pre-Increment /\\ Decrement\end{tabular}        & \begin{tabular}[c]{@{}l@{}}Similar to post-increment/decrement, these operators change a variable’s value by one. In contrast to the\\ other operators, pre-increment/decrement first update the variable then return the new value.\end{tabular}                                                                                        & V1 = ++V2                                                                                                        & \begin{tabular}[c]{@{}l@{}}V2 += 1\\ V1 = V2\end{tabular}                                                                                 \\
    \begin{tabular}[c]{@{}l@{}}Post-Increment /\\ Decrement\end{tabular}       & \begin{tabular}[c]{@{}l@{}}The post-increment and decrement operators change the value of their operands by 1 and return the original\\ value. Confusion arises because the value of the expression is different from the value of the variable.\end{tabular}                                                                            & V1 = V2++                                                                                                        & \begin{tabular}[c]{@{}l@{}}V1 = V2\\ V2 += 1\end{tabular}                                                                                 \\
    Type Conversion                                                            & \begin{tabular}[c]{@{}l@{}}The C compiler will often implicitly convert types when there is a mismatch. Sometimes this conversion also\\ results in a different outcome than a reader expects\end{tabular}                                                                                                                               & \{\} + 4                                                                                                         & \{\}.toString() + (4).toString()                                                                                                          \\
    Conditional Operator                                                       & \begin{tabular}[c]{@{}l@{}}The conditional operator is the only ternary operator in C, and functions similarly to an if/else block. However,\\ it is an expression for which the value is that of the executed branch.\end{tabular}                                                                                                      & V2 = (V1==3)?2:V2                                                                                                & \begin{tabular}[c]{@{}l@{}}if (V1 == 3) \{ \\   V2 = 2;\\ \\ \}\end{tabular}                                                              \\
    Comma Operator                                                             & \begin{tabular}[c]{@{}l@{}}The comma operator is used to sequence series of computations. Whether due to its eccentricity, or its odd\\ precedence, the comma operator is commonly misinterpreted.\end{tabular}                                                                                                                          & V3 = (V1 += 1; V1)                                                                                               & \begin{tabular}[c]{@{}l@{}}V1 += 1;\\ V3 = V1;\end{tabular}                                                                               \\
    Implicit Predicate                                                         & \begin{tabular}[c]{@{}l@{}}The semantics of a predicate are easily mistaken. The most common example happens when assignment is\\ used inside a predicate or when a success state is represented as 0.\end{tabular}                                                                                                                      & if (4 \% 2)                                                                                                      & if (4 \% 2 != 0)                                                                                                                          \\
    Repurposed Variables                                                       & When a variable is used in different roles in a program, its current meaning can be difficult to follow                                                                                                                                                                                                                                  & \begin{tabular}[c]{@{}l@{}}let n = 1;\\ n = "OK";\end{tabular}                                                   & \begin{tabular}[c]{@{}l@{}}let n = 1;\\ let s = "OK";\end{tabular}                                                                        \\
    Omitted Curly Braces                                                       & When control statements omit curly braces, the scope of their influence can be difficult to discern                                                                                                                                                                                                                                      & if (V) F(); G();                                                                                                 & if (V) \{F();\} G();                                                                                                                      \\
    \begin{tabular}[c]{@{}l@{}}Infix Operator \\ Precedence\end{tabular}       & \begin{tabular}[c]{@{}l@{}}JavaScript has more than 40 operators, the majority of them infix, each in one of 17 precedence levels with either right or left associativity. Most\\ programmers know only a functional subset of these rules.\end{tabular}                                                                                 & 0 \&\& 1 || 2                                                                                                    & (0 \&\& 1) || 2                                                                                                                           \\
    Arithmetic as Logic                                                        & \begin{tabular}[c]{@{}l@{}}Arithmetic operators are capable of mimicking any predicate formulated with logical operators. Arithmetic,\\ however, implies a non-boolean range, which may be confusing to a reader\end{tabular}                                                                                                            & (V1-3) * (V2-4)                                                                                                  & V1!=3 \&\& V2!=4                                                                                                                          \\
    Constant Variables                                                         & \begin{tabular}[c]{@{}l@{}}Constant variables are a layer of abstraction that emphasize a concept rather than a particular value itself.\\ When trying to hand evaluate a piece of code having a layer of indirection can obscure the value of the data.\end{tabular}                                                                    & \begin{tabular}[c]{@{}l@{}}let V1 = 5\\ console.log(V1)\end{tabular}                                             & console.log(5)                                                                                                                            \\
    \begin{tabular}[c]{@{}l@{}}Dead, Unreachable,\\ Repeated\end{tabular}      & Redundant code is executed to no functional effect, but its appearance may imply that meaningful changes are being made                                                                                                                                                                                                                  & \begin{tabular}[c]{@{}l@{}}V1 = 1;\\ V1 = 2;\end{tabular}                                                        & V1 = 2;                                                                                                                                   \\
    \begin{tabular}[c]{@{}l@{}}Lack of Indentation, \\ no Braces\end{tabular}  & \begin{tabular}[c]{@{}l@{}}Block delimiters and indentation are very important to understand nested structures in programs. If block delimiters are missing and\\ indentation is incorrect, developers may misunderstand the nested structure of the program.\end{tabular}                                                               & \begin{tabular}[c]{@{}l@{}}if (V1)\\     if (V2) \\         V3 = V3 + 2;\\ else \\     V3 = V3 - 1;\end{tabular} & \begin{tabular}[c]{@{}l@{}}if (V1) \{\\     if (V2) \{ V3 = V3 + 2; \}\\     else \{ V3 = V3 - 1; \}\\ \}\end{tabular}                    \\
    \begin{tabular}[c]{@{}l@{}}Lack of Indentation,\\ with Braces\end{tabular} & \begin{tabular}[c]{@{}l@{}}Similar to Lack of Indentation, no Braces, but the block delimiters are present. The lack of indentation still has the potential to cause\\ confusion.\end{tabular}                                                                                                                                           & \begin{tabular}[c]{@{}l@{}}if (V1) \{\};\\     V1 = V1 - 1;\end{tabular}                                         & \begin{tabular}[c]{@{}l@{}}if (V1) \{\};\\ V1 = V1 - 1;\end{tabular}                                                                      \\
    \begin{tabular}[c]{@{}l@{}}Automatic Semicolon\\ Insertion\end{tabular}    & \begin{tabular}[c]{@{}l@{}}Semicolons in JavaScript are optional in some contexts. However, some cases where semicolons are optional can confuse the JavaScript\\ parser. This may lead to surprising behavior.\end{tabular}                                                                                                             & \begin{tabular}[c]{@{}l@{}}let V1 = 9\\ (console.log(V1))\end{tabular}                                           & \begin{tabular}[c]{@{}l@{}}let V1 = 9;\\ (console.log(V1));\end{tabular}                                                                  \\
    Property Access                                                            & \begin{tabular}[c]{@{}l@{}}Object properties can be accessed using the dot (".") notation used in most object oriented languages and also using notation similar to \\ how elements in collections like dictionaries and maps are accessed in other languages. The latter is much less popular than the former.\end{tabular}             & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1:0, P2:5 \}\\ V1{[}'P1'{]} = 9;\end{tabular}                             & \begin{tabular}[c]{@{}l@{}}let V2 = \{P3:0, P4:5 \}\\ V2{[}'P3'{]} = 9;\end{tabular}                                                      \\
    Arrow Function                                                             & The usage of arrow function notation, instead of the more traditional JavaScript notations for anonymous functions.                                                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}let V1 = x =\textgreater x * x * x;\\ let V2 = V1(5);\end{tabular}                    & \begin{tabular}[c]{@{}l@{}}let V3 = (x) =\textgreater \{ \\   return x * x * x; \\ \}\\ let V4 = V3(5);\end{tabular}                      \\
    Array Spread                                                               & \begin{tabular}[c]{@{}l@{}}Spread facilitates copying, extracting elements and concatenating arrays. Some caveats exist as the copying is shallow so this can \\ cause misunderstanding between programmers.\end{tabular}                                                                                                                & \begin{tabular}[c]{@{}l@{}}let V1 = {[}1,2{]};\\ let V2 = {[}5,6{]};\\ V3 = {[}...V1, ...V2{]}\end{tabular}      & \begin{tabular}[c]{@{}l@{}}let V1 = {[}1,2{]};\\ let V2 = {[}5,6{]};\\ V3 = {[}V1{[}0{]}, V1{[}1{]}, V2{[}0{]}, V2{[}1{]}{]}\end{tabular} \\
    Object Spread                                                              & \begin{tabular}[c]{@{}l@{}}Similar to the spread in arrays, but with slightly more complex semantics because it may lead to the values of properties being\\ overwritten if properties with the same names exist in two objects combined by means of spread.\end{tabular}                                                                & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1:5, P2:8, P3:13\};\\ let V2 = \{...V1, P3:99\};\end{tabular}             & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1:5, P2:8, P3:13\};\\ let V2 = \{P3:99\};\\ V2.P1 = V1.P1\\ V2.P2 = V1.P2\end{tabular}             \\
    Array Destructuring                                                        & \begin{tabular}[c]{@{}l@{}}This feature provides an approach to perform a limited form of pattern matching based on the elements of an array. The syntax mixes the \\ introduction of new variables with array literals and is arguably less explicit than directly initializing new variables with array elements.\end{tabular}         & \begin{tabular}[c]{@{}l@{}}let {[}V2,,,V4{]} = {[}1,2,3,4{]};\\ console.log(V2, V4);\end{tabular}                & \begin{tabular}[c]{@{}l@{}}let V1 = {[}1,2,3,4{]};\\ console.log(V1{[}0{]}, V1{[}3{]});\end{tabular}                                      \\
    Object Destructuring                                                       & \begin{tabular}[c]{@{}l@{}}Similar to array destructuring, but uses the syntax associated with object literals to introduce new variables by matching their names with the \\ names of properties of the destructured object. This is less explicit than initializing the new variables with the values of said properties.\end{tabular} & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1: 5\};\\ let \{P1, P2 = 20\} = V1;\end{tabular}                          & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1: 5\};\\ let P1=V1.P1, P2=20;\end{tabular}                                                       
    \end{tabular}
\end{tiny}
    \end{table}

### 2022-07-27 3_Method.tex
Put the paper in the elsarticle template. It still missing authors and affiliations. Furthermore, we still need to add the descriptions of the atoms and the discussion.
Added the two tables with the descriptions and examples for all the atoms investigated in the two experiments. They are ugly and in landscape mode, but that was the best I could do.

@@
 Overall, the repeated measures study analyzes 24 atoms, 15 from the original study and 9 that are new.
-Table~\ref{tab:atoms_replication} presents the complete list.
+Some of these atoms are also analyzed in the Latin square study (Section~\ref{sec:meth:survey}).
+Table~\ref{tab:atomsBoth} presents the atom candidates that both studies investigate and Table~\ref{tab:atomsRepeated} presents the ones that only the repeated measures study analyzes.
+
+\input{atom_descriptions}
 
@@
 \begin{table}[htb!]
+
   \centering

### 2022-07-27 atom_descriptions.tex
Added the two tables with the descriptions and examples for all the atoms investigated in the two experiments. They are ugly and in landscape mode, but that was the best I could do.

@@
-\footnote{https://developer.mozilla.org/en-US/docs/Web/JavaScript/Guide/Expressions\_and\_Operators}
-
-\cite{Alves:2022:BPS}
-
 % Please add the following required packages to your document preamble:
@@
 % \useunder{\uline}{\ul}{}
-\begin{table}[]
+\begin{sidewaystable}[]
+%    \begin{adjustbox}{angle=90}
+%    \begin{sideways}
          \begin{tiny}
+            \caption{Atom candidates analyzed in the two experiments.
+Nine of these atoms come from the original study of Gopstein and colleagues~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
+For these atoms, both the descriptions and the examples, with the required adaptations, are taken directly from their paper.}
+            \label{tab:atomsBoth}        
             \begin{tabular}{llll}
@@
                 Atom Name                                                                     & Description                                                                                                                                                                                                                                                                                                                                                         & Obfuscated                                                                                                       & Clean                                                                                                                         \\ \hline
-    \begin{tabular}[c]{@{}l@{}}Change of Literal\\ Encoding\end{tabular}       & \begin{tabular}[c]{@{}l@{}}All numbers are stored as binary, but for convenience we represent numbers in decimal, hex, or octal.\\ Depending on the circumstance, certain representations are more understandable.\end{tabular}                                                                                                          & \begin{tabular}[c]{@{}l@{}}let V1 = 013\\ console.log(V1)\end{tabular}                                           & \begin{tabular}[c]{@{}l@{}}let V1 = 11\\ console.log(V1)\end{tabular}                                                                     \\
-    Logic as Control Flow                                                      & \begin{tabular}[c]{@{}l@{}}Traditionally, the \&\& and || operators are used for logical conjunction and disjunction, respectively.
-Due to\\ short-circuiting, they can also be used for conditional execution.\end{tabular}                                                                                                             & V1 \&\& F2();                                                                                                    & if (V1) \{ F2() \};                                                                                                                       \\
-    Assignment as Value                                                        & \begin{tabular}[c]{@{}l@{}}The assignment expression changes the state of the program when it executes, however it also returns a\\ value.
-When reading an assignment expression people may forget one of the two effects of the expression\end{tabular}                                                                                 & V1 = V2 = 3                                                                                                      & \begin{tabular}[c]{@{}l@{}}V2 = 3\\ V1 = V2\end{tabular}                                                                                  \\
-    \begin{tabular}[c]{@{}l@{}}Pre-Increment /\\ Decrement\end{tabular}        & \begin{tabular}[c]{@{}l@{}}Similar to post-increment/decrement, these operators change a variable’s value by one.
-In contrast to the\\ other operators, pre-increment/decrement first update the variable then return the new value.\end{tabular}                                                                                        & V1 = ++V2                                                                                                        & \begin{tabular}[c]{@{}l@{}}V2 += 1\\ V1 = V2\end{tabular}                                                                                 \\
-    \begin{tabular}[c]{@{}l@{}}Post-Increment /\\ Decrement\end{tabular}       & \begin{tabular}[c]{@{}l@{}}The post-increment and decrement operators change the value of their operands by 1 and return the original\\ value.
-Confusion arises because the value of the expression is different from the value of the variable.\end{tabular}                                                                            & V1 = V2++                                                                                                        & \begin{tabular}[c]{@{}l@{}}V1 = V2\\ V2 += 1\end{tabular}                                                                                 \\
-    Type Conversion                                                            & \begin{tabular}[c]{@{}l@{}}The C compiler will often implicitly convert types when there is a mismatch.
-Sometimes this conversion also\\ results in a different outcome than a reader expects\end{tabular}                                                                                                                               & \{\} + 4                                                                                                         & \{\}.toString() + (4).toString()                                                                                                          \\
-    Conditional Operator                                                       & \begin{tabular}[c]{@{}l@{}}The conditional operator is the only ternary operator in C, and functions similarly to an if/else block.
-However,\\ it is an expression for which the value is that of the executed branch.\end{tabular}                                                                                                      & V2 = (V1==3)?2:V2                                                                                                & \begin{tabular}[c]{@{}l@{}}if (V1 == 3) \{ \\   V2 = 2;\\ \\ \}\end{tabular}                                                              \\
+                \begin{tabular}[c]{@{}l@{}}Logic as Control\\ Flow\end{tabular}               & \begin{tabular}[c]{@{}l@{}}Traditionally, the \&\& and || operators are used for logical conjunction \\ and disjunction, respectively.
+Due to short-circuiting, they can also be used for \\ conditional execution.\end{tabular}                                                                                                                                    & V1 \&\& F2();                                                                                                    & if (V1) \{ F2() \};                                                                                                           \\
+                \begin{tabular}[c]{@{}l@{}}Assignment as \\ Value\end{tabular}                & \begin{tabular}[c]{@{}l@{}}The assignment expression changes the state of the program when it executes, \\ however it also returns a value.
+When reading an assignment expression people\\ may forget one of the two effects of the expression\end{tabular}                                                                                                         & V1 = V2 = 3                                                                                                      & \begin{tabular}[c]{@{}l@{}}V2 = 3\\ V1 = V2\end{tabular}                                                                      \\
+                \begin{tabular}[c]{@{}l@{}}Pre-Increment /\\ Decrement\end{tabular}           & \begin{tabular}[c]{@{}l@{}}Similar to post-increment/decrement, these operators change a variable’s value\\ by one.
+In contrast to the other operators, pre-increment/decrement first update \\ the variable then return the new value.\end{tabular}                                                                                                                & V1 = ++V2                                                                                                        & \begin{tabular}[c]{@{}l@{}}V2 += 1\\ V1 = V2\end{tabular}                                                                     \\
+                \begin{tabular}[c]{@{}l@{}}Post-Increment /\\ Decrement\end{tabular}          & \begin{tabular}[c]{@{}l@{}}The post-increment and decrement operators change the value of their operands\\ by 1 and return the original value.
+Confusion arises because the value of the \\ expression is different from the value of the variable.\end{tabular}                                                                                                    & V1 = V2++                                                                                                        & \begin{tabular}[c]{@{}l@{}}V1 = V2\\ V2 += 1\end{tabular}                                                                     \\
                 Comma Operator                                                                & \begin{tabular}[c]{@{}l@{}}The comma operator is used to sequence series of computations.
-Whether due to its eccentricity, or its odd\\ precedence, the comma operator is commonly misinterpreted.\end{tabular}                                                                                                                          & V3 = (V1 += 1; V1)                                                                                               & \begin{tabular}[c]{@{}l@{}}V1 += 1;\\ V3 = V1;\end{tabular}                                                                               \\
+Whether due \\ to its eccentricity, or its odd precedence, the comma operator is commonly \\ misinterpreted.\end{tabular}                                                                                                                                                 & V3 = (V1 += 1; V1)                                                                                               & \begin{tabular}[c]{@{}l@{}}V1 += 1;\\ V3 = V1;\end{tabular}                                                                   \\
                 Implicit Predicate                                                            & \begin{tabular}[c]{@{}l@{}}The semantics of a predicate are easily mistaken.
-The most common example happens when assignment is\\ used inside a predicate or when a success state is represented as 0.\end{tabular}                                                                                                                      & if (4 \% 2)                                                                                                      & if (4 \% 2 != 0)                                                                                                                          \\
-    Repurposed Variables                                                       & When a variable is used in different roles in a program, its current meaning can be difficult to follow                                                                                                                                                                                                                                  & \begin{tabular}[c]{@{}l@{}}let n = 1;\\ n = "OK";\end{tabular}                                                   & \begin{tabular}[c]{@{}l@{}}let n = 1;\\ let s = "OK";\end{tabular}                                                                        \\
-    Omitted Curly Braces                                                       & When control statements omit curly braces, the scope of their influence can be difficult to discern                                                                                                                                                                                                                                      & if (V) F(); G();                                                                                                 & if (V) \{F();\} G();                                                                                                                      \\
-    \begin{tabular}[c]{@{}l@{}}Infix Operator \\ Precedence\end{tabular}       & \begin{tabular}[c]{@{}l@{}}JavaScript has more than 40 operators, the majority of them infix, each in one of 17 precedence levels with either right or left associativity.
-Most\\ programmers know only a functional subset of these rules.\end{tabular}                                                                                 & 0 \&\& 1 || 2                                                                                                    & (0 \&\& 1) || 2                                                                                                                           \\
-    Arithmetic as Logic                                                        & \begin{tabular}[c]{@{}l@{}}Arithmetic operators are capable of mimicking any predicate formulated with logical operators.
-Arithmetic,\\ however, implies a non-boolean range, which may be confusing to a reader\end{tabular}                                                                                                            & (V1-3) * (V2-4)                                                                                                  & V1!=3 \&\& V2!=4                                                                                                                          \\
-    Constant Variables                                                         & \begin{tabular}[c]{@{}l@{}}Constant variables are a layer of abstraction that emphasize a concept rather than a particular value itself.\\ When trying to hand evaluate a piece of code having a layer of indirection can obscure the value of the data.\end{tabular}                                                                    & \begin{tabular}[c]{@{}l@{}}let V1 = 5\\ console.log(V1)\end{tabular}                                             & console.log(5)                                                                                                                            \\
-    \begin{tabular}[c]{@{}l@{}}Dead, Unreachable,\\ Repeated\end{tabular}      & Redundant code is executed to no functional effect, but its appearance may imply that meaningful changes are being made                                                                                                                                                                                                                  & \begin{tabular}[c]{@{}l@{}}V1 = 1;\\ V1 = 2;\end{tabular}                                                        & V1 = 2;                                                                                                                                   \\
-    \begin{tabular}[c]{@{}l@{}}Lack of Indentation, \\ no Braces\end{tabular}  & \begin{tabular}[c]{@{}l@{}}Block delimiters and indentation are very important to understand nested structures in programs.
-If block delimiters are missing and\\ indentation is incorrect, developers may misunderstand the nested structure of the program.\end{tabular}                                                               & \begin{tabular}[c]{@{}l@{}}if (V1)\\     if (V2) \\         V3 = V3 + 2;\\ else \\     V3 = V3 - 1;\end{tabular} & \begin{tabular}[c]{@{}l@{}}if (V1) \{\\     if (V2) \{ V3 = V3 + 2; \}\\     else \{ V3 = V3 - 1; \}\\ \}\end{tabular}                    \\
-    \begin{tabular}[c]{@{}l@{}}Lack of Indentation,\\ with Braces\end{tabular} & \begin{tabular}[c]{@{}l@{}}Similar to Lack of Indentation, no Braces, but the block delimiters are present.
-The lack of indentation still has the potential to cause\\ confusion.\end{tabular}                                                                                                                                           & \begin{tabular}[c]{@{}l@{}}if (V1) \{\};\\     V1 = V1 - 1;\end{tabular}                                         & \begin{tabular}[c]{@{}l@{}}if (V1) \{\};\\ V1 = V1 - 1;\end{tabular}                                                                      \\
-    \begin{tabular}[c]{@{}l@{}}Automatic Semicolon\\ Insertion\end{tabular}    & \begin{tabular}[c]{@{}l@{}}Semicolons in JavaScript are optional in some contexts.
-However, some cases where semicolons are optional can confuse the JavaScript\\ parser.
-This may lead to surprising behavior.\end{tabular}                                                                                                             & \begin{tabular}[c]{@{}l@{}}let V1 = 9\\ (console.log(V1))\end{tabular}                                           & \begin{tabular}[c]{@{}l@{}}let V1 = 9;\\ (console.log(V1));\end{tabular}                                                                  \\
-    Property Access                                                            & \begin{tabular}[c]{@{}l@{}}Object properties can be accessed using the dot (".") notation used in most object oriented languages and also using notation similar to \\ how elements in collections like dictionaries and maps are accessed in other languages.
-The latter is much less popular than the former.\end{tabular}             & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1:0, P2:5 \}\\ V1{[}'P1'{]} = 9;\end{tabular}                             & \begin{tabular}[c]{@{}l@{}}let V2 = \{P3:0, P4:5 \}\\ V2{[}'P3'{]} = 9;\end{tabular}                                                      \\
-    Arrow Function                                                             & The usage of arrow function notation, instead of the more traditional JavaScript notations for anonymous functions.
-& \begin{tabular}[c]{@{}l@{}}let V1 = x =\textgreater x * x * x;\\ let V2 = V1(5);\end{tabular}                    & \begin{tabular}[c]{@{}l@{}}let V3 = (x) =\textgreater \{ \\   return x * x * x; \\ \}\\ let V4 = V3(5);\end{tabular}                      \\
-    Array Spread                                                               & \begin{tabular}[c]{@{}l@{}}Spread facilitates copying, extracting elements and concatenating arrays.
-Some caveats exist as the copying is shallow so this can \\ cause misunderstanding between programmers.\end{tabular}                                                                                                                & \begin{tabular}[c]{@{}l@{}}let V1 = {[}1,2{]};\\ let V2 = {[}5,6{]};\\ V3 = {[}...V1, ...V2{]}\end{tabular}      & \begin{tabular}[c]{@{}l@{}}let V1 = {[}1,2{]};\\ let V2 = {[}5,6{]};\\ V3 = {[}V1{[}0{]}, V1{[}1{]}, V2{[}0{]}, V2{[}1{]}{]}\end{tabular} \\
-    Object Spread                                                              & \begin{tabular}[c]{@{}l@{}}Similar to the spread in arrays, but with slightly more complex semantics because it may lead to the values of properties being\\ overwritten if properties with the same names exist in two objects combined by means of spread.\end{tabular}                                                                & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1:5, P2:8, P3:13\};\\ let V2 = \{...V1, P3:99\};\end{tabular}             & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1:5, P2:8, P3:13\};\\ let V2 = \{P3:99\};\\ V2.P1 = V1.P1\\ V2.P2 = V1.P2\end{tabular}             \\
-    Array Destructuring                                                        & \begin{tabular}[c]{@{}l@{}}This feature provides an approach to perform a limited form of pattern matching based on the elements of an array.
-The syntax mixes the \\ introduction of new variables with array literals and is arguably less explicit than directly initializing new variables with array elements.\end{tabular}         & \begin{tabular}[c]{@{}l@{}}let {[}V2,,,V4{]} = {[}1,2,3,4{]};\\ console.log(V2, V4);\end{tabular}                & \begin{tabular}[c]{@{}l@{}}let V1 = {[}1,2,3,4{]};\\ console.log(V1{[}0{]}, V1{[}3{]});\end{tabular}                                      \\
-    Object Destructuring                                                       & \begin{tabular}[c]{@{}l@{}}Similar to array destructuring, but uses the syntax associated with object literals to introduce new variables by matching their names with the \\ names of properties of the destructured object.
-This is less explicit than initializing the new variables with the values of said properties.\end{tabular} & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1: 5\};\\ let \{P1, P2 = 20\} = V1;\end{tabular}                          & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1: 5\};\\ let P1=V1.P1, P2=20;\end{tabular}
+The most common example \\ happens when assignment is used inside a predicate or when a success state is \\ represented as 0.\end{tabular}                                                                                                                                             & if (4 \% 2)                                                                                                      & if (4 \% 2 != 0)                                                                                                              \\
+                \begin{tabular}[c]{@{}l@{}}Conditional \\ Operator\end{tabular}               & \begin{tabular}[c]{@{}l@{}}The conditional operator is the only ternary operator in C, and functions \\ similarly to an if/else block.
+However, it is an expression for which the value is \\ that of the executed branch.\end{tabular}                                                                                                                             & V2 = (V1==3)?2:V2                                                                                                & \begin{tabular}[c]{@{}l@{}}if (V1 == 3) \{ \\   V2 = 2;\\ \\ \}\end{tabular}                                                  \\
+                \begin{tabular}[c]{@{}l@{}}Automatic \\ Semicolon\\ Insertion\end{tabular}    & \begin{tabular}[c]{@{}l@{}}Semicolons in JavaScript are optional in some contexts.
+However, some cases \\ where semicolons are optional can confuse the JavaScript parser.
+This may lead \\ to surprising behavior.\end{tabular}                                                                                                                                    & \begin{tabular}[c]{@{}l@{}}let V1 = 9\\ (console.log(V1))\end{tabular}                                           & \begin{tabular}[c]{@{}l@{}}let V1 = 9;\\ (console.log(V1));\end{tabular}                                                      \\
+                \begin{tabular}[c]{@{}l@{}}Omitted Curly \\ Braces\end{tabular}               & \begin{tabular}[c]{@{}l@{}}When control statements omit curly braces, the scope of their influence can be \\ difficult to discern\end{tabular}                                                                                                                                                                                                                      & if (V) F(); G();                                                                                                 & if (V) \{F();\} G();                                                                                                          \\
+                \begin{tabular}[c]{@{}l@{}}Arithmetic as \\ Logic\end{tabular}                & \begin{tabular}[c]{@{}l@{}}Arithmetic operators are capable of mimicking any predicate formulated with \\ logical operators.
+Arithmetic, however, implies a non-boolean range, which may \\ be confusing to a reader\end{tabular}                                                                                                                                   & (V1-3) * (V2-4)                                                                                                  & V1!=3 \&\& V2!=4                                                                                                              \\
+                \hline
             \end{tabular}
         \end{tiny}
-    \end{table}
+\end{sidewaystable}
+
+\begin{sidewaystable}[]
+%    \begin{adjustbox}{angle=90}
+%    \begin{sideways}
+        \begin{tiny}
+            \caption{Atom candidates analyzed only in the repeated measures experiment.
+Six of these atoms come from the original study of Gopstein and colleagues~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
+For these atoms, both the descriptions and the examples, with the required adaptations, are taken directly from their paper.}
+            \label{tab:atomsRepeated}            
+          \begin{tabular}{llll}
+                \hline
+                Atom Name                                                                     & Description                                                                                                                                                                                  & Obfuscated                                                                                                       & Clean                                                                                                                         \\ \hline
+                Type Conversion                                                               & \begin{tabular}[c]{@{}l@{}}JavaScript will often implicitly convert types when there is a mismatch.
+\\ Sometimes this conversion also results in a different outcome than a reader expects\end{tabular}                                                                                                                                                         & \{\} + 4                                                                                                         & \{\}.toString() + (4).toString()                                                                                              \\
+                \begin{tabular}[c]{@{}l@{}}Repurposed \\ Variables\end{tabular}               & \begin{tabular}[c]{@{}l@{}}When a variable is used in different roles in a program, its current meaning can \\ be difficult to follow\end{tabular}                                                                                                                                                                                                                  & \begin{tabular}[c]{@{}l@{}}let n = 1;\\ n = "OK";\end{tabular}                                                   & \begin{tabular}[c]{@{}l@{}}let n = 1;\\ let s = "OK";\end{tabular}                                                            \\
+                \begin{tabular}[c]{@{}l@{}}Change of Literal\\ Encoding\end{tabular}          & \begin{tabular}[c]{@{}l@{}}All numbers are stored as binary, but for convenience we represent numbers in \\ decimal, hex, or octal.
+Depending on the circumstance, certain representations \\ are more understandable.\end{tabular}                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}let V1 = 013\\ console.log(V1)\end{tabular}                                           & \begin{tabular}[c]{@{}l@{}}let V1 = 11\\ console.log(V1)\end{tabular}                                                         \\
+                \begin{tabular}[c]{@{}l@{}}Infix Operator \\ Precedence\end{tabular}          & \begin{tabular}[c]{@{}l@{}}JavaScript has more than 40 operators\footnote{https://developer.mozilla.org/en-US/docs/Web/JavaScript/Guide/Expressions\_and\_Operators}
+                    , the majority of them infix, each in one of \\ 17 precedence levels with either right or left associativity.
+Most programmers \\ know only a functional subset of these rules.\end{tabular}                                                                                                        & 0 \&\& 1 || 2                                                                                                    & (0 \&\& 1) || 2                                                                                                               \\
+                \begin{tabular}[c]{@{}l@{}}Constant \\ Variables\end{tabular}                 & \begin{tabular}[c]{@{}l@{}}Constant variables are a layer of abstraction that emphasize a concept rather than \\ a particular value itself.
+When trying to hand evaluate a piece of code having a \\ layer of indirection can obscure the value of the data.\end{tabular}                                                                                           & \begin{tabular}[c]{@{}l@{}}let V1 = 5\\ console.log(V1)\end{tabular}                                             & console.log(5)                                                                                                                \\
+                \begin{tabular}[c]{@{}l@{}}Dead, \\ Unreachable,\\ Repeated\end{tabular}      & \begin{tabular}[c]{@{}l@{}}Redundant code is executed to no functional effect, but its appearance may imply \\ that meaningful changes are being made\end{tabular}                                                                                                                                                                                                  & \begin{tabular}[c]{@{}l@{}}V1 = 1;\\ V1 = 2;\end{tabular}                                                        & V1 = 2;                                                                                                                       \\
+                \begin{tabular}[c]{@{}l@{}}Lack of \\ Indentation, \\ no Braces\end{tabular}  & \begin{tabular}[c]{@{}l@{}}Block delimiters and indentation are very important to understand nested \\ structures in programs.
+If block delimiters are missing and indentation is incorrect,\\ developers may misunderstand the nested structure of the program.\end{tabular}                                                                                       & \begin{tabular}[c]{@{}l@{}}if (V1)\\     if (V2) \\         V3 = V3 + 2;\\ else \\     V3 = V3 - 1;\end{tabular} & \begin{tabular}[c]{@{}l@{}}if (V1) \{\\     if (V2) \{ V3 = V3 + 2; \}\\     else \{ V3 = V3 - 1; \}\\ \}\end{tabular}        \\
+                \begin{tabular}[c]{@{}l@{}}Lack of \\ Indentation,\\ with Braces\end{tabular} & \begin{tabular}[c]{@{}l@{}}Similar to Lack of Indentation, no Braces, but the block delimiters are present.
+\\ The lack of indentation still has the potential to cause confusion.\end{tabular}                                                                                                                                                                     & \begin{tabular}[c]{@{}l@{}}if (V1) \{\};\\     V1 = V1 - 1;\end{tabular}                                         & \begin{tabular}[c]{@{}l@{}}if (V1) \{\};\\ V1 = V1 - 1;\end{tabular}                                                          \\
+                Property Access                                                               & \begin{tabular}[c]{@{}l@{}}Object properties can be accessed using the dot (".") notation used in most object-\\ oriented languages and using notation similar to how elements in collections like \\ dictionaries and maps are accessed in other languages.
+The latter is less popular\cite{Alves:2022:BPS}
+                    \\ than the former.
+Lack of familiarity may lead to confusion.\end{tabular} & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1:0, P2:5 \}\\ V1{[}'P1'{]} = 9;\end{tabular}                             & \begin{tabular}[c]{@{}l@{}}let V2 = \{P3:0, P4:5 \}\\ V2{[}'P3'{]} = 9;\end{tabular}                                          \\
+                Arrow Function                                                                & \begin{tabular}[c]{@{}l@{}}The usage of arrow function notation, instead of the more traditional JavaScript \\ notations for anonymous functions.\end{tabular}                                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}let V1 = x =\textgreater x * x * x;\\ let V2 = V1(5);\end{tabular}                    & \begin{tabular}[c]{@{}l@{}}let V3 = (x) =\textgreater \{ \\   return x * x * x; \\ \}\\ let V4 = V3(5);\end{tabular}          \\
+                Array Spread                                                                  & \begin{tabular}[c]{@{}l@{}}Spread facilitates copying, extracting elements and concatenating arrays.
+Some \\ caveats exist as the copying is shallow so this can cause misunderstanding \\ between programmers.\end{tabular}                                                                                                                                        & \begin{tabular}[c]{@{}l@{}}let V1 = {[}1,2{]};\\ let V2 = {[}5{]};\\ V3 = {[}...V1, ...V2{]}\end{tabular}        & \begin{tabular}[c]{@{}l@{}}let V1 = {[}1,2{]};\\ let V2 = {[}5{]};\\ V3 = {[}V1{[}0{]}, V1{[}1{]}, V2{[}0{]}{]}\end{tabular}  \\
+                Object Spread                                                                 & \begin{tabular}[c]{@{}l@{}}Similar to the spread in arrays, but with slightly more complex semantics \\ because it may lead to the values of properties being overwritten if properties with \\ the same names exist in two objects combined by means of spread.\end{tabular}                                                                                       & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1:5, P2:8, P3:13\};\\ let V2 = \{...V1, P3:99\};\end{tabular}             & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1:5, P2:8, P3:13\};\\ let V2 = \{P3:99\};\\ V2.P1 = V1.P1\\ V2.P2 = V1.P2\end{tabular} \\
+                \begin{tabular}[c]{@{}l@{}}Array \\ Destructuring\end{tabular}                & \begin{tabular}[c]{@{}l@{}}This feature provides an approach to perform a limited form of pattern matching \\ based on the elements of an array.
+The syntax mixes the introduction of new \\ variables with array literals and is arguably less explicit than directly initializing \\ new variables with array elements.\end{tabular}                              & \begin{tabular}[c]{@{}l@{}}let {[}V2,,,V4{]} = {[}1,2,3,4{]};\\ console.log(V2, V4);\end{tabular}                & \begin{tabular}[c]{@{}l@{}}let V1 = {[}1,2,3,4{]};\\ console.log(V1{[}0{]}, V1{[}3{]});\end{tabular}                          \\
+                \begin{tabular}[c]{@{}l@{}}Object \\ Destructuring\end{tabular}               & \begin{tabular}[c]{@{}l@{}}Similar to array destructuring, but uses the syntax associated with object literals\\ to introduce new variables by matching their names with the names of properties\\ of the destructured object.
+This is less explicit than initializing the new variables \\ with the values of said properties.\end{tabular}                        & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1: 5\};\\ let \{P1, P2 = 20\} = V1;\end{tabular}                          & \begin{tabular}[c]{@{}l@{}}let V1 = \{P1: 5\};\\ let P1=V1.P1, P2=20;\end{tabular}    \\ \hline
+                \end{tabular}
+            \end{tiny}
+%        \end{sideways}
+%\end{adjustbox}
+    \end{sidewaystable}

### 2022-07-27 0_Abstract.tex

--- 2020-11-09 0_Abstract.tex
@@
 \begin{abstract}
-{\bf Context.} Engineering maintainable software is essential and must take into account \emph{how understandable the code is} from a human perspective, and spending much cognitive effort might discourage a developer when trying to familiarize herself with a legacy codebase.
-\noindent{\bf Problem.} 
-%Besides lack of experience and other factors, 
-Previous research works have discussed the impact of small code patterns that contribute to misunderstanding the source code written in statically typed languages (e.g., \clang, \cpplang, and Swift).
-However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamically typed languages.
-\noindent{\bf Research Method and Goal.} In this work, we fill this research gap by presenting the results of a mixed-methods research:
-a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
-Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
-\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead to a code that is hard to understand: 
-%removing the atoms of confusion lead to an improvement above 15\% in 6 out of 9 atoms we consider in our study when developers try to predict the outcomes of code examples. 
-developers correctly predict the outcome of at least 15\% more when an atom of confusion is not present for 6 out of 9 atoms.
-Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the \texttt{comma operator} is present in 11\% of the projects in our dataset; while five other atoms appear in more than 60\% of the projects---including the atoms \texttt{assignment as value} and \texttt{omitted curly braces}, which appear \num{8253} and \num{53357} times, respectively.
-Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code.
-For instance, even experienced JavaScript developers agree that it is hard to understand code that uses callback functions, even though we found widespread use of \texttt{callbacks} in our projects dataset.
-\noindent{\bf Implications.} Altogether, our findings could help developers write cleaner JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code.
+Evolving software is particularly challenging when the code has been poorly written or
+uses confusing idioms and language constructs, which might increase maintenance efforts
+and impose a significant cognitive load on developers. For this reason, researchers have
+investigated possible sources of confusion in codebases, including the impact of small code
+patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code.
+Although researchers have explored atoms of confusion in code written in \clang, \cpplang,
+and Java, different languages have different features, developer communities, and development cultures. This
+justifies the exploration of other languages to verify whether they also exhibit confusion-inducing patterns. 
+In this paper we investigate the impact of atoms of confusion on understanding JavaScript code---a
+dynamically typed language whose popularity is growing in the most diverse application domains. To this end,
+we carry out a mixed-methods research comprising two experiments, a set of interviews with practitioners,
+and a mining software repositories (MSR) effort. 
+The results of both experiments suggest that two code patterns that have been previously observed to confuse \clang
+programmers also confuse JavaScript programmers: the comma operator and assignments
+being used as values. In addition, pre- and post-increment operators, when used as values,
+omitted curly braces, changes of literal encoding, and type conversions have led to confusion in at least one
+of the two experiments. For all these cases effect sizes were either medium or high. The interviews we conducted
+reinforce some of these findings while pointing to other constructs and idioms that merit investigation in the future;
+while our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems:
+four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code
+in the dataset.
+%Our findings provide significant evidence that some code patterns should be avoided if code is expected to be understood by a wide audience. 
 \end{abstract}
-
-\begin{CCSXML}
-<ccs2012>
- <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{datasets, neural networks, gaze detection, text tagging}

### 2022-07-27 3_Method.tex
First feature-complete version of the paper. It is enormous.

@@
 % \item \rqc  
-\item \rqd\castor{This one will probably be removed.}
+\item \rqd
 \end{enumerate}
@@
 By answering these research questions, we can either generalize or refute the findings about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). 
-\castor{The following pertains to the third RQ.}
-In addition, answering the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice\castor{This should probably be removed.}.
+In addition, answering the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice.
 Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions.
@@
 We conduct a mixed-methods study to answer these questions. 
-It includes two independently designed and conducted experiments: a repeated measures/within subjects study and one using a Latin square, counterbalancing experimental design to control for subject experience.
+It includes two independently designed and conducted experiments: a repeated measures/within subjects study (Section~\ref{sec:meth:replication}) and one using a Latin square, counterbalancing experimental design to control for subject experience (Section~\ref{sec:meth:survey}).
 By using two different experiments, designed by different researchers, with different designs, and conducted with different participants, we hope to rigorously verify the impact of atoms of confusion in code comprehension activities.
@@
 They investigate whether or not programs that contain atom candidates tend to produce more misunderstanding for programmers trying to predict their outcome.% We leverage two independently-developed and conducted experiments with different designs to more
-We also conduct a set of interviews with developers to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2).
+We also conduct a set of interviews with developers (Section~\ref{sec:method:interview}) to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2).
+Finally, we perform a repository mining study (Section~\ref{sec:method:msr}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
 
@@
 
-\castor{We must have introduced the atoms before.}
 The first experiment we perform uses a repeated measures, or within-subject, design~\cite{Keselman:2001:ARM}.
@@
 
-\subsection{Interview Study Setting}
+\subsection{Interview Study Setting}\label{sec:method:interview}
 
@@
 
-\subsection{Mining JavaScript Software Repositories Setting}
+\subsection{Mining JavaScript Software Repositories Setting}\label{sec:method:msr}
 

### 2022-07-27 4_Result.tex
First feature-complete version of the paper. It is enormous.

@@
   \begin{scriptsize}
-\begin{tabular}{lccc} \toprule
+\begin{tabular}{lrrr} \toprule
   Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
@@
 Array Destructuring       &           15 &      28 &                 +87 \\
-Indentation no Braces     &           30 &      50 &                 +67 \\
+Lack of Indentation no Braces     &           30 &      50 &                 +67 \\
 Omitted Curly Braces      &           35 &      51 &                 +46 \\
@@
 Implicit Predicate        &           41 &      44 &                  +7 \\
-Indentation With Braces   &           53 &      55 &                  +4 \\
+Lack of Indentation With Braces   &           53 &      55 &                  +4 \\
 Conditional Operator      &           60 &      62 &                  +3 \\
@@
 One of the atom candidates, Constant Variables, is omitted from the table because, due to the very low number of mistakes, it is not possible to perform the test.
-The table shows that eight of the 24 analyzed atom candidates can be considered atoms of confusion based on the results of our study, Object Destructuring, Type Conversion, Change of Literal Encoding, Comma Operator, Indentation No Braces, Assignment as Value, Pre-Increment, and Ommitted Curly Braces.
+The table shows that eight of the 24 analyzed atom candidates can be considered atoms of confusion based on the results of our study, Object Destructuring, Type Conversion, Change of Literal Encoding, Comma Operator, Lack of Indentation No Braces, Assignment as Value, Pre-Increment, and Ommitted Curly Braces.
 Among these eight, only Object Destructuring is JavaScript-specific.
-Six of the other atoms have been previously reported for the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and five for Java~\cite{Langhout:2021:ACJ}.
-Indentation No Braces is not specific to JavaScript, but has not been studied previously.
+Six of the other atoms have been previously reported for the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and six for Java~\cite{Langhout:2021:ACJ}.
+% Lack of Indentation No Braces is not specific to JavaScript and has also been studied by Langhout and Aniche~\cite{Langhout:2021:ACJ}. 
 For some candidates that have been previously identified as atoms, there was not a statistically significant difference.
 This is the case for Repurposed Variables, Logic as Control Flow, Implicit Predicate, and Conditional Operator.
+We examine this in Section~\ref{sec:discussion}.
 
@@
    \toprule
-Atom candidate & McNemar test &      odds ratio \\
+Atom candidate & McNemar test &      Odds Ratio \\
 \midrule
@@
 Comma Operator            &   \textbf{$<$ 0.0001*} & 11.5 \\
-Indentation No Braces     &   \textbf{$<$ 0.0001*} &  11.0 \\
+Lack of Indentation No Braces     &   \textbf{$<$ 0.0001*} &  11.0 \\
 Assignment As Value       &   \textbf{$<$ 0.0001*} &  9.3 \\
@@
 Implicit Predicate        &  0.6900 &  1.3 \\
-Indentation With Braces   &  0.7266 &  1.7 \\
+Lack of Indentation With Braces   &  0.7266 &  1.7 \\
 Conditional Operator      &  0.7905 &  1.3 \\
@@
  \centering
- \includegraphics[width=\columnwidth]{images/wrong-answers-plot-1.pdf}
+ \includegraphics[width=0.6\columnwidth]{images/wrong-answers-plot-1.pdf}
  \caption{Number of wrong answers of each subject.}
@@
 \begin{figure}
+  \centering
   \includegraphics[scale=0.3]{images/correctness-by-experience}
@@
   \begin{scriptsize}
-\begin{tabular}{lccc} \toprule
+\begin{tabular}{lrrr} \toprule
 Atom & Obfuscated  & Clean & $\Delta$(\%) \\ \midrule
@@
  {\scriptsize
- \begin{tabular}{lccc|cc}
+ \begin{tabular}{lrrr|rr}
    \toprule
-Atom  & Chi-square & Odds          & CI & Mann-Whitey& Cliff's\\ 
- &  test &Ratio          & CI &  U test & Delta \\  \midrule
+Atom  & Chi-square & Odds          & Confidence & Mann-Whitey& Cliff's\\ 
+ &  test &Ratio          & Interval &  U test & Delta \\  \midrule
 Comma Operator & \textbf{$<$ 0.0001*} & 19.02 & (6.60, 68.22) & \textbf{$<$ 0.0001*} & -0.53 \\ 
@@
       \emph{without} atom candidates. The participants were only presented the code snippets, without any indication about whether one was confusing or not. 
-      Participants were also allowed to choose Neutral when they thought both sides were equally readable.}
+      Participants were also allowed to choose Neutral when they thought both sides were equally readable.}\label{tab:interview-results1}
     \begin{tabular}{lrrr}\toprule
@@
     }
-    \label{tab:interview-results1}
+
 \end{table}
@@
 However, there was a significant difference in terms of the time required to predict the output correctly in favor of the obfuscated version.
-This highlights how factors such as the experience of the participants and the selection of code snippets can have considerable impact on the results of studies about code readability~\cite{Feitelson:2021:CPC}.
-\castor{Should the last sentence above be in the discussion?}
-\rb{Agree. We can move it to the discussion section.}
 %% \rb{pela discuss\~{a}o anterior, parece-me que alguns
@@
 \setlength\tabcolsep{2pt} % default value: 6pt
-\begin{tabular}{lrc}%c}
+\begin{tabular}{lrr}%c}
   \toprule

### 2022-07-27 discussion.tex
Put the paper in the elsarticle template. It still missing authors and affiliations. Furthermore, we still need to add the descriptions of the atoms and the discussion.
Added the two tables with the descriptions and examples for all the atoms investigated in the two experiments. They are ugly and in landscape mode, but that was the best I could do.
First feature-complete version of the paper. It is enormous.

@@
 
-% From the paper. I think it makes more sense to discuss this in the context of teh two studies. 
-Comparing the results of the latin square study to previous research, three 
-atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed
-for \clang and \cpplang do not lead to a statistically significant impact
-on correctness in our study: Logic as Control Flow,
-Conditional Operator, and Pre-Increment. Our results
-support the Gopstein et al. findings for the remaining
-atoms. The work of Langhout and Aniche~\cite{Langhout:2021:ACJ} investigates six atom candidates for Java that we also explore here for JavaScript. Among these atom
-candidates, we achieve similar conclusions to their
-study for four atom candidates: Post Increment,
-Conditional Operator, Arithmetic as Logic, and
-Omitted Curly Braces---both studies confirmed only the former
-as an atom of confusion. Our findings diverged from
-the work of Langhout and Aniche for two atoms: Pre Increment and
-Logic as Control Flow.
+The results of the multiple studies we have conducted, combined and contrasted with previous work on atoms of confusion, provide an in-depth perspective on which code patterns tend to lead to confusion and which ones may require additional investigation.
+Across different studies, targeting different languages and groups of participants, these patterns have produced confusion, even in constrained scenarios without other interfering factors.
+Existing studies judge whether a code pattern is an atom of confusion or not by evaluating whether participants incorrectly predict the behavior of programs including that pattern significantly more often than the behavior of functionally equivalent programs that do not include it.
+According to this criterion, there are five atoms of confusion that have been confirmed in the studies of Langhout and Aniche~\cite{Langhout:2021:ACJ}, Gopstein and colleagues~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, and at least one of the experiments we conducted, \textbf{Type Conversion}, \textbf{Change of Literal Encoding}, \textbf{Omitted Curly Braces}, \textbf{Pre-Increment}, and \textbf{Post-Increment}.
+They are small code patterns that have been shown to cause confusion for different samples, in programs written in different languages (Java, C, and JavaScript).
+In addition, \textbf{Comma Operator}, and \textbf{Assignment as Value} have been shown to be atoms in both of our experiments and in the study of Gopstein et al.\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
+Two code patterns have caused confusion in the participants of one of our experiments and also in one prior study, \textbf{Indentation No Braces}~\cite{Langhout:2021:ACJ} and \textbf{Implicit Predicate}~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
 
-\castor{Which atoms?
-The weight of experience.
-Statistical significance vs.
-practical relevance.
-Most of them, with the exception of Comma Operator and Object Destructuring, have also been observed for Java programs~\cite{Langhout:2021:ACJ}.
+Our interviews and the study of Langhout and Aniche~\cite{Langhout:2021:ACJ} have also analyzed the propensity of some of the atoms to induce confusion from a qualitative perspective, by asking developers about their preferences.
+In both cases, the obfuscated versions of \textbf{Omitted Curly Braces}, \textbf{Pre-Increment}, \textbf{Post-Increment}, \textbf{Implicit Predicate} were considered more confusing than their clean counterparts.
+Furthermore, the interview results are aligned with the experimental results: the clean versions were consistently considered less confusing than the obfuscated ones, except for one code pattern, Conditional Operator, which has not been confirmed as an atom by either of our experiments.
 
-In addition, many of the coding idioms that have been previously identified as atoms of confusion in previous work~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ}, such as Logic as Control Flow
+Considering the accumulated evidence for the aforementioned atoms, not only in terms of statistical significance but also effect sizes, it is reasonable to suggest that coding style guides avoid them if they aim to cater to a wide range of developer experience levels.
+These guides usually exist in contexts where multiple developers need to maintain the same codebase and they aim to promote uniformity\footnote{https://google.github.io/styleguide/} and readability\footnote{https://github.com/airbnb/javascript}.
+Although personal preference and development experience have an impact on individual developers' ability to understand these code patterns, this is arguably not the audience that coding style guides should target.
 
-Differences among the two studies, for example, post increment was not significant in the first one but it was significant in the second.
-}
+In the repeated measures study, we identified one JavaScript-specific atom, Object Destructuring.
+In the Latin square study, Automatic Semicolon Insertion was identified as another JavaScript-specific atom.
+For both cases, p-values were much lower than the threshold and effect sizes were high.
+However, in the repeated measures study obfuscated versions of Automatic Semicolon Insertion were not significantly different from their clean counterparts.
+Our interviewees favored the use of semicolon over Automatic Semicolon Insertion.
+Finally, as discussed in Section~\ref{sec:repeated:correct}, Array Destructuring exhibited a large effect size, in spite of the non-statistically significant difference (after correction).
+As discussed elsehwere~\cite{Ellis:2010:EGE}, this is a scenario that hints at practical relevance, specially considering the very low p-value ($< 0.003$).
+Further studies investigating these code patterns, potentially using different response variables~\cite{Oliveira:2020:ECR,Feitelson:2021:CPC}, are left for future work.
 
-Our work has several implications.
-First, it adds external validity to the
-work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17},
-which investigates
-the impact of atom candidates on
-understanding \clang,\cpplang code. That is,
-similarly to their work, the atom candidates
-Comma Operator, Post/Pre Increment, Omitted Curly Braces,
-Assignment as Value, Implicit Predicate, and Logic as
-Control Flow seem to make 
-JavaScript code harder to understand. For five of them, the difference is statistically significant, with a large effect size for four atoms. Our results also 
-refute the hypothesis that Arithmetic as Logic is an atom of confusion (i.e., a source of misunderstanding).
-In comparison to the original
-work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, 
-our study
-led to some differences in the effect size
-of the atom candidates.
-Altogether, we answer our first research question.
+Some code patterns that have been identified as atoms in previous work were not confirmed in our experiments, at least from a correctness perspective.
+Comparing the results of the Latin square study to previous research, two atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed for \clang and \cpplang do not lead to a statistically significant impact on correctness in our study, Logic as Control Flow, and Conditional Operator.
+Conditional Operator was the only atom candidate in our interviews whose obfuscated version was considered less confusing than the corresponding clean version.
+
+Logic as Control Flow is a more interesting case.
+Considering the correctness perspective, none of our experiments has identified Logic as Control Flow as an atom of confusion.
+However, in the Latin square study there was a significant difference in the time required to make a correct prediction in favor of the clean version of this atom candidate, with a medium effect size.
+Furthermore, both Langhout and Aniche~\cite{Langhout:2021:ACJ} and Gopstein and colleagues~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} have identified it as an atom.
+Two other atoms have exhibited a statistically significant difference in time, in favor of their obfuscated versions (with a small effect size).
+One of them, Pre-Increment, has been confirmed as an atom of confusion in multiple experiments, i.e., participants are less confused by the clean versions.
+These conflicting results reinforce the importance of conducting experiments that consider multiple response variables~\cite{Oliveira:2020:ECR}.
+In addition, it is clear that factors such as the experience of the participants and the selection of code snippets can have considerable impact on the results of studies about code readability~\cite{Feitelson:2021:CPC}.
+
+Altogether, we answer our first research question as follows.
 

### 2022-07-29 4_Result.tex
small fixes

@@
 The p-values for these tests are reported in the ``Chi-square test'' column of Table~\ref{tab:hypothesis-testing}.
-The results indicate that for five atom candidates (Comma Operator, Automatic Semicolon Insertion, Post-Increment, Assignment as Value, Implicit Predicate) the obfuscated versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$).
+For five atom candidates (Comma Operator, Automatic Semicolon Insertion, Post-Increment, Assignment as Value, Implicit Predicate), the results indicate that the obfuscated versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$).
 This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
@@
 Finally, we also use a \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impacts correctness.
-The findings suggest that \emph{participant experience} impacts the results related to correctness (Figure~\ref{fig:correctness-over-experience}).
+The findings suggest that \emph{participant experience} impacts the results related to correctness (p-value $<$ 0.001 and $\chi^2 = 24.18$).
+Figure~\ref{fig:correctness-over-experience} summarizes the number of correct and incorrect answers over the four experience groups.
 Even though the presence of atoms of confusion reduces the number of correct answers for all \emph{experience groups}, this effect is not uniform.
@@
 It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.
-We replicate the \emph{Pearson's Chi-squared test} for each experience group and find that the statistical difference is less significant for those developers with more than ten years of experience.
-Differently from \emph{participant experience}, the factor \emph{participant education} does not significantly change the correctness of the answers.
-\castor{I'd like to have the numbers for the results above. What are the p-values, regression coefficients, etc.?}
+We replicate the \emph{Pearson's Chi-squared test} for each experience group and find that the statistical difference is less significant for those developers with more than ten years of experience (p-value = 0.007 and $\chi^2$ = 7.08).
+Differently from \emph{participant experience}, the factor \emph{participant education} does not significantly change the correctness of the answers (p-value = 0.236 and $\chi^2$ = 5.54).
+
+% \castor{I'd like to have the numbers for the results above. What are the p-values, regression coefficients, etc.?}
+
 \begin{figure}
   \centering
-  \includegraphics[scale=0.3]{images/correctness-by-experience}
+  \includegraphics[scale=0.5]{images/correctness-by-experience}
   \caption{Impact of atoms of confusion on the correctness (over the four participant experience groups)}

### 2022-08-01 0_Abstract.tex
JSS submitted version

@@
 uses confusing idioms and language constructs, which might increase maintenance efforts
-and impose a significant cognitive load on developers. For this reason, researchers have
-investigated possible sources of confusion in codebases, including the impact of small code
+and impose a significant cognitive load on developers. Previous research has
+investigated possible sources of confusion in programs, including the impact of small code
 patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code.
@@
 In this paper we investigate the impact of atoms of confusion on understanding JavaScript code---a
-dynamically typed language whose popularity is growing in the most diverse application domains. To this end,
-we carry out a mixed-methods research comprising two experiments, a set of interviews with practitioners,
-and a mining software repositories (MSR) effort. 
+dynamically typed language whose popularity is growing in the most diverse application domains. In this paper
+we report the results of a mixed-methods research comprising two experiments, a set of interviews with practitioners,
+and a mining software repository (MSR) effort. 
 The results of both experiments suggest that two code patterns that have been previously observed to confuse \clang
@@
 being used as values. In addition, pre- and post-increment operators, when used as values,
-omitted curly braces, changes of literal encoding, and type conversions have led to confusion in at least one
+omitted curly braces, changes of literal encoding, and type conversions have been confirmed as confuse in at least one
 of the two experiments. For all these cases effect sizes were either medium or high. The interviews we conducted
 reinforce some of these findings while pointing to other constructs and idioms that merit investigation in the future;
-while our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems:
+while our MSR effort reveals that the small code patterns we explore here are frequent and used intensively in 72 popular open-source JavaScript systems:
 four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code
@@
 \end{abstract}
+
+\begin{keyword}
+software maintenance, program comprehension, program understanding, atoms of confusion, JavaScript code
+\end{keyword}

### 2022-08-01 main.tex
Added the two tables with the descriptions and examples for all the atoms investigated in the two experiments. They are ugly and in landscape mode, but that was the best I could do.
First feature-complete version of the paper. It is enormous.
JSS submitted version

@@
 \usepackage{balance}
+\usepackage{adjustbox}
+\usepackage{rotating}
 \usepackage{caption}
+\usepackage[normalem]{ulem}
+\useunder{\uline}{\ul}{}
 %\usepackage{cite}
@@
 \usepackage{booktabs}
-%\usepackage{longtable}
+\usepackage{longtable}
 %\usepackage{tabularx} 
@@
 
-\newcommand{\rqa}{What is the impact of atoms of confusion on the comprehension of
-  JavaScript code?}
+\newcommand{\rqa}{What is the impact of the analyzed code patterns% atoms of confusion 
+on the comprehension of JavaScript code?}
 
@@
 
-\title{On The Impact of Atoms of Confusion in JavaScriptCode}
+\title{An Investigation of Confusing Code Patterns in JavaScript}
+
+\author[1]{Adriano Torres}
+\author[1]{Caio Oliveira}
+\author[1]{M\'{a}rcio Okimoto}
+\author[2]{Diego Marc\'{i}lio} 
+\author[3]{Pedro Queiroga}
+\author[4,5]{Fernando Castor}
+\author[1]{Rodrigo Bonif\'{a}cio}
+\author[1]{Edna Dias Canedo}
+\author[5]{M\'{a}rcio Ribeiro}
+\author[6]{Eduardo Monteiro}
+
+\address[1]{Computer Science Department, University of Bras\'{i}lia, Brazil}
+\address[2]{Universit\'{a} Della Svizzera Italiana, Switzerland}
+\address[3]{Informatics Center, University Federal of Pernambuco, Brazil}
+\address[4]{Information and Computing Sciences Department, Utrecht University, The Netherlands}
+\address[5]{Institute of Computing, Federal University of Alagoas, Brazil}
+\address[6]{Statistics Department, University of Bras\'{i}lia, Brazil}
 
@@
 \input{2_Background}
+%\input{7_Atoms}
 \input{3_Method}
@@
 \input{6_Conclusion}
-% \input{7_Atoms}
 

### 2022-09-23 declaration.tex
adding letters

\documentclass{letter}
\usepackage{hyperref}
\signature{Rodrigo Bonif\'{a}cio}
\address{Computer Science Department \\ University of Bras\'{i}lia \\ Brazil}
\begin{document}

\begin{letter}{Journal of Systems and Software Editors-in-Chief}

\opening{Dear Profs. P. Avgeriou and D. Shepherd}

On behalf of the authors of the
paper ``\emph{An Investigation of Confusing Code Patterns in JavaScript}'',
I declare that we have no known competing financial interests or personal relationships
that could have appeared to influence the work reported in this paper.

\closing{Yours Faithfully,}

\end{letter}
\end{document}

### 2022-09-23 letter.tex
adding letters

\documentclass{letter}
\usepackage{hyperref}
\signature{Rodrigo Bonif\'{a}cio}
\address{Computer Science Department \\ University of Bras\'{i}lia \\ Brazil}
\begin{document}

\begin{letter}{Journal of Systems and Software Editors-in-Chief}

\opening{Dear Profs. P. Avgeriou and D. Shepherd}

I am writing to you on behalf of the authors of the
paper ``\emph{An Investigation of Confusing Code Patterns in JavaScript}''.
The paper presents the results of an original study that explores the impact of atoms of confusion in
JavaScript code. In more details, we report the results of a mixed methods research
that includes two experiments with 210 JavaScript developers, an interview
study with 15 JavaScript developers, and a software repository mining effort.
The highlights of the paper are as follows:

\begin{itemize}
 \item The experiments confirm that two code patterns previously validated as atoms of confusion for C and C++ languages
  also confuse JavaScript programmers: The Comma Operator and Assignment as Value patterns.

 \item The code patterns Pre- and Post-increment, Omitted Curly Braces, Change of Literal
   Encoding, Implicit Predicate, and Type Conversions have been confirmed as confusing in at least one of our experiments.

 \item Code patterns that we validated as confusing frequently appear in open source programs. In particular,
   we found 19.89 occurrences of the atom Implicit Predicate per KLOC.

 \item Our experiments did not confirm that the patterns Logic as Control Flow and Conditional Operator are confusing for
   JavaScript developers, contrasting with previous research works that present evidence these code patterns are atoms of confusion for C and Java.

 \item Practitioners also consider JavaScript specific idioms (such as Property Access via Array Subscription and Arrow Functions)
   confusing---they are possible atom candidates that deserve further research for validation. 
\end{itemize}

We hope our paper is within the scope of the Journal of Systems and Software and
we kindly ask you to consider it for review.

\closing{Yours Faithfully,}

\end{letter}
\end{document}

### 2022-10-04 3_Method.tex
Making some modifications to address the reviwers comments, in particular pertaining to the reasons for having multiple studies.

@@
 We conduct a mixed-methods study to answer these questions. 
-It includes two independently designed and conducted experiments: a repeated measures/within subjects study (Section~\ref{sec:meth:replication}) and one using a Latin square, counterbalancing experimental design to control for subject experience (Section~\ref{sec:meth:survey}).
+It includes two independently designed and conducted experiments: a repeated measures/within subjects study (Section~\ref{sec:meth:replication}) and one using a Latin square, between subjects, counterbalancing experimental design to control for subject experience (Section~\ref{sec:meth:survey}).
 By using two different experiments, designed by different researchers, with different designs, and conducted with different participants, we hope to rigorously verify the impact of atoms of confusion in code comprehension activities.
+\revised{Conducting multiple studies, with different designs and samples, is a typical approach in other areas of science to obtain more solid evidence and also to contest evidence obtained by single study, for example, Weggemans et al.~\cite{Weggemans:2001:DCE}.
+We discuss this matter further in Section~\ref{sec:whytwo}}.
+
 These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ1).

### 2022-10-04 discussion.tex
Making some modifications to address the reviwers comments, in particular pertaining to the reasons for having multiple studies.

@@
 \label{sec:discussion}
+
+In this section we discuss the results of the studies in combination.
+
+%----------------------------------------------------------------------
+\subsection{\revised{Contrasting the results of the studies}}\label{sec:contrast}
 
@@
 In addition, it is clear that factors such as the experience of the participants and the selection of code snippets can have considerable impact on the results of studies about code readability~\cite{Feitelson:2021:CPC}.
+
+%----------------------------------------------------------------------
+\subsection{\revised{Addressing the research questions}}
 
@@
 
+%----------------------------------------------------------------------
+\subsection{\revised{Why multiple studies?}}\label{sec:whytwo}
+
+\revised{It is not reasonable to consider something a scientific fact in an area as inexact as Software Engineering based on a single experiment.
+We present two independently-designed and conducted studies, with diverse code snippets, samples, and experimental designs, to provide stronger evidence.
+We believe that atom candidates confirmed to be atoms of confusion by both experiments arguably provide more reliable evidence of the potential to cause confusion of these code patterns for a wide range of developers, in diverse scenarios.
+That is the case for the Comma Operator and Assignment as Value atoms.
+Considering our work and others', there is little doubt that these are code patterns that lead to confusion.
+Furthermore, we complement the two experiments with interviews with 15 developers.
+Generally, in more mature areas of science, conducting multiple independent studies, potentially with different experimental designs, e.g., controlled vs.
+free-living experiments in the medical sciences~\cite{Weggemans:2001:DCE}, and examining their results in combination is the most reliable way of establishing truth.}
+
+\castor{For the pros and cons of each approach, do you prefer lists of items or regular text paragraphs?}
+\revised{The experimental designs we have employed in the two experiments have different trade-offs. The main advantages of having a repeated measures (or within subjects) study are the following: 
+\begin{itemize}
+  \item[+] They are better able to detect an effect size, given that one exists, since it is possible to collect data about both treatments from each participant. 
+  \item[+] There tends to be less variance between groups, since the same participant is subject to both treatments
+  \item[+] Individual differences between participants are controlled for because each participant acts as their own control
+\end{itemize}
+}
+\revised{
+\noindent
+Conversely, the main limitations of a repeated measures study are the following: 
+\begin{itemize}
+\item[--] Data is not independent. This precludes the use of some statistical techniques and tests. 
+\item [--] There is a potential learning/practice effect, since going through multiple examples may make the participants able to predict results of subsequent cases.
+This can be mitigated by separating examples pertaining to the same atom candidate, but not eliminated.
+\item [--] Fatigue effect, since participants have to examine more code snippets. This was exacerbated in our case, due to the larger number of atom candidates.
+\end{itemize}}
+\revised{
+The Latin square study is an example of a between subjects study.
+In a way, its main advantages are precisely the absence (or a reduction) of the disadvantages of the repeated measures study and its main disadvantages are related to the advantages of a repeated measures study:
+\begin{itemize}
+\item[+] Data is independent. 
+\item[+] There is no possibility of learning/practice effect because each participant is only exposed to one treatment. 
+\item[+] Fatigue effect is reduced, since each participant is only exposed to code snippets related to one of the treatments. 
+\end{itemize}}
+\revised{
+\noindent
+The main disadvantages of a between subjects study are the following:
+\begin{itemize}
+  \item[-] Effect sizes tend to be lower. 
+  \item[-] Potentially more variance between groups since each participant is only required to be exposed to one treatment. 
+  \item[-] Does not account for variance within groups.
+\end{itemize}
+The use of a Latin square mitigates some of the disadvantages intrinsic to a between subjects design.
+For example, since less work is required from participants, it was possible to obtain more data for the latin square study than we could get for the repeated measures study.
+This increases its statistical power, making it possible to detect smaller effect sizes.}
+
+\revised{
+In addition, to capture aspects that none of these studies can capture, e.g., personal impressions, we also leverage the interview study.
+Section~\ref{sec:contrast} highlights that some atoms that are perceived as confusing did not cause participants of the experiments to commit mistakes, e.g., Arithmetic as Logic.
+This combination of three different studies, with different samples and designs provides a more in-depth perspective on the extent to which these code patterns confuse JavaScript developers than we could obtain by having a single study.
+}
+
 %% We conducted a non-exact replication of the three

### 2022-10-31 3_Method.tex
Fernando tackled the reviewer comments assigned to him by Rodrigo.

@@
 That excludes atoms related to pointers and preprocessor directives.
-Furthermore, we conduct an informal search of programming forums, solutions to Code Golf~\footnote{https://codegolf.stackexchange.com/} problems, code style guidelines, and expert knowledge to identify new atom candidates.
-Overall, the repeated measures study analyzes 24 atoms, 15 from the original study and 9 that are new.
+Furthermore, we conduct an informal search of \revised{JavaScript questions on StackOverflow about constructs that are not typical of other languages, such as array and object spread, solutions to Code Golf~\footnote{https://codegolf.stackexchange.com/} problems, and code style guides\footnote{AirBNB JavaScript Style Guide: https://github.com/airbnb/javascript}\footnote{Google JavaScript Style Guide: https://google.github.io/styleguide/jsguide.html} to identify new atom candidates.
+We leverage previous work~\cite{castor2018} on the identification of atom candidates to guide us in this informal search.} Overall, the repeated measures study analyzes 24 atoms, 15 from the original study and 9 that are new.
 Some of these atoms are also analyzed in the Latin square study (Section~\ref{sec:meth:survey}).

### 2022-12-02 abstract.tex
changing the paper organization

@@
 \begin{abstract}
-  Evolving legacy code is a challenging task, particularly when the code has been poorly written or uses confuse idioms and language constructs, which might increase maintenance efforts and impose a significant cognitive load on developers.
-For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in statically typed languages such as \clang, \cpplang, and Java.
-In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language.
-We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and an activity of mining open source JavaScript repositories (MSR).
-Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript.
-Considering ten atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates.
-For five of these atoms, the difference is statistically significant (accounting for p-value correction).
-Effect size is large for four atoms and medium for one of them.
-In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems.
-Four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code in the dataset.
-Altogether, our findings might help practitioners: (1) better understand the implications of atoms of confusion on understanding JavaScript code, (2) avoid writing code that is unnecessarily difficult to maintain, and (3) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
+Evolving software is particularly challenging when the code has been poorly written or
+uses confusing idioms and language constructs, which might increase maintenance efforts
+and impose a significant cognitive load on developers. Previous research has
+investigated possible sources of confusion in programs, including the impact of small code
+patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code.
+Although researchers have explored atoms of confusion in code written in \clang, \cpplang,
+and Java, different languages have different features, developer communities, and development cultures. This
+justifies the exploration of other languages to verify whether they also exhibit confusion-inducing patterns. 
+In this paper we investigate the impact of atoms of confusion on understanding JavaScript code---a
+dynamically typed language whose popularity is growing in the most diverse application domains. In this paper
+we report the results of a mixed-methods research comprising two experiments, a set of interviews with practitioners,
+and a mining software repository (MSR) effort. 
+The results of both experiments suggest that two code patterns that have been previously observed to confuse \clang
+programmers also confuse JavaScript programmers: the comma operator and assignments
+being used as values.
+In addition, some code patterns, such as omitted curly braces and change of literal encoding, have caused confusion in participants in one of the experiments.
+We discover that some JavaScript-specific elements, such as automatic semicolon insertion and object destructuring, also have the potential to cause confusion.
+%In addition, at least one experiment has confirmed that some atoms detected for other languages, such as  and type conversion, also 
+%In addition, pre- and post-increment operators, when used as values, omitted curly braces, changes of literal encoding, and type conversions have been confirmed as confuse in at least one
+%of the two experiments. 
+For all these cases effect sizes were either medium or high. The interviews we conducted
+%reinforce some of these findings while pointing to
+indicate other constructs and idioms that merit investigation in the future.
+Finally, our MSR effort reveals that the examined code patterns are frequent and used intensively in 72 popular open-source JavaScript projects:
+four atom candidates appear in 90\% of them and two of them occur more than once for every 100 lines of code.
+%Our findings provide significant evidence that some code patterns should be avoided if code is expected to be understood by a wide audience. 
 \end{abstract}
@@
 \begin{keyword}
-  code readability, program comprehension, program understanding, atoms of confusion, JavaScript code
+software maintenance, program comprehension, program understanding, atoms of confusion, JavaScript code
 \end{keyword}

### 2022-12-02 conclusion.tex
changing the paper organization

--- 2022-03-29 6_Conclusion.tex
@@
 This paper reports the results of a mixed-method research effort that allowed us to better understand the impact of atoms of confusion in JavaScript code.
-First, we conducted a survey with 140 JavaScript developers, asking them to predict the output of some code snippets (with and without atoms).
-We provide evidence that five atom candidates significantly impact program comprehension activities and can be considered atoms of confusion.
+First, we conducted two independent experiments, which involve 210 JavaScript developers in total, asking them to predict the output of some code snippets (with and without atoms candidates).
+In at least one of the experiments, five atom candidates (Type Conversion, Change of Literal Encoding, Omitted Curly Braces, Pre-Increment, and Post- Increment) significantly impact program comprehension activities and can be considered atoms of confusion.
 %% : Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate. 
@@
 Our efforts have two implications for practice.
-First, we present evidence that not all atoms of confusion validated to statically typed languages (such as \clang, \cpplang, or Java) led to a statistically significant impact on program understanding for JavaScript (a dynamically typed language).
+First, we present evidence that not all atoms of confusion validated to \clang, \cpplang, or Java led to a statistically significant impact on program understanding for JavaScript.
 Besides that, our findings might be used to alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment/Decrement, Assignment as Value, and Implicit Predicate).
@@
 reproduce our survey to validate whether or not these additional
-\emph{atom candidates}
+idioms 
 %, which the interviewees pointed out, 

### 2022-12-02 discussion.tex
changing the paper organization

@@
 \item[+] Data is independent. 
-\item[+] There is no possibility of learning/practice effect because each participant is only exposed to one treatment. 
-\item[+] Fatigue effect is reduced, since each participant is only exposed to code snippets related to one of the treatments. 
+\item[+] There is no possibility of learning/practice effect and the fatigue effect is reduced since each participant is only exposed to one version of each code snippet.
+There is no possibility of learning/practice effect because each participant is only exposed to one treatment.
 \end{itemize}}
@@
   \item[-] Effect sizes tend to be lower. 
-  \item[-] Potentially more variance between groups since each participant is only required to be exposed to one treatment. 
   \item[-] Does not account for variance within groups.

### 2022-12-02 methodology.tex
changing the paper organization

--- 2022-10-04 3_Method.tex
@@
-\section{Methodology}
+\section{Research Methods}
 \label{method}
@@
 Finally, we perform a repository mining study (Section~\ref{sec:method:msr}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
-
-\subsection{Repeated Measures Study}\label{sec:meth:replication}
-
-The first experiment we perform uses a repeated measures, or within-subject, design~\cite{Keselman:2001:ARM}.
-This experiment is inspired by the work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, but targeting JavaScript instead of the C language, and including some methodological differences.
-It covers all the atom candidates analyzed in the work of Gopstein et al.
-that can be transposed to JavaScript and adds candidates that are specific to this language.
-%Overall, we select 24 atom candidates in the repeated measures study.
-
-\subsubsection*{Experimental design}
-
-In this experiment, the control variable consists of a tiny program containing a single atom candidate and the treatment is a functionally-equivalent version that does not contain the atom.
-The main dependent variable is whether subjects are able to correctly determine the output of the programs.
-The other dependent variable is the time required to correctly vs.
-incorrectly determine the outputs, independently of the presence of atoms.
-We use the same null hypothesis as Gopstein et al.: \textit{``code from both control and treatment groups can be hand-evaluated with equal accuracy''} and the alternate hypothesis that \textit{``the existence of an atom of confusion causes more errors than other code in hand-evaluated outputs.''}.
-For each atom candidate, we built six programs, three versions including the atom candidate and three versions not including it.
-%Each subject is exposed to four programs for every atom candidate, two including an atom and two not including an atom.
-We have also built six programs for each atom candidate, three with and three without the atom.
-Considering the elements that can vary in a replication study, defined by Juristo and Gomez~\cite{Juristo2012}, in this study we employ the same experimental protocol and construct operationalizations.
-We have different population properties, experimenters, and a different site.
-In addition, since we focus on a different language, some atoms that exist in C cannot by construction exist in JavaScript programs.
-Conversely, JavaScript exhibits atom candidates that do not exist in C.
-
-% The design of the repeated measures study considers two treatments: the presence or absence of atom candidates within the programs.
-To identify atom candidates, we start out by selecting every atom candidate from Gopstein's study that can also happen in JavaScript.
-That excludes atoms related to pointers and preprocessor directives.
-Furthermore, we conduct an informal search of programming forums, solutions to Code Golf~\footnote{https://codegolf.stackexchange.com/} problems, code style guidelines, and expert knowledge to identify new atom candidates.
-Overall, the repeated measures study analyzes 24 atoms, 15 from the original study and 9 that are new.
-Some of these atoms are also analyzed in the Latin square study (Section~\ref{sec:meth:survey}).
-Table~\ref{tab:atomsBoth} presents the atom candidates that both studies investigate and Table~\ref{tab:atomsRepeated} presents the ones that only the repeated measures study analyzes.
-
-\input{atom_descriptions}
-
-For each atom candidate, we write six short programs, three including and three not including the atom candidate.
-The former are the control for the study, or the \textbf{obfuscated} versions of the programs.
-The alternative, atom-free, versions of the programs are the treatment, or \textbf{clean}, versions.
-Each subject is exposed to one obfuscated and one clean version of each atom, totalling 48 programs, and should determine their outputs.
-As mentioned before, we measure answer correctness and the time to answer questions correctly vs.
-incorrectly.
-We control for learning effects~\cite{Neely:1991:SPE} in three manners.
-First, by having multiple obfuscated and clean versions for each atom and presenting only one of each per subject.
-Second, by presenting the programs in a random order.
-Third, by presenting obfuscated and clean versions corresponding to the same atom candidate with at least 11 other programs in between, in  accordance to the original experimental protocol~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
-
-\subsubsection*{Study instrument}
-
-The experiment is conducted by means of a questionnaire made available as a web application.
-As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the questionnaire would take on average.
-Based on the feedback of the participants of the pilot, we present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.% Initially, our plan was to present two of each.
-
-We organize the study in three sections.
-In the first one we present instructions and also a check button whose checking means users agree that all collected data will be used anonymously and solely for research purposes.
-The instructions explain how the study works and asks them to dedicate their attention to it.
-We stress to participants the importance of not using any aids during the experiment, such as online or console interpreters.
-
-The second section presents the programs, one at a time.
-For each one, there is a text box where the answer should be written and a ``Next'' button.
-If the subject leaves the text box empty and presses ``Next'', this is treated as a wrong answer.
-To reduce the likelihood of a subject attempting to execute the programs, the web application verifies whether the subject tries to change tabs or windows and presents a pop-up message if that is the case.
-Furthermore, it presents the programs as images, instead of text, to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter.
-Upon submitting an answer for a particular program, the subject is automatically led to a similar page, containing the next program.
-
-We do not provide feedback about the time subjects take to answer each question.
-We also do not tell them whether their answers are correct or not.
-This aims at avoiding introducing bias for future respondents.
-Since we posted information about the study on social media platforms, possible threats could have arisen if we gave respondents instant feedback.
-
-\subsubsection*{Study audience}
-
-We posted the link to the questionnaire on the authors' social media platforms, on Brazilian CS departments' mailing lists, and on two programming subreddits.
-We explained our research purposes and asked developers to take part in the study only if they have some familiarity with JavaScript.
-We collected 70 responses.
-%The average survey respondent has a mean 26 months of experience in JavaScript, with a median of 12.\castor{We should not say this here unless we are able to do the same for the Latin square study.
-}
-
-\subsection{Latin Square Study}\label{sec:meth:survey}
-
-This experiment uses a different experimental design that makes it possible to control for the experience of the subjects.
-It employs a subset of the atom candidates used in the repeated measures study.
-More specifically, this study investigates ten of the atoms investigated in the repeated measures study:
-%Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first 
-%select a set of 9+1 atom candidates: 
-nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion).
-%Our supplementary material discusses and shows examples of the atoms we consider in our research.\castor{We should probably do this here or just refer to the table that I intend to insert for the replication study.}
-
-\subsubsection*{Experimental design}
-
-The design of the Latin square study blocks two variables (subject experience and the programs) and considers two treatments: the presence or absence of atom candidates within the programs.
-To achieve such a design goal, controlling the effect of experience and individual programs, we resort to the \textit{Latin Square Design}~\cite{Hunter-Experimenters}.
-Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of programs.
-The design of each square (a replica) is such that no treatment is repeated in the same row or column.
-For example, considering that we have a set of 10
-code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about clean programs, they will only be presented with clean versions of the code samples $s_6, s_7,..., s_{10}$.
-Furthermore, a given subject (P2), who constitutes the second row of our example square, will be asked questions about the clean versions for $s_1, s_2, ..., s_5$, and will answer questions about obfuscated programs for $s_6, s_7,..., s_{10}$.
-By doing that, we guarantee that all versions of the programs are contained within each square, and that each configuration occurs only once within a square.
-Figure \ref{fig:latinsquare} offers a visual representation of the concept.
-
-  \begin{figure}[htb!]
-      \noindent
-      \centering
-      \includegraphics[scale=.50]{images/latin-square.pdf}
-      \caption{Latin square design. Each ``square'' corresponds to 
-      a replica in our study. Each replica comprises two participants (square rows, e.g., P1 and P2) 
-      and two sets of programs (CS-1 and CS-2). We randomly apply the 
-      treatments (atom or non-atom code) to the cells of the squares.} 
-      \label{fig:latinsquare}
-  \end{figure}
-
-For each of the 10 selected atom candidates, we write one short program containing the atom.
-As in the repeated measures study, we call them the obfuscated versions of the programs.
-We also write corresponding short programs without the atom candidates and call them the clean versions of the programs.
-Overall, the experiment employed 20 programs, 10 obfuscated and 10 clean.
-In order to reduce the cognitive effort, each subject was asked to predict the output of 10 programs, five obfuscated and five clean.
-The order in which they are presented is randomized.
-By doing this, we seek to minimize the chances of subjects being aware that the current listing they are analyzing contains (or not) atoms of confusion.
-That is, each subject should indicate what would be the outcomes of the programs, some of them having atoms of confusion (while other programs do not).
-Since participants are not exposed to obfuscated and clean versions corresponding to the same atom candidate, learning effect is not possible.
-We measure answer correctness and the total time each participant requires to participate in the experiment.
-
-\subsubsection*{Study instrument}
-
-We implemented our experiment as a questionnaire in a web application and 
-%carried out a pilot. 
-%As part of this effort, we 
-carried out an informal pilot.% whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
-Undergraduate students and professional colleagues took part in the pilot.
-Some users reported layout defects, and many reported that the landing page did not explain the study well enough.
-We also spotted minor issues with our routines to create and populate the latin squares.
-
-We organize the questionnaire in three sections.
-The first section aims to characterize the subjects, asking their age, education level, and programming experience.
-We also include a check button, whose checking means users agree that all collected data will be used solely for research purposes.
-In the second section, we present to the participants a small set of instructions, where we explain how the study works and ask them to dedicate their attention to it.
-We stress to participants the importance of not using any aids during the experiment, such as online or console interpreters.
-For each question page, we kept track of whether or not the subjects switch windows.
-
-The last section of the questionnaire presents a sequence of ten questions, each containing a program.
-For each question, there is a text box where the answer should be written.
-There is also an ``I do not know'' button, which, when clicked, leads the subject to the next question.
-In our setting, ``I do not know'' is treated as a wrong answer.
-The programs are also presented as images copied from a text editor.
-Upon submitting their answer for a particular question, the subjects are automatically led to a similar page, containing another program.
-Similarly to the repeated measures study, we do not provide feedback about time and correctness to the subjects.
-
-% \rb{nao acho esse paragrafo necessario} 
-% \adriano{o de baixo ne? eu tambem acho}
-
-% As we mentioned before, we first wrote the code listings in a text editor, and took pictures of it.
-In the case of an atom of confusion that was exclusive to JavaScript, which we called \textit{Automatic Semicolon Insertion} (see Appendix~\ref{sec:appendix-atoms}), it was necessary to remove the syntax highlighter.
-Even though semicolons at the end of statements are optional to programmers in JavaScript, the interpreter automatically inserts them into the code.
-Our text editor was incorrectly highlighting a line break after a return statement, even thought it was valid JavaScript syntax.
-We had thus to turn the highlighter off to take the picture of this atom.
-
-\subsubsection*{Study audience}
-
-%% \diego{Esses detalhes são possíveis candidatos para remoção, se precisarmos de espaço}
-%% {\color{blue}We invited developers to answer our survey by sending invitations to communities of JavaScript programmers on the Internet.
-Our initial plan was to try to engage contributors for two major JavaScript projects, namely Node.js and NPM.
-None of the projects, though, offered a direct way to interact with the community, such as an online discussion forums or mailing list.
-We would be required to first make contact with the projects' leaders, and only then would we have a chance to approach potential respondents.
-Due to time constraints, we decided to look for respondents elsewhere.}
-
-We posted the invitation to take part in the study on a JavaScript Reddit channel.\footnote{https://www.reddit.com/r/javascript/} We explained our research purposes and asked developers of any level of expertise to take part.
-Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares.
-We collected significant data on time taken and discrepancies in answer correctness between obfuscated and clean versions of the programs.
-Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the questionnaire.
-We discarded from our study all squares that contained incomplete rows (a common approach for data imputation).
-%Since we had a large enough number of samples, the squares we had to discard did not impact our analysis.
-
-\subsection{Interview Study Setting}\label{sec:method:interview}
-
-To complement the two experiments, we perform semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding programs containing atoms of confusion (research question RQ2).
-We also ask each participant if they know of any other JavaScript-specific construct or idiom they think is likely to make the code hard to understand.
-% (research question RQ3).
-In this section we detail the protocol we followed to conduct the interviews and to analyze the results.
-
-% Nós realizamos entrevistas semi-estruturas com o objetivo de identificar a percepção dos desenvolvedores com experiência em JavaScript sobre algumas questões relacionadas à compreensão de código em JavaScript.
-Assim, nesta Seção nós descrevemos os procedimentos adotados para selecionar os participantes para as entrevistas e detalhamos como as entrevistas foram conduzidas.
-Além disso, detalhamos como os resultados das entrevistas foram analisados.
-
-\subsubsection*{Participant Selection} We invite the participants of the interviews using a snowballing technique.
-That is, starting from our network of contacts, we invite an initial set of candidates to take part in our experiments.
-From this initial list, we ask for an indication of additional candidates.
-Our main selection criterion is that all participants should have been working with JavaScript in their daily professional activities.
-We invited a total of 17 developers, and 15 of them agreed to participate.
-%We conducted the interviews for a period of two weeks.
-
-\subsubsection*{Interview Process} We conduct semi-structured interviews using web conferencing software.
-We record all the interviews with the consent of the participants.
-On average, the interviews last 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes.
-Two of the researchers conduct the interviews, and a third one listens to all the recordings to cross-validate the collected data.
-The interviews have three main parts.
-In the first one, we ask the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years, and other programming languages they have worked with.
-Table \ref{pinterview} summarizes this demographic information.
-
-\begin{table}[htb!]
-
-  \centering
-  \caption{Demographic information of the participants}
-\begin{scriptsize}  
-\begin{tabular}{clcl}
-\toprule
-ID & Education & JS Experience & Other Languages \\ \midrule 
-P1 & BSc Degree & 9 years & Java, PHP, C, Go  
-\\ 
-P2 & HS Degree & 3 years & Python, Go, Dart, Lua, C++, C\#
-\\ 
-P3 & BSc Degree & 4 years & Java, C
-\\ 
-P4 & Undergraduate & 3 years & Python, C, C++, Java, Go
-\\ 
-P5 & Master Student & 3 years & Python, SQL
-\\ 
-P6 & Bsc Degree & 15  years  & Java, PHP, C, Python , Ruby, C\#
-\\ 
-P7 & BSc Degree  & 6 years & C, C++, Java, Assembly, Kotlin
-\\ 
-P8 & PhD Degree & 2 years & Java, Python
-\\ 
-P9 & BSc Degree  & 5 years & PHP
-\\ 
-P10 & Master Student & 4 years & Java, C\# and Python
-\\ 
-P11 & Master Student & 4 years & Java, Erlang, C\#, Cobol
-\\ 
-P12 & Master Degree & 1 year & C
-\\ 
-P13 & Master Degree & 13 years & Java, PHP
-\\ 
-P14 & Master Student & 3 years & C, Python, Ruby
-\\ 
-P15 &  BSc degree  & 2 years & Java, PHP
-\\ \bottomrule
-\end{tabular}
-\end{scriptsize}
-    \label{pinterview}
-\end{table}
-
-In the second part of the interview our aim is to allow the subjects to describe their JavaScript experience, as well as to allow them to reveal any JavaScript constructs they regard as innately confusing.
-This allows us to identify potential atom candidates that are more specific to the JavaScript language.
-Examples of questions we explore in this section include: \emph{Does JavaScript favor developers to produce code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as especially confusing?}
-
-In the third part of the interview, participants are shown pairs of programs that were used in the experiments, where each pair contains a clean and an obfuscated version of the same atom candidate.
-For this part we only use atoms that appear in both experiments.
-The participants
-%were told both sides of each pair had the same output, and 
-are asked to evaluate which version of the code is easier to understand.
-To avoid introducing bias in the answers, the interviewers do not explain that one of the versions in each pair contains the atom under investigation.
-Subjects are just presented the pairs of programs and allowed to take the necessary time to decide on the most readable one.
-
-\subsubsection*{Interview Analysis}
-
-We first transcribe each interview and then examine the broad distribution of the answers.
-Our goal is to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular.
-We follow-up with an open-coding procedure, highlighting the main themes and quoting the answers of the participants.
-We present these results in Section~\ref{sec:interview-results}.
-%Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.
-
-%\subsection{Third Study: Additional Survey and Interviews}
-
-% \rb{acho que aqui podemos incluir o segundo survey e a segunda
-%  rodada das entrevistas, explorando os \'{a}tomos adicionais}
-
-\subsection{Mining JavaScript Software Repositories Setting}\label{sec:method:msr}
-
-To understand how often the analyzed atom candidates appear in open source projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mine a set of GitHub open source repositories.
-To this end, we first collect the most popular GitHub repositories that are primarily written in JavaScript.
-We measure popularity using the project's stargazers.
-This metric, available through the GitHub API, represents the number of stars a project received from users of the platform.
-The same metric has been used in a number of previous studies as a proxy to estimate a project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}.
-We then select the top 100 most popular repositories and remove projects that do not reach the first quartile of the distribution of lines of code.
-
-After filtering out JavaScript project candidates, in the second step we build a curated dataset comprising the top \minedprojects repositories.
-Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}.
-Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research.
-The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
-All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars.
-We automate all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.
-
-We mine the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}.
-CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript).
-We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study).
-Finally, we compute some descriptive statistics to measure the prevalence of atoms of confusion in practice.
-
-\begin{table}[ht]
-  \centering
-   \caption{Some descriptive statistics about the projects used in the MSR study}
-  \begin{scriptsize}
- \begin{tabular}{rrrrr}
-   \toprule
-                       & Min.             & Median        & Mean             & Max. \\ \midrule
- Lines of Code         & \num{5543}       & \num{36161.5} & \num{111432.33}  & \num{1278405} \\
- Num. of Forks         & \num{1244}       & \num{6078}    & \num{8906.24}    & \num{68849} \\
- Num. of Contributors  & \num{6}          & \num{285.5}   & \num{533.44}     & \num{4047} \\
- Num. of Stars         & \num{23672}      & \num{34990}   & \num{46919.51}   & \num{310935} \\
-
-    \bottomrule
- \end{tabular}
- \end{scriptsize}
- \label{tab:projects-statistics} 
- \end{table}

### 2022-12-02 s01.tex
changing the paper organization

--- 2022-07-27 4_Result.tex
@@
-%% \section{Results and Discussion}
-%% \label{sec:results}
+\section{Repeated Measures Study}
 
-%% In this section we present the main findings of our research.
-We first present the results of each study (Section~\ref{sec:survey-resuts}, Sections~\ref{sec:msr-results}, and Section~\ref{sec:interview-results}).
-After that, in Section~\ref{sec:discussion} we consolidate our findings and present some implications of our study.
+The first experiment we perform uses a repeated measures, or within-subject, design~\cite{Keselman:2001:ARM}.
+This experiment is inspired by the work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, but targeting JavaScript instead of the C language, and including some methodological differences.
+It covers all the atom candidates analyzed in the work of Gopstein et al.
+that can be transposed to JavaScript and adds candidates that are specific to this language.
+%Overall, we select 24 atom candidates in the repeated measures study.
 
-%\section{Results of the Survey}
-%\label{sec:survey-resuts}
+\subsection{Experimental design}
 
-%-----------------------------------------------------------------------
-\section{Results of the Repeated Measures Study}\label{sec:repeated}
+In this experiment, the control variable consists of a tiny program containing a single atom candidate and the treatment is a functionally-equivalent version that does not contain the atom.
+The main dependent variable is whether subjects are able to correctly determine the output of the programs.
+The other dependent variable is the time required to correctly vs.
+incorrectly determine the outputs, independently of the presence of atoms.
+We use the same null hypothesis as Gopstein et al.: \textit{``code from both control and treatment groups can be hand-evaluated with equal accuracy''} and the alternate hypothesis that \textit{``the existence of an atom of confusion causes more errors than other code in hand-evaluated outputs.''}.
+For each atom candidate, we built six programs, three versions including the atom candidate and three versions not including it.
+%Each subject is exposed to four programs for every atom candidate, two including an atom and two not including an atom.
+We have also built six programs for each atom candidate, three with and three without the atom.
+Considering the elements that can vary in a replication study, defined by Juristo and Gomez~\cite{Juristo2012}, in this study we employ the same experimental protocol and construct operationalizations.
+We have different population properties, experimenters, and a different site.
+In addition, since we focus on a different language, some atoms that exist in C cannot by construction exist in JavaScript programs.
+Conversely, JavaScript exhibits atom candidates that do not exist in C.
+
+% The design of the repeated measures study considers two treatments: the presence or absence of atom candidates within the programs.
+To identify atom candidates, we start out by selecting every atom candidate from Gopstein's study that can also happen in JavaScript.
+That excludes atoms related to pointers and preprocessor directives.
+Furthermore, we conduct an informal search of \revised{JavaScript questions on StackOverflow about constructs that are not typical of other languages, such as array and object spread, solutions to Code Golf~\footnote{https://codegolf.stackexchange.com/} problems, and code style guides\footnote{AirBNB JavaScript Style Guide: https://github.com/airbnb/javascript}\footnote{Google JavaScript Style Guide: https://google.github.io/styleguide/jsguide.html} to identify new atom candidates.
+We leverage previous work~\cite{castor2018} on the identification of atom candidates to guide us in this informal search.} Overall, the repeated measures study analyzes 24 atoms, 15 from the original study and 9 that are new.
+Some of these atoms are also analyzed in the Latin square study (Section~\ref{sec:meth:survey}).
+Table~\ref{tab:atomsBoth} presents the atom candidates that both studies investigate and Table~\ref{tab:atomsRepeated} presents the ones that only the repeated measures study analyzes.
+
+\input{atom_descriptions}
+
+For each atom candidate, we write six short programs, three including and three not including the atom candidate.
+The former are the control for the study, or the \textbf{obfuscated} versions of the programs.
+The alternative, atom-free, versions of the programs are the treatment, or \textbf{clean}, versions.
+Each subject is exposed to one obfuscated and one clean version of each atom, totalling 48 programs, and should determine their outputs.
+As mentioned before, we measure answer correctness and the time to answer questions correctly vs.
+incorrectly.
+We control for learning effects~\cite{Neely:1991:SPE} in three manners.
+First, by having multiple obfuscated and clean versions for each atom and presenting only one of each per subject.
+Second, by presenting the programs in a random order.
+Third, by presenting obfuscated and clean versions corresponding to the same atom candidate with at least 11 other programs in between, in  accordance to the original experimental protocol~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
+
+{\bf Study instrument.} The experiment is conducted by means of a questionnaire made available as a web application.
+As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the questionnaire would take on average.
+Based on the feedback of the participants of the pilot, we present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.% Initially, our plan was to present two of each.
+
+We organize the study in three sections.
+In the first one we present instructions and also a check button whose checking means users agree that all collected data will be used anonymously and solely for research purposes.
+The instructions explain how the study works and asks them to dedicate their attention to it.
+We stress to participants the importance of not using any aids during the experiment, such as online or console interpreters.
+
+The second section presents the programs, one at a time.
+For each one, there is a text box where the answer should be written and a ``Next'' button.
+If the subject leaves the text box empty and presses ``Next'', this is treated as a wrong answer.
+To reduce the likelihood of a subject attempting to execute the programs, the web application verifies whether the subject tries to change tabs or windows and presents a pop-up message if that is the case.
+Furthermore, it presents the programs as images, instead of text, to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter.
+Upon submitting an answer for a particular program, the subject is automatically led to a similar page, containing the next program.
+
+We do not provide feedback about the time subjects take to answer each question.
+We also do not tell them whether their answers are correct or not.
+This aims at avoiding introducing bias for future respondents.
+Since we posted information about the study on social media platforms, possible threats could have arisen if we gave respondents instant feedback.
+
+{\bf Study audience.} We posted the link to the questionnaire on the authors' social media platforms, on Brazilian CS departments' mailing lists, and on two programming subreddits.
+We explained our research purposes and asked developers to take part in the study only if they have some familiarity with JavaScript.
+We collected 70 responses.
+%The average survey respondent has a mean 26 months of experience in JavaScript, with a median of 12.\castor{We should not say this here unless we are able to do the same for the Latin square study.
+}
+
+\subsection{Results}\label{sec:repeated}
 
@@
 %-----------------------------------------------------------------------
-\subsection{Correctness Analysis}\label{sec:repeated:correct}
+\subsubsection{Correctness Analysis}\label{sec:repeated:correct}
 
-\subsubsection*{Exploratory Data Analysis}
+{\bf Exploratory Data Analysis.}
 For each of the 24 atom candidates, the participants predicted the outputs of two versions, one clean (control) and one obfuscated (treatment), as per the definition of a repeated measures design~\cite{Hunter-Experimenters}.
@@
 
-\subsubsection*{Statistical analysis}
-
+{\bf Statistical analysis.}
 Since this study uses a within-subject design, the data is paired and dependent.
@@
 %-----------------------------------------------------------------------
-\subsection{Time Analysis}\label{sec:repeated:time}
+\subsubsection{Time Analysis}\label{sec:repeated:time}
 
-\subsubsection*{Exploratory Data Analysis}
-
+{\bf Exploratory Data Analysis.}
 In this section we investigate whether there is a trade-off between correctness and time required to predict program outputs, similarly to the analysis performed by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
@@
 
-\subsubsection*{Statistical Analysis}
-
+{\bf Statistical Analysis.}
 We employ the non-parametric Mann-Whitney U test, also known as Wilcoxon's rank-sum test, to verify the aforementioned null hypothesis, i.e., whether the times for correct and incorrect answers can be considered identical.
@@
 This produces a result of \textbf{0.1027}, which suggests that, although participants tend to provide correct predictions more quickly, they do that only slightly so.
-
-%-----------------------------------------------------------------------
-\section{Results of the Latin-Square Study}\label{sec:latin}
-
-In the latin-square study, we estimate the impact of atoms considering the same two perspectives of the previous experiment: \emph{correctness} (number of wrong answers)
-and \emph{time} (how long to provide a correct answer).
-In particular, in the time analysis, differently from the repeated  measures study, we compare the time required by participants to submit a correct response when analyzing code snippets with and without atom candidates.
-This study is based on the responses of 140 participants (a total of 70 replicas).
-All participants had taken at least some university course or hold a bachelor degree or equivalent.
-In addition, 21 participants hold a master's degree and three a doctorate degree.
-Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.
-Accordingly, we characterize the effect of atoms of confusion in JavaScript code taking into account the perceptions of both novice and experienced developers.
-
-\subsection{Correctness Analysis}
-% \castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me to which rate this refers.}
-
-\subsubsection*{Exploratory Data Analysis}
-
-As mentioned in Section~\ref{sec:meth:survey}, each participant in this experiment evaluated ten code snippets, from which five were in their obfuscated versions, whilst the other five contained clean versions of the code snippets (i.e., without the atom candidates).
-As in the repeated measures study, the participants should provide the expected outcomes of the code snippets.
-Differently from that study, in this one each participant only predicts the outcome of one version of each code snippet, either clean or obfuscated.
-As discussed, we collect information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{time}.
-Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the correctness experiment.
-
-Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the obfuscated versions.
-In particular, the presence of the \emph{Comma Operator} atom exhibits the highest impact on misunderstanding.
-Frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also result in many mistakes.
-%introduce high degrees of confusion.
-The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the clean versions of the code snippets.
-Also, the sample of responses with no atoms had almost no dispersion, which supports the argument that the clean versions of the code snippets are easier to evaluate correctly.
-
-% \begin{table}[htbp]
-% \caption{Difference in answer correctness between confusing and non-confusing pairs}
-% \begin{center}
-% \begin{small}
-% \begin{tabularx}
-% {{\linewidth}}{l p{1.5cm} p{1.1cm} p{1.1cm} p{1.2cm} }
-% \textbf{Atom} & \textbf{\%Correct} & \textbf{\%Correct} \\
-% &  \multicolumn{1}{l}{With AOC} \multicolumn{2}{l}{Without AOC}  & \Delta (\%)    \\
-%  \hline
-% Comma Operator & 40 & 93 & +132\%                 \\     
-% Automatic Semicolon  Insertion & 46 & 97 & +110\% \\
-% Post Increment & 69 & 91 & +31\%         \\
-% Omitted Curly Braces & 67 & 83 & +23\%   \\ 
-% Assignment as Value & 80 & 97 & +21\%    \\
-% Implicit Predicate & 83 & 97 & +16\%     \\
-% Logic as Control Flow & 59 & 68 & +15\%  \\
-% Ternary Operator & 86 & 94 & +9\%        \\
-% Pre-Increment & 71 & 76 & +7\%           \\
-% Arithmetic as Logic & 91 & 90 & -1\%    \\
-% \end{tabularx}
-% \end{small}
-% \end{center}
-% \label{results_correctness}
-% \end{table}
-
-\begin{table}[htbp]
-\caption{Summary of the correctness analysis}
-\label{tab:difference-correctness}
-\centering{
-  {\scriptsize
-  \begin{scriptsize}
-\begin{tabular}{lrrr} \toprule
-  Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
-  Comma Operator                &  28 &  65 & +132   \\ 
-  Automatic Semicolon Insertion &  32 &  68 & +112   \\ 
-  Post-Increment                &  48 &  64 & +33    \\ 
-  Omitted Curly Braces          &  47 &  58 &  +23   \\
-  Assignment as Value           &  56 &  68 &  +21   \\ 
-  Implicit Predicate            &  58 &  68 &  +17   \\ 
-  Logic as Control Flow         &  41 &  48 &  +17   \\ 
-  Conditional Operator              &  60 &  66 &  +10   \\ 
-  Pre-Increment                 &  50 &  53 &  +6    \\ 
-  Arithmetic as Logic           &  64 &  63 &  -2    \\ \bottomrule
-
-\end{tabular}
-\end{scriptsize}
-}}
-\end{table}
-
- %% Comma Operator          & 40 & 93  & +132 \\
- %% Post Increment          & 69 & 91  & + 31  \\
- %% Omitted Curly Braces    & 67 & 83  & +23 \\
- %% Assignment as Value     & 80 & 97  & +21 \\
- %% Implicit Predicate      & 83 & 97  & +16 \\
- %% Logic as Control Flow   & 59 & 68  & +15 \\
- %% Ternary Operator        & 86 & 94  & +9  \\
- %% Pre-Increment           & 71 & 76  & +7  \\
- %% Arithmetic as Logic     & 91 & 90  & +1  \\ \bottomrule
-
-\begin{figure}[b!]
-\noindent
- \centering
- \includegraphics[width=0.6\columnwidth]{images/wrong-answers-plot-1.pdf}
- \caption{Number of wrong answers of each subject.}
- \label{fig:boxplotcorrectness}
- \end{figure}% \castor{Fiz de tudo para ajeitar a figura porque o texto embaixo deveria ser "Clean versions" e "Confusing versions" mas não sei porque não funcionou.
-Até mudei a figura.
-Não sei o que está havendo.}
-
-% \rb{verificar a consistencia nas questoes de pesquisa mais especificas}
-
-\subsubsection*{Statistical analysis}
-
-We first use the \emph{Pearson's Chi-squared test}
-to investigate if there is a statistically significant difference in the frequency of correct and incorrect answers---due to the versions of the code snippets (obfuscated and clean code).
-The p-values for these tests are reported in the ``Chi-square test'' column of Table~\ref{tab:hypothesis-testing}.
-The results indicate that for five atom candidates (Comma Operator, Automatic Semicolon Insertion, Post-Increment, Assignment as Value, Implicit Predicate) the obfuscated versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$).
-This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
-
-We measure the effect size of the clean version of the code snippets into the answers' correctness using the \emph{Odds Ratio} (OR).
-The results are reported in the ``Odds Ratio'' column of Table~\ref{tab:hypothesis-testing}.
-In the table, we also report the confidence interval (CI) for the OR in the ``CI'' column.
-Although many of the intervals are wide, for six of the atom candidates the lower bound of the confidence interval is greater than or equal to 1.
-This indicates that, at a 95\% confidence level, a developer is likely to commit less mistakes when using the clean versions.
-%Considering only the atoms for which there was a significant difference in correctness for confusing and clean versions and u
-%Using the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), 
-For four of the atoms, Comma Operator, Automatic Semicolon Insertion, Assignment as Value, and Implicit Predicate, effect size can be considered large. For Implicit Predicate, it is medium. 
-%We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. 
-For instance, when comparing clean and obfuscated code snippets pertaining to the Comma Operator atom candidate, we observed an \emph{Odds Ratio} of \num{19.02}.
-This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding obfuscated version of the code snippet.
-Furthermore, with a 95\% confidence level, the odds ratio is between 6.6 and 68.22, i.e., although the error margin is wide, it is strongly in favor of the clean version.
-If there is a significant likelihood of a developer committing a mistake when analyzing a clean version, the lower bound of the CI should be a number between 0 and 1 (since the OR is a ratio).
-This is the case for the four atom candidates in the lower part of the table.
-%In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a participant was actually more likely to misunderstand the clean version.
-
-Finally, we also use a \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impacts correctness.
-The findings suggest that \emph{participant experience} impacts the results related to correctness (Figure~\ref{fig:correctness-over-experience}).
-Even though the presence of atoms of confusion reduces the number of correct answers for all \emph{experience groups}, this effect is not uniform.
-For instance, novice developers (under one year of experience) provide a higher number of wrong answers for the obfuscated version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the obfuscated version of the code snippets.
-It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.
-We replicate the \emph{Pearson's Chi-squared test} for each experience group and find that the statistical difference is less significant for those developers with more than ten years of experience.
-Differently from \emph{participant experience}, the factor \emph{participant education} does not significantly change the correctness of the answers.
-\castor{I'd like to have the numbers for the results above. What are the p-values, regression coefficients, etc.?}
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.3]{images/correctness-by-experience}
-  \caption{Impact of atoms of confusion on the correctness (over the four participant experience groups)}
-  \label{fig:correctness-over-experience}
-\end{figure}
-
-%% \begin{mh}
-%%   The results of our exploratory data analysis and   hypothesis testing suggest that the atom candidates Comma Operator, Automatic Semicolon Insertion, Post Increment, Omitted Curly Braces, Assignment as Value, and Implicit Predicate %, and Ternary Operator
-%%   introduce some degree of misunderstanding in JavaScript code. %Based on the results of this study, 
-%%   Therefore, they can be deemed atoms of confusion for JavaScript.
-%% \end{mh}
-
-%% \begin{mh}
-%% {\color{red}Comparing the results of different studies,
-%%   the impact of atoms candidates on source code
-%%   (mis)understanding differ according to the programming
-%%   language.}
-%% \end{mh}
-%% Regarding the first question we 
-%% address in the survey (\emph{Do code snippets that contain atoms of confusion produce a higher error rate than snippets where the atom is removed?}), we found evidence that the atoms of confusion lead programmers to misunderstand JavaScript code.
-We also realized that just one atom whose correction has a non-significant improvement in the percentage of correct answers---we found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of ten atoms we consider in the survey).
-
-\subsection{Time Analysis}
-
-\subsubsection*{Exploratory Data Analysis}
-Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both obfuscated and clean versions.
-We do not consider cases where the participants made mistakes.
-Seven atom candidates required more time from participants to predict a correct response.
-For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet.
-In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
-time on average to find the correct answer for the clean version of the code snippets. 
-Not all atom candidates, though, require less time for the participants to predict the answer.
-In fact, for  Arithmetic as Logic and Pre-Increment, the time to give a correct answer for the clean versions was
-29.06\% and 38.19\% than for the obfuscated versions---we did not confirm these atom candidates as atoms of confusion in the previous section.
-
-\begin{table}[htbp]
-\caption{Time in seconds to submit a correct answer}
-\centering{
-  {\scriptsize
-  \begin{scriptsize}
-\begin{tabular}{lrrr} \toprule
-Atom & Obfuscated  & Clean & $\Delta$(\%) \\ \midrule
-Comma Operator & 87.67 & 20.84 & -76.23 \\ 
-Automatic Semicolon Insertion & 46.08 & 22.04 & -52.17 \\ 
-Post-Increment & 28.70 & 25.67 & -10.56 \\ 
-Omitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
-Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
-Implicit Predicate & 36.24 & 24.01 & -33.75 \\ 
-Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
-Conditional Operator & 41.80 & 39.34 & -5.90 \\ 
-Pre-Increment & 30.71 & 42.45 & 38.19 \\  
-Arithmetic as Logic & 28.82 & 37.20 & 29.06 \\ \bottomrule
-\end{tabular}
-\label{tab:difference-time-taken}
-\end{scriptsize}
-}}
-\end{table}
-
- % latex table generated in R 4.0.4 by xtable 1.8-4 package
- % Thu Apr 22 07:58:21 2021
- \begin{table*}[ht]
-\caption{Hypotheses Testing (misunderstanding and time). Asterisks ($^{*}$) indicate a statistically significant difference.}
- \centering
- {\scriptsize
- \begin{tabular}{lrrr|rr}
-   \toprule
-Atom  & Chi-square & Odds          & Confidence & Mann-Whitey& Cliff's\\ 
- &  test &Ratio          & Interval &  U test & Delta \\  \midrule
-Comma Operator & \textbf{$<$ 0.0001*} & 19.02 & (6.60, 68.22) & \textbf{$<$ 0.0001*} & -0.53 \\ 
-Automatic Semicolon Insertion & \textbf{$<$ 0.0001*} & 39.33 & (9.21, 356.62) & 0.0980 & -0.16 \\ 
-Post-Increment & \textbf{0.0015*} & 4.83 & (1.73, 15.74) & 0.1201 & 0.15 \\ 
-Omitted Curly Braces & 0.0510 & 2.35 & (1.00, 5.77) & 0.9253 & -0.01 \\ 
-Assignment as Value & \textbf{0.0035*} & 8.39 & (1.81, 79.12) & 0.8268 & 0.02 \\ 
-Implicit Predicate & \textbf{0.0112*} & 6.95 & (1.46, 66.56) & \textbf{0.0029*} & -0.29 \\ 
-Logic as Control Flow & 0.2920 & 1.54 & (0.73, 3.28) & \textbf{$<$ 0.0001*} & -0.39 \\ 
-Conditional Operator & 0.1590 & 2.73 & (0.74, 12.57) & 0.2407 & -0.12 \\ 
-Pre-Increment & 0.7015 & 1.25 & (0.55, 2.85) & \textbf{0.0062*} & 0.27 \\ 
-Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & \textbf{0.0172*} & 0.23 \\ 
-    \bottomrule
- \end{tabular}}
- \label{tab:hypothesis-testing}
- \end{table*}
-
-\subsubsection*{Statistical analysis}
-We use the \emph{Mann-Whitney U test} to
-investigate the null hypothesis that developers 
-spend the same amount of time to correctly
-predict the outcome of clean and obfuscated versions of a code snippet.
-%, regardless of evaluating the confusing or clean version of the code.
-The results of Table~\ref{tab:hypothesis-testing}
-suggest that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
-For the first three, the analysis suggests that the participants need more time to predict the outcome of the code snippet in the obfuscated code version.
-For the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time to predict the outcome of the code snippets in the obfuscated version.
-We also computed the effect size using Cliff's Delta statistic.
-In the table, negative effect sizes suggest that it took less time to correctly predict the output of clean versions of the snippets.
-Considering the thresholds established by Romano et al.~\cite{Romano:2006:ASO} (0.147 $<$ $|$d$|$ $<$ 0.33 small, $|$d$|$ $<$ 0.474 medium, otherwise large), we found a large effect for the Comma Operator atom candidate; a medium effect size for Logic as Control Flow, and small effect sizes for Automatic Semicolon Insertion, Implicit Predicate, Arithmetic as Logic, Post-, and Pre-Increment.
-
-%% \begin{mh}
-%%   The results suggest that developers tend to spend
-%%   more time to predict the correct answer in the presence of the atom candidates Comma Operator, Logic as Control Flow, and Implicit Predicate.
-Conversely, the code snippets with the atom candidates Arithmetic as Logic and Pre-Increment demand less time to predict the correct answers.
-%% \end{mh}
-
-%%  Comma Operator        & 60 & 21  & -65  \\
-%% Logic as Control Flow & 85 & 49  & -42  \\
-%% Implicit Predicate    & 33 & 24  & -27  \\
-%% Omitted Curly Braces  & 43 & 31  & -27  \\
-%% Assignment as Value   & 53 & 49  & -7   \\
-%% Ternary Operator      & 42 & 42  & 0    \\
-%% Post Increment        & 27 & 27  & 0    \\
-%% Arithmetic as Logic   & 29 & 36  & +24  \\
-%% Pre-Increment         & 34 & 49  & +44  \\
-
-% \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}
-
-\section{Results of the Interview Study}
-\label{sec:interview-results}
-
-In this section we present the results of the interviews with the practitioners.
-We contrast the findings of the experiments of Sections~\ref{sec:repeated} and~\ref{sec:latin} with the opinion of software developers about their preferences regarding the obfuscated or clean versions of the code snippets.
-For the interview study, we only considered atoms that were analyzed in both the repeated measures (Section~\ref{sec:repeated}) and the latin square (Section~\ref{sec:latin}) studies.
-
-%% \rb{acho que essa separa\c c\~{a}o em dois rounds est\'{a}
-%%   gerando uma complexidade desnecess\'{a}ria. vou tentar
-%%   remover isso}.
-
-%In this section, we present the results of the 
-% interviews with the 15 practitioners who agreed to participate in the study. 
-A total of 15 practitioners took part of the interviews.
-We collected information regarding programming
-experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in the both experiments.
-We first presented them pairs of clean and obfuscated code snippets and then asked them to discuss their preferences towards one version.
-We did not indicate in any way whether any version was assumed to be confusing or not.
-% ask the participants to discuss their preferences (i.e., code with or without atom).
-
-%% In the second round of interviews, ten of the fifteen developers
-%% who participated in the first round were shown another 8 atom candidates to atom of confusion whose frequency we observed during the repository mining phase.
-The same procedure of showing code that behaves exactly the same and surveying for how easy to understand each snippet was applied in this phase.
-
-\begin{table}[!htb]
-    \centering
-    {\scriptsize
-    \caption{Summary of participants' preferences for code snippets \emph{with} and
-      \emph{without} atom candidates. The participants were only presented the code snippets, without any indication about whether one was confusing or not. 
-      Participants were also allowed to choose Neutral when they thought both sides were equally readable.}\label{tab:interview-results1}
-    \begin{tabular}{lrrr}\toprule
-      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
-      \cmidrule(lr){2-4}
-         Atom           & \multicolumn{1}{c}{Obfuscated}
-                                      &  \multicolumn{1}{c}{Clean}
-                                               & \multicolumn{1}{c}{Neutral} \\ \midrule
-         Comma Operator                  & 0  & 100    & 0     \\
-         Automatic Semicolon Insertion  & 0  & 80     & 0     \\
-         Post-Increment                  & 20 & 73.33  & 6.67  \\
-         Omitted Curly Braces            & 0  & 100    & 0     \\
-         Assignment as Value             & 20 & 60     & 20    \\
-         Implicit Predicate              & 20 & 73.33  & 6.67  \\
-         Logic as Control Flow           & 20 & 60     & 20    \\
-         Conditional Operator                & 60 & 26.67  & 13.3  \\
-         Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
-         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \midrule
-         \textsc{overall}                & 18 & 71.33  & 8.64  \\
-         \bottomrule
-    \end{tabular}
-    }
-
-\end{table}
-
-\subsubsection*{The Participants' perceptions of the atom candidates}
-
-Supporting the results of Sections~\ref{sec:repeated} and~\ref{sec:latin}, Table~\ref{tab:interview-results1}
- shows that for eight out of the
-10 scenarios surveyed, the majority of the respondents prefer the version of the
-code without the atom candidate.
-In no case the \emph{neutral}
-ratio was higher than the option for the clean version.
-%% An example entry is contained in Appendix~\ref{}. 
-%% For a full listing of the code snippets, visit the paper repository. 
-Only for the Conditional Operator atom candidate the participants preferred the obfuscated version, instead of the clean one.
-Figure~\ref{code:ternary} shows the code snippets for the corresponding obfuscated and clean versions.
-For this atom candidate, some participants who preferred the left-hand side version (obfuscated) still believed that the right-hand side version (clean) was more readable.
-The following quotes were extracted from the transcripts with
-% three interviewees:
-two interviewees:
-
-\begin{figure*}
-
-\noindent\begin{minipage}{.45\textwidth}
-\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Conditional Operator} atom).]
-let config = {size: 3, isActive: false};
-const_config = config.isActive === true 
-             ? config 
-             : {size: 10};
-console.log(_config.size);
-\end{lstlisting}
-\end{minipage}\hfill
-\begin{minipage}{.45\textwidth}
-\begin{lstlisting}[language=JavaScript, caption=\emph{Right-hand side} (without the atom).]
-let config = {size: 3, isActive: false}
-let _config;
-if(config.isActive === true) {
-  _config = config;
-}
-else{
-   _config = {size: 10};
-}
-console.log(_config.size);
-\end{lstlisting}
-\end{minipage}
-\caption{Example of a pair of code snippets used in the interview pertaining to the \emph{Conditional Operator} atom candidate}
-\label{code:ternary}
-\vspace{-0.2cm}
-\end{figure*}
-
-\begin{mq}
-\emph{``I prefer to write [code using the \lhs version], but I think [the \rhs version] is easier to read, especially for newer programmers''}.
-\end{mq}
-
-\begin{mq}
-\emph{``When I am programming, I write code with the conditional operator, [...], but, to be honest, I still think that 
-[the code using the \rhs version] 
-is easier to understand''}.
-\end{mq}
-
-%% \begin{mq}
-%% \emph{``I think [the \lhs version] is easier to understand, but [the \rhs version] is what I would write"}
-%% \end{mq}
-
-%% The atom candidate Ternary Operator also opens up
-%% the possibility for a derivative construct that JavaScript
-%% allows which can be rather confusing, and that is the nested
-%% ternaries construct, in which the right-hand side of the a
-%% ternary operator can be another ternary construct.
-%% While nested ternaries remove the number of lines that
-%% would be necessary to construct using nested if-then-else statements,
-%% they can become quite taxing to understand.
-%% Nested ternaries are a choice of atoms to be analyzed in future work.
-
-The Pre-Increment atom candidate also caused a
-conflict in one of the interviewees, who
-regarded the clean version as simpler to understand, but would still opt to write
-code with the atom. In contrast, %to such opinion,
-one of the participants found the version with the atom candidate more elegant, but recognized it was less readable, and was willing to sacrifice elegance for readability.
-In the repeated measures study, we found a significant difference in favor of the clean version, with a large effect size in its favor.
-In the latin square study, no difference could be observed in terms of correctness.
-However, there was a significant difference in terms of the time required to predict the output correctly in favor of the obfuscated version.
-%% \rb{pela discuss\~{a}o anterior, parece-me que alguns
-%%   desenvolvedores reconhecem que esses candidatos
-%%   a \'{a}tomos introduzem certa dificuldade de compreens\~{a}o;
-%%   mas mesmo assim optam por usar a constru\c c\~{a}o correspondente.
-%%   Acho que cabe um box aqui com essa discuss\~{a}o sobre isso.}
-
-When analyzing the Logic as Control Flow atom candidate, one of the interviewees gave an example of personal experience that might motivate one to avoid writing code using this construct:
-
-\begin{mq}
-  \emph{``This one is interesting, because I have written code that looks like the left-hand side [obfuscated version], and my colleagues complained that it was difficult to understand.
-Nowadays I prefer to write code using the version on the right-hand side [clean version]''}.
-\end{mq}
-\noindent
-None of the two studies found a significant difference in correctness between clean and obfuscated versions of snippets related to this atom candidate.
-However, in the latin square study the participants analyzing clean versions of the snippets required significantly less time.
-
-For two of the atom candidates, the interviewees were unanimous in their preference for the clean version: Comma Operator and Omitted Curly Braces.
-Not coincidentally, in both experiments there were significant differences in correctness in favor of the clean versions for both atoms.
-Regarding the first one, we could often notice during the interviews that the \emph{\lhs} (with the atom of confusion) caused significant confusion among the participants.
-One remark about the \emph{Comma Operator} atom of confusion is listed below:
-
-%% \begin{mq}
-%% \emph{``I just learned that [the \lhs version of this code] is possible. I did not even know it worked''}
-%% \end{mq}
-
-\begin{mq}
-\emph{``The code in [the \lhs version] is unlikely to be understood unless the programmer knows \clang or \cpplang''}.
-\end{mq}
-
-As for the atom candidate Omitted Curly Braces, one of the interviewees mentioned that, although they understand why one would opt not to use braces for simple if-then-else statements, they still advised against it, on grounds that:
-
-\begin{mq}
-\emph{``[I prefer the \rhs version of the code \ldots]
-If I want to see well-written, easily understandable code, then I also have to do my job.
-Therefore I believe that, since I do not know who is on the other end maintaining this code, and it could be any person with any level of expertise, then I try to write readable, easy-to-understand code''}.
-\end{mq}
-
-%This is an interesting perspective, because such a perhaps simple decision,  might hinder novice developers in the task of maintaining code; and one cannot make any assumption about the programmers' experience of who is going to maintain a code in a long run.
-
-%This is an interesting perspective because such a simple decision might hinder novice developers in maintaining code; one cannot assume the programmers' experience of who will maintain a code in the long run.
-
-\subsubsection*{Confusion in JavaScript Code}
-
-% The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confusing code.
-
-The final remarks drawn from the interviews are related to potentially confusing constructs suggested by the participants and their perspectives on JavaScript as a language.
-From the latter, we can also uncover some other forms in which the language itself might contribute to writing confusing code.
-
-One of the participants mentioned that the use of JavaScript's prototype-based inheritance can make it difficult to understand code, particularly when involving deep prototype chains.
-%This is a core feature of the language, and most high-level tools and frameworks abstract it away.
-%Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.
-% When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate.
-When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern---potentially leading to several levels of nested function calls---as extremely difficult to assimilate.
-One of the respondents stated:
-
-\begin{mq}
-\emph{``Nested callbacks are very confusing. Even writing them can be confusing, let alone understanding them.''}
-\end{mq}
-
-Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript.
-%Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.
-We also found other idioms that are JavaScript atom candidates, including
-\emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).
-The former was investigated in the repeated measures study but we did not find a significant difference between clean and obfuscated versions.
-As future work, we aim at investigating whether or not these atom candidates are more likely to introduce confusion in JavaScript code.
-
-\begin{figure}
-\begin{small}
-\begin{lstlisting}[language=JavaScript,caption=Example of arrow function.,label=arr1]
-let inc = (x) => x + 1;
-\end{lstlisting}
-\begin{lstlisting}[language=JavaScript,caption=Alternative version without arrow function.,label=arr2]
-let inc = function(x) {
-  return x + 1
-}
-\end{lstlisting}
-\vspace*{-0.6cm}
-\end{small}
-\end{figure}
-
-% citar uninitialized objects' types. floating point precision issues, 'this', type coercion
-
-% \subsection{Common Misunderstandings About the Language}
-
-% TODO: I don't see that our results support these claims.
-% Let us comment for now.
-
-%% Six practitioners regarded the JavaScript language as easy to learn, while four
-%% participants mentioned that the abundance of frameworks,
-%% coupled with the abstractions they provide, make it so that one
-%% can develop quickly in the language. However, as mentioned by one
-%% of the interviewees, ``\emph{Developers who have started
-%% programming in JavaScript might miss out on important lower-level
-%% aspects, such as garbage collection, memory management,
-%% and process management.}''.
-
-%% Moreover, such abstractions might lead a developer to start to program
-%% using the language without actually understanding some of its basic aspects.
-%% For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
-%% For the multi-purpose language that is one of the most, if not the most used one by practitioners,
-%% a lack of understanding of the languages core
-%% features may be a problem worth investigating in future studies.
-
-%% \begin{table}[!htb]
-%%     \centering
-%%     \caption{Interviews Round 2 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}).
-Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
-%%     \label{tab:interview-results2}
-%%     \begin{tabular}{lrrr}\toprule
-%%       & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
-%%       \cmidrule(lr){2-4}
-%%          \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
-%%                                       &  \multicolumn{1}{c}{\textsc{without aoc}}
-%%                                                & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
-%%          Property access         & 10  & 60  & 30  \\
-%%          Object spread           & 50 & 50     & 0    \\
-%%          Array spread            & 60  & 40    & 0     \\
-%%          Arrow function          & 40 & 40     & 20  \\
-%%          Array destructuring     & 40 & 40  & 20  \\
-%%          Object destructuring    & 40 & 40     & 20    \\
-%%          Type conversion         & 30  & 60    & 10     \\
-%%          Change Literal encoding & 40 & 50  & 10  \\ \midrule
-%%          \textsc{overall}        & 42.8 & 47.6  & 9.6  \\
-%%          \bottomrule
-%%     \end{tabular}
-%% \end{table}
-
-\section{Results of the MSR Effort}
-\label{sec:msr-results}
-
-In this section we present the results of the  software repository mining effort, whose goal is to reveal how prevalent atom candidates are in open source JavaScript projects. 
-We mined \minedprojects open source JavaScript repositories to understand how
-often atoms arise in real software. As in the interview study, this study focused on atoms that were investigated in both experiments.
-
-Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19} which investigate
-the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
-frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
-occur in at least 80\% of the projects. Considering the extremes, atom candidates Conditional Operator and Implicit Predicate were found in all repositories we mined,
-while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
-
-% Figure~\ref{fig:rate} summarizes this finding, showing that the occurrence of atoms of confusion in JavaScript systems range from 11.34\% (Comma Operator) to 89.69\% (Ternary Operator).
-
-%Figure~\ref{fig:atoms-occurrence} shows that the occurrence of in JavaScript systems range from
-
-% latex table generated in R 4.0.3 by xtable 1.8-4 package
-% Fri Nov 13 11:28:56 2020
-%%\begin{table}[!htb]
-%%\centering
-%% \caption{Summary of atoms occurrences on our dataset}
-%% 
-%%\setlength\tabcolsep{2pt} % default value: 6pt
-%%\label{tab:occurrences-summary}
-%%\begin{tabular}{lrcc}
-%%
-%%  \hline
-%%Atom & Projects & Occurrences/KLOC & Category \\ 
-%% \hline
-%%Ternary Operator & 100\% & 10.16 &  highly used \\ 
-%%Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
-%%Post Increment & 90.28\% & 5.62 &  highly used \\ 
-%%Pre-Increment & 83.33\% & 1.04 & commonly used \\ 
-%%Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
-%%Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
-%%Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
-%%Comma Operator & 15.28\% & 0.03 & little used \\ 
-%%   \hline
-%%\end{tabular}
-%%\end{table}
-
-% latex table generated in R 4.0.3 by xtable 1.8-4 package
-% Tue Apr 13 09:34:35 2021
-\begin{table}[ht]
-  \centering
-  {\scriptsize 
-\caption{Summary of atom candidate occurrences in our dataset.}
-\setlength\tabcolsep{2pt} % default value: 6pt
-\begin{tabular}{lrr}%c}
-  \toprule
-Atom & Projects & Occurr./KLOC \\%& Frequency \\ 
-  \midrule
-Implicit Predicate & 100\% & 19.89 \\%&  intensively used \\ 
-  Conditional Operator & 100\% & 10.16 \\%&  intensively used \\ 
-  Omitted Curly Braces & 91.67\% & 6.61 \\%&  intensively used \\ 
-  Post Increment & 90.28\% & 5.62 \\%&  intensively used \\ 
-  Pre Increment & 83.33\% & 1.04 \\%& commonly used \\ 
-  Assignment as Value & 81.94\% & 1.02 \\%& commonly used \\ 
-  Logic as Control Flow & 56.94\% & 0.95 \\%& commonly used \\ 
-  Automatic Semicolon Insertion & 33.33\% & 0.13 \\%& commonly used \\ 
-  Arithmetic as Logic & 23.61\% & 0.02 \\%& rarely used \\ 
-  Comma Operator & 15.28\% & 0.03 \\%& rarely used \\
-   \bottomrule
-\end{tabular}
-  }
-\label{tab:occurrences-summary}
-\end{table}
-
-Considering all JavaScript projects in our dataset, we found a total of \num{364873} atom candidates, though four atoms are responsible
-for 92.97\% of this total: Implicit Predicate, Post Increment, Conditional Operator, and Omitted Curly Braces.
-The first two, based on the results of Section~\ref{sec:latin} and Section~\ref{sec:repeated}, can be considered atoms of confusion.
-The remaining atom
-candidates comprise \num{25621} occurrences combined, 7.03\% of the total number of occurrences.
-The Comma Operator and Arithmetic as Logic atoms are the ones that arise less
-frequently (0.001\% in total with 232 and 165 occurrences, respectively).
-
-% Regarding frequency, we proceeded with a classification of \emph{atom occurrence}
-% based on thresholds used in a previous
-% study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. 
-% Atoms that occur 1000 times are classified as \textit{rarely used} while atoms with more than \num{10000} occurrences are classified as \textit{highly used}; \textit{commonly used} atoms occur in between.
-We calculate the frequency of each atom per KLOC.
-The four atoms mentioned in the previous paragraph,
-Implicit Predicate, Post Increment, Conditional Operator, and Omitted Curly Braces, occur frequently in the analyzed projects.
-All of them have more than 5 occurrences per KLOC on average.
-In particular, two atom candidates occur very frequently, Implicit Predicate (19.89 occurrences per KLOC) and Conditional Operator (10.16 occurrences per KLOC).
-On the other hand, Arithmetic as Logic and Comma Operator occur less than once for every 50 KLOC.
-
-%In our dataset of JavaScript projects, 
-%We found the Ternary Operator atom occurring more frequently than the Omitted Curly Braces, differently from what has been reported in a previous work~\cite{DBLP:conf/msr/GopsteinZFC18}. %
-
-The results of our experiments and interviews suggest that Conditional Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to the Latin-square study).
-Nonetheless, the Post-Increment Expression and Automatic Semicolon Insertion atoms are listed in the top four sources of misunderstanding  (Table~\ref{tab:difference-correctness}), and also frequently appear in open source systems.
-As such, refactoring existing systems to avoid Post-Increment and Automatic Semicolon Insertion might improve the readability of large amounts of code in JavaScript projects.

### 2022-12-02 s02.tex
changing the paper organization

--- 2022-07-29 4_Result.tex
@@
-%% \section{Results and Discussion}
-%% \label{sec:results}
+\section{Latin Square Study}\label{sec:meth:survey}
 
-%% In this section we present the main findings of our research.
-We first present the results of each study (Section~\ref{sec:survey-resuts}, Sections~\ref{sec:msr-results}, and Section~\ref{sec:interview-results}).
-After that, in Section~\ref{sec:discussion} we consolidate our findings and present some implications of our study.
+This experiment uses a different experimental design that makes it possible to control for the experience of the subjects.
+It employs a subset of the atom candidates used in the repeated measures study.
+More specifically, this study investigates ten of the atoms investigated in the repeated measures study:
+%Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first 
+%select a set of 9+1 atom candidates: 
+nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion).
+%Our supplementary material discusses and shows examples of the atoms we consider in our research.\castor{We should probably do this here or just refer to the table that I intend to insert for the replication study.}
 
-%\section{Results of the Survey}
-%\label{sec:survey-resuts}
+\subsection{Experimental design}
 
-%-----------------------------------------------------------------------
-\section{Results of the Repeated Measures Study}\label{sec:repeated}
+The design of the Latin square study blocks two variables (subject experience and the programs) and considers two treatments: the presence or absence of atom candidates within the programs.
+To achieve such a design goal, controlling the effect of experience and individual programs, we resort to the \textit{Latin Square Design}~\cite{Hunter-Experimenters}.
+Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of programs.
+The design of each square (a replica) is such that no treatment is repeated in the same row or column.
+For example, considering that we have a set of 10
+code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about clean programs, they will only be presented with clean versions of the code samples $s_6, s_7,..., s_{10}$.
+Furthermore, a given subject (P2), who constitutes the second row of our example square, will be asked questions about the clean versions for $s_1, s_2, ..., s_5$, and will answer questions about obfuscated programs for $s_6, s_7,..., s_{10}$.
+By doing that, we guarantee that all versions of the programs are contained within each square, and that each configuration occurs only once within a square.
+Figure \ref{fig:latinsquare} offers a visual representation of the concept.
 
-In this experiment, we investigate whether 24 atom candidates impact readability.
-We examine how the performance of the participants differ when predicting the outputs of the clean and obfuscated versions of small code snippets (Section~\ref{sec:repeated:correct}).
-Based on those results we establish which candidates can actually be considered atoms of confusion.
-In addition, similarly to Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, we verify whether there is a trade-off between correctness and time irrespective of the presence of atoms; we check if participants who correctly predict the outputs of the programs tend to do so more slowly than the ones that make incorrect predictions (Section~\ref{sec:repeated:time}).
+  \begin{figure}[htb!]
+      \noindent
+      \centering
+      \includegraphics[scale=.50]{images/latin-square.pdf}
+      \caption{Latin square design. Each ``square'' corresponds to 
+      a replica in our study. Each replica comprises two participants (square rows, e.g., P1 and P2) 
+      and two sets of programs (CS-1 and CS-2). We randomly apply the 
+      treatments (atom or non-atom code) to the cells of the squares.} 
+      \label{fig:latinsquare}
+  \end{figure}
 
-This study had 70 participants. Out of these, 67 have received university-level training in programming, 4 hold a PhD degree, 7 hold a master's degree, and 11 hold a bachelor's degree. 
-The median programming experience of the participants is four years (mean 6.5 years) and the median experience with JavaScript programming is one year (mean 2.5 years). 
-% General programming: median: 48, 1st quartile: 22,5, 3rd quartile: 91,5
-% JavaScript programming: median: 12, 1st quartile: 4, 3rd: 24
+For each of the 10 selected atom candidates, we write one short program containing the atom.
+As in the repeated measures study, we call them the obfuscated versions of the programs.
+We also write corresponding short programs without the atom candidates and call them the clean versions of the programs.
+Overall, the experiment employed 20 programs, 10 obfuscated and 10 clean.
+In order to reduce the cognitive effort, each subject was asked to predict the output of 10 programs, five obfuscated and five clean.
+The order in which they are presented is randomized.
+By doing this, we seek to minimize the chances of subjects being aware that the current listing they are analyzing contains (or not) atoms of confusion.
+That is, each subject should indicate what would be the outcomes of the programs, some of them having atoms of confusion (while other programs do not).
+Since participants are not exposed to obfuscated and clean versions corresponding to the same atom candidate, learning effect is not possible.
+We measure answer correctness and the total time each participant requires to participate in the experiment.
 
-%-----------------------------------------------------------------------
-\subsection{Correctness Analysis}\label{sec:repeated:correct}
+{\bf Study instrument.} 
+We implemented our experiment as a questionnaire in a web application and 
+%carried out a pilot. 
+%As part of this effort, we 
+carried out an informal pilot.% whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
+Undergraduate students and professional colleagues took part in the pilot.
+Some users reported layout defects, and many reported that the landing page did not explain the study well enough.
+We also spotted minor issues with our routines to create and populate the latin squares.
 
-\subsubsection*{Exploratory Data Analysis}
-For each of the 24 atom candidates, the participants predicted the outputs of two versions, one clean (control) and one obfuscated (treatment), as per the definition of a repeated measures design~\cite{Hunter-Experimenters}.
-Thus, each participant examined 48 code snippets.
-We also measured the time required by the participants to provide their answers.
-For most cases, participants made more mistakes when attempting to predict the outcomes of the obfuscated versions.
+We organize the questionnaire in three sections.
+The first section aims to characterize the subjects, asking their age, education level, and programming experience.
+We also include a check button, whose checking means users agree that all collected data will be used solely for research purposes.
+In the second section, we present to the participants a small set of instructions, where we explain how the study works and ask them to dedicate their attention to it.
+We stress to participants the importance of not using any aids during the experiment, such as online or console interpreters.
+For each question page, we kept track of whether or not the subjects switch windows.
 
-Table~\ref{tab:diff-correctness-repeated} presents the number of correct predictions for the clean and obfuscated versions of each atom candidate, as well as the percentage difference.
-In only three cases the participants predicted results correctly more often for the obfuscated versions, Arithmetic as Logic, Logic as Control Flow, and Property Access.
-The former two were empirically validated as atoms of confusion in the study of Gopstein and colleagues~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
-For seven of the atom candidates, the number of correct predictions for the clean versions is more than twice the corresponding number for the obfuscated versions.
-Some of these atoms are common programming patterns in JavaScript programs, such as pre- and post-increments and object destructuring.
-The difference was even greater for Type Conversion (+440\%) and Change of Literal Encoding (+300\%).
+The last section of the questionnaire presents a sequence of ten questions, each containing a program.
+For each question, there is a text box where the answer should be written.
+There is also an ``I do not know'' button, which, when clicked, leads the subject to the next question.
+In our setting, ``I do not know'' is treated as a wrong answer.
+The programs are also presented as images copied from a text editor.
+Upon submitting their answer for a particular question, the subjects are automatically led to a similar page, containing another program.
+Similarly to the repeated measures study, we do not provide feedback about time and correctness to the subjects.
 
-\begin{table}[t!]
-\caption{Summary of the correctness analysis.}
-\label{tab:diff-correctness-repeated}
-\centering{
-  \begin{scriptsize}
-\begin{tabular}{lrrr} \toprule
-  Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
-Type Conversion           &            5 &      27 &                 +440 \\
-Change of Literal Encoding &           9 &      36 &                 +300 \\
-Comma Operator            &            8 &      29 &                 +262 \\
-Object Destructuring      &           15 &      42 &                 +180 \\
-Assignment As Value       &           18 &      43 &                 +139 \\
-Pre-Increment             &           13 &      28 &                 +115 \\
-Post-Increment            &            8 &      17 &                 +112 \\
-Array Destructuring       &           15 &      28 &                 +87 \\
-Lack of Indentation no Braces     &           30 &      50 &                 +67 \\
-Omitted Curly Braces      &           35 &      51 &                 +46 \\
-Object Spread             &           21 &      30 &                 +43 \\
-Array Spread              &           22 &      31 &                 +41 \\
-Automatic Semicolon Insertion &       15 &      21 &                 +40 \\
-Infix Operator Precedence &           47 &      53 &                 +13 \\
-Arrow Function            &           54 &      58 &                  +7 \\
-Implicit Predicate        &           41 &      44 &                  +7 \\
-Lack of Indentation With Braces   &           53 &      55 &                  +4 \\
-Conditional Operator      &           60 &      62 &                  +3 \\
-Constant Variables        &           59 &      61 &                  +3 \\
-Dead Unreachable Repeated &           63 &      64 &                  +2 \\
-Repurposed Variables      &           23 &      23 &                   0 \\
-Arithmetic as Logic       &           52 &      47 &                 -10 \\
-Logic as Control Flow     &           33 &      28 &                 -15 \\
-Property Access           &           51 &      42 &                 -18 \\
-\bottomrule  
-\end{tabular}
-\end{scriptsize}
-}
-\end{table}
+% \rb{nao acho esse paragrafo necessario} 
+% \adriano{o de baixo ne? eu tambem acho}
 
-\subsubsection*{Statistical analysis}
+% As we mentioned before, we first wrote the code listings in a text editor, and took pictures of it.
+In the case of an atom of confusion that was exclusive to JavaScript, which we called \textit{Automatic Semicolon Insertion} (see Appendix~\ref{sec:appendix-atoms}), it was necessary to remove the syntax highlighter.
+Even though semicolons at the end of statements are optional to programmers in JavaScript, the interpreter automatically inserts them into the code.
+Our text editor was incorrectly highlighting a line break after a return statement, even thought it was valid JavaScript syntax.
+We had thus to turn the highlighter off to take the picture of this atom.
 
-Since this study uses a within-subject design, the data is paired and dependent.
-Our goal is to verify whether there is a significant difference in the performance of the same participants when examining clean and obfuscated versions of programs.
-The analyzed data is also dichotomous, since each participant either correctly predicts the output of a program or does not.
-Considering this scenario, we employ McNemar's test~\cite{McNemar:1947:NSE}, which is aimed at \textit{``judging the significance of the difference between correlated proportions''}.
+{\bf Study audience.}
+We posted the invitation to take part in the study on a JavaScript Reddit channel.\footnote{https://www.reddit.com/r/javascript/} We explained our research purposes and asked developers of any level of expertise to take part.
+Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares.
+We collected significant data on time taken and discrepancies in answer correctness between obfuscated and clean versions of the programs.
+Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the questionnaire.
+We discarded from our study all squares that contained incomplete rows.
 
-The ``McNemar test'' column of Table~\ref{tab:repeated:misunderstanding} indicates the p-values for the tests.
-These p-values are compared against an alpha of 0.0021, after applying Bonferroni correction (0.05/24).
-One of the atom candidates, Constant Variables, is omitted from the table because, due to the very low number of mistakes, it is not possible to perform the test.
-The table shows that eight of the 24 analyzed atom candidates can be considered atoms of confusion based on the results of our study, Object Destructuring, Type Conversion, Change of Literal Encoding, Comma Operator, Lack of Indentation No Braces, Assignment as Value, Pre-Increment, and Ommitted Curly Braces.
-Among these eight, only Object Destructuring is JavaScript-specific.
-Six of the other atoms have been previously reported for the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and six for Java~\cite{Langhout:2021:ACJ}.
-% Lack of Indentation No Braces is not specific to JavaScript and has also been studied by Langhout and Aniche~\cite{Langhout:2021:ACJ}. 
-For some candidates that have been previously identified as atoms, there was not a statistically significant difference.
-This is the case for Repurposed Variables, Logic as Control Flow, Implicit Predicate, and Conditional Operator.
-We examine this in Section~\ref{sec:discussion}.
+%Since we had a large enough number of samples, the squares we had to discard did not impact our analysis.
 
-We leverage the odds ratio (OR) as a measure of effect size, to quantify the meaningfulness~\cite{Ellis:2010:EGE} of these results.
-The odds ratio is an appropriate measure of effect size for scenarios where McNemar's test is applicable~\cite{Mangiafico:2016:SAE}.
-For example, the odds ratio for Type Conversion is 12.
-This suggests that the odds of a participant correctly predicting the output of a clean version while incorrectly predicting the output of the corresponding obfuscated version is 12 times higher than the odds of correctly predicting the output of the obfuscated version and incorrectly predicting the output of the correct one.
-
-According to the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), for most of the atoms where there is a statistically significant difference the odds ratio can be considered high.
-The value of the odds ratio for Object Destructuring is $\infty$ because no participant correctly guessed the output for the obfuscated version while incorrectly guessing the output for the clean one, which results in a denominator of 0.
-Omitted Curly Braces exhibits an OR of 6.3, which is borderline between medium and high~\cite{Chen:2010:HBB}.
-Array Destructuring and Post-Increment are two atom candidates that deserve a special note.
-The difference in the performance of participants examining clean and obfuscated versions of these atoms was not statistically significant, after correction.
-Notwithstanding, effect sizes were large and medium, respectively.
-This is a result that hints at practical relevance in spite of the p-values~\cite{Ellis:2010:EGE}.
-Further investigation is required for these two atom candidates.
-
-\begin{table}[th!]
-\caption{Hypotheses Testing (correctness).
-Asterisks ($^{*}$) indicate a statistically significant difference.
-Using Bonferroni correction, a p-value is considered statistically significant if it is lower than 0.0021.}
- \centering
- {\scriptsize 
- \begin{tabular}{lrr}
-   \toprule
-Atom candidate & McNemar test &      Odds Ratio \\
-\midrule
-Object Destructuring      &   \textbf{$<$ 0.0001*} & $\infty$ \\
-Type Conversion           &   \textbf{$<$ 0.0001*} & 12.0 \\
-Change of Literal Encoding   &   \textbf{$<$ 0.0001*} & 28.0 \\
-Comma Operator            &   \textbf{$<$ 0.0001*} & 11.5 \\
-Lack of Indentation No Braces     &   \textbf{$<$ 0.0001*} &  11.0 \\
-Assignment As Value       &   \textbf{$<$ 0.0001*} &  9.3 \\
-Pre-Increment              &  \textbf{0.0003*} &    16.0 \\
-Omitted Curly Braces      &  \textbf{0.0009*} &  6.3 \\
-Array Destructuring       &  0.0023 &     7.5 \\
-Post-Increment              &  0.0225 &     5.5 \\
-Array Spread              &  0.0490 &    3.3 \\
-Object Spread             &  0.0636 &     2.8 \\
-Property Access           &  0.0636 &     2.8 \\
-Infix Operator Precedence &  0.1796 &     2.5 \\
-Arrow Function            &  0.2891 &     3.0 \\
-Arithmetic As Logic       &  0.3323 &  1.8 \\
-Automated Semicolon Insertion  &  0.3915 &  1.4 \\
-Logic As Control Flow     &  0.4421 &  1.5 \\
-%Constant Variables        &  0.5000 &     0.0 \\
-Implicit Predicate        &  0.6900 &  1.3 \\
-Lack of Indentation With Braces   &  0.7266 &  1.7 \\
-Conditional Operator      &  0.7905 &  1.3 \\
-Dead Unreachable Repeated &  1.0000 &     1.5 \\
-Repurposed Variables      &   1.0000 &     1.0 \\
-\bottomrule
-\end{tabular}
-}
-\label{tab:repeated:misunderstanding}
-\end{table}
-
-%-----------------------------------------------------------------------
-\subsection{Time Analysis}\label{sec:repeated:time}
-
-\subsubsection*{Exploratory Data Analysis}
-
-In this section we investigate whether there is a trade-off between correctness and time required to predict program outputs, similarly to the analysis performed by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
-For this analysis, we consider only whether participants correctly predict the outputs of programs or not, independently of the presence or absence of atoms, and the time required to do so.
-The null hypothesis we investigate is that there is no significant difference in time between correct and incorrect predictions.
-Intuitively, we would expect correct responses to take longer, on the basis that participants would spend more time carefully analyzing the programs before providing an answer.
-
-Table~\ref{tab:repeated:time} presents the mean and median times required by the participants to provide answers.
-The data is based on the 1648 correct and 1572 incorrect answers collected in the study.
-Contrary to our intuition, the mean time required by participants to provide an incorrect answer is 25\% greater than the time required to provide a correct one and the median time is 20\% greater.
-The standard deviations suggest that there is great variation in the times required by different participants, for different code snippets.
-Notwithstanding, the standard deviation for incorrect answers is 62.58\% greater when compared to the same statistic for correct answers.
-
-\begin{table}[tb]
-  \caption{Time in seconds required by the participants to correctly and incorrectly predict the outcomes of the programs.}\label{tab:repeated:time}
-  \centering
-  {\scriptsize 
-  \begin{tabular}{lrrr}
-    \toprule  
-             & Mean  & Median & Std. Dev. \\
-    \midrule
-  Correct   & 35.44 & 24.36  & 38.43     \\
-  Incorrect & 44.50 & 29.29  & 62.48    \\
-  \bottomrule
-  \end{tabular}
-}
-  \end{table}
-
-\subsubsection*{Statistical Analysis}
-
-We employ the non-parametric Mann-Whitney U test, also known as Wilcoxon's rank-sum test, to verify the aforementioned null hypothesis, i.e., whether the times for correct and incorrect answers can be considered identical.
-That test produces a p-value of \textbf{0.00000046}, which shows that the difference is statistically significant and we can reject the null hypothesis.
-To gauge the magnitude of that difference we employ Cliff's Delta as a measure of effect size.
-This produces a result of \textbf{0.1027}, which suggests that, although participants tend to provide correct predictions more quickly, they do that only slightly so.
-
-%-----------------------------------------------------------------------
-\section{Results of the Latin-Square Study}\label{sec:latin}
+\subsection{Results}\label{sec:latin}
 
@@
 
-\subsection{Correctness Analysis}
+\subsubsection{Correctness Analysis}
 % \castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me to which rate this refers.}
 
-\subsubsection*{Exploratory Data Analysis}
-
+{\bf Exploratory Data Analysis.}
 As mentioned in Section~\ref{sec:meth:survey}, each participant in this experiment evaluated ten code snippets, from which five were in their obfuscated versions, whilst the other five contained clean versions of the code snippets (i.e., without the atom candidates).
@@
 
-\subsubsection*{Statistical analysis}
-
+{\bf Statistical analysis.}
 We first use the \emph{Pearson's Chi-squared test}
@@
 
-\subsection{Time Analysis}
+\subsubsection{Time Analysis}
 
-\subsubsection*{Exploratory Data Analysis}
+{\bf Exploratory Data Analysis.}
 Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both obfuscated and clean versions.
@@
 In fact, for  Arithmetic as Logic and Pre-Increment, the time to give a correct answer for the clean versions was
-29.06\% and 38.19\% than for the obfuscated versions---we did not confirm these atom candidates as atoms of confusion in the previous section.
+29.06\% and 38.19\% smaller than for the obfuscated versions---we did not confirm these atom candidates as atoms of confusion in the previous section.
 
@@
 
-\subsubsection*{Statistical analysis}
+{\bf Statistical analysis}
 We use the \emph{Mann-Whitney U test} to
@@
 % \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}
-
-\section{Results of the Interview Study}
-\label{sec:interview-results}
-
-In this section we present the results of the interviews with the practitioners.
-We contrast the findings of the experiments of Sections~\ref{sec:repeated} and~\ref{sec:latin} with the opinion of software developers about their preferences regarding the obfuscated or clean versions of the code snippets.
-For the interview study, we only considered atoms that were analyzed in both the repeated measures (Section~\ref{sec:repeated}) and the latin square (Section~\ref{sec:latin}) studies.
-
-%% \rb{acho que essa separa\c c\~{a}o em dois rounds est\'{a}
-%%   gerando uma complexidade desnecess\'{a}ria. vou tentar
-%%   remover isso}.
-
-%In this section, we present the results of the 
-% interviews with the 15 practitioners who agreed to participate in the study. 
-A total of 15 practitioners took part of the interviews.
-We collected information regarding programming
-experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in the both experiments.
-We first presented them pairs of clean and obfuscated code snippets and then asked them to discuss their preferences towards one version.
-We did not indicate in any way whether any version was assumed to be confusing or not.
-% ask the participants to discuss their preferences (i.e., code with or without atom).
-
-%% In the second round of interviews, ten of the fifteen developers
-%% who participated in the first round were shown another 8 atom candidates to atom of confusion whose frequency we observed during the repository mining phase.
-The same procedure of showing code that behaves exactly the same and surveying for how easy to understand each snippet was applied in this phase.
-
-\begin{table}[!htb]
-    \centering
-    {\scriptsize
-    \caption{Summary of participants' preferences for code snippets \emph{with} and
-      \emph{without} atom candidates. The participants were only presented the code snippets, without any indication about whether one was confusing or not. 
-      Participants were also allowed to choose Neutral when they thought both sides were equally readable.}\label{tab:interview-results1}
-    \begin{tabular}{lrrr}\toprule
-      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
-      \cmidrule(lr){2-4}
-         Atom           & \multicolumn{1}{c}{Obfuscated}
-                                      &  \multicolumn{1}{c}{Clean}
-                                               & \multicolumn{1}{c}{Neutral} \\ \midrule
-         Comma Operator                  & 0  & 100    & 0     \\
-         Automatic Semicolon Insertion  & 0  & 80     & 0     \\
-         Post-Increment                  & 20 & 73.33  & 6.67  \\
-         Omitted Curly Braces            & 0  & 100    & 0     \\
-         Assignment as Value             & 20 & 60     & 20    \\
-         Implicit Predicate              & 20 & 73.33  & 6.67  \\
-         Logic as Control Flow           & 20 & 60     & 20    \\
-         Conditional Operator                & 60 & 26.67  & 13.3  \\
-         Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
-         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \midrule
-         \textsc{overall}                & 18 & 71.33  & 8.64  \\
-         \bottomrule
-    \end{tabular}
-    }
-
-\end{table}
-
-\subsubsection*{The Participants' perceptions of the atom candidates}
-
-Supporting the results of Sections~\ref{sec:repeated} and~\ref{sec:latin}, Table~\ref{tab:interview-results1}
- shows that for eight out of the
-10 scenarios surveyed, the majority of the respondents prefer the version of the
-code without the atom candidate.
-In no case the \emph{neutral}
-ratio was higher than the option for the clean version.
-%% An example entry is contained in Appendix~\ref{}. 
-%% For a full listing of the code snippets, visit the paper repository. 
-Only for the Conditional Operator atom candidate the participants preferred the obfuscated version, instead of the clean one.
-Figure~\ref{code:ternary} shows the code snippets for the corresponding obfuscated and clean versions.
-For this atom candidate, some participants who preferred the left-hand side version (obfuscated) still believed that the right-hand side version (clean) was more readable.
-The following quotes were extracted from the transcripts with
-% three interviewees:
-two interviewees:
-
-\begin{figure*}
-
-\noindent\begin{minipage}{.45\textwidth}
-\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Conditional Operator} atom).]
-let config = {size: 3, isActive: false};
-const_config = config.isActive === true 
-             ? config 
-             : {size: 10};
-console.log(_config.size);
-\end{lstlisting}
-\end{minipage}\hfill
-\begin{minipage}{.45\textwidth}
-\begin{lstlisting}[language=JavaScript, caption=\emph{Right-hand side} (without the atom).]
-let config = {size: 3, isActive: false}
-let _config;
-if(config.isActive === true) {
-  _config = config;
-}
-else{
-   _config = {size: 10};
-}
-console.log(_config.size);
-\end{lstlisting}
-\end{minipage}
-\caption{Example of a pair of code snippets used in the interview pertaining to the \emph{Conditional Operator} atom candidate}
-\label{code:ternary}
-\vspace{-0.2cm}
-\end{figure*}
-
-\begin{mq}
-\emph{``I prefer to write [code using the \lhs version], but I think [the \rhs version] is easier to read, especially for newer programmers''}.
-\end{mq}
-
-\begin{mq}
-\emph{``When I am programming, I write code with the conditional operator, [...], but, to be honest, I still think that 
-[the code using the \rhs version] 
-is easier to understand''}.
-\end{mq}
-
-%% \begin{mq}
-%% \emph{``I think [the \lhs version] is easier to understand, but [the \rhs version] is what I would write"}
-%% \end{mq}
-
-%% The atom candidate Ternary Operator also opens up
-%% the possibility for a derivative construct that JavaScript
-%% allows which can be rather confusing, and that is the nested
-%% ternaries construct, in which the right-hand side of the a
-%% ternary operator can be another ternary construct.
-%% While nested ternaries remove the number of lines that
-%% would be necessary to construct using nested if-then-else statements,
-%% they can become quite taxing to understand.
-%% Nested ternaries are a choice of atoms to be analyzed in future work.
-
-The Pre-Increment atom candidate also caused a
-conflict in one of the interviewees, who
-regarded the clean version as simpler to understand, but would still opt to write
-code with the atom. In contrast, %to such opinion,
-one of the participants found the version with the atom candidate more elegant, but recognized it was less readable, and was willing to sacrifice elegance for readability.
-In the repeated measures study, we found a significant difference in favor of the clean version, with a large effect size in its favor.
-In the latin square study, no difference could be observed in terms of correctness.
-However, there was a significant difference in terms of the time required to predict the output correctly in favor of the obfuscated version.
-%% \rb{pela discuss\~{a}o anterior, parece-me que alguns
-%%   desenvolvedores reconhecem que esses candidatos
-%%   a \'{a}tomos introduzem certa dificuldade de compreens\~{a}o;
-%%   mas mesmo assim optam por usar a constru\c c\~{a}o correspondente.
-%%   Acho que cabe um box aqui com essa discuss\~{a}o sobre isso.}
-
-When analyzing the Logic as Control Flow atom candidate, one of the interviewees gave an example of personal experience that might motivate one to avoid writing code using this construct:
-
-\begin{mq}
-  \emph{``This one is interesting, because I have written code that looks like the left-hand side [obfuscated version], and my colleagues complained that it was difficult to understand.
-Nowadays I prefer to write code using the version on the right-hand side [clean version]''}.
-\end{mq}
-\noindent
-None of the two studies found a significant difference in correctness between clean and obfuscated versions of snippets related to this atom candidate.
-However, in the latin square study the participants analyzing clean versions of the snippets required significantly less time.
-
-For two of the atom candidates, the interviewees were unanimous in their preference for the clean version: Comma Operator and Omitted Curly Braces.
-Not coincidentally, in both experiments there were significant differences in correctness in favor of the clean versions for both atoms.
-Regarding the first one, we could often notice during the interviews that the \emph{\lhs} (with the atom of confusion) caused significant confusion among the participants.
-One remark about the \emph{Comma Operator} atom of confusion is listed below:
-
-%% \begin{mq}
-%% \emph{``I just learned that [the \lhs version of this code] is possible. I did not even know it worked''}
-%% \end{mq}
-
-\begin{mq}
-\emph{``The code in [the \lhs version] is unlikely to be understood unless the programmer knows \clang or \cpplang''}.
-\end{mq}
-
-As for the atom candidate Omitted Curly Braces, one of the interviewees mentioned that, although they understand why one would opt not to use braces for simple if-then-else statements, they still advised against it, on grounds that:
-
-\begin{mq}
-\emph{``[I prefer the \rhs version of the code \ldots]
-If I want to see well-written, easily understandable code, then I also have to do my job.
-Therefore I believe that, since I do not know who is on the other end maintaining this code, and it could be any person with any level of expertise, then I try to write readable, easy-to-understand code''}.
-\end{mq}
-
-%This is an interesting perspective, because such a perhaps simple decision,  might hinder novice developers in the task of maintaining code; and one cannot make any assumption about the programmers' experience of who is going to maintain a code in a long run.
-
-%This is an interesting perspective because such a simple decision might hinder novice developers in maintaining code; one cannot assume the programmers' experience of who will maintain a code in the long run.
-
-\subsubsection*{Confusion in JavaScript Code}
-
-% The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confusing code.
-
-The final remarks drawn from the interviews are related to potentially confusing constructs suggested by the participants and their perspectives on JavaScript as a language.
-From the latter, we can also uncover some other forms in which the language itself might contribute to writing confusing code.
-
-One of the participants mentioned that the use of JavaScript's prototype-based inheritance can make it difficult to understand code, particularly when involving deep prototype chains.
-%This is a core feature of the language, and most high-level tools and frameworks abstract it away.
-%Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.
-% When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate.
-When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern---potentially leading to several levels of nested function calls---as extremely difficult to assimilate.
-One of the respondents stated:
-
-\begin{mq}
-\emph{``Nested callbacks are very confusing. Even writing them can be confusing, let alone understanding them.''}
-\end{mq}
-
-Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript.
-%Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.
-We also found other idioms that are JavaScript atom candidates, including
-\emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).
-The former was investigated in the repeated measures study but we did not find a significant difference between clean and obfuscated versions.
-As future work, we aim at investigating whether or not these atom candidates are more likely to introduce confusion in JavaScript code.
-
-\begin{figure}
-\begin{small}
-\begin{lstlisting}[language=JavaScript,caption=Example of arrow function.,label=arr1]
-let inc = (x) => x + 1;
-\end{lstlisting}
-\begin{lstlisting}[language=JavaScript,caption=Alternative version without arrow function.,label=arr2]
-let inc = function(x) {
-  return x + 1
-}
-\end{lstlisting}
-\vspace*{-0.6cm}
-\end{small}
-\end{figure}
-
-% citar uninitialized objects' types. floating point precision issues, 'this', type coercion
-
-% \subsection{Common Misunderstandings About the Language}
-
-% TODO: I don't see that our results support these claims.
-% Let us comment for now.
-
-%% Six practitioners regarded the JavaScript language as easy to learn, while four
-%% participants mentioned that the abundance of frameworks,
-%% coupled with the abstractions they provide, make it so that one
-%% can develop quickly in the language. However, as mentioned by one
-%% of the interviewees, ``\emph{Developers who have started
-%% programming in JavaScript might miss out on important lower-level
-%% aspects, such as garbage collection, memory management,
-%% and process management.}''.
-
-%% Moreover, such abstractions might lead a developer to start to program
-%% using the language without actually understanding some of its basic aspects.
-%% For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
-%% For the multi-purpose language that is one of the most, if not the most used one by practitioners,
-%% a lack of understanding of the languages core
-%% features may be a problem worth investigating in future studies.
-
-%% \begin{table}[!htb]
-%%     \centering
-%%     \caption{Interviews Round 2 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}).
-Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
-%%     \label{tab:interview-results2}
-%%     \begin{tabular}{lrrr}\toprule
-%%       & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
-%%       \cmidrule(lr){2-4}
-%%          \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
-%%                                       &  \multicolumn{1}{c}{\textsc{without aoc}}
-%%                                                & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
-%%          Property access         & 10  & 60  & 30  \\
-%%          Object spread           & 50 & 50     & 0    \\
-%%          Array spread            & 60  & 40    & 0     \\
-%%          Arrow function          & 40 & 40     & 20  \\
-%%          Array destructuring     & 40 & 40  & 20  \\
-%%          Object destructuring    & 40 & 40     & 20    \\
-%%          Type conversion         & 30  & 60    & 10     \\
-%%          Change Literal encoding & 40 & 50  & 10  \\ \midrule
-%%          \textsc{overall}        & 42.8 & 47.6  & 9.6  \\
-%%          \bottomrule
-%%     \end{tabular}
-%% \end{table}
-
-\section{Results of the MSR Effort}
-\label{sec:msr-results}
-
-In this section we present the results of the  software repository mining effort, whose goal is to reveal how prevalent atom candidates are in open source JavaScript projects. 
-We mined \minedprojects open source JavaScript repositories to understand how
-often atoms arise in real software. As in the interview study, this study focused on atoms that were investigated in both experiments.
-
-Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19} which investigate
-the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
-frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
-occur in at least 80\% of the projects. Considering the extremes, atom candidates Conditional Operator and Implicit Predicate were found in all repositories we mined,
-while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
-
-% Figure~\ref{fig:rate} summarizes this finding, showing that the occurrence of atoms of confusion in JavaScript systems range from 11.34\% (Comma Operator) to 89.69\% (Ternary Operator).
-
-%Figure~\ref{fig:atoms-occurrence} shows that the occurrence of in JavaScript systems range from
-
-% latex table generated in R 4.0.3 by xtable 1.8-4 package
-% Fri Nov 13 11:28:56 2020
-%%\begin{table}[!htb]
-%%\centering
-%% \caption{Summary of atoms occurrences on our dataset}
-%% 
-%%\setlength\tabcolsep{2pt} % default value: 6pt
-%%\label{tab:occurrences-summary}
-%%\begin{tabular}{lrcc}
-%%
-%%  \hline
-%%Atom & Projects & Occurrences/KLOC & Category \\ 
-%% \hline
-%%Ternary Operator & 100\% & 10.16 &  highly used \\ 
-%%Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
-%%Post Increment & 90.28\% & 5.62 &  highly used \\ 
-%%Pre-Increment & 83.33\% & 1.04 & commonly used \\ 
-%%Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
-%%Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
-%%Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
-%%Comma Operator & 15.28\% & 0.03 & little used \\ 
-%%   \hline
-%%\end{tabular}
-%%\end{table}
-
-% latex table generated in R 4.0.3 by xtable 1.8-4 package
-% Tue Apr 13 09:34:35 2021
-\begin{table}[ht]
-  \centering
-  {\scriptsize 
-\caption{Summary of atom candidate occurrences in our dataset.}
-\setlength\tabcolsep{2pt} % default value: 6pt
-\begin{tabular}{lrr}%c}
-  \toprule
-Atom & Projects & Occurr./KLOC \\%& Frequency \\ 
-  \midrule
-Implicit Predicate & 100\% & 19.89 \\%&  intensively used \\ 
-  Conditional Operator & 100\% & 10.16 \\%&  intensively used \\ 
-  Omitted Curly Braces & 91.67\% & 6.61 \\%&  intensively used \\ 
-  Post Increment & 90.28\% & 5.62 \\%&  intensively used \\ 
-  Pre Increment & 83.33\% & 1.04 \\%& commonly used \\ 
-  Assignment as Value & 81.94\% & 1.02 \\%& commonly used \\ 
-  Logic as Control Flow & 56.94\% & 0.95 \\%& commonly used \\ 
-  Automatic Semicolon Insertion & 33.33\% & 0.13 \\%& commonly used \\ 
-  Arithmetic as Logic & 23.61\% & 0.02 \\%& rarely used \\ 
-  Comma Operator & 15.28\% & 0.03 \\%& rarely used \\
-   \bottomrule
-\end{tabular}
-  }
-\label{tab:occurrences-summary}
-\end{table}
-
-Considering all JavaScript projects in our dataset, we found a total of \num{364873} atom candidates, though four atoms are responsible
-for 92.97\% of this total: Implicit Predicate, Post Increment, Conditional Operator, and Omitted Curly Braces.
-The first two, based on the results of Section~\ref{sec:latin} and Section~\ref{sec:repeated}, can be considered atoms of confusion.
-The remaining atom
-candidates comprise \num{25621} occurrences combined, 7.03\% of the total number of occurrences.
-The Comma Operator and Arithmetic as Logic atoms are the ones that arise less
-frequently (0.001\% in total with 232 and 165 occurrences, respectively).
-
-% Regarding frequency, we proceeded with a classification of \emph{atom occurrence}
-% based on thresholds used in a previous
-% study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. 
-% Atoms that occur 1000 times are classified as \textit{rarely used} while atoms with more than \num{10000} occurrences are classified as \textit{highly used}; \textit{commonly used} atoms occur in between.
-We calculate the frequency of each atom per KLOC.
-The four atoms mentioned in the previous paragraph,
-Implicit Predicate, Post Increment, Conditional Operator, and Omitted Curly Braces, occur frequently in the analyzed projects.
-All of them have more than 5 occurrences per KLOC on average.
-In particular, two atom candidates occur very frequently, Implicit Predicate (19.89 occurrences per KLOC) and Conditional Operator (10.16 occurrences per KLOC).
-On the other hand, Arithmetic as Logic and Comma Operator occur less than once for every 50 KLOC.
-
-%In our dataset of JavaScript projects, 
-%We found the Ternary Operator atom occurring more frequently than the Omitted Curly Braces, differently from what has been reported in a previous work~\cite{DBLP:conf/msr/GopsteinZFC18}. %
-
-The results of our experiments and interviews suggest that Conditional Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to the Latin-square study).
-Nonetheless, the Post-Increment Expression and Automatic Semicolon Insertion atoms are listed in the top four sources of misunderstanding  (Table~\ref{tab:difference-correctness}), and also frequently appear in open source systems.
-As such, refactoring existing systems to avoid Post-Increment and Automatic Semicolon Insertion might improve the readability of large amounts of code in JavaScript projects.

### 2022-12-02 s03.tex
changing the paper organization

--- 2022-07-27 4_Result.tex
@@
-%% \section{Results and Discussion}
-%% \label{sec:results}
+\section{Interview Study}\label{sec:method:interview}
 
-%% In this section we present the main findings of our research.
-We first present the results of each study (Section~\ref{sec:survey-resuts}, Sections~\ref{sec:msr-results}, and Section~\ref{sec:interview-results}).
-After that, in Section~\ref{sec:discussion} we consolidate our findings and present some implications of our study.
+To complement the two experiments, we perform semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding programs containing atoms of confusion (research question RQ2).
+We also ask each participant if they know of any other JavaScript-specific construct or idiom they think is likely to make the code hard to understand.
+% (research question RQ3).
+In this section we detail the protocol we followed to conduct the interviews and to analyze the results.
 
-%\section{Results of the Survey}
-%\label{sec:survey-resuts}
+% Nós realizamos entrevistas semi-estruturas com o objetivo de identificar a percepção dos desenvolvedores com experiência em JavaScript sobre algumas questões relacionadas à compreensão de código em JavaScript.
+Assim, nesta Seção nós descrevemos os procedimentos adotados para selecionar os participantes para as entrevistas e detalhamos como as entrevistas foram conduzidas.
+Além disso, detalhamos como os resultados das entrevistas foram analisados.
 
-%-----------------------------------------------------------------------
-\section{Results of the Repeated Measures Study}\label{sec:repeated}
+\subsection{Participant Selection} We invite the participants of the interviews using a snowballing technique.
+That is, starting from our network of contacts, we invite an initial set of candidates to take part in our experiments.
+From this initial list, we ask for an indication of additional candidates.
+Our main selection criterion is that all participants should have been working with JavaScript in their daily professional activities.
+We invited a total of 17 developers, and 15 of them agreed to participate.
+%We conducted the interviews for a period of two weeks.
 
-In this experiment, we investigate whether 24 atom candidates impact readability.
-We examine how the performance of the participants differ when predicting the outputs of the clean and obfuscated versions of small code snippets (Section~\ref{sec:repeated:correct}).
-Based on those results we establish which candidates can actually be considered atoms of confusion.
-In addition, similarly to Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, we verify whether there is a trade-off between correctness and time irrespective of the presence of atoms; we check if participants who correctly predict the outputs of the programs tend to do so more slowly than the ones that make incorrect predictions (Section~\ref{sec:repeated:time}).
+\subsection{Interview Process} We conduct semi-structured interviews using web conferencing software.
+We record all the interviews with the consent of the participants.
+On average, the interviews last 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes.
+Two of the researchers conduct the interviews, and a third one listens to all the recordings to cross-validate the collected data.
+The interviews have three main parts.
+In the first one, we ask the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years, and other programming languages they have worked with.
+Table \ref{pinterview} summarizes this demographic information.
 
-This study had 70 participants. Out of these, 67 have received university-level training in programming, 4 hold a PhD degree, 7 hold a master's degree, and 11 hold a bachelor's degree. 
-The median programming experience of the participants is four years (mean 6.5 years) and the median experience with JavaScript programming is one year (mean 2.5 years). 
-% General programming: median: 48, 1st quartile: 22,5, 3rd quartile: 91,5
-% JavaScript programming: median: 12, 1st quartile: 4, 3rd: 24
+\begin{table}[htb!]
 
-%-----------------------------------------------------------------------
-\subsection{Correctness Analysis}\label{sec:repeated:correct}
-
-\subsubsection*{Exploratory Data Analysis}
-For each of the 24 atom candidates, the participants predicted the outputs of two versions, one clean (control) and one obfuscated (treatment), as per the definition of a repeated measures design~\cite{Hunter-Experimenters}.
-Thus, each participant examined 48 code snippets.
-We also measured the time required by the participants to provide their answers.
-For most cases, participants made more mistakes when attempting to predict the outcomes of the obfuscated versions.
-
-Table~\ref{tab:diff-correctness-repeated} presents the number of correct predictions for the clean and obfuscated versions of each atom candidate, as well as the percentage difference.
-In only three cases the participants predicted results correctly more often for the obfuscated versions, Arithmetic as Logic, Logic as Control Flow, and Property Access.
-The former two were empirically validated as atoms of confusion in the study of Gopstein and colleagues~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
-For seven of the atom candidates, the number of correct predictions for the clean versions is more than twice the corresponding number for the obfuscated versions.
-Some of these atoms are common programming patterns in JavaScript programs, such as pre- and post-increments and object destructuring.
-The difference was even greater for Type Conversion (+440\%) and Change of Literal Encoding (+300\%).
-
-\begin{table}[t!]
-\caption{Summary of the correctness analysis.}
-\label{tab:diff-correctness-repeated}
-\centering{
+  \centering
+  \caption{Demographic information of the participants}
 \begin{scriptsize}  
-\begin{tabular}{lrrr} \toprule
-  Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
-Type Conversion           &            5 &      27 &                 +440 \\
-Change of Literal Encoding &           9 &      36 &                 +300 \\
-Comma Operator            &            8 &      29 &                 +262 \\
-Object Destructuring      &           15 &      42 &                 +180 \\
-Assignment As Value       &           18 &      43 &                 +139 \\
-Pre-Increment             &           13 &      28 &                 +115 \\
-Post-Increment            &            8 &      17 &                 +112 \\
-Array Destructuring       &           15 &      28 &                 +87 \\
-Lack of Indentation no Braces     &           30 &      50 &                 +67 \\
-Omitted Curly Braces      &           35 &      51 &                 +46 \\
-Object Spread             &           21 &      30 &                 +43 \\
-Array Spread              &           22 &      31 &                 +41 \\
-Automatic Semicolon Insertion &       15 &      21 &                 +40 \\
-Infix Operator Precedence &           47 &      53 &                 +13 \\
-Arrow Function            &           54 &      58 &                  +7 \\
-Implicit Predicate        &           41 &      44 &                  +7 \\
-Lack of Indentation With Braces   &           53 &      55 &                  +4 \\
-Conditional Operator      &           60 &      62 &                  +3 \\
-Constant Variables        &           59 &      61 &                  +3 \\
-Dead Unreachable Repeated &           63 &      64 &                  +2 \\
-Repurposed Variables      &           23 &      23 &                   0 \\
-Arithmetic as Logic       &           52 &      47 &                 -10 \\
-Logic as Control Flow     &           33 &      28 &                 -15 \\
-Property Access           &           51 &      42 &                 -18 \\
-\bottomrule  
+\begin{tabular}{clcl}
+\toprule
+ID & Education & JS Experience & Other Languages \\ \midrule 
+P1 & BSc Degree & 9 years & Java, PHP, C, Go  
+\\ 
+P2 & HS Degree & 3 years & Python, Go, Dart, Lua, C++, C\#
+\\ 
+P3 & BSc Degree & 4 years & Java, C
+\\ 
+P4 & Undergraduate & 3 years & Python, C, C++, Java, Go
+\\ 
+P5 & Master Student & 3 years & Python, SQL
+\\ 
+P6 & Bsc Degree & 15  years  & Java, PHP, C, Python , Ruby, C\#
+\\ 
+P7 & BSc Degree  & 6 years & C, C++, Java, Assembly, Kotlin
+\\ 
+P8 & PhD Degree & 2 years & Java, Python
+\\ 
+P9 & BSc Degree  & 5 years & PHP
+\\ 
+P10 & Master Student & 4 years & Java, C\# and Python
+\\ 
+P11 & Master Student & 4 years & Java, Erlang, C\#, Cobol
+\\ 
+P12 & Master Degree & 1 year & C
+\\ 
+P13 & Master Degree & 13 years & Java, PHP
+\\ 
+P14 & Master Student & 3 years & C, Python, Ruby
+\\ 
+P15 &  BSc degree  & 2 years & Java, PHP
+\\ \bottomrule
 \end{tabular}
 \end{scriptsize}
-}
+    \label{pinterview}
 \end{table}
 
-\subsubsection*{Statistical analysis}
+In the second part of the interview our aim is to allow the subjects to describe their JavaScript experience, as well as to allow them to reveal any JavaScript constructs they regard as innately confusing.
+This allows us to identify potential atom candidates that are more specific to the JavaScript language.
+Examples of questions we explore in this section include: \emph{Does JavaScript favor developers to produce code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as especially confusing?}
 
-Since this study uses a within-subject design, the data is paired and dependent.
-Our goal is to verify whether there is a significant difference in the performance of the same participants when examining clean and obfuscated versions of programs.
-The analyzed data is also dichotomous, since each participant either correctly predicts the output of a program or does not.
-Considering this scenario, we employ McNemar's test~\cite{McNemar:1947:NSE}, which is aimed at \textit{``judging the significance of the difference between correlated proportions''}.
+In the third part of the interview, participants are shown pairs of programs that were used in the experiments, where each pair contains a clean and an obfuscated version of the same atom candidate.
+For this part we only use atoms that appear in both experiments.
+The participants
+are asked to evaluate which version of the code is easier to understand.
+To avoid introducing bias in the answers, the interviewers do not explain that one of the versions in each pair contains the atom under investigation.
+Subjects are just presented the pairs of programs and allowed to take the necessary time to decide on the most readable one.
 
-The ``McNemar test'' column of Table~\ref{tab:repeated:misunderstanding} indicates the p-values for the tests.
-These p-values are compared against an alpha of 0.0021, after applying Bonferroni correction (0.05/24).
-One of the atom candidates, Constant Variables, is omitted from the table because, due to the very low number of mistakes, it is not possible to perform the test.
-The table shows that eight of the 24 analyzed atom candidates can be considered atoms of confusion based on the results of our study, Object Destructuring, Type Conversion, Change of Literal Encoding, Comma Operator, Lack of Indentation No Braces, Assignment as Value, Pre-Increment, and Ommitted Curly Braces.
-Among these eight, only Object Destructuring is JavaScript-specific.
-Six of the other atoms have been previously reported for the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and six for Java~\cite{Langhout:2021:ACJ}.
-% Lack of Indentation No Braces is not specific to JavaScript and has also been studied by Langhout and Aniche~\cite{Langhout:2021:ACJ}. 
-For some candidates that have been previously identified as atoms, there was not a statistically significant difference.
-This is the case for Repurposed Variables, Logic as Control Flow, Implicit Predicate, and Conditional Operator.
-We examine this in Section~\ref{sec:discussion}.
+\subsection{Interview Analysis}
 
-We leverage the odds ratio (OR) as a measure of effect size, to quantify the meaningfulness~\cite{Ellis:2010:EGE} of these results.
-The odds ratio is an appropriate measure of effect size for scenarios where McNemar's test is applicable~\cite{Mangiafico:2016:SAE}.
-For example, the odds ratio for Type Conversion is 12.
-This suggests that the odds of a participant correctly predicting the output of a clean version while incorrectly predicting the output of the corresponding obfuscated version is 12 times higher than the odds of correctly predicting the output of the obfuscated version and incorrectly predicting the output of the correct one.
+We first transcribe each interview and then examine the broad distribution of the answers.
+Our goal is to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular.
+We follow-up with an open-coding procedure, highlighting the main themes and quoting the answers of the participants.
+We present these results in Section~\ref{sec:interview-results}.
 
-According to the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), for most of the atoms where there is a statistically significant difference the odds ratio can be considered high.
-The value of the odds ratio for Object Destructuring is $\infty$ because no participant correctly guessed the output for the obfuscated version while incorrectly guessing the output for the clean one, which results in a denominator of 0.
-Omitted Curly Braces exhibits an OR of 6.3, which is borderline between medium and high~\cite{Chen:2010:HBB}.
-Array Destructuring and Post-Increment are two atom candidates that deserve a special note.
-The difference in the performance of participants examining clean and obfuscated versions of these atoms was not statistically significant, after correction.
-Notwithstanding, effect sizes were large and medium, respectively.
-This is a result that hints at practical relevance in spite of the p-values~\cite{Ellis:2010:EGE}.
-Further investigation is required for these two atom candidates.
-
-\begin{table}[th!]
-\caption{Hypotheses Testing (correctness).
-Asterisks ($^{*}$) indicate a statistically significant difference.
-Using Bonferroni correction, a p-value is considered statistically significant if it is lower than 0.0021.}
- \centering
- {\scriptsize 
- \begin{tabular}{lrr}
-   \toprule
-Atom candidate & McNemar test &      Odds Ratio \\
-\midrule
-Object Destructuring      &   \textbf{$<$ 0.0001*} & $\infty$ \\
-Type Conversion           &   \textbf{$<$ 0.0001*} & 12.0 \\
-Change of Literal Encoding   &   \textbf{$<$ 0.0001*} & 28.0 \\
-Comma Operator            &   \textbf{$<$ 0.0001*} & 11.5 \\
-Lack of Indentation No Braces     &   \textbf{$<$ 0.0001*} &  11.0 \\
-Assignment As Value       &   \textbf{$<$ 0.0001*} &  9.3 \\
-Pre-Increment              &  \textbf{0.0003*} &    16.0 \\
-Omitted Curly Braces      &  \textbf{0.0009*} &  6.3 \\
-Array Destructuring       &  0.0023 &     7.5 \\
-Post-Increment              &  0.0225 &     5.5 \\
-Array Spread              &  0.0490 &    3.3 \\
-Object Spread             &  0.0636 &     2.8 \\
-Property Access           &  0.0636 &     2.8 \\
-Infix Operator Precedence &  0.1796 &     2.5 \\
-Arrow Function            &  0.2891 &     3.0 \\
-Arithmetic As Logic       &  0.3323 &  1.8 \\
-Automated Semicolon Insertion  &  0.3915 &  1.4 \\
-Logic As Control Flow     &  0.4421 &  1.5 \\
-%Constant Variables        &  0.5000 &     0.0 \\
-Implicit Predicate        &  0.6900 &  1.3 \\
-Lack of Indentation With Braces   &  0.7266 &  1.7 \\
-Conditional Operator      &  0.7905 &  1.3 \\
-Dead Unreachable Repeated &  1.0000 &     1.5 \\
-Repurposed Variables      &   1.0000 &     1.0 \\
-\bottomrule
-\end{tabular}
-}
-\label{tab:repeated:misunderstanding}
-\end{table}
-
-%-----------------------------------------------------------------------
-\subsection{Time Analysis}\label{sec:repeated:time}
-
-\subsubsection*{Exploratory Data Analysis}
-
-In this section we investigate whether there is a trade-off between correctness and time required to predict program outputs, similarly to the analysis performed by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
-For this analysis, we consider only whether participants correctly predict the outputs of programs or not, independently of the presence or absence of atoms, and the time required to do so.
-The null hypothesis we investigate is that there is no significant difference in time between correct and incorrect predictions.
-Intuitively, we would expect correct responses to take longer, on the basis that participants would spend more time carefully analyzing the programs before providing an answer.
-
-Table~\ref{tab:repeated:time} presents the mean and median times required by the participants to provide answers.
-The data is based on the 1648 correct and 1572 incorrect answers collected in the study.
-Contrary to our intuition, the mean time required by participants to provide an incorrect answer is 25\% greater than the time required to provide a correct one and the median time is 20\% greater.
-The standard deviations suggest that there is great variation in the times required by different participants, for different code snippets.
-Notwithstanding, the standard deviation for incorrect answers is 62.58\% greater when compared to the same statistic for correct answers.
-
-\begin{table}[tb]
-  \caption{Time in seconds required by the participants to correctly and incorrectly predict the outcomes of the programs.}\label{tab:repeated:time}
-  \centering
-  {\scriptsize 
-  \begin{tabular}{lrrr}
-    \toprule  
-             & Mean  & Median & Std. Dev. \\
-    \midrule
-  Correct   & 35.44 & 24.36  & 38.43     \\
-  Incorrect & 44.50 & 29.29  & 62.48    \\
-  \bottomrule
-  \end{tabular}
-}
-  \end{table}
-
-\subsubsection*{Statistical Analysis}
-
-We employ the non-parametric Mann-Whitney U test, also known as Wilcoxon's rank-sum test, to verify the aforementioned null hypothesis, i.e., whether the times for correct and incorrect answers can be considered identical.
-That test produces a p-value of \textbf{0.00000046}, which shows that the difference is statistically significant and we can reject the null hypothesis.
-To gauge the magnitude of that difference we employ Cliff's Delta as a measure of effect size.
-This produces a result of \textbf{0.1027}, which suggests that, although participants tend to provide correct predictions more quickly, they do that only slightly so.
-
-%-----------------------------------------------------------------------
-\section{Results of the Latin-Square Study}\label{sec:latin}
-
-In the latin-square study, we estimate the impact of atoms considering the same two perspectives of the previous experiment: \emph{correctness} (number of wrong answers)
-and \emph{time} (how long to provide a correct answer).
-In particular, in the time analysis, differently from the repeated  measures study, we compare the time required by participants to submit a correct response when analyzing code snippets with and without atom candidates.
-This study is based on the responses of 140 participants (a total of 70 replicas).
-All participants had taken at least some university course or hold a bachelor degree or equivalent.
-In addition, 21 participants hold a master's degree and three a doctorate degree.
-Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.
-Accordingly, we characterize the effect of atoms of confusion in JavaScript code taking into account the perceptions of both novice and experienced developers.
-
-\subsection{Correctness Analysis}
-% \castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me to which rate this refers.}
-
-\subsubsection*{Exploratory Data Analysis}
-
-As mentioned in Section~\ref{sec:meth:survey}, each participant in this experiment evaluated ten code snippets, from which five were in their obfuscated versions, whilst the other five contained clean versions of the code snippets (i.e., without the atom candidates).
-As in the repeated measures study, the participants should provide the expected outcomes of the code snippets.
-Differently from that study, in this one each participant only predicts the outcome of one version of each code snippet, either clean or obfuscated.
-As discussed, we collect information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{time}.
-Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the correctness experiment.
-
-Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the obfuscated versions.
-In particular, the presence of the \emph{Comma Operator} atom exhibits the highest impact on misunderstanding.
-Frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also result in many mistakes.
-%introduce high degrees of confusion.
-The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the clean versions of the code snippets.
-Also, the sample of responses with no atoms had almost no dispersion, which supports the argument that the clean versions of the code snippets are easier to evaluate correctly.
-
-% \begin{table}[htbp]
-% \caption{Difference in answer correctness between confusing and non-confusing pairs}
-% \begin{center}
-% \begin{small}
-% \begin{tabularx}
-% {{\linewidth}}{l p{1.5cm} p{1.1cm} p{1.1cm} p{1.2cm} }
-% \textbf{Atom} & \textbf{\%Correct} & \textbf{\%Correct} \\
-% &  \multicolumn{1}{l}{With AOC} \multicolumn{2}{l}{Without AOC}  & \Delta (\%)    \\
-%  \hline
-% Comma Operator & 40 & 93 & +132\%                 \\     
-% Automatic Semicolon  Insertion & 46 & 97 & +110\% \\
-% Post Increment & 69 & 91 & +31\%         \\
-% Omitted Curly Braces & 67 & 83 & +23\%   \\ 
-% Assignment as Value & 80 & 97 & +21\%    \\
-% Implicit Predicate & 83 & 97 & +16\%     \\
-% Logic as Control Flow & 59 & 68 & +15\%  \\
-% Ternary Operator & 86 & 94 & +9\%        \\
-% Pre-Increment & 71 & 76 & +7\%           \\
-% Arithmetic as Logic & 91 & 90 & -1\%    \\
-% \end{tabularx}
-% \end{small}
-% \end{center}
-% \label{results_correctness}
-% \end{table}
-
-\begin{table}[htbp]
-\caption{Summary of the correctness analysis}
-\label{tab:difference-correctness}
-\centering{
-  {\scriptsize
-  \begin{scriptsize}
-\begin{tabular}{lrrr} \toprule
-  Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
-  Comma Operator                &  28 &  65 & +132   \\ 
-  Automatic Semicolon Insertion &  32 &  68 & +112   \\ 
-  Post-Increment                &  48 &  64 & +33    \\ 
-  Omitted Curly Braces          &  47 &  58 &  +23   \\
-  Assignment as Value           &  56 &  68 &  +21   \\ 
-  Implicit Predicate            &  58 &  68 &  +17   \\ 
-  Logic as Control Flow         &  41 &  48 &  +17   \\ 
-  Conditional Operator              &  60 &  66 &  +10   \\ 
-  Pre-Increment                 &  50 &  53 &  +6    \\ 
-  Arithmetic as Logic           &  64 &  63 &  -2    \\ \bottomrule
-
-\end{tabular}
-\end{scriptsize}
-}}
-\end{table}
-
- %% Comma Operator          & 40 & 93  & +132 \\
- %% Post Increment          & 69 & 91  & + 31  \\
- %% Omitted Curly Braces    & 67 & 83  & +23 \\
- %% Assignment as Value     & 80 & 97  & +21 \\
- %% Implicit Predicate      & 83 & 97  & +16 \\
- %% Logic as Control Flow   & 59 & 68  & +15 \\
- %% Ternary Operator        & 86 & 94  & +9  \\
- %% Pre-Increment           & 71 & 76  & +7  \\
- %% Arithmetic as Logic     & 91 & 90  & +1  \\ \bottomrule
-
-\begin{figure}[b!]
-\noindent
- \centering
- \includegraphics[width=0.6\columnwidth]{images/wrong-answers-plot-1.pdf}
- \caption{Number of wrong answers of each subject.}
- \label{fig:boxplotcorrectness}
- \end{figure}% \castor{Fiz de tudo para ajeitar a figura porque o texto embaixo deveria ser "Clean versions" e "Confusing versions" mas não sei porque não funcionou.
-Até mudei a figura.
-Não sei o que está havendo.}
-
-% \rb{verificar a consistencia nas questoes de pesquisa mais especificas}
-
-\subsubsection*{Statistical analysis}
-
-We first use the \emph{Pearson's Chi-squared test}
-to investigate if there is a statistically significant difference in the frequency of correct and incorrect answers---due to the versions of the code snippets (obfuscated and clean code).
-The p-values for these tests are reported in the ``Chi-square test'' column of Table~\ref{tab:hypothesis-testing}.
-The results indicate that for five atom candidates (Comma Operator, Automatic Semicolon Insertion, Post-Increment, Assignment as Value, Implicit Predicate) the obfuscated versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$).
-This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
-
-We measure the effect size of the clean version of the code snippets into the answers' correctness using the \emph{Odds Ratio} (OR).
-The results are reported in the ``Odds Ratio'' column of Table~\ref{tab:hypothesis-testing}.
-In the table, we also report the confidence interval (CI) for the OR in the ``CI'' column.
-Although many of the intervals are wide, for six of the atom candidates the lower bound of the confidence interval is greater than or equal to 1.
-This indicates that, at a 95\% confidence level, a developer is likely to commit less mistakes when using the clean versions.
-%Considering only the atoms for which there was a significant difference in correctness for confusing and clean versions and u
-%Using the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), 
-For four of the atoms, Comma Operator, Automatic Semicolon Insertion, Assignment as Value, and Implicit Predicate, effect size can be considered large. For Implicit Predicate, it is medium. 
-%We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. 
-For instance, when comparing clean and obfuscated code snippets pertaining to the Comma Operator atom candidate, we observed an \emph{Odds Ratio} of \num{19.02}.
-This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding obfuscated version of the code snippet.
-Furthermore, with a 95\% confidence level, the odds ratio is between 6.6 and 68.22, i.e., although the error margin is wide, it is strongly in favor of the clean version.
-If there is a significant likelihood of a developer committing a mistake when analyzing a clean version, the lower bound of the CI should be a number between 0 and 1 (since the OR is a ratio).
-This is the case for the four atom candidates in the lower part of the table.
-%In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a participant was actually more likely to misunderstand the clean version.
-
-Finally, we also use a \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impacts correctness.
-The findings suggest that \emph{participant experience} impacts the results related to correctness (Figure~\ref{fig:correctness-over-experience}).
-Even though the presence of atoms of confusion reduces the number of correct answers for all \emph{experience groups}, this effect is not uniform.
-For instance, novice developers (under one year of experience) provide a higher number of wrong answers for the obfuscated version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the obfuscated version of the code snippets.
-It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.
-We replicate the \emph{Pearson's Chi-squared test} for each experience group and find that the statistical difference is less significant for those developers with more than ten years of experience.
-Differently from \emph{participant experience}, the factor \emph{participant education} does not significantly change the correctness of the answers.
-\castor{I'd like to have the numbers for the results above. What are the p-values, regression coefficients, etc.?}
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.3]{images/correctness-by-experience}
-  \caption{Impact of atoms of confusion on the correctness (over the four participant experience groups)}
-  \label{fig:correctness-over-experience}
-\end{figure}
-
-%% \begin{mh}
-%%   The results of our exploratory data analysis and   hypothesis testing suggest that the atom candidates Comma Operator, Automatic Semicolon Insertion, Post Increment, Omitted Curly Braces, Assignment as Value, and Implicit Predicate %, and Ternary Operator
-%%   introduce some degree of misunderstanding in JavaScript code. %Based on the results of this study, 
-%%   Therefore, they can be deemed atoms of confusion for JavaScript.
-%% \end{mh}
-
-%% \begin{mh}
-%% {\color{red}Comparing the results of different studies,
-%%   the impact of atoms candidates on source code
-%%   (mis)understanding differ according to the programming
-%%   language.}
-%% \end{mh}
-%% Regarding the first question we 
-%% address in the survey (\emph{Do code snippets that contain atoms of confusion produce a higher error rate than snippets where the atom is removed?}), we found evidence that the atoms of confusion lead programmers to misunderstand JavaScript code.
-We also realized that just one atom whose correction has a non-significant improvement in the percentage of correct answers---we found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of ten atoms we consider in the survey).
-
-\subsection{Time Analysis}
-
-\subsubsection*{Exploratory Data Analysis}
-Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both obfuscated and clean versions.
-We do not consider cases where the participants made mistakes.
-Seven atom candidates required more time from participants to predict a correct response.
-For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet.
-In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
-time on average to find the correct answer for the clean version of the code snippets. 
-Not all atom candidates, though, require less time for the participants to predict the answer.
-In fact, for  Arithmetic as Logic and Pre-Increment, the time to give a correct answer for the clean versions was
-29.06\% and 38.19\% than for the obfuscated versions---we did not confirm these atom candidates as atoms of confusion in the previous section.
-
-\begin{table}[htbp]
-\caption{Time in seconds to submit a correct answer}
-\centering{
-  {\scriptsize
-  \begin{scriptsize}
-\begin{tabular}{lrrr} \toprule
-Atom & Obfuscated  & Clean & $\Delta$(\%) \\ \midrule
-Comma Operator & 87.67 & 20.84 & -76.23 \\ 
-Automatic Semicolon Insertion & 46.08 & 22.04 & -52.17 \\ 
-Post-Increment & 28.70 & 25.67 & -10.56 \\ 
-Omitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
-Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
-Implicit Predicate & 36.24 & 24.01 & -33.75 \\ 
-Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
-Conditional Operator & 41.80 & 39.34 & -5.90 \\ 
-Pre-Increment & 30.71 & 42.45 & 38.19 \\  
-Arithmetic as Logic & 28.82 & 37.20 & 29.06 \\ \bottomrule
-\end{tabular}
-\label{tab:difference-time-taken}
-\end{scriptsize}
-}}
-\end{table}
-
- % latex table generated in R 4.0.4 by xtable 1.8-4 package
- % Thu Apr 22 07:58:21 2021
- \begin{table*}[ht]
-\caption{Hypotheses Testing (misunderstanding and time). Asterisks ($^{*}$) indicate a statistically significant difference.}
- \centering
- {\scriptsize
- \begin{tabular}{lrrr|rr}
-   \toprule
-Atom  & Chi-square & Odds          & Confidence & Mann-Whitey& Cliff's\\ 
- &  test &Ratio          & Interval &  U test & Delta \\  \midrule
-Comma Operator & \textbf{$<$ 0.0001*} & 19.02 & (6.60, 68.22) & \textbf{$<$ 0.0001*} & -0.53 \\ 
-Automatic Semicolon Insertion & \textbf{$<$ 0.0001*} & 39.33 & (9.21, 356.62) & 0.0980 & -0.16 \\ 
-Post-Increment & \textbf{0.0015*} & 4.83 & (1.73, 15.74) & 0.1201 & 0.15 \\ 
-Omitted Curly Braces & 0.0510 & 2.35 & (1.00, 5.77) & 0.9253 & -0.01 \\ 
-Assignment as Value & \textbf{0.0035*} & 8.39 & (1.81, 79.12) & 0.8268 & 0.02 \\ 
-Implicit Predicate & \textbf{0.0112*} & 6.95 & (1.46, 66.56) & \textbf{0.0029*} & -0.29 \\ 
-Logic as Control Flow & 0.2920 & 1.54 & (0.73, 3.28) & \textbf{$<$ 0.0001*} & -0.39 \\ 
-Conditional Operator & 0.1590 & 2.73 & (0.74, 12.57) & 0.2407 & -0.12 \\ 
-Pre-Increment & 0.7015 & 1.25 & (0.55, 2.85) & \textbf{0.0062*} & 0.27 \\ 
-Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & \textbf{0.0172*} & 0.23 \\ 
-    \bottomrule
- \end{tabular}}
- \label{tab:hypothesis-testing}
- \end{table*}
-
-\subsubsection*{Statistical analysis}
-We use the \emph{Mann-Whitney U test} to
-investigate the null hypothesis that developers 
-spend the same amount of time to correctly
-predict the outcome of clean and obfuscated versions of a code snippet.
-%, regardless of evaluating the confusing or clean version of the code.
-The results of Table~\ref{tab:hypothesis-testing}
-suggest that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
-For the first three, the analysis suggests that the participants need more time to predict the outcome of the code snippet in the obfuscated code version.
-For the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time to predict the outcome of the code snippets in the obfuscated version.
-We also computed the effect size using Cliff's Delta statistic.
-In the table, negative effect sizes suggest that it took less time to correctly predict the output of clean versions of the snippets.
-Considering the thresholds established by Romano et al.~\cite{Romano:2006:ASO} (0.147 $<$ $|$d$|$ $<$ 0.33 small, $|$d$|$ $<$ 0.474 medium, otherwise large), we found a large effect for the Comma Operator atom candidate; a medium effect size for Logic as Control Flow, and small effect sizes for Automatic Semicolon Insertion, Implicit Predicate, Arithmetic as Logic, Post-, and Pre-Increment.
-
-%% \begin{mh}
-%%   The results suggest that developers tend to spend
-%%   more time to predict the correct answer in the presence of the atom candidates Comma Operator, Logic as Control Flow, and Implicit Predicate.
-Conversely, the code snippets with the atom candidates Arithmetic as Logic and Pre-Increment demand less time to predict the correct answers.
-%% \end{mh}
-
-%%  Comma Operator        & 60 & 21  & -65  \\
-%% Logic as Control Flow & 85 & 49  & -42  \\
-%% Implicit Predicate    & 33 & 24  & -27  \\
-%% Omitted Curly Braces  & 43 & 31  & -27  \\
-%% Assignment as Value   & 53 & 49  & -7   \\
-%% Ternary Operator      & 42 & 42  & 0    \\
-%% Post Increment        & 27 & 27  & 0    \\
-%% Arithmetic as Logic   & 29 & 36  & +24  \\
-%% Pre-Increment         & 34 & 49  & +44  \\
-
-% \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}
-
-\section{Results of the Interview Study}
-\label{sec:interview-results}
+\subsection{Results}\label{sec:interview-results}
 
@@
 
-%% \rb{acho que essa separa\c c\~{a}o em dois rounds est\'{a}
-%%   gerando uma complexidade desnecess\'{a}ria. vou tentar
-%%   remover isso}.
-
-%In this section, we present the results of the 
-% interviews with the 15 practitioners who agreed to participate in the study. 
 A total of 15 practitioners took part of the interviews.
@@
 We did not indicate in any way whether any version was assumed to be confusing or not.
-% ask the participants to discuss their preferences (i.e., code with or without atom).
-
-%% In the second round of interviews, ten of the fifteen developers
-%% who participated in the first round were shown another 8 atom candidates to atom of confusion whose frequency we observed during the repository mining phase.
-The same procedure of showing code that behaves exactly the same and surveying for how easy to understand each snippet was applied in this phase.
 
@@
 
-\subsubsection*{The Participants' perceptions of the atom candidates}
+\subsubsection{The Participants' perceptions of the atom candidates}
 
@@
 
-\subsubsection*{Confusion in JavaScript Code}
+\subsubsection{Confusion in JavaScript Code}
 
@@
 \end{figure}
-
-% citar uninitialized objects' types. floating point precision issues, 'this', type coercion
-
-% \subsection{Common Misunderstandings About the Language}
-
-% TODO: I don't see that our results support these claims.
-% Let us comment for now.
-
-%% Six practitioners regarded the JavaScript language as easy to learn, while four
-%% participants mentioned that the abundance of frameworks,
-%% coupled with the abstractions they provide, make it so that one
-%% can develop quickly in the language. However, as mentioned by one
-%% of the interviewees, ``\emph{Developers who have started
-%% programming in JavaScript might miss out on important lower-level
-%% aspects, such as garbage collection, memory management,
-%% and process management.}''.
-
-%% Moreover, such abstractions might lead a developer to start to program
-%% using the language without actually understanding some of its basic aspects.
-%% For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
-%% For the multi-purpose language that is one of the most, if not the most used one by practitioners,
-%% a lack of understanding of the languages core
-%% features may be a problem worth investigating in future studies.
-
-%% \begin{table}[!htb]
-%%     \centering
-%%     \caption{Interviews Round 2 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}).
-Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
-%%     \label{tab:interview-results2}
-%%     \begin{tabular}{lrrr}\toprule
-%%       & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
-%%       \cmidrule(lr){2-4}
-%%          \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
-%%                                       &  \multicolumn{1}{c}{\textsc{without aoc}}
-%%                                                & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
-%%          Property access         & 10  & 60  & 30  \\
-%%          Object spread           & 50 & 50     & 0    \\
-%%          Array spread            & 60  & 40    & 0     \\
-%%          Arrow function          & 40 & 40     & 20  \\
-%%          Array destructuring     & 40 & 40  & 20  \\
-%%          Object destructuring    & 40 & 40     & 20    \\
-%%          Type conversion         & 30  & 60    & 10     \\
-%%          Change Literal encoding & 40 & 50  & 10  \\ \midrule
-%%          \textsc{overall}        & 42.8 & 47.6  & 9.6  \\
-%%          \bottomrule
-%%     \end{tabular}
-%% \end{table}
-
-\section{Results of the MSR Effort}
-\label{sec:msr-results}
-
-In this section we present the results of the  software repository mining effort, whose goal is to reveal how prevalent atom candidates are in open source JavaScript projects. 
-We mined \minedprojects open source JavaScript repositories to understand how
-often atoms arise in real software. As in the interview study, this study focused on atoms that were investigated in both experiments.
-
-Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19} which investigate
-the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
-frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
-occur in at least 80\% of the projects. Considering the extremes, atom candidates Conditional Operator and Implicit Predicate were found in all repositories we mined,
-while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
-
-% Figure~\ref{fig:rate} summarizes this finding, showing that the occurrence of atoms of confusion in JavaScript systems range from 11.34\% (Comma Operator) to 89.69\% (Ternary Operator).
-
-%Figure~\ref{fig:atoms-occurrence} shows that the occurrence of in JavaScript systems range from
-
-% latex table generated in R 4.0.3 by xtable 1.8-4 package
-% Fri Nov 13 11:28:56 2020
-%%\begin{table}[!htb]
-%%\centering
-%% \caption{Summary of atoms occurrences on our dataset}
-%% 
-%%\setlength\tabcolsep{2pt} % default value: 6pt
-%%\label{tab:occurrences-summary}
-%%\begin{tabular}{lrcc}
-%%
-%%  \hline
-%%Atom & Projects & Occurrences/KLOC & Category \\ 
-%% \hline
-%%Ternary Operator & 100\% & 10.16 &  highly used \\ 
-%%Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
-%%Post Increment & 90.28\% & 5.62 &  highly used \\ 
-%%Pre-Increment & 83.33\% & 1.04 & commonly used \\ 
-%%Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
-%%Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
-%%Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
-%%Comma Operator & 15.28\% & 0.03 & little used \\ 
-%%   \hline
-%%\end{tabular}
-%%\end{table}
-
-% latex table generated in R 4.0.3 by xtable 1.8-4 package
-% Tue Apr 13 09:34:35 2021
-\begin{table}[ht]
-  \centering
-  {\scriptsize 
-\caption{Summary of atom candidate occurrences in our dataset.}
-\setlength\tabcolsep{2pt} % default value: 6pt
-\begin{tabular}{lrr}%c}
-  \toprule
-Atom & Projects & Occurr./KLOC \\%& Frequency \\ 
-  \midrule
-Implicit Predicate & 100\% & 19.89 \\%&  intensively used \\ 
-  Conditional Operator & 100\% & 10.16 \\%&  intensively used \\ 
-  Omitted Curly Braces & 91.67\% & 6.61 \\%&  intensively used \\ 
-  Post Increment & 90.28\% & 5.62 \\%&  intensively used \\ 
-  Pre Increment & 83.33\% & 1.04 \\%& commonly used \\ 
-  Assignment as Value & 81.94\% & 1.02 \\%& commonly used \\ 
-  Logic as Control Flow & 56.94\% & 0.95 \\%& commonly used \\ 
-  Automatic Semicolon Insertion & 33.33\% & 0.13 \\%& commonly used \\ 
-  Arithmetic as Logic & 23.61\% & 0.02 \\%& rarely used \\ 
-  Comma Operator & 15.28\% & 0.03 \\%& rarely used \\
-   \bottomrule
-\end{tabular}
-  }
-\label{tab:occurrences-summary}
-\end{table}
-
-Considering all JavaScript projects in our dataset, we found a total of \num{364873} atom candidates, though four atoms are responsible
-for 92.97\% of this total: Implicit Predicate, Post Increment, Conditional Operator, and Omitted Curly Braces.
-The first two, based on the results of Section~\ref{sec:latin} and Section~\ref{sec:repeated}, can be considered atoms of confusion.
-The remaining atom
-candidates comprise \num{25621} occurrences combined, 7.03\% of the total number of occurrences.
-The Comma Operator and Arithmetic as Logic atoms are the ones that arise less
-frequently (0.001\% in total with 232 and 165 occurrences, respectively).
-
-% Regarding frequency, we proceeded with a classification of \emph{atom occurrence}
-% based on thresholds used in a previous
-% study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. 
-% Atoms that occur 1000 times are classified as \textit{rarely used} while atoms with more than \num{10000} occurrences are classified as \textit{highly used}; \textit{commonly used} atoms occur in between.
-We calculate the frequency of each atom per KLOC.
-The four atoms mentioned in the previous paragraph,
-Implicit Predicate, Post Increment, Conditional Operator, and Omitted Curly Braces, occur frequently in the analyzed projects.
-All of them have more than 5 occurrences per KLOC on average.
-In particular, two atom candidates occur very frequently, Implicit Predicate (19.89 occurrences per KLOC) and Conditional Operator (10.16 occurrences per KLOC).
-On the other hand, Arithmetic as Logic and Comma Operator occur less than once for every 50 KLOC.
-
-%In our dataset of JavaScript projects, 
-%We found the Ternary Operator atom occurring more frequently than the Omitted Curly Braces, differently from what has been reported in a previous work~\cite{DBLP:conf/msr/GopsteinZFC18}. %
-
-The results of our experiments and interviews suggest that Conditional Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to the Latin-square study).
-Nonetheless, the Post-Increment Expression and Automatic Semicolon Insertion atoms are listed in the top four sources of misunderstanding  (Table~\ref{tab:difference-correctness}), and also frequently appear in open source systems.
-As such, refactoring existing systems to avoid Post-Increment and Automatic Semicolon Insertion might improve the readability of large amounts of code in JavaScript projects.

### 2022-12-02 s04.tex
changing the paper organization

--- 2022-07-27 4_Result.tex
@@
-%% \section{Results and Discussion}
-%% \label{sec:results}
+\section{Prevalence Study}\label{sec:method:msr}
 
-%% In this section we present the main findings of our research.
-We first present the results of each study (Section~\ref{sec:survey-resuts}, Sections~\ref{sec:msr-results}, and Section~\ref{sec:interview-results}).
-After that, in Section~\ref{sec:discussion} we consolidate our findings and present some implications of our study.
+To understand how often the analyzed atom candidates appear in open source projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mine a set of GitHub open source repositories and compute the prevalence of atom candidates.
 
-%\section{Results of the Survey}
-%\label{sec:survey-resuts}
+\subsection{Study Settings}
 
-%-----------------------------------------------------------------------
-\section{Results of the Repeated Measures Study}\label{sec:repeated}
+We first collect the most popular GitHub repositories that are primarily written in JavaScript.
+We measure popularity using the project's stargazers.
+This metric, available through the GitHub API, represents the number of stars a project received from users of the platform.
+The same metric has been used in a number of previous studies as a proxy to estimate a project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}.
+We then select the top 100 most popular repositories and remove projects that do not reach the first quartile of the distribution of lines of code.
 
-In this experiment, we investigate whether 24 atom candidates impact readability.
-We examine how the performance of the participants differ when predicting the outputs of the clean and obfuscated versions of small code snippets (Section~\ref{sec:repeated:correct}).
-Based on those results we establish which candidates can actually be considered atoms of confusion.
-In addition, similarly to Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, we verify whether there is a trade-off between correctness and time irrespective of the presence of atoms; we check if participants who correctly predict the outputs of the programs tend to do so more slowly than the ones that make incorrect predictions (Section~\ref{sec:repeated:time}).
+After filtering out JavaScript project candidates, we build a curated dataset comprising the top \minedprojects repositories.
+Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}.
+Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research.
+The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
+All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars.
+We automate all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.
 
-This study had 70 participants. Out of these, 67 have received university-level training in programming, 4 hold a PhD degree, 7 hold a master's degree, and 11 hold a bachelor's degree. 
-The median programming experience of the participants is four years (mean 6.5 years) and the median experience with JavaScript programming is one year (mean 2.5 years). 
-% General programming: median: 48, 1st quartile: 22,5, 3rd quartile: 91,5
-% JavaScript programming: median: 12, 1st quartile: 4, 3rd: 24
+We mine the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}.
+CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript).
+We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study).
+Finally, we compute some descriptive statistics to measure the prevalence of atoms of confusion in practice.
 
-%-----------------------------------------------------------------------
-\subsection{Correctness Analysis}\label{sec:repeated:correct}
+\begin{table}[ht]
+  \centering
+   \caption{Some descriptive statistics about the projects used in the MSR study}
+  \begin{scriptsize}
+ \begin{tabular}{rrrrr}
+   \toprule
+                       & Min.             & Median        & Mean             & Max. \\ \midrule
+ Lines of Code         & \num{5543}       & \num{36161.5} & \num{111432.33}  & \num{1278405} \\
+ Num. of Forks         & \num{1244}       & \num{6078}    & \num{8906.24}    & \num{68849} \\
+ Num. of Contributors  & \num{6}          & \num{285.5}   & \num{533.44}     & \num{4047} \\
+ Num. of Stars         & \num{23672}      & \num{34990}   & \num{46919.51}   & \num{310935} \\
 
-\subsubsection*{Exploratory Data Analysis}
-For each of the 24 atom candidates, the participants predicted the outputs of two versions, one clean (control) and one obfuscated (treatment), as per the definition of a repeated measures design~\cite{Hunter-Experimenters}.
-Thus, each participant examined 48 code snippets.
-We also measured the time required by the participants to provide their answers.
-For most cases, participants made more mistakes when attempting to predict the outcomes of the obfuscated versions.
-
-Table~\ref{tab:diff-correctness-repeated} presents the number of correct predictions for the clean and obfuscated versions of each atom candidate, as well as the percentage difference.
-In only three cases the participants predicted results correctly more often for the obfuscated versions, Arithmetic as Logic, Logic as Control Flow, and Property Access.
-The former two were empirically validated as atoms of confusion in the study of Gopstein and colleagues~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
-For seven of the atom candidates, the number of correct predictions for the clean versions is more than twice the corresponding number for the obfuscated versions.
-Some of these atoms are common programming patterns in JavaScript programs, such as pre- and post-increments and object destructuring.
-The difference was even greater for Type Conversion (+440\%) and Change of Literal Encoding (+300\%).
-
-\begin{table}[t!]
-\caption{Summary of the correctness analysis.}
-\label{tab:diff-correctness-repeated}
-\centering{
-  \begin{scriptsize}
-\begin{tabular}{lrrr} \toprule
-  Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
-Type Conversion           &            5 &      27 &                 +440 \\
-Change of Literal Encoding &           9 &      36 &                 +300 \\
-Comma Operator            &            8 &      29 &                 +262 \\
-Object Destructuring      &           15 &      42 &                 +180 \\
-Assignment As Value       &           18 &      43 &                 +139 \\
-Pre-Increment             &           13 &      28 &                 +115 \\
-Post-Increment            &            8 &      17 &                 +112 \\
-Array Destructuring       &           15 &      28 &                 +87 \\
-Lack of Indentation no Braces     &           30 &      50 &                 +67 \\
-Omitted Curly Braces      &           35 &      51 &                 +46 \\
-Object Spread             &           21 &      30 &                 +43 \\
-Array Spread              &           22 &      31 &                 +41 \\
-Automatic Semicolon Insertion &       15 &      21 &                 +40 \\
-Infix Operator Precedence &           47 &      53 &                 +13 \\
-Arrow Function            &           54 &      58 &                  +7 \\
-Implicit Predicate        &           41 &      44 &                  +7 \\
-Lack of Indentation With Braces   &           53 &      55 &                  +4 \\
-Conditional Operator      &           60 &      62 &                  +3 \\
-Constant Variables        &           59 &      61 &                  +3 \\
-Dead Unreachable Repeated &           63 &      64 &                  +2 \\
-Repurposed Variables      &           23 &      23 &                   0 \\
-Arithmetic as Logic       &           52 &      47 &                 -10 \\
-Logic as Control Flow     &           33 &      28 &                 -15 \\
-Property Access           &           51 &      42 &                 -18 \\
     \bottomrule
@@
  \end{scriptsize}
-}
+ \label{tab:projects-statistics} 
 \end{table}
 
-\subsubsection*{Statistical analysis}
-
-Since this study uses a within-subject design, the data is paired and dependent.
-Our goal is to verify whether there is a significant difference in the performance of the same participants when examining clean and obfuscated versions of programs.
-The analyzed data is also dichotomous, since each participant either correctly predicts the output of a program or does not.
-Considering this scenario, we employ McNemar's test~\cite{McNemar:1947:NSE}, which is aimed at \textit{``judging the significance of the difference between correlated proportions''}.
-
-The ``McNemar test'' column of Table~\ref{tab:repeated:misunderstanding} indicates the p-values for the tests.
-These p-values are compared against an alpha of 0.0021, after applying Bonferroni correction (0.05/24).
-One of the atom candidates, Constant Variables, is omitted from the table because, due to the very low number of mistakes, it is not possible to perform the test.
-The table shows that eight of the 24 analyzed atom candidates can be considered atoms of confusion based on the results of our study, Object Destructuring, Type Conversion, Change of Literal Encoding, Comma Operator, Lack of Indentation No Braces, Assignment as Value, Pre-Increment, and Ommitted Curly Braces.
-Among these eight, only Object Destructuring is JavaScript-specific.
-Six of the other atoms have been previously reported for the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and six for Java~\cite{Langhout:2021:ACJ}.
-% Lack of Indentation No Braces is not specific to JavaScript and has also been studied by Langhout and Aniche~\cite{Langhout:2021:ACJ}. 
-For some candidates that have been previously identified as atoms, there was not a statistically significant difference.
-This is the case for Repurposed Variables, Logic as Control Flow, Implicit Predicate, and Conditional Operator.
-We examine this in Section~\ref{sec:discussion}.
-
-We leverage the odds ratio (OR) as a measure of effect size, to quantify the meaningfulness~\cite{Ellis:2010:EGE} of these results.
-The odds ratio is an appropriate measure of effect size for scenarios where McNemar's test is applicable~\cite{Mangiafico:2016:SAE}.
-For example, the odds ratio for Type Conversion is 12.
-This suggests that the odds of a participant correctly predicting the output of a clean version while incorrectly predicting the output of the corresponding obfuscated version is 12 times higher than the odds of correctly predicting the output of the obfuscated version and incorrectly predicting the output of the correct one.
-
-According to the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), for most of the atoms where there is a statistically significant difference the odds ratio can be considered high.
-The value of the odds ratio for Object Destructuring is $\infty$ because no participant correctly guessed the output for the obfuscated version while incorrectly guessing the output for the clean one, which results in a denominator of 0.
-Omitted Curly Braces exhibits an OR of 6.3, which is borderline between medium and high~\cite{Chen:2010:HBB}.
-Array Destructuring and Post-Increment are two atom candidates that deserve a special note.
-The difference in the performance of participants examining clean and obfuscated versions of these atoms was not statistically significant, after correction.
-Notwithstanding, effect sizes were large and medium, respectively.
-This is a result that hints at practical relevance in spite of the p-values~\cite{Ellis:2010:EGE}.
-Further investigation is required for these two atom candidates.
-
-\begin{table}[th!]
-\caption{Hypotheses Testing (correctness).
-Asterisks ($^{*}$) indicate a statistically significant difference.
-Using Bonferroni correction, a p-value is considered statistically significant if it is lower than 0.0021.}
- \centering
- {\scriptsize 
- \begin{tabular}{lrr}
-   \toprule
-Atom candidate & McNemar test &      Odds Ratio \\
-\midrule
-Object Destructuring      &   \textbf{$<$ 0.0001*} & $\infty$ \\
-Type Conversion           &   \textbf{$<$ 0.0001*} & 12.0 \\
-Change of Literal Encoding   &   \textbf{$<$ 0.0001*} & 28.0 \\
-Comma Operator            &   \textbf{$<$ 0.0001*} & 11.5 \\
-Lack of Indentation No Braces     &   \textbf{$<$ 0.0001*} &  11.0 \\
-Assignment As Value       &   \textbf{$<$ 0.0001*} &  9.3 \\
-Pre-Increment              &  \textbf{0.0003*} &    16.0 \\
-Omitted Curly Braces      &  \textbf{0.0009*} &  6.3 \\
-Array Destructuring       &  0.0023 &     7.5 \\
-Post-Increment              &  0.0225 &     5.5 \\
-Array Spread              &  0.0490 &    3.3 \\
-Object Spread             &  0.0636 &     2.8 \\
-Property Access           &  0.0636 &     2.8 \\
-Infix Operator Precedence &  0.1796 &     2.5 \\
-Arrow Function            &  0.2891 &     3.0 \\
-Arithmetic As Logic       &  0.3323 &  1.8 \\
-Automated Semicolon Insertion  &  0.3915 &  1.4 \\
-Logic As Control Flow     &  0.4421 &  1.5 \\
-%Constant Variables        &  0.5000 &     0.0 \\
-Implicit Predicate        &  0.6900 &  1.3 \\
-Lack of Indentation With Braces   &  0.7266 &  1.7 \\
-Conditional Operator      &  0.7905 &  1.3 \\
-Dead Unreachable Repeated &  1.0000 &     1.5 \\
-Repurposed Variables      &   1.0000 &     1.0 \\
-\bottomrule
-\end{tabular}
-}
-\label{tab:repeated:misunderstanding}
-\end{table}
-
-%-----------------------------------------------------------------------
-\subsection{Time Analysis}\label{sec:repeated:time}
-
-\subsubsection*{Exploratory Data Analysis}
-
-In this section we investigate whether there is a trade-off between correctness and time required to predict program outputs, similarly to the analysis performed by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
-For this analysis, we consider only whether participants correctly predict the outputs of programs or not, independently of the presence or absence of atoms, and the time required to do so.
-The null hypothesis we investigate is that there is no significant difference in time between correct and incorrect predictions.
-Intuitively, we would expect correct responses to take longer, on the basis that participants would spend more time carefully analyzing the programs before providing an answer.
-
-Table~\ref{tab:repeated:time} presents the mean and median times required by the participants to provide answers.
-The data is based on the 1648 correct and 1572 incorrect answers collected in the study.
-Contrary to our intuition, the mean time required by participants to provide an incorrect answer is 25\% greater than the time required to provide a correct one and the median time is 20\% greater.
-The standard deviations suggest that there is great variation in the times required by different participants, for different code snippets.
-Notwithstanding, the standard deviation for incorrect answers is 62.58\% greater when compared to the same statistic for correct answers.
-
-\begin{table}[tb]
-  \caption{Time in seconds required by the participants to correctly and incorrectly predict the outcomes of the programs.}\label{tab:repeated:time}
-  \centering
-  {\scriptsize 
-  \begin{tabular}{lrrr}
-    \toprule  
-             & Mean  & Median & Std. Dev. \\
-    \midrule
-  Correct   & 35.44 & 24.36  & 38.43     \\
-  Incorrect & 44.50 & 29.29  & 62.48    \\
-  \bottomrule
-  \end{tabular}
-}
-  \end{table}
-
-\subsubsection*{Statistical Analysis}
-
-We employ the non-parametric Mann-Whitney U test, also known as Wilcoxon's rank-sum test, to verify the aforementioned null hypothesis, i.e., whether the times for correct and incorrect answers can be considered identical.
-That test produces a p-value of \textbf{0.00000046}, which shows that the difference is statistically significant and we can reject the null hypothesis.
-To gauge the magnitude of that difference we employ Cliff's Delta as a measure of effect size.
-This produces a result of \textbf{0.1027}, which suggests that, although participants tend to provide correct predictions more quickly, they do that only slightly so.
-
-%-----------------------------------------------------------------------
-\section{Results of the Latin-Square Study}\label{sec:latin}
-
-In the latin-square study, we estimate the impact of atoms considering the same two perspectives of the previous experiment: \emph{correctness} (number of wrong answers)
-and \emph{time} (how long to provide a correct answer).
-In particular, in the time analysis, differently from the repeated  measures study, we compare the time required by participants to submit a correct response when analyzing code snippets with and without atom candidates.
-This study is based on the responses of 140 participants (a total of 70 replicas).
-All participants had taken at least some university course or hold a bachelor degree or equivalent.
-In addition, 21 participants hold a master's degree and three a doctorate degree.
-Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.
-Accordingly, we characterize the effect of atoms of confusion in JavaScript code taking into account the perceptions of both novice and experienced developers.
-
-\subsection{Correctness Analysis}
-% \castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me to which rate this refers.}
-
-\subsubsection*{Exploratory Data Analysis}
-
-As mentioned in Section~\ref{sec:meth:survey}, each participant in this experiment evaluated ten code snippets, from which five were in their obfuscated versions, whilst the other five contained clean versions of the code snippets (i.e., without the atom candidates).
-As in the repeated measures study, the participants should provide the expected outcomes of the code snippets.
-Differently from that study, in this one each participant only predicts the outcome of one version of each code snippet, either clean or obfuscated.
-As discussed, we collect information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{time}.
-Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the correctness experiment.
-
-Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the obfuscated versions.
-In particular, the presence of the \emph{Comma Operator} atom exhibits the highest impact on misunderstanding.
-Frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also result in many mistakes.
-%introduce high degrees of confusion.
-The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the clean versions of the code snippets.
-Also, the sample of responses with no atoms had almost no dispersion, which supports the argument that the clean versions of the code snippets are easier to evaluate correctly.
-
-% \begin{table}[htbp]
-% \caption{Difference in answer correctness between confusing and non-confusing pairs}
-% \begin{center}
-% \begin{small}
-% \begin{tabularx}
-% {{\linewidth}}{l p{1.5cm} p{1.1cm} p{1.1cm} p{1.2cm} }
-% \textbf{Atom} & \textbf{\%Correct} & \textbf{\%Correct} \\
-% &  \multicolumn{1}{l}{With AOC} \multicolumn{2}{l}{Without AOC}  & \Delta (\%)    \\
-%  \hline
-% Comma Operator & 40 & 93 & +132\%                 \\     
-% Automatic Semicolon  Insertion & 46 & 97 & +110\% \\
-% Post Increment & 69 & 91 & +31\%         \\
-% Omitted Curly Braces & 67 & 83 & +23\%   \\ 
-% Assignment as Value & 80 & 97 & +21\%    \\
-% Implicit Predicate & 83 & 97 & +16\%     \\
-% Logic as Control Flow & 59 & 68 & +15\%  \\
-% Ternary Operator & 86 & 94 & +9\%        \\
-% Pre-Increment & 71 & 76 & +7\%           \\
-% Arithmetic as Logic & 91 & 90 & -1\%    \\
-% \end{tabularx}
-% \end{small}
-% \end{center}
-% \label{results_correctness}
-% \end{table}
-
-\begin{table}[htbp]
-\caption{Summary of the correctness analysis}
-\label{tab:difference-correctness}
-\centering{
-  {\scriptsize
-  \begin{scriptsize}
-\begin{tabular}{lrrr} \toprule
-  Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
-  Comma Operator                &  28 &  65 & +132   \\ 
-  Automatic Semicolon Insertion &  32 &  68 & +112   \\ 
-  Post-Increment                &  48 &  64 & +33    \\ 
-  Omitted Curly Braces          &  47 &  58 &  +23   \\
-  Assignment as Value           &  56 &  68 &  +21   \\ 
-  Implicit Predicate            &  58 &  68 &  +17   \\ 
-  Logic as Control Flow         &  41 &  48 &  +17   \\ 
-  Conditional Operator              &  60 &  66 &  +10   \\ 
-  Pre-Increment                 &  50 &  53 &  +6    \\ 
-  Arithmetic as Logic           &  64 &  63 &  -2    \\ \bottomrule
-
-\end{tabular}
-\end{scriptsize}
-}}
-\end{table}
-
- %% Comma Operator          & 40 & 93  & +132 \\
- %% Post Increment          & 69 & 91  & + 31  \\
- %% Omitted Curly Braces    & 67 & 83  & +23 \\
- %% Assignment as Value     & 80 & 97  & +21 \\
- %% Implicit Predicate      & 83 & 97  & +16 \\
- %% Logic as Control Flow   & 59 & 68  & +15 \\
- %% Ternary Operator        & 86 & 94  & +9  \\
- %% Pre-Increment           & 71 & 76  & +7  \\
- %% Arithmetic as Logic     & 91 & 90  & +1  \\ \bottomrule
-
-\begin{figure}[b!]
-\noindent
- \centering
- \includegraphics[width=0.6\columnwidth]{images/wrong-answers-plot-1.pdf}
- \caption{Number of wrong answers of each subject.}
- \label{fig:boxplotcorrectness}
- \end{figure}% \castor{Fiz de tudo para ajeitar a figura porque o texto embaixo deveria ser "Clean versions" e "Confusing versions" mas não sei porque não funcionou.
-Até mudei a figura.
-Não sei o que está havendo.}
-
-% \rb{verificar a consistencia nas questoes de pesquisa mais especificas}
-
-\subsubsection*{Statistical analysis}
-
-We first use the \emph{Pearson's Chi-squared test}
-to investigate if there is a statistically significant difference in the frequency of correct and incorrect answers---due to the versions of the code snippets (obfuscated and clean code).
-The p-values for these tests are reported in the ``Chi-square test'' column of Table~\ref{tab:hypothesis-testing}.
-The results indicate that for five atom candidates (Comma Operator, Automatic Semicolon Insertion, Post-Increment, Assignment as Value, Implicit Predicate) the obfuscated versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$).
-This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
-
-We measure the effect size of the clean version of the code snippets into the answers' correctness using the \emph{Odds Ratio} (OR).
-The results are reported in the ``Odds Ratio'' column of Table~\ref{tab:hypothesis-testing}.
-In the table, we also report the confidence interval (CI) for the OR in the ``CI'' column.
-Although many of the intervals are wide, for six of the atom candidates the lower bound of the confidence interval is greater than or equal to 1.
-This indicates that, at a 95\% confidence level, a developer is likely to commit less mistakes when using the clean versions.
-%Considering only the atoms for which there was a significant difference in correctness for confusing and clean versions and u
-%Using the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), 
-For four of the atoms, Comma Operator, Automatic Semicolon Insertion, Assignment as Value, and Implicit Predicate, effect size can be considered large. For Implicit Predicate, it is medium. 
-%We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. 
-For instance, when comparing clean and obfuscated code snippets pertaining to the Comma Operator atom candidate, we observed an \emph{Odds Ratio} of \num{19.02}.
-This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding obfuscated version of the code snippet.
-Furthermore, with a 95\% confidence level, the odds ratio is between 6.6 and 68.22, i.e., although the error margin is wide, it is strongly in favor of the clean version.
-If there is a significant likelihood of a developer committing a mistake when analyzing a clean version, the lower bound of the CI should be a number between 0 and 1 (since the OR is a ratio).
-This is the case for the four atom candidates in the lower part of the table.
-%In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a participant was actually more likely to misunderstand the clean version.
-
-Finally, we also use a \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impacts correctness.
-The findings suggest that \emph{participant experience} impacts the results related to correctness (Figure~\ref{fig:correctness-over-experience}).
-Even though the presence of atoms of confusion reduces the number of correct answers for all \emph{experience groups}, this effect is not uniform.
-For instance, novice developers (under one year of experience) provide a higher number of wrong answers for the obfuscated version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the obfuscated version of the code snippets.
-It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.
-We replicate the \emph{Pearson's Chi-squared test} for each experience group and find that the statistical difference is less significant for those developers with more than ten years of experience.
-Differently from \emph{participant experience}, the factor \emph{participant education} does not significantly change the correctness of the answers.
-\castor{I'd like to have the numbers for the results above. What are the p-values, regression coefficients, etc.?}
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.3]{images/correctness-by-experience}
-  \caption{Impact of atoms of confusion on the correctness (over the four participant experience groups)}
-  \label{fig:correctness-over-experience}
-\end{figure}
-
-%% \begin{mh}
-%%   The results of our exploratory data analysis and   hypothesis testing suggest that the atom candidates Comma Operator, Automatic Semicolon Insertion, Post Increment, Omitted Curly Braces, Assignment as Value, and Implicit Predicate %, and Ternary Operator
-%%   introduce some degree of misunderstanding in JavaScript code. %Based on the results of this study, 
-%%   Therefore, they can be deemed atoms of confusion for JavaScript.
-%% \end{mh}
-
-%% \begin{mh}
-%% {\color{red}Comparing the results of different studies,
-%%   the impact of atoms candidates on source code
-%%   (mis)understanding differ according to the programming
-%%   language.}
-%% \end{mh}
-%% Regarding the first question we 
-%% address in the survey (\emph{Do code snippets that contain atoms of confusion produce a higher error rate than snippets where the atom is removed?}), we found evidence that the atoms of confusion lead programmers to misunderstand JavaScript code.
-We also realized that just one atom whose correction has a non-significant improvement in the percentage of correct answers---we found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of ten atoms we consider in the survey).
-
-\subsection{Time Analysis}
-
-\subsubsection*{Exploratory Data Analysis}
-Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both obfuscated and clean versions.
-We do not consider cases where the participants made mistakes.
-Seven atom candidates required more time from participants to predict a correct response.
-For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet.
-In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
-time on average to find the correct answer for the clean version of the code snippets. 
-Not all atom candidates, though, require less time for the participants to predict the answer.
-In fact, for  Arithmetic as Logic and Pre-Increment, the time to give a correct answer for the clean versions was
-29.06\% and 38.19\% than for the obfuscated versions---we did not confirm these atom candidates as atoms of confusion in the previous section.
-
-\begin{table}[htbp]
-\caption{Time in seconds to submit a correct answer}
-\centering{
-  {\scriptsize
-  \begin{scriptsize}
-\begin{tabular}{lrrr} \toprule
-Atom & Obfuscated  & Clean & $\Delta$(\%) \\ \midrule
-Comma Operator & 87.67 & 20.84 & -76.23 \\ 
-Automatic Semicolon Insertion & 46.08 & 22.04 & -52.17 \\ 
-Post-Increment & 28.70 & 25.67 & -10.56 \\ 
-Omitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
-Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
-Implicit Predicate & 36.24 & 24.01 & -33.75 \\ 
-Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
-Conditional Operator & 41.80 & 39.34 & -5.90 \\ 
-Pre-Increment & 30.71 & 42.45 & 38.19 \\  
-Arithmetic as Logic & 28.82 & 37.20 & 29.06 \\ \bottomrule
-\end{tabular}
-\label{tab:difference-time-taken}
-\end{scriptsize}
-}}
-\end{table}
-
- % latex table generated in R 4.0.4 by xtable 1.8-4 package
- % Thu Apr 22 07:58:21 2021
- \begin{table*}[ht]
-\caption{Hypotheses Testing (misunderstanding and time). Asterisks ($^{*}$) indicate a statistically significant difference.}
- \centering
- {\scriptsize
- \begin{tabular}{lrrr|rr}
-   \toprule
-Atom  & Chi-square & Odds          & Confidence & Mann-Whitey& Cliff's\\ 
- &  test &Ratio          & Interval &  U test & Delta \\  \midrule
-Comma Operator & \textbf{$<$ 0.0001*} & 19.02 & (6.60, 68.22) & \textbf{$<$ 0.0001*} & -0.53 \\ 
-Automatic Semicolon Insertion & \textbf{$<$ 0.0001*} & 39.33 & (9.21, 356.62) & 0.0980 & -0.16 \\ 
-Post-Increment & \textbf{0.0015*} & 4.83 & (1.73, 15.74) & 0.1201 & 0.15 \\ 
-Omitted Curly Braces & 0.0510 & 2.35 & (1.00, 5.77) & 0.9253 & -0.01 \\ 
-Assignment as Value & \textbf{0.0035*} & 8.39 & (1.81, 79.12) & 0.8268 & 0.02 \\ 
-Implicit Predicate & \textbf{0.0112*} & 6.95 & (1.46, 66.56) & \textbf{0.0029*} & -0.29 \\ 
-Logic as Control Flow & 0.2920 & 1.54 & (0.73, 3.28) & \textbf{$<$ 0.0001*} & -0.39 \\ 
-Conditional Operator & 0.1590 & 2.73 & (0.74, 12.57) & 0.2407 & -0.12 \\ 
-Pre-Increment & 0.7015 & 1.25 & (0.55, 2.85) & \textbf{0.0062*} & 0.27 \\ 
-Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & \textbf{0.0172*} & 0.23 \\ 
-    \bottomrule
- \end{tabular}}
- \label{tab:hypothesis-testing}
- \end{table*}
-
-\subsubsection*{Statistical analysis}
-We use the \emph{Mann-Whitney U test} to
-investigate the null hypothesis that developers 
-spend the same amount of time to correctly
-predict the outcome of clean and obfuscated versions of a code snippet.
-%, regardless of evaluating the confusing or clean version of the code.
-The results of Table~\ref{tab:hypothesis-testing}
-suggest that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.
-For the first three, the analysis suggests that the participants need more time to predict the outcome of the code snippet in the obfuscated code version.
-For the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time to predict the outcome of the code snippets in the obfuscated version.
-We also computed the effect size using Cliff's Delta statistic.
-In the table, negative effect sizes suggest that it took less time to correctly predict the output of clean versions of the snippets.
-Considering the thresholds established by Romano et al.~\cite{Romano:2006:ASO} (0.147 $<$ $|$d$|$ $<$ 0.33 small, $|$d$|$ $<$ 0.474 medium, otherwise large), we found a large effect for the Comma Operator atom candidate; a medium effect size for Logic as Control Flow, and small effect sizes for Automatic Semicolon Insertion, Implicit Predicate, Arithmetic as Logic, Post-, and Pre-Increment.
-
-%% \begin{mh}
-%%   The results suggest that developers tend to spend
-%%   more time to predict the correct answer in the presence of the atom candidates Comma Operator, Logic as Control Flow, and Implicit Predicate.
-Conversely, the code snippets with the atom candidates Arithmetic as Logic and Pre-Increment demand less time to predict the correct answers.
-%% \end{mh}
-
-%%  Comma Operator        & 60 & 21  & -65  \\
-%% Logic as Control Flow & 85 & 49  & -42  \\
-%% Implicit Predicate    & 33 & 24  & -27  \\
-%% Omitted Curly Braces  & 43 & 31  & -27  \\
-%% Assignment as Value   & 53 & 49  & -7   \\
-%% Ternary Operator      & 42 & 42  & 0    \\
-%% Post Increment        & 27 & 27  & 0    \\
-%% Arithmetic as Logic   & 29 & 36  & +24  \\
-%% Pre-Increment         & 34 & 49  & +44  \\
-
-% \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}
-
-\section{Results of the Interview Study}
-\label{sec:interview-results}
-
-In this section we present the results of the interviews with the practitioners.
-We contrast the findings of the experiments of Sections~\ref{sec:repeated} and~\ref{sec:latin} with the opinion of software developers about their preferences regarding the obfuscated or clean versions of the code snippets.
-For the interview study, we only considered atoms that were analyzed in both the repeated measures (Section~\ref{sec:repeated}) and the latin square (Section~\ref{sec:latin}) studies.
-
-%% \rb{acho que essa separa\c c\~{a}o em dois rounds est\'{a}
-%%   gerando uma complexidade desnecess\'{a}ria. vou tentar
-%%   remover isso}.
-
-%In this section, we present the results of the 
-% interviews with the 15 practitioners who agreed to participate in the study. 
-A total of 15 practitioners took part of the interviews.
-We collected information regarding programming
-experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in the both experiments.
-We first presented them pairs of clean and obfuscated code snippets and then asked them to discuss their preferences towards one version.
-We did not indicate in any way whether any version was assumed to be confusing or not.
-% ask the participants to discuss their preferences (i.e., code with or without atom).
-
-%% In the second round of interviews, ten of the fifteen developers
-%% who participated in the first round were shown another 8 atom candidates to atom of confusion whose frequency we observed during the repository mining phase.
-The same procedure of showing code that behaves exactly the same and surveying for how easy to understand each snippet was applied in this phase.
-
-\begin{table}[!htb]
-    \centering
-    {\scriptsize
-    \caption{Summary of participants' preferences for code snippets \emph{with} and
-      \emph{without} atom candidates. The participants were only presented the code snippets, without any indication about whether one was confusing or not. 
-      Participants were also allowed to choose Neutral when they thought both sides were equally readable.}\label{tab:interview-results1}
-    \begin{tabular}{lrrr}\toprule
-      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
-      \cmidrule(lr){2-4}
-         Atom           & \multicolumn{1}{c}{Obfuscated}
-                                      &  \multicolumn{1}{c}{Clean}
-                                               & \multicolumn{1}{c}{Neutral} \\ \midrule
-         Comma Operator                  & 0  & 100    & 0     \\
-         Automatic Semicolon Insertion  & 0  & 80     & 0     \\
-         Post-Increment                  & 20 & 73.33  & 6.67  \\
-         Omitted Curly Braces            & 0  & 100    & 0     \\
-         Assignment as Value             & 20 & 60     & 20    \\
-         Implicit Predicate              & 20 & 73.33  & 6.67  \\
-         Logic as Control Flow           & 20 & 60     & 20    \\
-         Conditional Operator                & 60 & 26.67  & 13.3  \\
-         Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
-         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \midrule
-         \textsc{overall}                & 18 & 71.33  & 8.64  \\
-         \bottomrule
-    \end{tabular}
-    }
-
-\end{table}
-
-\subsubsection*{The Participants' perceptions of the atom candidates}
-
-Supporting the results of Sections~\ref{sec:repeated} and~\ref{sec:latin}, Table~\ref{tab:interview-results1}
- shows that for eight out of the
-10 scenarios surveyed, the majority of the respondents prefer the version of the
-code without the atom candidate.
-In no case the \emph{neutral}
-ratio was higher than the option for the clean version.
-%% An example entry is contained in Appendix~\ref{}. 
-%% For a full listing of the code snippets, visit the paper repository. 
-Only for the Conditional Operator atom candidate the participants preferred the obfuscated version, instead of the clean one.
-Figure~\ref{code:ternary} shows the code snippets for the corresponding obfuscated and clean versions.
-For this atom candidate, some participants who preferred the left-hand side version (obfuscated) still believed that the right-hand side version (clean) was more readable.
-The following quotes were extracted from the transcripts with
-% three interviewees:
-two interviewees:
-
-\begin{figure*}
-
-\noindent\begin{minipage}{.45\textwidth}
-\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Conditional Operator} atom).]
-let config = {size: 3, isActive: false};
-const_config = config.isActive === true 
-             ? config 
-             : {size: 10};
-console.log(_config.size);
-\end{lstlisting}
-\end{minipage}\hfill
-\begin{minipage}{.45\textwidth}
-\begin{lstlisting}[language=JavaScript, caption=\emph{Right-hand side} (without the atom).]
-let config = {size: 3, isActive: false}
-let _config;
-if(config.isActive === true) {
-  _config = config;
-}
-else{
-   _config = {size: 10};
-}
-console.log(_config.size);
-\end{lstlisting}
-\end{minipage}
-\caption{Example of a pair of code snippets used in the interview pertaining to the \emph{Conditional Operator} atom candidate}
-\label{code:ternary}
-\vspace{-0.2cm}
-\end{figure*}
-
-\begin{mq}
-\emph{``I prefer to write [code using the \lhs version], but I think [the \rhs version] is easier to read, especially for newer programmers''}.
-\end{mq}
-
-\begin{mq}
-\emph{``When I am programming, I write code with the conditional operator, [...], but, to be honest, I still think that 
-[the code using the \rhs version] 
-is easier to understand''}.
-\end{mq}
-
-%% \begin{mq}
-%% \emph{``I think [the \lhs version] is easier to understand, but [the \rhs version] is what I would write"}
-%% \end{mq}
-
-%% The atom candidate Ternary Operator also opens up
-%% the possibility for a derivative construct that JavaScript
-%% allows which can be rather confusing, and that is the nested
-%% ternaries construct, in which the right-hand side of the a
-%% ternary operator can be another ternary construct.
-%% While nested ternaries remove the number of lines that
-%% would be necessary to construct using nested if-then-else statements,
-%% they can become quite taxing to understand.
-%% Nested ternaries are a choice of atoms to be analyzed in future work.
-
-The Pre-Increment atom candidate also caused a
-conflict in one of the interviewees, who
-regarded the clean version as simpler to understand, but would still opt to write
-code with the atom. In contrast, %to such opinion,
-one of the participants found the version with the atom candidate more elegant, but recognized it was less readable, and was willing to sacrifice elegance for readability.
-In the repeated measures study, we found a significant difference in favor of the clean version, with a large effect size in its favor.
-In the latin square study, no difference could be observed in terms of correctness.
-However, there was a significant difference in terms of the time required to predict the output correctly in favor of the obfuscated version.
-%% \rb{pela discuss\~{a}o anterior, parece-me que alguns
-%%   desenvolvedores reconhecem que esses candidatos
-%%   a \'{a}tomos introduzem certa dificuldade de compreens\~{a}o;
-%%   mas mesmo assim optam por usar a constru\c c\~{a}o correspondente.
-%%   Acho que cabe um box aqui com essa discuss\~{a}o sobre isso.}
-
-When analyzing the Logic as Control Flow atom candidate, one of the interviewees gave an example of personal experience that might motivate one to avoid writing code using this construct:
-
-\begin{mq}
-  \emph{``This one is interesting, because I have written code that looks like the left-hand side [obfuscated version], and my colleagues complained that it was difficult to understand.
-Nowadays I prefer to write code using the version on the right-hand side [clean version]''}.
-\end{mq}
-\noindent
-None of the two studies found a significant difference in correctness between clean and obfuscated versions of snippets related to this atom candidate.
-However, in the latin square study the participants analyzing clean versions of the snippets required significantly less time.
-
-For two of the atom candidates, the interviewees were unanimous in their preference for the clean version: Comma Operator and Omitted Curly Braces.
-Not coincidentally, in both experiments there were significant differences in correctness in favor of the clean versions for both atoms.
-Regarding the first one, we could often notice during the interviews that the \emph{\lhs} (with the atom of confusion) caused significant confusion among the participants.
-One remark about the \emph{Comma Operator} atom of confusion is listed below:
-
-%% \begin{mq}
-%% \emph{``I just learned that [the \lhs version of this code] is possible. I did not even know it worked''}
-%% \end{mq}
-
-\begin{mq}
-\emph{``The code in [the \lhs version] is unlikely to be understood unless the programmer knows \clang or \cpplang''}.
-\end{mq}
-
-As for the atom candidate Omitted Curly Braces, one of the interviewees mentioned that, although they understand why one would opt not to use braces for simple if-then-else statements, they still advised against it, on grounds that:
-
-\begin{mq}
-\emph{``[I prefer the \rhs version of the code \ldots]
-If I want to see well-written, easily understandable code, then I also have to do my job.
-Therefore I believe that, since I do not know who is on the other end maintaining this code, and it could be any person with any level of expertise, then I try to write readable, easy-to-understand code''}.
-\end{mq}
-
-%This is an interesting perspective, because such a perhaps simple decision,  might hinder novice developers in the task of maintaining code; and one cannot make any assumption about the programmers' experience of who is going to maintain a code in a long run.
-
-%This is an interesting perspective because such a simple decision might hinder novice developers in maintaining code; one cannot assume the programmers' experience of who will maintain a code in the long run.
-
-\subsubsection*{Confusion in JavaScript Code}
-
-% The final remarks that were drawn from the interviews are related to potentially confusing constructs that were suggested by the participants, as well as their perspective on JavaScript as a language, from which we can also uncover some other forms in which the language itself might contribute to writing confusing code.
-
-The final remarks drawn from the interviews are related to potentially confusing constructs suggested by the participants and their perspectives on JavaScript as a language.
-From the latter, we can also uncover some other forms in which the language itself might contribute to writing confusing code.
-
-One of the participants mentioned that the use of JavaScript's prototype-based inheritance can make it difficult to understand code, particularly when involving deep prototype chains.
-%This is a core feature of the language, and most high-level tools and frameworks abstract it away.
-%Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.
-% When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate.
-When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern---potentially leading to several levels of nested function calls---as extremely difficult to assimilate.
-One of the respondents stated:
-
-\begin{mq}
-\emph{``Nested callbacks are very confusing. Even writing them can be confusing, let alone understanding them.''}
-\end{mq}
-
-Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript.
-%Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.
-We also found other idioms that are JavaScript atom candidates, including
-\emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).
-The former was investigated in the repeated measures study but we did not find a significant difference between clean and obfuscated versions.
-As future work, we aim at investigating whether or not these atom candidates are more likely to introduce confusion in JavaScript code.
-
-\begin{figure}
-\begin{small}
-\begin{lstlisting}[language=JavaScript,caption=Example of arrow function.,label=arr1]
-let inc = (x) => x + 1;
-\end{lstlisting}
-\begin{lstlisting}[language=JavaScript,caption=Alternative version without arrow function.,label=arr2]
-let inc = function(x) {
-  return x + 1
-}
-\end{lstlisting}
-\vspace*{-0.6cm}
-\end{small}
-\end{figure}
-
-% citar uninitialized objects' types. floating point precision issues, 'this', type coercion
-
-% \subsection{Common Misunderstandings About the Language}
-
-% TODO: I don't see that our results support these claims.
-% Let us comment for now.
-
-%% Six practitioners regarded the JavaScript language as easy to learn, while four
-%% participants mentioned that the abundance of frameworks,
-%% coupled with the abstractions they provide, make it so that one
-%% can develop quickly in the language. However, as mentioned by one
-%% of the interviewees, ``\emph{Developers who have started
-%% programming in JavaScript might miss out on important lower-level
-%% aspects, such as garbage collection, memory management,
-%% and process management.}''.
-
-%% Moreover, such abstractions might lead a developer to start to program
-%% using the language without actually understanding some of its basic aspects.
-%% For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
-%% For the multi-purpose language that is one of the most, if not the most used one by practitioners,
-%% a lack of understanding of the languages core
-%% features may be a problem worth investigating in future studies.
-
-%% \begin{table}[!htb]
-%%     \centering
-%%     \caption{Interviews Round 2 - Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}).
-Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
-%%     \label{tab:interview-results2}
-%%     \begin{tabular}{lrrr}\toprule
-%%       & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
-%%       \cmidrule(lr){2-4}
-%%          \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
-%%                                       &  \multicolumn{1}{c}{\textsc{without aoc}}
-%%                                                & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
-%%          Property access         & 10  & 60  & 30  \\
-%%          Object spread           & 50 & 50     & 0    \\
-%%          Array spread            & 60  & 40    & 0     \\
-%%          Arrow function          & 40 & 40     & 20  \\
-%%          Array destructuring     & 40 & 40  & 20  \\
-%%          Object destructuring    & 40 & 40     & 20    \\
-%%          Type conversion         & 30  & 60    & 10     \\
-%%          Change Literal encoding & 40 & 50  & 10  \\ \midrule
-%%          \textsc{overall}        & 42.8 & 47.6  & 9.6  \\
-%%          \bottomrule
-%%     \end{tabular}
-%% \end{table}
-
-\section{Results of the MSR Effort}
+\subsection{Results}
 \label{sec:msr-results}
 
-In this section we present the results of the  software repository mining effort, whose goal is to reveal how prevalent atom candidates are in open source JavaScript projects. 
+In this section we present the results of the  software repository
+mining effort, whose goal is to reveal how prevalent atom candidates are in open source JavaScript projects. 
 We mined \minedprojects open source JavaScript repositories to understand how
@@
 \caption{Summary of atom candidate occurrences in our dataset.}
+\label{tab:occurrences-summary}
 \setlength\tabcolsep{2pt} % default value: 6pt
@@
   }
-\label{tab:occurrences-summary}
 \end{table}
@@
 for 92.97\% of this total: Implicit Predicate, Post Increment, Conditional Operator, and Omitted Curly Braces.
-The first two, based on the results of Section~\ref{sec:latin} and Section~\ref{sec:repeated}, can be considered atoms of confusion.
+The first two, based on the results of Section~\ref{sec:repeated} and Section~\ref{sec:latin}, can be considered atoms of confusion.
 The remaining atom

### 2022-12-02 methodology.tex
presenting an overview of the studies

@@
-\section{Research Methods}
+\section{Research Method Overview}
 \label{method}
@@
 
-By answering these research questions, we can either generalize or refute the findings about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). 
+By answering these research questions, we can either generalize or refute the findings about atoms of confusion already discussed in the literature
+(goal of research questions (RQ1) and (RQ2)). 
 In addition, answering the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice.
@@
 %---though we postpone these results to future research work.
+
 We conduct a mixed-methods study to answer these questions. 
-It includes two independently designed and conducted experiments: a repeated measures/within subjects study (Section~\ref{sec:meth:replication}) and one using a Latin square, between subjects, counterbalancing experimental design to control for subject experience (Section~\ref{sec:meth:survey}).
-By using two different experiments, designed by different researchers, with different designs, and conducted with different participants, we hope to rigorously verify the impact of atoms of confusion in code comprehension activities.
+It includes two independently designed and conducted experiments: a repeated measures/within subjects study (Section~\ref{sec:s01}) and one using a Latin square, between subjects, counterbalancing experimental design to control for subject experience (Section~\ref{sec:s02}).
+By using two different experiments, conceived by different researchers, with different designs, and conducted with different participants, we hope to rigorously verify the impact of atoms of confusion in code comprehension activities.
 \revised{Conducting multiple studies, with different designs and samples, is a typical approach in other areas of science to obtain more solid evidence and also to contest evidence obtained by single study, for example, Weggemans et al.~\cite{Weggemans:2001:DCE}.
@@
 They investigate whether or not programs that contain atom candidates tend to produce more misunderstanding for programmers trying to predict their outcome.% We leverage two independently-developed and conducted experiments with different designs to more
-We also conduct a set of interviews with developers (Section~\ref{sec:method:interview}) to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2).
-Finally, we perform a repository mining study (Section~\ref{sec:method:msr}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
+We also conduct a set of interviews with developers (Section~\ref{sec:s03}) to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2).
+Finally, we perform a repository mining study (Section~\ref{sec:s04}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
+
+Table~\ref{tab:atomsBoth} presents the atom candidates that we explore in all stduies, while Table~\ref{tab:atomsRepeated} presents the additional 
+ones that only the repeated measures study analyzes.
+
+\input{atom_descriptions}

### 2022-12-02 s01.tex
presenting an overview of the studies

@@
-\section{Repeated Measures Study}
+\section{Repeated Measures Study}\label{sec:s01}
 
@@
 Furthermore, we conduct an informal search of \revised{JavaScript questions on StackOverflow about constructs that are not typical of other languages, such as array and object spread, solutions to Code Golf~\footnote{https://codegolf.stackexchange.com/} problems, and code style guides\footnote{AirBNB JavaScript Style Guide: https://github.com/airbnb/javascript}\footnote{Google JavaScript Style Guide: https://google.github.io/styleguide/jsguide.html} to identify new atom candidates.
-We leverage previous work~\cite{castor2018} on the identification of atom candidates to guide us in this informal search.} Overall, the repeated measures study analyzes 24 atoms, 15 from the original study and 9 that are new.
-Some of these atoms are also analyzed in the Latin square study (Section~\ref{sec:meth:survey}).
-Table~\ref{tab:atomsBoth} presents the atom candidates that both studies investigate and Table~\ref{tab:atomsRepeated} presents the ones that only the repeated measures study analyzes.
-
-\input{atom_descriptions}
+We leverage previous work~\cite{castor2018} on the identification of atom candidates to guide us in this informal search.} Overall, the repeated measures study analyzes 24 atoms, 15 from the original study and 9 that are new (see Tables~\ref{tab:atomsBoth} and ~\ref{tab:atomsRepeated}.
 

### 2022-12-14 background.tex
review os s04

--- 2021-10-25 2_Background.tex
@@
 One example that occurs in many programming languages is the Change of Literal Encoding.
-For instance, the \clang code statement \lstinline{printf("\%d", 013);}
+For instance, the \clang code statement \lstinline{printf("%d", 013);}
 often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}.

### 2022-12-14 main.tex
Making some modifications to address the reviwers comments, in particular pertaining to the reasons for having multiple studies.
minor changes
changing the paper organization
review os s04

@@
 
-\newcommand{\cpplang}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\small\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\small\bf +}\xspace}
+\newcommand{\cpplang}{C\nolinebreak\hspace{-.05em}\raisebox{.6ex}{\tiny\bf +}\nolinebreak\hspace{-.10em}\raisebox{.6ex}{\tiny\bf +}\xspace}
 
@@
 
-\newcommand{\rqa}{What is the impact of the analyzed code patterns% atoms of confusion 
+\newcommand{\rqa}{What is the impact of the analyzed code patterns % atoms of confusion 
  on the comprehension of JavaScript code?}
@@
 \newcommand{\rqd}{What is the frequency of occurrence of atoms of confusion in practice (i.e., in open-source JavaScript projects)?}
+
+\newcommand{\revised}[1]{\textcolor{blue}{#1}}
 
@@
 \lstdefinelanguage{JavaScript}[]{Java}{
-  morekeywords={let, console.log},
+  morekeywords={var, let, console.log},
   moredelim=[is][\textcolor{darkgray}]{\%\%}{\%\%},
@@
 
+\lstdefinelanguage{CodeQL}[]{Java}{
+  morekeywords={select, from, where, not, and}
+}
 \newcommand{\na}{10\xspace}
@@
 
-\input{0_Abstract}
+\input{abstract}
 
@@
 
-\input{1_Introduction_alt}
-\input{2_Background}
-%\input{7_Atoms}
-\input{3_Method}
-\input{4_Result}
+\input{introduction}
+\input{background}
+\input{methodology}
+\input{s01}
+\input{s02}
+\input{s03}
+\input{s04}
 \input{discussion}
-\input{5_Threats}
-\input{6_Conclusion}
+\input{threats}
+\input{conclusion}
 

### 2022-12-14 s04.tex
presenting an overview of the studies
review os s04

@@
-\section{Prevalence Study}\label{sec:method:msr}
+\section{Prevalence Study}\label{sec:s04}
 
-To understand how often the analyzed atom candidates appear in open source projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mine a set of GitHub open source repositories and compute the prevalence of atom candidates.
+To understand how often the analyzed atom candidates appear in open source JavaScript projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mine a set of GitHub open source repositories and compute the prevalence of atom candidates.
 
@@
 This metric, available through the GitHub API, represents the number of stars a project received from users of the platform.
-The same metric has been used in a number of previous studies as a proxy to estimate a project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}.
-We then select the top 100 most popular repositories and remove projects that do not reach the first quartile of the distribution of lines of code.
+The same metric has been used in a number of previous studies as a proxy to estimate projects' popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}.
+We then select the top 100 most popular repositories and remove projects that do not reach the first quartile of the distribution of lines of code (excluding from our dataset small projects).
 
@@
 
-We mine the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}.
+We mine the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we specified using the Semmle's CodeQL language~\cite{moor:gttse2007}.
 CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript).
-We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study).
+All CodeQL queries deal with a specific atom candidate, and collect the exact location in the source code where the atoms appear.
+Figure~\ref{lst:codeql} shows an example of a CodeQL query that mine occurrences of the conditional operator.
+
+Although most of our queries are simple, we have to circumvent a few corner cases to avoid false positives.
+For instance, we do not consider atom candidates the pre-increment occurrences that appear within
+\texttt{for statement} definitions.
+We automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study).
 Finally, we compute some descriptive statistics to measure the prevalence of atoms of confusion in practice.
@@
    \caption{Some descriptive statistics about the projects used in the MSR study}
-  \begin{scriptsize}
+%  \begin{scriptsize}
  \begin{tabular}{rrrrr}
@@
  \end{tabular}
- \end{scriptsize}
+ %\end{scriptsize}
  \label{tab:projects-statistics} 
 \end{table}
+
+\begin{figure}
+\begin{lstlisting}[language=CodeQL]
+import javascript
+
+from ConditionalExpr e
+where not e.getTopLevel().isMinified() 
+select e, e.getLocation().getStartLine() as Location,e.getLocation().getFile()
+\end{lstlisting}
+\caption{CodeQL query that searches for Conditional Expressions.}
+\label{lst:codeql}
+\end{figure}
+
+%% \begin{lstlisting}[language=CodeQL]
+%% import javascript
+
+%% from ControlStmt c
+%% where not c.getTopLevel().isMinified() and
+%%  (c instanceof IfStmt and not c.(IfStmt).getCondition().mayHaveBooleanValue(true)) or
+%%  (c instanceof LoopStmt and  not c.(LoopStmt).getTest().mayHaveBooleanValue(true) ) or 
+%%  (c instanceof WithStmt and not  c.(WithStmt).getExpr().mayHaveBooleanValue(true))  
+%% select c, c.getLocation().getStartLine() as Location,c.getLocation().getFile()
+%% \end{lstlisting}
 
@@
 We mined \minedprojects open source JavaScript repositories to understand how
-often atoms arise in real software. As in the interview study, this study focused on atoms that were investigated in both experiments.
+often atoms arise in real software. As in the interview study, this study focused on atoms that were investigated in both
+experiments (detailed in Section~\ref{sec:s01} and Section~\ref{sec:s02}).
 
@@
 frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
-occur in at least 80\% of the projects. Considering the extremes, atom candidates Conditional Operator and Implicit Predicate were found in all repositories we mined,
-while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
+occur in at least 80\% of the projects. Considering the extremes, atom candidates Conditional Operator and
+Implicit Predicate were found in all repositories we mined (see examples in Figure~\ref{lst:conditional-operator-sample}
+and Figure~\ref{lst:implicit-predicate-sample}), while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).
+Figure~\ref{lst:comma-operator-sample} shows an example of the use of the Comma Operator in the \texttt{styled-components}
+project.
+
+\begin{figure}[htb]
+  \begin{lstlisting}[language=JavaScript]
+function prependModifierMarker (symbol, name, dynamic) {
+  return dynamic
+  ? ("_p(" + name + ",\"" + symbol + "\")")
+  : symbol + name // mark the event as captured
+}
+
+function concat (a, b) {
+  return a ? b ? (a + ' ' + b) : a : (b || '')
+}
+  \end{lstlisting}
+  \caption{Examples of the Conditional Operator atom present in the \texttt{vue} project (urls: shorturl.at/bJPS4 and shorturl.at/tJOS8)}
+   \label{lst:conditional-operator-sample}
+\end{figure}
+
+\begin{figure}[htb]
+  \begin{lstlisting}[language=JavaScript]
+React.useEffect(() => {
+  if (props.config.selectedLines) { // Implicit Predicate
+    dispatch({
+      type: 'MULTILINE',
+      selectedLines: props.config.selectedLines,
+    })
+  }
+}, [props.config.selectedLines])
+  \end{lstlisting}
+  \caption{Example of the Implicit Predicate atom present in the \texttt{carbon} project (url: shorturl.at/aloqU)}
+  \label{lst:implicit-predicate-sample}
+\end{figure}
+
+\begin{figure}[htb]
+  \begin{lstlisting}[language=JavaScript]
+var props = ((_props = {}),
+ (_props[SC_ATTR] = stringifyNames(names)),
+ (_props[SC_VERSION_ATTR] = '4.3.1'),
+ _props);
+  \end{lstlisting}
+  \caption{Example of the Comma Operator atom present in the \texttt{styled-components} project (url: shorturl.at/sCKM0)}
+  \label{lst:comma-operator-sample}
+\end{figure}
 
@@
   \centering
-  {\scriptsize 
 \caption{Summary of atom candidate occurrences in our dataset.}
@@
 \end{tabular}
-  }
+
 \end{table}

### 2022-12-14 s02.tex
presenting an overview of the studies
review os s04
remove unused final_ds

@@
-\section{Latin Square Study}\label{sec:meth:survey}
+\section{Latin Square Study}\label{sec:s02}
 
-This experiment uses a different experimental design that makes it possible to control for the experience of the subjects.
+This study uses a different experimental design that makes it possible to control for the experience of the subjects.
 It employs a subset of the atom candidates used in the repeated measures study.
@@
    \toprule
+      & \multicolumn{3}{c}{Correcness analysis} & \multicolumn{2}{c}{Time analysis} \\ \toprule
 Atom  & Chi-square & Odds           & Confidence & Mann-Whitey& Cliff's\\ 

### 2022-12-14 discussion.tex
review os s04

@@
 Finally, as discussed in Section~\ref{sec:repeated:correct}, Array Destructuring exhibited a large effect size, in spite of the non-statistically significant difference (after correction).
-As discussed elsehwere~\cite{Ellis:2010:EGE}, this is a scenario that hints at practical relevance, specially considering the very low p-value ($< 0.003$).
+As discussed elsewhere~\cite{Ellis:2010:EGE}, this is a scenario that hints at practical relevance, especially considering the very low p-value ($< 0.003$).
 Further studies investigating these code patterns, potentially using different response variables~\cite{Oliveira:2020:ECR,Feitelson:2021:CPC}, are left for future work.
@@
 
-\castor{For the pros and cons of each approach, do you prefer lists of items or regular text paragraphs?}
+% \castor{For the pros and cons of each approach, do you prefer lists of items or regular text paragraphs?}
 \revised{The experimental designs we have employed in the two experiments have different trade-offs. The main advantages of having a repeated measures (or within subjects) study are the following: 

### 2022-12-14 package/s03_mining_software_repositories/analysis.Rmd

---
title: "Atoms JS - Prevalence Study"
author: "Rodrigo Bonifácio et al."
date: '2022-12-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(".")
require(sqldf)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
prevalence <- read.csv("dataset.csv")
repositories <- read.csv("repositories.csv")

##  [1] "language"         "id"               "full_name"        "name"            
##  [5] "size"             "stargazers_count" "forks"            "created_at"      
##  [9] "git_url"          "num_lines"        "num_contributors"

colnames(repositories)
colnames(prevalence)
nrow(prevalence)

## [1] "atom"      "atom_name" "full_name" "location"  "url"

sqldf("select atom_name, count(*) 
       from prevalence 
       group by atom_name 
      order by 2 desc")

sqldf("select count(distinct full_name) 
       from prevalence")

sqldf("select atom_name, count(distinct full_name) * 100 / 72.00 as Prevalence
      from prevalence 
       group by atom_name 
       order by 2 desc")

summary(repositories$num_lines)

full_ds <- sqldf("select r.name, r.size, r.num_lines, p.* 
                  from repositories r, prevalence p 
                  where r.full_name = p.full_name")

sqldf("select count(distinct name) from full_ds")

sqldf("select name, size, num_lines, count(distinct atom_name), count(*)
       from full_ds 
       group by name, size, num_lines 
       order by 3 desc")

##                      atom_name Prevalence
## 1              TernaryOperator  100.00000
## 2            ImplicitPredicate  100.00000
## 3           OmittedCurlyBraces   91.66667
## 4      PostIncrementExpression   90.27778
## 5        PreIncremenExpression   83.33333
## 6            AssignmentasValue   81.94444
## 7           LogicasControlFlow   56.94444
## 8  AutomaticSemicolonInsertion   33.33333
## 9            ArithmeticasLogic   23.61111
## 10               CommaOperator   15.27778

sqldf("select * from 
       full_ds where  name = 'carbon' and atom_name = 'ImplicitPredicate'
       order by RANDOM() 
       LIMIT 100")
```

### 2022-12-15 response-letter01.tex
early draft version of the response letter

\documentclass{letter}

\usepackage{xcolor}

\usepackage{hyperref}
\signature{Rodrigo Bonif\'{a}cio}
\address{Computer Science Department \\ University of Bras\'{i}lia \\ Brazil}
\begin{document}

\begin{letter}{Journal of Systems and Software Editors-in-Chief}

\opening{Dear Profs. P. Avgeriou and D. Shepherd}

First of all, we would like to thank the anonymous reviewers of
our paper.

{\bf R1 concern about the paper's structure}

We changed the structure of the paper. In the previous version,
we organized the paper in terms of Study Settings, Results, and Discussion.
Each one of those sections present information of the four studies.
We now orgnized the paper in terms of one section per study, and each
one of these sections detail the Study Settings and Results. We
hope that this new structure might have improved the paper understanding.
This new structure also makes easier to change the order of the
studies' presentation. So, if the reviewers agree that we should
present our fourth study first (as R1 suggested), we can easily
apply this change. Thanks for the comment.

{\bf R1 concern about using effect size}

We are computing effect size even when the differences are
not statistically relevant because this might reveal practical
relevance, in particular in border lines scenarios---as is the case
that we show in the paper. If the reviewers do not agree with our
decision, we can remove these calculations from the paper in the
final version. Thanks for the comment.

{\bf R2 concern about the relevance of studying atoms of confusion in JavaScript}

We changed the first section of the paper (Introduction, page 3) to motivate our study better. The
changes are in {\color{blue}blue}. Thanks for the comment.

{\bf R2 concern about the procedures we use to identify atom candidates}

We clarified this point on the current section 4.1 of the paper (page 10). The changes are in
{\color{blue}blue}. Thanks for the comment.

{\bf R3 concern about the lengthy of the MSR study}

Considering the new organization of the paper, we have actually devoted almost 5 pages
of the paper to the MSR study on Prevalence (excluding the discussion section). This is comparable to the
other three studies. Nonetheless, if the reviewers agree that we should put more emphasis
on specific parts of this study, we can enrich this section in the final
version of the paper. Thanks for the comment.

{\bf R3 concerns about the lack of replication package}

We are sharing our replication package at Zenodo. {\color{red}Incluir um link aqui}.
We totally agree that we should have shared it before. We hope that our package might
help understand our work better. We have included examples of the CodeQL queries we used
to search for atoms of confusion in open source JavaScript projects---both in the paper
as well as in the replication package. The replication package also contains all atoms of
confusion we collected using the CodeQL queris (MSR study) and the atoms candidates
we explored in the survey and experiments. Some of our queries might lead to
either false positives and false negatives, although we worked hard to improve
accuracy. We include a new threat to validity regarding this issue, that might threaten
any comparison of atoms prevalence in JavaScript and C/C++. Thanks again for this comment.

{\bf R3 concerns about the presentation of both experiments (RM/LS)}

We justify better the report of both experiments in Section 3 (page 7). The changes
are in {\color{blue}blue}. We revisited this issue again on Section 8.3 (pages 36, 37, and
38). The changes are in {\color{blue}blue}.

Unfortunatly, we cannot use the same test (Chi-square) for both experiments. The McNemar test targets paired,
dependent samples (\url{https://en.wikipedia.org/wiki/McNemar_test}). This is the case of the repeated measures study.
Chi-square is for independent samples. The Chi-square test assumes that the groups are not related
(\url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3900058/}). This is precisely what happens in the Latin square study,
where each participant is exposed to a single treatment. This is not the case in the repeated measures study,
where the same subjects receive both treatments.

We have also fixed all minor issues the reviewers reported. Thanks one more time.

\closing{Yours Faithfully,}

\end{letter}
\end{document}

### 2022-12-16 s02.tex
improving threats to validity

@@
    \toprule
-      & \multicolumn{3}{c}{Correcness analysis} & \multicolumn{2}{c}{Time analysis} \\ \toprule
+      & \multicolumn{3}{c}{Correctness analysis} & \multicolumn{2}{c}{Time analysis} \\ \toprule
 Atom  & Chi-square & Odds           & Confidence & Mann-Whitey& Cliff's\\ 

### 2022-12-16 threats.tex
changing the paper organization
improving threats to validity

--- 2021-09-27 5_Threats.tex
@@
 \label{threat}
+
+In this section we discuss some of the main threats to the validity of this work.
 
@@
 Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code.
+
+\revised{\textbf{Precision of our CodeQL queries.}
+  We relied on existing CodeQL JavaScript predicates to implement most of our source code queries. However,
+  we still had to implement our own predicates for three specific queries: the queries that search for the atoms Automatic Semicolon Insertion, Comma Operator, and Omitted Curly Braces.
+  Figure~\ref{lst:codeql-automatic} shows examples of predicates we have implemented. Our test cases and manual assessments revealed that our queries correctly identified the
+  atoms in the source code. Unfortunately, we observed that a predefined CodeQL predicate (named mayHaveBooleanValue) we used to mine the Implicit Predicate atom might have
+  produced many false positives. That is, there are situations where this predicate wrongly states that an expression does not have a boolean value.
+  This issue might have inflated the number of Implicit Predicate occurrences we reported in Section~\ref{sec:s04}}.
+
+\begin{figure}[htb]
+\begin{lstlisting}[language=CodeQL, basicstyle=\scriptsize]
+import javascript
+
+predicate hasTokenAfter(ReturnStmt r) {
+  r.getLastToken().getNextToken().getValue() != "}"
+}
+
+predicate isNextTokenVariable(ReturnStmt r) {
+  r.getLastToken().getNextToken() instanceof IdentifierToken and 
+  r.getLastToken().getNextToken().getNextToken().getValue() != "="
+}
+
+predicate isNextTokenLiteral(ReturnStmt r) {
+  ( r.getLastToken().getNextToken() instanceof StringLiteralToken or
+    r.getLastToken().getNextToken() instanceof NumericLiteralToken ) and 
+  r.getLastToken().getNextToken().getNextToken().getValue() != "="
+ }
+
+from ReturnStmt r
+where not r.getTopLevel().isMinified() and
+ r.isSubjectToSemicolonInsertion() and 
+ r.getNumChildStmt() = 0 and
+ r.getNumChildExpr() = 0 and
+ not r.toString().matches("%;") and
+ hasTokenAfter(r) 
+select r, r.getLocation().getStartLine(), r.getLocation().getFile()
+\end{lstlisting}
+  \caption{CodeQL query that searches for the atom Automatic Semicolon Insertion. Note that, for this query,
+    we had to implement three CodeQL predicates.}
+  \label{lst:codeql-automatic}
+\end{figure}
 % \castor{A forma como os átomos foram detectados é uma ameaça que não discutimos.}

### 2022-12-16 response-letter01.tex
improving threats to validity
minor fixes

@@
 
-\opening{Dear Profs. P. Avgeriou and D. Shepherd}
+  \opening{Dear JSS editors and reviewers}
 
-First of all, we would like to thank the anonymous reviewers of
-our paper.
+  We thank the associate editor and reviewers
+  for their detailed, constructive comments. We took great care to address all points raised
+  with a thorough revision of the manuscript. In this response, we detail the changes we introduced for major
+  review comments. We hope you will find the revised version of the manuscript satisfactory,
+  and we look forward to any further comments.
 
@@
 
-We changed the structure of the paper. In the previous version,
+We changed the whole structure of the paper. In the previous version,
 we organized the paper in terms of Study Settings, Results, and Discussion.
-Each one of those sections present information of the four studies.
-We now orgnized the paper in terms of one section per study, and each
-one of these sections detail the Study Settings and Results. We
-hope that this new structure might have improved the paper understanding.
-This new structure also makes easier to change the order of the
+Each one of those sections presented detailed information of the four studies.
+We now organized the paper in terms of one section per study, and each
+one of these sections details specific Study Settings and Results. We
+hope that this new structure might have improved the paper readability.
+This new structure also makes it easier to change the order of the
 studies' presentation. So, if the reviewers agree that we should
-present our fourth study first (as R1 suggested), we can easily
+present our Prevalence Study first (as R1 suggested), we can easily
 apply this change. Thanks for the comment.
@@
 relevance, in particular in border lines scenarios---as is the case
-that we show in the paper. If the reviewers do not agree with our
+that we show in the paper. If the reviewers disagree with our
 decision, we can remove these calculations from the paper in the
@@
 
-We clarified this point on the current section 4.1 of the paper (page 10). The changes are in
+We clarified this point in the current Section 4.1 of the paper (page 10). The changes are in
 {\color{blue}blue}. Thanks for the comment.
 
-{\bf R3 concern about the lengthy of the MSR study}
+{\bf R3 concern about the length of the MSR study}
 
-Considering the new organization of the paper, we have actually devoted almost 5 pages
-of the paper to the MSR study on Prevalence (excluding the discussion section). This is comparable to the
-other three studies. Nonetheless, if the reviewers agree that we should put more emphasis
-on specific parts of this study, we can enrich this section in the final
-version of the paper. Thanks for the comment.
+Considering the paper's new organization, we have devoted almost five pages of the paper to the MSR study on
+Prevalence (excluding the discussion section). This length is comparable to the other three studies.
+Nonetheless, if the reviewers agree that we should emphasize specific parts of this study,
+we can enrich this section in the final version of the paper. Thanks for the comment.
 
@@
 
-We are sharing our replication package at Zenodo. {\color{red}Incluir um link aqui}.
+We are sharing our replication \href{https://github.com/rbonifacio/AtomsJS/tree/main/package}{package at GitHub}.
 We totally agree that we should have shared it before. We hope that our package might
-help understand our work better. We have included examples of the CodeQL queries we used
+help a reader to understand our work better. We have included examples of the CodeQL queries we used
 to search for atoms of confusion in open source JavaScript projects---both in the paper
 as well as in the replication package. The replication package also contains all atoms of
-confusion we collected using the CodeQL queris (MSR study) and the atoms candidates
+confusion we collected using the CodeQL queries (MSR study) and the atom candidates
 we explored in the survey and experiments. Some of our queries might lead to
 either false positives and false negatives, although we worked hard to improve
-accuracy. We include a new threat to validity regarding this issue, that might threaten
+accuracy. We included a new threat to validity regarding this issue that might threaten
 any comparison of atoms prevalence in JavaScript and C/C++. Thanks again for this comment.
@@
 
-Unfortunatly, we cannot use the same test (Chi-square) for both experiments. The McNemar test targets paired,
+Unfortunately, we cannot use the same test (Chi-square) for both experiments (as R3 suggested). The McNemar test targets paired,
 dependent samples (\url{https://en.wikipedia.org/wiki/McNemar_test}). This is the case of the repeated measures study.
@@
 
-We have also fixed all minor issues the reviewers reported. Thanks one more time.
+We have also fixed all other minor issues the reviewers reported. Thanks one more time.
 

### 2022-12-16 s02.tex
final revision version

@@
 %As part of this effort, we 
-carried out an informal pilot.% whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
+carried out an informal pilot.
 Undergraduate students and professional colleagues took part in the pilot.

### 2022-12-20 biography.tex
adding paper's metadata

\documentclass{article}

\usepackage[brazil]{babel}
\usepackage[utf8]{inputenc}

\title{Authors Biography}

\begin{document}

\maketitle

\begin{itemize}

\item Adriano Torres is a PhD student at The University of Adelaide, Australia. His alma mater is the University of Brasilia, where he completed his Bachelors and MsC. His research focuses on the development of methods and tools that automate the discovery and correction of software faults. He also has industry experience with distributed systems, and currently works as engineering lead for Luma Health Inc.

\item M\'{a}rcio Okimoto is a Database Administrator and junior researcher. He's a specialist in Business Intelligence and Data Warehousing. He also has a Master's Degree in Computer Science from the University of Brasília (UnB) and his major research interest is the Mining Software Repositories (MSR) field with a focus on productivity.

\item Diego Marc\'{i}lio is a Ph.D. student in the Software institute at USI Università della Svizzera italiana in Lugano, Switzerland.
He received his M.S. in Informatics from the University of Brasília, Brazil, in 2019. His research interests center around aiding developers when understanding, writing, evolving, and maintaining source code.

\item Caio Oliveira has a Computer Science Bachelor's Degree from the University of Brasilia. He is a front-end focused developer with nearly 7
  years of experience with Angular. He has worked in projects as full-stack developer with technologies such as JavaEE, Django, Flask, Spring boot, Laravel, and all of
  them had Angular as the frontend technology. Has familiarity with Javascript and Typescript languages.

\item Fernando Castor is an assistant professor at the Department of Information and Computing Sciences, Utrecht University, The Netherlands,  working under prof. Sjaak Brinkkemper, and an associate professor (on leave) at the Informatics Center, Federal University of Pernambuco, Brazil. His research interests center around program understanding, energy consumption, and exception handling.

\item Rodrigo Bonif\'{a}cio is an associate professor at the Computer Science Department, University of Brasília, Brazil, and a member of the Software Productivity Group, led by Prof. Paulo Borba. His research interests lie in understanding
  the common usage of programming language features, program transformations, software architecture and modularity, and software security.

\item M\'{a}rcio Ribeiro is an associate professor at the Computing Institute, Federal University of Alagoas, Brazil. His research interests lie in program refactoring, program understanding, program testing, software modularity, and high configurable systems.

\item Edna Dias Canedo is currently an associate professor at University of Brasilia (UnB - Brazil). She received her Ph.D. (2012) degree in Electrical Engineering from the University of Brasilia (UnB).
  Her current research interests include Software Engineering, Requirements Engineering, Gender Diversity, Software Systems, Empirical Software Engineering, and Usability.

\item Eduardo Monteiro is an associate professor at the Statistics Department in Universidade de Brasília, Brazil, and focuses his research in statistical modeling and applications.  
\end{itemize}

\end{document}

### 2022-12-20 contributions.tex
adding paper's metadata

\documentclass{article}

\usepackage[brazil]{babel}
\usepackage[utf8]{inputenc}

\title{Authors Contributions}

\begin{document}

\maketitle

\begin{itemize}

\item {\bf Adriano Torres} and {\bf Caio Ribeiro} led all steps in the second study (experiment using Latin Square design). {\bf Adriano Torres} also contributed to the third study (interviews),
  under the supervision of {\bf Edna Dias Canedo}.

\item {\bf M\'{a}rcio Okimoto} and {\bf Diego Marc\'{i}lio} led all steps in the fourth study (MSR effort), implementing the CodeQL queries---under the
  supervision of {\bf Rodrigo Bonif\'{a}cio} and {\bf Edna Dias Canedo}.

\item {\bf Fernando Castor} and {\bf Pedro Queiroga} led all steps in the first study (experiment using the Repeated Measures design). {\bf Fernando Castor}, {\bf Rodrigo Bonif\'{a}cio}, {\bf M\'{a}rcio Ribeiro}, {\bf Edna Dias Canedo}, and {\bf Eduardo Monteiro} co-supervised the whole research.

\item {\bf Eduardo Monteiro} contributed to defining the statistical methods used in this research and implementing several R scripts for evaluating the outcomes of this research.

\item {\bf All authors} have contributed to the writting of the manuscript.    
\end{itemize}

\end{document}

### 2022-12-20 highlights.tex
adding paper's metadata

\documentclass{article}

\title{Paper Highlights}

\begin{document}

\maketitle

\begin{itemize}

\item We present the results of four empirical studies that explores the impact and
  prevalence of atoms of confusion in JavaScript.

\item We bring evidence that two code patterns that have been previously observed to confuse C
  programmers also confuse JavaScript programmers: the comma operator and assignments being used as values.

\item We reveal that some JavaScript-specific elements, such as automatic semicolon insertion and object destructuring,
  also have the potential to cause confusion.

\item We shhow that examined code patterns are frequent and used intensively in 72 popular open-source JavaScript projects:
  four atom candidates appear in 90\% of them and two of them occur more than once for every 100 lines of code.

\end{itemize}

\end{document}

### 2023-01-11 response_reviewer_1_2nd_round.md
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

In general, I do not understand why some of my concerns are addressed while others are purposefully ignored. Even minor issues, such as missing thousands separators, are ignored. I feel that my review was not respected, and the authors' changes to my review were not highlighted, causing me to spend a lot of time trying to find out if they had made changes. The major revision decision will remain until all my comments have been responded to by the authors.

----------------------------------//-----------------------------------

Reviewer #1: Summary

General Comments

Thank you for providing me with an enjoyable review experience. The authors clearly have a good capacity to construct and organize their work on a professional level. Generally, I am satisfied with the paper, but I still have some concerns that I hope the authors will address.

---> We thank the reviewer for the kinds words.

Detailed Comments

- What concerns me the most are the paper's contributions. In some ways, the paper is like an extended work from C to JavaScript. There are only a few new atom candidates being researched that are specific to JavaScript. Given this concern, I believe it is necessary to reorder the RQs in order to boost the work's contributions. RQ3 should be the first RQ to uncover the prevalence of atom candidates in real-world js programming. There would be no need for this study to be done if these atom candidates were infrequently observed in the real world. Based on this discovery, the other experiments could be continued.

---> We have re-organized the studies as suggested by the reviewer. The repository mining study now comes first and is presented as a motivating study for the remaining studies, highlighting their relevance. Besides the re-ordering of the sections, changes have been applied to multiple parts of the paper.

- Pages 1-2: The research's conclusions and findings, which are summarized in the abstract, are too detailed and somewhat tedious. I would suggest that the authors revisit the answers to the three RQs and rewrite the back half of the abstract.

---> We have cut about 85 words from the abstract, specifically from the part about the contributions. We avoided cutting more because the paper presents the results of four different studies and the abstract to a certain extent reflects that.

- Page 6: At the beginning of section 3, I expect a figure introducing the entire workflow. Given that the paper includes several experiments, it would be helpful to use a figure to create a broad overview of the paper.

*Make the workflow.*

- Page 11: As stated by the authors on this page,  for the second experiment, they only use ten of the atom candidates that were used in the first experiment. Table 2 shows that 14 atom candidates were not studied in the second experiment. Can the authors explain how the atom candidate was chosen for the second experiment?

---> In the second experiment, we have tackled the nine atoms that have been validated for the C language in the original paper by Gopstein et al. (2017). In addition, we have studied a single atom that is specific to JavaScript programs, Automated Semicolon Insertion. The explain this in the beginning of Section 6. The reason for analyzing a smaller number of atoms is to attempt to increase the number of participants, since a large number of tasks in a study usually results in subjects not participating or dropping out of the study. This strategy was successful, as can be seen by the considerably larger number of participants in the Latin Square Study, twice as large as that of the repeated measures study. We emphasize that the two experiments were designed and conducted independently, by two independent groups without knowledge about the other. This explains why some atom candidates are in one study and not the other.

- Page 25: It seems that developers need more time to submit the correct answer using the obfuscated versions of two atom candidates than the clean versions. Is there a rationale behind this?

---> This could only be observed with statistical significance for three of the atoms, after p-value correction. Effect sizes are also mostly very small and they are not affected by sample sizes, differently from the p-values. This combination of results makes it difficult to formulate hypotheses beyond the obvious: that obfuscated snippets are slightly more confusing than the cleaned ones, with few exceptions.

- Page 34: I do not think it is a good idea to answer all of the RQs on one page. Since the experiments involve so many details, it is difficult for readers to recall all of them when reading these answers. I would suggest treating RQs as separate sections, with the answers to RQ serving as a summary of the section. They could have a look at the following paper:
Li, H., Shang, W. & Hassan, A.E. Which log level should developers choose for a new logging statement?. Empir Software Eng 22, 1684-1716 (2017).

---> We thank the reviewer for the comment. We have restructured the sections of the paper so that each study is wholly contained in a single section, both the methodology and the results. This addresses a comment from another review but we believe it improves the cohesiveness of the sections. As for the summaries of the responses, we would prefer to keep it in a single place, though, for ease of reference. The goal is to enable readers to have a high-level overview of the results in a single place.

Minor issues:

- Page 6: the analyzed code patternson -> the analyzed code patterns on

---> Fixed. Thanks.

- Page 13: pilot.Undergraduate -> pilot. Undergraduate

---> Fixed. Thanks.

- Thousand separators are missing in numbers.

---> Fixed. Thanks. We gave a pass through the entire paper but only found numbers requiring this correction in Section 4 (the MSR study). Please let us know if we have missed any others.

- I would suggest rounding the decimals to the same number of decimal places.

---> We have partially incorporated this suggestion. We focused mainly on the tables, where we believe the uniformity matters more for presentation and interpretation purposes. For p-values, we kept the four decimal places because it is relevant to be precise, as some p-values are very small. For all the other numbers where decimal places matter, we used two. In the text itself, we did not change the number of decimal places because we do not think much is lost. For example, for a p-value of 0.007, we prefer to use 0.007 instead of saying 0.0070 just to preserve uniformity.

- Page 25: Seven atom candidates are mentioned in the text, but the table lists eight candidates.

---> Fixed. Thanks.

### 2023-01-11 3_Method.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

--- 2022-10-04 3_Method.tex
@@
 \begin{enumerate}[(RQ1)]
+\item \rqd
 \item \rqa 
 \item \rqb 
-% \item \rqc  
-\item \rqd
 \end{enumerate}
 
-By answering these research questions, we can either generalize or refute the findings about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). 
-In addition, answering the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice.
+\revised{Answering RQ1 enables us to better motivate the other two questions by showing that studying these code patterns has practical relevance.}
+Futhermore, by addressing research questions RQ2 and RQ3, we can either generalize or refute the findings about atoms of confusion already discussed in the literature. 
 Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions.
@@
 
-These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ1).
+We perform a repository mining study (Section~\ref{sec:method:msr}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
+This helps us motivate the remaining three studies.
+The two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ2).
 They investigate whether or not programs that contain atom candidates tend to produce more misunderstanding for programmers trying to predict their outcome.% We leverage two independently-developed and conducted experiments with different designs to more
-We also conduct a set of interviews with developers (Section~\ref{sec:method:interview}) to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2).
-Finally, we perform a repository mining study (Section~\ref{sec:method:msr}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
-
-\subsection{Repeated Measures Study}\label{sec:meth:replication}
-
-The first experiment we perform uses a repeated measures, or within-subject, design~\cite{Keselman:2001:ARM}.
-This experiment is inspired by the work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, but targeting JavaScript instead of the C language, and including some methodological differences.
-It covers all the atom candidates analyzed in the work of Gopstein et al.
-that can be transposed to JavaScript and adds candidates that are specific to this language.
-%Overall, we select 24 atom candidates in the repeated measures study.
-
-\subsubsection*{Experimental design}
-
-In this experiment, the control variable consists of a tiny program containing a single atom candidate and the treatment is a functionally-equivalent version that does not contain the atom.
-The main dependent variable is whether subjects are able to correctly determine the output of the programs.
-The other dependent variable is the time required to correctly vs.
-incorrectly determine the outputs, independently of the presence of atoms.
-We use the same null hypothesis as Gopstein et al.: \textit{``code from both control and treatment groups can be hand-evaluated with equal accuracy''} and the alternate hypothesis that \textit{``the existence of an atom of confusion causes more errors than other code in hand-evaluated outputs.''}.
-For each atom candidate, we built six programs, three versions including the atom candidate and three versions not including it.
-%Each subject is exposed to four programs for every atom candidate, two including an atom and two not including an atom.
-We have also built six programs for each atom candidate, three with and three without the atom.
-Considering the elements that can vary in a replication study, defined by Juristo and Gomez~\cite{Juristo2012}, in this study we employ the same experimental protocol and construct operationalizations.
-We have different population properties, experimenters, and a different site.
-In addition, since we focus on a different language, some atoms that exist in C cannot by construction exist in JavaScript programs.
-Conversely, JavaScript exhibits atom candidates that do not exist in C.
-
-% The design of the repeated measures study considers two treatments: the presence or absence of atom candidates within the programs.
-To identify atom candidates, we start out by selecting every atom candidate from Gopstein's study that can also happen in JavaScript.
-That excludes atoms related to pointers and preprocessor directives.
-Furthermore, we conduct an informal search of programming forums, solutions to Code Golf~\footnote{https://codegolf.stackexchange.com/} problems, code style guidelines, and expert knowledge to identify new atom candidates.
-Overall, the repeated measures study analyzes 24 atoms, 15 from the original study and 9 that are new.
-Some of these atoms are also analyzed in the Latin square study (Section~\ref{sec:meth:survey}).
-Table~\ref{tab:atomsBoth} presents the atom candidates that both studies investigate and Table~\ref{tab:atomsRepeated} presents the ones that only the repeated measures study analyzes.
-
-\input{atom_descriptions}
-
-For each atom candidate, we write six short programs, three including and three not including the atom candidate.
-The former are the control for the study, or the \textbf{obfuscated} versions of the programs.
-The alternative, atom-free, versions of the programs are the treatment, or \textbf{clean}, versions.
-Each subject is exposed to one obfuscated and one clean version of each atom, totalling 48 programs, and should determine their outputs.
-As mentioned before, we measure answer correctness and the time to answer questions correctly vs.
-incorrectly.
-We control for learning effects~\cite{Neely:1991:SPE} in three manners.
-First, by having multiple obfuscated and clean versions for each atom and presenting only one of each per subject.
-Second, by presenting the programs in a random order.
-Third, by presenting obfuscated and clean versions corresponding to the same atom candidate with at least 11 other programs in between, in  accordance to the original experimental protocol~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
-
-\subsubsection*{Study instrument}
-
-The experiment is conducted by means of a questionnaire made available as a web application.
-As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the questionnaire would take on average.
-Based on the feedback of the participants of the pilot, we present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.% Initially, our plan was to present two of each.
-
-We organize the study in three sections.
-In the first one we present instructions and also a check button whose checking means users agree that all collected data will be used anonymously and solely for research purposes.
-The instructions explain how the study works and asks them to dedicate their attention to it.
-We stress to participants the importance of not using any aids during the experiment, such as online or console interpreters.
-
-The second section presents the programs, one at a time.
-For each one, there is a text box where the answer should be written and a ``Next'' button.
-If the subject leaves the text box empty and presses ``Next'', this is treated as a wrong answer.
-To reduce the likelihood of a subject attempting to execute the programs, the web application verifies whether the subject tries to change tabs or windows and presents a pop-up message if that is the case.
-Furthermore, it presents the programs as images, instead of text, to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter.
-Upon submitting an answer for a particular program, the subject is automatically led to a similar page, containing the next program.
-
-We do not provide feedback about the time subjects take to answer each question.
-We also do not tell them whether their answers are correct or not.
-This aims at avoiding introducing bias for future respondents.
-Since we posted information about the study on social media platforms, possible threats could have arisen if we gave respondents instant feedback.
-
-\subsubsection*{Study audience}
-
-We posted the link to the questionnaire on the authors' social media platforms, on Brazilian CS departments' mailing lists, and on two programming subreddits.
-We explained our research purposes and asked developers to take part in the study only if they have some familiarity with JavaScript.
-We collected 70 responses.
-%The average survey respondent has a mean 26 months of experience in JavaScript, with a median of 12.\castor{We should not say this here unless we are able to do the same for the Latin square study.
-}
-
-\subsection{Latin Square Study}\label{sec:meth:survey}
-
-This experiment uses a different experimental design that makes it possible to control for the experience of the subjects.
-It employs a subset of the atom candidates used in the repeated measures study.
-More specifically, this study investigates ten of the atoms investigated in the repeated measures study:
-%Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first 
-%select a set of 9+1 atom candidates: 
-nine previously-validated atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs plus one atom candidate that is specific to the JavaScript language (Automatic Semicolon Insertion).
-%Our supplementary material discusses and shows examples of the atoms we consider in our research.\castor{We should probably do this here or just refer to the table that I intend to insert for the replication study.}
-
-\subsubsection*{Experimental design}
-
-The design of the Latin square study blocks two variables (subject experience and the programs) and considers two treatments: the presence or absence of atom candidates within the programs.
-To achieve such a design goal, controlling the effect of experience and individual programs, we resort to the \textit{Latin Square Design}~\cite{Hunter-Experimenters}.
-Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of programs.
-The design of each square (a replica) is such that no treatment is repeated in the same row or column.
-For example, considering that we have a set of 10
-code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about clean programs, they will only be presented with clean versions of the code samples $s_6, s_7,..., s_{10}$.
-Furthermore, a given subject (P2), who constitutes the second row of our example square, will be asked questions about the clean versions for $s_1, s_2, ..., s_5$, and will answer questions about obfuscated programs for $s_6, s_7,..., s_{10}$.
-By doing that, we guarantee that all versions of the programs are contained within each square, and that each configuration occurs only once within a square.
-Figure \ref{fig:latinsquare} offers a visual representation of the concept.
-
-  \begin{figure}[htb!]
-      \noindent
-      \centering
-      \includegraphics[scale=.50]{images/latin-square.pdf}
-      \caption{Latin square design. Each ``square'' corresponds to 
-      a replica in our study. Each replica comprises two participants (square rows, e.g., P1 and P2) 
-      and two sets of programs (CS-1 and CS-2). We randomly apply the 
-      treatments (atom or non-atom code) to the cells of the squares.} 
-      \label{fig:latinsquare}
-  \end{figure}
-
-For each of the 10 selected atom candidates, we write one short program containing the atom.
-As in the repeated measures study, we call them the obfuscated versions of the programs.
-We also write corresponding short programs without the atom candidates and call them the clean versions of the programs.
-Overall, the experiment employed 20 programs, 10 obfuscated and 10 clean.
-In order to reduce the cognitive effort, each subject was asked to predict the output of 10 programs, five obfuscated and five clean.
-The order in which they are presented is randomized.
-By doing this, we seek to minimize the chances of subjects being aware that the current listing they are analyzing contains (or not) atoms of confusion.
-That is, each subject should indicate what would be the outcomes of the programs, some of them having atoms of confusion (while other programs do not).
-Since participants are not exposed to obfuscated and clean versions corresponding to the same atom candidate, learning effect is not possible.
-We measure answer correctness and the total time each participant requires to participate in the experiment.
-
-\subsubsection*{Study instrument}
-
-We implemented our experiment as a questionnaire in a web application and 
-%carried out a pilot. 
-%As part of this effort, we 
-carried out an informal pilot.% whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average.
-Undergraduate students and professional colleagues took part in the pilot.
-Some users reported layout defects, and many reported that the landing page did not explain the study well enough.
-We also spotted minor issues with our routines to create and populate the latin squares.
-
-We organize the questionnaire in three sections.
-The first section aims to characterize the subjects, asking their age, education level, and programming experience.
-We also include a check button, whose checking means users agree that all collected data will be used solely for research purposes.
-In the second section, we present to the participants a small set of instructions, where we explain how the study works and ask them to dedicate their attention to it.
-We stress to participants the importance of not using any aids during the experiment, such as online or console interpreters.
-For each question page, we kept track of whether or not the subjects switch windows.
-
-The last section of the questionnaire presents a sequence of ten questions, each containing a program.
-For each question, there is a text box where the answer should be written.
-There is also an ``I do not know'' button, which, when clicked, leads the subject to the next question.
-In our setting, ``I do not know'' is treated as a wrong answer.
-The programs are also presented as images copied from a text editor.
-Upon submitting their answer for a particular question, the subjects are automatically led to a similar page, containing another program.
-Similarly to the repeated measures study, we do not provide feedback about time and correctness to the subjects.
-
-% \rb{nao acho esse paragrafo necessario} 
-% \adriano{o de baixo ne? eu tambem acho}
-
-% As we mentioned before, we first wrote the code listings in a text editor, and took pictures of it.
-In the case of an atom of confusion that was exclusive to JavaScript, which we called \textit{Automatic Semicolon Insertion} (see Appendix~\ref{sec:appendix-atoms}), it was necessary to remove the syntax highlighter.
-Even though semicolons at the end of statements are optional to programmers in JavaScript, the interpreter automatically inserts them into the code.
-Our text editor was incorrectly highlighting a line break after a return statement, even thought it was valid JavaScript syntax.
-We had thus to turn the highlighter off to take the picture of this atom.
-
-\subsubsection*{Study audience}
-
-%% \diego{Esses detalhes são possíveis candidatos para remoção, se precisarmos de espaço}
-%% {\color{blue}We invited developers to answer our survey by sending invitations to communities of JavaScript programmers on the Internet.
-Our initial plan was to try to engage contributors for two major JavaScript projects, namely Node.js and NPM.
-None of the projects, though, offered a direct way to interact with the community, such as an online discussion forums or mailing list.
-We would be required to first make contact with the projects' leaders, and only then would we have a chance to approach potential respondents.
-Due to time constraints, we decided to look for respondents elsewhere.}
-
-We posted the invitation to take part in the study on a JavaScript Reddit channel.\footnote{https://www.reddit.com/r/javascript/} We explained our research purposes and asked developers of any level of expertise to take part.
-Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares.
-We collected significant data on time taken and discrepancies in answer correctness between obfuscated and clean versions of the programs.
-Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the questionnaire.
-We discarded from our study all squares that contained incomplete rows (a common approach for data imputation).
-%Since we had a large enough number of samples, the squares we had to discard did not impact our analysis.
-
-\subsection{Interview Study Setting}\label{sec:method:interview}
-
-To complement the two experiments, we perform semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding programs containing atoms of confusion (research question RQ2).
-We also ask each participant if they know of any other JavaScript-specific construct or idiom they think is likely to make the code hard to understand.
-% (research question RQ3).
-In this section we detail the protocol we followed to conduct the interviews and to analyze the results.
-
-% Nós realizamos entrevistas semi-estruturas com o objetivo de identificar a percepção dos desenvolvedores com experiência em JavaScript sobre algumas questões relacionadas à compreensão de código em JavaScript.
-Assim, nesta Seção nós descrevemos os procedimentos adotados para selecionar os participantes para as entrevistas e detalhamos como as entrevistas foram conduzidas.
-Além disso, detalhamos como os resultados das entrevistas foram analisados.
-
-\subsubsection*{Participant Selection} We invite the participants of the interviews using a snowballing technique.
-That is, starting from our network of contacts, we invite an initial set of candidates to take part in our experiments.
-From this initial list, we ask for an indication of additional candidates.
-Our main selection criterion is that all participants should have been working with JavaScript in their daily professional activities.
-We invited a total of 17 developers, and 15 of them agreed to participate.
-%We conducted the interviews for a period of two weeks.
-
-\subsubsection*{Interview Process} We conduct semi-structured interviews using web conferencing software.
-We record all the interviews with the consent of the participants.
-On average, the interviews last 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes.
-Two of the researchers conduct the interviews, and a third one listens to all the recordings to cross-validate the collected data.
-The interviews have three main parts.
-In the first one, we ask the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years, and other programming languages they have worked with.
-Table \ref{pinterview} summarizes this demographic information.
-
-\begin{table}[htb!]
-
-  \centering
-  \caption{Demographic information of the participants}
-\begin{scriptsize}  
-\begin{tabular}{clcl}
-\toprule
-ID & Education & JS Experience & Other Languages \\ \midrule 
-P1 & BSc Degree & 9 years & Java, PHP, C, Go  
-\\ 
-P2 & HS Degree & 3 years & Python, Go, Dart, Lua, C++, C\#
-\\ 
-P3 & BSc Degree & 4 years & Java, C
-\\ 
-P4 & Undergraduate & 3 years & Python, C, C++, Java, Go
-\\ 
-P5 & Master Student & 3 years & Python, SQL
-\\ 
-P6 & Bsc Degree & 15  years  & Java, PHP, C, Python , Ruby, C\#
-\\ 
-P7 & BSc Degree  & 6 years & C, C++, Java, Assembly, Kotlin
-\\ 
-P8 & PhD Degree & 2 years & Java, Python
-\\ 
-P9 & BSc Degree  & 5 years & PHP
-\\ 
-P10 & Master Student & 4 years & Java, C\# and Python
-\\ 
-P11 & Master Student & 4 years & Java, Erlang, C\#, Cobol
-\\ 
-P12 & Master Degree & 1 year & C
-\\ 
-P13 & Master Degree & 13 years & Java, PHP
-\\ 
-P14 & Master Student & 3 years & C, Python, Ruby
-\\ 
-P15 &  BSc degree  & 2 years & Java, PHP
-\\ \bottomrule
-\end{tabular}
-\end{scriptsize}
-    \label{pinterview}
-\end{table}
-
-In the second part of the interview our aim is to allow the subjects to describe their JavaScript experience, as well as to allow them to reveal any JavaScript constructs they regard as innately confusing.
-This allows us to identify potential atom candidates that are more specific to the JavaScript language.
-Examples of questions we explore in this section include: \emph{Does JavaScript favor developers to produce code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as especially confusing?}
-
-In the third part of the interview, participants are shown pairs of programs that were used in the experiments, where each pair contains a clean and an obfuscated version of the same atom candidate.
-For this part we only use atoms that appear in both experiments.
-The participants
-%were told both sides of each pair had the same output, and 
-are asked to evaluate which version of the code is easier to understand.
-To avoid introducing bias in the answers, the interviewers do not explain that one of the versions in each pair contains the atom under investigation.
-Subjects are just presented the pairs of programs and allowed to take the necessary time to decide on the most readable one.
-
-\subsubsection*{Interview Analysis}
-
-We first transcribe each interview and then examine the broad distribution of the answers.
-Our goal is to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular.
-We follow-up with an open-coding procedure, highlighting the main themes and quoting the answers of the participants.
-We present these results in Section~\ref{sec:interview-results}.
-%Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.
-
-%\subsection{Third Study: Additional Survey and Interviews}
-
-% \rb{acho que aqui podemos incluir o segundo survey e a segunda
-%  rodada das entrevistas, explorando os \'{a}tomos adicionais}
-
-\subsection{Mining JavaScript Software Repositories Setting}\label{sec:method:msr}
-
-To understand how often the analyzed atom candidates appear in open source projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mine a set of GitHub open source repositories.
-To this end, we first collect the most popular GitHub repositories that are primarily written in JavaScript.
-We measure popularity using the project's stargazers.
-This metric, available through the GitHub API, represents the number of stars a project received from users of the platform.
-The same metric has been used in a number of previous studies as a proxy to estimate a project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}.
-We then select the top 100 most popular repositories and remove projects that do not reach the first quartile of the distribution of lines of code.
-
-After filtering out JavaScript project candidates, in the second step we build a curated dataset comprising the top \minedprojects repositories.
-Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}.
-Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research.
-The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
-All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars.
-We automate all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.
-
-We mine the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}.
-CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript).
-We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study).
-Finally, we compute some descriptive statistics to measure the prevalence of atoms of confusion in practice.
-
-\begin{table}[ht]
-  \centering
-   \caption{Some descriptive statistics about the projects used in the MSR study}
-  \begin{scriptsize}
- \begin{tabular}{rrrrr}
-   \toprule
-                       & Min.             & Median        & Mean             & Max. \\ \midrule
- Lines of Code         & \num{5543}       & \num{36161.5} & \num{111432.33}  & \num{1278405} \\
- Num. of Forks         & \num{1244}       & \num{6078}    & \num{8906.24}    & \num{68849} \\
- Num. of Contributors  & \num{6}          & \num{285.5}   & \num{533.44}     & \num{4047} \\
- Num. of Stars         & \num{23672}      & \num{34990}   & \num{46919.51}   & \num{310935} \\
-
-    \bottomrule
- \end{tabular}
- \end{scriptsize}
- \label{tab:projects-statistics} 
- \end{table}
+We also conduct a set of interviews with developers (Section~\ref{sec:method:interview}) to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ3).

### 2023-01-11 abstract.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

@@
 dynamically typed language whose popularity is growing in the most diverse application domains. In this paper
-we report the results of a mixed-methods research comprising two experiments, a set of interviews with practitioners,
-and a mining software repository (MSR) effort. 
+we report the results of a mixed-methods research comprising a mining software repositories (MSR) study, two experiments, and a set of interviews with practitioners.  
+Our MSR effort shows that atom candidastes are frequent and used intensively in 72 popular open-source JavaScript projects:
+four atom candidates appear in 90\% of them and two of them occur more than once for every 100 lines of code. This helps motivate the other three studies. 
 The results of both experiments suggest that two code patterns that have been previously observed to confuse \clang
@@
 indicate other constructs and idioms that merit investigation in the future. 
-Finally, our MSR effort reveals that the examined code patterns are frequent and used intensively in 72 popular open-source JavaScript projects:
-four atom candidates appear in 90\% of them and two of them occur more than once for every 100 lines of code.
 %Our findings provide significant evidence that some code patterns should be avoided if code is expected to be understood by a wide audience. 

### 2023-01-11 conclusion.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

@@
 This paper reports the results of a mixed-method research effort that allowed us to better understand the impact of atoms of confusion in JavaScript code.
-First, we conducted two independent experiments, which involve 210 JavaScript developers in total, asking them to predict the output of some code snippets (with and without atoms candidates).
+This involved a repository mining study, two experiments, and an interview study.
+The two independently-conducted experiments, which involved 210 JavaScript developers in total, asked them to predict the output of code snippets with and without atoms candidates.
 In at least one of the experiments, five atom candidates (Type Conversion, Change of Literal Encoding, Omitted Curly Braces, Pre-Increment, and Post- Increment) significantly impact program comprehension activities and can be considered atoms of confusion.

### 2023-01-11 introduction.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

--- 2021-09-27 1_Introduction.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Understanding code is an expensive and important task.
+In real-world software development, it is commonplace for developers to spend more time understanding code than actually writing it~\cite{CitationRequired}.Figuring out what parts of an existing system do by reading its source code is necessary for important software development tasks such as testing, debugging, and adding new functionality to a system.
 
-Source code misunderstandings can occur due to many reasons; for instance, the wrong choice of names for arguments, variables, and methods might hinder developers from understanding a piece of code~\cite{avidan:icpc2017}.
-Besides that, programming languages can also have specific idioms and constructs that may lead to misunderstandings.
-That is the case of atoms of confusion: small pieces of code that prevent developers from program comprehension, even though functionally equivalent alternatives exist and are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.
+Developers are often confused while reading unfamiliar code~\cite{Ebert:2021:ESC}.
+According to Hermans~\cite{ProgrammersBrain}, that confusion can stem from three sources: (i) lack of knowledge, i.e., not knowing what an element in the program does; (ii) lack of information, i.e., not knowing how a program element works; and (iii) lack of processing power, i.e., the inability to combine all the elements in a program and make sense of its execution in one's head.
+Even small pieces of code can be confusing for developers~\cite{Ajami:2017:SPI,DBLP:conf/sigsoft/GopsteinIYDZYC17} at first glance.
 
-Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability.
 For instance, consider the pair of JavaScript code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
-The code at the left-hand side contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: Logic as Control Flow and Post-increment.
-At the right-hand side, we present a clean version of the code, that is, without atoms of confusion.
-According to the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms Logic as Control Flow and Post-increment improved the accuracy of program understanding tasks in 41\% and 34\%, respectively.
+The code snippet on the left-hand side is arguably harder to understand than the one on the right-hand side.
+The former uses the logical AND operator (\texttt{\&\&}) beyond its lexical meaning~\cite{castor2018} to determine the control flow of the program.
+Furthermore, the expression has a potential side effect, due to the use of the post-increment operator (\texttt{s++}).
+The snippet on the right-hand side uses an \texttt{if} statement for control flow and increments the value of variable \texttt{s} using an assignment, making it easier to understand.
+Notwithstanding, confusion is a direct consequence of the experience of the person who is reading the code.
+A developer who customarily employs the idiom on the left-hand side may find it easier to understand.
 
 \begin{figure*}[thb]
-\noindent\begin{minipage}{.45\textwidth}  
+\noindent\begin{minipage}{.48\textwidth}  
 \begin{lstlisting}[language=JavaScript]
-res = properties.reduce((s, b) => {
+res = ps.reduce((s, b) => {
   b && s++
@@
 \end{minipage}\hfill
-\begin{minipage}{.45\textwidth}
+\begin{minipage}{.48\textwidth}
 \begin{lstlisting}[language=JavaScript]
-res = properties.reduce((s, b) => {
+res = ps.reduce((s, b) => {
   if(b) {
@@
 \end{minipage}
-\captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project. Line 2 on the left-hand side of the figure contains two
-  atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: Logic
-  as Control Flow and Post-increment.
-On the right-hand side, we present an alternative implementation without atoms of confusion.
-We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p.length})).}\label{fig:lst01}
+\captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project (left-hand side) and an alternative implementation (right-hand size).}\label{fig:lst01}
+% On the right-hand side, we present an alternative implementation without atoms of confusion.
+We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p.length})).}
 \end{figure*}
 
-Nonetheless, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java).
-This way, there is a knowledge gap with respect to the impact of such atoms in dynamic languages such as JavaScript.
-In this sense, because JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic capabilities and weaker type system, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa.
-Also, besides being the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}, JavaScript contains features that combined make the source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.
+%Line 2 on the left-hand side of the figure contains two atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: Logic as Control Flow and Post-increment.
+
+Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ,TheEyesDoNotLie} has attempted to elicit and catalog simple code patterns and language constructs that tend to confuse developers when reading them and for which there are less confusing, functionally-equivalent alternatives.
+These confusing constructs and patterns are called ``atoms of confusion''~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} when it is possible to experimentally ascertain that there is a less confusing alternative (otherwise they are called ``atom candidates'').
+The snippet on the left-hand side of Figure~\ref{fig:lst01} contains two atoms of confusion discussed in previous work~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} focusing on the \clang language: Logic as Control Flow and Post-increment.
+Previous work has identified atoms of confusion in two languages, \clang and Java.
+According to the studies of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms Logic as Control Flow and Post-increment from small \clang code snippets improved the ability of study participants to predict their outcomes by 41\% and 34\%, respectively.
+For small Java code snippets, Langhout and Aniche~\cite{Langhout:2021:ACJ} report improvements of 53\% and 46.27\%, respectively.
+These results show that atoms of confusion may be \textbf{challenging} for developers using these languages.
+In addition, a study~\cite{DBLP:conf/msr/GopsteinZFC18} with 14 large-scale open source projects written in \clang found that 4.38\% of the lines of code have an atom and their presence has a strong correlation with bug fixing commits and long code comments.
+These results show that atoms of confusion are \textbf{prevalent} in large systems.
+They are also \textbf{relevant}, even for experienced software developers.
+%According to the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms Logic as Control Flow and Post-increment improved the accuracy of program understanding tasks in 41\% and 34\%, respectively.
+
+%factors such as the wrong choice of names for variables and methods~\cite{avidan:icpc2017} and the use of uncommon language constructs~\cite{VEM2018} may confused developers might hinder developers from understanding a piece of code.
+Besides that, programming languages can also have specific idioms and constructs that may lead to misunderstandings.
+That is the case of atoms of confusion: small pieces of code that prevent developers from program comprehension, even though functionally equivalent alternatives exist and are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.
+
+The goal of this paper is twofold.
+First, it aims to investigate whether atoms of confusion that were identified in Java and C also cause confusion to JavaScript developers.
+The latter, in spite of the syntactic similarities to Java, is a dynamically-typed language that is arguably more related to Scheme than it is to Java~\cite{Eich:2018:BHJ}.
+\revised{Previous work has shown that, even for similar languages such as C\# and Java, programmer behavior may differ significantly~\cite{Cabral:2007:EHF,Ebad:2016:ECJ}.}
+% This way, there is a knowledge gap with respect to the impact of such atoms in dynamic languages such as JavaScript. In this sense, because 
+JavaScript's programming culture differs from that of languages such as \clang and Java, particularly due to its dynamic capabilities and weaker type system.
+\revised{Furthermore, it includes elements such as array and object destructuring and spread that have no parallel in Java or C.
+At the same time, the language is not so different that many} previously identified atoms cannot be represented in it.
+Second, the paper aims to verify whether other code patterns, not investigated in previous work, also tend to confuse JavaScript developers.
+Some of these code patterns are specific to JavaScript, e.g., object destructuring.
+% \castor{The previous sentence is really bad.} 
+%Second, we want to analyze whether these atoms also cause confusion when we control for developer experience.
+As mentioned before, experience seems to play a significant role in whether a certain construct or code pattern can be considered an atom of confusion or not.
+Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} report that, in their studies, experience is positively correlated with better results.
+In this paper, we aim to use an experimental design that controls for experience, so as to have greater confidence in our results.
+%, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. 
+%Also, besides being the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}, 
+%JavaScript contains features that combined make the source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.\castor{I think we should only say this if we explicitly point out examples of such features}.
 %, such as weakly typed language~\cite{DynamicBehaviorJS:PLDI}, object-oriented, functional,
@@
 
-In this way, to better understand the impact of atoms of confusion in JavaScript, in this paper we present a mixed-method research effort based on a survey, on interviews, and on a mining software repositories study.
-First, we surveyed 140 JavaScript developers.
-In the survey, their task was to predict the output of small code snippets, with half of them containing atoms of confusion.
-The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate).
-The presence of an atom candidate might also affect the time necessary for submitting a correct answer.
+\revised{We present the results of a mixed-method research effort based on a repository mining study, two experiments, and a set of interviews. 
+%, and a mining software repositories study.
+The repository mining study measures the prevalence of the atom candidates and helps us motivate the importance of studying these code patterns in more detail.
+We mined popular JavaScript projects from GitHub looking for the atom candidates that were investigated in both experiments.
+We found out
+%We found out that atoms frequently appear in practice. Finally, 
+%our MSR effort reveals 
+that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems.} 
+The two experiments ask the participants to predict the output of small code snippets, where some of them have atom candidates and others do not.
+These two experiments were designed and conducted independently, by different researchers, with different samples, and complementary methodologies; one uses a repeated measures design~\cite{Keselman:2001:ARM} whereas the other uses a Latin square design~\cite{Hunter-Experimenters}.
+Both our experiments suggest that two code patterns that have been previously observed to confuse C programmers also confuse JavaScript programmers (with statistical significance, after correction): the comma operator and assignments being used as values.
+In addition, pre- and post-increment operators, when used as values, omitted curly braces, changes of literal encoding, and type conversions have led to confusion in at least one of the two experiments.
+For all these cases effect sizes were either medium or high.
+In addition, we have discovered that object destructuring and automatic semicolon insertion are potential JavaScript-specific atoms that have not been discussed in previous work.
+
+To learn more about the misunderstandings in the snippets used in the experiments, we also interviewed 15 experienced professional developers.
+Participants preferred the code snippets without the atoms in 70\% of the cases.
+The frequency with which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code.
+
+%% The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate).
+For 5 of these atoms, the difference is statistically significant (accounting for p-value correction).
+Effect size is large for 4 atoms and medium for 1 of them.
+%% The presence of an atom candidate might also affect the time necessary for submitting a correct answer.
 For instance, the presence of the Comma Operator atom increases the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.
-To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers.
-Participants prefered the code snippets without the atom candidates in 70\% of the cases, therefore corroborating the survey findings.
+%% To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers.
+Participants preferred the code snippets without the atoms in 70\% of the cases, therefore corroborating the survey findings.
 The participants also discussed other idioms and constructs that might lead developers to misunderstand JavaScript code.
 
-The frequency in which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code.
-To understand the incidence of atoms, we mined popular JavaScript projects from GitHub and found that atoms of confusion frequently appear in practice.
-Indeed, four atom candidates are highly used in JavaScript code---the Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.
+%% The frequency in which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code.
+To understand the incidence of atoms, we mined popular JavaScript projects from GitHub and found that atoms frequently appear in practice.
+%% Our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems.
+Four atom candidates appear in 90\% of the analyzed projects and two of them, Implicit Predicate and Ternary Operator, occur more than once for every 100 lines of code in the dataset.
+%Indeed, four atom candidates are frequently used in JavaScript code---the Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.
 
-Altogether, the main contributions of this paper are:
+%\castor{I would rather remove the following. We don't even talk about the "library and infrastructure".}
+%Altogether, the main contributions of this paper are:
 
-\begin{itemize}
+%\begin{itemize}
 
-\item A mixed-method research effort---based on a survey, interviews, and mining software
-  repositories---about the impact of atoms of confusion on JavaScript (a
-    dynamic language).
+%\item A mixed-method research effort---based on two experiments, interviews, and a software
+%  repository mining study---about the impact of atoms of confusion on JavaScript programs.
+  % (a dynamic language), controlling for subject experience.
 
-  \item A comparison of the impact and prevalence of atoms of confusion in
-    JavaScript with what has been previously reported for other languages
-    (\clang, \cpplang, and Java).
+%  \item A comparison of the impact and prevalence of atoms of confusion in
+%    JavaScript with what has been previously reported for other languages
+%    (\clang, \cpplang, and Java).
+%    \item A library and infrastructure to mine atoms of confusion using 
+%    code queries (implemented in the .QL language). This library and 
+%    infrastructure are publicly available and might help researchers and 
+%    practitioners to automate program analysis at a large scale.
 
-    \item A library and infrastructure to mine atoms of confusion using 
-    code queries (implemented in the .QL language). This library and 
-    infrastructure are publicly available and might help researchers and 
-    practitioners to automate program analysis at a large scale.
+%\end{itemize}
 
-\end{itemize}
-
-As implications, our results can help researchers and practitioners to better understand atoms of confusion in JavaScript.
+%As implications, our results can help researchers and practitioners to better understand atoms of confusion in JavaScript.
 For example, our findings might alert the JavaScript community to avoid writing code with certain atoms of confusion.
 Moreover, our mixed study might help researchers and practitioners to design program transformation tools to remove atoms of confusion in JavaScript.
-They might prioritize atoms that cause a lot of confusion and are very common in the repositories, for instance.
+
+%They might prioritize atoms that cause a lot of confusion and are very common in the repositories, for instance.
 

### 2023-01-11 main.tex
improving threats to validity
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

@@
 
-\newcommand{\rqd}{What is the frequency of occurrence of atoms of confusion in practice (i.e., in open-source JavaScript projects)?}
+\newcommand{\rqd}{What is the frequency of occurrence of atom candidates in practice (i.e., in open-source JavaScript projects)?}
 
@@
 \lstdefinelanguage{CodeQL}[]{Java}{
-  morekeywords={select, from, where, not, and}
+  morekeywords={predicate, select, from, where, not, and, or}
 }
@@
 \input{methodology}
+\input{s04}
 \input{s01}
@@
 \input{s03}
-\input{s04}
 \input{discussion}

### 2023-01-11 methodology.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

@@
 \begin{enumerate}[(RQ1)]
+\item \revised{\rqd}
 \item \rqa 
@@
 % \item \rqc  
-\item \rqd
 \end{enumerate}
 
-By answering these research questions, we can either generalize or refute the findings about atoms of confusion already discussed in the literature
-(goal of research questions (RQ1) and (RQ2)). 
+\revised{Answering RQ1 enables us to better motivate the other two questions by showing that studying these code patterns has practical relevance.}
+Futhermore, by addressing research questions RQ2 and RQ3, we can either generalize or refute the findings about atoms of confusion already discussed in the literature. 
 In addition, answering the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice.
@@
 
+\revised{We perform a repository mining study (Section~\ref{sec:s04}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
+This helps us motivate the remaining three studies.}
 These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ1).
@@
 We also conduct a set of interviews with developers (Section~\ref{sec:s03}) to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2).
-Finally, we perform a repository mining study (Section~\ref{sec:s04}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
+%Finally, we perform a repository mining study (Section~\ref{sec:s04}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
 

### 2023-01-11 s01.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

@@
 \section{Repeated Measures Study}\label{sec:s01}
+
+\revised{The results of the repository mining study show that some of the atom candidates we aim to study occur very frequently.
+Taking that as a starting point, we conduct two experiments and an interview study to gauge their potential to cause confusion.
+Furthermore, since we are focusing on JavaScript code, we also analyze a number of atom candidates that are specific to this language in one of the experiments.}
 
@@
 Object Destructuring      &   \textbf{$<$ 0.0001*} & $\infty$ \\
-Type Conversion           &   \textbf{$<$ 0.0001*} & 12.0 \\
-Change of Literal Encoding   &   \textbf{$<$ 0.0001*} & 28.0 \\
-Comma Operator            &   \textbf{$<$ 0.0001*} & 11.5 \\
+Type Conversion           &   \textbf{$<$ 0.0001*} & 12.00 \\
+Change of Literal Encoding   &   \textbf{$<$ 0.0001*} & 28.00 \\
+Comma Operator            &   \textbf{$<$ 0.0001*} & 11.50 \\
 Lack of Indentation No Braces     &   \textbf{$<$ 0.0001*} &  11.0 \\
-Assignment As Value       &   \textbf{$<$ 0.0001*} &  9.3 \\
-Pre-Increment              &  \textbf{0.0003*} &    16.0 \\
-Omitted Curly Braces      &  \textbf{0.0009*} &  6.3 \\
-Array Destructuring       &  0.0023 &     7.5 \\
-Post-Increment              &  0.0225 &     5.5 \\
-Array Spread              &  0.0490 &    3.3 \\
-Object Spread             &  0.0636 &     2.8 \\
-Property Access           &  0.0636 &     2.8 \\
-Infix Operator Precedence &  0.1796 &     2.5 \\
-Arrow Function            &  0.2891 &     3.0 \\
-Arithmetic As Logic       &  0.3323 &  1.8 \\
-Automated Semicolon Insertion  &  0.3915 &  1.4 \\
-Logic As Control Flow     &  0.4421 &  1.5 \\
+Assignment As Value       &   \textbf{$<$ 0.0001*} &  9.30 \\
+Pre-Increment              &  \textbf{0.0003*} &    16.00 \\
+Omitted Curly Braces      &  \textbf{0.0009*} &  6.30 \\
+Array Destructuring       &  0.0023 &     7.50 \\
+Post-Increment              &  0.0225 &     5.50 \\
+Array Spread              &  0.0490 &    3.30 \\
+Object Spread             &  0.0636 &     2.80 \\
+Property Access           &  0.0636 &     2.80 \\
+Infix Operator Precedence &  0.1796 &     2.50 \\
+Arrow Function            &  0.2891 &     3.00 \\
+Arithmetic As Logic       &  0.3323 &  1.80 \\
+Automated Semicolon Insertion  &  0.3915 &  1.40 \\
+Logic As Control Flow     &  0.4421 &  1.50 \\
 %Constant Variables        &  0.5000 &     0.0 \\
-Implicit Predicate        &  0.6900 &  1.3 \\
-Lack of Indentation With Braces   &  0.7266 &  1.7 \\
-Conditional Operator      &  0.7905 &  1.3 \\
-Dead Unreachable Repeated &  1.0000 &     1.5 \\
-Repurposed Variables      &   1.0000 &     1.0 \\
+Implicit Predicate        &  0.6900 &  1.30 \\
+Lack of Indentation With Braces   &  0.7266 &  1.70 \\
+Conditional Operator      &  0.7905 &  1.30 \\
+Dead Unreachable Repeated &  1.0000 &     1.50 \\
+Repurposed Variables      &   1.0000 &     1.00 \\
 \bottomrule

### 2023-01-11 s02.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

@@
 Seven atom candidates required more time from participants to predict a correct response.
-For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet.
+For these eight atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet.
 In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less

### 2023-01-11 s04.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

@@
 
-To understand how often the analyzed atom candidates appear in open source JavaScript projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mine a set of GitHub open source repositories and compute the prevalence of atom candidates.
+%To understand how often the analyzed atom candidates appear in open source JavaScript projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mine a set of GitHub open source repositories and compute the prevalence of atom candidates.
 
-\subsection{Study Settings}
+We mine a set of GitHub open source repositories and compute how often these atom candidates occur.
+We do this with the goal of quantifying the prevalence of the investigated atom candidates.
+\revised{If they rarely occur in real-world systems, this would suggest that cannot be a significant source of confusion for developers.
+This study addresses our first research question (\emph{What is the frequency of occurrence of atom candidates in practice?}).}
+
+\subsection{Study Setting}
 
@@
 The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
-All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars.
+All projects in our dataset have at least \num[group-separator = {,}]{1244} forks and at least \num[group-separator = {,}]{23672} stars.
 We automate all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.
@@
                        & Min.             & Median        & Mean             & Max. \\ \midrule
- Lines of Code         & \num{5543}       & \num{36161.5} & \num{111432.33}  & \num{1278405} \\
- Num. of Forks         & \num{1244}       & \num{6078}    & \num{8906.24}    & \num{68849} \\
- Num. of Contributors  & \num{6}          & \num{285.5}   & \num{533.44}     & \num{4047} \\
- Num. of Stars         & \num{23672}      & \num{34990}   & \num{46919.51}   & \num{310935} \\
+ Lines of Code         & \num[group-separator = {,}]{5543}       & \num[group-separator = {,}]{36161.5} & \num[group-separator = {,}]{111432.33}  & \num[group-separator = {,}]{1278405} \\
+ Num. of Forks         & \num[group-separator = {,}]{1244}       & \num[group-separator = {,}]{6078}    & \num[group-separator = {,}]{8906.24}    & \num[group-separator = {,}]{68849} \\
+ Num. of Contributors  & \num[group-separator = {,}]{6}          & \num[group-separator = {,}]{285.5}   & \num[group-separator = {,}]{533.44}     & \num[group-separator = {,}]{4047} \\
+ Num. of Stars         & \num[group-separator = {,}]{23672}      & \num[group-separator = {,}]{34990}   & \num[group-separator = {,}]{46919.51}   & \num[group-separator = {,}]{310935} \\
 
@@
 
-Considering all JavaScript projects in our dataset, we found a total of \num{364873} atom candidates, though four atoms are responsible
+Considering all JavaScript projects in our dataset, we found a total of \num[group-separator = {,}]{364873} atom candidates, though four atoms are responsible
 for 92.97\% of this total: Implicit Predicate, Post Increment, Conditional Operator, and Omitted Curly Braces. 
-The first two, based on the results of Section~\ref{sec:repeated} and Section~\ref{sec:latin}, can be considered atoms of confusion.
+%The first two, based on the results of Section~\ref{sec:repeated} and Section~\ref{sec:latin}, can be considered atoms of confusion. 
 The remaining atom
-candidates comprise \num{25621} occurrences combined, 7.03\% of the total number of occurrences.
+candidates comprise \num[group-separator = {,}]{25621} occurrences combined, 7.03\% of the total number of occurrences.
 The Comma Operator and Arithmetic as Logic atoms are the ones that arise less
@@
 % study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. 
-% Atoms that occur 1000 times are classified as \textit{rarely used} while atoms with more than \num{10000} occurrences are classified as \textit{highly used}; \textit{commonly used} atoms occur in between.
+% Atoms that occur 1000 times are classified as \textit{rarely used} while atoms with more than \num[group-separator = {,}]{10000} occurrences are classified as \textit{highly used}; \textit{commonly used} atoms occur in between.
 We calculate the frequency of each atom per KLOC.
@@
 
-The results of our experiments and interviews suggest that Conditional Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to the Latin-square study).
+%%%%%%%%%%%%%%%%%%%%%% MAYBE the following can go to a discussion section. 
+%%%%%%%%%%%%%%%%%%%%%% 
+%The first two, based on the results of Section~\ref{sec:repeated} and Section~\ref{sec:latin}, can be considered atoms of confusion.
+
+%The results of our experiments and interviews suggest that Conditional Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to the Latin-square study).
 Nonetheless, the Post-Increment Expression and Automatic Semicolon Insertion atoms are listed in the top four sources of misunderstanding  (Table~\ref{tab:difference-correctness}), and also frequently appear in open source systems.

### 2023-02-13 s01.tex
Fixed a couple of very minor inconsistencies.

@@
 Arithmetic As Logic       &  0.3323 &  1.80 \\
-Automated Semicolon Insertion  &  0.3915 &  1.40 \\
+Automatic Semicolon Insertion  &  0.3915 &  1.40 \\
 Logic As Control Flow     &  0.4421 &  1.50 \\

### 2023-02-13 s02.tex
Fixed a couple of very minor inconsistencies.

@@
 \begin{tabular}{lrrr} \toprule
-  Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
+  Atom candidate& Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
   Comma Operator                &  28 &  65 & +132   \\ 

### 2023-02-21 abstract.tex
endereçando os comentários de R1 e R3 após o segundo round de revisões

@@
 In this paper we investigate the impact of atoms of confusion on understanding JavaScript code---a
-dynamically typed language whose popularity is growing in the most diverse application domains. In this paper
-we report the results of a mixed-methods research comprising a mining software repositories (MSR) study, two experiments, and a set of interviews with practitioners.  
+dynamically typed language whose popularity is growing in the most diverse application domains. We present
+the results of a mixed-methods research comprising a mining software repositories (MSR) study, two experiments, and a set of interviews with practitioners.  
 Our MSR effort shows that atom candidastes are frequent and used intensively in 72 popular open-source JavaScript projects:

### 2023-02-21 discussion.tex
endereçando os comentários de R1 e R3 após o segundo round de revisões

@@
 
-In the repeated measures study, we identified one JavaScript-specific atom, Object Destructuring.
-In the Latin square study, Automatic Semicolon Insertion was identified as another JavaScript-specific atom.
+In the repeated measures study, we identified one JavaScript-specific atom, \textbf{Object Destructuring}.
+In the Latin square study, \textbf{Automatic Semicolon Insertion} was identified as another JavaScript-specific atom.
 For both cases, p-values were much lower than the threshold and effect sizes were high.
@@
 Our interviewees favored the use of semicolon over Automatic Semicolon Insertion.
-Finally, as discussed in Section~\ref{sec:repeated:correct}, Array Destructuring exhibited a large effect size, in spite of the non-statistically significant difference (after correction).
+Finally, as discussed in Section~\ref{sec:repeated:correct}, \textbf{Array Destructuring} exhibited a large effect size, in spite of the non-statistically significant difference (after correction).
 As discussed elsewhere~\cite{Ellis:2010:EGE}, this is a scenario that hints at practical relevance, especially considering the very low p-value ($< 0.003$).
@@
 Some code patterns that have been identified as atoms in previous work were not confirmed in our experiments, at least from a correctness perspective.
-Comparing the results of the Latin square study to previous research, two atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed for \clang and \cpplang do not lead to a statistically significant impact on correctness in our study, Logic as Control Flow, and Conditional Operator.
-Conditional Operator was the only atom candidate in our interviews whose obfuscated version was considered less confusing than the corresponding clean version.
+Comparing the results of the Latin square study to previous research, two atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed for \clang and \cpplang do not lead to a statistically significant impact on correctness in our study, \textbf{Logic as Control Flow}, and \textbf{Conditional Operator}.
+\textbf{Conditional Operator} was the only atom candidate in our interviews whose obfuscated version was considered less confusing than the corresponding clean version.
 
-Logic as Control Flow is a more interesting case.
+\textbf{Logic as Control Flow} is a more interesting case.
 Considering the correctness perspective, none of our experiments has identified Logic as Control Flow as an atom of confusion.
@@
 Two other atoms have exhibited a statistically significant difference in time, in favor of their obfuscated versions (with a small effect size).
-One of them, Pre-Increment, has been confirmed as an atom of confusion in multiple experiments, i.e., participants are less confused by the clean versions.
+One of them, \textbf{Pre-Increment}, has been confirmed as an atom of confusion in multiple experiments, i.e., participants are less confused by the clean versions.
 These conflicting results reinforce the importance of conducting experiments that consider multiple response variables~\cite{Oliveira:2020:ECR}.
 In addition, it is clear that factors such as the experience of the participants and the selection of code snippets can have considerable impact on the results of studies about code readability~\cite{Feitelson:2021:CPC}.
+\revised{Table~\ref{tab:summary-of-findings} summarizes the correctness assessment results in our two experiments (columns Exp01 and Exp02) and the experiments of
+Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and Langhout and Aniche~\cite{Langhout:2021:ACJ}.}
+
+\begin{table}[htb]
+  \caption{Summary of the findings accross the different studies. Here we show 
+    only the atom candidates that have been explored in at least one
+    of our experiments. A {\bf True} value means that the experiment
+    confirms the code pattern as an atom candidate ({\bf False} otherwise).}
+  \label{tab:summary-of-findings}
+\begin{scriptsize}
+\begin{tabular}{lcccc}\toprule 
+Atom Candidate	                 & Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} & Langhout and Aniche~\cite{Langhout:2021:ACJ}  & Exp01     &  Exp02   \\ \midrule
+
+Change of Literal Encoding       &   TRUE   &   TRUE   &   TRUE    &   -      \\
+Assignment as Value              &   TRUE   &   -      &   TRUE    &   TRUE   \\
+Logic as Control Flow            &   TRUE   &   TRUE   &   FALSE   &   FALSE  \\
+Post-Increment                   &   TRUE   &   TRUE   &   FALSE   &   TRUE   \\
+Type Conversion                  &   TRUE   &   TRUE   &   TRUE    &   -      \\
+Conditional Operator             &   TRUE   &   FALSE  &   FALSE   &   FALSE  \\
+Infix Operator Precedence        &   TRUE   &   FALSE  &   FALSE   &   -      \\
+Comma Operator                   &   TRUE   &   -      &   TRUE    &   TRUE   \\
+Pre-Increment                    &   TRUE   &   TRUE   &   TRUE    &   FALSE  \\
+Implicit Predicate               &   TRUE   &   -      &   FALSE   &   TRUE   \\
+Repurposed Variable              &   TRUE   &   FALSE  &   FALSE   &   -      \\ 
+Omitted Curly Braces             &   TRUE   &   TRUE   &   TRUE    &   FALSE  \\
+Dead Unreachable Repeated        &   FALSE  &   FALSE  &   FALSE   &   -      \\
+Arithmetic as Logic              &   FALSE  &   FALSE  &   TRUE    &   FALSE  \\
+Object Destructuring             &   -      &   -      &   TRUE    &   -      \\
+Lack of Indentation no Braces    &   -      &   TRUE   &   TRUE    &   -      \\
+Array Destructuring              &   -      &   -      &   FALSE   &   -      \\
+Array Spread                     &   -      &   -      &   FALSE   &   -      \\
+Object Spread                    &   -      &   -      &   FALSE   &   -      \\
+Property Access                  &   -      &   -      &   FALSE   &   -      \\
+Arrow Function                   &   -      &   -      &   FALSE   &   -      \\
+Automatic Semicolon Insertion    &   -      &   -      &   FALSE   &   TRUE   \\
+Lack of Indentation with Braces  &   -      &   -      &   FALSE   &   -      \\ \bottomrule
+
+\end{tabular}
+\end{scriptsize}
+\end{table}
+
+\revised{It is not feasible to combine the results of the statistical assessments coming from our two experiments---because their design differs significantly.
+  Nonetheless, we can reason about the relationships between correct/wrong answers in both experiments and combine the results of the exploratory data analysis
+  we presented in Sections~\ref{sec:repeated:correct} and~\ref{sec:latin:correct}.
+  Considering the ten atoms explored in our two experiments, for six atom candidates the number of correct answers for the clean
+  version of the code snippets is at least 25\% higher than the number of correct answers for the obfuscated version of the code (see Table~\ref{tab:merge-exploratory-study}).}
+
+\begin{table}
+  \caption{Results of the exploratory data analysis for the experiments (Repeated Measurea and Latin Square)}
+  \label{tab:merge-exploratory-study}
+  \begin{scriptsize}
+  \begin{tabular}{lccccc} \toprule
+                                 & \multicolumn{2}{c}{Repeated Measures} & \multicolumn{2}{c}{Latin Square} & \\  
+  Atom Candidate                 &   Obfuscated   &   Clean   &   Obfuscated   &   Clean   &  $\delta$ (\%)    \\ \midrule
+Comma Operator                   &      8         &   29      &       28       &     65    &   161.11  \\
+Automatic Semicolong Insertion   &     15         &   21      &       32       &     68    &    89.36  \\
+Assignment as Value              &     18         &   43      &       56       &     68    &    50.00  \\
+Post-Increment                   &      8         &   17      &       48       &     64    &    44.64  \\
+Omitted Curly Braces             &     35         &   51      &       47       &     58    &    32.92  \\
+Pre-Increment                    &     13         &   28      &       50       &     53    &    28.57  \\
+Implicit Predicate               &     41         &   44      &       58       &     68    &    13.13  \\
+Conditional Operator             &     60         &   62      &       60       &     66    &     6.66  \\
+Logic as Control Flow            &     33         &   28      &       41       &     48    &     2.70  \\ 
+Arithmetic as Logic              &     52         &   47      &       64       &     63    &    -5.17  \\ \bottomrule
+  \end{tabular}
+\end{scriptsize}  
+\end{table}
+
+\revised{Besides that, the results of the two experiments reveal findings that diverge. That is, for six atom candidates (out of ten explored in both experiments),
+  the results of the statistical tests differ. Three main reasons might explain the differences: (a) the code snippets used in the experiments are different,
+  (b) the participants of the experiments are different, and (c) the design of the experiments and thus the analysis procedures are different.
+  As such, additional research would be necessary to understand the impact of these items on the results of empirical assessments of atoms of confusion in general.}
+
+\revised{Our findings here suggest the need for replication studies in this field. Here we left an additional question:
+  What would be the results of the original studies on atoms of confusion if an independent group reproduces the original research?
+Although the divergences in the two experiments might threaten our study's conclusions, we complement the experiments with a qualitative study based on interviews.
+We return to this issue of reporting the results
+of different experiments in Section~\ref{{sec:whytwo}}.}
 
@@
 
-Altogether, we answer our first research question as follows.
+The results of the first study (mining open source
+JavaScript repositories) provides evidence that,
+although atoms of confusion compromise program
+comprehension, they frequently appear in open
+source JavaScript projects. In particular,
+seven, out of 10 atoms considered
+  in our study, appear in more than 50\% of
+the 72 projects we analyzed. Furthermore, at least two of them are used intensively, more than once for every 200 lines of code. In summary, the first study
+allows us to answer the third research
+question.
 
 \begin{mh}
-  {\bf Answer to RQ1:} Our experiments using repeated measures and the Latin square designs give evidence that some of the atom candidates in \clang and \cpplang programs that also exist in JavaScript correspond to a source of misunderstanding in
-  JavaScript code. 
+  {\bf Answer to RQ1:} The MSR study reveals that
+  several atom candidates explored in our research
+  appear frequently in practice, and cleaning up the use of 
+  Post-Increment/Decrement and the Automatic Semicolon Insertion
+  might improve the readabiliy of JavaScript code substantially. 
 \end{mh}
 
-The results of the interview study complement the understanding of atoms of confusion because the participants make clear the existence of a trade-off between code comprehension and other quality attributes.
-For instance, most of the participants prefer the version of the code with the Conditional Operator, even though they agree that its use might contribute to the misunderstanding of JavaScript code, particularly when novices are maintaining the codebase.
-The participants of the interview study also
+Altogether, the results of our experiments allow us
+to answer the second research question of this
+study as follows.
+
+\begin{mh}
+  {\bf Answer to RQ2:} Our experiments using repeated measures and the Latin square designs give evidence that some of
+  the atom candidates in \clang and \cpplang programs that also exist in JavaScript correspond to a source of misunderstanding in
+  JavaScript code. Two JavaScript-specific code patterns hava been identified as atoms of confusion by at least
+  one of our experiments: Object Descructuring and Automatic Semicolon Insertion. 
+\end{mh}
+
+The results of the interview study complement the understanding of atoms of confusion because the participants make clear the existence of a trade-off between
+code comprehension and other quality attributes. For instance, most of the participants prefer the version of the code with the
+Conditional Operator, even though they agree that its use might contribute to the misunderstanding of JavaScript code,
+particularly when novices are maintaining the codebase. The participants of the interview study also
 mentioned other possible sources of misunderstanding in JavaScript,
@@
 \begin{mh}
-  {\bf Answer to RQ2:} The qualitative analysis of the
+  {\bf Answer to RQ3:} The qualitative analysis of the
   interviews supports the results of our experiments---
@@
 %% \end{mh}
-
-The results of the third study (mining open source
-JavaScript repositories) provides evidence that,
-although atoms of confusion compromise program
-comprehension, they frequently appear in open
-source JavaScript projects. In particular,
-seven, out of 10 atoms considered
-  in our study, appear in more than 50\% of
-the 72 projects we analyzed. Furthermore, at least two of them are used intensively, more than once for every 200 lines of code. In summary, the third study
-allows us to answer the third research
-question.
-
-\begin{mh}
-  {\bf Answer to RQ3:} The MSR study reveals that
-  several atom candidates explored in our research
-  appear frequently in practice, and cleaning up the use of 
-  Post-Increment/Decrement and the Automatic Semicolon Insertion
-  might improve the readabiliy of JavaScript code substantially. 
-\end{mh}
 

### 2023-02-21 methodology.tex
endereçando os comentários de R1 e R3 após o segundo round de revisões

@@
 
-\revised{Answering RQ1 enables us to better motivate the other two questions by showing that studying these code patterns has practical relevance.}
+\revised{Answering RQ1 enables us to better motivate the other two questions by showing that studying these code patterns has practical relevance and discussing how often they occur in practice.}
 Futhermore, by addressing research questions RQ2 and RQ3, we can either generalize or refute the findings about atoms of confusion already discussed in the literature. 
-In addition, answering the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice.
-Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions.
+In addition, answering the third question allows us to enrich existing catalogs about atom candidates.
+Answering these research questions might also lay the foundations for the implementation of tools that can automatically transform code into cleaner versions.
 %---though we postpone these results to future research work.
@@
 We conduct a mixed-methods study to answer these questions.
-It includes two independently designed and conducted experiments: a repeated measures/within subjects study (Section~\ref{sec:s01}) and one using a Latin square, between subjects, counterbalancing experimental design to control for subject experience (Section~\ref{sec:s02}).
+\revised{We perform a repository mining study (Section~\ref{sec:s04}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
+This helps us motivate the remaining three studies.}
+Our research also includes two independently designed and conducted experiments: a repeated measures/within subjects study (Section~\ref{sec:s01}) and one using a Latin square, between subjects, counterbalancing experimental design to control for subject experience (Section~\ref{sec:s02}).
 By using two different experiments, conceived by different researchers, with different designs, and conducted with different participants, we hope to rigorously verify the impact of atoms of confusion in code comprehension activities.
@@
 
-\revised{We perform a repository mining study (Section~\ref{sec:s04}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
-This helps us motivate the remaining three studies.}
-These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ1).
+These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ2).
 They investigate whether or not programs that contain atom candidates tend to produce more misunderstanding for programmers trying to predict their outcome.% We leverage two independently-developed and conducted experiments with different designs to more
-We also conduct a set of interviews with developers (Section~\ref{sec:s03}) to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2).
+We also conduct a set of interviews with developers (Section~\ref{sec:s03}) to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ3).
 %Finally, we perform a repository mining study (Section~\ref{sec:s04}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.
 
-Table~\ref{tab:atomsBoth} presents the atom candidates that we explore in all stduies, while Table~\ref{tab:atomsRepeated} presents the additional 
+Table~\ref{tab:atomsBoth} presents the atom candidates that we explore in all studies, while Table~\ref{tab:atomsRepeated} presents the additional 
 ones that only the repeated measures study analyzes.

### 2023-02-21 s01.tex
endereçando os comentários de R1 e R3 após o segundo round de revisões

@@
 Furthermore, we conduct an informal search of \revised{JavaScript questions on StackOverflow about constructs that are not typical of other languages, such as array and object spread, solutions to Code Golf~\footnote{https://codegolf.stackexchange.com/} problems, and code style guides\footnote{AirBNB JavaScript Style Guide: https://github.com/airbnb/javascript}\footnote{Google JavaScript Style Guide: https://google.github.io/styleguide/jsguide.html} to identify new atom candidates.
-We leverage previous work~\cite{castor2018} on the identification of atom candidates to guide us in this informal search.} Overall, the repeated measures study analyzes 24 atoms, 15 from the original study and 9 that are new (see Tables~\ref{tab:atomsBoth} and ~\ref{tab:atomsRepeated}.
+We leverage previous work~\cite{castor2018} on the identification of atom candidates to guide us in this informal search.} Overall, the repeated measures study analyzes 24 atoms, 15 from the original study and 9 that are new (see Tables~\ref{tab:atomsBoth} and ~\ref{tab:atomsRepeated}).
 
@@
 As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the questionnaire would take on average.
-Based on the feedback of the participants of the pilot, we present only one obfuscated and one clean version of each atom candidate, to reduce to total time of the study and potentially increase participation.% Initially, our plan was to present two of each.
+Based on the feedback of the participants of the pilot, we present only one obfuscated and one clean version of each atom candidate, to reduce the total time of the study and potentially increase participation.% Initially, our plan was to present two of each.
 

### 2023-02-21 s02.tex
endereçando os comentários de R1 e R3 após o segundo round de revisões

@@
 
-\subsection{Experimental design}
+\subsection{Experimental design}\label{sec:latin-design}
 
@@
 The order in which they are presented is randomized.
-By doing this, we seek to minimize the chances of subjects being aware that the current listing they are analyzing contains (or not) atoms of confusion.
-That is, each subject should indicate what would be the outcomes of the programs, some of them having atoms of confusion (while other programs do not).
+By doing this, we seek to minimize the chances of subjects being aware that the current listing they are analyzing contains (or not) atom candidates.
+That is, each subject should indicate what would be the outcomes of the programs, some of them having obfuscated code (while other programs do not).
 Since participants are not exposed to obfuscated and clean versions corresponding to the same atom candidate, learning effect is not possible.
@@
 
-\subsection{Results}\label{sec:latin}
+\subsection{Results}\label{sec:latin-results}
 
@@
 
-\subsubsection{Correctness Analysis}
+\subsubsection{Correctness Analysis}\label{sec:latin:correct}
 % \castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me to which rate this refers.}
@@
 {\bf Exploratory Data Analysis.}
-As mentioned in Section~\ref{sec:meth:survey}, each participant in this experiment evaluated ten code snippets, from which five were in their obfuscated versions, whilst the other five contained clean versions of the code snippets (i.e., without the atom candidates).
+As mentioned in Section~\ref{sec:latin-design}, each participant in this experiment evaluated ten code snippets, from which five were in their obfuscated versions, whilst the other five contained clean versions of the code snippets (i.e., without the atom candidates).
 As in the repeated measures study, the participants should provide the expected outcomes of the code snippets.

### 2023-02-21 s03.tex
presenting an overview of the studies
review os s04
endereçando os comentários de R1 e R3 após o segundo round de revisões

@@
-\section{Interview Study}\label{sec:method:interview}
+\section{Interview Study}\label{sec:s03}
 
-To complement the two experiments, we perform semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding programs containing atoms of confusion (research question RQ2).
+To complement the two experiments, we perform semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding programs containing atoms of confusion (research question RQ3).
 We also ask each participant if they know of any other JavaScript-specific construct or idiom they think is likely to make the code hard to understand.
@@
 In this section we present the results of the interviews with the practitioners.
-We contrast the findings of the experiments of Sections~\ref{sec:repeated} and~\ref{sec:latin} with the opinion of software developers about their preferences regarding the obfuscated or clean versions of the code snippets.
-For the interview study, we only considered atoms that were analyzed in both the repeated measures (Section~\ref{sec:repeated}) and the latin square (Section~\ref{sec:latin}) studies.
+We contrast the findings of the experiments of Sections~\ref{sec:s01} and~\ref{sec:s02} with the opinion of software developers about their preferences regarding the obfuscated or clean versions of the code snippets.
+For the interview study, we only considered atoms that were analyzed in both the repeated measures (Section~\ref{sec:s01}) and the latin square (Section~\ref{sec:s02}) studies.
 
@@
 
-Supporting the results of Sections~\ref{sec:repeated} and~\ref{sec:latin}, Table~\ref{tab:interview-results1}
+Supporting the results of Sections~\ref{sec:repeated} and~\ref{sec:latin-results}, Table~\ref{tab:interview-results1}
  shows that for eight out of the
@@
 
-\noindent\begin{minipage}{.45\textwidth}
+%\noindent\begin{minipage}{.45\textwidth}
 \begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Conditional Operator} atom).]
@@
 \end{lstlisting}
-\end{minipage}\hfill
-\begin{minipage}{.45\textwidth}
+%\end{minipage}\hfill
+%\begin{minipage}{.45\textwidth}
 \begin{lstlisting}[language=JavaScript, caption=\emph{Right-hand side} (without the atom).]
@@
 \end{lstlisting}
-\end{minipage}
+%\end{minipage}
 \caption{Example of a pair of code snippets used in the interview pertaining to the \emph{Conditional Operator} atom candidate}
@@
 \noindent
-None of the two studies found a significant difference in correctness between clean and obfuscated versions of snippets related to this atom candidate.
-However, in the latin square study the participants analyzing clean versions of the snippets required significantly less time.
+None of the two experiments we detailed in Sections~\ref{sec:s01} and~\ref{sec:s02} found a significant difference in correctness between clean and obfuscated versions of snippets related to the Logic as Control Flow atom candidate.
+However, in the latin square study the participants analyzing clean versions of the snippets required significantly less time to predict a correct answer.
 

### 2023-02-21 s04.tex
endereçando os comentários de R1 e R3 após o segundo round de revisões

@@
 
-After filtering out JavaScript project candidates, we build a curated dataset comprising the top \minedprojects repositories.
+After filtering out JavaScript project candidates, we build a curated dataset comprising a total of \minedprojects repositories.
 Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}.
@@
 where not e.getTopLevel().isMinified() 
-select e, e.getLocation().getStartLine() as Location,e.getLocation().getFile()
+select e, e.getLocation().getStartLine() as Location,
+       e.getLocation().getFile()
 \end{lstlisting}
@@
 We mined \minedprojects open source JavaScript repositories to understand how
-often atoms arise in real software. As in the interview study, this study focused on atoms that were investigated in both
-experiments (detailed in Section~\ref{sec:s01} and Section~\ref{sec:s02}).
+often atoms arise in real software. This study focused on atoms that were investigated in
+previous research~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and that we explored in the remaining
+studies of this research (detailed in Section~\ref{sec:s01}, Section~\ref{sec:s02}, and Section~\ref{sec:s03}).
 
-Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19} which investigate
+Similarly to previous studies~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/msr/GopsteinZFC18} which investigate
 the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
@@
 On the other hand, Arithmetic as Logic and Comma Operator occur less than once for every 50 KLOC.
-
+\revised{Given the frequency of these code patterns in open-source JavaScript programs, it is worth it to investigate their impact on understanding JavaScript code.
+This is the goal of the studies we present in the following sections.}
 %In our dataset of JavaScript projects, 

### 2023-02-22 letters/response-letter02.tex
adicionando a response-letter e pequenos ajustes na secao de discussao

\documentclass{letter}

\usepackage{xcolor}

\usepackage{hyperref}
\signature{Rodrigo Bonif\'{a}cio and Fernando Castor}
\address{Computer Science Department \\ University of Bras\'{i}lia \\ Brazil}

\begin{document}
\newcommand{\comment}[1]{\emph{#1}}

\begin{letter}{Journal of Systems and Software Editors-in-Chief}

  \opening{}

  We thank the associate editor and reviewers again for their valuable feedback.
  We hope we have correctly addressed the reviewers' main concerns in this second round.
  In what follows, we present our actions to handle the reviewers' comments and answer some of the reviewers' questions.

{\bf R1 argued that we have purposefully ignored some of his/her comments}

We are really sorry about that. Nonetheless, we have changed
the whole structure of the paper in the previous revision, in order
to address one of the main concerns R1 raised in the first round.
We used to believe that, in doing so, we have also handled
most of the R1 comments. That was not true, and we hope that we
have fixed such a mistake in the new version of this paper.

In what follows, we present a detailed response on how we have addressed
the R1 comments in the current version of the paper.

\begin{itemize}

\item \comment{What concerns me the most are the paper's contributions. In some ways, the paper is like an extended work from C to JavaScript. There are only a few new atom candidates being researched that are specific to JavaScript. Given this concern, I believe it is necessary to reorder the RQs in order to boost the work's contributions. RQ3 should be the first RQ to uncover the prevalence of atom candidates in real-world js programming. There would be no need for this study to be done if these atom candidates were infrequently observed in the real world. Based on this discovery, the other experiments could be continued.}

{\bf Action.} We have re-organized the studies as suggested by the reviewer. The repository mining study now comes first and is presented as a motivating study for the remaining studies, highlighting their relevance. Besides the re-ordering of the sections, changes have been applied to multiple parts of the paper. For instance, we present a better motivation for this work in the first section (Introduction).

\item \comment{Pages 1-2: The research's conclusions and findings, which are summarized in the abstract, are too detailed and somewhat tedious. I would suggest that the authors revisit the answers to the three RQs and rewrite the back half of the abstract.}

{\bf Action.} We have cut about 85 words from the abstract, specifically from the part about the contributions. We avoided cutting more because the paper presents the results of four different studies and the abstract to a certain extent reflects that.

\item \comment{age 6: At the beginning of section 3, I expect a figure introducing the entire workflow. Given that the paper includes several experiments, it would be helpful to use a figure to create a broad overview of the paper.}

{\bf Action.} We discussed this comment for a long time and we agree that a workflow would be necessary to understand the first version of the paper better. Thanks for raising this issue. However, we believe that, in the current organization of the paper, a workflow is not necessary since the order of the studies is flowing better now.

\item \comment{As stated by the authors on this page,  for the second experiment, they only use ten of the atom candidates that were used in the first experiment. Table 2 shows that 14 atom candidates were not studied in the second experiment. Can the authors explain how the atom candidate was chosen for the second experiment?}

{\bf Answer.} In the second experiment, we have tackled the nine atoms that have been validated for the C language in the original paper by Gopstein et al. (2017). In addition, we have studied a single atom that is specific to JavaScript programs, Automated Semicolon Insertion. We explain this in the beginning of Section 6. The reason for analyzing a smaller number of atoms is to attempt to increase the number of participants, since a large number of tasks in a study usually results in subjects not participating or dropping out of the study. This strategy was successful, as can be seen by the considerably larger number of participants in the Latin Square Study, twice as large as that of the repeated measures study. We emphasize that the two experiments were designed and conducted independently, by two independent groups without knowledge about the other. This explains why some atom candidates are in one study and not the other.

\item \comment{It seems that developers need more time to submit the correct answer using the obfuscated versions of two atom candidates than the clean versions. Is there a rationale behind this?}

{\bf Answer.} This could only be observed with statistical significance for three of the atoms, after p-value correction. Effect sizes are also mostly very small and they are not affected by sample sizes, differently from the p-values. This combination of results makes it difficult to formulate hypotheses beyond the obvious: that obfuscated snippets are slightly more confusing than the cleaned ones, with few exceptions.

\item \comment{I do not think it is a good idea to answer all of the RQs on one page. Since the experiments involve so many details, it is difficult for readers to recall all of them when reading these answers. I would suggest treating RQs as separate sections, with the answers to RQ serving as a summary of the section. They could have a look at the following paper: Li, H., Shang, W. \& Hassan, A.E. Which log level should developers choose for a new logging statement?. Empir Software Eng 22, 1684-1716 (2017).}

{\bf Action.} We thank the reviewer for the comment. We have restructured the sections of the paper so that each study is wholly contained in a single section, both the methodology and the results. This addresses a comment from another review but we believe it improves the cohesiveness of the sections. As for the summaries of the responses, we would prefer to keep it in a single place, though, for ease of reference. The goal is to enable readers to have a high-level overview of the results in a single place.

\item \comment{Minor Issues}

\begin{itemize}
  \item \comment{Page 6: the analyzed code patternson -> the analyzed code patterns on}
    \begin{itemize}
     \item Fixed. Thanks.
    \end{itemize}  
  \item \comment{Page 13: pilot.Undergraduate -> pilot. Undergraduate} 
    \begin{itemize}
      \item Fixed. Thanks. 
    \end{itemize}
  \item \comment{Thousand separators are missing in numbers.}
    \begin{itemize}
      \item Fixed. Thanks. We gave a pass through the entire paper but only found numbers requiring this correction in Section 4 (the MSR study). Please let us know if we have missed any others. 
    \end{itemize}
 \item \comment{I would suggest rounding the decimals to the same number of decimal places.}
   \begin{itemize}
   \item We have partially incorporated this suggestion. We focused mainly on the tables, where we believe the uniformity matters more for presentation and interpretation purposes. For p-values, we kept the four decimal places because it is relevant to be precise, as some p-values are very small. For all the other numbers where decimal places matter, we used two. In the text itself, we did not change the number of decimal places because we do not think much is lost. For example, for a p-value of 0.007, we prefer to use 0.007 instead of saying 0.0070 just to preserve uniformity.
  \end{itemize}   
 \item \comment{Page 25: Seven atom candidates are mentioned in the text, but the table lists eight candidates.}
  \begin{itemize}
    \item Fixed. Thanks.
  \end{itemize}  
\end{itemize}

\end{itemize}

{\bf R2 recommended our paper for publication}

Thank you for your comments and feedback. Your approval means a lot to our study.

{\bf R3 concerns about the presentation of both experiments (RM/LS)}

{\bf Action.} We included two additional tables in Section 8.1 (Tables 13 and 14) that
aim to better compare the results of the two experiments. We also
discuss possible reasons that might explain the different findings at the end of Section 8.1.
We still believe that reporting the results of two independent experiments is worth. However,
if the reviewers and editors agree that we should present only the results of one
experiment, we would accept the recommendation and remove our second experiment
based on the Latin Square design.

{\bf R3 concerns about the replication package}

We improved the replication package to address those comments. Thanks so much for the
recommendations.

{\bf R3 concerns about the code snippets}

\comment{It might be worth mentioning some summary of the following in the paper:}

\begin{itemize}
 \item \comment{Comma Operator A - Only reporting the sum of the values allows certain types of misunderstanding to go unreported. E.g. what if a subject thinks the second line operates like multiple assignment assigning 1 to V1 and 2 to V2? The output would be 3 just as the correct interpretation.}
 \item \comment{Ternary Operator A - I would love to see an investigation of the confusingness of === vs ==. I know there's a difference, but I have no idea what it is.}
 \item \comment{Implicit Predicate A - This snippet contains the modulo operator which is likely an atom of confusion as well as per [Gopstein 2020 - Thinking Aloud \ldots]}
\item  \comment{Logic Control Flow A - This snippet contains both Pre-Increment and Implicit Predicate atoms as well as Logic Control Flow. This violates the concept of "minimality" in an atom of confusion.}
\end{itemize}

Many thanks for raising these questions. It is hard to devise code snippets that can avoid
all different sources of misunderstanding. Regarding the third and fourth bullets above,
we have updated our threats to the validity section (page 41, paragraph starting with ``We manually devised \ldots'')
to recognize this issue as a possible threat to the validity of our study.

{\bf R3 reflections on Section 8.3 (``Why multiple studies'')}

We agree with R3's clever observations and changed several sentences in this section.

\closing{Yours Faithfully,}

\end{letter}
\end{document}

### 2023-02-22 discussion.tex
minor
adicionando a response-letter e pequenos ajustes na secao de discussao

@@
 We return to this issue of reporting the results
-of different experiments in Section~\ref{{sec:whytwo}}.}
+of different experiments in Section~\ref{sec:whytwo}.}
 
@@
 We present two independently-designed and conducted studies, with diverse code snippets, samples, and experimental designs, to provide stronger evidence.
-We believe that atom candidates confirmed to be atoms of confusion by both experiments arguably provide more reliable evidence of the potential to cause confusion of these code patterns for a wide range of developers, in diverse scenarios.
+Due to the independence of the studies, the experiment using the Latin Square design does not cover all atom candidates the Repeated Measures experiment explores.
+Nonetheless, we believe that atom candidates confirmed to be atoms of confusion by both experiments arguably provide more reliable evidence of the potential to cause confusion of these code patterns for a wide range of developers, in diverse scenarios.
 That is the case for the Comma Operator and Assignment as Value atoms.
@@
 This can be mitigated by separating examples pertaining to the same atom candidate, but not eliminated.
-\item [--] Fatigue effect, since participants have to examine more code snippets. This was exacerbated in our case, due to the larger number of atom candidates.
+\item [--] Fatigue effect, since participants have to examine more code snippets.
+This was exacerbated in our case, due to the larger number of atom candidates.
+Although we could have mitigated this problem by limiting the number of atom candidates shown to the subjects, our decision was to ask the participants to evaluate pairs of code snippets for every atom candidate---and thus collecting a large number of answers.
 \end{itemize}}
@@
 \begin{itemize}
-\item[+] Data is independent. 
+\item[+] Although there is some data dependency (since we collect different observations from a given subject),	analyzing the data in the Latin square	design allows us to isolate confounding variables better.
 \item[+] There is no possibility of learning/practice effect and the fatigue effect is reduced since each participant is only exposed to one version of each code snippet. 
-There is no possibility of learning/practice effect because each participant is only exposed to one treatment.
 \end{itemize}}
@@
 \begin{itemize}
-  \item[-] Effect sizes tend to be lower. 
+  \item[-] Statistical significance tends to be lower. 
   \item[-] Does not account for variance within groups.
 \end{itemize}
-The use of a Latin square mitigates some of the disadvantages intrinsic to a between subjects design.
+The use of a Latin square mitigates some of the disadvantages intrinsic to within subjects design.
 For example, since less work is required from participants, it was possible to obtain more data for the latin square study than we could get for the repeated measures study.

### 2023-02-22 threats.tex
adicionando a response-letter e pequenos ajustes na secao de discussao

@@
 Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code.
+
+\revised{We manually devised sets of code snippets to use in our experiments.
+These code snippets should have been designed considering the concept of ``minimality''~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}
+  to avoid other sources of misunderstanding.
+After running the experiments, we observed that this concept was not well addressed in a few cases.
+In particular, the code snippet contained the atom candidate Logic as Control Flow in our second experiment (based on the Latin Square design) has additional atom candidates.
+This could have inflated the number of wrong answers in the obfuscated version of this code pattern.
+However, the experiment results show the contrary, and the findings of the second experiment do not confirm Logic as Control Flow as an atom of confusion.
+Further analysis should be
+  conducted to understand the effect of multiple atoms in the same code snippet on program comprehension tasks.}
 

### 2023-02-23 letters/response-letter02.tex
submitted version

@@
 
-{\bf Action.} We discussed this comment for a long time and we agree that a workflow would be necessary to understand the first version of the paper better.
-Thanks for raising this issue.
-However, we believe that, in the current organization of the paper, a workflow is not necessary since the order of the studies is flowing better now.
+{\bf Action.} As we mentioned, to address this issue raised by the reviewer, we reorganized several parts of the paper and considered the studies in different sections.
+We now believe that the settings of each study are better organized.
+Thanks for raising this issue!
+We believe that a workflow is not necessary anymore.
 

### 2023-02-23 discussion.tex
submitted version

@@
                                  & \multicolumn{2}{c}{Repeated Measures} & \multicolumn{2}{c}{Latin Square} & \\  
-  Atom Candidate                 &   Obfuscated   &   Clean   &   Obfuscated   &   Clean   &  $\delta$ (\%)    \\ \midrule
+  Atom Candidate                 &   Obfuscated   &   Clean   &   Obfuscated   &   Clean   &  $\Delta$ (\%)    \\ \midrule
 Comma Operator                   &      8         &   29      &       28       &     65    &   161.11  \\

### 2023-04-04 discussion.tex
cr candidate

@@
 In addition, it is clear that factors such as the experience of the participants and the selection of code snippets can have considerable impact on the results of studies about code readability~\cite{Feitelson:2021:CPC}.
-\revised{Table~\ref{tab:summary-of-findings} summarizes the correctness assessment results in our two experiments (columns Exp01 and Exp02) and the experiments of
+\revised{Table~\ref{tab:summary-of-findings} summarizes the correctness assessment results in our two experiments: Repeated Measures Study (column Exp01) and Latin Squares Study (column Exp02) and the experiments of
 Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and Langhout and Aniche~\cite{Langhout:2021:ACJ}.}
@@
 \begin{table}
-  \caption{Results of the exploratory data analysis for the experiments (Repeated Measurea and Latin Square)}
+  \caption{Results of the exploratory data analysis for the experiments (Repeated Measures and Latin Squares design). The Obfuscated and
+    Clean columns show the number of correct answers for the obfuscated and clean versions of the code samples in both experiments.
+We compute the {\bf total} number of correct answers for the obfuscated
+    version of the code in both experiments (TO) and the {\bf total} number of correct answers for the clean version of the code in
+    both experiments (TC). The $\Delta$ column is given by $\Delta = \frac{TC - TO}{TO}$.}
   \label{tab:merge-exploratory-study}
   \begin{scriptsize}
-  \begin{tabular}{lccccc} \toprule
+  \begin{tabular}{lcccc|ccc} \toprule
                                  & \multicolumn{2}{c}{Repeated Measures} & \multicolumn{2}{c}{Latin Square} & \\  
-  Atom Candidate                 &   Obfuscated   &   Clean   &   Obfuscated   &   Clean   &  $\Delta$ (\%)    \\ \midrule
-Comma Operator                   &      8         &   29      &       28       &     65    &   161.11  \\
-Automatic Semicolong Insertion   &     15         &   21      &       32       &     68    &    89.36  \\
-Assignment as Value              &     18         &   43      &       56       &     68    &    50.00  \\
-Post-Increment                   &      8         &   17      &       48       &     64    &    44.64  \\
-Omitted Curly Braces             &     35         &   51      &       47       &     58    &    32.92  \\
-Pre-Increment                    &     13         &   28      &       50       &     53    &    28.57  \\
-Implicit Predicate               &     41         &   44      &       58       &     68    &    13.13  \\
-Conditional Operator             &     60         &   62      &       60       &     66    &     6.66  \\
-Logic as Control Flow            &     33         &   28      &       41       &     48    &     2.70  \\ 
-Arithmetic as Logic              &     52         &   47      &       64       &     63    &    -5.17  \\ \bottomrule
+  Atom Candidate                 &   Obfuscated   &   Clean   &   Obfuscated   &   Clean   & TO  & TC  & $\Delta$ (\%)    \\ \midrule
+Comma Operator                   &      8         &   29      &       28       &     65    & 36  & 94  & 161.11  \\
+Automatic Semicolong Insertion   &     15         &   21      &       32       &     68    & 47  & 89  & 89.36  \\
+Assignment as Value              &     18         &   43      &       56       &     68    & 74  & 111 & 50.00  \\
+Post-Increment                   &      8         &   17      &       48       &     64    & 56  & 81  & 44.64  \\
+Omitted Curly Braces             &     35         &   51      &       47       &     58    & 82  & 109 & 32.92  \\
+Pre-Increment                    &     13         &   28      &       50       &     53    & 63  & 81  & 28.57  \\
+Implicit Predicate               &     41         &   44      &       58       &     68    & 99  & 112 & 13.13  \\
+Conditional Operator             &     60         &   62      &       60       &     66    & 120 & 128 &  6.66  \\
+Logic as Control Flow            &     33         &   28      &       41       &     48    & 74  & 76  & 2.70  \\ 
+Arithmetic as Logic              &     52         &   47      &       64       &     63    & 116 & 110 & -5.17  \\ \bottomrule
   \end{tabular}

### 2023-04-04 s04.tex
cr candidate

@@
 We do this with the goal of quantifying the prevalence of the investigated atom candidates.
-\revised{If they rarely occur in real-world systems, this would suggest that cannot be a significant source of confusion for developers.
+\revised{If they rarely occur in real-world systems, this would suggest that they cannot be a significant source of confusion for developers.
 This study addresses our first research question (\emph{What is the frequency of occurrence of atom candidates in practice?}).}
@@
 Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research.
-The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC.
+The size of the projects range from small ones (5,543 lines of code) to complex systems with more than 1 MLOC.
 All projects in our dataset have at least \num[group-separator = {,}]{1244} forks and at least \num[group-separator = {,}]{23672} stars.
